### 2019-04-07 TCGA_E9-A1N5_article.Rmd
uploading

---
title: "article_supplementary"
author: "Selcen Ari"
date: "20 Mart 2019"
output: html_document
bibliography: ref.bib
---

# Introduction

This file describes how to apply `ceRNAnetsim` package on a breast cancer patient miRNA:target interaction dataset. Before the approach, we obtained three datasets and combined them.

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# Install the packages which is necessary to obtain and process datasets.
BiocManager::install("SummarizedExperiment", version = "3.8")
library(TCGAbiolinks)
library(SummarizedExperiment)
#install devtools
devtools::install.github(selcenari/ceRNAnetsim)
library(ceRNAnetsim)

```

## 1. How to get gene expression counts of TCGA-E9-A1N5 patient.

We have obtained the gene expression values of patient using [TCGAbiolinks](https://www.bioconductor.org/packages/release/bioc/html/TCGAbiolinks.html) package from [Bioconductor](https://www.bioconductor.org/). For this process, we have followed the instructions of the package. `TCGAbiolinks` package provides to obtain data for whole number of given barcode(s) at once. But, we preferred to download them seperately to show datasets.

+ Obtain the tumor tissue gene expression counts.

```{r,message= FALSE, warning=FALSE}

BCP_tumor <- GDCquery(project = "TCGA-BRCA",
                  data.category = "Transcriptome Profiling",
                  data.type = "Gene Expression Quantification", 
                  workflow.type = "HTSeq - Counts",
                  barcode = "TCGA-E9-A1N5-01A-11R-A14D-07")

GDCdownload(BCP_tumor)
BCPGE_tumor <- GDCprepare(BCP_tumor)

as.data.frame(assay(BCPGE_tumor))%>%
  mutate(ensembl_gene_id = rownames(.))%>%
  dplyr::inner_join(as.data.frame(rowData(BCPGE_tumor)), by = "ensembl_gene_id")%>%
  dplyr::select(ensembl_gene_id, external_gene_name, 1)-> TCGA_E9_A1N5_tumor

colnames(TCGA_E9_A1N5_tumor)[3] <- "GE_tumor"

head(TCGA_E9_A1N5_tumor)
```

+ Obtain the normal tissue gene expression counts.

```{r, message= FALSE, warning=FALSE}
BCP_normal <- GDCquery(project = "TCGA-BRCA",
                  data.category = "Transcriptome Profiling",
                  data.type = "Gene Expression Quantification", 
                  workflow.type = "HTSeq - Counts",
                  barcode = "TCGA-E9-A1N5-11A-41R-A14D-07")
GDCdownload(BCP_normal)

BCPGE_normal <- GDCprepare(BCP_normal)

as.data.frame(assay(BCPGE_normal))%>%
  mutate(ensembl_gene_id = rownames(.))%>%
  dplyr::inner_join(as.data.frame(rowData(BCPGE_normal)), by = "ensembl_gene_id")%>%
  dplyr::select(ensembl_gene_id, external_gene_name, 1)-> TCGA_E9_A1N5_normal

colnames(TCGA_E9_A1N5_normal)[3] <- "GE_normal"

head(TCGA_E9_A1N5_normal)

```

```{r, echo=FALSE, message= FALSE, warning=FALSE}
#remove:
rm(BCPGE_normal, BCPGE_tumor, BCP_normal, BCP_tumor)
```

## 2. How to get miRNA expression counts of TCGA-E9-A1N5 patient.

We used `TCGAbiolinks` package to obtain miRNA expression quantification. The query gives read count of miRNA as isoform chromosome coordination. The data also contains mature miRNA information. So, we processed data to attain -5p -3p isoform information using mirbase release21 dataset.

+ Get the mirbase id of mature miRNA:

We downloaded the mirbase release 21 dataset from [mirbase](http://www.mirbase.org/ftp.shtml) and processed the patient mirna expression datasets as following:

```{r, message= FALSE, warning=FALSE}
library(readr)
read_tsv("hsa_mirna.txt", comment = "#", col_names = FALSE)%>%
  dplyr::select(mirna_type= X3, definition = X9)%>%
  filter(!endsWith(mirna_type, "primary_transcript"))%>%
  tidyr::separate(definition, c("ID", "Alias", "Name", "Derivated"), sep = ";")%>%
  dplyr::select(Alias, Name)%>%
  tidyr::separate(Alias, c("trash1", "ID"), sep = "=")%>%
  tidyr::separate(Name, c("trash2", "Name"), sep = "=")%>%
  dplyr::select(-trash1, -trash2)-> mirbase_id_conv

head(mirbase_id_conv)
```

+ Obtain the miRNA expression of tumor tissue of patient:

```{r, message= FALSE, warning=FALSE}
BCP_mirnatumor <- GDCquery(project = "TCGA-BRCA",
                  data.category = "Transcriptome Profiling",
                  data.type = "Isoform Expression Quantification",
                  workflow.type = "BCGSC miRNA Profiling",
                  barcode = "TCGA-E9-A1N5-01A-11R-A14C-13")

GDCdownload(BCP_mirnatumor)

GDCprepare(BCP_mirnatumor)%>%
  as.data.frame()%>%
  dplyr::select(miRNA_ID, read_count, reads_per_million_miRNA_mapped, miRNA_region)%>%
  dplyr::filter(startsWith(miRNA_region, "mature"))%>%
  dplyr::mutate(mirbase_id =str_remove(miRNA_region, "mature,"))%>%
  dplyr::select(-miRNA_region)%>%
  dplyr::inner_join(mirbase_id_conv, by = c("mirbase_id"="ID"))%>%
  dplyr::select(miRNA_name = Name, read_count, reads_per_million_miRNA_mapped)%>%
  dplyr::group_by(miRNA_name)%>%
  mutate(read_count= sum(read_count), reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
  dplyr::ungroup()%>%
  distinct() -> BCPME_mirnatumor

head(BCPME_mirnatumor)
```

+ Obtain the miRNA expression of normal tissue of patient:

```{r, message= FALSE, warning=FALSE}
BCP_mirnanormal <- GDCquery(project = "TCGA-BRCA",
                  data.category = "Transcriptome Profiling",
                  data.type = "Isoform Expression Quantification",
                  workflow.type = "BCGSC miRNA Profiling",
                  barcode = "TCGA-E9-A1N5-11A-41R-A14C-13")

GDCdownload(BCP_mirnanormal)
# a616435d-0b69-48ac-813d-5d75ad9b85eb.mirbase21.isoforms.quantification.txt

GDCprepare(BCP_mirnanormal)%>%
  as.data.frame()%>%
  dplyr::select(miRNA_ID, read_count, reads_per_million_miRNA_mapped, miRNA_region)%>%
  dplyr::filter(startsWith(miRNA_region, "mature"))%>%
  dplyr::mutate(mirbase_id =str_remove(miRNA_region, "mature,"))%>%
  dplyr::select(-miRNA_region)%>%
  dplyr::inner_join(mirbase_id_conv, by = c("mirbase_id"="ID"))%>%
  dplyr::select(miRNA_name = Name, read_count, reads_per_million_miRNA_mapped)%>%
  dplyr::group_by(miRNA_name)%>%
  mutate(read_count= sum(read_count), reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
  dplyr::ungroup()%>%
  distinct() -> BCPME_mirnanormal

head(BCPME_mirnanormal)
```

## 3. Get the high-throughput experimental miRNA:target dataset.

There are various datasets about miRNA:target pairs such as miRTarBase, DianaTools, miRecords, miRWalk etc. Some of these present the experimentally supported miRNA target pairs or only predicted ones. The experimentally supported datasets generally provides weak evidence for interactions. For these reasons, we obtained the high-throughput experimental miRNA:target dataset from two studies performed by Helwak et al. and Moore et al. These steps were not handle in this file because they contain many processes.

Briefly these datasets contain various common information about miRNA:target interactions such as the miRNA name, miRNAsequence, target name, target sequence, their chromosomal locations, binding location on the target sequence, binding free energy, seed structure. But these datasets provides the informations with different data structures.So we followed the steps:

+ The datasets were directly downloaded from supplementary data files of the studies.
+ It was provided that the datasets are converted to same human genome build.
+ The seed type information was organized as the same style.
+ The datasets were combined.
+ We commited the interaction factors as numeric values according to previous studies.

Finally, we have obtained the experimentally supported miRNA:target dataset.

```{r, message= FALSE, warning=FALSE}
data("experimentalmirnagene")

head(experimentalmirnagene)

```

The methods about miRNA:target interactions are based a basic principle that is reading after isolation of miRNA:target chimerics. The datasets contain all the chimeric miRNA:target structures found in the environment during the experiment. On the other hand, it could be said that the reading is performed as snapshot.Because of that, the methods can provide different chimeric interactions the same miRNA:target pair. We have preferred to select most effective interaction parameters for the same miRNA:target pairs that can exhibit various interactions. The step is performed as:

```{r, message= FALSE, warning=FALSE}
experimentalmirnagene%>%
  dplyr::select(miRNA, Ensembl_Gene_Id, hgnc_symbol, Energy, seed_type_effect, region_effect)%>%
  distinct()%>%
  group_by(Ensembl_Gene_Id, miRNA)%>%
  mutate(seed_type_effect= ifelse(seed_type_effect==max(seed_type_effect), seed_type_effect, max(seed_type_effect)), Energy = ifelse(Energy==min(Energy), Energy, min(Energy)), region_effect= ifelse(region_effect==max(region_effect), region_effect, max(region_effect)))%>%
  distinct()-> tocombine_mirnagene

head(tocombine_mirnagene)
```

## 4. Combine the dataset

```{r, message= FALSE, warning=FALSE}
BCPME_mirnanormal%>%
  dplyr::inner_join(tocombine_mirnagene, by = c("miRNA_name"="miRNA"))%>%
  dplyr::inner_join(TCGA_E9_A1N5_normal, by = c("Ensembl_Gene_Id"="ensembl_gene_id", "hgnc_symbol"="external_gene_name"))%>%
  distinct()%>%
  dplyr::select(hgnc_symbol, miRNA_name, mirna_RPM= reads_per_million_miRNA_mapped, GE_normal, Energy, seed_type_effect, region_effect)-> E9GE_mirnagenenormal

head(E9GE_mirnagenenormal)
```

```{r, message= FALSE, warning=FALSE}
BCPME_mirnatumor%>%
  dplyr::inner_join(tocombine_mirnagene, by = c("miRNA_name"="miRNA"))%>%
  dplyr::inner_join(TCGA_E9_A1N5_tumor, by = c("Ensembl_Gene_Id"="ensembl_gene_id", "hgnc_symbol"="external_gene_name"))%>%
  distinct()%>%
  dplyr::select(hgnc_symbol, miRNA_name, mirna_RPM= reads_per_million_miRNA_mapped, GE_tumor, Energy, seed_type_effect, region_effect)-> E9GE_mirnagenetumor

head(E9GE_mirnagenetumor)

```

## Comparison of two datasets

We have compared two datasets that are obtained for the tumor and normal tissue samples of same patient. We tried to change expression of a gene in normal tissue as the same level in the tumor tissue. Of course, we do not know that the selected gene is responsible for all regulations in the interactions. Moreover, we do not think that one gene causes different regulations in whole interactions. But we wanted to show that a single gene can cause changing regulations of interaction of whole miRNA target network.

For this step, we have determined the changes of the gene expression in terms of fold change:

```{r}
E9GE_mirnagenetumor%>%
  dplyr::select(hgnc_symbol, GE_tumor)%>%
  dplyr::inner_join((E9GE_mirnagenenormal%>%dplyr::select(hgnc_symbol, GE_normal)), by ="hgnc_symbol")%>%
  dplyr::mutate(FC= GE_tumor/GE_normal)%>%
  distinct()

# ABCC1 gene has 5827 read count in tumor tissue although 1420 in normal tissue (FC=4.10)
```

# Approach of Method into Combined Datasets

We selected ABCC1 gene for simulation of regulation on network.

## Find iteration of simulation

```{r}
as.data.frame(E9GE_mirnagenenormal)%>%
iteration_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect, "ABCC1", .iter =200 , how = 4.10, limit = 1)

```

The node amount of changed gene on the system in terms of percentage were shown in above. As seen, firstly, the changed gene count increase, decreasing started in the one point. The system which contains the hundreds of miRNAs and thousands of genes can slowly gain the steady-state again. At first glance, it can be assumed that when all nodes in the system are reached, a stable state will be provided. However, although all nodes are reached, the nodes competing with each other cause the edits to continue for a while.

The dynamics of the approach are shown in package viggnettes [link](https://github.com/selcenari/ceRNAnetsim/blob/master/vignettes/convenient_iteration.Rmd).

So, we offered an approach about to find iteration. `find_iteration` function does not give the iteration to gain steady-state, but it gives the iteration which has maximum affected node counts. The function is applied as following:

```{r}
find_iteration(as.data.frame(E9GE_mirnagenenormal), GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect, "ABCC1", .iter =100 , how = 4.10, limit = 10)

#26
```

We tried to apply the output of the `find_iteration` in simulation.

```{r}

as.data.frame(E9GE_mirnagenenormal)%>%
  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
  update_nodes(once=TRUE)%>%
  update_how("ABCC1", 4.10)%>%
  update_nodes()%>%
  simulate(26)

```

Actually, we have developed to provide a new approach mirna mediated regulation networks. This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and cohorent miRNA:target regulation approach.

### 2019-04-07 article.Rmd
uploading

---
title: "Network based multifactorial modelling microRNA-target interaction."
author: 
- Selcen Ari^[Yildiz Technical University Department of Bioengineering]
- Alper Yilmaz^[Yildiz Technical University Department of Bioengineering]
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document: default
bibliography: ref.bib
biblio-style: NAR-natbib.sty
---

## Abstract

## Introduction

MicroRNAs (miRNAs) that are key regulator of gene expression through various post-transcriptional mechanisms are a family of short non-coding RNAs. Although repression mechanisms of miRNA are not fully understood, exact effect of miRNAs is regulation of direct target/targets negatively. Repressive activities of miRNAs vary depending on many factors that are significant to microRNA:target interactions. These factors include microRNA:target binding energy, binding location in target sequence, base pairing types between miRNA and target, abundance of miRNAs and targets [@grimson_microrna_2007]. Binding energy of miRNA:target complexes varies based on nucleotide context and determines folding stability of complex [@cao_predicting_2012]. It has been demonstrated that the binding energy between miRNA and target indicates stability or affinity of complex [@helwak_mapping_2013] and does not directly determine repressive activity of miRNA [@cao_predicting_2012]. Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5’UTR bind to specific sequence located in 3’UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005]. In recent studies, it has been shown miRNAs can interact with targets via sequence that located in different region such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013]moore_mirnatarget_2015. These studies also shown that binding location could indicate functionality of miRNA:target interaction or affect profile of target abundance. It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is prefectly complementary with targets, seed region at the 5’UTR of miRNAs [@bartel_micrornas:_2009; @grimson_microrna_2007]. On the other hand, some researchers have reported that seed sequence of miRNA can have mismatches or bulged/wobble nucleotides and locates in different region of miRNAs [@hafner_transcriptome-wide_2010; @helwak_mapping_2013]. In addition to all of these, abundance of miRNAs and targets and miRNA:target ratio in cells are determinant for efficiency of miRNA:target interaction [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014]. 
As it is possible for miRNAs to suppress multiple targets, an individual RNA molecule can also be targeted by various miRNAs. In that case, the targeted RNAs exhibit competitor behavior, that is hypothesized as competing endogenous RNAs (ceRNAs) [@ala_integrated_2013; @cesana_deciphering_2013], against their miRNAs. Briefly, Ala et al. have explained the ceRNA hypothesis as disturbtion of the other target when a steady-state system that included one miRNA and two target, was perturbed with expression changes at one of the targets [@ala_integrated_2013]. When interaction between miRNAs and their target were dealed in the cell, explanation and prediction of results of perturbation is difficult due to varieties of interactions. Various computational and experimental studies have applied to be understood for ceRNA:miRNA interactions. It has been observed that when abundance of one of the targets of miR-122 was increased, the other target expression also slightly increased as a result of decreasing repressive activity of miR-122 on them [@denzler_assessing_2014]. Bosson et al. have developed a mathematical model for changes on total target pool concentration after grouping to affinity of targets. They have demonstrated that miRNA activity correlated with affinity between miRNA and target [@bosson_endogenous_2014]. Cooperative efficiency of miRNAs as well as competitor behaviors of targets also were studied and it has been demonstrated crucial for regulations at expression of targets [@denzler_impact_2016]. MiRNA:target interactions have modelled as stoichiometric and catalytic mechanisms and Figliuzzi et al. have recommended to handling models in network structures [@figliuzzi_micrornas_2013]. The model that can explain miRNA target interaction through topological features has been applied at bipartite network by Nitzan et al. [@nitzan_interactions_2014]. Robinson and Henderson applied the model that handles miRNA-target direct and indirect interactions via common miRNA of gene and target of miRNA, at bipartite network. It has been demonstrated that whole miRNAs and targets in network can interacted through common miRNAs and genes [@robinson_modelling_2018. Associated genes that is targets of the same miRNAs have been found with help of correlation of gene expression changes in recent study [@markus_list_sponge_2017]. List et al. have specified that their approach can be useful for ceRNA’s studies and published these approach as a package that can use with R programming.

## Methods

### miRNA:target ratio theoretical action in network based ceRNA model

We developed a novel network based approach that can use to simulation of gene regulation after any perturbation at miRNA:target interactions. If a perturbation or change induces on graph, whole interacted nodes vary because of changes in miRNA:target ratios. The system was assumed as steady state that miRNAs affect according to the ratio of the individual target to all targets (Fig.1a, expression values shown as integer for comprehensible sampling). When expression of gene 2 (G2) increased to two fold (Fig. 1b), distribution of miRNA (M1) expression changes (Fig. 1c). Consequently, expression of common target (G4) of miRNA1 and miRNA2 also varies. The change at expression of common target triggers to changes of proportional distribution of miRNA2 (Fig. 1d).

![Shematic presentation of mechanism of network based model. a) In steady state, miRNAs (M: triangle grey) repress targets (G: circle coconut) according to proportion of their targets' expression. b-c-d) MiRNA expression distributions to targets vary when expression of one of the targets perturbed. ](Fig1.jpg)

The model was developed by making some assumptions. Transcription rates and degradation rates of miRNAs were assumed as stationary and equal. Targets also have stabil transcription and degradation rates, at the same time they are equal. The factors that could be affect the miRNA:target interaction are disregarded in this approach. It is considered that miRNAs degradates or inhibits them via any mechanisms after binding to target while the miRNAs were used reversibly. In initial conditions, gene expression amounts in system have been considered after repression activity of miRNAs. If a change occurs in the system, variation on miRNA proportional distribution are reflected as change of repression activity.

$$Eff_{gi}= C_m * C_{gi}/\sum_{0}^{i} C_g $$
$$R_{gi}= Eff_{i1}+ Eff_{i2} ... $$

The efficiency of a miRNA on the individual target (Effgi) is calculated according to i; Cm, miRNA expression in the group; Cgi, symbolize individual gene expression; Cg gene expressions in the group (Groups are determined depending on miRNAs. Targets that are interacted with the same miRNA find in a group.). In this approach, because a gene can be found in different groups, the cooperative activity of microRNAs also can be calculated. Total repression activity of different miRNAs on the common target, R, is calculated based on ii; sum of efficiency of different miRNAs on the target.

![Calculations for determine to miRNA binding and repression efficiency. G, Gene; M, miRNA; STE, seed type effect; RE, Region Effect; E, Energy; STE’, normalized values of seed type efficiency coefficient; RE’, normalized values of region efficiency coefficient; E’, normalized values of energy coefficient.](fig2.jpg)

### miRNA action in network based multifactorial ceRNA model

Interactions between miRNAs and their targets can be affected from different parameters. For this reason, we improved model with factors that can affect to miRNA activity. In our approach, the parameters are classified as binding and efficiency factors and all of these are evaluated in their groups. Binding factors determine interaction between miRNA and target. The efficiency factors decide functionality of binding. In the literature, binding free energy [@cao_predicting_2012 ; @helwak_mapping_2013] and seed type [@werfel_preferential_2017]in miRNA:target interactions are represented as binding affinity. Efficiency factors determine how many of microRNA:target complexes will result in inhibition. For example, binding region on the target shows whether the target can be inhibited (functionality)[@hausser_analysis_2013; @helwak_mapping_2013]. Both of binding and efficiency factors are evaluated in the groups and normalized according to iii; each factor is normalized based on maximum variables that found in their groups.

$$F`= F/F_{max}$$

The normalized values of factors take into account to determine binding activity and miRNA efficiency on targets (Fig 2).  Binding affinities (activity, Eff) of miRNAs on each individual gene are calculated as shown in iv; Cm, miRNA expression in the group; Cg, Gene expression; Cgi, individual gene expression; gi, individual gene; g, whole genes in a group (Fig 2c).

$$Eff_{gi}= C_m * E`_{gi}*STE`_{gi}*C_{gi}/(\sum_{0}^{i} E`_{gi}*STE`_{gi}*C_{gi}) $$

$$Eff_{gi}= Eff_{gi}*RE`_{gi} $$

After miRNA binds to its target, but might not repress to bound target. The functionality of bound miRNA on target depends on efficiency factors like region that is binding sequence of miRNA on its target. Exact repression efficiency of miRNA is calculated according to equation v (Fig 2d); RE’gi, normalized values of region efficiency coefficient between miRNA and gene. The cooperative repression activity of miRNAs to their common targets is figured out as shown in Fig 2e.

![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow.jpg)

### Defined functions for ceRNA models

We defined the functions that can be used with R programming. Briefly, these functions process a given miRNA:gene dataset and convert to graph object. All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous sections. The functions and steps of approach are explained as following:
1-2. Convertion of dataset: priming_graph function process the given dataset that includes competing elements in first variable and repressive element in second variable. If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa. The formulations that are given in equations (i-v) are performed in this function. This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount. It is assumed that the initial target amounts in the dataset is observed after the reppressive activity of miRNAs in steady-state. 
2-3. Transition of variables in graph: In the previous step, the calculations are performed in the edge variables of the graph object. However, the graph object allows to use node variables, while the node features are handled to the graph. In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.
3-4. Trigger change in graph: The dataset are assumed as steady-state in previous step and the efficieny coefficients are calculated according to this acceptance. In the network that is found in steady-state conditions, the change is applied to the graph object for disturbtion of steady-state. To provide the disturbtion in the network the workflow offer two methods: update_variables and update_how. The fist, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables. The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
4-5. Updating current values of variables: After variables updating in edge varibles, these are carried to node variables. Current and previous values of variables are stored as node variables with update_variables function.
5-6. Simulation of competing behavior of targets: After the change in the steady- state conditions, the network elements try to gain steady-state again. This process progresses as repeating of regulations after the spreading the changes in the network. In this step, simulation of regulations according to given cycle count in simulate function is applied. After each simulation cycle, the miRNA repression values are re-calculated and the current values of competing elements are found and saved. The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node data.
The node elements in the dataset are handled as two type; repressives (miRNAs) and competings (targets). It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used. If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user. On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger. The functions that are used in the approach are developed with R programming so as can be used with other packages. These are can be found in the github repository [ceRNAnetsim github page](https://github.com/selcenari/ceRNAnetsim) and improved with contributions of others.

## Results and Conclusions

We developed a model approach and workflow for competitive ceRNA regulation in this study. The basic mode of miRNA repression activity has been based on miRNA and target abundance in various researches[@arvey_target_2010; @denzler_assessing_2014]. For this reason, we performed first application as considering only abundance of miRNAs and targets (Fig 1). After the increasing in a gene expression value, expression values of other genes also changed slightly in comparison with alteration in perturbation started gene node. Different studies have shown that the other gene expression values also rise differently, if a gene abundance increases in ceRNA system [@lai_understanding_2016; @salmena_cerna_2011; @tay_multilayered_2014]. It was observed that primary neighborhoods of the trigger gene change with the same ratio, but cooperative efficiency of miRNAs on common target causes differently change in common target (Fig 1d). The common target in system displayed like a trigger for the other group and induced changes of expression values of genes on the other group. Therefore, effect of primary change of the gene (G2) on genes of the other group was observed slightly (Fig 1d). In addition, like shown by ceRNA hypothesis model of Ala et al., after the increasing of a gene expression (G2), the miRNA that is found in the same group (M1) tended to be more repressive on this target. This caused to decrease in increased expression value of primer triggering gene.
When the expression of a gene (G2) was increased to two fold, we observed that the expression changes were more evident. So, we considered that high miRNA:target ratio results in stronger miRNA repression activity. This results also are coherent with previous reports that were offered by Arvey et al. and Denzler et al.[@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014]. When the simulation was ran in sample dataset that includes lower target abundance, the first response of primary neihgborhood of the same group was determined as similar with in system that have high target abudance. But regulations of ceRNAs were observed differences in the folowing steps of simulation. In overall state in the end of simulation, we established that all regulations more prominent when compared with low miRNA:target ratio datset. The model appeared the importance of mRNA:target ratio in regulation of competing RNAs.

![Target regulations with interaction parameters. a) In the steady-state the repression activity of miRNAs on the targets after binding and repression efficiency. b) The changes the repression activities after increasing of G2 expression. c) Perturbation of primary neighborhoods of M1 miRNA (M1 miRNA group). d) Regulation of gene expression of other gene group via triggering target (common target between M1 and M2). ](Fig4.jpg)

It has been reported in several studies that miRNA regulatory interactions also is affected different parameters. For example, Xu et. al have investigated the importance of seed pairing type between miRNAs and their targets and target site location by using proteomic dataset [@xu_characterization_2014]. They have proposed that the features of binding between miRNA and target can be critical efficiency of miRNA and can be used to find effective miRNA molecules. In addition, binding energy between miRNA and targets also could be determinant for miRNA efficiency. It has been reported by Breda et. al. that strength of miRNA:target interactions is depended on binding energy of complexes [@breda_quantifying_2015]. As similar to these, Bosson et al. have associated affinity and seed pairing of miRNA target pairs and suggested that number of canonical base pairing is correlated with affinity. For these reasons, we improved the model with optional interaction parameters that could be useful for understanding of miRNA repression activity. We explained the model in the same sample dataset with using different factors and simulated via the same change (Fig 4). When the factors were taken into account in the system, miRNA efficiencies varied as shown in Fig 4a. Although the miRNA target ratios in initial states were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed. When expression of a gene (G2) increased, expression values of all genes also change differentially because of contribution of efficiency factors (Fig 4 b,c,d).
In regulation of miRNA target sample system (Fig 4), miRNA (M1) repressive efficiency on the primary triggering gene (G2) is low in steady-state. So it has been observed that the regulatory activities of miRNA efficiencies on the targets are weak after the increase of miRNA target G2.  When we ran to model with two fold increasing expression value of common target (G4), the changes of other gene expressions were observed more prominent. Furthermore, folding of expression of target gene that has strong miRNA repression efficiency also resulted in evident perturbation.
The defined miRNA’ number is very low in comparison with defined protein coding genes’ number. So, this results in that the miRNA amounts in the system must be high or targeted RNAs’ amount must be low for effective miRNA activity. For these reasons, we tried to simulate a real dataset which contains thousands of genes and hundreds of miRNAs. Briefly, we have obtained the datasets and combined these via R programming. We have applied the our approach into integrated dataset using a selected gene. Firstly we observed the activities in the iterations and determined behaviours of the network after simulation. As expected, we have observed that the the huge dataset, compared to those containing a few genes and miRNAs, exhibited cumbersome behaviours and complexity to gain steady-state. We realized that a small change in a system with a large change actually brought many arrangements. This can be explained with the competing groups. Because the genes are targeted with many common miRNAs, and this results in that the common targets exhibits competing behaviours in different groups. On the other hand, the changes of gene expressions were very different. While the change in some of genes were highly evident, somes have changed slightly. So, we have compared two datasets: the tumor gene expression and simulation results. When we analysed comparison of two datasets, we observed that the results of simulation is not convenient with tumor tissue expression values. The simulation results were not expected to be concordant with tumor tissue expression values because of a single gene could not be responsible for gene regulation in all tissue. Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored. Of course, it would be more useful to test a dataset that includes gene expression values before and after the regulation of a gene. When we evaluated the suitability of approach, the approach allows different miRNA targets to be considered together and can work with different regulatory elements due to the network structure. 
The network based approaches have been developed in previous studies. Figliuzzi et al. have tried to explain the ceRNA crosstalk in a network-like minimal interaction structure with concentrations of ceRNA and miRNAs. They have pointed out that the larger miRNA number can cause to evident crosstalk between ceRNAs. It has been thought that miRNAs and their targets interactions depend on rates of transcription, degradation, binding and unbinding in the network based kinetic model which has been developed by Nitzan et al. Nitzan and collaborators have used the high throughput experimental dataset [@helwak_mapping_2013] for miRNA:target pair that occur high free energy and microarray dataset of a miR-92a depletion for expression values. They have been demonstrated that distant ceRNAs can interact with each other via indirect links, and the interactions change to depend on distance between ceRNAs, and topological features of network [@nitzan_interactions_2014]. 
In our approach, we have not taken into account transcription, degradation or binding rates of elements in network. Because, although it is known as the miRNAs are highly stable, the transcription and degradation rates of miRNAs change to depend on according to cellular conditions [@ruegger_microrna_2012]. According to our search, there is no dataset that was includes to degradation and transcription rates on specific cellular conditions for each miRNA. So, we preferred to accept that degradation and transcription rates of each miRNAs are equal to each other, instead of determining a constant value for all miRNAs in the system (The same applies to mRNA.). On the other hand other regulation parameters such as gene-gene interactions and transcription factors are ignored but the network structure can promote to integration of other regulation elements. In the future, with development in experimental techniques about features of miRNAs and their targets, more consistent and useful results can be obtained from our approach. Additionally, this can provide to contribution to prediction of abnormal regulations and pathways with the studies that will be developed.

### 2019-04-07 fig1_2app.Rmd
uploading

---
title: "supplement"
author: "Selcen Arı"
date: "7 Nisan 2019"
output: html_document
---

```{r}
data("minsamp")

minsamp
```

```{r}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
  update_nodes(once=TRUE)%>%
  update_how("Gene2", 2)%>%
  update_nodes()%>%
  simulate(3)%>%activate(edges)%>%
  as_tibble()%>%
  select(Competing_name,comp_count_list,effect_list)%>%
  unnest()

```

```{r}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_nodes(once=TRUE)%>%
  update_how("Gene2", 2)%>%
  update_nodes()%>%
  simulate(3)%>%activate(edges)%>%
  as_tibble()%>%
  select(Competing_name,comp_count_list,effect_list)%>%
  unnest()
```

### 2019-04-07 article.Rmd
cross-ref

@@
 These factors include microRNA:target binding energy, binding location in target sequence, base pairing types between miRNA and target, abundance of miRNAs and targets [@grimson_microrna_2007].
-Binding energy of miRNA:target complexes varies based on nucleotide context and determines folding stability of complex [@cao_predicting_2012].
+Binding energies of miRNA:target complexes vary based on nucleotide context and determine folding stability of complex [@cao_predicting_2012].
 It has been demonstrated that the binding energy between miRNA and target indicates stability or affinity of complex [@helwak_mapping_2013] and does not directly determine repressive activity of miRNA [@cao_predicting_2012].
@@
 [@nitzan_interactions_2014].
-Robinson and Henderson applied the model that handles miRNA-target direct and indirect interactions via common miRNA of gene and target of miRNA, at bipartite network.
-It has been demonstrated that whole miRNAs and targets in network can interacted through common miRNAs and genes [@robinson_modelling_2018.
-Associated genes that is targets of the same miRNAs have been found with help of correlation of gene expression changes in recent study [@markus_list_sponge_2017].
+Robinson and Henderson applied the model that handles miRNA:target direct and indirect interactions via common miRNA of gene and target of miRNA, at bipartite network.
+It has been demonstrated that whole miRNAs and targets in network can interacted through common miRNAs and genes [@robinson_modelling_2018].
+Associated genes that are targets of the same miRNAs have been found with help of correlation of gene expression changes in recent study [@markus_list_sponge_2017].
 List et al.
@@
 If a perturbation or change induces on graph, whole interacted nodes vary because of changes in miRNA:target ratios.
-The system was assumed as steady state that miRNAs affect according to the ratio of the individual target to all targets (Fig.1a, expression values shown as integer for comprehensible sampling).
-When expression of gene 2 (G2) increased to two fold (Fig.
-1b), distribution of miRNA (M1) expression changes (Fig.
-1c).
+The system was assumed as steady state that miRNAs affect according to the ratio of the individual target to all targets (\autoref{fig1}a, expression values shown as integer for comprehensible sampling).
+When expression of gene 2 (G2) increased to two fold (\autoref{fig1}b), distribution of miRNA (M1) expression changes (\autoref{fig1}c).
 Consequently, expression of common target (G4) of miRNA1 and miRNA2 also varies.
-The change at expression of common target triggers to changes of proportional distribution of miRNA2 (Fig.
-1d).
+The change at expression of common target triggers to changes of proportional distribution of miRNA2 (\autoref{fig1}d).
 
@@
 b-c-d) MiRNA expression distributions to targets vary when expression of one of the targets perturbed.
-](Fig1.jpg)
+](Fig1.jpg){#fig1}
 
@@
 
-$$Eff_{gi}= C_m * C_{gi}/\sum_{0}^{i} C_g $$
-$$R_{gi}= Eff_{i1}+ Eff_{i2} ... $$
+\begin{equation} 
+    Eff_{gi}= C_m * C_{gi}/\sum_{1}^{i} C_g \tag{1}\label{eq:1}
+\end{equation}
 
-The efficiency of a miRNA on the individual target (Effgi) is calculated according to i; Cm, miRNA expression in the group; Cgi, symbolize individual gene expression; Cg gene expressions in the group (Groups are determined depending on miRNAs.
+\begin{equation}
+   R_{gi}= Eff_{i1}+ Eff_{i2} ... \tag{2}\label{eq:2}
+\end{equation}
+
+The efficiency of a miRNA on the individual target ($Eff_{gi}$) is calculated according to equation \eqref{eq:1}; $C_m$, miRNA expression in the group; $C_{gi}$, symbolize individual gene expression; $C_g$ gene expressions in the group (Groups are determined depending on miRNAs.
 Targets that are interacted with the same miRNA find in a group.).
 In this approach, because a gene can be found in different groups, the cooperative activity of microRNAs also can be calculated.
-Total repression activity of different miRNAs on the common target, R, is calculated based on ii; sum of efficiency of different miRNAs on the target.
+Total repression activity of different miRNAs on the common target, $R$, is calculated based on equation \eqref{eq:2}; sum of efficiency of different miRNAs on the target.
 
 ![Calculations for determine to miRNA binding and repression efficiency.
-G, Gene; M, miRNA; STE, seed type effect; RE, Region Effect; E, Energy; STE’, normalized values of seed type efficiency coefficient; RE’, normalized values of region efficiency coefficient; E’, normalized values of energy coefficient.](fig2.jpg)
+$G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE’$, normalized values of seed type efficiency coefficient; $RE’$, normalized values of region efficiency coefficient; $E’$, normalized values of energy coefficient.](fig2.jpg){#fig2}
 
@@
 For example, binding region on the target shows whether the target can be inhibited (functionality)[@hausser_analysis_2013; @helwak_mapping_2013].
-Both of binding and efficiency factors are evaluated in the groups and normalized according to iii; each factor is normalized based on maximum variables that found in their groups.
+Both of binding and efficiency factors are evaluated in the groups and normalized according to equation \eqref{eq:3}; each factor is normalized based on maximum variables that found in their groups.
 
-$$F`= F/F_{max}$$
+\begin{equation}
+F`= F/F_{max} \tag{3}\label{eq:3}
+\end{equation}
 
-The normalized values of factors take into account to determine binding activity and miRNA efficiency on targets (Fig 2).
-Binding affinities (activity, Eff) of miRNAs on each individual gene are calculated as shown in iv; Cm, miRNA expression in the group; Cg, Gene expression; Cgi, individual gene expression; gi, individual gene; g, whole genes in a group (Fig 2c).
+The normalized values of factors take into account to determine binding activity and miRNA efficiency on targets (\autoref{fig2}).
+Binding affinities (activity, $Eff$) of miRNAs on each individual gene are calculated as shown in equation \eqref{eq:4}; $C_m$, miRNA expression in the group; $C_g$, Gene expression; $C_{gi}$, individual gene expression; gi, individual gene; g, whole genes in a group (\autoref{fig2}c).
 
-$$Eff_{gi}= C_m * E`_{gi}*STE`_{gi}*C_{gi}/(\sum_{0}^{i} E`_{gi}*STE`_{gi}*C_{gi}) $$
+\begin{equation}
+Eff_{gi}= C_m * E`_{gi}*STE`_{gi}*C_{gi}/(\sum_{0}^{i} E`_{gi}*STE`_{gi}*C_{gi}) \tag{4}\label{eq:4}
+\end{equation}
 
-$$Eff_{gi}= Eff_{gi}*RE`_{gi} $$
+\begin{equation}
+Eff_{gi}= Eff_{gi}*RE`_{gi} \tag{5}\label{eq:5}
+\end{equation}
 
@@
 The functionality of bound miRNA on target depends on efficiency factors like region that is binding sequence of miRNA on its target.
-Exact repression efficiency of miRNA is calculated according to equation v (Fig 2d); RE’gi, normalized values of region efficiency coefficient between miRNA and gene.
-The cooperative repression activity of miRNAs to their common targets is figured out as shown in Fig 2e.
+Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:5} (\autoref{fig2}d); $RE’_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
+The cooperative repression activity of miRNAs to their common targets is figured out as shown in \autoref{fig2}e.
 
-![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow.jpg)
+![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow.jpg){#fig3}
 
@@
 All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous sections.
-The functions and steps of approach are explained as following:
+The functions and steps of approach are explained as following \autoref{fig3}:
 1-2.
@@
 The basic mode of miRNA repression activity has been based on miRNA and target abundance in various researches[@arvey_target_2010; @denzler_assessing_2014].
-For this reason, we performed first application as considering only abundance of miRNAs and targets (Fig 1).
+For this reason, we performed first application as considering only abundance of miRNAs and targets (\autoref{fig1}).
 After the increasing in a gene expression value, expression values of other genes also changed slightly in comparison with alteration in perturbation started gene node.
 Different studies have shown that the other gene expression values also rise differently, if a gene abundance increases in ceRNA system [@lai_understanding_2016; @salmena_cerna_2011; @tay_multilayered_2014].
-It was observed that primary neighborhoods of the trigger gene change with the same ratio, but cooperative efficiency of miRNAs on common target causes differently change in common target (Fig 1d).
+It was observed that primary neighborhoods of the trigger gene change with the same ratio, but cooperative efficiency of miRNAs on common target causes differently change in common target (\autoref{fig1}d).
 The common target in system displayed like a trigger for the other group and induced changes of expression values of genes on the other group.
-Therefore, effect of primary change of the gene (G2) on genes of the other group was observed slightly (Fig 1d).
+Therefore, effect of primary change of the gene (G2) on genes of the other group was observed slightly (\autoref{fig1}d).
 In addition, like shown by ceRNA hypothesis model of Ala et al., after the increasing of a gene expression (G2), the miRNA that is found in the same group (M1) tended to be more repressive on this target.
@@
 d) Regulation of gene expression of other gene group via triggering target (common target between M1 and M2).
-](Fig4.jpg)
+](Fig4.jpg){#fig4}
 
@@
 For these reasons, we improved the model with optional interaction parameters that could be useful for understanding of miRNA repression activity.
-We explained the model in the same sample dataset with using different factors and simulated via the same change (Fig 4).
-When the factors were taken into account in the system, miRNA efficiencies varied as shown in Fig 4a.
+We explained the model in the same sample dataset with using different factors and simulated via the same change (\autoref{fig4}).
+When the factors were taken into account in the system, miRNA efficiencies varied as shown in \autoref{fig4}a.
 Although the miRNA target ratios in initial states were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed.
-When expression of a gene (G2) increased, expression values of all genes also change differentially because of contribution of efficiency factors (Fig 4 b,c,d).
-In regulation of miRNA target sample system (Fig 4), miRNA (M1) repressive efficiency on the primary triggering gene (G2) is low in steady-state.
+When expression of a gene (G2) increased, expression values of all genes also change differentially because of contribution of efficiency factors (\autoref{fig4} b,c,d).
+In regulation of miRNA target sample system (\autoref{fig4}), miRNA (M1) repressive efficiency on the primary triggering gene (G2) is low in steady-state.
 So it has been observed that the regulatory activities of miRNA efficiencies on the targets are weak after the increase of miRNA target G2.
@@
 Additionally, this can provide to contribution to prediction of abnormal regulations and pathways with the studies that will be developed.
+
+## References

### 2019-04-08 article.Rmd
check

@@
 In addition to all of these, abundance of miRNAs and targets and miRNA:target ratio in cells are determinant for efficiency of miRNA:target interaction [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
+
 As it is possible for miRNAs to suppress multiple targets, an individual RNA molecule can also be targeted by various miRNAs.
@@
 ## Methods
-
 ### miRNA:target ratio theoretical action in network based ceRNA model
-
 We developed a novel network based approach that can use to simulation of gene regulation after any perturbation at miRNA:target interactions.
@@
 a) In steady state, miRNAs (M: triangle grey) repress targets (G: circle coconut) according to proportion of their targets' expression.
-b-c-d) MiRNA expression distributions to targets vary when expression of one of the targets perturbed.
-](Fig1.jpg){#fig1}
+b-c-d) MiRNA expression distributions to targets vary when expression of one of the targets perturbed.](Fig1.jpg){#fig1}
 
@@
 If a change occurs in the system, variation on miRNA proportional distribution are reflected as change of repression activity.
-
 \begin{equation} 
@@
 \end{equation}
-
 \begin{equation}
@@
 
-![Calculations for determine to miRNA binding and repression efficiency.
+![Calculations to determine of miRNA binding and repression efficiency.
 $G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE’$, normalized values of seed type efficiency coefficient; $RE’$, normalized values of region efficiency coefficient; $E’$, normalized values of energy coefficient.](fig2.jpg){#fig2}
@@
 The efficiency factors decide functionality of binding.
-In the literature, binding free energy [@cao_predicting_2012 ; @helwak_mapping_2013] and seed type [@werfel_preferential_2017]in miRNA:target interactions are represented as binding affinity.
+In the literature, binding free energy [@cao_predicting_2012 ; @helwak_mapping_2013] and seed type [@werfel_preferential_2017] in miRNA:target interactions are represented as binding affinity.
 Efficiency factors determine how many of microRNA:target complexes will result in inhibition.
@@
 Both of binding and efficiency factors are evaluated in the groups and normalized according to equation \eqref{eq:3}; each factor is normalized based on maximum variables that found in their groups.
-
 \begin{equation}
@@
 Binding affinities (activity, $Eff$) of miRNAs on each individual gene are calculated as shown in equation \eqref{eq:4}; $C_m$, miRNA expression in the group; $C_g$, Gene expression; $C_{gi}$, individual gene expression; gi, individual gene; g, whole genes in a group (\autoref{fig2}c).
-
 \begin{equation}
-Eff_{gi}= C_m * E`_{gi}*STE`_{gi}*C_{gi}/(\sum_{0}^{i} E`_{gi}*STE`_{gi}*C_{gi}) \tag{4}\label{eq:4}
+Eff_{gi}= C_m * E`_{gi}*STE`_{gi}*C_{gi}/(\sum_{1}^{i} E`_{gi}*STE`_{gi}*C_{gi}) \tag{4}\label{eq:4}
 \end{equation}
-
 \begin{equation}
@@
 ### Defined functions for ceRNA models
-
 We defined the functions that can be used with R programming.
 Briefly, these functions process a given miRNA:gene dataset and convert to graph object.
-All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous sections.
-The functions and steps of approach are explained as following \autoref{fig3}:
-1-2.
-Convertion of dataset: priming_graph function process the given dataset that includes competing elements in first variable and repressive element in second variable.
+All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
+The functions and steps of approach are explained as following (\autoref{fig3}) :
+*Convertion of dataset*: priming_graph function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
 If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
-The formulations that are given in equations (i-v) are performed in this function.
+The formulations that are given in equations \eqref{eq:1}-\eqref{eq:5} are performed in this function.
 This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
 It is assumed that the initial target amounts in the dataset is observed after the reppressive activity of miRNAs in steady-state.
-2-3.
-Transition of variables in graph: In the previous step, the calculations are performed in the edge variables of the graph object.
+*Transition of variables in graph*: In the previous step, the calculations are performed in the edge variables of the graph object.
 However, the graph object allows to use node variables, while the node features are handled to the graph.
 In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.
-3-4.
-Trigger change in graph: The dataset are assumed as steady-state in previous step and the efficieny coefficients are calculated according to this acceptance.
+*Trigger change in graph*: The dataset are assumed as steady-state in previous step and the efficieny coefficients are calculated according to this acceptance.
 In the network that is found in steady-state conditions, the change is applied to the graph object for disturbtion of steady-state.
 To provide the disturbtion in the network the workflow offer two methods: update_variables and update_how.
-The fist, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
+The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
 The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
-4-5.
-Updating current values of variables: After variables updating in edge varibles, these are carried to node variables.
+*Updating current values of variables*: After variables updating in edge varibles, these are carried to node variables.
 Current and previous values of variables are stored as node variables with update_variables function.
-5-6.
-Simulation of competing behavior of targets: After the change in the steady- state conditions, the network elements try to gain steady-state again.
+*Simulation of competing behavior of targets*: After the change in the steady-state conditions, the network elements try to gain steady-state again.
 This process progresses as repeating of regulations after the spreading the changes in the network.
@@
 
-We developed a model approach and workflow for competitive ceRNA regulation in this study.
+We have developed a model approach and workflow for competitive ceRNA regulation in this study.
 The basic mode of miRNA repression activity has been based on miRNA and target abundance in various researches[@arvey_target_2010; @denzler_assessing_2014].
@@
 It was observed that primary neighborhoods of the trigger gene change with the same ratio, but cooperative efficiency of miRNAs on common target causes differently change in common target (\autoref{fig1}d).
-The common target in system displayed like a trigger for the other group and induced changes of expression values of genes on the other group.
+The common target in system has displayed like a trigger for the other group and induced changes of expression values of genes on the other group.
 Therefore, effect of primary change of the gene (G2) on genes of the other group was observed slightly (\autoref{fig1}d).
 In addition, like shown by ceRNA hypothesis model of Ala et al., after the increasing of a gene expression (G2), the miRNA that is found in the same group (M1) tended to be more repressive on this target.
-This caused to decrease in increased expression value of primer triggering gene.
+This has caused to decrease in increased expression value of primer triggering gene.
 When the expression of a gene (G2) was increased to two fold, we observed that the expression changes were more evident.
-So, we considered that high miRNA:target ratio results in stronger miRNA repression activity.
+So, we have considered that high miRNA:target ratio results in stronger miRNA repression activity.
 This results also are coherent with previous reports that were offered by Arvey et al.
 and Denzler et al.[@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
-When the simulation was ran in sample dataset that includes lower target abundance, the first response of primary neihgborhood of the same group was determined as similar with in system that have high target abudance.
+When the simulation was run in sample dataset that includes lower target abundance, the first response of primary neihgborhood of the same group was determined as similar with in system that have high target abudance.
 But regulations of ceRNAs were observed differences in the folowing steps of simulation.
-In overall state in the end of simulation, we established that all regulations more prominent when compared with low miRNA:target ratio datset.
-The model appeared the importance of mRNA:target ratio in regulation of competing RNAs.
+In overall state in the end of simulation, we have established that all regulations more prominent when compared with low miRNA:target ratio dataset.
+The model has appeared the importance of mRNA:target ratio in regulation of competing RNAs.
 
@@
 have associated affinity and seed pairing of miRNA target pairs and suggested that number of canonical base pairing is correlated with affinity.
-For these reasons, we improved the model with optional interaction parameters that could be useful for understanding of miRNA repression activity.
-We explained the model in the same sample dataset with using different factors and simulated via the same change (\autoref{fig4}).
+For these reasons, we have improved the model with optional interaction parameters that could be useful for understanding of miRNA repression activity.
+We have explained the model in the same sample dataset with using different factors and simulated via the same change (\autoref{fig4}).
 When the factors were taken into account in the system, miRNA efficiencies varied as shown in \autoref{fig4}a.
@@
 When expression of a gene (G2) increased, expression values of all genes also change differentially because of contribution of efficiency factors (\autoref{fig4} b,c,d).
+
 In regulation of miRNA target sample system (\autoref{fig4}), miRNA (M1) repressive efficiency on the primary triggering gene (G2) is low in steady-state.
 So it has been observed that the regulatory activities of miRNA efficiencies on the targets are weak after the increase of miRNA target G2.
-When we ran to model with two fold increasing expression value of common target (G4), the changes of other gene expressions were observed more prominent.
+When we ran to model with two fold increasing expression value of common target (G4), the changes of other gene expressions have observed more prominent.
 Furthermore, folding of expression of target gene that has strong miRNA repression efficiency also resulted in evident perturbation.
+
 The defined miRNA’ number is very low in comparison with defined protein coding genes’ number.
 So, this results in that the miRNA amounts in the system must be high or targeted RNAs’ amount must be low for effective miRNA activity.
-For these reasons, we tried to simulate a real dataset which contains thousands of genes and hundreds of miRNAs.
+We have tried to simulate a real dataset which contains thousands of genes and hundreds of miRNAs.
 Briefly, we have obtained the datasets and combined these via R programming.
 We have applied the our approach into integrated dataset using a selected gene.
-Firstly we observed the activities in the iterations and determined behaviours of the network after simulation.
-As expected, we have observed that the the huge dataset, compared to those containing a few genes and miRNAs, exhibited cumbersome behaviours and complexity to gain steady-state.
+Firstly, we have observed the activities in the iterations and determined behaviours of the network after simulation.
+As expected, we have observed that the big dataset, compared to those containing a few genes and miRNAs, exhibited cumbersome behaviours and complexity to gain steady-state.
 We realized that a small change in a system with a large change actually brought many arrangements.
@@
 So, we have compared two datasets: the tumor gene expression and simulation results.
-When we analysed comparison of two datasets, we observed that the results of simulation is not convenient with tumor tissue expression values.
+When we analysed comparison of two datasets, we have observed that the results of simulation is not convenient with tumor tissue expression values.
 The simulation results were not expected to be concordant with tumor tissue expression values because of a single gene could not be responsible for gene regulation in all tissue.
@@
 When we evaluated the suitability of approach, the approach allows different miRNA targets to be considered together and can work with different regulatory elements due to the network structure.
+
 The network based approaches have been developed in previous studies.
@@
 They have been demonstrated that distant ceRNAs can interact with each other via indirect links, and the interactions change to depend on distance between ceRNAs, and topological features of network [@nitzan_interactions_2014].
+
 In our approach, we have not taken into account transcription, degradation or binding rates of elements in network.

### 2019-04-10 TCGA_E9-A1N5_article.Rmd
article rev

@@
 date: "20 Mart 2019"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 bibliography: ref.bib
@@
 # Install the packages which is necessary to obtain and process datasets.
-BiocManager::install("SummarizedExperiment", version = "3.8")
+#BiocManager::install("SummarizedExperiment", version = "3.8")
 library(TCGAbiolinks)
@@
 library(ceRNAnetsim)
-
 ```
@@
 
-+ Obtain the tumor tissue gene expression counts.
++ Obtain to gene expression counts of tumor tissue.
 
@@
 
-+ Obtain the normal tissue gene expression counts.
++ Obtain to gene expression counts of normal tissue.
 
@@
 
-We used `TCGAbiolinks` package to obtain miRNA expression quantification.
+We have used `TCGAbiolinks` package to obtain miRNA expression quantification.
 The query gives read count of miRNA as isoform chromosome coordination.

### 2019-04-10 article.Rmd
article rev

@@
 Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5’UTR bind to specific sequence located in 3’UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005].
-In recent studies, it has been shown miRNAs can interact with targets via sequence that located in different region such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013]moore_mirnatarget_2015.
+In recent studies, it has been shown miRNAs can interact with targets via sequence that located in different region such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013] **moore_mirnatarget_2015**.
 These studies also shown that binding location could indicate functionality of miRNA:target interaction or affect profile of target abundance.
@@
 
-![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow.jpg){#fig3}
+### Application of model in a breast cancer patient dataset
 
-### Defined functions for ceRNA models
-We defined the functions that can be used with R programming.
-Briefly, these functions process a given miRNA:gene dataset and convert to graph object.
-All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
-The functions and steps of approach are explained as following (\autoref{fig3}) :
-*Convertion of dataset*: priming_graph function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
-If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
-The formulations that are given in equations \eqref{eq:1}-\eqref{eq:5} are performed in this function.
-This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
-It is assumed that the initial target amounts in the dataset is observed after the reppressive activity of miRNAs in steady-state.
-*Transition of variables in graph*: In the previous step, the calculations are performed in the edge variables of the graph object.
-However, the graph object allows to use node variables, while the node features are handled to the graph.
-In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.
-*Trigger change in graph*: The dataset are assumed as steady-state in previous step and the efficieny coefficients are calculated according to this acceptance.
-In the network that is found in steady-state conditions, the change is applied to the graph object for disturbtion of steady-state.
-To provide the disturbtion in the network the workflow offer two methods: update_variables and update_how.
-The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
-The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
-*Updating current values of variables*: After variables updating in edge varibles, these are carried to node variables.
-Current and previous values of variables are stored as node variables with update_variables function.
-*Simulation of competing behavior of targets*: After the change in the steady-state conditions, the network elements try to gain steady-state again.
-This process progresses as repeating of regulations after the spreading the changes in the network.
-In this step, simulation of regulations according to given cycle count in simulate function is applied.
-After each simulation cycle, the miRNA repression values are re-calculated and the current values of competing elements are found and saved.
-The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node data.
-The node elements in the dataset are handled as two type; repressives (miRNAs) and competings (targets).
-It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used.
-If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user.
-On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger.
-The functions that are used in the approach are developed with R programming so as can be used with other packages.
-These are can be found in the github repository [ceRNAnetsim github page](https://github.com/selcenari/ceRNAnetsim) and improved with contributions of others.
+To determine suitability of method, we have tried to apply model in a real dataset.
+Firstly, we obtained the miRNA and gene expression datasets of the patient's tumor and normal tissue.
+We downloaded the high-throughtput experimental datasets which are provided miRNA:gene target pairs with interaction factors by **moore_mirnatarget_2015** and [@helwak_mapping_2013].
+We have combined miRNA and gene expression datasets via miRNA:target gene dataset **(Whole steps can be found in Supplementary data(TCGA_E9-A1N5_article.Rmd)**).
+We selected a gene which is over expressed in tumor tissue with compared to normal tissue.
+It is pointed out that ABCC1 gene, *Multidrug resistance-associated protein 1*, is one of the most significant factor to develop resistance against chemotherapotic agents[@atalay2008role; @lu2015microrna; @atalay2006multidrug].
+So, we selected ABCC1 gene for simulation of integrated dataset.
+We have detected the iteration that is suitable for ABCC1 gene simulation.
+After simulation of system via ABCC1 gene with over expression, we have compared simulation results and tumor tissue expression levels.
 
@@
 d) Regulation of gene expression of other gene group via triggering target (common target between M1 and M2).
-](Fig4.jpg){#fig4}
+](Fig3.jpg){#fig3}
 
@@
 For these reasons, we have improved the model with optional interaction parameters that could be useful for understanding of miRNA repression activity.
-We have explained the model in the same sample dataset with using different factors and simulated via the same change (\autoref{fig4}).
-When the factors were taken into account in the system, miRNA efficiencies varied as shown in \autoref{fig4}a.
+We have explained the model in the same sample dataset with using different factors and simulated via the same change (\autoref{fig3}).
+When the factors were taken into account in the system, miRNA efficiencies varied as shown in \autoref{fig3}a.
 Although the miRNA target ratios in initial states were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed.
-When expression of a gene (G2) increased, expression values of all genes also change differentially because of contribution of efficiency factors (\autoref{fig4} b,c,d).
+When expression of a gene (G2) increased, expression values of all genes also change differentially because of contribution of efficiency factors (\autoref{fig3} b,c,d).
 
-In regulation of miRNA target sample system (\autoref{fig4}), miRNA (M1) repressive efficiency on the primary triggering gene (G2) is low in steady-state.
+In regulation of miRNA target sample system (\autoref{fig3}), miRNA (M1) repressive efficiency on the primary triggering gene (G2) is low in steady-state.
 So it has been observed that the regulatory activities of miRNA efficiencies on the targets are weak after the increase of miRNA target G2.
@@
 Furthermore, folding of expression of target gene that has strong miRNA repression efficiency also resulted in evident perturbation.
+On the other hand, It was observed that Gene2 was weakly affected from change of Gene4 expression because of its weak interaction factors.This also shows that the perturbation iniated element in the system is important parameter.
+So, we also have developed an approach that is found perturbation efficiencies of elements in dataset.
+**(see details in Supplementary File(fig1_2app.Rmd)**.
+When we applied the method on the minimal dataset, while Gene4 is the most efficient element in terms of number of perturbed elements, it have observed that Mir2 is most potent element in the mean of expression change.
 
@@
 We have tried to simulate a real dataset which contains thousands of genes and hundreds of miRNAs.
-Briefly, we have obtained the datasets and combined these via R programming.
-We have applied the our approach into integrated dataset using a selected gene.
 Firstly, we have observed the activities in the iterations and determined behaviours of the network after simulation.
-As expected, we have observed that the big dataset, compared to those containing a few genes and miRNAs, exhibited cumbersome behaviours and complexity to gain steady-state.
+As expected, we have observed that the big dataset, compared to those containing a few genes and miRNAs, exhibited cumbersome behaviours and complexity to gain steady-state **(see details in Supplementary File(TCGA_E9-A1N5_article.Rmd)**.
 We realized that a small change in a system with a large change actually brought many arrangements.
 This can be explained with the competing groups.
-Because the genes are targeted with many common miRNAs, and this results in that the common targets exhibits competing behaviours in different groups.
+Because the genes are targeted with many common miRNAs, and this results in that the common targets exhibit competing behaviours in different groups.
 On the other hand, the changes of gene expressions were very different.

### 2019-04-10 fig1_2app.Rmd
article rev

@@
 title: "supplement"
-author: "Selcen Arı"
+author: "Selcen Ari"
 date: "7 Nisan 2019"
@@
 ---
+
+### Defined functions for ceRNA models and workflow of method
+We defined the functions that can be used with R programming.
+Briefly, these functions process a given miRNA:gene dataset and convert to graph object.
+All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
+The functions and steps of approach are explained as following (Figure 1) :
+
+*Convertion of dataset*: priming_graph function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
+If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
+The formulations that are given in equations (1-5) are performed in this function.
+This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
+It is assumed that the initial target amounts in the dataset is observed after the reppressive activity of miRNAs in steady-state.
+
+![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow1.jpg){#fig1}
+
+*Transition of variables in graph*: In the previous step, the calculations are performed in the edge variables of the graph object.
+However, the graph object allows to use node variables, while the node features are handled to the graph.
+In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.
+
+*Trigger change in graph*: The dataset are assumed as steady-state in previous step and the efficieny coefficients are calculated according to this acceptance.
+In the network that is found in steady-state conditions, the change is applied to the graph object for disturbtion of steady-state.
+To provide the disturbtion in the network the workflow offer two methods: update_variables and update_how.
+The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
+The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
+
+*Updating current values of variables*: After variables updating in edge varibles, these are carried to node variables.
+Current and previous values of variables are stored as node variables with update_variables function.
+
+*Simulation of competing behavior of targets*: After the change in the steady-state conditions, the network elements try to gain steady-state again.
+This process progresses as repeating of regulations after the spreading the changes in the network.
+In this step, simulation of regulations according to given cycle count in simulate function is applied.
+After each simulation cycle, the miRNA repression values are re-calculated and the current values of competing elements are found and saved.
+The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node data.
+
+The node elements in the dataset are handled as two type; repressives (miRNAs) and competings (targets).
+It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used.
+If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user.
+On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger.
+The functions that are used in the approach are developed with R programming so as can be used with other packages.
+These are can be found in the github repository [ceRNAnetsim github page](https://github.com/selcenari/ceRNAnetsim) and improved with contributions of others.
+
+```{r, message=FALSE, warning=FALSE}
+library(ceRNAnetsim)
+```
+
++ load *minsamp* data
 
@@
 
-```{r}
+## *minsamp* dataset analysis in lack of interaction factors.
+
+Firstly, we have analysed minimal data without interaction factors between miRNA:target.
+
+- 1. We have evaluated graph in the steady state conditions as followings:
+
+```{r, message=FALSE, warning=FALSE}
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
+  update_nodes(once=TRUE)%>%
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset in steady-state conditions")
+```
+
+- 2. We have obtained graph after change on Gene2 expression as followings:
+
+```{r, message=FALSE, warning=FALSE}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
@@
   update_nodes()%>%
-  simulate(3)%>%activate(edges)%>%
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors")
+```
+
+- 3. We have determined regulations after Gene2 Upregulation:
+
+```{r, message=FALSE, warning=FALSE}
+
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
+  update_nodes(once=TRUE)%>%
+  update_how("Gene2", 2)%>%
+  update_nodes()%>%
+  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "Regulations after Gene2 Upregulation", 2)
+
+```
+Note that the regulations are colored according to expression changes of present and a previous value.
+So, it can be observed that whole gene expressions increase in comparison of initial steady-state.
+The overall regulations of gene expressions are as followings:
+
+```{r, message=FALSE, warning=FALSE}
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
+  update_nodes(once=TRUE)%>%
+  update_how("Gene2", 2)%>%
+  update_nodes()%>%
+  simulate(2)%>%
+  activate(edges)%>%
   as_tibble()%>%
@@
   unnest()
+```
+
+## *minsamp* dataset analysis with interaction factors.
+
+We have made the same analysis in present of interaction factors.
+
+```{r, message=FALSE, warning=FALSE}
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+  update_nodes(once=TRUE)%>%
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset in steady-state conditions")
 
@@
 
-```{r}
+```{r, message=FALSE, warning=FALSE}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
@@
   update_nodes()%>%
-  simulate(3)%>%activate(edges)%>%
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors")
+```
+
+```{r, message=FALSE, warning=FALSE}
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+  update_nodes(once=TRUE)%>%
+  update_how("Gene2", 2)%>%
+  update_nodes()%>%
+  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors",2)
+```
+
+When the graphs which were resulted from analyses were examined, it was observed that behaviours were same.
+But, when the results were analysed in terms of expression values, the regulation differences can be observed.
+
+```{r, message=FALSE, warning=FALSE}
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+  update_nodes(once=TRUE)%>%
+  update_how("Gene2", 2)%>%
+  update_nodes()%>%
+  simulate(3)%>%
+  activate(edges)%>%
   as_tibble()%>%
@@
 ```
+
+## Common target perturbation in *minsamp* dataset.
+
+There are hundreds of defined miRNAs for human, so this results in presence of common targets of miRNAs in cells.
+Therefore, we have analysed perturbation efficiency of common target in *minsamp* dataset.
+
+```{r, message=FALSE, warning=FALSE}
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+  update_nodes(once=TRUE)%>%
+  update_how("Gene4", 2)%>%
+  update_nodes()%>%
+  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors",2)
+```
+
+The common target perturbation  (increasing to two fold at Gene4 expression in presence of interaction factors) resulted in more prominent efficiency at the same conditions (shown in following).
+
+```{r, message=FALSE, warning=FALSE}
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+  update_nodes(once=TRUE)%>%
+  update_how("Gene4", 2)%>%
+  update_nodes()%>%
+  simulate(3)%>%
+  activate(edges)%>%
+  as_tibble()%>%
+  select(Competing_name,comp_count_list,effect_list)%>%
+  unnest()
+```
+
+## Determination of perturbation efficiencies efficiencies of elements in system.
+
+```{r, message=FALSE, warning=FALSE}
+
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)-> sample_graph
+
+find_node_perturbation(sample_graph, how = 2, cycle = 3, limit = 0.1)
+
+```

### 2019-04-10 ceRNAnetsim/R/find_iteration.R

@@ -1,0 +1,156 @@
+#' @title find_iteration
+#'
+#' @description find_iteration searches the iteration that provides maximum affected node number. The user defines a symbolic iteration with .iter. The function calculates the number of affected nodes for each iteration and than selects the iteration that has maximum affected nodes' number.
+#'
+#' @return It gives an iteration number to use in simulate() function.
+#'
+#' @param df A data frame that includes the miRNA and competing targets.
+#' @param competing_count The counts (expression) of competing elements of the dataset.
+#' @param miRNA_count The counts (expression) of repressive element (miRNA) of the dataset.
+#' @param aff_factor The parameter/s of binding between miRNA and targets.
+#' @param deg_factor The parameter/s for degradation of bound miRNA:target complex.
+#' @param node_name The name f node that is used as trigger.
+#' @param how The change of count (expression) of the given node in terms of fold change.
+#' @param .iter The maximum iteration prescribed by the user.
+#' @param limit The minimum amount of change of any node.
+#'
+#' @details
+#'
+#' @examples
+#'
+#' find_iteration(minsamp, Competing_expression, miRNA_expression, aff_factor = energy, deg_factor = region, node_name = "Gene2", how= 3)
+#'
+#' @export
+
+
+find_iteration <- function(df, competing_count, miRNA_count, aff_factor=dummy, deg_factor=dummy, node_name, how, .iter = 100, limit= 0.1){
+
+  competing_exp <- enquo(competing_count)
+  mirna_exp <- enquo(miRNA_count)
+  affinity <- enquos(aff_factor)
+  degradation <- enquos(deg_factor)
+
+  iteration <- data.frame(iter = seq(1,.iter, 1), effect= rep(0))
+
+  df <- df%>%
+    dplyr::mutate(competing = df[,1], miRNA= df[,2], Competing_name = df[,1], miRNA_name= df[,2], dummy=1)%>%
+    dplyr::select(competing, miRNA, Competing_name, miRNA_name, !!competing_exp, !!mirna_exp, !!!affinity, !!!degradation, dummy)
+
+
+  df%>%
+    dplyr::group_by(miRNA)%>%
[…]

### 2019-04-10 ceRNAnetsim/R/priming_graph.R

@@ -1,0 +1,79 @@
+#' @title normalize
+#'
+#' @description normalizes the values according to maximum values inside a group. The helper function of priming_graph.
+#'
+#' @param x The variable name that is normalized.
+#'
+#' @return NULL
+#'
+
+
+normalize <- function(x){abs(x)/max(abs(x))}
+
+#' @title priming_graph
+#'
+#' @description The function converts the given dataframe using first variable as competing and the second as miRNA. If user defines interaction factors as affinity or degradation, the factors are taken into account.
+#'
+#' @return the graph object.
+#'
+#' @param df A data frame that includes the miRNA and competing targets.
+#' @param competing_count The counts (or expression) of competing elements of the dataset.
+#' @param miRNA_count The counts (or expression) of repressive element (miRNA) of the dataset.
+#' @param aff_factor The parameter/s of binding between miRNA and targets.
+#' @param deg_factor The parameter/s for degradation of bound miRNA:target complex.
+#'
+#' @details priming_graph provides grouping of competing targets and evaluation of targets within the groups taking into account miRNA:target, target:total target, interaction and degradation parameters. The target groups are determined according to miRNAs. If the factors that are important in target interactions are specified as arguments, the factors also are evaluated separately within each group.
+#' priming_graph also calculates the miRNA efficiency in steady-state conditions. It is assumed that quantity of competing targets and miRNAs are shown in the steady-state system after the miRNAs exhibit repressive efficiency.
+#' Note that the data must not include missing values such as NA or "-".
+#'
+#'
+#' @examples
+#'
+#' priming_graph(minsamp, Competing_expression, miRNA_expression, aff_factor = c(seed_type,energy), deg_factor = region)
+#'
+#' @export
+
+priming_graph <- function(df, competing_count, miRNA_count, aff_factor=dummy, deg_factor=dummy){
+
+
+  if(sum(is.na.data.frame(df)) !=0){
+
+    stop("Missing or NA value in dataframe")
+  }
+
+  else{
+
+  competing_exp <- enquo(competing_count)
+  mirna_exp <- enquo(miRNA_count)
[…]

### 2019-04-10 ceRNAnetsim/R/simulate_vis.R

@@ -1,0 +1,57 @@
+#' @title simulate_vis
+#'
+#' @description  simulate_vis provides visualisation of the graph in addition to simulate function.
+#'
+#' @details simulate_vis gives the last graph object and each iterations' image.
+#'
+#' @return It gives a graph and the images of states in each iteration until the end of the simulation.
+#'
+#' @param input_graph The graph object that processed in previous steps.
+#' @param cycle Optimal iteration number for gaining steady-state.
+#' @param Competing_color The color of competing elements on the graph with "green" default.
+#' @param mirna_color The color of miRNAs on the graph with "orange" default.
+#' @param Upregulation The color of Upregulated elements on the graph with "red" default.
+#' @param Downregulation The color of Downregulated elements on the graph with "blue" default.
+#' @param title Title of the given graph.
+#' @param layout The layout that will be used for visualisation of the graph.
+#'
+#' @examples
+#'
+#' # When does the system gain steady-state conditions again?
+#'
+#' ## new_counts, the dataset that includes the current counts of nodes.
+#'
+#' priming_graph(minsamp, Competing_expression, miRNA_expression, aff_factor = c(seed_type,energy), deg_factor = c(region))%>%
+#'    update_nodes(once = TRUE)%>%
+#'    update_variables(new_counts)%>%
+#'    update_nodes()%>%
+#'    simulate_vis(cycle = 12)
+#'
+#' @export
+
+
+simulate_vis <- function(input_graph, cycle=1, Competing_color = "green", mirna_color = "orange", Upregulation = "red", Downregulation = "blue", title = "GRAPH", layout= "kk"){
+
+  for(i in seq_along(1:cycle)){
+
+    input_graph%E>%
+      group_by(to)%>%
+      mutate(mirna_count_per_comp = mirna_count_current*comp_count_current*afff_factor/sum(comp_count_current*afff_factor), mirna_count_per_comp = ifelse(is.na(mirna_count_per_comp), 0, mirna_count_per_comp))%>%
+      ungroup()%>%
+      mutate(effect_pre = effect_current, effect_current = mirna_count_per_comp*degg_factor)%>%
+      group_by(from)%>%
[…]

### 2019-04-10 ceRNAnetsim/tests.R

@@ -1,0 +1,2 @@
+library("testthat")
+test_check("ceRNAnetsim")

### 2019-04-10 ceRNAnetsim/tests/testthat/test-data_convertion.R

@@ -1,0 +1,81 @@
+context("Convertion to the graph.")
+
+
+testdata <- data.frame(competing = c("TP53", "Gene2", "ABCC1", "Gene4", "Gene4", "Gene5", "PTEN"),
+                       miRNA = c("mir1", "mir1","hsa-mir3", "mir1", "MiR2", "MiR2", "miR4"),
+                       Competing_expression = c(10000, 10000, 5000, 10000, 10000, 5000, 10000),
+                       miRNA_expression = c(1000, 1000, 1000, 1000, 2000, 2000, 2000),
+                       stringsAsFactors = FALSE)
+
+testdata2 <- data.frame(competing = c("TP53", "Gene-2", "ABCC1", "Gene 4", "Gene 4", "Gene5", "PTEN"),
+                        miRNA = c("mir1", "mir1","hsa-mir3", "mir1", "MiR2", "MiR2", "miR4"),
+                        Competing_expression = c(10000, 10000, 5000, 10000, 10000, 5000, 10000),
+                        miRNA_expression = c(1000, 1000, 1000, 1000, 2000, 2000, 2000),
+                        stringsAsFactors = FALSE)
+
+testdata3 <- data.frame(competing = c("TP53", "Gene-2", "ABCC1", "Gene 4", "Gene 4", "Gene5", "PTEN"),
+                        miRNA = c("mir1", "mir1","hsa-mir3", "mir1", "MiR2", "MiR2", "miR4"),
+                        Competing_expression = c(NA, 10000, 5000, 10000, 10000, 5000, 10000),
+                        miRNA_expression = c(1000, 1000, 1000, 1000, 2000, 2000, 2000),
+                        stringsAsFactors = FALSE)
+
+test_that("Can dataset be correctly converted to graph ? ", {
+
+
+ prime_type_1 <-  (priming_graph(testdata, competing_count = Competing_expression, miRNA_count= miRNA_expression)%N>%
+      as_tibble()%>%
+      filter(type == "miRNA")%>%
+      count())%>%
+   pull()
+
+ prime_type_2 <-priming_graph(testdata, competing_count = Competing_expression, miRNA_count= miRNA_expression)%N>%
+     as_tibble()%>%
+     filter(type == "Competing")%>%
+     count()%>%
+   pull()
+
+
+ prime_type_3 <-  (priming_graph(testdata2, competing_count = Competing_expression, miRNA_count= miRNA_expression)%N>%
+                     as_tibble()%>%
+                     filter(type == "miRNA")%>%
[…]

### 2019-04-10 ceRNAnetsim/tests/testthat/test-updating_variables.R

@@ -1,0 +1,71 @@
+context("Check to Update")
+
+
+
+test_that("Check the functions which update the values on the datasets.", {
+
+  data("midsamp")
+
+  primerg<- midsamp%>%
+    priming_graph(competing_count = Gene_expression, miRNA_count = miRNA_expression)
+
+  seconderg <- primerg%>%
+    update_nodes(once= TRUE)
+
+  thirdg<- seconderg%>%
+    update_how(sample_n(midsamp, 1)$Genes, 2)
+
+  forth<- thirdg%>%
+    update_nodes()
+
+
+  expect_equal(ncol(as_tibble(primerg%>%activate(nodes))), 3)
+  expect_equal(ncol(as_tibble(seconderg%>%activate(nodes))), 7)
+  expect_equal(ncol(as_tibble(thirdg%>%activate(nodes))), 7)
+  expect_equal(ncol(as_tibble(forth%>%activate(nodes))), 7)
+  expect_equal(ncol(as_tibble(seconderg%>%activate(edges))), 19)
+  expect_equal(ncol(as_tibble(thirdg%>%activate(edges))), 19)
+  expect_equal(ncol(as_tibble(forth%>%activate(edges))), 19)
+
+
+})
+
+
+test_that("Check the functions which update the values on the datasets with interaction factors.", {
+
+  data("midsamp")
+
+  primergf <- midsamp%>%
+    priming_graph(competing_count = Gene_expression, miRNA_count = miRNA_expression, aff_factor = c(seeds, Energy), deg_factor = targeting_region)
+
+  secondergf <- primergf%>%
+    update_nodes(once= TRUE)
+
+  thirdgf <- secondergf%>%
+    update_how(sample_n(midsamp, 1)$Genes, 2)
+
+  forthf <- thirdgf%>%
+    update_nodes()
+
+  midsamp%>%
+    priming_graph(competing_count = Gene_expression, miRNA_count = miRNA_expression)%>%
+    update_nodes(once=TRUE)%>%
+    update_how("Gene15", 0.3)%>%
+    update_nodes()%>%
+    simulate(10)%>%
+    as_tibble()%>%
+    dplyr::mutate(count_current= round(count_current, 2))%>%
+    dplyr::select(count_current)%>%
+    pull() -> outputall
+
+
[…]

### 2019-04-11 article.Rmd
article rev

@@
 output:
-  pdf_document: default
+  pdf_document: 
   html_document: default
 bibliography: ref.bib
-biblio-style: NAR-natbib.sty
 ---
@@
 ## Abstract
+
+Competing endogeneous RNA (ceRNA) regulations and crosstalk in human are remarkable issue in mean of regulations via miRNAs.
+Many studies have pointed out that an alteration in miRNA:target interaction can result in unexpected regulations because of indirect and complex interaction.
+In this paper, we defined a new network-based model that handles miRNA:ceRNA interactions with expression values and interaction factors optionally.
+We applied approach in sample dataset and defined dynamics of it.
+We have observed that a perturbation in an element of network affects whole competing elements differently and cooperative efficiencies of miRNAs on common targets could be shown.
+Importance of miRNA:target ratios was also observed consistently with previous reports.
+We have obtained that the competing elements which have the same or close expression values may not affected equal from the perturbation because of repression functionality of interactions.
+To apply model to the real sample that obtained combining  breeast cancer gene and miRNA expression dataset via high throughput experimental miRNA:target dataset, we selected a gene, *ABCC1*, which over expressed in tumor tissue and utilized it as perturbating element.
+We have observed that one change can be sufficient to perturb regulations, and may not predicted all of regulations correctly, with this change.
+Therefore, this model which helps to understand crosstalking between elements could be tested and improved via results of well-constructed experimental dataset and takes to hope for future research.
 
@@
 When expression of a gene (G2) increased, expression values of all genes also change differentially because of contribution of efficiency factors (\autoref{fig3} b,c,d).
+We have considered that in our approach energy and seed type of pairs is significant for binding and targeted region is important for repression.
+But these factors can be handled optionally and differently.
 
@@
 Furthermore, folding of expression of target gene that has strong miRNA repression efficiency also resulted in evident perturbation.
-On the other hand, It was observed that Gene2 was weakly affected from change of Gene4 expression because of its weak interaction factors.This also shows that the perturbation iniated element in the system is important parameter.
+On the other hand, It was observed that Gene2 was weakly affected from change of Gene4 expression because of its weak interaction factors.This also shows that the perturbation-initiation element in the system is important parameter.
 So, we also have developed an approach that is found perturbation efficiencies of elements in dataset.
@@
 When we applied the method on the minimal dataset, while Gene4 is the most efficient element in terms of number of perturbed elements, it have observed that Mir2 is most potent element in the mean of expression change.
+In our approach, the elements behave differently based on type (competing or miRNA).
+When a competing element was changed, whole elements out of miRNAs are re-regulated.
+But, if a miRNA was perturbed, whole competing elements are re-regulated and miRNA preserves expression at after perturbation.
 
@@
 While the change in some of genes were highly evident, somes have changed slightly.
-So, we have compared two datasets: the tumor gene expression and simulation results.
+So, we have compared two datasets; the tumor gene expression and simulation results.
 When we analysed comparison of two datasets, we have observed that the results of simulation is not convenient with tumor tissue expression values.
@@
 Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored.
-Of course, it would be more useful to test a dataset that includes gene expression values before and after the regulation of a gene.
+Of course, it would be more useful to test a dataset that includes gene expression values before and after the regulation of a gene and miRNA expression values at initial conditions.
 When we evaluated the suitability of approach, the approach allows different miRNA targets to be considered together and can work with different regulatory elements due to the network structure.
@@
 They have been demonstrated that distant ceRNAs can interact with each other via indirect links, and the interactions change to depend on distance between ceRNAs, and topological features of network [@nitzan_interactions_2014].
+List et al.
+have developed an approach to detect ceRNA interaction by using the miRNA expression, gene expression and common miRNAs between gene targets [@markus_list_sponge_2017].
+Their approach could be useful to analysis interacted genes through miRNAs.
+When it has been throught that a miRNA can exhibit strong functionality to a target but may not against an other, common miRNA based approach can be inconvenient to understand regulations of ceRNA interactions.
 

### 2019-05-26 article.Rmd
fix problematic reference moore

@@
   pdf_document: 
+    keep_tex: true
   html_document: default
@@
 Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5’UTR bind to specific sequence located in 3’UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005].
-In recent studies, it has been shown miRNAs can interact with targets via sequence that located in different region such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013] **moore_mirnatarget_2015**.
+In recent studies, it has been shown miRNAs can interact with targets via sequence that located in different region such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013; @moore_mirnatarget_2015].
 These studies also shown that binding location could indicate functionality of miRNA:target interaction or affect profile of target abundance.

### 2019-05-27 article.tex
untrack, remove log file

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Network based multifactorial modelling microRNA-target interaction.},
            pdfauthor={Selcen Ari; Alper Yilmaz},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Network based multifactorial modelling microRNA-target interaction.}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Selcen Ari\footnote{Yildiz Technical University Department of
  Bioengineering} \\ Alper Yilmaz\footnote{Yildiz Technical University Department of
  Bioengineering}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{17 May 2019}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Competing endogeneous RNA (ceRNA) regulations and crosstalk in human are
remarkable issue in mean of regulations via miRNAs. Many studies have
pointed out that an alteration in miRNA:target interaction can result in
unexpected regulations because of indirect and complex interaction. In
this paper, we defined a new network-based model that handles
miRNA:ceRNA interactions with expression values and interaction factors
optionally. We applied approach in sample dataset and defined dynamics
of it. We have observed that a perturbation in an element of network
affects whole competing elements differently and cooperative
efficiencies of miRNAs on common targets could be shown. Importance of
miRNA:target ratios was also observed consistently with previous
reports. We have obtained that the competing elements which have the
same or close expression values may not affected equal from the
perturbation because of repression functionality of interactions. To
apply model to the real sample that obtained combining breeast cancer
gene and miRNA expression dataset via high throughput experimental
miRNA:target dataset, we selected a gene, \emph{ABCC1}, which over
expressed in tumor tissue and utilized it as perturbating element. We
have observed that one change can be sufficient to perturb regulations,
and may not predicted all of regulations correctly, with this change.
Therefore, this model which helps to understand crosstalking between
elements could be tested and improved via results of well-constructed
experimental dataset and takes to hope for future research.

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

MicroRNAs (miRNAs) that are key regulator of gene expression through
various post-transcriptional mechanisms are a family of short non-coding
RNAs. Although repression mechanisms of miRNA are not fully understood,
exact effect of miRNAs is regulation of direct target/targets
negatively. Repressive activities of miRNAs vary depending on many
factors that are significant to microRNA:target interactions. These
factors include microRNA:target binding energy, binding location in
target sequence, base pairing types between miRNA and target, abundance
of miRNAs and targets (Grimson et al. 2007). Binding energies of
miRNA:target complexes vary based on nucleotide context and determine
folding stability of complex (Cao and Chen 2012). It has been
demonstrated that the binding energy between miRNA and target indicates
stability or affinity of complex (Helwak et al. 2013) and does not
directly determine repressive activity of miRNA (Cao and Chen 2012).
Early studies have argued that 2-8 nt sequence, seed, located in miRNA
5'UTR bind to specific sequence located in 3'UTR of its target (Bartel
2004; Lewis, Burge, and Bartel 2005). In recent studies, it has been
shown miRNAs can interact with targets via sequence that located in
different region such as 5'UTR or CDS (Hausser et al. 2013; Helwak et
al. 2013) \textbf{moore\_mirnatarget\_2015} (Moore et al. 2015). These
studies also shown that binding location could indicate functionality of
miRNA:target interaction or affect profile of target abundance. It has
been shown that miRNAs exhibit repressive activity via, 6-8 nt long
sequence that is prefectly complementary with targets, seed region at
the 5'UTR of miRNAs (Bartel 2009; Grimson et al. 2007). On the other
hand, some researchers have reported that seed sequence of miRNA can
have mismatches or bulged/wobble nucleotides and locates in different
region of miRNAs (Hafner et al. 2010; Helwak et al. 2013). In addition
to all of these, abundance of miRNAs and targets and miRNA:target ratio
in cells are determinant for efficiency of miRNA:target interaction
(Arvey et al. 2010; Bosson, Zamudio, and Sharp 2014; Denzler et al.
2014).

As it is possible for miRNAs to suppress multiple targets, an individual
RNA molecule can also be targeted by various miRNAs. In that case, the
targeted RNAs exhibit competitor behavior, that is hypothesized as
competing endogenous RNAs (ceRNAs) (Ala et al. 2013; Cesana and Daley
2013), against their miRNAs. Briefly, Ala et al.~have explained the
ceRNA hypothesis as disturbtion of the other target when a steady-state
system that included one miRNA and two target, was perturbed with
expression changes at one of the targets (Ala et al. 2013). When
interaction between miRNAs and their target were dealed in the cell,
explanation and prediction of results of perturbation is difficult due
to varieties of interactions. Various computational and experimental
studies have applied to be understood for ceRNA:miRNA interactions. It
has been observed that when abundance of one of the targets of miR-122
was increased, the other target expression also slightly increased as a
result of decreasing repressive activity of miR-122 on them (Denzler et
al. 2014). Bosson et al.~have developed a mathematical model for changes
on total target pool concentration after grouping to affinity of
targets. They have demonstrated that miRNA activity correlated with
affinity between miRNA and target (Bosson, Zamudio, and Sharp 2014).
Cooperative efficiency of miRNAs as well as competitor behaviors of
targets also were studied and it has been demonstrated crucial for
regulations at expression of targets (Denzler et al. 2016). MiRNA:target
interactions have modelled as stoichiometric and catalytic mechanisms
and Figliuzzi et al.~have recommended to handling models in network
structures (Figliuzzi, Marinari, and De Martino 2013). The model that
can explain miRNA target interaction through topological features has
been applied at bipartite network by Nitzan et al.~(Nitzan et al. 2014).
Robinson and Henderson applied the model that handles miRNA:target
direct and indirect interactions via common miRNA of gene and target of
miRNA, at bipartite network. It has been demonstrated that whole miRNAs
and targets in network can interacted through common miRNAs and genes
(Robinson and Henderson 2018). Associated genes that are targets of the
same miRNAs have been found with help of correlation of gene expression
changes in recent study (Markus List 2017). List et al.~have specified
that their approach can be useful for ceRNA's studies and published
these approach as a package that can use with R programming.

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{mirnatarget-ratio-theoretical-action-in-network-based-cerna-model}{%
\subsubsection{miRNA:target ratio theoretical action in network based
ceRNA
model}\label{mirnatarget-ratio-theoretical-action-in-network-based-cerna-model}}

We developed a novel network based approach that can use to simulation
of gene regulation after any perturbation at miRNA:target interactions.
If a perturbation or change induces on graph, whole interacted nodes
vary because of changes in miRNA:target ratios. The system was assumed
as steady state that miRNAs affect according to the ratio of the
individual target to all targets (\autoref{fig1}a, expression values
shown as integer for comprehensible sampling). When expression of gene 2
(G2) increased to two fold (\autoref{fig1}b), distribution of miRNA (M1)
expression changes (\autoref{fig1}c). Consequently, expression of common
target (G4) of miRNA1 and miRNA2 also varies. The change at expression
of common target triggers to changes of proportional distribution of
miRNA2 (\autoref{fig1}d).

\begin{figure}
\hypertarget{fig1}{%
\centering
\includegraphics{Fig1.jpg}
\caption{Shematic presentation of mechanism of network based model. a)
In steady state, miRNAs (M: triangle grey) repress targets (G: circle
coconut) according to proportion of their targets' expression. b-c-d)
MiRNA expression distributions to targets vary when expression of one of
the targets perturbed.}\label{fig1}
}
\end{figure}

The model was developed by making some assumptions. Transcription rates
and degradation rates of miRNAs were assumed as stationary and equal.
Targets also have stabil transcription and degradation rates, at the
same time they are equal. The factors that could be affect the
miRNA:target interaction are disregarded in this approach. It is
considered that miRNAs degradates or inhibits them via any mechanisms
after binding to target while the miRNAs were used reversibly. In
initial conditions, gene expression amounts in system have been
considered after repression activity of miRNAs. If a change occurs in
the system, variation on miRNA proportional distribution are reflected
as change of repression activity. \begin{equation} 
    Eff_{gi}= C_m * C_{gi}/\sum_{1}^{i} C_g \tag{1}\label{eq:1}
\end{equation} \begin{equation}
   R_{gi}= Eff_{i1}+ Eff_{i2} ... \tag{2}\label{eq:2}
\end{equation}

The efficiency of a miRNA on the individual target (\(Eff_{gi}\)) is
calculated according to equation \eqref{eq:1}; \(C_m\), miRNA expression
in the group; \(C_{gi}\), symbolize individual gene expression; \(C_g\)
gene expressions in the group (Groups are determined depending on
miRNAs. Targets that are interacted with the same miRNA find in a
group.). In this approach, because a gene can be found in different
groups, the cooperative activity of microRNAs also can be calculated.
Total repression activity of different miRNAs on the common target,
\(R\), is calculated based on equation \eqref{eq:2}; sum of efficiency
of different miRNAs on the target.

\begin{figure}
\hypertarget{fig2}{%
\centering
\includegraphics{fig2.jpg}
\caption{Calculations to determine of miRNA binding and repression
efficiency. \(G\), Gene; \(M\), miRNA; \(STE\), seed type effect;
\(RE\), Region Effect; \(E\), Energy; \(STE’\), normalized values of
seed type efficiency coefficient; \(RE’\), normalized values of region
efficiency coefficient; \(E’\), normalized values of energy
coefficient.}\label{fig2}
}
\end{figure}

\hypertarget{mirna-action-in-network-based-multifactorial-cerna-model}{%
\subsubsection{miRNA action in network based multifactorial ceRNA
model}\label{mirna-action-in-network-based-multifactorial-cerna-model}}

Interactions between miRNAs and their targets can be affected from
different parameters. For this reason, we improved model with factors
that can affect to miRNA activity. In our approach, the parameters are
classified as binding and efficiency factors and all of these are
evaluated in their groups. Binding factors determine interaction between
miRNA and target. The efficiency factors decide functionality of
binding. In the literature, binding free energy (Cao and Chen 2012 ;
Helwak et al. 2013) and seed type (Werfel et al. 2017) in miRNA:target
interactions are represented as binding affinity. Efficiency factors
determine how many of microRNA:target complexes will result in
inhibition. For example, binding region on the target shows whether the
target can be inhibited (functionality)(Hausser et al. 2013; Helwak et
al. 2013). Both of binding and efficiency factors are evaluated in the
groups and normalized according to equation \eqref{eq:3}; each factor is
normalized based on maximum variables that found in their groups.
\begin{equation}
F`= F/F_{max} \tag{3}\label{eq:3}
\end{equation}

The normalized values of factors take into account to determine binding
activity and miRNA efficiency on targets (\autoref{fig2}). Binding
affinities (activity, \(Eff\)) of miRNAs on each individual gene are
calculated as shown in equation \eqref{eq:4}; \(C_m\), miRNA expression
in the group; \(C_g\), Gene expression; \(C_{gi}\), individual gene
expression; gi, individual gene; g, whole genes in a group
(\autoref{fig2}c). \begin{equation}
Eff_{gi}= C_m * E`_{gi}*STE`_{gi}*C_{gi}/(\sum_{1}^{i} E`_{gi}*STE`_{gi}*C_{gi}) \tag{4}\label{eq:4}
\end{equation} \begin{equation}
Eff_{gi}= Eff_{gi}*RE`_{gi} \tag{5}\label{eq:5}
\end{equation}

After miRNA binds to its target, but might not repress to bound target.
The functionality of bound miRNA on target depends on efficiency factors
like region that is binding sequence of miRNA on its target. Exact
repression efficiency of miRNA is calculated according to equation
\eqref{eq:5} (\autoref{fig2}d); \(RE’_{gi}\), normalized values of
region efficiency coefficient between miRNA and gene. The cooperative
repression activity of miRNAs to their common targets is figured out as
shown in \autoref{fig2}e.

\hypertarget{application-of-model-in-a-breast-cancer-patient-dataset}{%
\subsubsection{Application of model in a breast cancer patient
dataset}\label{application-of-model-in-a-breast-cancer-patient-dataset}}

To determine suitability of method, we have tried to apply model in a
real dataset. Firstly, we obtained the miRNA and gene expression
datasets of the patient's tumor and normal tissue. We downloaded the
high-throughtput experimental datasets which are provided miRNA:gene
target pairs with interaction factors by
\textbf{moore\_mirnatarget\_2015} and (Helwak et al. 2013). We have
combined miRNA and gene expression datasets via miRNA:target gene
dataset \textbf{(Whole steps can be found in Supplementary
data(TCGA\_E9-A1N5\_article.Rmd)}). We selected a gene which is over
expressed in tumor tissue with compared to normal tissue. It is pointed
out that ABCC1 gene, \emph{Multidrug resistance-associated protein 1},
is one of the most significant factor to develop resistance against
chemotherapotic agents(Atalay, Demirkazik, and Gunduz 2008; Lu et al.
2015; Atalay et al. 2006). So, we selected ABCC1 gene for simulation of
integrated dataset. We have detected the iteration that is suitable for
ABCC1 gene simulation. After simulation of system via ABCC1 gene with
over expression, we have compared simulation results and tumor tissue
expression levels.

\hypertarget{results-and-conclusions}{%
\subsection{Results and Conclusions}\label{results-and-conclusions}}

We have developed a model approach and workflow for competitive ceRNA
regulation in this study. The basic mode of miRNA repression activity
has been based on miRNA and target abundance in various researches(Arvey
et al. 2010; Denzler et al. 2014). For this reason, we performed first
application as considering only abundance of miRNAs and targets
(\autoref{fig1}). After the increasing in a gene expression value,
expression values of other genes also changed slightly in comparison
with alteration in perturbation started gene node. Different studies
have shown that the other gene expression values also rise differently,
if a gene abundance increases in ceRNA system (Lai, Wolkenhauer, and
Vera 2016; Salmena et al. 2011; Tay, Rinn, and Pandolfi 2014). It was
observed that primary neighborhoods of the trigger gene change with the
same ratio, but cooperative efficiency of miRNAs on common target causes
differently change in common target (\autoref{fig1}d). The common target
in system has displayed like a trigger for the other group and induced
changes of expression values of genes on the other group. Therefore,
effect of primary change of the gene (G2) on genes of the other group
was observed slightly (\autoref{fig1}d). In addition, like shown by
ceRNA hypothesis model of Ala et al., after the increasing of a gene
expression (G2), the miRNA that is found in the same group (M1) tended
to be more repressive on this target. This has caused to decrease in
increased expression value of primer triggering gene. When the
expression of a gene (G2) was increased to two fold, we observed that
the expression changes were more evident. So, we have considered that
high miRNA:target ratio results in stronger miRNA repression activity.
This results also are coherent with previous reports that were offered
by Arvey et al.~and Denzler et al.(Arvey et al. 2010; Bosson, Zamudio,
and Sharp 2014; Denzler et al. 2014). When the simulation was run in
sample dataset that includes lower target abundance, the first response
of primary neihgborhood of the same group was determined as similar with
in system that have high target abudance. But regulations of ceRNAs were
observed differences in the folowing steps of simulation. In overall
state in the end of simulation, we have established that all regulations
more prominent when compared with low miRNA:target ratio dataset. The
model has appeared the importance of mRNA:target ratio in regulation of
competing RNAs.

\begin{figure}
\hypertarget{fig3}{%
\centering
\includegraphics{Fig3.jpg}
\caption{Target regulations with interaction parameters. a) In the
steady-state the repression activity of miRNAs on the targets after
binding and repression efficiency. b) The changes the repression
activities after increasing of G2 expression. c) Perturbation of primary
neighborhoods of M1 miRNA (M1 miRNA group). d) Regulation of gene
expression of other gene group via triggering target (common target
between M1 and M2).}\label{fig3}
}
\end{figure}

It has been reported in several studies that miRNA regulatory
interactions also is affected different parameters. For example, Xu et.
al have investigated the importance of seed pairing type between miRNAs
and their targets and target site location by using proteomic dataset
(Xu, Wang, and Liu 2014). They have proposed that the features of
binding between miRNA and target can be critical efficiency of miRNA and
can be used to find effective miRNA molecules. In addition, binding
energy between miRNA and targets also could be determinant for miRNA
efficiency. It has been reported by Breda et. al.~that strength of
miRNA:target interactions is depended on binding energy of complexes
(Breda et al. 2015). As similar to these, Bosson et al.~have associated
affinity and seed pairing of miRNA target pairs and suggested that
number of canonical base pairing is correlated with affinity. For these
reasons, we have improved the model with optional interaction parameters
that could be useful for understanding of miRNA repression activity. We
have explained the model in the same sample dataset with using different
factors and simulated via the same change (\autoref{fig3}). When the
factors were taken into account in the system, miRNA efficiencies varied
as shown in \autoref{fig3}a. Although the miRNA target ratios in initial
states were same in comparison with the sample dataset without factors,
efficiency of binding and repression have changed. When expression of a
gene (G2) increased, expression values of all genes also change
differentially because of contribution of efficiency factors
(\autoref{fig3} b,c,d). We have considered that in our approach energy
and seed type of pairs is significant for binding and targeted region is
important for repression. But these factors can be handled optionally
and differently.

In regulation of miRNA target sample system (\autoref{fig3}), miRNA (M1)
repressive efficiency on the primary triggering gene (G2) is low in
steady-state. So it has been observed that the regulatory activities of
miRNA efficiencies on the targets are weak after the increase of miRNA
target G2. When we ran to model with two fold increasing expression
value of common target (G4), the changes of other gene expressions have
observed more prominent. Furthermore, folding of expression of target
gene that has strong miRNA repression efficiency also resulted in
evident perturbation. On the other hand, It was observed that Gene2 was
weakly affected from change of Gene4 expression because of its weak
interaction factors.This also shows that the perturbation-initiation
element in the system is important parameter. So, we also have developed
an approach that is found perturbation efficiencies of elements in
dataset. \textbf{(see details in Supplementary File(fig1\_2app.Rmd)}.
When we applied the method on the minimal dataset, while Gene4 is the
most efficient element in terms of number of perturbed elements, it have
observed that Mir2 is most potent element in the mean of expression
change. In our approach, the elements behave differently based on type
(competing or miRNA). When a competing element was changed, whole
elements out of miRNAs are re-regulated. But, if a miRNA was perturbed,
whole competing elements are re-regulated and miRNA preserves expression
at after perturbation.

The defined miRNA' number is very low in comparison with defined protein
coding genes' number. So, this results in that the miRNA amounts in the
system must be high or targeted RNAs' amount must be low for effective
miRNA activity. We have tried to simulate a real dataset which contains
thousands of genes and hundreds of miRNAs. Firstly, we have observed the
activities in the iterations and determined behaviours of the network
after simulation. As expected, we have observed that the big dataset,
compared to those containing a few genes and miRNAs, exhibited
cumbersome behaviours and complexity to gain steady-state \textbf{(see
details in Supplementary File(TCGA\_E9-A1N5\_article.Rmd)}. We realized
that a small change in a system with a large change actually brought
many arrangements. This can be explained with the competing groups.
Because the genes are targeted with many common miRNAs, and this results
in that the common targets exhibit competing behaviours in different
groups. On the other hand, the changes of gene expressions were very
different. While the change in some of genes were highly evident, somes
have changed slightly. So, we have compared two datasets; the tumor gene
expression and simulation results. When we analysed comparison of two
datasets, we have observed that the results of simulation is not
convenient with tumor tissue expression values. The simulation results
were not expected to be concordant with tumor tissue expression values
because of a single gene could not be responsible for gene regulation in
all tissue. Besides, the all other factors such as up/down regulation of
miRNAs or other genes were ignored. Of course, it would be more useful
to test a dataset that includes gene expression values before and after
the regulation of a gene and miRNA expression values at initial
conditions. When we evaluated the suitability of approach, the approach
allows different miRNA targets to be considered together and can work
with different regulatory elements due to the network structure.

The network based approaches have been developed in previous studies.
Figliuzzi et al.~have tried to explain the ceRNA crosstalk in a
network-like minimal interaction structure with concentrations of ceRNA
and miRNAs. They have pointed out that the larger miRNA number can cause
to evident crosstalk between ceRNAs. It has been thought that miRNAs and
their targets interactions depend on rates of transcription,
degradation, binding and unbinding in the network based kinetic model
which has been developed by Nitzan et al.~Nitzan and collaborators have
used the high throughput experimental dataset (Helwak et al. 2013) for
miRNA:target pair that occur high free energy and microarray dataset of
a miR-92a depletion for expression values. They have been demonstrated
that distant ceRNAs can interact with each other via indirect links, and
the interactions change to depend on distance between ceRNAs, and
topological features of network (Nitzan et al. 2014). List et al.~have
developed an approach to detect ceRNA interaction by using the miRNA
expression, gene expression and common miRNAs between gene targets
(Markus List 2017). Their approach could be useful to analysis
interacted genes through miRNAs. When it has been throught that a miRNA
can exhibit strong functionality to a target but may not against an
other, common miRNA based approach can be inconvenient to understand
regulations of ceRNA interactions.

In our approach, we have not taken into account transcription,
degradation or binding rates of elements in network. Because, although
it is known as the miRNAs are highly stable, the transcription and
degradation rates of miRNAs change to depend on according to cellular
conditions (Rüegger and Großhans 2012). According to our search, there
is no dataset that was includes to degradation and transcription rates
on specific cellular conditions for each miRNA. So, we preferred to
accept that degradation and transcription rates of each miRNAs are equal
to each other, instead of determining a constant value for all miRNAs in
the system (The same applies to mRNA.). On the other hand other
regulation parameters such as gene-gene interactions and transcription
factors are ignored but the network structure can promote to integration
of other regulation elements. In the future, with development in
experimental techniques about features of miRNAs and their targets, more
consistent and useful results can be obtained from our approach.
Additionally, this can provide to contribution to prediction of abnormal
regulations and pathways with the studies that will be developed.

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-ala_integrated_2013}{}%
Ala, U., F. A. Karreth, C. Bosia, A. Pagnani, R. Taulli, V. Leopold, Y.
Tay, P. Provero, R. Zecchina, and P. P. Pandolfi. 2013. ``Integrated
Transcriptional and Competitive Endogenous RNA Networks Are
Cross-Regulated in Permissive Molecular Environments.''
\emph{Proceedings of the National Academy of Sciences} 110 (18):
7154--9. \url{https://doi.org/10.1073/pnas.1222509110}.

\leavevmode\hypertarget{ref-arvey_target_2010}{}%
Arvey, Aaron, Erik Larsson, Chris Sander, Christina S Leslie, and Debora
S Marks. 2010. ``Target mRNA Abundance Dilutes microRNA and siRNA
Activity.'' \emph{Molecular Systems Biology} 6 (April).
\url{https://doi.org/10.1038/msb.2010.24}.

\leavevmode\hypertarget{ref-atalay2006multidrug}{}%
Atalay, Can, Ismet Deliloglu Gurhan, Cigdem Irkkan, and Ufuk Gunduz.
2006. ``Multidrug Resistance in Locally Advanced Breast Cancer.''
\emph{Tumor Biology} 27 (6): 309--18.

\leavevmode\hypertarget{ref-atalay2008role}{}%
Atalay, C, A Demirkazik, and U Gunduz. 2008. ``Role of Abcb1 and Abcc1
Gene Induction on Survival in Locally Advanced Breast Cancer.''
\emph{Journal of Chemotherapy} 20 (6): 734--39.

\leavevmode\hypertarget{ref-bartel_micrornas_2004}{}%
Bartel, David P. 2004. ``MicroRNAs.'' \emph{Cell} 116 (2): 281--97.
\url{https://doi.org/10.1016/S0092-8674(04)00045-5}.

\leavevmode\hypertarget{ref-bartel_micrornas:_2009}{}%
Bartel, David P. 2009. ``MicroRNAs: Target Recognition and Regulatory
Functions.'' \emph{Cell} 136 (2): 215--33.
\url{https://doi.org/10.1016/j.cell.2009.01.002}.

\leavevmode\hypertarget{ref-bosson_endogenous_2014}{}%
Bosson, Andrew D., Jesse R. Zamudio, and Phillip A. Sharp. 2014.
``Endogenous miRNA and Target Concentrations Determine Susceptibility to
Potential ceRNA Competition.'' \emph{Molecular Cell} 56 (3): 347--59.
\url{https://doi.org/10.1016/j.molcel.2014.09.018}.

\leavevmode\hypertarget{ref-breda_quantifying_2015}{}%
Breda, Jeremie, Andrzej J. Rzepiela, Rafal Gumienny, Erik van Nimwegen,
and Mihaela Zavolan. 2015. ``Quantifying the Strength of miRNA--Target
Interactions.'' \emph{Methods} 85 (September): 90--99.
\url{https://doi.org/10.1016/j.ymeth.2015.04.012}.

\leavevmode\hypertarget{ref-cao_predicting_2012}{}%
Cao, Song, and Shi-Jie Chen. 2012. ``Predicting Kissing Interactions in
microRNA--Target Complex and Assessment of microRNA Activity.''
\emph{Nucleic Acids Research} 40 (10): 4681--90.
\url{https://doi.org/10.1093/nar/gks052}.

\leavevmode\hypertarget{ref-cesana_deciphering_2013}{}%
Cesana, Marcella, and George Q. Daley. 2013. ``Deciphering the Rules of
ceRNA Networks.'' \emph{Proceedings of the National Academy of Sciences
of the United States of America} 110 (18): 7112--3.
\url{https://doi.org/10.1073/pnas.1305322110}.

\leavevmode\hypertarget{ref-denzler_assessing_2014}{}%
Denzler, Rémy, Vikram Agarwal, Joanna Stefano, David P. Bartel, and
Markus Stoffel. 2014. ``Assessing the ceRNA Hypothesis with Quantitative
Measurements of miRNA and Target Abundance.'' \emph{Molecular Cell} 54
(5): 766--76. \url{https://doi.org/10.1016/j.molcel.2014.03.045}.

\leavevmode\hypertarget{ref-denzler_impact_2016}{}%
Denzler, Rémy, Sean E. McGeary, Alexandra C. Title, Vikram Agarwal,
David P. Bartel, and Markus Stoffel. 2016. ``Impact of MicroRNA Levels,
Target-Site Complementarity, and Cooperativity on Competing Endogenous
RNA-Regulated Gene Expression.'' \emph{Molecular Cell} 64 (3): 565--79.
\url{https://doi.org/10.1016/j.molcel.2016.09.027}.

\leavevmode\hypertarget{ref-figliuzzi_micrornas_2013}{}%
Figliuzzi, Matteo, Enzo Marinari, and De MartinoAndrea. 2013.
``MicroRNAs as a Selective Channel of Communication Between Competing
RNAs: A Steady-State Theory.'' \emph{Biophysical Journal} 104 (5):
1203--13. \url{https://doi.org/10.1016/j.bpj.2013.01.012}.

\leavevmode\hypertarget{ref-grimson_microrna_2007}{}%
Grimson, Andrew, Kyle Kai-How Farh, Wendy K. Johnston, Philip
Garrett-Engele, Lee P. Lim, and David P. Bartel. 2007. ``MicroRNA
Targeting Specificity in Mammals: Determinants Beyond Seed Pairing.''
\emph{Molecular Cell} 27 (1): 91--105.
\url{https://doi.org/10.1016/j.molcel.2007.06.017}.

\leavevmode\hypertarget{ref-hafner_transcriptome-wide_2010}{}%
Hafner, Markus, Markus Landthaler, Lukas Burger, Mohsen Khorshid, Jean
Hausser, Philipp Berninger, Andrea Rothballer, et al. 2010.
``Transcriptome-Wide Identification of RNA-Binding Protein and MicroRNA
Target Sites by PAR-CLIP.'' \emph{Cell} 141 (1): 129--41.
\url{https://doi.org/10.1016/j.cell.2010.03.009}.

\leavevmode\hypertarget{ref-hausser_analysis_2013}{}%
Hausser, J., A. P. Syed, B. Bilen, and M. Zavolan. 2013. ``Analysis of
CDS-Located miRNA Target Sites Suggests That They Can Effectively
Inhibit Translation.'' \emph{Genome Research} 23 (4): 604--15.
\url{https://doi.org/10.1101/gr.139758.112}.

\leavevmode\hypertarget{ref-helwak_mapping_2013}{}%
Helwak, Aleksandra, Grzegorz Kudla, Tatiana Dudnakova, and David
Tollervey. 2013. ``Mapping the Human miRNA Interactome by CLASH Reveals
Frequent Noncanonical Binding.'' \emph{Cell} 153 (3): 654--65.
\url{https://doi.org/10.1016/j.cell.2013.03.043}.

\leavevmode\hypertarget{ref-lai_understanding_2016}{}%
Lai, Xin, Olaf Wolkenhauer, and Julio Vera. 2016. ``Understanding
microRNA-Mediated Gene Regulatory Networks Through Mathematical
Modelling.'' \emph{Nucleic Acids Research} 44 (13): 6019--35.
\url{https://doi.org/10.1093/nar/gkw550}.

\leavevmode\hypertarget{ref-lewis_conserved_2005}{}%
Lewis, Benjamin P., Christopher B. Burge, and David P. Bartel. 2005.
``Conserved Seed Pairing, Often Flanked by Adenosines, Indicates That
Thousands of Human Genes Are MicroRNA Targets.'' \emph{Cell} 120 (1):
15--20. \url{https://doi.org/10.1016/j.cell.2004.12.035}.

\leavevmode\hypertarget{ref-lu2015microrna}{}%
Lu, Lin, Fang Ju, Hui Zhao, and Xuezhen Ma. 2015. ``MicroRNA-134
Modulates Resistance to Doxorubicin in Human Breast Cancer Cells by
Downregulating Abcc1.'' \emph{Biotechnology Letters} 37 (12): 2387--94.

\leavevmode\hypertarget{ref-markus_list_sponge_2017}{}%
Markus List, Marcel Schulz. 2017. ``SPONGE.'' Bioconductor.
\url{https://doi.org/10.18129/B9.bioc.SPONGE}.

\leavevmode\hypertarget{ref-moore_mirnatarget_2015}{}%
Moore, Michael J., Troels K. H. Scheel, Joseph M. Luna, Christopher Y.
Park, John J. Fak, Eiko Nishiuchi, Charles M. Rice, and Robert B.
Darnell. 2015. ``miRNA-Target Chimeras Reveal miRNA 3'-End Pairing as a
Major Determinant of Argonaute Target Specificity.'' \emph{Nature
Communications} 6 (November): 8864.
\url{https://doi.org/10.1038/ncomms9864}.

\leavevmode\hypertarget{ref-nitzan_interactions_2014}{}%
Nitzan, Mor, Avital Steiman-Shimony, Yael Altuvia, Ofer Biham, and Hanah
Margalit. 2014. ``Interactions Between Distant ceRNAs in Regulatory
Networks.'' \emph{Biophysical Journal} 106 (10): 2254--66.
\url{https://doi.org/10.1016/j.bpj.2014.03.040}.

\leavevmode\hypertarget{ref-robinson_modelling_2018}{}%
Robinson, J. M., and W. A. Henderson. 2018. ``Modelling the Structure of
a ceRNA-Theoretical, Bipartite microRNA-mRNA Interaction Network
Regulating Intestinal Epithelial Cellular Pathways Using R
Programming.'' \emph{BMC Research Notes} 11 (1): 19.
\url{https://doi.org/10.1186/s13104-018-3126-y}.

\leavevmode\hypertarget{ref-ruegger_microrna_2012}{}%
Rüegger, Stefan, and Helge Großhans. 2012. ``MicroRNA Turnover: When,
How, and Why.'' \emph{Trends in Biochemical Sciences} 37 (10): 436--46.
\url{https://doi.org/10.1016/j.tibs.2012.07.002}.

\leavevmode\hypertarget{ref-salmena_cerna_2011}{}%
Salmena, Leonardo, Laura Poliseno, Yvonne Tay, Lev Kats, and Pier Paolo
Pandolfi. 2011. ``A ceRNA Hypothesis: The Rosetta Stone of a Hidden RNA
Language?'' \emph{Cell} 146 (3): 353--58.
\url{https://doi.org/10.1016/j.cell.2011.07.014}.

\leavevmode\hypertarget{ref-tay_multilayered_2014}{}%
Tay, Yvonne, John Rinn, and Pier Paolo Pandolfi. 2014. ``The
Multilayered Complexity of ceRNA Crosstalk and Competition.''
\emph{Nature} 505 (7483): 344--52.
\url{https://doi.org/10.1038/nature12986}.

\leavevmode\hypertarget{ref-werfel_preferential_2017}{}%
Werfel, Stanislas, Simon Leierseder, Benjamin Ruprecht, Bernhard Kuster,
and Stefan Engelhardt. 2017. ``Preferential microRNA Targeting Revealed
by in Vivo Competitive Binding and Differential Argonaute
Immunoprecipitation.'' \emph{Nucleic Acids Research} 45 (17): 10218--28.
\url{https://doi.org/10.1093/nar/gkx640}.

\leavevmode\hypertarget{ref-xu_characterization_2014}{}%
Xu, Wenlong, Zixing Wang, and Yin Liu. 2014. ``The Characterization of
microRNA-Mediated Gene Regulation as Impacted by Both Target Site
Location and Seed Match Type.'' Edited by Thomas Preiss. \emph{PLoS ONE}
9 (9): e108260. \url{https://doi.org/10.1371/journal.pone.0108260}.

\end{document}

### 2019-05-27 article.Rmd
abstract edited, into first paragraph edited

@@
 
-Competing endogeneous RNA (ceRNA) regulations and crosstalk in human are remarkable issue in mean of regulations via miRNAs.
-Many studies have pointed out that an alteration in miRNA:target interaction can result in unexpected regulations because of indirect and complex interaction.
-In this paper, we defined a new network-based model that handles miRNA:ceRNA interactions with expression values and interaction factors optionally.
-We applied approach in sample dataset and defined dynamics of it.
-We have observed that a perturbation in an element of network affects whole competing elements differently and cooperative efficiencies of miRNAs on common targets could be shown.
-Importance of miRNA:target ratios was also observed consistently with previous reports.
-We have obtained that the competing elements which have the same or close expression values may not affected equal from the perturbation because of repression functionality of interactions.
-To apply model to the real sample that obtained combining  breeast cancer gene and miRNA expression dataset via high throughput experimental miRNA:target dataset, we selected a gene, *ABCC1*, which over expressed in tumor tissue and utilized it as perturbating element.
-We have observed that one change can be sufficient to perturb regulations, and may not predicted all of regulations correctly, with this change.
-Therefore, this model which helps to understand crosstalking between elements could be tested and improved via results of well-constructed experimental dataset and takes to hope for future research.
+Competing endogenous RNA (ceRNA) regulations and crosstalk between various types of non-coding RNA in human is remarkable in means of miRNA regulation.
+Many studies have pointed out that an alteration in miRNA:target interaction can result in unexpected changes due to indirect and complex interactions.
+In this paper, we defined a new network-based model that handles miRNA:ceRNA interactions with expression values.
+Our model is able to handle miRNA interaction factors such as seed type, binding energy, if provided.
+Our approach is able to reveal that a perturbation in an element of network affects whole competing elements differently and cooperative efficiencies of miRNAs on common targets could be calculated.
+Our findings emphasized importance of miRNA:target ratios being crucial, as reported by previous studies.
+We have showed that the competing elements which have the same or close expression values may not be affected equally from the perturbation because of repression functionality of interactions(TOFIX son kısmı anlayamadım).
+We applied the model to real sample consisting of breast cancer gene and miRNA expression dataset and experimental miRNA:target interaction dataset all generated via high throughput sequencing methods.
+A gene over-expressed in tumor tissue, namely *ABCC1*, is used as perturbing element.
+We have observed that change in expression level of single gene in miRNA:target network is sufficient to perturb regulations in whole network, due to unforeseen and unpredicted regulation which are only visible when considered in network context.
+Therefore, this model helps unveiling the crosstalk between elements in miRNA:target network where abundance of target and sponge effect are taken into account.
+The model is scalable and can be plugged in with emerging miRNA effectors such as circRNAs.
+The model is available as R package at https://github.com/selcenari/ceRNAnetsim.
+(TODO the link will change)
 
@@
 
-MicroRNAs (miRNAs) that are key regulator of gene expression through various post-transcriptional mechanisms are a family of short non-coding RNAs.
-Although repression mechanisms of miRNA are not fully understood, exact effect of miRNAs is regulation of direct target/targets negatively.
+MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key regulator of gene expression through various post-transcriptional mechanisms.
+Although the mechanisms by which miRNA represses are not fully understood,  miRNAs predominantly repress their targets.
 Repressive activities of miRNAs vary depending on many factors that are significant to microRNA:target interactions.
@@
 It has been demonstrated that the binding energy between miRNA and target indicates stability or affinity of complex [@helwak_mapping_2013] and does not directly determine repressive activity of miRNA [@cao_predicting_2012].
-Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5’UTR bind to specific sequence located in 3’UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005].
-In recent studies, it has been shown miRNAs can interact with targets via sequence that located in different region such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013; @moore_mirnatarget_2015].
-These studies also shown that binding location could indicate functionality of miRNA:target interaction or affect profile of target abundance.
-It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is prefectly complementary with targets, seed region at the 5’UTR of miRNAs [@bartel_micrornas:_2009; @grimson_microrna_2007].
-On the other hand, some researchers have reported that seed sequence of miRNA can have mismatches or bulged/wobble nucleotides and locates in different region of miRNAs [@hafner_transcriptome-wide_2010; @helwak_mapping_2013].
-In addition to all of these, abundance of miRNAs and targets and miRNA:target ratio in cells are determinant for efficiency of miRNA:target interaction [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
+Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5’UTR (TOFIX mirna için UTR ifadesi geçerli mi, sadec 5 ucu dense yeterli mi?) bind to specific sequence located in 3’UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005].
+In recent studies, it has been shown that miRNAs can interact with targets via sequences located in regions such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013; @moore_mirnatarget_2015].
+These studies also showed that binding location could indicate functionality of miRNA:target interaction or affect profile of target abundance(TOFIX abundance etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı?).
+It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is perfectly complementary with targets, seed region at the 5’UTR of miRNAs(TOFIX mirna için UTR geçerli mi) [@bartel_micrornas:_2009; @grimson_microrna_2007].
+On the other hand, some researchers have reported that seed sequence of miRNA can have mismatches or bulged/wobble nucleotides and may locate in region other than 5'end of miRNAs [@hafner_transcriptome-wide_2010; @helwak_mapping_2013].
+On top of all these factors, abundance of miRNAs and targets and miRNA:target ratio in cells predominantly affect efficiency of miRNA:target interaction [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
 
-As it is possible for miRNAs to suppress multiple targets, an individual RNA molecule can also be targeted by various miRNAs.
-In that case, the targeted RNAs exhibit competitor behavior, that is hypothesized as competing endogenous RNAs (ceRNAs) [@ala_integrated_2013; @cesana_deciphering_2013], against their miRNAs.
+As it is possible for miRNAs to suppress multiple targets, an individual mRNA molecule can also be targeted by multiple miRNAs.
+In that case, the targeted mRNAs exhibit competitor behavior, that is hypothesized as competing endogenous RNAs (ceRNAs) [@ala_integrated_2013; @cesana_deciphering_2013], against their miRNAs.
 Briefly, Ala et al.
-have explained the ceRNA hypothesis as disturbtion of the other target when a steady-state system that included one miRNA and two target, was perturbed with expression changes at one of the targets [@ala_integrated_2013].
+have explained the ceRNA hypothesis as distribution of the other target when a steady-state system that included one miRNA and two target, was perturbed with expression changes at one of the targets [@ala_integrated_2013].
 When interaction between miRNAs and their target were dealed in the cell, explanation and prediction of results of perturbation is difficult due to varieties of interactions.
@@
 Cooperative efficiency of miRNAs as well as competitor behaviors of targets also were studied and it has been demonstrated crucial for regulations at expression of targets [@denzler_impact_2016].
-MiRNA:target interactions have modelled as stoichiometric and catalytic mechanisms and Figliuzzi et al.
+MiRNA:target interactions have modeled as stoichiometric and catalytic mechanisms and Figliuzzi et al.
 have recommended to handling models in network structures [@figliuzzi_micrornas_2013].

### 2019-05-27 article.Rmd
editing second half of introduction

@@
 Briefly, Ala et al.
-have explained the ceRNA hypothesis as distribution of the other target when a steady-state system that included one miRNA and two target, was perturbed with expression changes at one of the targets [@ala_integrated_2013].
-When interaction between miRNAs and their target were dealed in the cell, explanation and prediction of results of perturbation is difficult due to varieties of interactions.
-Various computational and experimental studies have applied to be understood for ceRNA:miRNA interactions.
+have explained the ceRNA hypothesis as distribution (TOFIX disturbance mı kastedildi acaba?) of the second target when a steady-state system that included one miRNA and two target when perturbed with expression changes at one of the targets [@ala_integrated_2013].
+Regarding interaction between miRNAs and their target in a cell, explaining and predicting results of an individual perturbation is difficult due to complexity of interactions.
+Various computational and experimental studies have tackled the problem of unraveling ceRNA:miRNA interactions.
 It has been observed that when abundance of one of the targets of miR-122 was increased, the other target expression also slightly increased as a result of decreasing repressive activity of miR-122 on them [@denzler_assessing_2014].
 Bosson et al.
-have developed a mathematical model for changes on total target pool concentration after grouping to affinity of targets.
-They have demonstrated that miRNA activity correlated with affinity between miRNA and target [@bosson_endogenous_2014].
-Cooperative efficiency of miRNAs as well as competitor behaviors of targets also were studied and it has been demonstrated crucial for regulations at expression of targets [@denzler_impact_2016].
-MiRNA:target interactions have modeled as stoichiometric and catalytic mechanisms and Figliuzzi et al.
-have recommended to handling models in network structures [@figliuzzi_micrornas_2013].
+have developed a mathematical model for changes on total target pool concentration after grouping targets according to affinity (TOFIX orijinal cümle ile aynı manaya geldi mi emin olamadım) and demonstrated that miRNA activity correlated with affinity between miRNA and target [@bosson_endogenous_2014].
+Cooperative efficiency of miRNAs as well as competitor behaviors of targets were also studied and it has been demonstrated to be crucial for regulating available mRNA levels (TOFIX expression deyince sanki transcriptional bir etki var gibi anlaşılacak diye rna level deme ihtiyacı hissettim) of targets [@denzler_impact_2016].
+MiRNA:target interactions have been modeled as stoichiometric and catalytic mechanisms and Figliuzzi et al.
+have recommended handling models in network context [@figliuzzi_micrornas_2013].
 The model that can explain miRNA target interaction through topological features has been applied at bipartite network by Nitzan et al.
@@
 Robinson and Henderson applied the model that handles miRNA:target direct and indirect interactions via common miRNA of gene and target of miRNA, at bipartite network.
-It has been demonstrated that whole miRNAs and targets in network can interacted through common miRNAs and genes [@robinson_modelling_2018].
-Associated genes that are targets of the same miRNAs have been found with help of correlation of gene expression changes in recent study [@markus_list_sponge_2017].
+It has been demonstrated that all miRNAs and targets in network can interact each other, bia not necessarily direct interactions, through common miRNAs and genes [@robinson_modelling_2018].
+Associated genes that are targets of the same miRNAs have been found with help of correlation of gene expression changes in recent study(TOFIX bu R paketine atıf veriyor, makale atıfı varsa daha iyi olurdu, makalesi yoksa, recent study yerine, recent algoritm deriz) [@markus_list_sponge_2017].
 List et al.
-have specified that their approach can be useful for ceRNA’s studies and published these approach as a package that can use with R programming.
+have specified that their approach can be useful for ceRNA studies and published their approach as an R package.
 

### 2019-05-27 ceRNAnetsim/R/calc_perturbation.R
remove package

@@ -1,39 +1,0 @@
-#' @title calc_perturbation
-#'
-#' @description  calculates mean of expression changes of all  nodes and finds the perturbed node count for a given node.
-#'
-#' @return gives a vector about the perturbation efficiency and number of perturbed nodes.
-#'
-#' @details calc_perturbation calculates mean expression changes of elements after the change in the network in terms of percentage. It also calculates the number of nodes that have expression changes after the change occur in the network.
-#'  The function determines the perturbation efficiency and number of perturbed nodes after given change with how, cycle and limit parameter.
-#'
-#' @param input_graph the graph object that was processed with priming graph in previous step.
-#' @param node_name The node that is trigger for simulation.
-#' @param how The change of count of the given node in terms of fold change.
-#' @param cycle The iteration of simulation.
-#' @param limit The minimum fold change which can be taken into account for perturbation calculation on all nodes in terms of percentage.
-#'
-#' @examples
-#'
-#' minsamp%>%
-#'    priming_graph(competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy,seed_type), deg_factor = region)%>%
-#'    calc_perturbation("Gene6",3, cycle = 4)
-#'
-#'
-#' @export
-
-
-calc_perturbation <- function(input_graph, node_name, how=1 , cycle=1, limit=0){
-
-  input_graph%>%
-    update_nodes(once=TRUE)%>%
-    update_how(node_name, how)%>%
-    simulate(cycle)%>%
-    as_tibble()-> res
-
-  as.double((res%>%summarise(mean(abs(count_current-initial_count)*100/initial_count)))[[1]])-> perturbation_eff
-
-  as.double((res%>%filter((abs(count_current-initial_count)*100/initial_count) > limit)%>%count())[[1]])->perturbed_count
-
-  return(c(perturbation_efficiency= perturbation_eff, perturbed_count = perturbed_count))
-}

### 2019-05-27 ceRNAnetsim/R/find_iteration.R

@@ -1,156 +1,0 @@
-#' @title find_iteration
-#'
-#' @description find_iteration searches the iteration that provides maximum affected node number. The user defines a symbolic iteration with .iter. The function calculates the number of affected nodes for each iteration and than selects the iteration that has maximum affected nodes' number.
-#'
-#' @return It gives an iteration number to use in simulate() function.
-#'
-#' @param df A data frame that includes the miRNA and competing targets.
-#' @param competing_count The counts (expression) of competing elements of the dataset.
-#' @param miRNA_count The counts (expression) of repressive element (miRNA) of the dataset.
-#' @param aff_factor The parameter/s of binding between miRNA and targets.
-#' @param deg_factor The parameter/s for degradation of bound miRNA:target complex.
-#' @param node_name The name f node that is used as trigger.
-#' @param how The change of count (expression) of the given node in terms of fold change.
-#' @param .iter The maximum iteration prescribed by the user.
-#' @param limit The minimum amount of change of any node.
-#'
-#' @details
-#'
-#' @examples
-#'
-#' find_iteration(minsamp, Competing_expression, miRNA_expression, aff_factor = energy, deg_factor = region, node_name = "Gene2", how= 3)
-#'
-#' @export
-
-
-find_iteration <- function(df, competing_count, miRNA_count, aff_factor=dummy, deg_factor=dummy, node_name, how, .iter = 100, limit= 0.1){
-
-  competing_exp <- enquo(competing_count)
-  mirna_exp <- enquo(miRNA_count)
-  affinity <- enquos(aff_factor)
-  degradation <- enquos(deg_factor)
-
-  iteration <- data.frame(iter = seq(1,.iter, 1), effect= rep(0))
-
-  df <- df%>%
-    dplyr::mutate(competing = df[,1], miRNA= df[,2], Competing_name = df[,1], miRNA_name= df[,2], dummy=1)%>%
-    dplyr::select(competing, miRNA, Competing_name, miRNA_name, !!competing_exp, !!mirna_exp, !!!affinity, !!!degradation, dummy)
-
-
-  df%>%
-    dplyr::group_by(miRNA)%>%
[…]

### 2019-05-27 ceRNAnetsim/R/find_node_perturbation.R

@@ -1,38 +1,0 @@
-#' @title find_node_perturbation
-#'
-#' @description  calculates the mean of expression changes of all  nodes and finds the perturbed node count for all nodes in system.
-#'
-#' @return It gives a tibble form dataset that includes node names, perturbation efficiency and perturbed count of nodes.
-#'
-#' @details find_node_perturbation calculates mean expression changes of elements after the change in the network in terms of percentage. It also calculates the number of nodes that have expression changes after the change occur in the network.
-#'   The outputs of the function are the perturbation efficiency and perturbed count of nodes for each nodes.
-#'
-#' @param input_graph The graph object that was processed with priming_graph function.
-#' @param how The change of count (expression) of the given node in terms of fold change.
-#' @param cycle The iteration of simulation.
-#' @param limit The minimum fold change which can be taken into account for perturbation calculation on all nodes in terms of percentage.
-#'
-#' @examples
-#'
-#' minsamp%>%
-#'    priming_graph(competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy,seed_type), deg_factor = region)%>%
-#'    find_node_perturbation(how = 3, cycle = 4)
-#'
-#' @export
-
-find_node_perturbation <- function(input_graph, how = 2, cycle=1, limit= 0){
-
-  input_graph%N>%
-    mutate(eff_count =future_map(V(input_graph)$name, ~calc_perturbation(input_graph, .x, how, cycle, limit)))%>%
-    as_tibble()-> result
-
-  result%>%
-    unnest(eff_count)%>%
-    group_by(name) %>%
-    mutate(col=seq_along(name)) %>%
-    spread(key=col, value=eff_count)%>%
-    ungroup()%>%
-    dplyr::rename(perturbation_efficiency= "1", perturbed_count = "2") -> result
-
-  return(result)
-}

### 2019-05-27 ceRNAnetsim/R/priming_graph.R

@@ -1,79 +1,0 @@
-#' @title normalize
-#'
-#' @description normalizes the values according to maximum values inside a group. The helper function of priming_graph.
-#'
-#' @param x The variable name that is normalized.
-#'
-#' @return NULL
-#'
-
-
-normalize <- function(x){abs(x)/max(abs(x))}
-
-#' @title priming_graph
-#'
-#' @description The function converts the given dataframe using first variable as competing and the second as miRNA. If user defines interaction factors as affinity or degradation, the factors are taken into account.
-#'
-#' @return the graph object.
-#'
-#' @param df A data frame that includes the miRNA and competing targets.
-#' @param competing_count The counts (or expression) of competing elements of the dataset.
-#' @param miRNA_count The counts (or expression) of repressive element (miRNA) of the dataset.
-#' @param aff_factor The parameter/s of binding between miRNA and targets.
-#' @param deg_factor The parameter/s for degradation of bound miRNA:target complex.
-#'
-#' @details priming_graph provides grouping of competing targets and evaluation of targets within the groups taking into account miRNA:target, target:total target, interaction and degradation parameters. The target groups are determined according to miRNAs. If the factors that are important in target interactions are specified as arguments, the factors also are evaluated separately within each group.
-#' priming_graph also calculates the miRNA efficiency in steady-state conditions. It is assumed that quantity of competing targets and miRNAs are shown in the steady-state system after the miRNAs exhibit repressive efficiency.
-#' Note that the data must not include missing values such as NA or "-".
-#'
-#'
-#' @examples
-#'
-#' priming_graph(minsamp, Competing_expression, miRNA_expression, aff_factor = c(seed_type,energy), deg_factor = region)
-#'
-#' @export
-
-priming_graph <- function(df, competing_count, miRNA_count, aff_factor=dummy, deg_factor=dummy){
-
-
-  if(sum(is.na.data.frame(df)) !=0){
-
-    stop("Missing or NA value in dataframe")
-  }
-
-  else{
-
-  competing_exp <- enquo(competing_count)
-  mirna_exp <- enquo(miRNA_count)
[…]

### 2019-05-27 ceRNAnetsim/R/simulate.R

@@ -1,37 +1,0 @@
-#' @title simulate
-#'
-#' @description  simulate function uses the change in expression value/s as triggering.
-#'
-#' @return The graph.
-#'
-#' @details The steady-state conditions of the system are disturbed after the change in the graph (with update_how or update_variables). In this case, the system tend to be steady state again. The arrangement of competetive profiles of the targets continue until all nodes are updated and steady-state nearly.
-#'
-#' @param input_graph The graph object that processed in previous steps.
-#' @param cycle Optimal iteration number for gaining steady-state.
-#'
-#' @examples
-#'
-#' priming_graph(minsamp, Competing_expression, miRNA_expression, aff_factor = c(seed_type,energy), deg_factor = region)%>%
-#'    update_nodes(once = TRUE)%>%
-#'    update_variables(new_counts)%>%
-#'    simulate(cycle = 3)
-#'
-#' @export
-
-simulate <- function(input_graph, cycle=1){
-
-  for(i in seq_along(1:cycle)){
-
-    input_graph%E>%
-      group_by(to)%>%
-      mutate(mirna_count_per_comp = mirna_count_current*comp_count_current*afff_factor/sum(comp_count_current*afff_factor), mirna_count_per_comp = ifelse(is.na(mirna_count_per_comp), 0, mirna_count_per_comp))%>%
-      ungroup()%>%
-      mutate(effect_pre = effect_current, effect_current = mirna_count_per_comp*degg_factor)%>%
-      group_by(from)%>%
-      mutate(comp_count_pre = ifelse(comp_count_current< 0, 1, comp_count_current), comp_count_current = comp_count_pre - (sum(effect_current)-sum(effect_pre)), comp_count_current = ifelse(comp_count_current< 0, 1, comp_count_current))%>%
-      ungroup()%>%
-      mutate(comp_count_list = pmap(list(comp_count_list, comp_count_current), c), effect_list = pmap(list(effect_list, effect_current), c), mirna_count_list = pmap(list(mirna_count_list, mirna_count_current), c))%N>%
-      update_nodes()-> input_graph
-  }
-  input_graph
-}

### 2019-05-27 ceRNAnetsim/R/simulate_vis.R

@@ -1,56 +1,0 @@
-#' @title simulate_vis
-#'
-#' @description  simulate_vis provides visualisation of the graph in addition to simulate function.
-#'
-#' @details simulate_vis gives the last graph object and each iterations' image.
-#'
-#' @return It gives a graph and the images of states in each iteration until the end of the simulation.
-#'
-#' @param input_graph The graph object that processed in previous steps.
-#' @param cycle Optimal iteration number for gaining steady-state.
-#' @param Competing_color The color of competing elements on the graph with "green" default.
-#' @param mirna_color The color of miRNAs on the graph with "orange" default.
-#' @param Upregulation The color of Upregulated elements on the graph with "red" default.
-#' @param Downregulation The color of Downregulated elements on the graph with "blue" default.
-#' @param title Title of the given graph.
-#' @param layout The layout that will be used for visualisation of the graph.
-#'
-#' @examples
-#'
-#' # When does the system gain steady-state conditions again?
-#'
-#' ## new_counts, the dataset that includes the current counts of nodes.
-#'
-#' priming_graph(minsamp, Competing_expression, miRNA_expression, aff_factor = c(seed_type,energy), deg_factor = c(region))%>%
-#'    update_nodes(once = TRUE)%>%
-#'    update_variables(new_counts)%>%
-#'    update_nodes()%>%
-#'    simulate_vis(cycle = 12)
-#'
-#' @export
-
-
-simulate_vis <- function(input_graph, cycle=1, Competing_color = "green", mirna_color = "orange", Upregulation = "red", Downregulation = "blue", title = "GRAPH", layout= "kk"){
-
-  for(i in seq_along(1:cycle)){
-
-    input_graph%E>%
-      group_by(to)%>%
-      mutate(mirna_count_per_comp = mirna_count_current*comp_count_current*afff_factor/sum(comp_count_current*afff_factor), mirna_count_per_comp = ifelse(is.na(mirna_count_per_comp), 0, mirna_count_per_comp))%>%
-      ungroup()%>%
-      mutate(effect_pre = effect_current, effect_current = mirna_count_per_comp*degg_factor)%>%
-      group_by(from)%>%
[…]

### 2019-05-27 ceRNAnetsim/R/update_how.R

@@ -1,44 +1,0 @@
-#' @title update_how
-#'
-#' @description  update_how converts the count value of the given node.
-#'
-#' @return the graph object.
-#'
-#' @details update_how function calculates the current value of given mirna or gene node on the graph object. User must specify current value as fold change.
-#'
-#' @param input_graph The graph object that processed in previous step/s.
-#' @param node_name The name of the node whose count is to be changed.
-#' @param how The change in terms of fold change.
-#'
-#' @examples
-#'
-#' priming_graph(minsamp, Competing_expression, miRNA_expression, aff_factor = c(seed_type,energy), deg_factor = region)%>%
-#'    update_nodes(once = TRUE)%>%
-#'    update_how("Gene1", 3)
-#'
-#' @export
-
-update_how <- function (input_graph, node_name, how){
-
-
-  if(node_name %in% E(input_graph)$Competing_name){
-
-    input_graph <- input_graph%>%
-      activate(edges)%>%
-      mutate(comp_count_current = ifelse(node_name == E(input_graph)$Competing_name, comp_count_current*how, comp_count_current))
-
-  } else if(node_name %in% E(input_graph)$miRNA_name){
-
-   input_graph <- input_graph%>%
-     activate(edges)%>%
-      mutate(mirna_count_current = ifelse(node_name == E(input_graph)$miRNA_name, mirna_count_current*how, mirna_count_current))
-
-  } else{
-
-    stop("Given node name was not found! Please check it.")
-
-  }
-
-  return(input_graph)
-}
-

### 2019-05-27 ceRNAnetsim/R/update_nodes.R

@@ -1,89 +1,0 @@
-#' @title prepare_rhs_once
-#'
-#' @description  Carries the variables from edge to node.
-#'
-#' @details The function is a helper function for processing of graph object with update_nodes function.
-#'
-#' @param input_graph Processed graph object in previous step.
-#'
-
-
-prepare_rhs_once <- function(input_graph){
-
-  bind_rows((as_tibble(input_graph%>%
-                         activate(edges)%>%
-                         dplyr::select(comp_count_current, comp_count_pre))%>%
-               mutate(initial_count = comp_count_pre)%>%
-               dplyr::select(id = from, initial_count, count_pre = comp_count_pre,  count_current = comp_count_current)%>%
-               dplyr::distinct()),
-            (as_tibble(input_graph%>%
-                         activate(edges)%>%
-                         dplyr::select(mirna_count_current, mirna_count_pre))%>%
-               mutate(initial_count = mirna_count_pre)%>%
-               dplyr::select(id = to, initial_count, count_pre = mirna_count_pre,  count_current = mirna_count_current)%>%
-               dplyr::distinct()))
-
-
-}
-
-#' @title prepare_rhs
-#'
-#' @description  Carries the variables from edge to node.
-#'
-#' @details The function is a helper function for processing of graph object with update_nodes function.
-#'
-#' @param input_graph Processed graph object in previous step.
-#'
-
-
-prepare_rhs <- function(input_graph){
-
-  bind_rows((as_tibble(input_graph%>%
-                         activate(edges))%>%
-               dplyr::select(id = from, count_current = comp_count_current)%>%
-               dplyr::distinct()),
-            (as_tibble(input_graph%>%
-                         activate(edges))%>%
-               dplyr::select(id= to,  count_current = mirna_count_current)%>%
-               dplyr::distinct()))
-}
-
-#' @title update_nodes
-#'
-#' @description  The function carries variables from edge to node.
-#'
-#' @return the graph object.
-#'
-#' @param input_graph Processed graph object in previous step.
-#' @param once The argument is about when the carrying process runs.
-#'
-#' @details If the carrying process performs after priming_graph function, the argument must be TRUE.
-#'     The function helps to visualisation of processed graph object, especially that includes too many nodes.This step makes it easily to follow the processes.
-#'
[…]

### 2019-05-27 ceRNAnetsim/R/update_variables.R

@@ -1,41 +1,0 @@
-#' @title update_variables
-#'
-#' @description  The function replace new values with previous values of competing or miRNA counts.
-#'
-#' @return the graph object.
-#'
-#' @details update_variables function provides updating edge variables to current values. If the microRNA or competing expression (or both) change (decreasing or increasing), this function switches the values that are found in a new dataset provided by user. But the current value dataset must be equal with initial dataset in terms of node name.
-#'
-#' @param input_graph The processed graph object.
-#' @param current_counts The additional df that provided by user.
-#'
-#' @examples
-#'
-#' minsamp%>%
-#' priming_graph(Competing_expression, miRNA_expression, aff_factor = c(seed_type,energy), deg_factor = region)%>%
-#'    update_nodes(once = TRUE)%>%
-#'    update_variables(new_counts)  #the dataset that includes the current counts of nodes.
-#'
-#' @export
-
-update_variables <- function (input_graph, current_counts){
-
-  if(isTRUE(all.equal(current_counts$Competing, E(input_graph)$Competing_name))){
-
-    E(input_graph)$comp_count_current[match(current_counts$Competing, E(input_graph)$Competing_name)] <- current_counts$Competing_count
-  }
-
-  if (isTRUE(all.equal(current_counts$miRNA, E(input_graph)$miRNA_name))) {
-
-    E(input_graph)$mirna_count_current[match(current_counts$miRNA, E(input_graph)$miRNA_name)] <- current_counts$miRNA_count
-  }
-
-  else{
-
-    stop("Current values include one or more different node name!")
-  }
-
-
-  return(input_graph)
-
-}

### 2019-05-27 ceRNAnetsim/R/vis_graph.R

@@ -1,55 +1,0 @@
-#' @title vis_graph
-#'
-#' @description  vis_graph provides to visualisation of the graph.
-#'
-#' @return The graph object.
-#'
-#' @details vis_graph ensures the process to be followed.
-#'
-#' @param input_graph The graph object.
-#' @param Competing_color The color of competing elements on the graph with "green" default.
-#' @param mirna_color The color of miRNAs on the graph with "orange" default.
-#' @param Upregulation The color of Upregulated elements on the graph with "red" default.
-#' @param Downregulation The color of Downregulated elements on the graph with "blue" default.
-#' @param title Title of the given graph.
-#' @param layout The layout that will be used for visualisation of the graph.
-#'
-#' @examples
-#'
-#' # Visualisation of graph in steady-state.
-#'
-#' priming_graph(minsamp, Competing_expression, miRNA_expression, aff_factor = c(seed_type,energy), deg_factor = region)%>%
-#'    update_nodes(once = TRUE)%>%
-#'    vis_graph()
-#'
-#' # Visualisation of graph after the change.
-#'
-#' priming_graph(minsamp, Competing_expression, miRNA_expression, aff_factor = c(seed_type,energy), deg_factor = region)%>%
-#'    update_nodes(once = TRUE)%>%
-#'    update_variables(current_counts)%>%
-#'    update_nodes()%>%
-#'    vis_graph()
-#'
-#' @export
-
-vis_graph <-function(input_graph, Competing_color = "green", mirna_color = "orange", Upregulation = "red", Downregulation = "blue", title = "GRAPH", layout= "kk"){
-
-  sample_layout <- create_layout(graph = input_graph, layout = layout)
-
-  if(missing(layout)){
-    warning("Kamada-Kawai (kk) layout was used.
-")
-  }
-
-  input_graph%>%
-    ggraph("manual", node.position=sample_layout)+
-    geom_edge_link(colour = "#838B8B" , alpha = 0.5)+
-    geom_node_point(aes(filter= (changes_variable == "Competing"), color = "Competing", size= count_current), shape= 16)+
-    geom_node_point(aes(filter= (changes_variable == "miRNA"), color = "miRNA"), shape= 16, size= 2)+
-    geom_node_point(aes(filter= (changes_variable == "Up"), color= "Up", size= count_current), shape= 16)+
[…]

### 2019-05-27 ceRNAnetsim/tests/testthat/test-data_convertion.R

@@ -1,81 +1,0 @@
-context("Convertion to the graph.")
-
-
-testdata <- data.frame(competing = c("TP53", "Gene2", "ABCC1", "Gene4", "Gene4", "Gene5", "PTEN"),
-                       miRNA = c("mir1", "mir1","hsa-mir3", "mir1", "MiR2", "MiR2", "miR4"),
-                       Competing_expression = c(10000, 10000, 5000, 10000, 10000, 5000, 10000),
-                       miRNA_expression = c(1000, 1000, 1000, 1000, 2000, 2000, 2000),
-                       stringsAsFactors = FALSE)
-
-testdata2 <- data.frame(competing = c("TP53", "Gene-2", "ABCC1", "Gene 4", "Gene 4", "Gene5", "PTEN"),
-                        miRNA = c("mir1", "mir1","hsa-mir3", "mir1", "MiR2", "MiR2", "miR4"),
-                        Competing_expression = c(10000, 10000, 5000, 10000, 10000, 5000, 10000),
-                        miRNA_expression = c(1000, 1000, 1000, 1000, 2000, 2000, 2000),
-                        stringsAsFactors = FALSE)
-
-testdata3 <- data.frame(competing = c("TP53", "Gene-2", "ABCC1", "Gene 4", "Gene 4", "Gene5", "PTEN"),
-                        miRNA = c("mir1", "mir1","hsa-mir3", "mir1", "MiR2", "MiR2", "miR4"),
-                        Competing_expression = c(NA, 10000, 5000, 10000, 10000, 5000, 10000),
-                        miRNA_expression = c(1000, 1000, 1000, 1000, 2000, 2000, 2000),
-                        stringsAsFactors = FALSE)
-
-test_that("Can dataset be correctly converted to graph ? ", {
-
-
- prime_type_1 <-  (priming_graph(testdata, competing_count = Competing_expression, miRNA_count= miRNA_expression)%N>%
-      as_tibble()%>%
-      filter(type == "miRNA")%>%
-      count())%>%
-   pull()
-
- prime_type_2 <-priming_graph(testdata, competing_count = Competing_expression, miRNA_count= miRNA_expression)%N>%
-     as_tibble()%>%
-     filter(type == "Competing")%>%
-     count()%>%
-   pull()
-
-
- prime_type_3 <-  (priming_graph(testdata2, competing_count = Competing_expression, miRNA_count= miRNA_expression)%N>%
-                     as_tibble()%>%
-                     filter(type == "miRNA")%>%
[…]

### 2019-05-27 ceRNAnetsim/tests/testthat/test-iter_perturb.R

@@ -1,50 +1,0 @@
-context("Simulation and perturbation of network")
-
-test_that("Could the functions work? ",{
-
-  data("midsamp")
-
-calc_test<- midsamp%>%
-    priming_graph(competing_count = Gene_expression, miRNA_count = miRNA_expression)%>%
-    calc_perturbation(sample_n(midsamp, 1)$Genes,3, cycle = 30, limit = 0.1)
-
-perturb_test<- midsamp%>%
-  priming_graph(competing_count = Gene_expression, miRNA_count = miRNA_expression)
-
-
-  expect_is(find_iteration(midsamp,Gene_expression, miRNA_expression, node_name = sample_n(midsamp, 1)$Genes, how= 3), 'numeric')
-  expect_gt(find_iteration(midsamp,Gene_expression, miRNA_expression, node_name = sample_n(midsamp, 1)$Genes, how= 3),  0)
-  expect_equal(attr(calc_test, "names"), c("perturbation_efficiency", "perturbed_count" ))
-  expect_true(is.data.frame(find_node_perturbation(perturb_test)))
-
- })
-
-
-
-test_that("Outputs of the functions", {
-
-data("midsamp")
-
-  classofout<- midsamp%>%
-    priming_graph(competing_count = Gene_expression, miRNA_count = miRNA_expression)%>%
-    update_nodes(once = TRUE)%>%
-    update_how(sample_n(midsamp, 1)$Genes, 3)%>%
-    update_nodes()%>%
-    simulate(15)%>%
-    class()
-
-
-  classofout2<-midsamp%>%
-    priming_graph(competing_count = Gene_expression, miRNA_count = miRNA_expression)%>%
-    update_nodes(once = TRUE)%>%
-    update_how(sample_n(midsamp, 1)$Genes, 3)%>%
-    update_nodes()%>%
-    simulate_vis(4)%>%
-    class()
-
-  expect_equal(classofout, classofout2)
-
-
-})
-
-

### 2019-05-27 ceRNAnetsim/tests/testthat/test-updating_variables.R

@@ -1,71 +1,0 @@
-context("Check to Update")
-
-
-
-test_that("Check the functions which update the values on the datasets.", {
-
-  data("midsamp")
-
-  primerg<- midsamp%>%
-    priming_graph(competing_count = Gene_expression, miRNA_count = miRNA_expression)
-
-  seconderg <- primerg%>%
-    update_nodes(once= TRUE)
-
-  thirdg<- seconderg%>%
-    update_how(sample_n(midsamp, 1)$Genes, 2)
-
-  forth<- thirdg%>%
-    update_nodes()
-
-
-  expect_equal(ncol(as_tibble(primerg%>%activate(nodes))), 3)
-  expect_equal(ncol(as_tibble(seconderg%>%activate(nodes))), 7)
-  expect_equal(ncol(as_tibble(thirdg%>%activate(nodes))), 7)
-  expect_equal(ncol(as_tibble(forth%>%activate(nodes))), 7)
-  expect_equal(ncol(as_tibble(seconderg%>%activate(edges))), 19)
-  expect_equal(ncol(as_tibble(thirdg%>%activate(edges))), 19)
-  expect_equal(ncol(as_tibble(forth%>%activate(edges))), 19)
-
-
-})
-
-
-test_that("Check the functions which update the values on the datasets with interaction factors.", {
-
-  data("midsamp")
-
-  primergf <- midsamp%>%
-    priming_graph(competing_count = Gene_expression, miRNA_count = miRNA_expression, aff_factor = c(seeds, Energy), deg_factor = targeting_region)
-
-  secondergf <- primergf%>%
-    update_nodes(once= TRUE)
-
-  thirdgf <- secondergf%>%
-    update_how(sample_n(midsamp, 1)$Genes, 2)
-
-  forthf <- thirdgf%>%
-    update_nodes()
-
-  midsamp%>%
-    priming_graph(competing_count = Gene_expression, miRNA_count = miRNA_expression)%>%
-    update_nodes(once=TRUE)%>%
-    update_how("Gene15", 0.3)%>%
-    update_nodes()%>%
-    simulate(10)%>%
-    as_tibble()%>%
-    dplyr::mutate(count_current= round(count_current, 2))%>%
-    dplyr::select(count_current)%>%
-    pull() -> outputall
-
-
[…]

### 2019-05-27 article.Rmd

@@
 
-![Shematic presentation of mechanism of network based model.
-a) In steady state, miRNAs (M: triangle grey) repress targets (G: circle coconut) according to proportion of their targets' expression.
+![Schematic presentation of mechanism of network based model.
+a) In steady state, miRNAs (M: gray triangle) repress targets (G: circle coconut) according to proportion of their targets' expression.
 b-c-d) MiRNA expression distributions to targets vary when expression of one of the targets perturbed.](Fig1.jpg){#fig1}
@@
 Transcription rates and degradation rates of miRNAs were assumed as stationary and equal.
-Targets also have stabil transcription and degradation rates, at the same time they are equal.
+Targets also have stable transcription and degradation rates, at the same time they are equal.
 The factors that could be affect the miRNA:target interaction are disregarded in this approach.
-It is considered that miRNAs degradates or inhibits them via any mechanisms after binding to target while the miRNAs were used reversibly.
+It is considered that miRNAs degrades or inhibits them via any mechanisms after binding to target while the miRNAs were used reversibly.
 In initial conditions, gene expression amounts in system have been considered after repression activity of miRNAs.
@@
 Firstly, we obtained the miRNA and gene expression datasets of the patient's tumor and normal tissue.
-We downloaded the high-throughtput experimental datasets which are provided miRNA:gene target pairs with interaction factors by **moore_mirnatarget_2015** and [@helwak_mapping_2013].
+We downloaded the high-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors by **moore_mirnatarget_2015** and [@helwak_mapping_2013].
 We have combined miRNA and gene expression datasets via miRNA:target gene dataset **(Whole steps can be found in Supplementary data(TCGA_E9-A1N5_article.Rmd)**).
@@
 and Denzler et al.[@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
-When the simulation was run in sample dataset that includes lower target abundance, the first response of primary neihgborhood of the same group was determined as similar with in system that have high target abudance.
-But regulations of ceRNAs were observed differences in the folowing steps of simulation.
+When the simulation was run in sample dataset that includes lower target abundance, the first response of primary neighborhood of the same group was determined as similar with in system that have high target abundance.
+But regulations of ceRNAs were observed differences in the following steps of simulation.
 In overall state in the end of simulation, we have established that all regulations more prominent when compared with low miRNA:target ratio dataset.
@@
 For example, Xu et.
-al have investigated the importance of seed pairing type between miRNAs and their targets and target site location by using proteomic dataset [@xu_characterization_2014].
+al have investigated the importance of seed pairing type between miRNAs and their targets and target site location by using proteomics dataset [@xu_characterization_2014].
 They have proposed that the features of binding between miRNA and target can be critical efficiency of miRNA and can be used to find effective miRNA molecules.
@@
 We have tried to simulate a real dataset which contains thousands of genes and hundreds of miRNAs.
-Firstly, we have observed the activities in the iterations and determined behaviours of the network after simulation.
-As expected, we have observed that the big dataset, compared to those containing a few genes and miRNAs, exhibited cumbersome behaviours and complexity to gain steady-state **(see details in Supplementary File(TCGA_E9-A1N5_article.Rmd)**.
+Firstly, we have observed the activities in the iterations and determined behaviors of the network after simulation.
+As expected, we have observed that the big dataset, compared to those containing a few genes and miRNAs, exhibited cumbersome behaviors and complexity to gain steady-state **(see details in Supplementary File(TCGA_E9-A1N5_article.Rmd)**.
 We realized that a small change in a system with a large change actually brought many arrangements.
 This can be explained with the competing groups.
-Because the genes are targeted with many common miRNAs, and this results in that the common targets exhibit competing behaviours in different groups.
+Because the genes are targeted with many common miRNAs, and this results in that the common targets exhibit competing behaviors in different groups.
 On the other hand, the changes of gene expressions were very different.
-While the change in some of genes were highly evident, somes have changed slightly.
+While the change in some of genes were highly evident, some have changed slightly.
 So, we have compared two datasets; the tumor gene expression and simulation results.
-When we analysed comparison of two datasets, we have observed that the results of simulation is not convenient with tumor tissue expression values.
+When we analyzed comparison of two datasets, we have observed that the results of simulation is not convenient with tumor tissue expression values.
 The simulation results were not expected to be concordant with tumor tissue expression values because of a single gene could not be responsible for gene regulation in all tissue.
@@
 Their approach could be useful to analysis interacted genes through miRNAs.
-When it has been throught that a miRNA can exhibit strong functionality to a target but may not against an other, common miRNA based approach can be inconvenient to understand regulations of ceRNA interactions.
+When it has been thought that a miRNA can exhibit strong functionality to a target but may not against an other, common miRNA based approach can be inconvenient to understand regulations of ceRNA interactions.
 

### 2019-05-28 article.Rmd
todo-tofix

@@
 Our findings emphasized importance of miRNA:target ratios being crucial, as reported by previous studies.
-We have showed that the competing elements which have the same or close expression values may not be affected equally from the perturbation because of repression functionality of interactions(TOFIX son kısmı anlayamadım).
+We have showed that the competing elements which have the same or close expression values may not be affected equally from the perturbation because of repression functionality depended on interaction factors of miRNA target pairs (TOFIX: Hocam etkileşim faktörlerinden dolayı ekspresyonları yakın olsa bile baskılanma aktitesinin aynı sergilenmeyeceğini ifade etmeye çalıştım.
+Bu şekilde anlaşılıyor mu?).
 We applied the model to real sample consisting of breast cancer gene and miRNA expression dataset and experimental miRNA:target interaction dataset all generated via high throughput sequencing methods.
@@
 It has been demonstrated that the binding energy between miRNA and target indicates stability or affinity of complex [@helwak_mapping_2013] and does not directly determine repressive activity of miRNA [@cao_predicting_2012].
-Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5’UTR (TOFIX mirna için UTR ifadesi geçerli mi, sadec 5 ucu dense yeterli mi?) bind to specific sequence located in 3’UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005].
+Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5’end bind to specific sequence located in 3’UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005].
 In recent studies, it has been shown that miRNAs can interact with targets via sequences located in regions such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013; @moore_mirnatarget_2015].
-These studies also showed that binding location could indicate functionality of miRNA:target interaction or affect profile of target abundance(TOFIX abundance etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı?).
-It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is perfectly complementary with targets, seed region at the 5’UTR of miRNAs(TOFIX mirna için UTR geçerli mi) [@bartel_micrornas:_2009; @grimson_microrna_2007].
+These studies also showed that binding location could indicate functionality of miRNA:target interaction or be effective on abundance of targets(TOFIX abundance etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı?, evet hocam anlaşılmıyorsa değiştirebilirim).
+It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is perfectly complementary with targets, seed region at the 5’ end of miRNAs [@bartel_micrornas:_2009; @grimson_microrna_2007].
 On the other hand, some researchers have reported that seed sequence of miRNA can have mismatches or bulged/wobble nucleotides and may locate in region other than 5'end of miRNAs [@hafner_transcriptome-wide_2010; @helwak_mapping_2013].
@@
 Briefly, Ala et al.
-have explained the ceRNA hypothesis as distribution (TOFIX disturbance mı kastedildi acaba?) of the second target when a steady-state system that included one miRNA and two target when perturbed with expression changes at one of the targets [@ala_integrated_2013].
+have explained the ceRNA hypothesis as disturbance of the other target when one of the targets on a steady-state system that included one miRNA and two target was perturbed with expression change [@ala_integrated_2013].
 Regarding interaction between miRNAs and their target in a cell, explaining and predicting results of an individual perturbation is difficult due to complexity of interactions.
@@
 Bosson et al.
-have developed a mathematical model for changes on total target pool concentration after grouping targets according to affinity (TOFIX orijinal cümle ile aynı manaya geldi mi emin olamadım) and demonstrated that miRNA activity correlated with affinity between miRNA and target [@bosson_endogenous_2014].
-Cooperative efficiency of miRNAs as well as competitor behaviors of targets were also studied and it has been demonstrated to be crucial for regulating available mRNA levels (TOFIX expression deyince sanki transcriptional bir etki var gibi anlaşılacak diye rna level deme ihtiyacı hissettim) of targets [@denzler_impact_2016].
+have developed a mathematical model for changes on total target pool concentration after grouping targets according to affinity and demonstrated that miRNA activity correlated with affinity between miRNA and target [@bosson_endogenous_2014].
+Cooperative efficiency of miRNAs as well as competitor behaviors of targets were also studied and it has been demonstrated to be crucial for regulating available mRNA levels of targets [@denzler_impact_2016].
 MiRNA:target interactions have been modeled as stoichiometric and catalytic mechanisms and Figliuzzi et al.
@@
 Robinson and Henderson applied the model that handles miRNA:target direct and indirect interactions via common miRNA of gene and target of miRNA, at bipartite network.
-It has been demonstrated that all miRNAs and targets in network can interact each other, bia not necessarily direct interactions, through common miRNAs and genes [@robinson_modelling_2018].
-Associated genes that are targets of the same miRNAs have been found with help of correlation of gene expression changes in recent study(TOFIX bu R paketine atıf veriyor, makale atıfı varsa daha iyi olurdu, makalesi yoksa, recent study yerine, recent algoritm deriz) [@markus_list_sponge_2017].
+It has been demonstrated that all miRNAs and targets in the network can interact with each other through common miRNAs and genes, without interaction between the same type of nodes [@robinson_modelling_2018].Associated genes that are targets of the same miRNAs have been found with help of correlation of gene expression changes in recent algoritm [@markus_list_sponge_2017].
 List et al.

### 2019-05-28 article.tex
todo-tofix

@@
   \postdate{\par}
-    \date{17 May 2019}
+    \date{28 Mayıs 2019}
 
@@
 
-Competing endogeneous RNA (ceRNA) regulations and crosstalk in human are
-remarkable issue in mean of regulations via miRNAs. Many studies have
-pointed out that an alteration in miRNA:target interaction can result in
-unexpected regulations because of indirect and complex interaction. In
-this paper, we defined a new network-based model that handles
-miRNA:ceRNA interactions with expression values and interaction factors
-optionally. We applied approach in sample dataset and defined dynamics
-of it. We have observed that a perturbation in an element of network
-affects whole competing elements differently and cooperative
-efficiencies of miRNAs on common targets could be shown. Importance of
-miRNA:target ratios was also observed consistently with previous
-reports. We have obtained that the competing elements which have the
-same or close expression values may not affected equal from the
-perturbation because of repression functionality of interactions. To
-apply model to the real sample that obtained combining breeast cancer
-gene and miRNA expression dataset via high throughput experimental
-miRNA:target dataset, we selected a gene, \emph{ABCC1}, which over
-expressed in tumor tissue and utilized it as perturbating element. We
-have observed that one change can be sufficient to perturb regulations,
-and may not predicted all of regulations correctly, with this change.
-Therefore, this model which helps to understand crosstalking between
-elements could be tested and improved via results of well-constructed
-experimental dataset and takes to hope for future research.
+Competing endogenous RNA (ceRNA) regulations and crosstalk between
+various types of non-coding RNA in human is remarkable in means of miRNA
+regulation. Many studies have pointed out that an alteration in
+miRNA:target interaction can result in unexpected changes due to
+indirect and complex interactions. In this paper, we defined a new
+network-based model that handles miRNA:ceRNA interactions with
+expression values. Our model is able to handle miRNA interaction factors
+such as seed type, binding energy, if provided. Our approach is able to
+reveal that a perturbation in an element of network affects whole
+competing elements differently and cooperative efficiencies of miRNAs on
+common targets could be calculated. Our findings emphasized importance
+of miRNA:target ratios being crucial, as reported by previous studies.
+We have showed that the competing elements which have the same or close
+expression values may not be affected equally from the perturbation
+because of repression functionality depended on interaction factors of
+miRNA target pairs (TOFIX: Hocam etkileşim faktörlerinden dolayı
+ekspresyonları yakın olsa bile baskılanma aktitesinin aynı
+sergilenmeyeceğini ifade etmeye çalıştım. Bu şekilde anlaşılıyor mu?).
+We applied the model to real sample consisting of breast cancer gene and
+miRNA expression dataset and experimental miRNA:target interaction
+dataset all generated via high throughput sequencing methods. A gene
+over-expressed in tumor tissue, namely \emph{ABCC1}, is used as
+perturbing element. We have observed that change in expression level of
+single gene in miRNA:target network is sufficient to perturb regulations
+in whole network, due to unforeseen and unpredicted regulation which are
+only visible when considered in network context. Therefore, this model
+helps unveiling the crosstalk between elements in miRNA:target network
+where abundance of target and sponge effect are taken into account. The
+model is scalable and can be plugged in with emerging miRNA effectors
+such as circRNAs. The model is available as R package at
+\url{https://github.com/selcenari/ceRNAnetsim}. (TODO the link will
+change)
 
@@
 
-MicroRNAs (miRNAs) that are key regulator of gene expression through
-various post-transcriptional mechanisms are a family of short non-coding
-RNAs. Although repression mechanisms of miRNA are not fully understood,
-exact effect of miRNAs is regulation of direct target/targets
-negatively. Repressive activities of miRNAs vary depending on many
-factors that are significant to microRNA:target interactions. These
-factors include microRNA:target binding energy, binding location in
-target sequence, base pairing types between miRNA and target, abundance
-of miRNAs and targets (Grimson et al. 2007). Binding energies of
-miRNA:target complexes vary based on nucleotide context and determine
-folding stability of complex (Cao and Chen 2012). It has been
-demonstrated that the binding energy between miRNA and target indicates
-stability or affinity of complex (Helwak et al. 2013) and does not
-directly determine repressive activity of miRNA (Cao and Chen 2012).
-Early studies have argued that 2-8 nt sequence, seed, located in miRNA
-5'UTR bind to specific sequence located in 3'UTR of its target (Bartel
-2004; Lewis, Burge, and Bartel 2005). In recent studies, it has been
-shown miRNAs can interact with targets via sequence that located in
-different region such as 5'UTR or CDS (Hausser et al. 2013; Helwak et
-al. 2013) \textbf{moore\_mirnatarget\_2015} (Moore et al. 2015). These
-studies also shown that binding location could indicate functionality of
-miRNA:target interaction or affect profile of target abundance. It has
-been shown that miRNAs exhibit repressive activity via, 6-8 nt long
-sequence that is prefectly complementary with targets, seed region at
-the 5'UTR of miRNAs (Bartel 2009; Grimson et al. 2007). On the other
-hand, some researchers have reported that seed sequence of miRNA can
-have mismatches or bulged/wobble nucleotides and locates in different
-region of miRNAs (Hafner et al. 2010; Helwak et al. 2013). In addition
-to all of these, abundance of miRNAs and targets and miRNA:target ratio
-in cells are determinant for efficiency of miRNA:target interaction
-(Arvey et al. 2010; Bosson, Zamudio, and Sharp 2014; Denzler et al.
-2014).
+MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key
+regulator of gene expression through various post-transcriptional
+mechanisms. Although the mechanisms by which miRNA represses are not
+fully understood, miRNAs predominantly repress their targets. Repressive
+activities of miRNAs vary depending on many factors that are significant
+to microRNA:target interactions. These factors include microRNA:target
+binding energy, binding location in target sequence, base pairing types
+between miRNA and target, abundance of miRNAs and targets (Grimson et
+al. 2007). Binding energies of miRNA:target complexes vary based on
+nucleotide context and determine folding stability of complex (Cao and
+Chen 2012). It has been demonstrated that the binding energy between
+miRNA and target indicates stability or affinity of complex (Helwak et
+al. 2013) and does not directly determine repressive activity of miRNA
+(Cao and Chen 2012). Early studies have argued that 2-8 nt sequence,
+seed, located in miRNA 5'end bind to specific sequence located in 3'UTR
+of its target (Bartel 2004; Lewis, Burge, and Bartel 2005). In recent
+studies, it has been shown that miRNAs can interact with targets via
+sequences located in regions such as 5'UTR or CDS (Hausser et al. 2013;
+Helwak et al. 2013; Moore et al. 2015). These studies also showed that
+binding location could indicate functionality of miRNA:target
+interaction or be effective on abundance of targets(TOFIX abundance
+etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı?,
+evet hocam anlaşılmıyorsa değiştirebilirim). It has been shown that
+miRNAs exhibit repressive activity via, 6-8 nt long sequence that is
+perfectly complementary with targets, seed region at the 5' end of
+miRNAs (Bartel 2009; Grimson et al. 2007). On the other hand, some
+researchers have reported that seed sequence of miRNA can have
+mismatches or bulged/wobble nucleotides and may locate in region other
+than 5'end of miRNAs (Hafner et al. 2010; Helwak et al. 2013). On top of
+all these factors, abundance of miRNAs and targets and miRNA:target
+ratio in cells predominantly affect efficiency of miRNA:target
+interaction (Arvey et al. 2010; Bosson, Zamudio, and Sharp 2014; Denzler
+et al. 2014).
 
 As it is possible for miRNAs to suppress multiple targets, an individual
-RNA molecule can also be targeted by various miRNAs. In that case, the
-targeted RNAs exhibit competitor behavior, that is hypothesized as
+mRNA molecule can also be targeted by multiple miRNAs. In that case, the
+targeted mRNAs exhibit competitor behavior, that is hypothesized as
 competing endogenous RNAs (ceRNAs) (Ala et al. 2013; Cesana and Daley
 2013), against their miRNAs. Briefly, Ala et al.~have explained the
-ceRNA hypothesis as disturbtion of the other target when a steady-state
-system that included one miRNA and two target, was perturbed with
-expression changes at one of the targets (Ala et al. 2013). When
-interaction between miRNAs and their target were dealed in the cell,
-explanation and prediction of results of perturbation is difficult due
-to varieties of interactions. Various computational and experimental
-studies have applied to be understood for ceRNA:miRNA interactions. It
-has been observed that when abundance of one of the targets of miR-122
-was increased, the other target expression also slightly increased as a
-result of decreasing repressive activity of miR-122 on them (Denzler et
-al. 2014). Bosson et al.~have developed a mathematical model for changes
-on total target pool concentration after grouping to affinity of
-targets. They have demonstrated that miRNA activity correlated with
-affinity between miRNA and target (Bosson, Zamudio, and Sharp 2014).
-Cooperative efficiency of miRNAs as well as competitor behaviors of
-targets also were studied and it has been demonstrated crucial for
-regulations at expression of targets (Denzler et al. 2016). MiRNA:target
-interactions have modelled as stoichiometric and catalytic mechanisms
-and Figliuzzi et al.~have recommended to handling models in network
-structures (Figliuzzi, Marinari, and De Martino 2013). The model that
-can explain miRNA target interaction through topological features has
-been applied at bipartite network by Nitzan et al.~(Nitzan et al. 2014).
-Robinson and Henderson applied the model that handles miRNA:target
-direct and indirect interactions via common miRNA of gene and target of
-miRNA, at bipartite network. It has been demonstrated that whole miRNAs
-and targets in network can interacted through common miRNAs and genes
-(Robinson and Henderson 2018). Associated genes that are targets of the
-same miRNAs have been found with help of correlation of gene expression
-changes in recent study (Markus List 2017). List et al.~have specified
-that their approach can be useful for ceRNA's studies and published
-these approach as a package that can use with R programming.
+ceRNA hypothesis as disturbance of the other target when one of the
+targets on a steady-state system that included one miRNA and two target
+was perturbed with expression change (Ala et al. 2013). Regarding
+interaction between miRNAs and their target in a cell, explaining and
+predicting results of an individual perturbation is difficult due to
+complexity of interactions. Various computational and experimental
+studies have tackled the problem of unraveling ceRNA:miRNA interactions.
+It has been observed that when abundance of one of the targets of
+miR-122 was increased, the other target expression also slightly
+increased as a result of decreasing repressive activity of miR-122 on
+them (Denzler et al. 2014). Bosson et al.~have developed a mathematical
+model for changes on total target pool concentration after grouping
+targets according to affinity and demonstrated that miRNA activity
+correlated with affinity between miRNA and target (Bosson, Zamudio, and
+Sharp 2014). Cooperative efficiency of miRNAs as well as competitor
+behaviors of targets were also studied and it has been demonstrated to
+be crucial for regulating available mRNA levels of targets (Denzler et
+al. 2016). MiRNA:target interactions have been modeled as stoichiometric
+and catalytic mechanisms and Figliuzzi et al.~have recommended handling
+models in network context (Figliuzzi, Marinari, and De Martino 2013).
+The model that can explain miRNA target interaction through topological
+features has been applied at bipartite network by Nitzan et al.~(Nitzan
+et al. 2014). Robinson and Henderson applied the model that handles
+miRNA:target direct and indirect interactions via common miRNA of gene
+and target of miRNA, at bipartite network. It has been demonstrated that
+all miRNAs and targets in the network can interact with each other
+through common miRNAs and genes, without interaction between the same
+type of nodes (Robinson and Henderson 2018).Associated genes that are
+targets of the same miRNAs have been found with help of correlation of
+gene expression changes in recent algoritm (Markus List 2017). List et
+al.~have specified that their approach can be useful for ceRNA studies
+and published their approach as an R package.
 
@@
 \includegraphics{Fig1.jpg}
-\caption{Shematic presentation of mechanism of network based model. a)
-In steady state, miRNAs (M: triangle grey) repress targets (G: circle
+\caption{Schematic presentation of mechanism of network based model. a)
+In steady state, miRNAs (M: gray triangle) repress targets (G: circle
 coconut) according to proportion of their targets' expression. b-c-d)
@@
 and degradation rates of miRNAs were assumed as stationary and equal.
-Targets also have stabil transcription and degradation rates, at the
+Targets also have stable transcription and degradation rates, at the
 same time they are equal. The factors that could be affect the
 miRNA:target interaction are disregarded in this approach. It is
-considered that miRNAs degradates or inhibits them via any mechanisms
+considered that miRNAs degrades or inhibits them via any mechanisms
 after binding to target while the miRNAs were used reversibly. In
@@
 datasets of the patient's tumor and normal tissue. We downloaded the
-high-throughtput experimental datasets which are provided miRNA:gene
+high-throughput experimental datasets which are provided miRNA:gene
 target pairs with interaction factors by
@@
 sample dataset that includes lower target abundance, the first response
-of primary neihgborhood of the same group was determined as similar with
-in system that have high target abudance. But regulations of ceRNAs were
-observed differences in the folowing steps of simulation. In overall
-state in the end of simulation, we have established that all regulations
-more prominent when compared with low miRNA:target ratio dataset. The
-model has appeared the importance of mRNA:target ratio in regulation of
-competing RNAs.
+of primary neighborhood of the same group was determined as similar with
+in system that have high target abundance. But regulations of ceRNAs
+were observed differences in the following steps of simulation. In
+overall state in the end of simulation, we have established that all
+regulations more prominent when compared with low miRNA:target ratio
+dataset. The model has appeared the importance of mRNA:target ratio in
+regulation of competing RNAs.
 
@@
 al have investigated the importance of seed pairing type between miRNAs
-and their targets and target site location by using proteomic dataset
+and their targets and target site location by using proteomics dataset
 (Xu, Wang, and Liu 2014). They have proposed that the features of
@@
 thousands of genes and hundreds of miRNAs. Firstly, we have observed the
-activities in the iterations and determined behaviours of the network
+activities in the iterations and determined behaviors of the network
 after simulation. As expected, we have observed that the big dataset,
 compared to those containing a few genes and miRNAs, exhibited
-cumbersome behaviours and complexity to gain steady-state \textbf{(see
+cumbersome behaviors and complexity to gain steady-state \textbf{(see
 details in Supplementary File(TCGA\_E9-A1N5\_article.Rmd)}. We realized
@@
 Because the genes are targeted with many common miRNAs, and this results
-in that the common targets exhibit competing behaviours in different
+in that the common targets exhibit competing behaviors in different
 groups. On the other hand, the changes of gene expressions were very
-different. While the change in some of genes were highly evident, somes
+different. While the change in some of genes were highly evident, some
 have changed slightly. So, we have compared two datasets; the tumor gene
-expression and simulation results. When we analysed comparison of two
+expression and simulation results. When we analyzed comparison of two
 datasets, we have observed that the results of simulation is not
@@
 (Markus List 2017). Their approach could be useful to analysis
-interacted genes through miRNAs. When it has been throught that a miRNA
+interacted genes through miRNAs. When it has been thought that a miRNA
 can exhibit strong functionality to a target but may not against an
@@
 \leavevmode\hypertarget{ref-figliuzzi_micrornas_2013}{}%
-Figliuzzi, Matteo, Enzo Marinari, and De MartinoAndrea. 2013.
+Figliuzzi, Matteo, Enzo Marinari, and Andrea De Martino. 2013.
 ``MicroRNAs as a Selective Channel of Communication Between Competing

### 2019-07-22 article.Rmd
rewrite method section

@@
 author: 
-- Selcen Ari^[Yildiz Technical University Department of Bioengineering]
-- Alper Yilmaz^[Yildiz Technical University Department of Bioengineering]
+- Selcen Ari^[Department of Bioengineering, Yildiz Technical University, Istanbul, Turkey]
+- Alper Yilmaz^[Department of Bioengineering, Yildiz Technical University, Istanbul, Turkey]
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 Robinson and Henderson applied the model that handles miRNA:target direct and indirect interactions via common miRNA of gene and target of miRNA, at bipartite network.
-It has been demonstrated that all miRNAs and targets in the network can interact with each other through common miRNAs and genes, without interaction between the same type of nodes [@robinson_modelling_2018].Associated genes that are targets of the same miRNAs have been found with help of correlation of gene expression changes in recent algoritm [@markus_list_sponge_2017].
+It has been demonstrated that all miRNAs and targets in the network can interact with each other through common miRNAs and genes, without interaction between the same type of nodes [@robinson_modelling_2018].Associated genes that are targets of the same miRNAs have been found with help of correlation of gene expression changes in recent algorithm [@markus_list_sponge_2017].
 List et al.
@@
 ## Methods
-### miRNA:target ratio theoretical action in network based ceRNA model
-We developed a novel network based approach that can use to simulation of gene regulation after any perturbation at miRNA:target interactions.
-If a perturbation or change induces on graph, whole interacted nodes vary because of changes in miRNA:target ratios.
-The system was assumed as steady state that miRNAs affect according to the ratio of the individual target to all targets (\autoref{fig1}a, expression values shown as integer for comprehensible sampling).
-When expression of gene 2 (G2) increased to two fold (\autoref{fig1}b), distribution of miRNA (M1) expression changes (\autoref{fig1}c).
-Consequently, expression of common target (G4) of miRNA1 and miRNA2 also varies.
-The change at expression of common target triggers to changes of proportional distribution of miRNA2 (\autoref{fig1}d).
+### Construction of miRNA:target network
+miRNA and target pairs per line should be provided as edge list to construct the network.
+At each line minimum required information is expression levels of miRNA and the target.
+If available, additional data about factors effecting binding or efficiency of miRNA can be provided as separate columns.
+After construction of the network, amount of miRNA per target is calculated and kept as edge data.
+Simply, a target will sequester miRNA proportional to its ratio amount among other targets.
+If additional criteria effecting the binding of miRNA to its target is provided, distribution of miRNA will be calculated accordingly.
+Target can be mRNA or any other ceRNA (circRNA, ncRNA, etc.) thus, throughout the manuscript terms target, gene and ceRNA are used interchangeably.
+
+### Triggering perturbation and subsequent calculations
+Initially, the network is assumed in steady-state (\autoref{fig1}a) condition and needs a trigger for initiating calculations.
+The trigger can be a change in expression level of one or more genes (\autoref{fig1}b).
+After a trigger, the network undergoes iterative cycle of calculations at each of which distribution of miRNA in local neighborhood is recalculated (\autoref{fig1}c).
+Based on new miRNA distribution, expression level of each node (i.e.
+ceRNA) is updated.
+Due to nodes which are targeted by multiple miRNAs, the change in one neighborhood spreads to other neighborhoods (\autoref{fig1}d), consequently have potential to effect whole network due to "ripple effect".
 
@@
 a) In steady state, miRNAs (M: gray triangle) repress targets (G: circle coconut) according to proportion of their targets' expression.
-b-c-d) MiRNA expression distributions to targets vary when expression of one of the targets perturbed.](Fig1.jpg){#fig1}
+b) Two fold increase in transcript level of Gene2 (G2) acts as a trigger.
+c) Distribution of miRNA1 (M1) changes.
+d) The change at expression of common target effects changes of proportional distribution of miRNA2 (M2).
+Expression values are rounded to integers for simplicity](Fig1.jpg){#fig1}
 
-The model was developed by making some assumptions.
-Transcription rates and degradation rates of miRNAs were assumed as stationary and equal.
-Targets also have stable transcription and degradation rates, at the same time they are equal.
-The factors that could be affect the miRNA:target interaction are disregarded in this approach.
-It is considered that miRNAs degrades or inhibits them via any mechanisms after binding to target while the miRNAs were used reversibly.
-In initial conditions, gene expression amounts in system have been considered after repression activity of miRNAs.
-If a change occurs in the system, variation on miRNA proportional distribution are reflected as change of repression activity.
+(#TODO caption needs more text, every color should be explained)
+
+During calculations, following assumptions were adopted; 1) Transcription and degradation rates of miRNAs are steady and equal.
+2) All available miRNAs are recycled as in miRNA:ceRNA binding, target is degraded and miRNA is unaffected.
+3) ceRNA targets also have stable transcription and degradation rates and these rates are equal.
+
+The repression efficiency of a miRNA on the individual target ($Eff_{gi}$) is calculated according to equation \eqref{eq:1}; where miRNA expression ($C_m$) in local neighborhood is distributed among targets using individual gene expression levels ($C_{gi}$) (#TODO isn't $C_g$ gene expressions in the group is equal to total of $C_{gi}$ in group?).
+For the genes targeted by multiple miRNAs, cooperative activity of miRNAs, $R$, is calculated by summing repression activity of each miRNA (Equation \eqref{eq:2}).
+
 \begin{equation} 
-    Eff_{gi}= C_m * C_{gi}/\sum_{1}^{i} C_g \tag{1}\label{eq:1}
+    Eff_{gi}= C_m \times C_{gi}/\sum_{1}^{i} C_g \tag{1}\label{eq:1}
 \end{equation}
 \begin{equation}
-   R_{gi}= Eff_{i1}+ Eff_{i2} ... \tag{2}\label{eq:2}
+   R_{gi}= Eff_{i1} + Eff_{i2} ... \tag{2}\label{eq:2}
 \end{equation}
 
-The efficiency of a miRNA on the individual target ($Eff_{gi}$) is calculated according to equation \eqref{eq:1}; $C_m$, miRNA expression in the group; $C_{gi}$, symbolize individual gene expression; $C_g$ gene expressions in the group (Groups are determined depending on miRNAs.
-Targets that are interacted with the same miRNA find in a group.).
-In this approach, because a gene can be found in different groups, the cooperative activity of microRNAs also can be calculated.
-Total repression activity of different miRNAs on the common target, $R$, is calculated based on equation \eqref{eq:2}; sum of efficiency of different miRNAs on the target.
+### Multifactorial calculations in miRNA:target network
 
-![Calculations to determine of miRNA binding and repression efficiency.
-$G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE’$, normalized values of seed type efficiency coefficient; $RE’$, normalized values of region efficiency coefficient; $E’$, normalized values of energy coefficient.](fig2.jpg){#fig2}
-
-### miRNA action in network based multifactorial ceRNA model
-
-Interactions between miRNAs and their targets can be affected from different parameters.
-For this reason, we improved model with factors that can affect to miRNA activity.
-In our approach, the parameters are classified as binding and efficiency factors and all of these are evaluated in their groups.
-Binding factors determine interaction between miRNA and target.
-The efficiency factors decide functionality of binding.
-In the literature, binding free energy [@cao_predicting_2012 ; @helwak_mapping_2013] and seed type [@werfel_preferential_2017] in miRNA:target interactions are represented as binding affinity.
-Efficiency factors determine how many of microRNA:target complexes will result in inhibition.
-For example, binding region on the target shows whether the target can be inhibited (functionality)[@hausser_analysis_2013; @helwak_mapping_2013].
-Both of binding and efficiency factors are evaluated in the groups and normalized according to equation \eqref{eq:3}; each factor is normalized based on maximum variables that found in their groups.
-\begin{equation}
-F`= F/F_{max} \tag{3}\label{eq:3}
-\end{equation}
-
+Interactions between miRNAs and their targets can be affected from various factors.
+So, our model integrates multiple factors when calculating overall miRNA activity.
+We classified factors into two categories.
+Factors effecting binding determine interaction between miRNA and target and they alter amount of miRNA sequestered to target.
+Factors effecting efficiency determine proportion of target mRNA degraded by bound miRNA (#TODO fix this sentence).
+In other words, binding factors exert their influence before or during binding, efficiency factors exert their influence after binding.
+In the literature, binding free energy [@cao_predicting_2012 ; @helwak_mapping_2013] and seed type [@werfel_preferential_2017] in miRNA:target interactions are described as factors effecting binding affinity.
+Efficiency factors determine how many of miRNA:target complexes will result in inhibition and binding region on target drastically effect miRNA degradation efficiency [@hausser_analysis_2013; @helwak_mapping_2013].
+Both binding and efficiency factors are normalized to their maximum values and scaled to [0,1] interval.
 The normalized values of factors take into account to determine binding activity and miRNA efficiency on targets (\autoref{fig2}).
@@
 \begin{equation}
-Eff_{gi}= C_m * E`_{gi}*STE`_{gi}*C_{gi}/(\sum_{1}^{i} E`_{gi}*STE`_{gi}*C_{gi}) \tag{4}\label{eq:4}
+Eff_{gi}= C_m \times E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}/(\sum_{1}^{i} E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}) \tag{4}\label{eq:4}
 \end{equation}
 \begin{equation}
-Eff_{gi}= Eff_{gi}*RE`_{gi} \tag{5}\label{eq:5}
+Eff_{gi}= Eff_{gi}*RE^\prime_{gi} \tag{5}\label{eq:5}
 \end{equation}
+
+![Calculations to determine of miRNA binding and repression efficiency.
+$G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE^\prime$, normalized values of seed type efficiency coefficient; $RE^\prime$, normalized values of region efficiency coefficient; $E^\prime$, normalized values of energy coefficient.](fig2.jpg){#fig2}
 
@@
 The functionality of bound miRNA on target depends on efficiency factors like region that is binding sequence of miRNA on its target.
-Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:5} (\autoref{fig2}d); $RE’_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
+Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:5} (\autoref{fig2}d); $RE^\prime_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
 The cooperative repression activity of miRNAs to their common targets is figured out as shown in \autoref{fig2}e.
 
-### Application of model in a breast cancer patient dataset
+### Breast cancer patient dataset
 
-To determine suitability of method, we have tried to apply model in a real dataset.
-Firstly, we obtained the miRNA and gene expression datasets of the patient's tumor and normal tissue.
-We downloaded the high-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors by **moore_mirnatarget_2015** and [@helwak_mapping_2013].
-We have combined miRNA and gene expression datasets via miRNA:target gene dataset **(Whole steps can be found in Supplementary data(TCGA_E9-A1N5_article.Rmd)**).
-We selected a gene which is over expressed in tumor tissue with compared to normal tissue.
-It is pointed out that ABCC1 gene, *Multidrug resistance-associated protein 1*, is one of the most significant factor to develop resistance against chemotherapotic agents[@atalay2008role; @lu2015microrna; @atalay2006multidrug].
-So, we selected ABCC1 gene for simulation of integrated dataset.
-We have detected the iteration that is suitable for ABCC1 gene simulation.
-After simulation of system via ABCC1 gene with over expression, we have compared simulation results and tumor tissue expression levels.
+We have applied our model in a real dataset for which experimental measurements of various factors were available.
+Expression levels of miRNA and genes in tumor and normal tissue of single  patient are retrieved from TCGA(#TODO cite).
+High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015].
+We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from ...
+(#TODO name of database and its citation).
+Detailed description of network construction and its code is available in Supplementary data (#TODO burada link nasıl olacak, word dosyası veya PDF dosyası ismi mi vermemiz mi  gerekiyor, yoksa sadece Sup Data denilmesi yeterli mi?
+ilgili dosya: TCGA_E9-A1N5_article.Rmd).
+ABCC1 gene, over-expressed in tumor tissue, was selected as trigger for simulation of integrated dataset.
+After simulation of network, we have compared simulation results and tumor tissue expression levels.
+
+(#TODO about formulas, 1) eq:3 F/Fmax can be removed, 2) in formulas eq:1 and eq:2, C_m or Eff_gi is not intuitive, is there consensus terms about effect or nth gene?
+3) in formula eq:4, instead of, 4) in general, `*` sign should be replaced with `\times` and apostrophe should be replaced with `\prime` )
 
@@
 
+(#TODO BRCA results go here) ABCC1 gene, *Multidrug resistance-associated protein 1*, is one of the most significant factor to develop resistance against chemotherapotic agents[@atalay2008role; @lu2015microrna; @atalay2006multidrug]
+
 ## References

### 2019-07-25 article.Rmd
question about equation
editing results. used markdown inline comment

@@
 ### Multifactorial calculations in miRNA:target network
-
 Interactions between miRNAs and their targets can be affected from various factors.
@@
 $G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE^\prime$, normalized values of seed type efficiency coefficient; $RE^\prime$, normalized values of region efficiency coefficient; $E^\prime$, normalized values of energy coefficient.](fig2.jpg){#fig2}
+
+\iffalse
+%% (#TODO the equation is not considering binding factors, right?. We should explain the calculation made in  simulate function  here, in simple sentences. the function looks like this: 
+%%    input_graph%>%
+%%      tidygraph::activate(edges)%>%
+%%      group_by(to)%>%
+%%      mutate(mirna_count_per_comp = mirna_count_current*comp_count_current*afff_factor/sum(comp_count_current*afff_factor), mirna_count_per_comp = ifelse(is.na(mirna_count_per_comp), 0, mirna_count_per_comp))%>%
+%%      ungroup()%>%
+%%      mutate(effect_pre = effect_current, effect_current = mirna_count_per_comp*degg_factor)%>%
+%%      group_by(from)%>%
+%%      mutate(comp_count_pre = ifelse(comp_count_current< 0, 1, comp_count_current), comp_count_current = comp_count_pre - (sum(effect_current)-sum(effect_pre)), comp_count_current = ifelse(comp_count_current< 0, 1, comp_count_current))%>%
+%%      ungroup()%>%
+%%      mutate(comp_count_list = pmap(list(comp_count_list, comp_count_current), c), effect_list = pmap(list(effect_list, effect_current), c), mirna_count_list = pmap(list(mirna_count_list, mirna_count_current), c))
+%% )
+\fi
 
@@
 ## Results and Conclusions
+### Networks with single factor
+We have developed a network-based approach to assess effects of expression level changes in competitive ceRNA regulation.
+The basic mode of miRNA repression activity has been based on miRNA and target abundance in various researches[@arvey_target_2010; @denzler_assessing_2014].
+Our approach can effortlessly calculate effects of expression changes when abundance levels of miRNAs and targets is only available factor.
+In sample network given in Figure (\autoref{fig1}), after an increase in expression level of a gene (G2), expression values of other genes also changed due to redistribution of miRNA among its targets.
+Previous studies have shown that if a gene abundance increases in ceRNA system, expression levels of genes targeted by shared miRNA are also affected [@lai_understanding_2016; @salmena_cerna_2011; @tay_multilayered_2014].
+It was observed that primary neighborhoods of the trigger gene change with the same ratio, but cooperative efficiency of miRNAs on common target causes differently change in common target (#TODO last sentence needs some clarification, after "but cooperative...") (\autoref{fig1}d).
+Genes targeted by multiple miRNAs act as a trigger for adjacent local neighborhood of targeting miRNAs, causing changes in expression levels of genes outside the local neighborhood of original trigger gene.
+Therefore, primary expression level change in gene (G2) causes changes in other group of genes (G5 and G6) even though original trigger gene (G2) and genes in other group are not targeted by common miRNA.
+In addition, as shown by ceRNA hypothesis model of Ala et al., after the increase of gene expression level of G2, the miRNA that is found in the same group (M1) tended to be more (#TODO isn't it less?) repressive on its remaining targets (G1, G3 and G4).
+It's important to note that the changes in gene expression levels will have more pronounced effect if miRNA:target ratio is high, i.e., more miRNA available per target, which was reported in previous findings [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
+[][#TODO]
 
-We have developed a model approach and workflow for competitive ceRNA regulation in this study.
-The basic mode of miRNA repression activity has been based on miRNA and target abundance in various researches[@arvey_target_2010; @denzler_assessing_2014].
-For this reason, we performed first application as considering only abundance of miRNAs and targets (\autoref{fig1}).
-After the increasing in a gene expression value, expression values of other genes also changed slightly in comparison with alteration in perturbation started gene node.
-Different studies have shown that the other gene expression values also rise differently, if a gene abundance increases in ceRNA system [@lai_understanding_2016; @salmena_cerna_2011; @tay_multilayered_2014].
-It was observed that primary neighborhoods of the trigger gene change with the same ratio, but cooperative efficiency of miRNAs on common target causes differently change in common target (\autoref{fig1}d).
-The common target in system has displayed like a trigger for the other group and induced changes of expression values of genes on the other group.
-Therefore, effect of primary change of the gene (G2) on genes of the other group was observed slightly (\autoref{fig1}d).
-In addition, like shown by ceRNA hypothesis model of Ala et al., after the increasing of a gene expression (G2), the miRNA that is found in the same group (M1) tended to be more repressive on this target.
-This has caused to decrease in increased expression value of primer triggering gene.
-When the expression of a gene (G2) was increased to two fold, we observed that the expression changes were more evident.
-So, we have considered that high miRNA:target ratio results in stronger miRNA repression activity.
-This results also are coherent with previous reports that were offered by Arvey et al.
-and Denzler et al.[@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
+[#TODO]: <> (following 4 sentences can be deleted, not so clear.
+OR we can summarize it one-two sentences.
 When the simulation was run in sample dataset that includes lower target abundance, the first response of primary neighborhood of the same group was determined as similar with in system that have high target abundance.
@@
 In overall state in the end of simulation, we have established that all regulations more prominent when compared with low miRNA:target ratio dataset.
-The model has appeared the importance of mRNA:target ratio in regulation of competing RNAs.
+The model has appeared the importance of mRNA:target ratio in regulation of competing RNAs.)
 
@@
 
+### Multifactorial networks
+(#TODO title should be more informative)
 It has been reported in several studies that miRNA regulatory interactions also is affected different parameters.

### 2019-07-26 article.Rmd
finished results section

@@
 
+[#TODO]: <> "there are many et. al and et. al. , with or without dot(.) please make sure the formatting is correct and consistent"
+
 As it is possible for miRNAs to suppress multiple targets, an individual mRNA molecule can also be targeted by multiple miRNAs.
@@
 
-(#TODO caption needs more text, every color should be explained)
+[#TODO]: <> "caption needs more text, every color should be explained"
 
@@
 
-(#TODO about formulas, 1) eq:3 F/Fmax can be removed, 2) in formulas eq:1 and eq:2, C_m or Eff_gi is not intuitive, is there consensus terms about effect or nth gene?
-3) in formula eq:4, instead of, 4) in general, `*` sign should be replaced with `\times` and apostrophe should be replaced with `\prime` )
+[#TODO]: <> "about formulas, 1) eq:3 F/Fmax can be removed, 2) in formulas eq:1 and eq:2, C_m or Eff_gi is not intuitive, is there consensus terms about effect or nth gene?
+3) in formula eq:4, instead of, 4) in general, `*` sign should be replaced with `\times` and apostrophe should be replaced with `\prime` )"
 
@@
 ### Multifactorial networks
-(#TODO title should be more informative)
-It has been reported in several studies that miRNA regulatory interactions also is affected different parameters.
+
+[#TODO]: <> "title should be more informative"
+
+Earlier studies reported that miRNA regulatory interactions are affected by different parameters.
 For example, Xu et.
 al have investigated the importance of seed pairing type between miRNAs and their targets and target site location by using proteomics dataset [@xu_characterization_2014].
-They have proposed that the features of binding between miRNA and target can be critical efficiency of miRNA and can be used to find effective miRNA molecules.
-In addition, binding energy between miRNA and targets also could be determinant for miRNA efficiency.
-It has been reported by Breda et.
-al.
-that strength of miRNA:target interactions is depended on binding energy of complexes [@breda_quantifying_2015].
-As similar to these, Bosson et al.
-have associated affinity and seed pairing of miRNA target pairs and suggested that number of canonical base pairing is correlated with affinity.
-For these reasons, we have improved the model with optional interaction parameters that could be useful for understanding of miRNA repression activity.
-We have explained the model in the same sample dataset with using different factors and simulated via the same change (\autoref{fig3}).
+They have proposed that the features of binding between miRNA and target can be critical for miRNA efficiency.
+In addition, binding energy between miRNA and targets is a significant determinant for miRNA efficiency and it has been reported that strength of miRNA:target interactions is depended on binding energy of complexes [@breda_quantifying_2015].
+Similarly, another study revealed that affinity is correlated with seed pairing of miRNA:target pairs and suggested that affinity is correlated with number of canonical seed base pairing [@bosson_endogenous_2014].
+Therefore, we have integrated aforementioned interaction parameters that could be useful for predicting miRNA repression activity more accurately.
+The sample dataset used in  Figure \autoref{fig1} is recalculated with additional factors in effect [][#TODO] with same trigger, two fold increase in Gene2 (\autoref{fig3}).
 When the factors were taken into account in the system, miRNA efficiencies varied as shown in \autoref{fig3}a.
-Although the miRNA target ratios in initial states were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed.
-When expression of a gene (G2) increased, expression values of all genes also change differentially because of contribution of efficiency factors (\autoref{fig3} b,c,d).
+Although the miRNA:target expression ratios in steady-state were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed (compare miRNA distribution values between Figure \autoref{fig1}a and Figure \autoref{fig3}a).
+When expression of Gene2 (G2) increased, expression values of all genes also changed differentially because of contribution of efficiency factors (\autoref{fig3} b,c,d).
 We have considered that in our approach energy and seed type of pairs is significant for binding and targeted region is important for repression.
-But these factors can be handled optionally and differently.
+[][#TODO2]
+
+[#TODO]: <> "although data is visible at Figure2, since it's a small  table,  we might add the table showing expression levels, seed type, binding energy for M1, M2 and genes so that reader can see which miRNA  has low affinity and understand Figure 3 better"
+[#TODO2]: <> "we should compare and contrast Figure 1a and figure 3a, the reader should see how affinity effects distrbution G4:M2, and how efficiency is effected without change in affnity M1:G3"
 
@@
 So it has been observed that the regulatory activities of miRNA efficiencies on the targets are weak after the increase of miRNA target G2.
-When we ran to model with two fold increasing expression value of common target (G4), the changes of other gene expressions have observed more prominent.
-Furthermore, folding of expression of target gene that has strong miRNA repression efficiency also resulted in evident perturbation.
-On the other hand, It was observed that Gene2 was weakly affected from change of Gene4 expression because of its weak interaction factors.This also shows that the perturbation-initiation element in the system is important parameter.
-So, we also have developed an approach that is found perturbation efficiencies of elements in dataset.
+When model is triggered with two fold increase in expression level of common target (G4), the changes of other gene expressions have observed more prominent [][#TODO].
+Furthermore, change in expression level of target gene that has strong miRNA repression efficiency resulted in evident perturbation in network.
+On the other hand, it was observed that Gene2 was weakly affected from change of Gene4 expression because of its weak interaction factors.
+Since perturbation efficiency of each gene is different, we have developed a function which screens each gene in the network for their perturbation efficiencies.
 **(see details in Supplementary File(fig1_2app.Rmd)**.
-When we applied the method on the minimal dataset, while Gene4 is the most efficient element in terms of number of perturbed elements, it have observed that Mir2 is most potent element in the mean of expression change.
-In our approach, the elements behave differently based on type (competing or miRNA).
+When we applied the method on the minimal dataset, Gene4 has been found to be the most efficient element in terms of number of perturbed elements and miRNA2 (M2) has been found to be causing the highest mean in expression changes.
+[][#TODO2]
+
+[#TODO]: <> "didnt we perform upregulation of Gene2, why did we say Gene4 here?"
+[#TODO2]: <> "Removed these sentences because it was somewhat confusing: In our approach, the elements behave differently based on type (competing or miRNA).
 When a competing element was changed, whole elements out of miRNAs are re-regulated.
-But, if a miRNA was perturbed, whole competing elements are re-regulated and miRNA preserves expression at after perturbation.
+But, if a miRNA was perturbed, whole competing elements are re-regulated and miRNA preserves expression at after perturbation."
 
-The defined miRNA’ number is very low in comparison with defined protein coding genes’ number.
-So, this results in that the miRNA amounts in the system must be high or targeted RNAs’ amount must be low for effective miRNA activity.
-We have tried to simulate a real dataset which contains thousands of genes and hundreds of miRNAs.
-Firstly, we have observed the activities in the iterations and determined behaviors of the network after simulation.
-As expected, we have observed that the big dataset, compared to those containing a few genes and miRNAs, exhibited cumbersome behaviors and complexity to gain steady-state **(see details in Supplementary File(TCGA_E9-A1N5_article.Rmd)**.
-We realized that a small change in a system with a large change actually brought many arrangements.
-This can be explained with the competing groups.
-Because the genes are targeted with many common miRNAs, and this results in that the common targets exhibit competing behaviors in different groups.
-On the other hand, the changes of gene expressions were very different.
-While the change in some of genes were highly evident, some have changed slightly.
-So, we have compared two datasets; the tumor gene expression and simulation results.
+[][#TODO] We used our model to simulate a real dataset which contains thousands of genes and hundreds of miRNAs.
+Our  model can successfully simulate perturbations in such large network despite complex behaviors and struggle to reach steady-state **(see details in Supplementary File(TCGA_E9-A1N5_article.Rmd)**.
+Simulations show that change in expression level of single gene has potential to effect whole network, perturbing almost all nodes.
+These observations are in accordance with competing endogenous RNA hypothesis where genes targeted with many common miRNAs subsequently transmit perturbation to neighboring groups.
+
+[#TODO]: <> "removed this section, didnt understand, confusing: So, we have compared two datasets; the tumor gene expression and simulation results.
 When we analyzed comparison of two datasets, we have observed that the results of simulation is not convenient with tumor tissue expression values.
@@
 Of course, it would be more useful to test a dataset that includes gene expression values before and after the regulation of a gene and miRNA expression values at initial conditions.
-When we evaluated the suitability of approach, the approach allows different miRNA targets to be considered together and can work with different regulatory elements due to the network structure.
+When we evaluated the suitability of approach, the approach allows different miRNA targets to be considered together and can work with different regulatory elements due to the network structure."
 
-The network based approaches have been developed in previous studies.
-Figliuzzi et al.
-have tried to explain the ceRNA crosstalk in a network-like minimal interaction structure with concentrations of ceRNA and miRNAs.
-They have pointed out that the larger miRNA number can cause to evident crosstalk between ceRNAs.
-It has been thought that miRNAs and their targets interactions depend on rates of transcription, degradation, binding and unbinding in the network based kinetic model which has been developed by Nitzan et al.
-Nitzan and collaborators have used the high throughput experimental dataset [@helwak_mapping_2013] for miRNA:target pair that occur high free energy and microarray dataset of a miR-92a depletion for expression values.
-They have been demonstrated that distant ceRNAs can interact with each other via indirect links, and the interactions change to depend on distance between ceRNAs, and topological features of network [@nitzan_interactions_2014].
-List et al.
-have developed an approach to detect ceRNA interaction by using the miRNA expression, gene expression and common miRNAs between gene targets [@markus_list_sponge_2017].
-Their approach could be useful to analysis interacted genes through miRNAs.
-When it has been thought that a miRNA can exhibit strong functionality to a target but may not against an other, common miRNA based approach can be inconvenient to understand regulations of ceRNA interactions.
+[#TODO]: <> "removed this section, was confusing: The defined miRNA’ number is very low in comparison with defined protein coding genes’ number.
+So, this results in that the miRNA amounts in the system must be high or targeted RNAs’ amount must be low for effective miRNA activity.
+"
+
+Network based approaches for analyzing miRNA:target interactions have been developed in earlier studies.
+An initial  attempt demonstrated the ceRNA crosstalk in a network-like minimal interaction structure with concentrations of ceRNA and miRNAs [@figliuzzi_micrornas_2013].
+Next, a network based kinetic model integrating miRNA and target rates of transcription, degradation, binding and unbinding was developed [@nitzan_interactions_2014] using high throughput experimental dataset about miR-92a depletion [@helwak_mapping_2013].
+In this study, it was demonstrated that distant ceRNAs can interact with each other via indirect links, and the interactions are effected depending on distance between ceRNAs or topological features of network [@nitzan_interactions_2014].
+More recently, an approach to detect ceRNA interaction by using the miRNA expression, gene expression and common miRNAs between gene targets was developed [@markus_list_sponge_2017] which was effective in analyzing genes through miRNAs.
+Based on the observations that a miRNA can exhibit strong functionality to a target but may not against an other, the authors have concluded that common miRNA based approach may not be suitable for understanding regulations of ceRNA interactions.
+
+[#TODO]: <> "we can talk about BRCA results here: ABCC1 gene, *Multidrug resistance-associated protein 1*, is one of the most significant factor to develop resistance against chemotherapotic agents[@atalay2008role; @lu2015microrna; @atalay2006multidrug]"
 
 In our approach, we have not taken into account transcription, degradation or binding rates of elements in network.
-Because, although it is known as the miRNAs are highly stable, the transcription and degradation rates of miRNAs change to depend on according to cellular conditions [@ruegger_microrna_2012].
-According to our search, there is no dataset that was includes to degradation and transcription rates on specific cellular conditions for each miRNA.
-So, we preferred to accept that degradation and transcription rates of each miRNAs are equal to each other, instead of determining a constant value for all miRNAs in the system (The same applies to mRNA.).
-On the other hand other regulation parameters such as gene-gene interactions and transcription factors are ignored but the network structure can promote to integration of other regulation elements.
-In the future, with development in experimental techniques about features of miRNAs and their targets, more consistent and useful results can be obtained from our approach.
-Additionally, this can provide to contribution to prediction of abnormal regulations and pathways with the studies that will be developed.
-
-(#TODO BRCA results go here) ABCC1 gene, *Multidrug resistance-associated protein 1*, is one of the most significant factor to develop resistance against chemotherapotic agents[@atalay2008role; @lu2015microrna; @atalay2006multidrug]
+Because, although it is known as the miRNAs are highly stable, the transcription and degradation rates of miRNAs change depending on cellular conditions [@ruegger_microrna_2012].
+However, due to lack of datasets including degradation and transcription rates of miRNAs at specific cellular conditions, such factors are neglected in our model.
+Additionally, other regulation parameters such as gene-gene interactions and transcription factors are ignored but the network structure is flexible and can integrate additional regulation elements as they become available.
+In the future, with developments in experimental techniques about miRNAs:target interaction dynamica, more accurate results can be obtained from our model resulting in better understanding and predictions of abnormal regulations and pathways underlying diseases or conditions.
 

### 2019-08-09 article.Rmd
overview

@@
 Our findings emphasized importance of miRNA:target ratios being crucial, as reported by previous studies.
-We have showed that the competing elements which have the same or close expression values may not be affected equally from the perturbation because of repression functionality depended on interaction factors of miRNA target pairs (TOFIX: Hocam etkileşim faktörlerinden dolayı ekspresyonları yakın olsa bile baskılanma aktitesinin aynı sergilenmeyeceğini ifade etmeye çalıştım.
+We have showed that the competing elements which have the same or close expression values may not be affected equally from the perturbation because of repression functionality depended on interaction factors of miRNA target pairs (answer: Hocam etkileşim faktörlerinden dolayı ekspresyonları yakın olsa bile baskılanma aktitesinin aynı sergilenmeyeceğini ifade etmeye çalıştım.
 Bu şekilde anlaşılıyor mu?).
@@
 The model is scalable and can be plugged in with emerging miRNA effectors such as circRNAs.
-The model is available as R package at https://github.com/selcenari/ceRNAnetsim.
-(TODO the link will change)
+The model is available as R package [ceRNAnetsim](https://github.com/selcenari/ceRNAnetsim).
 
@@
 In recent studies, it has been shown that miRNAs can interact with targets via sequences located in regions such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013; @moore_mirnatarget_2015].
-These studies also showed that binding location could indicate functionality of miRNA:target interaction or be effective on abundance of targets(TOFIX abundance etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı?, evet hocam anlaşılmıyorsa değiştirebilirim).
+These studies also showed that binding location could indicate functionality of miRNA:target interaction or be effective on abundance of targets(TOFIX abundance etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı?
+answer: evet hocam anlaşılmıyorsa değiştirebilirim).
 It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is perfectly complementary with targets, seed region at the 5’ end of miRNAs [@bartel_micrornas:_2009; @grimson_microrna_2007].
@@
 On top of all these factors, abundance of miRNAs and targets and miRNA:target ratio in cells predominantly affect efficiency of miRNA:target interaction [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
-
-[#TODO]: <> "there are many et. al and et. al. , with or without dot(.) please make sure the formatting is correct and consistent"
 
@@
 a) In steady state, miRNAs (M: gray triangle) repress targets (G: circle coconut) according to proportion of their targets' expression.
-b) Two fold increase in transcript level of Gene2 (G2) acts as a trigger.
+b) Two fold increase in transcript level of Gene2 (G2) acts as a trigger (shown with red in figure).
 c) Distribution of miRNA1 (M1) changes.
-d) The change at expression of common target effects changes of proportional distribution of miRNA2 (M2).
+d) The change at expression (shown with light red because minor changes are occured on their expression values.) of common target effects changes of proportional distribution of miRNA2 (M2).
 Expression values are rounded to integers for simplicity](Fig1.jpg){#fig1}
@@
 [#TODO]: <> "caption needs more text, every color should be explained"
+[answer]: <> "explained in caption."
 
@@
 
-The repression efficiency of a miRNA on the individual target ($Eff_{gi}$) is calculated according to equation \eqref{eq:1}; where miRNA expression ($C_m$) in local neighborhood is distributed among targets using individual gene expression levels ($C_{gi}$) (#TODO isn't $C_g$ gene expressions in the group is equal to total of $C_{gi}$ in group?).
-For the genes targeted by multiple miRNAs, cooperative activity of miRNAs, $R$, is calculated by summing repression activity of each miRNA (Equation \eqref{eq:2}).
+The repression efficiency of a miRNA on the individual target ($Eff_{gi}$) is calculated according to equation \eqref{eq:1}; where miRNA expression ($C_m$) in local neighborhood is distributed among targets using individual gene expression levels ($C_{gi}$) (#TODO isn't $C_g$ gene expressions in the group is equal to total of $C_{gi}$ in group?
+answer: its my false.
+now its corect.).
+For the genes targeted by multiple miRNAs, cooperative activity of miRNAs on a target gene, $R$, is calculated by summing repression activity of each miRNA (Equation \eqref{eq:2}).
 
 \begin{equation} 
-    Eff_{gi}= C_m \times C_{gi}/\sum_{1}^{i} C_g \tag{1}\label{eq:1}
+    Eff_{gi}= C_m \times C_{gi}/\sum_{1}^{i} C_{gi} \tag{1}\label{eq:1}
 \end{equation}
@@
 Factors effecting binding determine interaction between miRNA and target and they alter amount of miRNA sequestered to target.
-Factors effecting efficiency determine proportion of target mRNA degraded by bound miRNA (#TODO fix this sentence).
+Degradation efficiency factors assign amount of degradated target amount in miRNA:target pairs.(#TODO fix this sentence: answer:).
 In other words, binding factors exert their influence before or during binding, efficiency factors exert their influence after binding.
@@
 We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from ...
-(#TODO name of database and its citation).
+(#TODO name of database and its citation)->suplementary datasets of [@helwak_mapping_2013; @moore_mirnatarget_2015] .
 Detailed description of network construction and its code is available in Supplementary data (#TODO burada link nasıl olacak, word dosyası veya PDF dosyası ismi mi vermemiz mi  gerekiyor, yoksa sadece Sup Data denilmesi yeterli mi?
@@
 Therefore, primary expression level change in gene (G2) causes changes in other group of genes (G5 and G6) even though original trigger gene (G2) and genes in other group are not targeted by common miRNA.
-In addition, as shown by ceRNA hypothesis model of Ala et al., after the increase of gene expression level of G2, the miRNA that is found in the same group (M1) tended to be more (#TODO isn't it less?) repressive on its remaining targets (G1, G3 and G4).
+In addition, as shown by ceRNA hypothesis model of Ala et al., after the increase of gene expression level of G2, the miRNA that is found in the same group (M1) tended to be less repressive on its remaining targets (G1, G3 and G4).
 It's important to note that the changes in gene expression levels will have more pronounced effect if miRNA:target ratio is high, i.e., more miRNA available per target, which was reported in previous findings [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
@@
 
-### Multifactorial networks
+### Competing Endogenous RNA:Target networks based on interaction factors
 
-[#TODO]: <> "title should be more informative"
+[#TODO]: <> "title should be more informative"-> answer.
 
 Earlier studies reported that miRNA regulatory interactions are affected by different parameters.
-For example, Xu et.
-al have investigated the importance of seed pairing type between miRNAs and their targets and target site location by using proteomics dataset [@xu_characterization_2014].
+For example, Xu et al.
+have investigated the importance of seed pairing type between miRNAs and their targets and target site location by using proteomics dataset [@xu_characterization_2014].
 They have proposed that the features of binding between miRNA and target can be critical for miRNA efficiency.
@@
 [][#TODO2]
+
+answer: knitr::kable(minsamp): figure ü fig1_2 dosyasında chunk2ye attım. Burada bir hata veriyor. Düzenleyemedim.
 
@@
 [#TODO]: <> "didnt we perform upregulation of Gene2, why did we say Gene4 here?" 
-[#TODO2]: <> "Removed these sentences because it was somewhat confusing: In our approach, the elements behave differently based on type (competing or miRNA).
-When a competing element was changed, whole elements out of miRNAs are re-regulated.
-But, if a miRNA was perturbed, whole competing elements are re-regulated and miRNA preserves expression at after perturbation."
+[answer]: <> "Because, in fact, We want to show that triggering element is significant in the same dataset, in the same conditions. Maybe I must add a conjuction here."
 

### 2019-08-09 article.tex
overview

@@
   \posttitle{\par}
-    \author{Selcen Ari\footnote{Yildiz Technical University Department of
-  Bioengineering} \\ Alper Yilmaz\footnote{Yildiz Technical University Department of
-  Bioengineering}}
+    \author{Selcen Ari\footnote{Department of Bioengineering, Yildiz Technical
+  University, Istanbul, Turkey} \\ Alper Yilmaz\footnote{Department of Bioengineering, Yildiz Technical
+  University, Istanbul, Turkey}}
     \preauthor{\centering\large\emph}
@@
   \postdate{\par}
-    \date{28 Mayıs 2019}
+    \date{09 Ağustos 2019}
 
@@
 model is scalable and can be plugged in with emerging miRNA effectors
-such as circRNAs. The model is available as R package at
-\url{https://github.com/selcenari/ceRNAnetsim}. (TODO the link will
-change)
+such as circRNAs. The model is available as R package
+\href{https://github.com/selcenari/ceRNAnetsim}{ceRNAnetsim}.
 
@@
 targets of the same miRNAs have been found with help of correlation of
-gene expression changes in recent algoritm (Markus List 2017). List et
+gene expression changes in recent algorithm (Markus List 2017). List et
 al.~have specified that their approach can be useful for ceRNA studies
@@
 
-\hypertarget{mirnatarget-ratio-theoretical-action-in-network-based-cerna-model}{%
-\subsubsection{miRNA:target ratio theoretical action in network based
-ceRNA
-model}\label{mirnatarget-ratio-theoretical-action-in-network-based-cerna-model}}
+\hypertarget{construction-of-mirnatarget-network}{%
+\subsubsection{Construction of miRNA:target
+network}\label{construction-of-mirnatarget-network}}
 
-We developed a novel network based approach that can use to simulation
-of gene regulation after any perturbation at miRNA:target interactions.
-If a perturbation or change induces on graph, whole interacted nodes
-vary because of changes in miRNA:target ratios. The system was assumed
-as steady state that miRNAs affect according to the ratio of the
-individual target to all targets (\autoref{fig1}a, expression values
-shown as integer for comprehensible sampling). When expression of gene 2
-(G2) increased to two fold (\autoref{fig1}b), distribution of miRNA (M1)
-expression changes (\autoref{fig1}c). Consequently, expression of common
-target (G4) of miRNA1 and miRNA2 also varies. The change at expression
-of common target triggers to changes of proportional distribution of
-miRNA2 (\autoref{fig1}d).
+miRNA and target pairs per line should be provided as edge list to
+construct the network. At each line minimum required information is
+expression levels of miRNA and the target. If available, additional data
+about factors effecting binding or efficiency of miRNA can be provided
+as separate columns. After construction of the network, amount of miRNA
+per target is calculated and kept as edge data. Simply, a target will
+sequester miRNA proportional to its ratio amount among other targets. If
+additional criteria effecting the binding of miRNA to its target is
+provided, distribution of miRNA will be calculated accordingly. Target
+can be mRNA or any other ceRNA (circRNA, ncRNA, etc.) thus, throughout
+the manuscript terms target, gene and ceRNA are used interchangeably.
+
+\hypertarget{triggering-perturbation-and-subsequent-calculations}{%
+\subsubsection{Triggering perturbation and subsequent
+calculations}\label{triggering-perturbation-and-subsequent-calculations}}
+
+Initially, the network is assumed in steady-state (\autoref{fig1}a)
+condition and needs a trigger for initiating calculations. The trigger
+can be a change in expression level of one or more genes
+(\autoref{fig1}b). After a trigger, the network undergoes iterative
+cycle of calculations at each of which distribution of miRNA in local
+neighborhood is recalculated (\autoref{fig1}c). Based on new miRNA
+distribution, expression level of each node (i.e.~ceRNA) is updated. Due
+to nodes which are targeted by multiple miRNAs, the change in one
+neighborhood spreads to other neighborhoods (\autoref{fig1}d),
+consequently have potential to effect whole network due to ``ripple
+effect''.
 
@@
 In steady state, miRNAs (M: gray triangle) repress targets (G: circle
-coconut) according to proportion of their targets' expression. b-c-d)
-MiRNA expression distributions to targets vary when expression of one of
-the targets perturbed.}\label{fig1}
+coconut) according to proportion of their targets' expression. b) Two
+fold increase in transcript level of Gene2 (G2) acts as a trigger. c)
+Distribution of miRNA1 (M1) changes. d) The change at expression of
+common target effects changes of proportional distribution of miRNA2
+(M2). Expression values are rounded to integers for
+simplicity}\label{fig1}
 }
@@
 
-The model was developed by making some assumptions. Transcription rates
-and degradation rates of miRNAs were assumed as stationary and equal.
-Targets also have stable transcription and degradation rates, at the
-same time they are equal. The factors that could be affect the
-miRNA:target interaction are disregarded in this approach. It is
-considered that miRNAs degrades or inhibits them via any mechanisms
-after binding to target while the miRNAs were used reversibly. In
-initial conditions, gene expression amounts in system have been
-considered after repression activity of miRNAs. If a change occurs in
-the system, variation on miRNA proportional distribution are reflected
-as change of repression activity. \begin{equation} 
-    Eff_{gi}= C_m * C_{gi}/\sum_{1}^{i} C_g \tag{1}\label{eq:1}
+During calculations, following assumptions were adopted; 1)
+Transcription and degradation rates of miRNAs are steady and equal. 2)
+All available miRNAs are recycled as in miRNA:ceRNA binding, target is
+degraded and miRNA is unaffected. 3) ceRNA targets also have stable
+transcription and degradation rates and these rates are equal.
+
+The repression efficiency of a miRNA on the individual target
+(\(Eff_{gi}\)) is calculated according to equation \eqref{eq:1}; where
+miRNA expression (\(C_m\)) in local neighborhood is distributed among
+targets using individual gene expression levels (\(C_{gi}\)) (\#TODO
+isn't \(C_g\) gene expressions in the group is equal to total of
+\(C_{gi}\) in group?). For the genes targeted by multiple miRNAs,
+cooperative activity of miRNAs, \(R\), is calculated by summing
+repression activity of each miRNA (Equation \eqref{eq:2}).
+
+\begin{equation} 
+    Eff_{gi}= C_m \times C_{gi}/\sum_{1}^{i} C_g \tag{1}\label{eq:1}
 \end{equation} \begin{equation}
-   R_{gi}= Eff_{i1}+ Eff_{i2} ... \tag{2}\label{eq:2}
+   R_{gi}= Eff_{i1} + Eff_{i2} ... \tag{2}\label{eq:2}
 \end{equation}
 
-The efficiency of a miRNA on the individual target (\(Eff_{gi}\)) is
-calculated according to equation \eqref{eq:1}; \(C_m\), miRNA expression
-in the group; \(C_{gi}\), symbolize individual gene expression; \(C_g\)
-gene expressions in the group (Groups are determined depending on
-miRNAs. Targets that are interacted with the same miRNA find in a
-group.). In this approach, because a gene can be found in different
-groups, the cooperative activity of microRNAs also can be calculated.
-Total repression activity of different miRNAs on the common target,
-\(R\), is calculated based on equation \eqref{eq:2}; sum of efficiency
-of different miRNAs on the target.
+\hypertarget{multifactorial-calculations-in-mirnatarget-network}{%
+\subsubsection{Multifactorial calculations in miRNA:target
+network}\label{multifactorial-calculations-in-mirnatarget-network}}
+
+Interactions between miRNAs and their targets can be affected from
+various factors. So, our model integrates multiple factors when
+calculating overall miRNA activity. We classified factors into two
+categories. Factors effecting binding determine interaction between
+miRNA and target and they alter amount of miRNA sequestered to target.
+Factors effecting efficiency determine proportion of target mRNA
+degraded by bound miRNA (\#TODO fix this sentence). In other words,
+binding factors exert their influence before or during binding,
+efficiency factors exert their influence after binding. In the
+literature, binding free energy (Cao and Chen 2012 ; Helwak et al. 2013)
+and seed type (Werfel et al. 2017) in miRNA:target interactions are
+described as factors effecting binding affinity. Efficiency factors
+determine how many of miRNA:target complexes will result in inhibition
+and binding region on target drastically effect miRNA degradation
+efficiency (Hausser et al. 2013; Helwak et al. 2013). Both binding and
+efficiency factors are normalized to their maximum values and scaled to
+{[}0,1{]} interval. The normalized values of factors take into account
+to determine binding activity and miRNA efficiency on targets
+(\autoref{fig2}). Binding affinities (activity, \(Eff\)) of miRNAs on
+each individual gene are calculated as shown in equation \eqref{eq:4};
+\(C_m\), miRNA expression in the group; \(C_g\), Gene expression;
+\(C_{gi}\), individual gene expression; gi, individual gene; g, whole
+genes in a group (\autoref{fig2}c). \begin{equation}
+Eff_{gi}= C_m \times E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}/(\sum_{1}^{i} E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}) \tag{4}\label{eq:4}
+\end{equation} \begin{equation}
+Eff_{gi}= Eff_{gi}*RE^\prime_{gi} \tag{5}\label{eq:5}
+\end{equation}
 
@@
 efficiency. \(G\), Gene; \(M\), miRNA; \(STE\), seed type effect;
-\(RE\), Region Effect; \(E\), Energy; \(STE’\), normalized values of
-seed type efficiency coefficient; \(RE’\), normalized values of region
-efficiency coefficient; \(E’\), normalized values of energy
+\(RE\), Region Effect; \(E\), Energy; \(STE^\prime\), normalized values
+of seed type efficiency coefficient; \(RE^\prime\), normalized values of
+region efficiency coefficient; \(E^\prime\), normalized values of energy
 coefficient.}\label{fig2}
@@
 
-\hypertarget{mirna-action-in-network-based-multifactorial-cerna-model}{%
-\subsubsection{miRNA action in network based multifactorial ceRNA
-model}\label{mirna-action-in-network-based-multifactorial-cerna-model}}
+\iffalse
 
-Interactions between miRNAs and their targets can be affected from
-different parameters. For this reason, we improved model with factors
-that can affect to miRNA activity. In our approach, the parameters are
-classified as binding and efficiency factors and all of these are
-evaluated in their groups. Binding factors determine interaction between
-miRNA and target. The efficiency factors decide functionality of
-binding. In the literature, binding free energy (Cao and Chen 2012 ;
-Helwak et al. 2013) and seed type (Werfel et al. 2017) in miRNA:target
-interactions are represented as binding affinity. Efficiency factors
-determine how many of microRNA:target complexes will result in
-inhibition. For example, binding region on the target shows whether the
-target can be inhibited (functionality)(Hausser et al. 2013; Helwak et
-al. 2013). Both of binding and efficiency factors are evaluated in the
-groups and normalized according to equation \eqref{eq:3}; each factor is
-normalized based on maximum variables that found in their groups.
-\begin{equation}
-F`= F/F_{max} \tag{3}\label{eq:3}
-\end{equation}
-
-The normalized values of factors take into account to determine binding
-activity and miRNA efficiency on targets (\autoref{fig2}). Binding
-affinities (activity, \(Eff\)) of miRNAs on each individual gene are
-calculated as shown in equation \eqref{eq:4}; \(C_m\), miRNA expression
-in the group; \(C_g\), Gene expression; \(C_{gi}\), individual gene
-expression; gi, individual gene; g, whole genes in a group
-(\autoref{fig2}c). \begin{equation}
-Eff_{gi}= C_m * E`_{gi}*STE`_{gi}*C_{gi}/(\sum_{1}^{i} E`_{gi}*STE`_{gi}*C_{gi}) \tag{4}\label{eq:4}
-\end{equation} \begin{equation}
-Eff_{gi}= Eff_{gi}*RE`_{gi} \tag{5}\label{eq:5}
-\end{equation}
+\%\% (\#TODO the equation is not considering binding factors, right?. We
+should explain the calculation made in simulate function here, in simple
+sentences. the function looks like this: \%\%
+input\_graph\%\textgreater{}\% \%\%
+tidygraph::activate(edges)\%\textgreater{}\% \%\%
+group\_by(to)\%\textgreater{}\% \%\% mutate(mirna\_count\_per\_comp =
+mirna\_count\_current\emph{comp\_count\_current}afff\_factor/sum(comp\_count\_current\emph{afff\_factor),
+mirna\_count\_per\_comp = ifelse(is.na(mirna\_count\_per\_comp), 0,
+mirna\_count\_per\_comp))\%\textgreater{}\% \%\%
+ungroup()\%\textgreater{}\% \%\% mutate(effect\_pre = effect\_current,
+effect\_current =
+mirna\_count\_per\_comp}degg\_factor)\%\textgreater{}\% \%\%
+group\_by(from)\%\textgreater{}\% \%\% mutate(comp\_count\_pre =
+ifelse(comp\_count\_current\textless{} 0, 1, comp\_count\_current),
+comp\_count\_current = comp\_count\_pre -
+(sum(effect\_current)-sum(effect\_pre)), comp\_count\_current =
+ifelse(comp\_count\_current\textless{} 0, 1,
+comp\_count\_current))\%\textgreater{}\% \%\%
+ungroup()\%\textgreater{}\% \%\% mutate(comp\_count\_list =
+pmap(list(comp\_count\_list, comp\_count\_current), c), effect\_list =
+pmap(list(effect\_list, effect\_current), c), mirna\_count\_list =
+pmap(list(mirna\_count\_list, mirna\_count\_current), c)) \%\% ) \fi
 
@@
 repression efficiency of miRNA is calculated according to equation
-\eqref{eq:5} (\autoref{fig2}d); \(RE’_{gi}\), normalized values of
+\eqref{eq:5} (\autoref{fig2}d); \(RE^\prime_{gi}\), normalized values of
 region efficiency coefficient between miRNA and gene. The cooperative
@@
 
-\hypertarget{application-of-model-in-a-breast-cancer-patient-dataset}{%
-\subsubsection{Application of model in a breast cancer patient
-dataset}\label{application-of-model-in-a-breast-cancer-patient-dataset}}
+\hypertarget{breast-cancer-patient-dataset}{%
+\subsubsection{Breast cancer patient
+dataset}\label{breast-cancer-patient-dataset}}
 
-To determine suitability of method, we have tried to apply model in a
-real dataset. Firstly, we obtained the miRNA and gene expression
-datasets of the patient's tumor and normal tissue. We downloaded the
-high-throughput experimental datasets which are provided miRNA:gene
-target pairs with interaction factors by
-\textbf{moore\_mirnatarget\_2015} and (Helwak et al. 2013). We have
-combined miRNA and gene expression datasets via miRNA:target gene
-dataset \textbf{(Whole steps can be found in Supplementary
-data(TCGA\_E9-A1N5\_article.Rmd)}). We selected a gene which is over
-expressed in tumor tissue with compared to normal tissue. It is pointed
-out that ABCC1 gene, \emph{Multidrug resistance-associated protein 1},
-is one of the most significant factor to develop resistance against
-chemotherapotic agents(Atalay, Demirkazik, and Gunduz 2008; Lu et al.
-2015; Atalay et al. 2006). So, we selected ABCC1 gene for simulation of
-integrated dataset. We have detected the iteration that is suitable for
-ABCC1 gene simulation. After simulation of system via ABCC1 gene with
-over expression, we have compared simulation results and tumor tissue
-expression levels.
+We have applied our model in a real dataset for which experimental
+measurements of various factors were available. Expression levels of
+miRNA and genes in tumor and normal tissue of single patient are
+retrieved from TCGA(\#TODO cite). High-throughput experimental datasets
+which are provided miRNA:gene target pairs with interaction factors
+(Helwak et al. 2013; Moore et al. 2015). We have combined miRNA and gene
+expression datasets via miRNA:target gene dataset retrieved from
+\ldots{} (\#TODO name of database and its
+citation)-\textgreater{}suplementary datasets of (Helwak et al. 2013;
+Moore et al. 2015) . Detailed description of network construction and
+its code is available in Supplementary data (\#TODO burada link nasıl
+olacak, word dosyası veya PDF dosyası ismi mi vermemiz mi gerekiyor,
+yoksa sadece Sup Data denilmesi yeterli mi? ilgili dosya:
+TCGA\_E9-A1N5\_article.Rmd). ABCC1 gene, over-expressed in tumor tissue,
+was selected as trigger for simulation of integrated dataset. After
+simulation of network, we have compared simulation results and tumor
+tissue expression levels.
 
@@
 
-We have developed a model approach and workflow for competitive ceRNA
-regulation in this study. The basic mode of miRNA repression activity
-has been based on miRNA and target abundance in various researches(Arvey
-et al. 2010; Denzler et al. 2014). For this reason, we performed first
-application as considering only abundance of miRNAs and targets
-(\autoref{fig1}). After the increasing in a gene expression value,
-expression values of other genes also changed slightly in comparison
-with alteration in perturbation started gene node. Different studies
-have shown that the other gene expression values also rise differently,
-if a gene abundance increases in ceRNA system (Lai, Wolkenhauer, and
-Vera 2016; Salmena et al. 2011; Tay, Rinn, and Pandolfi 2014). It was
-observed that primary neighborhoods of the trigger gene change with the
-same ratio, but cooperative efficiency of miRNAs on common target causes
-differently change in common target (\autoref{fig1}d). The common target
-in system has displayed like a trigger for the other group and induced
-changes of expression values of genes on the other group. Therefore,
-effect of primary change of the gene (G2) on genes of the other group
-was observed slightly (\autoref{fig1}d). In addition, like shown by
-ceRNA hypothesis model of Ala et al., after the increasing of a gene
-expression (G2), the miRNA that is found in the same group (M1) tended
-to be more repressive on this target. This has caused to decrease in
-increased expression value of primer triggering gene. When the
-expression of a gene (G2) was increased to two fold, we observed that
-the expression changes were more evident. So, we have considered that
-high miRNA:target ratio results in stronger miRNA repression activity.
-This results also are coherent with previous reports that were offered
-by Arvey et al.~and Denzler et al.(Arvey et al. 2010; Bosson, Zamudio,
-and Sharp 2014; Denzler et al. 2014). When the simulation was run in
-sample dataset that includes lower target abundance, the first response
-of primary neighborhood of the same group was determined as similar with
-in system that have high target abundance. But regulations of ceRNAs
-were observed differences in the following steps of simulation. In
-overall state in the end of simulation, we have established that all
-regulations more prominent when compared with low miRNA:target ratio
-dataset. The model has appeared the importance of mRNA:target ratio in
-regulation of competing RNAs.
+\hypertarget{networks-with-single-factor}{%
+\subsubsection{Networks with single
+factor}\label{networks-with-single-factor}}
+
+We have developed a network-based approach to assess effects of
+expression level changes in competitive ceRNA regulation. The basic mode
+of miRNA repression activity has been based on miRNA and target
+abundance in various researches(Arvey et al. 2010; Denzler et al. 2014).
+Our approach can effortlessly calculate effects of expression changes
+when abundance levels of miRNAs and targets is only available factor. In
+sample network given in Figure (\autoref{fig1}), after an increase in
+expression level of a gene (G2), expression values of other genes also
+changed due to redistribution of miRNA among its targets. Previous
+studies have shown that if a gene abundance increases in ceRNA system,
+expression levels of genes targeted by shared miRNA are also affected
+(Lai, Wolkenhauer, and Vera 2016; Salmena et al. 2011; Tay, Rinn, and
+Pandolfi 2014). It was observed that primary neighborhoods of the
+trigger gene change with the same ratio, but cooperative efficiency of
+miRNAs on common target causes differently change in common target
+(\#TODO last sentence needs some clarification, after ``but
+cooperative\ldots{}'') (\autoref{fig1}d). Genes targeted by multiple
+miRNAs act as a trigger for adjacent local neighborhood of targeting
+miRNAs, causing changes in expression levels of genes outside the local
+neighborhood of original trigger gene. Therefore, primary expression
+level change in gene (G2) causes changes in other group of genes (G5 and
+G6) even though original trigger gene (G2) and genes in other group are
+not targeted by common miRNA. In addition, as shown by ceRNA hypothesis
+model of Ala et al., after the increase of gene expression level of G2,
+the miRNA that is found in the same group (M1) tended to be less
+repressive on its remaining targets (G1, G3 and G4). It's important to
+note that the changes in gene expression levels will have more
+pronounced effect if miRNA:target ratio is high, i.e., more miRNA
+available per target, which was reported in previous findings (Arvey et
+al. 2010; Bosson, Zamudio, and Sharp 2014; Denzler et al. 2014).
+\href{}{}
 
@@
 
-It has been reported in several studies that miRNA regulatory
-interactions also is affected different parameters. For example, Xu et.
-al have investigated the importance of seed pairing type between miRNAs
-and their targets and target site location by using proteomics dataset
-(Xu, Wang, and Liu 2014). They have proposed that the features of
-binding between miRNA and target can be critical efficiency of miRNA and
-can be used to find effective miRNA molecules. In addition, binding
-energy between miRNA and targets also could be determinant for miRNA
-efficiency. It has been reported by Breda et. al.~that strength of
-miRNA:target interactions is depended on binding energy of complexes
-(Breda et al. 2015). As similar to these, Bosson et al.~have associated
-affinity and seed pairing of miRNA target pairs and suggested that
-number of canonical base pairing is correlated with affinity. For these
-reasons, we have improved the model with optional interaction parameters
-that could be useful for understanding of miRNA repression activity. We
-have explained the model in the same sample dataset with using different
-factors and simulated via the same change (\autoref{fig3}). When the
-factors were taken into account in the system, miRNA efficiencies varied
-as shown in \autoref{fig3}a. Although the miRNA target ratios in initial
-states were same in comparison with the sample dataset without factors,
-efficiency of binding and repression have changed. When expression of a
-gene (G2) increased, expression values of all genes also change
-differentially because of contribution of efficiency factors
+\hypertarget{competing-endogenous-rnatarget-networks-based-on-interaction-factors}{%
+\subsubsection{Competing Endogenous RNA:Target networks based on
+interaction
+factors}\label{competing-endogenous-rnatarget-networks-based-on-interaction-factors}}
+
+\href{}{\#TODO}: \textless{}\textgreater{} ``title should be more
+informative''-\textgreater{}
+
+Earlier studies reported that miRNA regulatory interactions are affected
+by different parameters. For example, Xu et al.~have investigated the
+importance of seed pairing type between miRNAs and their targets and
+target site location by using proteomics dataset (Xu, Wang, and Liu
+2014). They have proposed that the features of binding between miRNA and
+target can be critical for miRNA efficiency. In addition, binding energy
+between miRNA and targets is a significant determinant for miRNA
+efficiency and it has been reported that strength of miRNA:target
+interactions is depended on binding energy of complexes (Breda et al.
+2015). Similarly, another study revealed that affinity is correlated
+with seed pairing of miRNA:target pairs and suggested that affinity is
+correlated with number of canonical seed base pairing (Bosson, Zamudio,
+and Sharp 2014). Therefore, we have integrated aforementioned
+interaction parameters that could be useful for predicting miRNA
+repression activity more accurately. The sample dataset used in Figure
+\autoref{fig1} is recalculated with additional factors in effect
+\href{}{} with same trigger, two fold increase in Gene2
+(\autoref{fig3}). When the factors were taken into account in the
+system, miRNA efficiencies varied as shown in \autoref{fig3}a. Although
+the miRNA:target expression ratios in steady-state were same in
+comparison with the sample dataset without factors, efficiency of
+binding and repression have changed (compare miRNA distribution values
+between Figure \autoref{fig1}a and Figure \autoref{fig3}a). When
+expression of Gene2 (G2) increased, expression values of all genes also
+changed differentially because of contribution of efficiency factors
 (\autoref{fig3} b,c,d). We have considered that in our approach energy
 and seed type of pairs is significant for binding and targeted region is
-important for repression. But these factors can be handled optionally
-and differently.
+important for repression. \href{}{}
+
+knitr::kable(minsamp)
 
@@
 miRNA efficiencies on the targets are weak after the increase of miRNA
-target G2. When we ran to model with two fold increasing expression
-value of common target (G4), the changes of other gene expressions have
-observed more prominent. Furthermore, folding of expression of target
-gene that has strong miRNA repression efficiency also resulted in
-evident perturbation. On the other hand, It was observed that Gene2 was
-weakly affected from change of Gene4 expression because of its weak
-interaction factors.This also shows that the perturbation-initiation
-element in the system is important parameter. So, we also have developed
-an approach that is found perturbation efficiencies of elements in
-dataset. \textbf{(see details in Supplementary File(fig1\_2app.Rmd)}.
-When we applied the method on the minimal dataset, while Gene4 is the
-most efficient element in terms of number of perturbed elements, it have
-observed that Mir2 is most potent element in the mean of expression
-change. In our approach, the elements behave differently based on type
-(competing or miRNA). When a competing element was changed, whole
-elements out of miRNAs are re-regulated. But, if a miRNA was perturbed,
-whole competing elements are re-regulated and miRNA preserves expression
-at after perturbation.
+target G2. When model is triggered with two fold increase in expression
+level of common target (G4), the changes of other gene expressions have
+observed more prominent \href{}{}. Furthermore, change in expression
+level of target gene that has strong miRNA repression efficiency
+resulted in evident perturbation in network. On the other hand, it was
+observed that Gene2 was weakly affected from change of Gene4 expression
+because of its weak interaction factors. Since perturbation efficiency
+of each gene is different, we have developed a function which screens
+each gene in the network for their perturbation efficiencies.
+\textbf{(see details in Supplementary File(fig1\_2app.Rmd)}. When we
+applied the method on the minimal dataset, Gene4 has been found to be
+the most efficient element in terms of number of perturbed elements and
+miRNA2 (M2) has been found to be causing the highest mean in expression
+changes. \href{}{}
 
-The defined miRNA' number is very low in comparison with defined protein
-coding genes' number. So, this results in that the miRNA amounts in the
-system must be high or targeted RNAs' amount must be low for effective
-miRNA activity. We have tried to simulate a real dataset which contains
-thousands of genes and hundreds of miRNAs. Firstly, we have observed the
-activities in the iterations and determined behaviors of the network
-after simulation. As expected, we have observed that the big dataset,
-compared to those containing a few genes and miRNAs, exhibited
-cumbersome behaviors and complexity to gain steady-state \textbf{(see
-details in Supplementary File(TCGA\_E9-A1N5\_article.Rmd)}. We realized
-that a small change in a system with a large change actually brought
-many arrangements. This can be explained with the competing groups.
-Because the genes are targeted with many common miRNAs, and this results
-in that the common targets exhibit competing behaviors in different
-groups. On the other hand, the changes of gene expressions were very
-different. While the change in some of genes were highly evident, some
-have changed slightly. So, we have compared two datasets; the tumor gene
-expression and simulation results. When we analyzed comparison of two
-datasets, we have observed that the results of simulation is not
-convenient with tumor tissue expression values. The simulation results
-were not expected to be concordant with tumor tissue expression values
-because of a single gene could not be responsible for gene regulation in
-all tissue. Besides, the all other factors such as up/down regulation of
-miRNAs or other genes were ignored. Of course, it would be more useful
-to test a dataset that includes gene expression values before and after
-the regulation of a gene and miRNA expression values at initial
-conditions. When we evaluated the suitability of approach, the approach
-allows different miRNA targets to be considered together and can work
-with different regulatory elements due to the network structure.
+\href{}{} We used our model to simulate a real dataset which contains
+thousands of genes and hundreds of miRNAs. Our model can successfully
+simulate perturbations in such large network despite complex behaviors
+and struggle to reach steady-state \textbf{(see details in Supplementary
+File(TCGA\_E9-A1N5\_article.Rmd)}. Simulations show that change in
+expression level of single gene has potential to effect whole network,
+perturbing almost all nodes. These observations are in accordance with
+competing endogenous RNA hypothesis where genes targeted with many
+common miRNAs subsequently transmit perturbation to neighboring groups.
 
-The network based approaches have been developed in previous studies.
-Figliuzzi et al.~have tried to explain the ceRNA crosstalk in a
-network-like minimal interaction structure with concentrations of ceRNA
-and miRNAs. They have pointed out that the larger miRNA number can cause
-to evident crosstalk between ceRNAs. It has been thought that miRNAs and
-their targets interactions depend on rates of transcription,
-degradation, binding and unbinding in the network based kinetic model
-which has been developed by Nitzan et al.~Nitzan and collaborators have
-used the high throughput experimental dataset (Helwak et al. 2013) for
-miRNA:target pair that occur high free energy and microarray dataset of
-a miR-92a depletion for expression values. They have been demonstrated
-that distant ceRNAs can interact with each other via indirect links, and
-the interactions change to depend on distance between ceRNAs, and
-topological features of network (Nitzan et al. 2014). List et al.~have
-developed an approach to detect ceRNA interaction by using the miRNA
-expression, gene expression and common miRNAs between gene targets
-(Markus List 2017). Their approach could be useful to analysis
-interacted genes through miRNAs. When it has been thought that a miRNA
-can exhibit strong functionality to a target but may not against an
-other, common miRNA based approach can be inconvenient to understand
-regulations of ceRNA interactions.
+Network based approaches for analyzing miRNA:target interactions have
+been developed in earlier studies. An initial attempt demonstrated the
+ceRNA crosstalk in a network-like minimal interaction structure with
+concentrations of ceRNA and miRNAs (Figliuzzi, Marinari, and De Martino
+2013). Next, a network based kinetic model integrating miRNA and target
+rates of transcription, degradation, binding and unbinding was developed
+(Nitzan et al. 2014) using high throughput experimental dataset about
+miR-92a depletion (Helwak et al. 2013). In this study, it was
+demonstrated that distant ceRNAs can interact with each other via
+indirect links, and the interactions are effected depending on distance
+between ceRNAs or topological features of network (Nitzan et al. 2014).
+More recently, an approach to detect ceRNA interaction by using the
+miRNA expression, gene expression and common miRNAs between gene targets
+was developed (Markus List 2017) which was effective in analyzing genes
+through miRNAs. Based on the observations that a miRNA can exhibit
+strong functionality to a target but may not against an other, the
+authors have concluded that common miRNA based approach may not be
+suitable for understanding regulations of ceRNA interactions.
 
@@
 it is known as the miRNAs are highly stable, the transcription and
-degradation rates of miRNAs change to depend on according to cellular
-conditions (Rüegger and Großhans 2012). According to our search, there
-is no dataset that was includes to degradation and transcription rates
-on specific cellular conditions for each miRNA. So, we preferred to
-accept that degradation and transcription rates of each miRNAs are equal
-to each other, instead of determining a constant value for all miRNAs in
-the system (The same applies to mRNA.). On the other hand other
+degradation rates of miRNAs change depending on cellular conditions
+(Rüegger and Großhans 2012). However, due to lack of datasets including
+degradation and transcription rates of miRNAs at specific cellular
+conditions, such factors are neglected in our model. Additionally, other
 regulation parameters such as gene-gene interactions and transcription
-factors are ignored but the network structure can promote to integration
-of other regulation elements. In the future, with development in
-experimental techniques about features of miRNAs and their targets, more
-consistent and useful results can be obtained from our approach.
-Additionally, this can provide to contribution to prediction of abnormal
-regulations and pathways with the studies that will be developed.
+factors are ignored but the network structure is flexible and can
+integrate additional regulation elements as they become available. In
+the future, with developments in experimental techniques about
+miRNAs:target interaction dynamica, more accurate results can be
+obtained from our model resulting in better understanding and
+predictions of abnormal regulations and pathways underlying diseases or
+conditions.
 
@@
 \url{https://doi.org/10.1038/msb.2010.24}.
-
-\leavevmode\hypertarget{ref-atalay2006multidrug}{}%
-Atalay, Can, Ismet Deliloglu Gurhan, Cigdem Irkkan, and Ufuk Gunduz.
-2006. ``Multidrug Resistance in Locally Advanced Breast Cancer.''
-\emph{Tumor Biology} 27 (6): 309--18.
-
-\leavevmode\hypertarget{ref-atalay2008role}{}%
-Atalay, C, A Demirkazik, and U Gunduz. 2008. ``Role of Abcb1 and Abcc1
-Gene Induction on Survival in Locally Advanced Breast Cancer.''
-\emph{Journal of Chemotherapy} 20 (6): 734--39.
 
@@
 
-\leavevmode\hypertarget{ref-lu2015microrna}{}%
-Lu, Lin, Fang Ju, Hui Zhao, and Xuezhen Ma. 2015. ``MicroRNA-134
-Modulates Resistance to Doxorubicin in Human Breast Cancer Cells by
-Downregulating Abcc1.'' \emph{Biotechnology Letters} 37 (12): 2387--94.
-
 \leavevmode\hypertarget{ref-markus_list_sponge_2017}{}%

### 2019-08-09 fig1_2app.Rmd
overview

@@
 ```
+
+|competing| miRNA | Competing_expression | miRNA_expression | seed_type | region | energy
+|-----------|------------|----------|----------------|---------|-----------|----------|
+| `Gene1`   | `Mir1` | 10000 | 1000 | 0.43 | 0.30 | -20 |
+| `Gene2`   | `Mir1` | 10000 | 1000 | 0.43 | 0.01 | -15 |
+| `Gene3`   | `Mir1` | 5000  | 1000 | 0.32 | 0.40 | -14 |
+| `Gene4`   | `Mir1` | 10000 | 1000 | 0.23 | 0.50 | -10 |
+| `Gene4`   | `Mir2` | 10000 | 2000 | 0.35 | 0.90 | -12 |
+| `Gene5`   | `Mir2` | 5000  | 2000 | 0.05 | 0.40 | -11 |
+| `Gene6`   | `Mir2` | 10000 | 2000 | 0.01 | 0.80 | -25 |
 

### 2019-08-22 TCGA_E9-A1N5_article.Rmd
article files

@@
 ---
-title: "article_supplementary"
+title: "Supplementary file 2"
 author: "Selcen Ari"
@@
 # Install the packages which is necessary to obtain and process datasets.
-#BiocManager::install("SummarizedExperiment", version = "3.8")
+#install.packages("BiocManager")
+#BiocManager::install("SummarizedExperiment")
+#BiocManager::install("TCGAbiolinks")
 library(TCGAbiolinks)
@@
 #install devtools
-devtools::install.github(selcenari/ceRNAnetsim)
+#devtools::install_github("selcenari/ceRNAnetsim")
 library(ceRNAnetsim)
@@
 The methods about miRNA:target interactions are based a basic principle that is reading after isolation of miRNA:target chimerics.
-The datasets contain all the chimeric miRNA:target structures found in the environment during the experiment.
+The datasets contain all the chimeric miRNA:target structures found in the medium during the experiment.
 On the other hand, it could be said that the reading is performed as snapshot.Because of that, the methods can provide different chimeric interactions the same miRNA:target pair.
@@
 Of course, we do not know that the selected gene is responsible for all regulations in the interactions.
-Moreover, we do not think that one gene causes different regulations in whole interactions.
+Moreover, it is not considered that one gene causes different regulations in whole interactions.
 But we wanted to show that a single gene can cause changing regulations of interaction of whole miRNA target network.
@@
 as.data.frame(E9GE_mirnagenenormal)%>%
-iteration_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect, "ABCC1", .iter =200 , how = 4.10, limit = 1)
+ priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
+ update_how("ABCC1",4.10) %>%
+ simulate(100) %>%
+ find_iteration(limit=0, plot= TRUE)
 
@@
 ```{r}
-find_iteration(as.data.frame(E9GE_mirnagenenormal), GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect, "ABCC1", .iter =100 , how = 4.10, limit = 10)
+as.data.frame(E9GE_mirnagenenormal)%>%
+ priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
+ update_how("ABCC1",4.10) %>%
+ simulate(100) %>%
+ find_iteration(limit=0, plot= FALSE)
 
@@
   priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
-  update_nodes(once=TRUE)%>%
   update_how("ABCC1", 4.10)%>%
-  update_nodes()%>%
   simulate(26)

### 2019-08-22 article.Rmd
last thing
article files

@@
 author:
-- Selcen Ari^[Department of Bioengineering, Yildiz Technical University, Istanbul, Turkey]
-- Alper Yilmaz^[Department of Bioengineering, Yildiz Technical University, Istanbul, Turkey]
+- Selcen Ari^[Department of Bioengineering, Yildiz Technical University, Istanbul,
+  Turkey]
+- Alper Yilmaz^[Department of Bioengineering, Yildiz Technical University, Istanbul,
+  Turkey]
 date: "`r format(Sys.time(), '%d %B %Y')`"
 output:
+  word_document: default
   pdf_document:
-    keep_tex: true
+    keep_tex: yes
   html_document: default
+documentclass: article
+link-citations: yes
+cls: NAR.cls
+template: NAR-sample.tex
 bibliography: ref.bib
@@
 Although the mechanisms by which miRNA represses are not fully understood, miRNAs predominantly repress their targets.
-Repressive activities of miRNAs vary depending on many factors that are significant to microRNA:target interactions.
-These factors include microRNA:target binding energy, binding location in target sequence, base pairing types between miRNA and target, abundance of miRNAs and targets [@grimson_microrna_2007].
+Repressive activities of miRNAs vary depending on many factors that are significant to miRNA:target interactions.
+These factors include miRNA:target binding energy, binding location in target sequence, base pairing types between miRNA and target, abundance of miRNAs and targets [@grimson_microrna_2007].
 Binding energies of miRNA:target complexes vary based on nucleotide context and determine folding stability of complex [@cao_predicting_2012].
@@
 Various computational and experimental studies have tackled the problem of unraveling ceRNA:miRNA interactions.
-It has been observed that when abundance of one of the targets of miR-122 was increased, the other target expression also slightly increased as a result of decreasing repressive activity of miR-122 on them [@denzler_assessing_2014].
+It has been observed that when abundance of one of the targets of miR-122 was increased, the other targets' expression also slightly increased as a result of decreasing repressive activity of miR-122 on them [@denzler_assessing_2014].
 Bosson et al.
@@
 ### Triggering perturbation and subsequent calculations
-Initially, the network is assumed in steady-state (\autoref{fig1}a) condition and needs a trigger for initiating calculations.
+Initially, the network is assumed in steady-state (\autoref{fig1}a) condition and needs least one trigger for initiating calculations.
 The trigger can be a change in expression level of one or more genes (\autoref{fig1}b).
@@
 ceRNA) is updated.
-Due to nodes which are targeted by multiple miRNAs, the change in one neighborhood spreads to other neighborhoods (\autoref{fig1}d), consequently have potential to effect whole network due to "ripple effect".
+This update results in change of expression value of common target (G4) in system.
+In this case, the common target acts as trigger for the other group (the targets interacted with M2 miRNA) in the network.
+Due to common targeted elements, the change in one neighborhood spreads to other neighborhoods (\autoref{fig1}d), consequently have potential to effect whole network due to "ripple effect".
 
@@
 
-[#TODO]: <> "caption needs more text, every color should be explained"
+[#FIX]: <> "caption needs more text, every color should be explained"
 [answer]: <> "explained in caption."
@@
 
-The repression efficiency of a miRNA on the individual target ($Eff_{gi}$) is calculated according to equation \eqref{eq:1}; where miRNA expression ($C_m$) in local neighborhood is distributed among targets using individual gene expression levels ($C_{gi}$) (#TODO isn't $C_g$ gene expressions in the group is equal to total of $C_{gi}$ in group?
+The repression efficiency of a miRNA on the individual target ($Eff_{gi}$) is calculated according to equation \eqref{eq:1}; where miRNA expression ($C_m$) in local neighborhood is distributed among targets using individual gene expression levels ($C_{gi}$) (#FIX isn't $C_g$ gene expressions in the group is equal to total of $C_{gi}$ in group?
 answer: its my false.
@@
 \end{equation}
+
 \begin{equation}
@@
 Factors effecting binding determine interaction between miRNA and target and they alter amount of miRNA sequestered to target.
-Degradation efficiency factors assign amount of degradated target amount in miRNA:target pairs.(#TODO fix this sentence: answer:).
+Degradation efficiency factors assign amount of degradated target amount in miRNA:target pairs.(#FIX fix this sentence: answer:).
 In other words, binding factors exert their influence before or during binding, efficiency factors exert their influence after binding.
@@
 The normalized values of factors take into account to determine binding activity and miRNA efficiency on targets (\autoref{fig2}).
-Binding affinities (activity, $Eff$) of miRNAs on each individual gene are calculated as shown in equation \eqref{eq:4}; $C_m$, miRNA expression in the group; $C_g$, Gene expression; $C_{gi}$, individual gene expression; gi, individual gene; g, whole genes in a group (\autoref{fig2}c).
+Binding affinities (activity, $Eff$) of miRNAs on each individual gene are calculated as shown in equation \eqref{eq:3}; $C_m$, miRNA expression in the group; $C_{gi}$, individual gene expression; gi, individual gene (\autoref{fig2}c).
 \begin{equation}
-Eff_{gi}= C_m \times E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}/(\sum_{1}^{i} E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}) \tag{4}\label{eq:4}
+Eff_{gi}= C_m \times E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}/(\sum_{1}^{i} E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}) \tag{3}\label{eq:3}
 \end{equation}
+
 \begin{equation}
-Eff_{gi}= Eff_{gi}*RE^\prime_{gi} \tag{5}\label{eq:5}
+Eff_{gi}= Eff_{gi}\times RE^\prime_{gi} \tag{4}\label{eq:4}
 \end{equation}
@@
 
+<answer> I shown in figure and explained at following. I didn't fully understand whether it is enough or not.
+
 After miRNA binds to its target, but might not repress to bound target.
 The functionality of bound miRNA on target depends on efficiency factors like region that is binding sequence of miRNA on its target.
-Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:5} (\autoref{fig2}d); $RE^\prime_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
+Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:4} (\autoref{fig2}d); $RE^\prime_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
 The cooperative repression activity of miRNAs to their common targets is figured out as shown in \autoref{fig2}e.
@@
 We have applied our model in a real dataset for which experimental measurements of various factors were available.
-Expression levels of miRNA and genes in tumor and normal tissue of single  patient are retrieved from TCGA(#TODO cite).
+Expression levels of miRNA and genes in tumor and normal tissue of single  patient are retrieved from [TCGA](https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga)(#TODO).
 High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015].
 We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from ...
-(#TODO name of database and its citation)->suplementary datasets of [@helwak_mapping_2013; @moore_mirnatarget_2015] .
+(#FIX name of database and its citation) answer->suplementary datasets of [@helwak_mapping_2013; @moore_mirnatarget_2015].
 Detailed description of network construction and its code is available in Supplementary data (#TODO burada link nasıl olacak, word dosyası veya PDF dosyası ismi mi vermemiz mi  gerekiyor, yoksa sadece Sup Data denilmesi yeterli mi?
@@
 
-[#TODO]: <> "about formulas, 1) eq:3 F/Fmax can be removed, 2) in formulas eq:1 and eq:2, C_m or Eff_gi is not intuitive, is there consensus terms about effect or nth gene?
-3) in formula eq:4, instead of, 4) in general, `*` sign should be replaced with `\times` and apostrophe should be replaced with `\prime` )"
+[#TODO]: <> "about formulas, 1) eq:3 F/Fmax can be removed, 2) in formulas eq:1 and eq:2, C_m or Eff_gi is not intuitive, is there consensus terms about effect or nth gene?"
 
@@
 Previous studies have shown that if a gene abundance increases in ceRNA system, expression levels of genes targeted by shared miRNA are also affected [@lai_understanding_2016; @salmena_cerna_2011; @tay_multilayered_2014].
-It was observed that primary neighborhoods of the trigger gene change with the same ratio, but cooperative efficiency of miRNAs on common target causes differently change in common target (#TODO last sentence needs some clarification, after "but cooperative...") (\autoref{fig1}d).
+It was observed that expression levels of primary neighborhoods which do not interact with another miRNA of the trigger gene change in relation to their expression.
+However, expression level change of common target (G4) in network has been observed different from G1 that has the same expression value with G4 target at initial conditions because more than one miRNA repress this target (\autoref{fig1}d) (#FIX last sentence needs some clarification, after "but cooperative...").
+
 Genes targeted by multiple miRNAs act as a trigger for adjacent local neighborhood of targeting miRNAs, causing changes in expression levels of genes outside the local neighborhood of original trigger gene.
@@
 It's important to note that the changes in gene expression levels will have more pronounced effect if miRNA:target ratio is high, i.e., more miRNA available per target, which was reported in previous findings [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
-[][#TODO]
+[][#FIX]
 
-[#TODO]: <> (following 4 sentences can be deleted, not so clear.
+[#FIX]: <> (following 4 sentences can be deleted, not so clear.
 OR we can summarize it one-two sentences.
-When the simulation was run in sample dataset that includes lower target abundance, the first response of primary neighborhood of the same group was determined as similar with in system that have high target abundance.
-But regulations of ceRNAs were observed differences in the following steps of simulation.
-In overall state in the end of simulation, we have established that all regulations more prominent when compared with low miRNA:target ratio dataset.
-The model has appeared the importance of mRNA:target ratio in regulation of competing RNAs.)
+When the simulation was run in sample dataset that includes lower target abundance, the first response of primary neighborhood of the same group was determined as similar with in system that has higher total target abundance.
+In overall state in the end of simulation, we have established that all regulations more prominent when compared with low miRNA:target ratio dataset.)
 
@@
 
-[#TODO]: <> "title should be more informative"-> answer.
+[#FIX]: <> "title should be more informative"-> answer.
 
@@
 Therefore, we have integrated aforementioned interaction parameters that could be useful for predicting miRNA repression activity more accurately.
-The sample dataset used in  Figure \autoref{fig1} is recalculated with additional factors in effect [][#TODO] with same trigger, two fold increase in Gene2 (\autoref{fig3}).
+The sample dataset used in \autoref{fig1} is recalculated with additional factors \autoref{tab:one} in effect with same trigger, two fold increase in Gene2 (\autoref{fig3}).
 When the factors were taken into account in the system, miRNA efficiencies varied as shown in \autoref{fig3}a.
-Although the miRNA:target expression ratios in steady-state were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed (compare miRNA distribution values between Figure \autoref{fig1}a and Figure \autoref{fig3}a).
+Although the miRNA:target expression ratios in steady-state were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed.
+
+[#FIX2]: <> "we should compare and contrast Figure 1a and figure 3a, the reader should see how affinity effects distrbution G4:M2, and how efficiency is effected without change in affnity M1:G3"
+[answer2]: <> "I explained the mechanism, but i did not understand  how efficiency is effected without change in affnity M1:G3"
+It is considered that entire miRNAs in the system affect targets according to target:total target ratio \autoref{fig1}a, when interaction factors did not take into account.
+However, in presence of interaction factors in miRNA:target interaction network (shown in \autoref{tab:one}), miRNAs are distributed according to target:total target ratio and affinity factors in first like shown in \autoref{fig2}c.
+After affinity mediated propotional distribution of miRNA expression, degradation factor is considered to specify count of repressive miRNA in pairs because entire bound miRNA target pairs might not be resulted with degradation (It is figured out shown in \autoref{fig2}d).
+For this reason, miRNA1 (M1) has more weak repressive effect (shown as edge variables) on target Gene2 \autoref{fig3}a in comparison with miRNA1 (M1):Gene2 (G2) interaction in \autoref{fig1}a.
+On the other hand, it is possible that taking into accont of these interaction factors could cause increasing miRNA repression activity like in miRNA2:Gene4 interaction \autoref{fig3}a.
 When expression of Gene2 (G2) increased, expression values of all genes also changed differentially because of contribution of efficiency factors (\autoref{fig3} b,c,d).
 We have considered that in our approach energy and seed type of pairs is significant for binding and targeted region is important for repression.
-[][#TODO2]
 
-answer: knitr::kable(minsamp): figure ü fig1_2 dosyasında chunk2ye attım. Burada bir hata veriyor. Düzenleyemedim.
-
-[#TODO]: <> "although data is visible at Figure2, since it's a small  table,  we might add the table showing expression levels, seed type, binding energy for M1, M2 and genes so that reader can see which miRNA  has low affinity and understand Figure 3 better"
-[#TODO2]: <> "we should compare and contrast Figure 1a and figure 3a, the reader should see how affinity effects distrbution G4:M2, and how efficiency is effected without change in affnity M1:G3"
+\begin{table}[ht]
+\centering
+\caption{Expression values of elements and interaction factors of miRNA:target interactions in \autoref{fig3}} 
+\begin{tabular}{rllrrrrr}
+  \hline
+ & competing & miRNA & Competing expression & miRNA expression & seed type & region & energy \\ 
+  \hline
+   & Gene1 & Mir1 & 10000.00 & 1000.00 & 0.43 & 0.30 & -20.00 \\ 
+   & Gene2 & Mir1 & 10000.00 & 1000.00 & 0.43 & 0.01 & -15.00 \\ 
+   & Gene3 & Mir1 & 5000.00 & 1000.00 & 0.32 & 0.40 & -14.00 \\ 
+   & Gene4 & Mir1 & 10000.00 & 1000.00 & 0.23 & 0.50 & -10.00 \\ 
+   & Gene4 & Mir2 & 10000.00 & 2000.00 & 0.35 & 0.90 & -12.00 \\ 
+   & Gene5 & Mir2 & 5000.00 & 2000.00 & 0.05 & 0.40 & -11.00 \\ 
+   & Gene6 & Mir2 & 10000.00 & 2000.00 & 0.01 & 0.80 & -25.00 \\ 
+   \hline
+\end{tabular}
+\label{tab:one}
+\end{table}
 
@@
 **(see details in Supplementary File(fig1_2app.Rmd)**.
-When we applied the method on the minimal dataset, Gene4 has been found to be the most efficient element in terms of number of perturbed elements and miRNA2 (M2) has been found to be causing the highest mean in expression changes.
+On the other hand, when we applied the method on the minimal dataset, Gene4 has been found to be the most efficient element in terms of number of perturbed elements and miRNA2 (M2) has been found to be causing the highest mean in expression changes.
 [][#TODO2]
 
-[#TODO]: <> "didnt we perform upregulation of Gene2, why did we say Gene4 here?" 
+[#FIX]: <> "didnt we perform upregulation of Gene2, why did we say Gene4 here?" 
 [answer]: <> "Because, in fact, We want to show that triggering element is significant in the same dataset, in the same conditions. Maybe I must add a conjuction here."
 
-[][#TODO] We used our model to simulate a real dataset which contains thousands of genes and hundreds of miRNAs.
+We used our model to simulate a real dataset which contains thousands of genes and hundreds of miRNAs.
 Our  model can successfully simulate perturbations in such large network despite complex behaviors and struggle to reach steady-state **(see details in Supplementary File(TCGA_E9-A1N5_article.Rmd)**.
@@
 
-[#TODO]: <> "removed this section, didnt understand, confusing: So, we have compared two datasets; the tumor gene expression and simulation results.
-When we analyzed comparison of two datasets, we have observed that the results of simulation is not convenient with tumor tissue expression values.
-The simulation results were not expected to be concordant with tumor tissue expression values because of a single gene could not be responsible for gene regulation in all tissue.
-Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored.
-Of course, it would be more useful to test a dataset that includes gene expression values before and after the regulation of a gene and miRNA expression values at initial conditions.
-When we evaluated the suitability of approach, the approach allows different miRNA targets to be considered together and can work with different regulatory elements due to the network structure."
-
-[#TODO]: <> "removed this section, was confusing: The defined miRNA’ number is very low in comparison with defined protein coding genes’ number.
-So, this results in that the miRNA amounts in the system must be high or targeted RNAs’ amount must be low for effective miRNA activity.
-"
-
 Network based approaches for analyzing miRNA:target interactions have been developed in earlier studies.
@@
 Next, a network based kinetic model integrating miRNA and target rates of transcription, degradation, binding and unbinding was developed [@nitzan_interactions_2014] using high throughput experimental dataset about miR-92a depletion [@helwak_mapping_2013].
-In this study, it was demonstrated that distant ceRNAs can interact with each other via indirect links, and the interactions are effected depending on distance between ceRNAs or topological features of network [@nitzan_interactions_2014].
+It was demonstrated that distant ceRNAs can interact with each other via indirect links, and the interactions are effected depending on distance between ceRNAs or topological features of network [@nitzan_interactions_2014].
 More recently, an approach to detect ceRNA interaction by using the miRNA expression, gene expression and common miRNAs between gene targets was developed [@markus_list_sponge_2017] which was effective in analyzing genes through miRNAs.
-Based on the observations that a miRNA can exhibit strong functionality to a target but may not against an other, the authors have concluded that common miRNA based approach may not be suitable for understanding regulations of ceRNA interactions.
+Based on the observations that a miRNA can exhibit strong functionality to a target but may not against an other, the authors have concluded that existing miRNA based approach may not be suitable for understanding regulations of ceRNA interactions.
 
-[#TODO]: <> "we can talk about BRCA results here: ABCC1 gene, *Multidrug resistance-associated protein 1*, is one of the most significant factor to develop resistance against chemotherapotic agents[@atalay2008role; @lu2015microrna; @atalay2006multidrug]"
+[#TODO answer] We have observed that the results of simulation with ABCC1 gene, *Multidrug resistance-associated protein 1*, which is one of the most significant factor to develop resistance against chemotherapotic agents[@atalay2008role; @lu2015microrna; @atalay2006multidrug] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from [TCGA](https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga) and simulation results obtained via our approach.
+Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues.
+So, it might not said that a single gene is responsible for entire regulations in cancer cells.
+Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored.
+Therefore, it would be more useful to test our approach with an ideal dataset that includes entire informations of gene expression values before and after the regulation of a known gene at existing tissue, and miRNA expression values of this tissue at initial conditions.
+
+[#TODO]: <> "we can talk about BRCA results here: FIX at previous pragraph"
 

### 2019-08-22 article.tex
article files

@@
 }{}
-\usepackage[margin=1in]{geometry}
 \usepackage{hyperref}
@@
   \postdate{\par}
-    \date{09 Ağustos 2019}
+    \date{21 Ağustos 2019}
 
@@
 because of repression functionality depended on interaction factors of
-miRNA target pairs (TOFIX: Hocam etkileşim faktörlerinden dolayı
+miRNA target pairs (answer: Hocam etkileşim faktörlerinden dolayı
 ekspresyonları yakın olsa bile baskılanma aktitesinin aynı
@@
 between miRNA and target, abundance of miRNAs and targets (Grimson et
-al. 2007). Binding energies of miRNA:target complexes vary based on
-nucleotide context and determine folding stability of complex (Cao and
-Chen 2012). It has been demonstrated that the binding energy between
-miRNA and target indicates stability or affinity of complex (Helwak et
-al. 2013) and does not directly determine repressive activity of miRNA
-(Cao and Chen 2012). Early studies have argued that 2-8 nt sequence,
-seed, located in miRNA 5'end bind to specific sequence located in 3'UTR
-of its target (Bartel 2004; Lewis, Burge, and Bartel 2005). In recent
+al. \protect\hyperlink{ref-grimson_microrna_2007}{2007}). Binding
+energies of miRNA:target complexes vary based on nucleotide context and
+determine folding stability of complex (Cao and Chen
+\protect\hyperlink{ref-cao_predicting_2012}{2012}). It has been
+demonstrated that the binding energy between miRNA and target indicates
+stability or affinity of complex (Helwak et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}) and does not directly
+determine repressive activity of miRNA (Cao and Chen
+\protect\hyperlink{ref-cao_predicting_2012}{2012}). Early studies have
+argued that 2-8 nt sequence, seed, located in miRNA 5'end bind to
+specific sequence located in 3'UTR of its target (Bartel
+\protect\hyperlink{ref-bartel_micrornas_2004}{2004}; Lewis, Burge, and
+Bartel \protect\hyperlink{ref-lewis_conserved_2005}{2005}). In recent
 studies, it has been shown that miRNAs can interact with targets via
-sequences located in regions such as 5'UTR or CDS (Hausser et al. 2013;
-Helwak et al. 2013; Moore et al. 2015). These studies also showed that
-binding location could indicate functionality of miRNA:target
-interaction or be effective on abundance of targets(TOFIX abundance
-etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı?,
-evet hocam anlaşılmıyorsa değiştirebilirim). It has been shown that
-miRNAs exhibit repressive activity via, 6-8 nt long sequence that is
-perfectly complementary with targets, seed region at the 5' end of
-miRNAs (Bartel 2009; Grimson et al. 2007). On the other hand, some
-researchers have reported that seed sequence of miRNA can have
+sequences located in regions such as 5'UTR or CDS (Hausser et al.
+\protect\hyperlink{ref-hausser_analysis_2013}{2013}; Helwak et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
+\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). These studies
+also showed that binding location could indicate functionality of
+miRNA:target interaction or be effective on abundance of targets(TOFIX
+abundance etkisi derken, bazen azaltabilir, bazen arttırabilir,
+manasında mı? answer: evet hocam anlaşılmıyorsa değiştirebilirim). It
+has been shown that miRNAs exhibit repressive activity via, 6-8 nt long
+sequence that is perfectly complementary with targets, seed region at
+the 5' end of miRNAs (Bartel
+\protect\hyperlink{ref-bartel_micrornas:_2009}{2009}; Grimson et al.
+\protect\hyperlink{ref-grimson_microrna_2007}{2007}). On the other hand,
+some researchers have reported that seed sequence of miRNA can have
 mismatches or bulged/wobble nucleotides and may locate in region other
-than 5'end of miRNAs (Hafner et al. 2010; Helwak et al. 2013). On top of
-all these factors, abundance of miRNAs and targets and miRNA:target
-ratio in cells predominantly affect efficiency of miRNA:target
-interaction (Arvey et al. 2010; Bosson, Zamudio, and Sharp 2014; Denzler
-et al. 2014).
+than 5'end of miRNAs (Hafner et al.
+\protect\hyperlink{ref-hafner_transcriptome-wide_2010}{2010}; Helwak et
+al. \protect\hyperlink{ref-helwak_mapping_2013}{2013}). On top of all
+these factors, abundance of miRNAs and targets and miRNA:target ratio in
+cells predominantly affect efficiency of miRNA:target interaction (Arvey
+et al. \protect\hyperlink{ref-arvey_target_2010}{2010}; Bosson, Zamudio,
+and Sharp \protect\hyperlink{ref-bosson_endogenous_2014}{2014}; Denzler
+et al. \protect\hyperlink{ref-denzler_assessing_2014}{2014}).
 
@@
 targeted mRNAs exhibit competitor behavior, that is hypothesized as
-competing endogenous RNAs (ceRNAs) (Ala et al. 2013; Cesana and Daley
-2013), against their miRNAs. Briefly, Ala et al.~have explained the
-ceRNA hypothesis as disturbance of the other target when one of the
-targets on a steady-state system that included one miRNA and two target
-was perturbed with expression change (Ala et al. 2013). Regarding
+competing endogenous RNAs (ceRNAs) (Ala et al.
+\protect\hyperlink{ref-ala_integrated_2013}{2013}; Cesana and Daley
+\protect\hyperlink{ref-cesana_deciphering_2013}{2013}), against their
+miRNAs. Briefly, Ala et al.~have explained the ceRNA hypothesis as
+disturbance of the other target when one of the targets on a
+steady-state system that included one miRNA and two target was perturbed
+with expression change (Ala et al.
+\protect\hyperlink{ref-ala_integrated_2013}{2013}). Regarding
 interaction between miRNAs and their target in a cell, explaining and
@@
 increased as a result of decreasing repressive activity of miR-122 on
-them (Denzler et al. 2014). Bosson et al.~have developed a mathematical
-model for changes on total target pool concentration after grouping
-targets according to affinity and demonstrated that miRNA activity
-correlated with affinity between miRNA and target (Bosson, Zamudio, and
-Sharp 2014). Cooperative efficiency of miRNAs as well as competitor
-behaviors of targets were also studied and it has been demonstrated to
-be crucial for regulating available mRNA levels of targets (Denzler et
-al. 2016). MiRNA:target interactions have been modeled as stoichiometric
-and catalytic mechanisms and Figliuzzi et al.~have recommended handling
-models in network context (Figliuzzi, Marinari, and De Martino 2013).
-The model that can explain miRNA target interaction through topological
-features has been applied at bipartite network by Nitzan et al.~(Nitzan
-et al. 2014). Robinson and Henderson applied the model that handles
-miRNA:target direct and indirect interactions via common miRNA of gene
-and target of miRNA, at bipartite network. It has been demonstrated that
-all miRNAs and targets in the network can interact with each other
-through common miRNAs and genes, without interaction between the same
-type of nodes (Robinson and Henderson 2018).Associated genes that are
-targets of the same miRNAs have been found with help of correlation of
-gene expression changes in recent algorithm (Markus List 2017). List et
-al.~have specified that their approach can be useful for ceRNA studies
-and published their approach as an R package.
+them (Denzler et al.
+\protect\hyperlink{ref-denzler_assessing_2014}{2014}). Bosson et
+al.~have developed a mathematical model for changes on total target pool
+concentration after grouping targets according to affinity and
+demonstrated that miRNA activity correlated with affinity between miRNA
+and target (Bosson, Zamudio, and Sharp
+\protect\hyperlink{ref-bosson_endogenous_2014}{2014}). Cooperative
+efficiency of miRNAs as well as competitor behaviors of targets were
+also studied and it has been demonstrated to be crucial for regulating
+available mRNA levels of targets (Denzler et al.
+\protect\hyperlink{ref-denzler_impact_2016}{2016}). MiRNA:target
+interactions have been modeled as stoichiometric and catalytic
+mechanisms and Figliuzzi et al.~have recommended handling models in
+network context (Figliuzzi, Marinari, and De Martino
+\protect\hyperlink{ref-figliuzzi_micrornas_2013}{2013}). The model that
+can explain miRNA target interaction through topological features has
+been applied at bipartite network by Nitzan et al.~(Nitzan et al.
+\protect\hyperlink{ref-nitzan_interactions_2014}{2014}). Robinson and
+Henderson applied the model that handles miRNA:target direct and
+indirect interactions via common miRNA of gene and target of miRNA, at
+bipartite network. It has been demonstrated that all miRNAs and targets
+in the network can interact with each other through common miRNAs and
+genes, without interaction between the same type of nodes (Robinson and
+Henderson
+\protect\hyperlink{ref-robinson_modelling_2018}{2018}).Associated genes
+that are targets of the same miRNAs have been found with help of
+correlation of gene expression changes in recent algorithm (Markus List
+\protect\hyperlink{ref-markus_list_sponge_2017}{2017}). List et al.~have
+specified that their approach can be useful for ceRNA studies and
+published their approach as an R package.
 
@@
 coconut) according to proportion of their targets' expression. b) Two
-fold increase in transcript level of Gene2 (G2) acts as a trigger. c)
-Distribution of miRNA1 (M1) changes. d) The change at expression of
-common target effects changes of proportional distribution of miRNA2
-(M2). Expression values are rounded to integers for
-simplicity}\label{fig1}
+fold increase in transcript level of Gene2 (G2) acts as a trigger (shown
+with red in figure). c) Distribution of miRNA1 (M1) changes. d) The
+change at expression (shown with light red because minor changes are
+occured on their expression values.) of common target effects changes of
+proportional distribution of miRNA2 (M2). Expression values are rounded
+to integers for simplicity}\label{fig1}
 }
@@
 miRNA expression (\(C_m\)) in local neighborhood is distributed among
-targets using individual gene expression levels (\(C_{gi}\)) (\#TODO
+targets using individual gene expression levels (\(C_{gi}\)) (\#FIX
 isn't \(C_g\) gene expressions in the group is equal to total of
-\(C_{gi}\) in group?). For the genes targeted by multiple miRNAs,
-cooperative activity of miRNAs, \(R\), is calculated by summing
-repression activity of each miRNA (Equation \eqref{eq:2}).
+\(C_{gi}\) in group? answer: its my false. now its corect.). For the
+genes targeted by multiple miRNAs, cooperative activity of miRNAs on a
+target gene, \(R\), is calculated by summing repression activity of each
+miRNA (Equation \eqref{eq:2}).
 
 \begin{equation} 
-    Eff_{gi}= C_m \times C_{gi}/\sum_{1}^{i} C_g \tag{1}\label{eq:1}
-\end{equation} \begin{equation}
+    Eff_{gi}= C_m \times C_{gi}/\sum_{1}^{i} C_{gi} \tag{1}\label{eq:1}
+\end{equation}
+
+\begin{equation}
    R_{gi}= Eff_{i1} + Eff_{i2} ... \tag{2}\label{eq:2}
@@
 miRNA and target and they alter amount of miRNA sequestered to target.
-Factors effecting efficiency determine proportion of target mRNA
-degraded by bound miRNA (\#TODO fix this sentence). In other words,
-binding factors exert their influence before or during binding,
+Degradation efficiency factors assign amount of degradated target amount
+in miRNA:target pairs.(\#FIX fix this sentence: answer:). In other
+words, binding factors exert their influence before or during binding,
 efficiency factors exert their influence after binding. In the
-literature, binding free energy (Cao and Chen 2012 ; Helwak et al. 2013)
-and seed type (Werfel et al. 2017) in miRNA:target interactions are
-described as factors effecting binding affinity. Efficiency factors
-determine how many of miRNA:target complexes will result in inhibition
-and binding region on target drastically effect miRNA degradation
-efficiency (Hausser et al. 2013; Helwak et al. 2013). Both binding and
+literature, binding free energy (Cao and Chen
+\protect\hyperlink{ref-cao_predicting_2012}{2012} ; Helwak et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}) and seed type (Werfel
+et al. \protect\hyperlink{ref-werfel_preferential_2017}{2017}) in
+miRNA:target interactions are described as factors effecting binding
+affinity. Efficiency factors determine how many of miRNA:target
+complexes will result in inhibition and binding region on target
+drastically effect miRNA degradation efficiency (Hausser et al.
+\protect\hyperlink{ref-hausser_analysis_2013}{2013}; Helwak et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}). Both binding and
 efficiency factors are normalized to their maximum values and scaled to
@@
 (\autoref{fig2}). Binding affinities (activity, \(Eff\)) of miRNAs on
-each individual gene are calculated as shown in equation \eqref{eq:4};
-\(C_m\), miRNA expression in the group; \(C_g\), Gene expression;
-\(C_{gi}\), individual gene expression; gi, individual gene; g, whole
-genes in a group (\autoref{fig2}c). \begin{equation}
-Eff_{gi}= C_m \times E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}/(\sum_{1}^{i} E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}) \tag{4}\label{eq:4}
-\end{equation} \begin{equation}
-Eff_{gi}= Eff_{gi}*RE^\prime_{gi} \tag{5}\label{eq:5}
+each individual gene are calculated as shown in equation \eqref{eq:3};
+\(C_m\), miRNA expression in the group; \(C_{gi}\), individual gene
+expression; gi, individual gene (\autoref{fig2}c). \begin{equation}
+Eff_{gi}= C_m \times E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}/(\sum_{1}^{i} E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}) \tag{3}\label{eq:3}
+\end{equation}
+
+\begin{equation}
+Eff_{gi}= Eff_{gi}\times RE^\prime_{gi} \tag{4}\label{eq:4}
 \end{equation}
@@
 
+I shown in figure and explained at following. I didn't fully understand
+whether it is enough or not.
+
 After miRNA binds to its target, but might not repress to bound target.
@@
 repression efficiency of miRNA is calculated according to equation
-\eqref{eq:5} (\autoref{fig2}d); \(RE^\prime_{gi}\), normalized values of
+\eqref{eq:4} (\autoref{fig2}d); \(RE^\prime_{gi}\), normalized values of
 region efficiency coefficient between miRNA and gene. The cooperative
@@
 miRNA and genes in tumor and normal tissue of single patient are
-retrieved from TCGA(\#TODO cite). High-throughput experimental datasets
-which are provided miRNA:gene target pairs with interaction factors
-(Helwak et al. 2013; Moore et al. 2015). We have combined miRNA and gene
-expression datasets via miRNA:target gene dataset retrieved from
-\ldots{} (\#TODO name of database and its
-citation)-\textgreater{}suplementary datasets of (Helwak et al. 2013;
-Moore et al. 2015) . Detailed description of network construction and
-its code is available in Supplementary data (\#TODO burada link nasıl
-olacak, word dosyası veya PDF dosyası ismi mi vermemiz mi gerekiyor,
-yoksa sadece Sup Data denilmesi yeterli mi? ilgili dosya:
-TCGA\_E9-A1N5\_article.Rmd). ABCC1 gene, over-expressed in tumor tissue,
-was selected as trigger for simulation of integrated dataset. After
-simulation of network, we have compared simulation results and tumor
-tissue expression levels.
+retrieved from
+\href{https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga}{TCGA}(\#TODO).
+High-throughput experimental datasets which are provided miRNA:gene
+target pairs with interaction factors (Helwak et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
+\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). We have combined
+miRNA and gene expression datasets via miRNA:target gene dataset
+retrieved from \ldots{} (\#FIX name of database and its citation)
+answer-\textgreater{}suplementary datasets of (Helwak et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
+\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). Detailed
+description of network construction and its code is available in
+Supplementary data (\#TODO burada link nasıl olacak, word dosyası veya
+PDF dosyası ismi mi vermemiz mi gerekiyor, yoksa sadece Sup Data
+denilmesi yeterli mi? ilgili dosya: TCGA\_E9-A1N5\_article.Rmd). ABCC1
+gene, over-expressed in tumor tissue, was selected as trigger for
+simulation of integrated dataset. After simulation of network, we have
+compared simulation results and tumor tissue expression levels.
 
@@
 of miRNA repression activity has been based on miRNA and target
-abundance in various researches(Arvey et al. 2010; Denzler et al. 2014).
-Our approach can effortlessly calculate effects of expression changes
-when abundance levels of miRNAs and targets is only available factor. In
-sample network given in Figure (\autoref{fig1}), after an increase in
-expression level of a gene (G2), expression values of other genes also
-changed due to redistribution of miRNA among its targets. Previous
-studies have shown that if a gene abundance increases in ceRNA system,
-expression levels of genes targeted by shared miRNA are also affected
-(Lai, Wolkenhauer, and Vera 2016; Salmena et al. 2011; Tay, Rinn, and
-Pandolfi 2014). It was observed that primary neighborhoods of the
-trigger gene change with the same ratio, but cooperative efficiency of
-miRNAs on common target causes differently change in common target
-(\#TODO last sentence needs some clarification, after ``but
-cooperative\ldots{}'') (\autoref{fig1}d). Genes targeted by multiple
-miRNAs act as a trigger for adjacent local neighborhood of targeting
-miRNAs, causing changes in expression levels of genes outside the local
-neighborhood of original trigger gene. Therefore, primary expression
-level change in gene (G2) causes changes in other group of genes (G5 and
-G6) even though original trigger gene (G2) and genes in other group are
-not targeted by common miRNA. In addition, as shown by ceRNA hypothesis
-model of Ala et al., after the increase of gene expression level of G2,
-the miRNA that is found in the same group (M1) tended to be less
-repressive on its remaining targets (G1, G3 and G4). It's important to
-note that the changes in gene expression levels will have more
-pronounced effect if miRNA:target ratio is high, i.e., more miRNA
-available per target, which was reported in previous findings (Arvey et
-al. 2010; Bosson, Zamudio, and Sharp 2014; Denzler et al. 2014).
-\href{}{}
+abundance in various researches(Arvey et al.
+\protect\hyperlink{ref-arvey_target_2010}{2010}; Denzler et al.
+\protect\hyperlink{ref-denzler_assessing_2014}{2014}). Our approach can
+effortlessly calculate effects of expression changes when abundance
+levels of miRNAs and targets is only available factor. In sample network
+given in Figure (\autoref{fig1}), after an increase in expression level
+of a gene (G2), expression values of other genes also changed due to
+redistribution of miRNA among its targets. Previous studies have shown
+that if a gene abundance increases in ceRNA system, expression levels of
+genes targeted by shared miRNA are also affected (Lai, Wolkenhauer, and
+Vera \protect\hyperlink{ref-lai_understanding_2016}{2016}; Salmena et
+al. \protect\hyperlink{ref-salmena_cerna_2011}{2011}; Tay, Rinn, and
+Pandolfi \protect\hyperlink{ref-tay_multilayered_2014}{2014}). It was
+observed that expression levels of primary neighborhoods which do not
+interact with another miRNA of the trigger gene change in relation to
+their expression. However, expression level change of common target (G4)
+in network has been observed different from G1 that has the same
+expression value with G4 target at initial conditions because more than
+one miRNA repress this target (\autoref{fig1}d) (\#FIX last sentence
+needs some clarification, after ``but cooperative\ldots{}'').
+
+Genes targeted by multiple miRNAs act as a trigger for adjacent local
+neighborhood of targeting miRNAs, causing changes in expression levels
+of genes outside the local neighborhood of original trigger gene.
+Therefore, primary expression level change in gene (G2) causes changes
+in other group of genes (G5 and G6) even though original trigger gene
+(G2) and genes in other group are not targeted by common miRNA. In
+addition, as shown by ceRNA hypothesis model of Ala et al., after the
+increase of gene expression level of G2, the miRNA that is found in the
+same group (M1) tended to be less repressive on its remaining targets
+(G1, G3 and G4). It's important to note that the changes in gene
+expression levels will have more pronounced effect if miRNA:target ratio
+is high, i.e., more miRNA available per target, which was reported in
+previous findings (Arvey et al.
+\protect\hyperlink{ref-arvey_target_2010}{2010}; Bosson, Zamudio, and
+Sharp \protect\hyperlink{ref-bosson_endogenous_2014}{2014}; Denzler et
+al. \protect\hyperlink{ref-denzler_assessing_2014}{2014}). \href{}{}
 
@@
 
-\href{}{\#TODO}: \textless{}\textgreater{} ``title should be more
-informative''-\textgreater{}
+\href{}{\#FIX}: \textless{}\textgreater{} ``title should be more
+informative''-\textgreater{} answer.
 
@@
 target site location by using proteomics dataset (Xu, Wang, and Liu
-2014). They have proposed that the features of binding between miRNA and
-target can be critical for miRNA efficiency. In addition, binding energy
-between miRNA and targets is a significant determinant for miRNA
-efficiency and it has been reported that strength of miRNA:target
-interactions is depended on binding energy of complexes (Breda et al.
-2015). Similarly, another study revealed that affinity is correlated
-with seed pairing of miRNA:target pairs and suggested that affinity is
-correlated with number of canonical seed base pairing (Bosson, Zamudio,
-and Sharp 2014). Therefore, we have integrated aforementioned
-interaction parameters that could be useful for predicting miRNA
-repression activity more accurately. The sample dataset used in Figure
-\autoref{fig1} is recalculated with additional factors in effect
-\href{}{} with same trigger, two fold increase in Gene2
-(\autoref{fig3}). When the factors were taken into account in the
-system, miRNA efficiencies varied as shown in \autoref{fig3}a. Although
-the miRNA:target expression ratios in steady-state were same in
+\protect\hyperlink{ref-xu_characterization_2014}{2014}). They have
+proposed that the features of binding between miRNA and target can be
+critical for miRNA efficiency. In addition, binding energy between miRNA
+and targets is a significant determinant for miRNA efficiency and it has
+been reported that strength of miRNA:target interactions is depended on
+binding energy of complexes (Breda et al.
+\protect\hyperlink{ref-breda_quantifying_2015}{2015}). Similarly,
+another study revealed that affinity is correlated with seed pairing of
+miRNA:target pairs and suggested that affinity is correlated with number
+of canonical seed base pairing (Bosson, Zamudio, and Sharp
+\protect\hyperlink{ref-bosson_endogenous_2014}{2014}). Therefore, we
+have integrated aforementioned interaction parameters that could be
+useful for predicting miRNA repression activity more accurately. The
+sample dataset used in \autoref{fig1} is recalculated with additional
+factors \autoref{tab:one} in effect with same trigger, two fold increase
+in Gene2 (\autoref{fig3}). When the factors were taken into account in
+the system, miRNA efficiencies varied as shown in \autoref{fig3}a.
+Although the miRNA:target expression ratios in steady-state were same in
 comparison with the sample dataset without factors, efficiency of
-binding and repression have changed (compare miRNA distribution values
-between Figure \autoref{fig1}a and Figure \autoref{fig3}a). When
-expression of Gene2 (G2) increased, expression values of all genes also
-changed differentially because of contribution of efficiency factors
+binding and repression have changed.
+
+It is considered that entire miRNAs in the system affect targets
+according to target:total target ratio \autoref{fig1}a, when interaction
+factors did not take into account. However, in presence of interaction
+factors in miRNA:target interaction network (shown in
+\autoref{tab:one}), miRNAs are distributed according to target:total
+target ratio and affinity factors in first like shown in
+\autoref{fig2}c.~After affinity mediated propotional distribution of
+miRNA expression, degradation factor is considered to specify count of
+repressive miRNA in pairs because entire bound miRNA target pairs might
+not be resulted with degradation (It is figured out shown in
+\autoref{fig2}d). For this reason, miRNA1 (M1) has more weak repressive
+effect (shown as edge variables) on target Gene2 \autoref{fig3}a in
+comparison with miRNA1 (M1):Gene2 (G2) interaction in \autoref{fig1}a.
+On the other hand, it is possible that taking into accont of these
+interaction factors could cause increasing miRNA repression activity
+like in miRNA2:Gene4 interaction \autoref{fig3}a. When expression of
+Gene2 (G2) increased, expression values of all genes also changed
+differentially because of contribution of efficiency factors
 (\autoref{fig3} b,c,d). We have considered that in our approach energy
 and seed type of pairs is significant for binding and targeted region is
-important for repression. \href{}{}
+important for repression.
 
-knitr::kable(minsamp)
+\begin{table}[ht]
+\centering
+\caption{Expression values of elements and interaction factors of miRNA:target interactions in \autoref{fig3}} 
+\begin{tabular}{rllrrrrr}
+  \hline
+ & competing & miRNA & Competing expression & miRNA expression & seed type & region & energy \\ 
+  \hline
+   & Gene1 & Mir1 & 10000.00 & 1000.00 & 0.43 & 0.30 & -20.00 \\ 
+   & Gene2 & Mir1 & 10000.00 & 1000.00 & 0.43 & 0.01 & -15.00 \\ 
+   & Gene3 & Mir1 & 5000.00 & 1000.00 & 0.32 & 0.40 & -14.00 \\ 
+   & Gene4 & Mir1 & 10000.00 & 1000.00 & 0.23 & 0.50 & -10.00 \\ 
+   & Gene4 & Mir2 & 10000.00 & 2000.00 & 0.35 & 0.90 & -12.00 \\ 
+   & Gene5 & Mir2 & 5000.00 & 2000.00 & 0.05 & 0.40 & -11.00 \\ 
+   & Gene6 & Mir2 & 10000.00 & 2000.00 & 0.01 & 0.80 & -25.00 \\ 
+   \hline
+\end{tabular}
+\label{tab:one}
+\end{table}
 
@@
 each gene in the network for their perturbation efficiencies.
-\textbf{(see details in Supplementary File(fig1\_2app.Rmd)}. When we
-applied the method on the minimal dataset, Gene4 has been found to be
-the most efficient element in terms of number of perturbed elements and
-miRNA2 (M2) has been found to be causing the highest mean in expression
-changes. \href{}{}
+\textbf{(see details in Supplementary File(fig1\_2app.Rmd)}. On the
+other hand, when we applied the method on the minimal dataset, Gene4 has
+been found to be the most efficient element in terms of number of
+perturbed elements and miRNA2 (M2) has been found to be causing the
+highest mean in expression changes. {[}{]}{[}\#TODO2{]}
 
@@
 concentrations of ceRNA and miRNAs (Figliuzzi, Marinari, and De Martino
-2013). Next, a network based kinetic model integrating miRNA and target
-rates of transcription, degradation, binding and unbinding was developed
-(Nitzan et al. 2014) using high throughput experimental dataset about
-miR-92a depletion (Helwak et al. 2013). In this study, it was
-demonstrated that distant ceRNAs can interact with each other via
+\protect\hyperlink{ref-figliuzzi_micrornas_2013}{2013}). Next, a network
+based kinetic model integrating miRNA and target rates of transcription,
+degradation, binding and unbinding was developed (Nitzan et al.
+\protect\hyperlink{ref-nitzan_interactions_2014}{2014}) using high
+throughput experimental dataset about miR-92a depletion (Helwak et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}). In this study, it
+was demonstrated that distant ceRNAs can interact with each other via
 indirect links, and the interactions are effected depending on distance
-between ceRNAs or topological features of network (Nitzan et al. 2014).
-More recently, an approach to detect ceRNA interaction by using the
-miRNA expression, gene expression and common miRNAs between gene targets
-was developed (Markus List 2017) which was effective in analyzing genes
-through miRNAs. Based on the observations that a miRNA can exhibit
-strong functionality to a target but may not against an other, the
-authors have concluded that common miRNA based approach may not be
-suitable for understanding regulations of ceRNA interactions.
+between ceRNAs or topological features of network (Nitzan et al.
+\protect\hyperlink{ref-nitzan_interactions_2014}{2014}). More recently,
+an approach to detect ceRNA interaction by using the miRNA expression,
+gene expression and common miRNAs between gene targets was developed
+(Markus List \protect\hyperlink{ref-markus_list_sponge_2017}{2017})
+which was effective in analyzing genes through miRNAs. Based on the
+observations that a miRNA can exhibit strong functionality to a target
+but may not against an other, the authors have concluded that common
+miRNA based approach may not be suitable for understanding regulations
+of ceRNA interactions.
 
@@
 degradation rates of miRNAs change depending on cellular conditions
-(Rüegger and Großhans 2012). However, due to lack of datasets including
-degradation and transcription rates of miRNAs at specific cellular
-conditions, such factors are neglected in our model. Additionally, other
-regulation parameters such as gene-gene interactions and transcription
-factors are ignored but the network structure is flexible and can
-integrate additional regulation elements as they become available. In
-the future, with developments in experimental techniques about
-miRNAs:target interaction dynamica, more accurate results can be
-obtained from our model resulting in better understanding and
-predictions of abnormal regulations and pathways underlying diseases or
-conditions.
+(Rüegger and Großhans
+\protect\hyperlink{ref-ruegger_microrna_2012}{2012}). However, due to
+lack of datasets including degradation and transcription rates of miRNAs
+at specific cellular conditions, such factors are neglected in our
+model. Additionally, other regulation parameters such as gene-gene
+interactions and transcription factors are ignored but the network
+structure is flexible and can integrate additional regulation elements
+as they become available. In the future, with developments in
+experimental techniques about miRNAs:target interaction dynamica, more
+accurate results can be obtained from our model resulting in better
+understanding and predictions of abnormal regulations and pathways
+underlying diseases or conditions.
 

### 2019-08-22 fig1_2app.Rmd
article files

@@
 ---
-title: "supplement"
+title: "Supplementary File 1"
 author: "Selcen Ari"
 date: "7 Nisan 2019"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
+  word_document: default
 ---
@@
 
-*Convertion of dataset*: priming_graph function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
+**Convertion of dataset**: priming_graph function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
 If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
-The formulations that are given in equations (1-5) are performed in this function.
+The formulations that are given in equations (1-4) are performed in this function.
 This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
@@
 
-![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow1.jpg){#fig1}
+![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow1.png){#fig1}
 
-*Transition of variables in graph*: In the previous step, the calculations are performed in the edge variables of the graph object.
+**Transition of variables in graph**: In the previous step, the calculations are performed in the edge variables of the graph object.
 However, the graph object allows to use node variables, while the node features are handled to the graph.
@@
 
-*Trigger change in graph*: The dataset are assumed as steady-state in previous step and the efficieny coefficients are calculated according to this acceptance.
+**Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficieny coefficients are calculated according to this acceptance.
 In the network that is found in steady-state conditions, the change is applied to the graph object for disturbtion of steady-state.
@@
 
-*Updating current values of variables*: After variables updating in edge varibles, these are carried to node variables.
+**Updating current values of variables**: After variables updating in edge varibles, these are carried to node variables.
 Current and previous values of variables are stored as node variables with update_variables function.
 
-*Simulation of competing behavior of targets*: After the change in the steady-state conditions, the network elements try to gain steady-state again.
+**Simulation of competing behavior of targets**: After the change in the steady-state conditions, the network elements try to gain steady-state again.
 This process progresses as repeating of regulations after the spreading the changes in the network.
@@
 ```{r, message=FALSE, warning=FALSE}
+#install.packages("devtools")
+#devtools::install_github("selcenari/ceRNAnetsim")
 library(ceRNAnetsim)
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_nodes(once=TRUE)%>%
   vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset in steady-state conditions")
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_nodes(once=TRUE)%>%
   update_how("Gene2", 2)%>%
-  update_nodes()%>%
   vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors")
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_nodes(once=TRUE)%>%
   update_how("Gene2", 2)%>%
-  update_nodes()%>%
   simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "Regulations after Gene2 Upregulation", 2)
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_nodes(once=TRUE)%>%
   update_how("Gene2", 2)%>%
-  update_nodes()%>%
   simulate(2)%>%
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_nodes(once=TRUE)%>%
   vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset in steady-state conditions")
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_nodes(once=TRUE)%>%
   update_how("Gene2", 2)%>%
-  update_nodes()%>%
   vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors")
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_nodes(once=TRUE)%>%
   update_how("Gene2", 2)%>%
-  update_nodes()%>%
   simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors",2)
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_nodes(once=TRUE)%>%
   update_how("Gene2", 2)%>%
-  update_nodes()%>%
   simulate(3)%>%
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_nodes(once=TRUE)%>%
   update_how("Gene4", 2)%>%
-  update_nodes()%>%
   simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors",2)
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_nodes(once=TRUE)%>%
   update_how("Gene4", 2)%>%
-  update_nodes()%>%
   simulate(3)%>%

### 2019-08-22 article.Rmd
article todo

@@
 
-[#TODO]: <> "about formulas, 1) eq:3 F/Fmax can be removed, 2) in formulas eq:1 and eq:2, C_m or Eff_gi is not intuitive, is there consensus terms about effect or nth gene?"
+[#TODO]: <> "about formulas, 2) in formulas eq:1 and eq:2, C_m or Eff_gi is not intuitive, is there consensus terms about effect or nth gene?"
 

### 2019-08-26 article.Rmd
re-scan to literature

@@
 
+[#add]Similar with these approach, Silveira et al.
+have developed miRmapper package for using with R programme [@da2018mirmapper].
+Researchers have utilized the adjacency matrix to associate miRNAs using differentially expressed genes and found similarities of miRNAs with this approach.
+Authors who utilized the approach in dataset of bladder cancer cell line have identified importance of miRNAs in network, and miRNAs which work synergistically.
+
 [#TODO answer] We have observed that the results of simulation with ABCC1 gene, *Multidrug resistance-associated protein 1*, which is one of the most significant factor to develop resistance against chemotherapotic agents[@atalay2008role; @lu2015microrna; @atalay2006multidrug] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from [TCGA](https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga) and simulation results obtained via our approach.

### 2019-09-01 article.Rmd
git problem
Article edit

@@
 It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is perfectly complementary with targets, seed region at the 5’ end of miRNAs [@bartel_micrornas:_2009; @grimson_microrna_2007].
-On the other hand, some researchers have reported that seed sequence of miRNA can have mismatches or bulged/wobble nucleotides and may locate in region other than 5'end of miRNAs [@hafner_transcriptome-wide_2010; @helwak_mapping_2013].
+On the other hand, some researchers have reported that seed sequence of miRNA can have mismatches or bulged/wobble nucleotides [@chi2012alternative] and may locate in region other than 5'end of miRNAs [@hafner_transcriptome-wide_2010; @helwak_mapping_2013].
 On top of all these factors, abundance of miRNAs and targets and miRNA:target ratio in cells predominantly affect efficiency of miRNA:target interaction [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
@@
 
-[#add]Similar with these approach, Silveira et al.
+[#Add]Similar with this approach, Silveira et al.
 have developed miRmapper package for using with R programme [@da2018mirmapper].
@@
 Authors who utilized the approach in dataset of bladder cancer cell line have identified importance of miRNAs in network, and miRNAs which work synergistically.
+
+[#Add] Today, there are many of tools that provide predicted or experimentally validated miRNA:target interaction dataset.
+The datasets that contain weak, proteomic analysis like western blot, and strong, crosslinking and immunoprecipitation CLiP, experimentally evidence are more reliable sources than prediction datasets.
+While in experimentally weak supported datasets it is not known that indirect or direct target of miRNAs in cell, high-throughput methods ensure handling of bounded miRNA:target pairs.
+For this reason we have preferred to utilize high-throughput datasets in our application.
+We collected datasets that depend on chimeric [@helwak_mapping_2013; @moore_mirnatarget_2015] reading of miRNA: target pairs from among these datasets because they contain exact complementary seed regions.
+Although these experimental sources include exact binding sites, they do not provides which miRNA is functional or not [@liu2019prediction].
+We have used the features of miRNA:target pairs and scaled these parameters inside groups that is consisted according to ,ceRNA (Competing Endogenous RNA), genes targeted by the same miRNA.
+*In scaling process, we have considered previous studies.
+BURAYA O ÇALIŞMALARI ATIFLA*
 

### 2019-09-10 article.Rmd
news to article

@@
 output:
-  word_document: default
   pdf_document:
     keep_tex: yes
+  word_document: default
   html_document: default
@@
 [@nitzan_interactions_2014].
-Robinson and Henderson applied the model that handles miRNA:target direct and indirect interactions via common miRNA of gene and target of miRNA, at bipartite network.
+Robinson and Henderson applied the model that handles miRNA:target direct and indirect interactions via common targeting miRNA of genes and target of miRNAs, at bipartite network.
 It has been demonstrated that all miRNAs and targets in the network can interact with each other through common miRNAs and genes, without interaction between the same type of nodes [@robinson_modelling_2018].Associated genes that are targets of the same miRNAs have been found with help of correlation of gene expression changes in recent algorithm [@markus_list_sponge_2017].
@@
 We have applied our model in a real dataset for which experimental measurements of various factors were available.
-Expression levels of miRNA and genes in tumor and normal tissue of single  patient are retrieved from [TCGA](https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga)(#TODO).
+Expression levels of miRNA and genes in tumor and normal tissue of single  patient are retrieved from The Cancer Genome Atlas [TCGA](https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga)(#TODO).
 High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015].
@@
 We have developed a network-based approach to assess effects of expression level changes in competitive ceRNA regulation.
-The basic mode of miRNA repression activity has been based on miRNA and target abundance in various researches[@arvey_target_2010; @denzler_assessing_2014].
+The basic mode of miRNA repression activity has been based on miRNA and target abundance in various researches [@arvey_target_2010; @denzler_assessing_2014].
 Our approach can effortlessly calculate effects of expression changes when abundance levels of miRNAs and targets is only available factor.
@@
 In regulation of miRNA target sample system (\autoref{fig3}), miRNA (M1) repressive efficiency on the primary triggering gene (G2) is low in steady-state.
-So it has been observed that the regulatory activities of miRNA efficiencies on the targets are weak after the increase of miRNA target G2.
+So it has been observed that the regulatory activities of miRNA on the targets are weak after the increase of miRNA target G2.
 When model is triggered with two fold increase in expression level of common target (G4), the changes of other gene expressions have observed more prominent [][#TODO].
@@
 
-[#Add]Similar with this approach, Silveira et al.
+[#Add]More recently, Silveira et al.
 have developed miRmapper package for using with R programme [@da2018mirmapper].
 Researchers have utilized the adjacency matrix to associate miRNAs using differentially expressed genes and found similarities of miRNAs with this approach.
-Authors who utilized the approach in dataset of bladder cancer cell line have identified importance of miRNAs in network, and miRNAs which work synergistically.
+Authors who utilized the approach in dataset of bladder cancer cell line have identified significant miRNAs in network, and miRNAs which work synergistically.
 
@@
 The datasets that contain weak, proteomic analysis like western blot, and strong, crosslinking and immunoprecipitation CLiP, experimentally evidence are more reliable sources than prediction datasets.
-While in experimentally weak supported datasets it is not known that indirect or direct target of miRNAs in cell, high-throughput methods ensure handling of bounded miRNA:target pairs.
+While in experimentally weak supported datasets it is not known that indirect or direct target of miRNAs in cell, high-throughput methods ensure handling of AGO (Argounate Protein) interacted miRNAs.
 For this reason we have preferred to utilize high-throughput datasets in our application.
-We collected datasets that depend on chimeric [@helwak_mapping_2013; @moore_mirnatarget_2015] reading of miRNA: target pairs from among these datasets because they contain exact complementary seed regions.
-Although these experimental sources include exact binding sites, they do not provides which miRNA is functional or not [@liu2019prediction].
-We have used the features of miRNA:target pairs and scaled these parameters inside groups that is consisted according to ,ceRNA (Competing Endogenous RNA), genes targeted by the same miRNA.
-*In scaling process, we have considered previous studies.
-BURAYA O ÇALIŞMALARI ATIFLA*
+Especially, we collected datasets that depend on chimeric [@helwak_mapping_2013; @moore_mirnatarget_2015] reading of miRNA: target pairs from among these datasets because they contain exact complementary seed regions.
+
+[#Add] Although these experimental sources include exact binding sites, they do not provides which miRNA is functional or not [@liu2019prediction].
+We have used the features of bound miRNA:target pairs and scaled these parameters inside groups that is consisted according to, ceRNA (Competing Endogenous RNA), genes targeted by the same miRNA.
+In scaling process, we specified the interaction factors as numeric values in accordance with previous studies.
+Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study.
+On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study.
+Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions.
+Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values.
+We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
+With this process, we have handled this entegrated dataset in context of competitor behaviours and functionality of interactions.
+
+[#Add]The main advantage of the method, ceRNAnetsim package, is providing various functions which contain optionally arguments.
+So, users can run the functions in accordance with their own data.
+For example, the functions can work without interaction factors or with a simple interaction factor that is specified by user.
+Furthermore, it can also be operated according to a different parameter(s) provided by the user.In this study, we have preferred to run model, shown in *Multifactorial calculations in miRNA:target network* section.
 

### 2019-09-10 article.tex
news to article

@@
   \postdate{\par}
-    \date{21 Ağustos 2019}
+    \date{10 Eylül 2019}
 
@@
 activities of miRNAs vary depending on many factors that are significant
-to microRNA:target interactions. These factors include microRNA:target
-binding energy, binding location in target sequence, base pairing types
-between miRNA and target, abundance of miRNAs and targets (Grimson et
-al. \protect\hyperlink{ref-grimson_microrna_2007}{2007}). Binding
-energies of miRNA:target complexes vary based on nucleotide context and
-determine folding stability of complex (Cao and Chen
+to miRNA:target interactions. These factors include miRNA:target binding
+energy, binding location in target sequence, base pairing types between
+miRNA and target, abundance of miRNAs and targets (Grimson et al.
+\protect\hyperlink{ref-grimson_microrna_2007}{2007}). Binding energies
+of miRNA:target complexes vary based on nucleotide context and determine
+folding stability of complex (Cao and Chen
 \protect\hyperlink{ref-cao_predicting_2012}{2012}). It has been
@@
 some researchers have reported that seed sequence of miRNA can have
-mismatches or bulged/wobble nucleotides and may locate in region other
-than 5'end of miRNAs (Hafner et al.
+mismatches or bulged/wobble nucleotides (Chi, Hannon, and Darnell
+\protect\hyperlink{ref-chi2012alternative}{2012}) and may locate in
+region other than 5'end of miRNAs (Hafner et al.
 \protect\hyperlink{ref-hafner_transcriptome-wide_2010}{2010}; Helwak et
@@
 It has been observed that when abundance of one of the targets of
-miR-122 was increased, the other target expression also slightly
+miR-122 was increased, the other targets' expression also slightly
 increased as a result of decreasing repressive activity of miR-122 on
@@
 Henderson applied the model that handles miRNA:target direct and
-indirect interactions via common miRNA of gene and target of miRNA, at
-bipartite network. It has been demonstrated that all miRNAs and targets
-in the network can interact with each other through common miRNAs and
-genes, without interaction between the same type of nodes (Robinson and
-Henderson
+indirect interactions via common targeting miRNA of genes and target of
+miRNAs, at bipartite network. It has been demonstrated that all miRNAs
+and targets in the network can interact with each other through common
+miRNAs and genes, without interaction between the same type of nodes
+(Robinson and Henderson
 \protect\hyperlink{ref-robinson_modelling_2018}{2018}).Associated genes
@@
 Initially, the network is assumed in steady-state (\autoref{fig1}a)
-condition and needs a trigger for initiating calculations. The trigger
-can be a change in expression level of one or more genes
+condition and needs least one trigger for initiating calculations. The
+trigger can be a change in expression level of one or more genes
 (\autoref{fig1}b). After a trigger, the network undergoes iterative
@@
 neighborhood is recalculated (\autoref{fig1}c). Based on new miRNA
-distribution, expression level of each node (i.e.~ceRNA) is updated. Due
-to nodes which are targeted by multiple miRNAs, the change in one
-neighborhood spreads to other neighborhoods (\autoref{fig1}d),
-consequently have potential to effect whole network due to ``ripple
-effect''.
+distribution, expression level of each node (i.e.~ceRNA) is updated.
+This update results in change of expression value of common target (G4)
+in system. In this case, the common target acts as trigger for the other
+group (the targets interacted with M2 miRNA) in the network. Due to
+common targeted elements, the change in one neighborhood spreads to
+other neighborhoods (\autoref{fig1}d), consequently have potential to
+effect whole network due to ``ripple effect''.
 
@@
 miRNA and genes in tumor and normal tissue of single patient are
-retrieved from
+retrieved from The Cancer Genome Atlas
 \href{https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga}{TCGA}(\#TODO).
@@
 of miRNA repression activity has been based on miRNA and target
-abundance in various researches(Arvey et al.
+abundance in various researches (Arvey et al.
 \protect\hyperlink{ref-arvey_target_2010}{2010}; Denzler et al.
@@
 steady-state. So it has been observed that the regulatory activities of
-miRNA efficiencies on the targets are weak after the increase of miRNA
-target G2. When model is triggered with two fold increase in expression
-level of common target (G4), the changes of other gene expressions have
-observed more prominent \href{}{}. Furthermore, change in expression
-level of target gene that has strong miRNA repression efficiency
-resulted in evident perturbation in network. On the other hand, it was
-observed that Gene2 was weakly affected from change of Gene4 expression
-because of its weak interaction factors. Since perturbation efficiency
-of each gene is different, we have developed a function which screens
-each gene in the network for their perturbation efficiencies.
-\textbf{(see details in Supplementary File(fig1\_2app.Rmd)}. On the
-other hand, when we applied the method on the minimal dataset, Gene4 has
-been found to be the most efficient element in terms of number of
-perturbed elements and miRNA2 (M2) has been found to be causing the
-highest mean in expression changes. {[}{]}{[}\#TODO2{]}
+miRNA on the targets are weak after the increase of miRNA target G2.
+When model is triggered with two fold increase in expression level of
+common target (G4), the changes of other gene expressions have observed
+more prominent \href{}{}. Furthermore, change in expression level of
+target gene that has strong miRNA repression efficiency resulted in
+evident perturbation in network. On the other hand, it was observed that
+Gene2 was weakly affected from change of Gene4 expression because of its
+weak interaction factors. Since perturbation efficiency of each gene is
+different, we have developed a function which screens each gene in the
+network for their perturbation efficiencies. \textbf{(see details in
+Supplementary File(fig1\_2app.Rmd)}. On the other hand, when we applied
+the method on the minimal dataset, Gene4 has been found to be the most
+efficient element in terms of number of perturbed elements and miRNA2
+(M2) has been found to be causing the highest mean in expression
+changes. {[}{]}{[}\#TODO2{]}
 
-\href{}{} We used our model to simulate a real dataset which contains
-thousands of genes and hundreds of miRNAs. Our model can successfully
-simulate perturbations in such large network despite complex behaviors
-and struggle to reach steady-state \textbf{(see details in Supplementary
+We used our model to simulate a real dataset which contains thousands of
+genes and hundreds of miRNAs. Our model can successfully simulate
+perturbations in such large network despite complex behaviors and
+struggle to reach steady-state \textbf{(see details in Supplementary
 File(TCGA\_E9-A1N5\_article.Rmd)}. Simulations show that change in
@@
 throughput experimental dataset about miR-92a depletion (Helwak et al.
-\protect\hyperlink{ref-helwak_mapping_2013}{2013}). In this study, it
-was demonstrated that distant ceRNAs can interact with each other via
-indirect links, and the interactions are effected depending on distance
-between ceRNAs or topological features of network (Nitzan et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}). It was demonstrated
+that distant ceRNAs can interact with each other via indirect links, and
+the interactions are effected depending on distance between ceRNAs or
+topological features of network (Nitzan et al.
 \protect\hyperlink{ref-nitzan_interactions_2014}{2014}). More recently,
@@
 observations that a miRNA can exhibit strong functionality to a target
-but may not against an other, the authors have concluded that common
+but may not against an other, the authors have concluded that existing
 miRNA based approach may not be suitable for understanding regulations
 of ceRNA interactions.
+
+{[}\#Add{]}More recently, Silveira et al.~have developed miRmapper
+package for using with R programme (Silveira et al.
+\protect\hyperlink{ref-da2018mirmapper}{2018}). Researchers have
+utilized the adjacency matrix to associate miRNAs using differentially
+expressed genes and found similarities of miRNAs with this approach.
+Authors who utilized the approach in dataset of bladder cancer cell line
+have identified significant miRNAs in network, and miRNAs which work
+synergistically.
+
+{[}\#Add{]} Today, there are many of tools that provide predicted or
+experimentally validated miRNA:target interaction dataset. The datasets
+that contain weak, proteomic analysis like western blot, and strong,
+crosslinking and immunoprecipitation CLiP, experimentally evidence are
+more reliable sources than prediction datasets. While in experimentally
+weak supported datasets it is not known that indirect or direct target
+of miRNAs in cell, high-throughput methods ensure handling of AGO
+(Argounate Protein) interacted miRNAs. For this reason we have preferred
+to utilize high-throughput datasets in our application. Especially, we
+collected datasets that depend on chimeric (Helwak et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
+\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}) reading of miRNA:
+target pairs from among these datasets because they contain exact
+complementary seed regions.
+
+{[}\#Add{]} Although these experimental sources include exact binding
+sites, they do not provides which miRNA is functional or not (Liu and
+Wang \protect\hyperlink{ref-liu2019prediction}{2019}). We have used the
+features of bound miRNA:target pairs and scaled these parameters inside
+groups that is consisted according to, ceRNA (Competing Endogenous RNA),
+genes targeted by the same miRNA. In scaling process, we specified the
+interaction factors as numeric values in accordance with previous
+studies. Energy values in miRNA:target pairs are represented by
+high-throughput studies (Helwak et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
+\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}) which are utilized
+in this study. On the other hand, we have specified the other
+interaction factors, seed type and location of binding region on the
+target, as numeric values based on the previous studies.(Grimson et al.
+\protect\hyperlink{ref-grimson_microrna_2007}{2007}) have compared the
+seed types' effect on target repression with few miRNA had canonical
+seed pairing in their study. Additionally, (Bartel
+\protect\hyperlink{ref-bartel_micrornas:_2009}{2009}) and (Betel et al.
+\protect\hyperlink{ref-betel2010comprehensive}{2010}) have studied on
+functional and non-functional seed interactions. Based on results of
+these studies we have arranged seed types of miRNA:target interactions
+as numeric values. We also have redefined location of binding region on
+the target as numeric values, based on studies of (Hausser et al.
+\protect\hyperlink{ref-hausser_analysis_2013}{2013}) and (Helwak et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}). With this process,
+we have handled this entegrated dataset in context of competitor
+behaviours and functionality of interactions.
+
+{[}\#Add{]}The main advantage of the method, ceRNAnetsim package, is
+providing various functions which contain optionally arguments. So,
+users can run the functions in accordance with their own data. For
+example, the functions can work without interaction factors or with a
+simple interaction factor that is specified by user. Furthermore, it can
+also be operated according to a different parameter(s) provided by the
+user.In this study, we have preferred to run model, shown in
+\emph{Multifactorial calculations in miRNA:target network} section.
+
+{[}\#TODO answer{]} We have observed that the results of simulation with
+ABCC1 gene, \emph{Multidrug resistance-associated protein 1}, which is
+one of the most significant factor to develop resistance against
+chemotherapotic agents(Atalay, Demirkazik, and Gunduz
+\protect\hyperlink{ref-atalay2008role}{2008}; Lu et al.
+\protect\hyperlink{ref-lu2015microrna}{2015}; Atalay et al.
+\protect\hyperlink{ref-atalay2006multidrug}{2006}) is not convenient
+with tumor tissue expression values, when we compared the tumor gene
+expression dataset retrieved from
+\href{https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga}{TCGA}
+and simulation results obtained via our approach. Simulation results
+were not expected to be consistent with tumor tissue expression values
+because large number of abnormal regulations were observed in these
+tissues. So, it might not said that a single gene is responsible for
+entire regulations in cancer cells. Besides, the all other factors such
+as up/down regulation of miRNAs or other genes were ignored. Therefore,
+it would be more useful to test our approach with an ideal dataset that
+includes entire informations of gene expression values before and after
+the regulation of a known gene at existing tissue, and miRNA expression
+values of this tissue at initial conditions.
 
@@
 
+\leavevmode\hypertarget{ref-atalay2006multidrug}{}%
+Atalay, Can, Ismet Deliloglu Gurhan, Cigdem Irkkan, and Ufuk Gunduz.
+2006. ``Multidrug Resistance in Locally Advanced Breast Cancer.''
+\emph{Tumor Biology} 27 (6): 309--18.
+
+\leavevmode\hypertarget{ref-atalay2008role}{}%
+Atalay, C, A Demirkazik, and U Gunduz. 2008. ``Role of Abcb1 and Abcc1
+Gene Induction on Survival in Locally Advanced Breast Cancer.''
+\emph{Journal of Chemotherapy} 20 (6): 734--39.
+
 \leavevmode\hypertarget{ref-bartel_micrornas_2004}{}%
@@
 \url{https://doi.org/10.1016/j.cell.2009.01.002}.
+
+\leavevmode\hypertarget{ref-betel2010comprehensive}{}%
+Betel, Doron, Anjali Koppal, Phaedra Agius, Chris Sander, and Christina
+Leslie. 2010. ``Comprehensive Modeling of microRNA Targets Predicts
+Functional Non-Conserved and Non-Canonical Sites.'' \emph{Genome
+Biology} 11 (8): R90.
 
@@
 \url{https://doi.org/10.1073/pnas.1305322110}.
+
+\leavevmode\hypertarget{ref-chi2012alternative}{}%
+Chi, Sung Wook, Gregory J Hannon, and Robert B Darnell. 2012. ``An
+Alternative Mode of microRNA Target Recognition.'' \emph{Nature
+Structural \& Molecular Biology} 19 (3): 321.
 
@@
 
+\leavevmode\hypertarget{ref-liu2019prediction}{}%
+Liu, Weijun, and Xiaowei Wang. 2019. ``Prediction of Functional microRNA
+Targets by Integrative Modeling of microRNA Binding and Target
+Expression Data.'' \emph{Genome Biology} 20 (1): 18.
+
+\leavevmode\hypertarget{ref-lu2015microrna}{}%
+Lu, Lin, Fang Ju, Hui Zhao, and Xuezhen Ma. 2015. ``MicroRNA-134
+Modulates Resistance to Doxorubicin in Human Breast Cancer Cells by
+Downregulating Abcc1.'' \emph{Biotechnology Letters} 37 (12): 2387--94.
+
 \leavevmode\hypertarget{ref-markus_list_sponge_2017}{}%
@@
 
+\leavevmode\hypertarget{ref-da2018mirmapper}{}%
+Silveira, Willian da, Ludivine Renaud, Jonathan Simpson, William Glen,
+Edward Hazard, Dongjun Chung, and Gary Hardiman. 2018. ``MiRmapper: A
+Tool for Interpretation of miRNA--mRNA Interaction Networks.''
+\emph{Genes} 9 (9): 458.
+
 \leavevmode\hypertarget{ref-tay_multilayered_2014}{}%

### 2019-09-10 deneme.Rmd
news to article

---
title: "Network based multifactorial modelling microRNA-target interaction."
author: 
- Selcen Ari^[Department of Bioengineering, Yildiz Technical University, Istanbul, Turkey]
- Alper Yilmaz^[Department of Bioengineering, Yildiz Technical University, Istanbul, Turkey]
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: 
    keep_tex: true
  html_document: default
---

```{r echo= FALSE, results= 'asis'}
install.packages("xtable")
library(xtable)
data("minsamp")

print(xtable::xtable(minsamp, caption="Expression values of elements and interaction factors of miRNA:target interacions", label = "tab:one", 
             table.placement = "l!", 
             caption.placement = "top"), comment= FALSE)
```

\begin{table}[ht]
\centering
\caption{Expression values of elements and interaction factors of miRNA:target interacions} 
\begin{tabular}{rllrrrrr}
  \hline
 & competing & miRNA & Competing\_expression & miRNA\_expression & seed\_type & region & energy \\ 
  \hline
1 & Gene1 & Mir1 & 10000.00 & 1000.00 & 0.43 & 0.30 & -20.00 \\ 
  2 & Gene2 & Mir1 & 10000.00 & 1000.00 & 0.43 & 0.01 & -15.00 \\ 
  3 & Gene3 & Mir1 & 5000.00 & 1000.00 & 0.32 & 0.40 & -14.00 \\ 
  4 & Gene4 & Mir1 & 10000.00 & 1000.00 & 0.23 & 0.50 & -10.00 \\ 
  5 & Gene4 & Mir2 & 10000.00 & 2000.00 & 0.35 & 0.90 & -12.00 \\ 
  6 & Gene5 & Mir2 & 5000.00 & 2000.00 & 0.05 & 0.40 & -11.00 \\ 
  7 & Gene6 & Mir2 & 10000.00 & 2000.00 & 0.01 & 0.80 & -25.00 \\ 
   \hline
\end{tabular}

\end{table}

### 2019-09-12 article.Rmd
news to article
edited with selcen

@@
 Our findings emphasized importance of miRNA:target ratios being crucial, as reported by previous studies.
-We have showed that the competing elements which have the same or close expression values may not be affected equally from the perturbation because of repression functionality depended on interaction factors of miRNA target pairs (answer: Hocam etkileşim faktörlerinden dolayı ekspresyonları yakın olsa bile baskılanma aktitesinin aynı sergilenmeyeceğini ifade etmeye çalıştım.
-Bu şekilde anlaşılıyor mu?).
+We have showed that the competing elements which have the same or close expression values may not be affected equally from the perturbation because of repression functionality depended on interaction factors of miRNA target pairs.
 We applied the model to real sample consisting of breast cancer gene and miRNA expression dataset and experimental miRNA:target interaction dataset all generated via high throughput sequencing methods.
@@
 It has been demonstrated that the binding energy between miRNA and target indicates stability or affinity of complex [@helwak_mapping_2013] and does not directly determine repressive activity of miRNA [@cao_predicting_2012].
-Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5’end bind to specific sequence located in 3’UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005].
+Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5'end bind to specific sequence located in 3'UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005].
 In recent studies, it has been shown that miRNAs can interact with targets via sequences located in regions such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013; @moore_mirnatarget_2015].
-These studies also showed that binding location could indicate functionality of miRNA:target interaction or be effective on abundance of targets(TOFIX abundance etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı?
-answer: evet hocam anlaşılmıyorsa değiştirebilirim).
+These studies also showed that  binding location either indicate functionality of miRNA:target interaction or affect abundance of targets.
 It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is perfectly complementary with targets, seed region at the 5’ end of miRNAs [@bartel_micrornas:_2009; @grimson_microrna_2007].
@@
 ![Schematic presentation of mechanism of network based model.
-a) In steady state, miRNAs (M: gray triangle) repress targets (G: circle coconut) according to proportion of their targets' expression.
-b) Two fold increase in transcript level of Gene2 (G2) acts as a trigger (shown with red in figure).
+a) In steady state, miRNAs (M: triangles) repress targets (G: circles) according to proportion of target expression level.
+b) Two fold increase in transcript level of Gene2 (G2) acts as a trigger (shown in red).
 c) Distribution of miRNA1 (M1) changes.
-d) The change at expression (shown with light red because minor changes are occured on their expression values.) of common target effects changes of proportional distribution of miRNA2 (M2).
-Expression values are rounded to integers for simplicity](Fig1.jpg){#fig1}
-
-[#FIX]: <> "caption needs more text, every color should be explained"
-[answer]: <> "explained in caption."
+d) The change at expression of common target effects changes of proportional distribution of miRNA2 (M2).
+Expression values are rounded to integers for simplicity.
+Gray values on edges indicate initial distribution, red values indicate increase and blue values indicate decrease in distribution.
+Shades of circles indicate different levels of increase.](Fig1.jpg){#fig1}
 
@@
 
-The repression efficiency of a miRNA on the individual target ($Eff_{gi}$) is calculated according to equation \eqref{eq:1}; where miRNA expression ($C_m$) in local neighborhood is distributed among targets using individual gene expression levels ($C_{gi}$) (#FIX isn't $C_g$ gene expressions in the group is equal to total of $C_{gi}$ in group?
-answer: its my false.
-now its corect.).
+The repression efficiency of a miRNA on the individual target ($Eff_{gi}$) is calculated according to equation \eqref{eq:1}; where miRNA expression ($C_m$) in local neighborhood is distributed among targets using individual gene expression levels ($C_{gi}$).
 For the genes targeted by multiple miRNAs, cooperative activity of miRNAs on a target gene, $R$, is calculated by summing repression activity of each miRNA (Equation \eqref{eq:2}).
@@
 We classified factors into two categories.
-Factors effecting binding determine interaction between miRNA and target and they alter amount of miRNA sequestered to target.
-Degradation efficiency factors assign amount of degradated target amount in miRNA:target pairs.(#FIX fix this sentence: answer:).
+Binding factors determine interaction between miRNA and target and they alter amount of miRNA sequestered to target.
+Efficiency factors dictate degradation efficieny of miRNA on its target and define amount of target to be degraded.
 In other words, binding factors exert their influence before or during binding, efficiency factors exert their influence after binding.
 In the literature, binding free energy [@cao_predicting_2012 ; @helwak_mapping_2013] and seed type [@werfel_preferential_2017] in miRNA:target interactions are described as factors effecting binding affinity.
-Efficiency factors determine how many of miRNA:target complexes will result in inhibition and binding region on target drastically effect miRNA degradation efficiency [@hausser_analysis_2013; @helwak_mapping_2013].
-Both binding and efficiency factors are normalized to their maximum values and scaled to [0,1] interval.
-The normalized values of factors take into account to determine binding activity and miRNA efficiency on targets (\autoref{fig2}).
+Also, binding region (e.g, 5'UTR, 3'UTR, CDS) was shown to drastically effect miRNA degradation efficiency [@hausser_analysis_2013; @helwak_mapping_2013].
+Since binding or degradation efficiency from literature have various and broad ranges, both binding and efficiency factors are normalized to their maximum values and scaled to [0,1] interval.
+The normalized values of factors are used to determine binding activity and miRNA efficiency on targets (\autoref{fig2}).
 Binding affinities (activity, $Eff$) of miRNAs on each individual gene are calculated as shown in equation \eqref{eq:3}; $C_m$, miRNA expression in the group; $C_{gi}$, individual gene expression; gi, individual gene (\autoref{fig2}c).
@@
 ![Calculations to determine of miRNA binding and repression efficiency.
-$G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE^\prime$, normalized values of seed type efficiency coefficient; $RE^\prime$, normalized values of region efficiency coefficient; $E^\prime$, normalized values of energy coefficient.](fig2.jpg){#fig2}
+$G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE^\prime$, normalized values of seed type efficiency coefficient; $RE^\prime$, normalized values of region efficiency coefficient; $E^\prime$, normalized values of energy coefficient.
+Numbering on edges match the pair order in \autoref{tab:one}](fig2.jpg){#fig2}
 
-\iffalse
-%% (#TODO the equation is not considering binding factors, right?. We should explain the calculation made in  simulate function  here, in simple sentences. the function looks like this: 
-%%    input_graph%>%
-%%      tidygraph::activate(edges)%>%
-%%      group_by(to)%>%
-%%      mutate(mirna_count_per_comp = mirna_count_current*comp_count_current*afff_factor/sum(comp_count_current*afff_factor), mirna_count_per_comp = ifelse(is.na(mirna_count_per_comp), 0, mirna_count_per_comp))%>%
-%%      ungroup()%>%
-%%      mutate(effect_pre = effect_current, effect_current = mirna_count_per_comp*degg_factor)%>%
-%%      group_by(from)%>%
-%%      mutate(comp_count_pre = ifelse(comp_count_current< 0, 1, comp_count_current), comp_count_current = comp_count_pre - (sum(effect_current)-sum(effect_pre)), comp_count_current = ifelse(comp_count_current< 0, 1, comp_count_current))%>%
-%%      ungroup()%>%
-%%      mutate(comp_count_list = pmap(list(comp_count_list, comp_count_current), c), effect_list = pmap(list(effect_list, effect_current), c), mirna_count_list = pmap(list(mirna_count_list, mirna_count_current), c))
-%% )
-\fi
-
-<answer> I shown in figure and explained at following. I didn't fully understand whether it is enough or not.
-
-After miRNA binds to its target, but might not repress to bound target.
-The functionality of bound miRNA on target depends on efficiency factors like region that is binding sequence of miRNA on its target.
+Not all miRNA:target binding events result in degradation of target.
+The degradation of target by bound miRNA depends on efficiency factors such as binding region.
 Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:4} (\autoref{fig2}d); $RE^\prime_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
@@
 We have applied our model in a real dataset for which experimental measurements of various factors were available.
-Expression levels of miRNA and genes in tumor and normal tissue of single  patient are retrieved from The Cancer Genome Atlas [TCGA](https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga)(#TODO).
+Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network: https://www.cancer.gov/tcga.
 High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015].
-We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from ...
-(#FIX name of database and its citation) answer->suplementary datasets of [@helwak_mapping_2013; @moore_mirnatarget_2015].
-Detailed description of network construction and its code is available in Supplementary data (#TODO burada link nasıl olacak, word dosyası veya PDF dosyası ismi mi vermemiz mi  gerekiyor, yoksa sadece Sup Data denilmesi yeterli mi?
-ilgili dosya: TCGA_E9-A1N5_article.Rmd).
+We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from published datasets [@helwak_mapping_2013; @moore_mirnatarget_2015].
+Detailed description of network construction and its code is available in Supplementary Materials and Methods.
 ABCC1 gene, over-expressed in tumor tissue, was selected as trigger for simulation of integrated dataset.
@@
 It was observed that expression levels of primary neighborhoods which do not interact with another miRNA of the trigger gene change in relation to their expression.
-However, expression level change of common target (G4) in network has been observed different from G1 that has the same expression value with G4 target at initial conditions because more than one miRNA repress this target (\autoref{fig1}d) (#FIX last sentence needs some clarification, after "but cooperative...").
+However, changes in expression levels of G4 and G1 differ although they have same initial conditions.
+This is due to fact that G4 is targeted by two miRNAs, M1 and M2 (\autoref{fig1}d).
 
@@
 It's important to note that the changes in gene expression levels will have more pronounced effect if miRNA:target ratio is high, i.e., more miRNA available per target, which was reported in previous findings [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
-[][#FIX]
-
-[#FIX]: <> (following 4 sentences can be deleted, not so clear.
-OR we can summarize it one-two sentences.
-When the simulation was run in sample dataset that includes lower target abundance, the first response of primary neighborhood of the same group was determined as similar with in system that has higher total target abundance.
-In overall state in the end of simulation, we have established that all regulations more prominent when compared with low miRNA:target ratio dataset.)
 
@@
 
-### Competing Endogenous RNA:Target networks based on interaction factors
-
-[#FIX]: <> "title should be more informative"-> answer.
-
+### ceRNA:Target networks based on interaction factors
 Earlier studies reported that miRNA regulatory interactions are affected by different parameters.
@@
 
-[#FIX2]: <> "we should compare and contrast Figure 1a and figure 3a, the reader should see how affinity effects distrbution G4:M2, and how efficiency is effected without change in affnity M1:G3"
-[answer2]: <> "I explained the mechanism, but i did not understand  how efficiency is effected without change in affnity M1:G3"
+[#TODO]: <> "figure1, 2 ve 3 uzerinden farkli dagilim ve degradasyon açıklaması buraya. G2 uzerinden olmasa iyi olur"
+
 It is considered that entire miRNAs in the system affect targets according to target:total target ratio \autoref{fig1}a, when interaction factors did not take into account.
@@
 So it has been observed that the regulatory activities of miRNA on the targets are weak after the increase of miRNA target G2.
-When model is triggered with two fold increase in expression level of common target (G4), the changes of other gene expressions have observed more prominent [][#TODO].
+When model is triggered with two fold increase in expression level of common target (G4), more prominent changes were observed in gene expression levels in network compared to changes caused by G2 increase.
 Furthermore, change in expression level of target gene that has strong miRNA repression efficiency resulted in evident perturbation in network.
@@
 On the other hand, when we applied the method on the minimal dataset, Gene4 has been found to be the most efficient element in terms of number of perturbed elements and miRNA2 (M2) has been found to be causing the highest mean in expression changes.
-[][#TODO2]
 
-[#FIX]: <> "didnt we perform upregulation of Gene2, why did we say Gene4 here?" 
-[answer]: <> "Because, in fact, We want to show that triggering element is significant in the same dataset, in the same conditions. Maybe I must add a conjuction here."
+### Simulation of large and realistic network
 
-We used our model to simulate a real dataset which contains thousands of genes and hundreds of miRNAs.
+Apart from tools predicting miRNA:target pairs, there are databases that provide  experimentally validated miRNA:target interaction datasets.
+The reliability in these datasets range from weak (e.g, western blot) to strong (e.g, CLiP, CLASH).
+High-throughput methods ensure handling of AGO (Argounate Protein) interacted miRNAs (CLiP) and their targets as well (CLEAR-CLiP, CLASH).
+We constructed a large ceRNA network utilizing high-throughput datasets containing chimeric reading of miRNA:target pairs [@helwak_mapping_2013; @moore_mirnatarget_2015] which expose exact complementary seed regions.
+
+Although these experimental sources include exact binding sites, they do not imply that each miRNA binding event is functional [@liu2019prediction].
+Seed type, energy, binding region affects functionality of miRNA drastically.
+In order to construct realistic network, we incorporated such factors in the network.
+Energy  values in miRNA:target pairs represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015], seed type [@grimson_microrna_2007; @bartel_micrornas:_2009; @betel2010comprehensive] and binding region effect [@hausser_analysis_2013; @helwak_mapping_2013] data were collected and integrated  into network.
+
+The network was finalized by integrating gene and miRNA expression levels from TCGA determined by RNA-Seq.
 Our  model can successfully simulate perturbations in such large network despite complex behaviors and struggle to reach steady-state **(see details in Supplementary File(TCGA_E9-A1N5_article.Rmd)**.
@@
 These observations are in accordance with competing endogenous RNA hypothesis where genes targeted with many common miRNAs subsequently transmit perturbation to neighboring groups.
+
+[#TODO] We have observed that the results of simulation with ABCC1 gene, *Multidrug resistance-associated protein 1*, which is one of the most significant factors to develop resistance against chemotherapotic agents[@atalay2008role; @lu2015microrna; @atalay2006multidrug] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach.
+Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues.
+So, it might not said that a single gene is responsible for entire regulations in cancer cells.
+Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored.
+Therefore, it would be more useful to test our approach with an ideal dataset that includes entire informations of gene expression values before and after the regulation of a known gene at existing tissue, and miRNA expression values of this tissue at initial conditions.
+
+[#TODO]: <> "run perturbation screening on tcga set, find a gene then couple sentences about that gene"
+
+## Discussion
 
@@
 Based on the observations that a miRNA can exhibit strong functionality to a target but may not against an other, the authors have concluded that existing miRNA based approach may not be suitable for understanding regulations of ceRNA interactions.
-
-[#Add]More recently, Silveira et al.
-have developed miRmapper package for using with R programme [@da2018mirmapper].
-Researchers have utilized the adjacency matrix to associate miRNAs using differentially expressed genes and found similarities of miRNAs with this approach.
-Authors who utilized the approach in dataset of bladder cancer cell line have identified significant miRNAs in network, and miRNAs which work synergistically.
-
-[#Add] Today, there are many of tools that provide predicted or experimentally validated miRNA:target interaction dataset.
-The datasets that contain weak, proteomic analysis like western blot, and strong, crosslinking and immunoprecipitation CLiP, experimentally evidence are more reliable sources than prediction datasets.
-While in experimentally weak supported datasets it is not known that indirect or direct target of miRNAs in cell, high-throughput methods ensure handling of AGO (Argounate Protein) interacted miRNAs.
-For this reason we have preferred to utilize high-throughput datasets in our application.
-Especially, we collected datasets that depend on chimeric [@helwak_mapping_2013; @moore_mirnatarget_2015] reading of miRNA: target pairs from among these datasets because they contain exact complementary seed regions.
-
-[#Add] Although these experimental sources include exact binding sites, they do not provides which miRNA is functional or not [@liu2019prediction].
-We have used the features of bound miRNA:target pairs and scaled these parameters inside groups that is consisted according to, ceRNA (Competing Endogenous RNA), genes targeted by the same miRNA.
-In scaling process, we specified the interaction factors as numeric values in accordance with previous studies.
-Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study.
-On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study.
-Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions.
-Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values.
-We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
-With this process, we have handled this entegrated dataset in context of competitor behaviours and functionality of interactions.
-
-[#Add]The main advantage of the method, ceRNAnetsim package, is providing various functions which contain optionally arguments.
-So, users can run the functions in accordance with their own data.
-For example, the functions can work without interaction factors or with a simple interaction factor that is specified by user.
-Furthermore, it can also be operated according to a different parameter(s) provided by the user.In this study, we have preferred to run model, shown in *Multifactorial calculations in miRNA:target network* section.
-
-[#TODO answer] We have observed that the results of simulation with ABCC1 gene, *Multidrug resistance-associated protein 1*, which is one of the most significant factor to develop resistance against chemotherapotic agents[@atalay2008role; @lu2015microrna; @atalay2006multidrug] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from [TCGA](https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga) and simulation results obtained via our approach.
-Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues.
-So, it might not said that a single gene is responsible for entire regulations in cancer cells.
-Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored.
-Therefore, it would be more useful to test our approach with an ideal dataset that includes entire informations of gene expression values before and after the regulation of a known gene at existing tissue, and miRNA expression values of this tissue at initial conditions.
-
-[#TODO]: <> "we can talk about BRCA results here: FIX at previous pragraph"
+Additionally, miRmapper package [@da2018mirmapper] utilizes an adjacency matrix to associate miRNAs using differentially expressed genes and identifies significant miRNAs in network.
 
@@
 Additionally, other regulation parameters such as gene-gene interactions and transcription factors are ignored but the network structure is flexible and can integrate additional regulation elements as they become available.
-In the future, with developments in experimental techniques about miRNAs:target interaction dynamica, more accurate results can be obtained from our model resulting in better understanding and predictions of abnormal regulations and pathways underlying diseases or conditions.
+
+## Conclusion
+
+The main advantage of ceRNAnetsim package is providing various functions which can be run in accordance with user data.
+For example, the functions can work without interaction factors or with a simple interaction factor that is specified by user.
+Furthermore, it can also be operated according to a different parameter(s) provided by the user.In this study, we have preferred to run model, shown in *Multifactorial calculations in miRNA:target network* section.
+
+In the future, with developments in experimental techniques about miRNAs:target interaction dynamics, more accurate results can be obtained from our model.
+Also, our model can integrate emerging miRNA effectors such as circRNAs.
+Consequently, our package may lead to better understanding and predictions of abnormal regulations and pathways underlying diseases or conditions.
 

### 2019-09-15 Suplementary_mat1.Rmd
last meeting

---
title: "Supplementary File 1"
author: "Selcen Ari"
date: "7 Nisan 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
bibliography: ref.bib
---

#1. Defined functions for ceRNA models and workflow of method
We defined the functions that can be used with R programming. Briefly, these functions process a given miRNA:gene dataset and convert to graph object. All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section. The functions and steps of approach are explained as following (Figure 1) :

**Convertion of dataset**: priming_graph function processes the given dataset that includes competing elements in first variable and repressive element in second variable. If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa. The formulations that are given in equations (1-4) are performed in this function. This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount. It is assumed that the initial target amounts in the dataset is observed after the reppressive activity of miRNAs in steady-state.

![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow1.png){#fig1}

**Transition of variables in graph**: In the previous step, the calculations are performed in the edge variables of the graph object. However, the graph object allows to use node variables, while the node features are handled to the graph. In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.

**Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficieny coefficients are calculated according to this acceptance. In the network that is found in steady-state conditions, the change is applied to the graph object for disturbtion of steady-state. To provide the disturbtion in the network the workflow offer two methods: update_variables and update_how. The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables. The second option, the amount of the given node name in update_how function can be changed according to “how” argument.

**Updating current values of variables**: After variables updating in edge varibles, these are carried to node variables. Current and previous values of variables are stored as node variables with update_variables function.

**Simulation of competing behavior of targets**: After the change in the steady-state conditions, the network elements try to gain steady-state again. This process progresses as repeating of regulations after the spreading the changes in the network. In this step, simulation of regulations according to given cycle count in simulate function is applied. After each simulation cycle, the miRNA repression values are re-calculated and the current values of competing elements are found and saved. The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node data.

The node elements in the dataset are handled as two type; repressives (miRNAs) and competings (targets). It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used. If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user. On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger. The functions that are used in the approach are developed with R programming so as can be used with other packages. These are can be found in the github repository [ceRNAnetsim github page](https://github.com/selcenari/ceRNAnetsim) and improved with contributions of others.

```{r, message=FALSE, warning=FALSE}
#install.packages("devtools")
#devtools::install_github("selcenari/ceRNAnetsim")
library(ceRNAnetsim)
```

+ load *minsamp* data

```{r}
data("minsamp")

minsamp
```

|competing| miRNA | Competing_expression | miRNA_expression | seed_type | region | energy
|-----------|------------|----------|----------------|---------|-----------|----------|
| `Gene1`   | `Mir1` | 10000 | 1000 | 0.43 | 0.30 | -20 |
| `Gene2`   | `Mir1` | 10000 | 1000 | 0.43 | 0.01 | -15 |
| `Gene3`   | `Mir1` | 5000  | 1000 | 0.32 | 0.40 | -14 |
| `Gene4`   | `Mir1` | 10000 | 1000 | 0.23 | 0.50 | -10 |
| `Gene4`   | `Mir2` | 10000 | 2000 | 0.35 | 0.90 | -12 |
| `Gene5`   | `Mir2` | 5000  | 2000 | 0.05 | 0.40 | -11 |
| `Gene6`   | `Mir2` | 10000 | 2000 | 0.01 | 0.80 | -25 |

## *minsamp* dataset analysis in lack of interaction factors.

Firstly, we have analysed minimal data without interaction factors between miRNA:target.

- 1. We have evaluated graph in the steady state conditions as followings:

```{r, message=FALSE, warning=FALSE}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset in steady-state conditions")
```

- 2. We have obtained graph after change on Gene2 expression as followings:

```{r, message=FALSE, warning=FALSE}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
  update_how("Gene2", 2)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors")
```

- 3. We have determined regulations after Gene2 Upregulation:

```{r, message=FALSE, warning=FALSE}

priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
  update_how("Gene2", 2)%>%
  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "Regulations after Gene2 Upregulation", 2)

```
Note that the regulations are colored according to expression changes of present and a previous value. So, it can be observed that whole gene expressions increase in comparison of initial steady-state. The overall regulations of gene expressions are as followings:

```{r, message=FALSE, warning=FALSE}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
  update_how("Gene2", 2)%>%
  simulate(2)%>%
  activate(edges)%>%
  as_tibble()%>%
  select(Competing_name,comp_count_list,effect_list)%>%
  unnest()
```

## *minsamp* dataset analysis with interaction factors.

We have made the same analysis in present of interaction factors.

```{r, message=FALSE, warning=FALSE}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset in steady-state conditions")

```

```{r, message=FALSE, warning=FALSE}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_how("Gene2", 2)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors")
```

```{r, message=FALSE, warning=FALSE}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_how("Gene2", 2)%>%
  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors",2)
```

When the graphs which were resulted from analyses were examined, it was observed that behaviours were same. But, when the results were analysed in terms of expression values, the regulation differences can be observed.

```{r, message=FALSE, warning=FALSE}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_how("Gene2", 2)%>%
  simulate(3)%>%
  activate(edges)%>%
  as_tibble()%>%
  select(Competing_name,comp_count_list,effect_list)%>%
  unnest()
```

## Common target perturbation in *minsamp* dataset.

There are hundreds of defined miRNAs for human, so this results in presence of common targets of miRNAs in cells. Therefore, we have analysed perturbation efficiency of common target in *minsamp* dataset.

```{r, message=FALSE, warning=FALSE}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_how("Gene4", 2)%>%
  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors",2)
```

The common target perturbation  (increasing to two fold at Gene4 expression in presence of interaction factors) resulted in more prominent efficiency at the same conditions (shown in following).

```{r, message=FALSE, warning=FALSE}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_how("Gene4", 2)%>%
  simulate(3)%>%
  activate(edges)%>%
  as_tibble()%>%
  select(Competing_name,comp_count_list,effect_list)%>%
  unnest()
```

## Determination of perturbation efficiencies efficiencies of elements in system.

```{r, message=FALSE, warning=FALSE}

priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)-> sample_graph

find_node_perturbation(sample_graph, how = 2, cycle = 3, limit = 0.1)

```

#2. Obtaining breast cancer dataset and integration

This section describes how to apply `ceRNAnetsim` package on a breast cancer patient miRNA:target interaction dataset. Before the approach, we obtained three datasets and combined them.

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# Install the packages which is necessary to obtain and process datasets.
install.packages("BiocManager")
BiocManager::install("SummarizedExperiment")
BiocManager::install("TCGAbiolinks")
library(TCGAbiolinks)
library(SummarizedExperiment)
#install devtools
devtools::install_github("selcenari/ceRNAnetsim")
library(ceRNAnetsim)
```

## 2.1 How to get gene expression counts of TCGA-E9-A1N5 patient.

We have obtained the gene expression values of patient using [TCGAbiolinks](https://www.bioconductor.org/packages/release/bioc/html/TCGAbiolinks.html) package from [Bioconductor](https://www.bioconductor.org/). For this process, we have followed the instructions of the package. `TCGAbiolinks` package provides to obtain data for whole number of given barcode(s) at once. But, we preferred to download them seperately to show datasets.

+ Obtain to gene expression counts of tumor tissue.

```{r,message= FALSE, warning=FALSE}

BCP_tumor <- GDCquery(project = "TCGA-BRCA",
                  data.category = "Transcriptome Profiling",
                  data.type = "Gene Expression Quantification", 
                  workflow.type = "HTSeq - Counts",
                  barcode = "TCGA-E9-A1N5-01A-11R-A14D-07")

GDCdownload(BCP_tumor)
BCPGE_tumor <- GDCprepare(BCP_tumor)

as.data.frame(assay(BCPGE_tumor))%>%
  mutate(ensembl_gene_id = rownames(.))%>%
  dplyr::inner_join(as.data.frame(rowData(BCPGE_tumor)), by = "ensembl_gene_id")%>%
  dplyr::select(ensembl_gene_id, external_gene_name, 1)-> TCGA_E9_A1N5_tumor

colnames(TCGA_E9_A1N5_tumor)[3] <- "GE_tumor"

head(TCGA_E9_A1N5_tumor)
```

+ Obtain to gene expression counts of normal tissue.

```{r, message= FALSE, warning=FALSE}
BCP_normal <- GDCquery(project = "TCGA-BRCA",
                  data.category = "Transcriptome Profiling",
                  data.type = "Gene Expression Quantification", 
                  workflow.type = "HTSeq - Counts",
                  barcode = "TCGA-E9-A1N5-11A-41R-A14D-07")
GDCdownload(BCP_normal)

BCPGE_normal <- GDCprepare(BCP_normal)

as.data.frame(assay(BCPGE_normal))%>%
  mutate(ensembl_gene_id = rownames(.))%>%
  dplyr::inner_join(as.data.frame(rowData(BCPGE_normal)), by = "ensembl_gene_id")%>%
  dplyr::select(ensembl_gene_id, external_gene_name, 1)-> TCGA_E9_A1N5_normal

colnames(TCGA_E9_A1N5_normal)[3] <- "GE_normal"

head(TCGA_E9_A1N5_normal)

```

```{r, echo=FALSE, message= FALSE, warning=FALSE}
#remove:
rm(BCPGE_normal, BCPGE_tumor, BCP_normal, BCP_tumor)
```

## 2.2 How to get miRNA expression counts of TCGA-E9-A1N5 patient.

We have used `TCGAbiolinks` package to obtain miRNA expression quantification. The query gives read count of miRNA as isoform chromosome coordination. The data also contains mature miRNA information. So, we processed data to attain -5p -3p isoform information using mirbase release21 dataset.

+ Get the mirbase id of mature miRNA:

We downloaded the mirbase release 21 dataset from [mirbase](http://www.mirbase.org/ftp.shtml) and processed the patient mirna expression datasets as following:

```{r, message= FALSE, warning=FALSE}
library(readr)
read_tsv("hsa_mirna.txt", comment = "#", col_names = FALSE)%>%
  dplyr::select(mirna_type= X3, definition = X9)%>%
  filter(!endsWith(mirna_type, "primary_transcript"))%>%
  tidyr::separate(definition, c("ID", "Alias", "Name", "Derivated"), sep = ";")%>%
  dplyr::select(Alias, Name)%>%
  tidyr::separate(Alias, c("trash1", "ID"), sep = "=")%>%
  tidyr::separate(Name, c("trash2", "Name"), sep = "=")%>%
  dplyr::select(-trash1, -trash2)-> mirbase_id_conv

head(mirbase_id_conv)
```

+ Obtain the miRNA expression of tumor tissue of patient:

```{r, message= FALSE, warning=FALSE}
BCP_mirnatumor <- GDCquery(project = "TCGA-BRCA",
                  data.category = "Transcriptome Profiling",
                  data.type = "Isoform Expression Quantification",
                  workflow.type = "BCGSC miRNA Profiling",
                  barcode = "TCGA-E9-A1N5-01A-11R-A14C-13")

GDCdownload(BCP_mirnatumor)

GDCprepare(BCP_mirnatumor)%>%
  as.data.frame()%>%
  dplyr::select(miRNA_ID, read_count, reads_per_million_miRNA_mapped, miRNA_region)%>%
  dplyr::filter(startsWith(miRNA_region, "mature"))%>%
  dplyr::mutate(mirbase_id =str_remove(miRNA_region, "mature,"))%>%
  dplyr::select(-miRNA_region)%>%
  dplyr::inner_join(mirbase_id_conv, by = c("mirbase_id"="ID"))%>%
  dplyr::select(miRNA_name = Name, read_count, reads_per_million_miRNA_mapped)%>%
  dplyr::group_by(miRNA_name)%>%
  mutate(read_count= sum(read_count), reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
  dplyr::ungroup()%>%
  distinct() -> BCPME_mirnatumor

head(BCPME_mirnatumor)
```

+ Obtain the miRNA expression of normal tissue of patient:

```{r, message= FALSE, warning=FALSE}
BCP_mirnanormal <- GDCquery(project = "TCGA-BRCA",
                  data.category = "Transcriptome Profiling",
                  data.type = "Isoform Expression Quantification",
                  workflow.type = "BCGSC miRNA Profiling",
                  barcode = "TCGA-E9-A1N5-11A-41R-A14C-13")

GDCdownload(BCP_mirnanormal)
# a616435d-0b69-48ac-813d-5d75ad9b85eb.mirbase21.isoforms.quantification.txt

GDCprepare(BCP_mirnanormal)%>%
  as.data.frame()%>%
  dplyr::select(miRNA_ID, read_count, reads_per_million_miRNA_mapped, miRNA_region)%>%
  dplyr::filter(startsWith(miRNA_region, "mature"))%>%
  dplyr::mutate(mirbase_id =str_remove(miRNA_region, "mature,"))%>%
  dplyr::select(-miRNA_region)%>%
  dplyr::inner_join(mirbase_id_conv, by = c("mirbase_id"="ID"))%>%
  dplyr::select(miRNA_name = Name, read_count, reads_per_million_miRNA_mapped)%>%
  dplyr::group_by(miRNA_name)%>%
  mutate(read_count= sum(read_count), reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
  dplyr::ungroup()%>%
  distinct() -> BCPME_mirnanormal

head(BCPME_mirnanormal)
```

## 2.3 Get the high-throughput experimental miRNA:target dataset.

There are various datasets about miRNA:target pairs such as miRTarBase, DianaTools, miRecords, miRWalk etc. Some of these present the experimentally supported miRNA target pairs or only predicted ones. The experimentally supported datasets generally provides weak evidence for interactions. For these reasons, we obtained the high-throughput experimental miRNA:target dataset from two studies performed by Helwak et al. and Moore et al. These steps were not handle in this file because they contain many processes.

Briefly these datasets contain various common information about miRNA:target interactions such as the miRNA name, miRNAsequence, target name, target sequence, their chromosomal locations, binding location on the target sequence, binding free energy, seed structure. But these datasets provides the informations with different data structures.So we followed the steps:

+ The datasets were directly downloaded from supplementary data files of the studies.
+ It was provided that the datasets are converted to same human genome build.
+ The seed type information was organized as the same style.
+ The datasets were combined.
+ We commited the interaction factors as numeric values according to previous studies. (We added the interaction factors and their numeric values at following tables)

Note: Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study. On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study. Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions. Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values. We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013]. With this process, we have handled this entegrated dataset in context of competitor behaviours and functionality of interactions.

|seed_type| Coefficient |
|-----------|------------|
| `5-mer`   | `0.05` |
| `5-mer_noncanonical`   | `0.04` |
| `6-mer`   | `0.07` |
| `6-mer_noncanonical`   | `0.05` |
| `6-merA1`   | `0.07` |
| `6-merA1_noncanonical`   | `0.05` |
| `7-mer`   | `0.23` |
| `7-mer_noncanonical`   | `0.19` |
| `7-merA1`   | `0.19` |
| `7-merA1_noncanonical`   | `0.16` |
| `7-mer-8m`   | `0.25` |
| `7-mer-8m_nonacnonical`   | `0.21` |
| `8-mer`   | `0.43` |
| `8-mer_noncanonical`   | `0.35` |
| `9-mer`   | `0.43` |
| `9-mer_noncanonical`   | `0.35` |
| `none`   | `0.01` |

|location of binding region on target| Coefficient |
|-----------|------------|
| `3'UTR`   | `0.84` |
| `CDS`   | `0.42` |
| `5'UTR`   | `0.01` |

Finally, we have obtained the experimentally supported miRNA:target dataset.

```{r, message= FALSE, warning=FALSE}
data("experimentalmirnagene")

head(experimentalmirnagene)

```

The methods about miRNA:target interactions are based a basic principle that is reading after isolation of miRNA:target chimerics. The datasets contain all the chimeric miRNA:target structures found in the medium during the experiment. On the other hand, it could be said that the reading is performed as snapshot.Because of that, the methods can provide different chimeric interactions the same miRNA:target pair. We have preferred to select most effective interaction parameters for the same miRNA:target pairs that can exhibit various interactions. The step is performed as:

```{r, message= FALSE, warning=FALSE}
experimentalmirnagene%>%
  dplyr::select(miRNA, Ensembl_Gene_Id, hgnc_symbol, Energy, seed_type_effect, region_effect)%>%
  distinct()%>%
  group_by(Ensembl_Gene_Id, miRNA)%>%
  mutate(seed_type_effect= ifelse(seed_type_effect==max(seed_type_effect), seed_type_effect, max(seed_type_effect)), Energy = ifelse(Energy==min(Energy), Energy, min(Energy)), region_effect= ifelse(region_effect==max(region_effect), region_effect, max(region_effect)))%>%
  distinct()-> tocombine_mirnagene

head(tocombine_mirnagene)
```

## 2.4 Combine the dataset

```{r, message= FALSE, warning=FALSE}
BCPME_mirnanormal%>%
  dplyr::inner_join(tocombine_mirnagene, by = c("miRNA_name"="miRNA"))%>%
  dplyr::inner_join(TCGA_E9_A1N5_normal, by = c("Ensembl_Gene_Id"="ensembl_gene_id", "hgnc_symbol"="external_gene_name"))%>%
  distinct()%>%
  dplyr::select(hgnc_symbol, miRNA_name, mirna_RPM= reads_per_million_miRNA_mapped, GE_normal, Energy, seed_type_effect, region_effect)-> E9GE_mirnagenenormal

head(E9GE_mirnagenenormal)
```

```{r, message= FALSE, warning=FALSE}
BCPME_mirnatumor%>%
  dplyr::inner_join(tocombine_mirnagene, by = c("miRNA_name"="miRNA"))%>%
  dplyr::inner_join(TCGA_E9_A1N5_tumor, by = c("Ensembl_Gene_Id"="ensembl_gene_id", "hgnc_symbol"="external_gene_name"))%>%
  distinct()%>%
  dplyr::select(hgnc_symbol, miRNA_name, mirna_RPM= reads_per_million_miRNA_mapped, GE_tumor, Energy, seed_type_effect, region_effect)-> E9GE_mirnagenetumor

head(E9GE_mirnagenetumor)

```

## 2.5 Selection of trigger node

We have compared two datasets that are obtained for the tumor and normal tissue samples of same patient. We tried to change expression of a gene in normal tissue as the same level in the tumor tissue.

For this step, we have determined the changes of the gene expression in terms of fold change:

```{r}
E9GE_mirnagenetumor%>%
  dplyr::select(hgnc_symbol, GE_tumor)%>%
  dplyr::inner_join((E9GE_mirnagenenormal%>%dplyr::select(hgnc_symbol, GE_normal)), by ="hgnc_symbol")%>%
  dplyr::mutate(FC= GE_tumor/GE_normal)%>%
  distinct()%>%
  filter(FC>2.5,  FC<3.5)-> three_fold_change

# ABCC1 gene has 5827 read count in tumor tissue although 1420 in normal tissue (FC=4.10)
```
Secondly, we have determined the most important nodes of network. We applied `find_node_perturbation` function. We only defined nodes that affect the other nodes more than 1.05 fold change with 10 iteration when they increase 3 fold.

```{r}
E9GE_mirnagenenormal%>%
  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
  find_node_perturbation(how = 3, cycle = 10, fast = 5)%>%
  filter(!is.na(perturbation_efficiency))->perturbationofnetwork

perturbationofnetwork
```
This dataset, perturbationofnetwork, includes 382 genes of totally 423 nodes.

Result of this, we obtained common nodes of these two datasets (i.e perturbationofnetwork and three_fold_change) and selected a gene, SERPINE2.

```{r}
three_fold_change%>%
  inner_join(perturbationofnetwork, by = c("hgnc_symbol"="name"))

#Selected node is SERPINE2
```
## 2.5 Approach of Method into Combined Datasets

We selected SERPINE2 gene for simulation of regulation on network.

### 2.5.1 Find iteration of simulation

```{r}
as.data.frame(E9GE_mirnagenenormal)%>%
 priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
 update_how("SERPINE2",2.75) %>%
 simulate(250) %>%
 find_iteration(limit=1, plot= TRUE) #limit=1 describes the change that is not taken into account.
```

The node amount of changed gene on the system in terms of percentage were shown in above. As seen, firstly, the changed gene count increase. The system which contains the hundreds of miRNAs and thousands of genes can slowly gain the steady-state again. At first glance, it can be assumed that when all nodes in the system are reached, stable state will be provided. However, although all nodes are reached, the nodes competing with each other cause the edits to continue for a while.

The dynamics of the approach are shown in package viggnettes [link](https://selcenari.github.io/ceRNAnetsim/articles/convenient_iteration.html).

So, we offered an approach about to find iteration. `find_iteration` function does not give the iteration to gain steady-state, but it gives the iteration which has maximum affected node counts. The function is applied as following:

```{r}
as.data.frame(E9GE_mirnagenenormal)%>%
 priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
 update_how("SERPINE2",2.75) %>%
 simulate(100) %>%
 find_iteration(limit=1, plot= FALSE)

#31
```

### 2.5.2 Simulation of dataset

We tried to apply two fold of the point that SERPINE2 has maximum affected genes on network.

```{r}

as.data.frame(E9GE_mirnagenenormal)%>%
  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
  update_how("SERPINE2",2.75)%>%
  simulate(62)

```

```{r}
as.data.frame(E9GE_mirnagenenormal)%>%
  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
  update_how("SERPINE2",2.75)%>%
  simulate(62)%>%
  as_tibble()%>%
  select(name, initial_count, count_current)->simulation_results
```

###2.5.3 Comparison of simulation results and tumor tissue expression values
```{r}
E9GE_mirnagenetumor%>%
  dplyr::select(hgnc_symbol, GE_tumor)%>%
  dplyr::inner_join((E9GE_mirnagenenormal%>%dplyr::select(hgnc_symbol, GE_normal)), by ="hgnc_symbol")%>%
  inner_join(simulation_results, by= c("hgnc_symbol"="name"))%>%
  distinct()
```

Actually, we have developed to provide a new approach mirna mediated regulation networks. This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and cohorent miRNA:target regulation approach.

### 2019-09-15 article.Rmd
last meeting

@@
   html_document: default
-documentclass: article
 link-citations: yes
@@
 We applied the model to real sample consisting of breast cancer gene and miRNA expression dataset and experimental miRNA:target interaction dataset all generated via high throughput sequencing methods.
-A gene over-expressed in tumor tissue, namely *ABCC1*, is used as perturbing element.
+A gene over-expressed in tumor tissue, namely *SERPINE2*, is used as perturbing element.
 We have observed that change in expression level of single gene in miRNA:target network is sufficient to perturb regulations in whole network, due to unforeseen and unpredicted regulation which are only visible when considered in network context.
@@
 In this case, the common target acts as trigger for the other group (the targets interacted with M2 miRNA) in the network.
-Due to common targeted elements, the change in one neighborhood spreads to other neighborhoods (\autoref{fig1}d), consequently have potential to effect whole network due to "ripple effect".
+Due to common targeted elements, the change in one neighborhood spreads to other neighborhoods (\autoref{fig1}d), consequently have potential to affect whole network due to "ripple effect".
 
@@
 c) Distribution of miRNA1 (M1) changes.
-d) The change at expression of common target effects changes of proportional distribution of miRNA2 (M2).
+d) The change at expression of common target affects changes of proportional distribution of miRNA2 (M2).
 Expression values are rounded to integers for simplicity.
@@
 In the literature, binding free energy [@cao_predicting_2012 ; @helwak_mapping_2013] and seed type [@werfel_preferential_2017] in miRNA:target interactions are described as factors effecting binding affinity.
-Also, binding region (e.g, 5'UTR, 3'UTR, CDS) was shown to drastically effect miRNA degradation efficiency [@hausser_analysis_2013; @helwak_mapping_2013].
+Also, binding region (e.g, 5'UTR, 3'UTR, CDS) was shown to drastically affect miRNA degradation efficiency [@hausser_analysis_2013; @helwak_mapping_2013].
 Since binding or degradation efficiency from literature have various and broad ranges, both binding and efficiency factors are normalized to their maximum values and scaled to [0,1] interval.
@@
 Detailed description of network construction and its code is available in Supplementary Materials and Methods.
-ABCC1 gene, over-expressed in tumor tissue, was selected as trigger for simulation of integrated dataset.
-After simulation of network, we have compared simulation results and tumor tissue expression levels.
+We have determined the most effective nodes in network and fold changes of genes as compared tumor and normal tissue samples of the same patient.As a result of comparison of these two datasets, we selected a gene , SERPINE2, as trigger for simulation of model (see Supplementary Materials and Methods).
 
@@
 ## Results and Conclusions
-### Networks with single factor
+### Network with miRNA:ceRNA expression level
 We have developed a network-based approach to assess effects of expression level changes in competitive ceRNA regulation.
@@
 [#TODO]: <> "figure1, 2 ve 3 uzerinden farkli dagilim ve degradasyon açıklaması buraya. G2 uzerinden olmasa iyi olur" 
+[FIXED by selcen]
+When the model run based on expression values of miRNAs and ceRNAs (genes), proportional distribution of miRNA on targets is determined in accordance with \eqref{eq:1, eq:2}.
+In that case, the system is observed as shown in \autoref{fig1}a at steady-state.
+However, if interaction factors are taken into account in the approach, specified interaction factors (i.e, binding factors such as seed type and energy) affect distribution of miRNA on targets as calculated in \eqref{eq:3} (shown in \autoref{fig2}c).
+After distributed miRNA efficiency was re-calculated with degradation factor(s) (i.e region effect) according to \eqref{eq:4}, steady-state condition of network is obtained (shown in \autoref{fig2}d and \autoref{fig3}a).
+For instance, proportional distribution of G1:M1 interaction in \autoref{fig1}a is calculated lower than the same pair (G1:M1) in the other approach that includes interaction factors (\autoref{fig2}c).
+But, a decrease is observed in repression efficiency of M1 miRNA on G1 in interaction factor based network approach (\autoref{fig2}d and \autoref{fig3}a).
 
@@
 Since perturbation efficiency of each gene is different, we have developed a function which screens each gene in the network for their perturbation efficiencies.
-**(see details in Supplementary File(fig1_2app.Rmd)**.
+**(see details in Supplementary Materials and Methods**.
 On the other hand, when we applied the method on the minimal dataset, Gene4 has been found to be the most efficient element in terms of number of perturbed elements and miRNA2 (M2) has been found to be causing the highest mean in expression changes.
@@
 The network was finalized by integrating gene and miRNA expression levels from TCGA determined by RNA-Seq.
-Our  model can successfully simulate perturbations in such large network despite complex behaviors and struggle to reach steady-state **(see details in Supplementary File(TCGA_E9-A1N5_article.Rmd)**.
-Simulations show that change in expression level of single gene has potential to effect whole network, perturbing almost all nodes.
+Our  model can successfully simulate perturbations in such large network despite complex behaviors and struggle to reach steady-state **(see details in Supplementary Materials and Methods)**.
+Simulations show that change in expression level of single gene has potential to affect whole network, perturbing almost all nodes.
 These observations are in accordance with competing endogenous RNA hypothesis where genes targeted with many common miRNAs subsequently transmit perturbation to neighboring groups.
 
-[#TODO] We have observed that the results of simulation with ABCC1 gene, *Multidrug resistance-associated protein 1*, which is one of the most significant factors to develop resistance against chemotherapotic agents[@atalay2008role; @lu2015microrna; @atalay2006multidrug] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach.
+[#TODO] We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach.
 Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues.
@@
 [#TODO]: <> "run perturbation screening on tcga set, find a gene then couple sentences about that gene"
+[fixed by selcen]
 
@@
 Next, a network based kinetic model integrating miRNA and target rates of transcription, degradation, binding and unbinding was developed [@nitzan_interactions_2014] using high throughput experimental dataset about miR-92a depletion [@helwak_mapping_2013].
-It was demonstrated that distant ceRNAs can interact with each other via indirect links, and the interactions are effected depending on distance between ceRNAs or topological features of network [@nitzan_interactions_2014].
+It was demonstrated that distant ceRNAs can interact with each other via indirect links, and the interactions are affected depending on distance between ceRNAs or topological features of network [@nitzan_interactions_2014].
 More recently, an approach to detect ceRNA interaction by using the miRNA expression, gene expression and common miRNAs between gene targets was developed [@markus_list_sponge_2017] which was effective in analyzing genes through miRNAs.

### 2019-09-17 Supplementary_mat_met.Rmd
article

--- 2019-09-15 Suplementary_mat1.Rmd
@@
 ---
-title: "Supplementary File 1"
+title: "Supplementary Materials And Methods"
 author: "Selcen Ari"
@@
 bibliography: ref.bib
+header-includes: 
+  \usepackage{float} \floatplacement{figure}{H} 
+  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}     \renewcommand{\thefigure}{S\arabic{figure}}}
+  \usepackage{subfig}
 ---
 
-#1. Defined functions for ceRNA models and workflow of method
+\beginsupplement
+
+# 1. Defined functions for ceRNA models and workflow of method
 We defined the functions that can be used with R programming.
@@
 All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
-The functions and steps of approach are explained as following (Figure 1) :
+The functions and steps of approach are explained as following (Figure S1) :
 
@@
 
+```{r, include=FALSE, message=FALSE, warning=FALSE}
+
+library(tidygraph)
+library(dplyr)
+library(tidyr)
+library(stringr)
+
+```
+
 ```{r, message=FALSE, warning=FALSE}
@@
 
-|competing| miRNA | Competing_expression | miRNA_expression | seed_type | region | energy
-|-----------|------------|----------|----------------|---------|-----------|----------|
-| `Gene1`   | `Mir1` | 10000 | 1000 | 0.43 | 0.30 | -20 |
-| `Gene2`   | `Mir1` | 10000 | 1000 | 0.43 | 0.01 | -15 |
-| `Gene3`   | `Mir1` | 5000  | 1000 | 0.32 | 0.40 | -14 |
-| `Gene4`   | `Mir1` | 10000 | 1000 | 0.23 | 0.50 | -10 |
-| `Gene4`   | `Mir2` | 10000 | 2000 | 0.35 | 0.90 | -12 |
-| `Gene5`   | `Mir2` | 5000  | 2000 | 0.05 | 0.40 | -11 |
-| `Gene6`   | `Mir2` | 10000 | 2000 | 0.01 | 0.80 | -25 |
+See Figure S1 in Supplementary Tables file.
 
@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset in Steady-state", fig.width=5, fig.height=4}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE, fig.keep='all', fig.cap= "Gene2 Upregulation on Minimal Dataset", fig.width=5, fig.height=4}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.cap = "Sequential iteration of minsampdata", fig.subcap= c("First response of system to Gene2 upregulation (2nd iteration)", "Spreading of perturbation on system (3th iteration)")}
 
@@
 ```
+
 Note that the regulations are colored according to expression changes of present and a previous value.
@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset with interaction factors in Steady-state", fig.width=5, fig.height=4}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset in steady-state conditions")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset with interaction factors in steady-state conditions")
 
@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE, fig.cap= " When Gene2 is upregulated on Minimal Dataset with interaction factors", fig.width=5, fig.height=4}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
   update_how("Gene2", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation with interaction factors")
 ```
 
-```{r, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.cap = "Sequential iteration of minsampdata with interaction factors", fig.subcap= c("First response of system to Gene2 upregulation (2nd iteration)", "Spreading of perturbation on system (3th iteration)")}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
   update_how("Gene2", 2)%>%
-  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors",2)
+  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation with interaction factors",2)
 ```
@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.cap = "Perturbation of Gene4 on minsampdata with interaction factors", fig.subcap= c("1st iteration after Gene4 upregulation", "2nd iteration after Gene4 perturbation")}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
   update_how("Gene4", 2)%>%
-  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors",2)
+  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Common Gene-Gene4 Upregulation without interaction factors",2)
 ```
@@
 
-#2. Obtaining breast cancer dataset and integration
+# 2. Obtaining breast cancer dataset and integration
 
@@
 # Install the packages which is necessary to obtain and process datasets.
-install.packages("BiocManager")
-BiocManager::install("SummarizedExperiment")
-BiocManager::install("TCGAbiolinks")
+#install.packages("BiocManager")
+#BiocManager::install("SummarizedExperiment")
+#BiocManager::install("TCGAbiolinks")
 library(TCGAbiolinks)
 library(SummarizedExperiment)
-#install devtools
-devtools::install_github("selcenari/ceRNAnetsim")
-library(ceRNAnetsim)
 ```
@@
 
-```{r,message= FALSE, warning=FALSE}
+```{r,message= FALSE, warning=FALSE, eval=FALSE}
 
@@
 
-```{r, message= FALSE, warning=FALSE}
+```{r, message= FALSE, warning=FALSE, eval=FALSE}
 BCP_normal <- GDCquery(project = "TCGA-BRCA",
@@
 
-```
-
-```{r, echo=FALSE, message= FALSE, warning=FALSE}
-#remove:
-rm(BCPGE_normal, BCPGE_tumor, BCP_normal, BCP_tumor)
 ```
@@
 
-```{r, message= FALSE, warning=FALSE}
+```{r, message= FALSE, warning=FALSE, eval=FALSE}
 BCP_mirnatumor <- GDCquery(project = "TCGA-BRCA",
@@
 
-```{r, message= FALSE, warning=FALSE}
+```{r, message= FALSE, warning=FALSE, eval=FALSE}
 BCP_mirnanormal <- GDCquery(project = "TCGA-BRCA",
@@
 + The datasets were combined.
-+ We commited the interaction factors as numeric values according to previous studies. (We added the interaction factors and their numeric values at following tables)
-
-Note: Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study.
-On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study.
-Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions.
-Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values.
-We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
-With this process, we have handled this entegrated dataset in context of competitor behaviours and functionality of interactions.
-
-|seed_type| Coefficient |
-|-----------|------------|
-| `5-mer`   | `0.05` |
-| `5-mer_noncanonical`   | `0.04` |
-| `6-mer`   | `0.07` |
-| `6-mer_noncanonical`   | `0.05` |
-| `6-merA1`   | `0.07` |
-| `6-merA1_noncanonical`   | `0.05` |
-| `7-mer`   | `0.23` |
-| `7-mer_noncanonical`   | `0.19` |
-| `7-merA1`   | `0.19` |
-| `7-merA1_noncanonical`   | `0.16` |
-| `7-mer-8m`   | `0.25` |
-| `7-mer-8m_nonacnonical`   | `0.21` |
-| `8-mer`   | `0.43` |
-| `8-mer_noncanonical`   | `0.35` |
-| `9-mer`   | `0.43` |
-| `9-mer_noncanonical`   | `0.35` |
-| `none`   | `0.01` |
-
-|location of binding region on target| Coefficient |
-|-----------|------------|
-| `3'UTR`   | `0.84` |
-| `CDS`   | `0.42` |
-| `5'UTR`   | `0.01` |
++ We commited the interaction factors as numeric values according to previous studies. (We added the interaction factors and their numeric values at Supplementary tables S2-3)
 
@@
 
-```{r, message= FALSE, warning=FALSE}
+```{r, message= FALSE, warning=FALSE, eval=FALSE}
 experimentalmirnagene%>%
@@
 
-```{r, message= FALSE, warning=FALSE}
+```{r, message= FALSE, warning=FALSE, eval=FALSE}
 BCPME_mirnanormal%>%
@@
 
+```
+
+```{r}
+data("E9GE_mirnagenenormal")
 head(E9GE_mirnagenenormal)
@@
 
-```{r, message= FALSE, warning=FALSE}
+```{r, message= FALSE, warning=FALSE, eval=FALSE}
 BCPME_mirnatumor%>%
@@
 
+```
+
+```{r}
+data("E9GE_mirnagenetumor")
 head(E9GE_mirnagenetumor)
-
 ```
@@
 
-```{r}
-E9GE_mirnagenenormal%>%
+```{r, include=FALSE, eval=FALSE}
+as.data.frame(E9GE_mirnagenenormal)%>%
   priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
@@
 
-```{r}
+```{r, eval=FALSE}
 three_fold_change%>%
@@
 
-```{r}
+```{r, fig.cap= "Percentage of affected nodes of each iteration for SERPINE2 Gene"}
 as.data.frame(E9GE_mirnagenenormal)%>%
@@
  update_how("SERPINE2",2.75) %>%
- simulate(250) %>%
+ simulate(150) %>%
  find_iteration(limit=1, plot= TRUE) #limit=1 describes the change that is not taken into account.
@@
 
-###2.5.3 Comparison of simulation results and tumor tissue expression values
+### 2.5.3 Comparison of simulation results and tumor tissue expression values
 ```{r}
@@
 This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and cohorent miRNA:target regulation approach.
+
+# 3. REFERENCES

### 2019-09-17 Supplementary_tables.Rmd
article

--- 2019-09-15 Suplementary_mat1.Rmd
@@
 ---
-title: "Supplementary File 1"
-author: "Selcen Ari"
-date: "7 Nisan 2019"
+title: "Supplementary Tables"
+author: "Selcen Arı"
+date: "16 09 2019"
 output:
@@
 bibliography: ref.bib
+header-includes: 
+  \usepackage{float} \floatplacement{figure}{H} 
+  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
 ---
 
-#1. Defined functions for ceRNA models and workflow of method
-We defined the functions that can be used with R programming.
-Briefly, these functions process a given miRNA:gene dataset and convert to graph object.
-All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
-The functions and steps of approach are explained as following (Figure 1) :
-
-**Convertion of dataset**: priming_graph function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
-If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
-The formulations that are given in equations (1-4) are performed in this function.
-This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
-It is assumed that the initial target amounts in the dataset is observed after the reppressive activity of miRNAs in steady-state.
-
-![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow1.png){#fig1}
-
-**Transition of variables in graph**: In the previous step, the calculations are performed in the edge variables of the graph object.
-However, the graph object allows to use node variables, while the node features are handled to the graph.
-In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.
-
-**Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficieny coefficients are calculated according to this acceptance.
-In the network that is found in steady-state conditions, the change is applied to the graph object for disturbtion of steady-state.
-To provide the disturbtion in the network the workflow offer two methods: update_variables and update_how.
-The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
-The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
-
-**Updating current values of variables**: After variables updating in edge varibles, these are carried to node variables.
-Current and previous values of variables are stored as node variables with update_variables function.
-
-**Simulation of competing behavior of targets**: After the change in the steady-state conditions, the network elements try to gain steady-state again.
-This process progresses as repeating of regulations after the spreading the changes in the network.
-In this step, simulation of regulations according to given cycle count in simulate function is applied.
-After each simulation cycle, the miRNA repression values are re-calculated and the current values of competing elements are found and saved.
-The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node data.
-
-The node elements in the dataset are handled as two type; repressives (miRNAs) and competings (targets).
-It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used.
-If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user.
-On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger.
-The functions that are used in the approach are developed with R programming so as can be used with other packages.
-These are can be found in the github repository [ceRNAnetsim github page](https://github.com/selcenari/ceRNAnetsim) and improved with contributions of others.
-
-```{r, message=FALSE, warning=FALSE}
-#install.packages("devtools")
-#devtools::install_github("selcenari/ceRNAnetsim")
-library(ceRNAnetsim)
+```{r, eval=FALSE, include=FALSE}
+#install.packages("kable")
+library(xtable)
 ```
 
-+ load *minsamp* data
-
-```{r}
-data("minsamp")
-
-minsamp
-```
+\beginsupplement
 
@@
 
-## *minsamp* dataset analysis in lack of interaction factors.
-
-Firstly, we have analysed minimal data without interaction factors between miRNA:target.
-
-- 1. We have evaluated graph in the steady state conditions as followings:
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset in steady-state conditions")
-```
-
-- 2. We have obtained graph after change on Gene2 expression as followings:
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors")
-```
-
-- 3. We have determined regulations after Gene2 Upregulation:
-
-```{r, message=FALSE, warning=FALSE}
-
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)%>%
-  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "Regulations after Gene2 Upregulation", 2)
-
-```
-Note that the regulations are colored according to expression changes of present and a previous value.
-So, it can be observed that whole gene expressions increase in comparison of initial steady-state.
-The overall regulations of gene expressions are as followings:
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)%>%
-  simulate(2)%>%
-  activate(edges)%>%
-  as_tibble()%>%
-  select(Competing_name,comp_count_list,effect_list)%>%
-  unnest()
-```
-
-## *minsamp* dataset analysis with interaction factors.
-
-We have made the same analysis in present of interaction factors.
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset in steady-state conditions")
-
-```
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene2", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors")
-```
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene2", 2)%>%
-  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors",2)
-```
-
-When the graphs which were resulted from analyses were examined, it was observed that behaviours were same.
-But, when the results were analysed in terms of expression values, the regulation differences can be observed.
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene2", 2)%>%
-  simulate(3)%>%
-  activate(edges)%>%
-  as_tibble()%>%
-  select(Competing_name,comp_count_list,effect_list)%>%
-  unnest()
-```
-
-## Common target perturbation in *minsamp* dataset.
-
-There are hundreds of defined miRNAs for human, so this results in presence of common targets of miRNAs in cells.
-Therefore, we have analysed perturbation efficiency of common target in *minsamp* dataset.
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene4", 2)%>%
-  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors",2)
-```
-
-The common target perturbation  (increasing to two fold at Gene4 expression in presence of interaction factors) resulted in more prominent efficiency at the same conditions (shown in following).
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene4", 2)%>%
-  simulate(3)%>%
-  activate(edges)%>%
-  as_tibble()%>%
-  select(Competing_name,comp_count_list,effect_list)%>%
-  unnest()
-```
-
-## Determination of perturbation efficiencies efficiencies of elements in system.
-
-```{r, message=FALSE, warning=FALSE}
-
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)-> sample_graph
-
-find_node_perturbation(sample_graph, how = 2, cycle = 3, limit = 0.1)
-
-```
-
-#2. Obtaining breast cancer dataset and integration
-
-This section describes how to apply `ceRNAnetsim` package on a breast cancer patient miRNA:target interaction dataset. Before the approach, we obtained three datasets and combined them.
-
-```{r, echo=FALSE, message= FALSE, warning=FALSE}
-# Install the packages which is necessary to obtain and process datasets.
-install.packages("BiocManager")
-BiocManager::install("SummarizedExperiment")
-BiocManager::install("TCGAbiolinks")
-library(TCGAbiolinks)
-library(SummarizedExperiment)
-#install devtools
-devtools::install_github("selcenari/ceRNAnetsim")
-library(ceRNAnetsim)
-```
-
-## 2.1 How to get gene expression counts of TCGA-E9-A1N5 patient.
-
-We have obtained the gene expression values of patient using [TCGAbiolinks](https://www.bioconductor.org/packages/release/bioc/html/TCGAbiolinks.html) package from [Bioconductor](https://www.bioconductor.org/).
-For this process, we have followed the instructions of the package.
-`TCGAbiolinks` package provides to obtain data for whole number of given barcode(s) at once.
-But, we preferred to download them seperately to show datasets.
-
-+ Obtain to gene expression counts of tumor tissue.
-
-```{r,message= FALSE, warning=FALSE}
-
-BCP_tumor <- GDCquery(project = "TCGA-BRCA",
-                  data.category = "Transcriptome Profiling",
-                  data.type = "Gene Expression Quantification", 
-                  workflow.type = "HTSeq - Counts",
-                  barcode = "TCGA-E9-A1N5-01A-11R-A14D-07")
-
-GDCdownload(BCP_tumor)
-BCPGE_tumor <- GDCprepare(BCP_tumor)
-
-as.data.frame(assay(BCPGE_tumor))%>%
-  mutate(ensembl_gene_id = rownames(.))%>%
-  dplyr::inner_join(as.data.frame(rowData(BCPGE_tumor)), by = "ensembl_gene_id")%>%
-  dplyr::select(ensembl_gene_id, external_gene_name, 1)-> TCGA_E9_A1N5_tumor
-
-colnames(TCGA_E9_A1N5_tumor)[3] <- "GE_tumor"
-
-head(TCGA_E9_A1N5_tumor)
-```
-
-+ Obtain to gene expression counts of normal tissue.
-
-```{r, message= FALSE, warning=FALSE}
-BCP_normal <- GDCquery(project = "TCGA-BRCA",
-                  data.category = "Transcriptome Profiling",
-                  data.type = "Gene Expression Quantification", 
-                  workflow.type = "HTSeq - Counts",
-                  barcode = "TCGA-E9-A1N5-11A-41R-A14D-07")
-GDCdownload(BCP_normal)
-
-BCPGE_normal <- GDCprepare(BCP_normal)
-
-as.data.frame(assay(BCPGE_normal))%>%
-  mutate(ensembl_gene_id = rownames(.))%>%
-  dplyr::inner_join(as.data.frame(rowData(BCPGE_normal)), by = "ensembl_gene_id")%>%
-  dplyr::select(ensembl_gene_id, external_gene_name, 1)-> TCGA_E9_A1N5_normal
-
-colnames(TCGA_E9_A1N5_normal)[3] <- "GE_normal"
-
-head(TCGA_E9_A1N5_normal)
-
-```
-
-```{r, echo=FALSE, message= FALSE, warning=FALSE}
-#remove:
-rm(BCPGE_normal, BCPGE_tumor, BCP_normal, BCP_tumor)
-```
-
-## 2.2 How to get miRNA expression counts of TCGA-E9-A1N5 patient.
-
-We have used `TCGAbiolinks` package to obtain miRNA expression quantification.
-The query gives read count of miRNA as isoform chromosome coordination.
-The data also contains mature miRNA information.
-So, we processed data to attain -5p -3p isoform information using mirbase release21 dataset.
-
-+ Get the mirbase id of mature miRNA:
-
-We downloaded the mirbase release 21 dataset from [mirbase](http://www.mirbase.org/ftp.shtml) and processed the patient mirna expression datasets as following:
-
-```{r, message= FALSE, warning=FALSE}
-library(readr)
-read_tsv("hsa_mirna.txt", comment = "#", col_names = FALSE)%>%
-  dplyr::select(mirna_type= X3, definition = X9)%>%
-  filter(!endsWith(mirna_type, "primary_transcript"))%>%
-  tidyr::separate(definition, c("ID", "Alias", "Name", "Derivated"), sep = ";")%>%
-  dplyr::select(Alias, Name)%>%
-  tidyr::separate(Alias, c("trash1", "ID"), sep = "=")%>%
-  tidyr::separate(Name, c("trash2", "Name"), sep = "=")%>%
-  dplyr::select(-trash1, -trash2)-> mirbase_id_conv
-
-head(mirbase_id_conv)
-```
-
-+ Obtain the miRNA expression of tumor tissue of patient:
-
-```{r, message= FALSE, warning=FALSE}
-BCP_mirnatumor <- GDCquery(project = "TCGA-BRCA",
-                  data.category = "Transcriptome Profiling",
-                  data.type = "Isoform Expression Quantification",
-                  workflow.type = "BCGSC miRNA Profiling",
-                  barcode = "TCGA-E9-A1N5-01A-11R-A14C-13")
-
-GDCdownload(BCP_mirnatumor)
-
-GDCprepare(BCP_mirnatumor)%>%
-  as.data.frame()%>%
-  dplyr::select(miRNA_ID, read_count, reads_per_million_miRNA_mapped, miRNA_region)%>%
-  dplyr::filter(startsWith(miRNA_region, "mature"))%>%
-  dplyr::mutate(mirbase_id =str_remove(miRNA_region, "mature,"))%>%
-  dplyr::select(-miRNA_region)%>%
-  dplyr::inner_join(mirbase_id_conv, by = c("mirbase_id"="ID"))%>%
-  dplyr::select(miRNA_name = Name, read_count, reads_per_million_miRNA_mapped)%>%
-  dplyr::group_by(miRNA_name)%>%
-  mutate(read_count= sum(read_count), reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
-  dplyr::ungroup()%>%
-  distinct() -> BCPME_mirnatumor
-
-head(BCPME_mirnatumor)
-```
-
-+ Obtain the miRNA expression of normal tissue of patient:
-
-```{r, message= FALSE, warning=FALSE}
-BCP_mirnanormal <- GDCquery(project = "TCGA-BRCA",
-                  data.category = "Transcriptome Profiling",
-                  data.type = "Isoform Expression Quantification",
-                  workflow.type = "BCGSC miRNA Profiling",
-                  barcode = "TCGA-E9-A1N5-11A-41R-A14C-13")
-
-GDCdownload(BCP_mirnanormal)
-# a616435d-0b69-48ac-813d-5d75ad9b85eb.mirbase21.isoforms.quantification.txt
-
-GDCprepare(BCP_mirnanormal)%>%
-  as.data.frame()%>%
-  dplyr::select(miRNA_ID, read_count, reads_per_million_miRNA_mapped, miRNA_region)%>%
-  dplyr::filter(startsWith(miRNA_region, "mature"))%>%
-  dplyr::mutate(mirbase_id =str_remove(miRNA_region, "mature,"))%>%
-  dplyr::select(-miRNA_region)%>%
-  dplyr::inner_join(mirbase_id_conv, by = c("mirbase_id"="ID"))%>%
-  dplyr::select(miRNA_name = Name, read_count, reads_per_million_miRNA_mapped)%>%
-  dplyr::group_by(miRNA_name)%>%
-  mutate(read_count= sum(read_count), reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
-  dplyr::ungroup()%>%
-  distinct() -> BCPME_mirnanormal
-
-head(BCPME_mirnanormal)
-```
-
-## 2.3 Get the high-throughput experimental miRNA:target dataset.
-
-There are various datasets about miRNA:target pairs such as miRTarBase, DianaTools, miRecords, miRWalk etc.
-Some of these present the experimentally supported miRNA target pairs or only predicted ones.
-The experimentally supported datasets generally provides weak evidence for interactions.
-For these reasons, we obtained the high-throughput experimental miRNA:target dataset from two studies performed by Helwak et al.
-and Moore et al.
-These steps were not handle in this file because they contain many processes.
-
-Briefly these datasets contain various common information about miRNA:target interactions such as the miRNA name, miRNAsequence, target name, target sequence, their chromosomal locations, binding location on the target sequence, binding free energy, seed structure.
-But these datasets provides the informations with different data structures.So we followed the steps:
-
-+ The datasets were directly downloaded from supplementary data files of the studies.
-+ It was provided that the datasets are converted to same human genome build.
-+ The seed type information was organized as the same style.
-+ The datasets were combined.
-+ We commited the interaction factors as numeric values according to previous studies. (We added the interaction factors and their numeric values at following tables)
+Table: *minsamp* sample dataset that includes interaction factors.
 
@@
 
+Table: Efficiency factors for seed types.
+
 |location of binding region on target| Coefficient |
@@
 
-Finally, we have obtained the experimentally supported miRNA:target dataset.
+Table: Efficiency factors for binding regions on targets
 
-```{r, message= FALSE, warning=FALSE}
-data("experimentalmirnagene")
+```{r, eval=FALSE, include=FALSE}
+data("E9GE_mirnagenenormal")
 
-head(experimentalmirnagene)
-
+xtable(head(E9GE_mirnagenenormal),caption = "Example of E9GE_mirnagenenormal dataset" )
 ```
 
-The methods about miRNA:target interactions are based a basic principle that is reading after isolation of miRNA:target chimerics.
-The datasets contain all the chimeric miRNA:target structures found in the medium during the experiment.
-On the other hand, it could be said that the reading is performed as snapshot.Because of that, the methods can provide different chimeric interactions the same miRNA:target pair.
-We have preferred to select most effective interaction parameters for the same miRNA:target pairs that can exhibit various interactions.
-The step is performed as:
+|hgnc_symbol| miRNA_name | mirna_RPM | GE_normal | Energy | seed_type_effect | region_effect |
+|-----------|------------|----------|----------------|---------|-----------|----------|
+| CCNG1 | hsa-let-7a-5p | 111204.15 | 5245.00 | -25.10 | 0.05 | 0.84 |
+| DICER1 | hsa-let-7a-5p | 111204.15 | 3285.00 | -24.40 | 0.43 | 0.42 | 
+| SESN1 | hsa-let-7a-5p | 111204.15 | 1179.00 | -22.20 | 0.05 | 0.93 |
+| NIPBL | hsa-let-7a-5p | 111204.15 | 4503.00 | -22.10 | 0.05 | 0.42 | 
+| INTS12 | hsa-let-7a-5p | 111204.15 | 600.00 | -21.90 | 0.05 | 0.42 | 
+| FNIP1 | hsa-let-7a-5p | 111204.15 | 1248.00 | -21.80 | 0.43 | 0.84 |
 
-```{r, message= FALSE, warning=FALSE}
-experimentalmirnagene%>%
-  dplyr::select(miRNA, Ensembl_Gene_Id, hgnc_symbol, Energy, seed_type_effect, region_effect)%>%
-  distinct()%>%
-  group_by(Ensembl_Gene_Id, miRNA)%>%
-  mutate(seed_type_effect= ifelse(seed_type_effect==max(seed_type_effect), seed_type_effect, max(seed_type_effect)), Energy = ifelse(Energy==min(Energy), Energy, min(Energy)), region_effect= ifelse(region_effect==max(region_effect), region_effect, max(region_effect)))%>%
-  distinct()-> tocombine_mirnagene
+Table: Example of E9GE_mirnagenenormal dataset.
 
-head(tocombine_mirnagene)
-```
+As a result of simulation a dataset, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at followings.
 
-## 2.4 Combine the dataset
-
-```{r, message= FALSE, warning=FALSE}
-BCPME_mirnanormal%>%
-  dplyr::inner_join(tocombine_mirnagene, by = c("miRNA_name"="miRNA"))%>%
-  dplyr::inner_join(TCGA_E9_A1N5_normal, by = c("Ensembl_Gene_Id"="ensembl_gene_id", "hgnc_symbol"="external_gene_name"))%>%
-  distinct()%>%
-  dplyr::select(hgnc_symbol, miRNA_name, mirna_RPM= reads_per_million_miRNA_mapped, GE_normal, Energy, seed_type_effect, region_effect)-> E9GE_mirnagenenormal
-
-head(E9GE_mirnagenenormal)
-```
-
-```{r, message= FALSE, warning=FALSE}
-BCPME_mirnatumor%>%
-  dplyr::inner_join(tocombine_mirnagene, by = c("miRNA_name"="miRNA"))%>%
-  dplyr::inner_join(TCGA_E9_A1N5_tumor, by = c("Ensembl_Gene_Id"="ensembl_gene_id", "hgnc_symbol"="external_gene_name"))%>%
-  distinct()%>%
-  dplyr::select(hgnc_symbol, miRNA_name, mirna_RPM= reads_per_million_miRNA_mapped, GE_tumor, Energy, seed_type_effect, region_effect)-> E9GE_mirnagenetumor
-
-head(E9GE_mirnagenetumor)
-
-```
-
-## 2.5 Selection of trigger node
-
-We have compared two datasets that are obtained for the tumor and normal tissue samples of same patient. We tried to change expression of a gene in normal tissue as the same level in the tumor tissue.
-
-For this step, we have determined the changes of the gene expression in terms of fold change:
-
-```{r}
-E9GE_mirnagenetumor%>%
-  dplyr::select(hgnc_symbol, GE_tumor)%>%
-  dplyr::inner_join((E9GE_mirnagenenormal%>%dplyr::select(hgnc_symbol, GE_normal)), by ="hgnc_symbol")%>%
-  dplyr::mutate(FC= GE_tumor/GE_normal)%>%
-  distinct()%>%
-  filter(FC>2.5,  FC<3.5)-> three_fold_change
-
-# ABCC1 gene has 5827 read count in tumor tissue although 1420 in normal tissue (FC=4.10)
-```
-Secondly, we have determined the most important nodes of network.
-We applied `find_node_perturbation` function.
-We only defined nodes that affect the other nodes more than 1.05 fold change with 10 iteration when they increase 3 fold.
-
-```{r}
-E9GE_mirnagenenormal%>%
-  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
-  find_node_perturbation(how = 3, cycle = 10, fast = 5)%>%
-  filter(!is.na(perturbation_efficiency))->perturbationofnetwork
-
-perturbationofnetwork
-```
-This dataset, perturbationofnetwork, includes 382 genes of totally 423 nodes.
-
-Result of this, we obtained common nodes of these two datasets (i.e perturbationofnetwork and three_fold_change) and selected a gene, SERPINE2.
-
-```{r}
-three_fold_change%>%
-  inner_join(perturbationofnetwork, by = c("hgnc_symbol"="name"))
-
-#Selected node is SERPINE2
-```
-## 2.5 Approach of Method into Combined Datasets
-
-We selected SERPINE2 gene for simulation of regulation on network.
-
-### 2.5.1 Find iteration of simulation
-
-```{r}
-as.data.frame(E9GE_mirnagenenormal)%>%
- priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
- update_how("SERPINE2",2.75) %>%
- simulate(250) %>%
- find_iteration(limit=1, plot= TRUE) #limit=1 describes the change that is not taken into account.
-```
-
-The node amount of changed gene on the system in terms of percentage were shown in above.
-As seen, firstly, the changed gene count increase.
-The system which contains the hundreds of miRNAs and thousands of genes can slowly gain the steady-state again.
-At first glance, it can be assumed that when all nodes in the system are reached, stable state will be provided.
-However, although all nodes are reached, the nodes competing with each other cause the edits to continue for a while.
-
-The dynamics of the approach are shown in package viggnettes [link](https://selcenari.github.io/ceRNAnetsim/articles/convenient_iteration.html).
-
-So, we offered an approach about to find iteration.
-`find_iteration` function does not give the iteration to gain steady-state, but it gives the iteration which has maximum affected node counts.
-The function is applied as following:
-
-```{r}
-as.data.frame(E9GE_mirnagenenormal)%>%
- priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
- update_how("SERPINE2",2.75) %>%
- simulate(100) %>%
- find_iteration(limit=1, plot= FALSE)
-
-#31
-```
-
-### 2.5.2 Simulation of dataset
-
-We tried to apply two fold of the point that SERPINE2 has maximum affected genes on network.
-
-```{r}
-
-as.data.frame(E9GE_mirnagenenormal)%>%
-  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
-  update_how("SERPINE2",2.75)%>%
-  simulate(62)
-
-```
-
-```{r}
-as.data.frame(E9GE_mirnagenenormal)%>%
-  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
-  update_how("SERPINE2",2.75)%>%
-  simulate(62)%>%
-  as_tibble()%>%
-  select(name, initial_count, count_current)->simulation_results
-```
-
-###2.5.3 Comparison of simulation results and tumor tissue expression values
-```{r}
-E9GE_mirnagenetumor%>%
-  dplyr::select(hgnc_symbol, GE_tumor)%>%
-  dplyr::inner_join((E9GE_mirnagenenormal%>%dplyr::select(hgnc_symbol, GE_normal)), by ="hgnc_symbol")%>%
-  inner_join(simulation_results, by= c("hgnc_symbol"="name"))%>%
-  distinct()
-```
-
-Actually, we have developed to provide a new approach mirna mediated regulation networks.
-This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and cohorent miRNA:target regulation approach.
+## REFERENCES

### 2019-09-17 article.tex
news to article
last meeting
article

@@
 }{}
+\usepackage[margin=1in]{geometry}
 \usepackage{hyperref}
@@
   \postdate{\par}
-    \date{10 Eylül 2019}
+    \date{15 Eylül 2019}
 
@@
 because of repression functionality depended on interaction factors of
-miRNA target pairs (answer: Hocam etkileşim faktörlerinden dolayı
-ekspresyonları yakın olsa bile baskılanma aktitesinin aynı
-sergilenmeyeceğini ifade etmeye çalıştım. Bu şekilde anlaşılıyor mu?).
-We applied the model to real sample consisting of breast cancer gene and
-miRNA expression dataset and experimental miRNA:target interaction
-dataset all generated via high throughput sequencing methods. A gene
-over-expressed in tumor tissue, namely \emph{ABCC1}, is used as
-perturbing element. We have observed that change in expression level of
-single gene in miRNA:target network is sufficient to perturb regulations
-in whole network, due to unforeseen and unpredicted regulation which are
-only visible when considered in network context. Therefore, this model
-helps unveiling the crosstalk between elements in miRNA:target network
-where abundance of target and sponge effect are taken into account. The
-model is scalable and can be plugged in with emerging miRNA effectors
-such as circRNAs. The model is available as R package
+miRNA target pairs. We applied the model to real sample consisting of
+breast cancer gene and miRNA expression dataset and experimental
+miRNA:target interaction dataset all generated via high throughput
+sequencing methods. A gene over-expressed in tumor tissue, namely
+\emph{SERPINE2}, is used as perturbing element. We have observed that
+change in expression level of single gene in miRNA:target network is
+sufficient to perturb regulations in whole network, due to unforeseen
+and unpredicted regulation which are only visible when considered in
+network context. Therefore, this model helps unveiling the crosstalk
+between elements in miRNA:target network where abundance of target and
+sponge effect are taken into account. The model is scalable and can be
+plugged in with emerging miRNA effectors such as circRNAs. The model is
+available as R package
 \href{https://github.com/selcenari/ceRNAnetsim}{ceRNAnetsim}.
@@
 \protect\hyperlink{ref-cao_predicting_2012}{2012}). Early studies have
-argued that 2-8 nt sequence, seed, located in miRNA 5'end bind to
-specific sequence located in 3'UTR of its target (Bartel
+argued that 2-8 nt sequence, seed, located in miRNA 5' end bind to
+specific sequence located in 3' UTR of its target (Bartel
 \protect\hyperlink{ref-bartel_micrornas_2004}{2004}; Lewis, Burge, and
@@
 studies, it has been shown that miRNAs can interact with targets via
-sequences located in regions such as 5'UTR or CDS (Hausser et al.
+sequences located in regions such as 5' UTR or CDS (Hausser et al.
 \protect\hyperlink{ref-hausser_analysis_2013}{2013}; Helwak et al.
@@
 \protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). These studies
-also showed that binding location could indicate functionality of
-miRNA:target interaction or be effective on abundance of targets(TOFIX
-abundance etkisi derken, bazen azaltabilir, bazen arttırabilir,
-manasında mı? answer: evet hocam anlaşılmıyorsa değiştirebilirim). It
-has been shown that miRNAs exhibit repressive activity via, 6-8 nt long
-sequence that is perfectly complementary with targets, seed region at
-the 5' end of miRNAs (Bartel
-\protect\hyperlink{ref-bartel_micrornas:_2009}{2009}; Grimson et al.
-\protect\hyperlink{ref-grimson_microrna_2007}{2007}). On the other hand,
-some researchers have reported that seed sequence of miRNA can have
-mismatches or bulged/wobble nucleotides (Chi, Hannon, and Darnell
-\protect\hyperlink{ref-chi2012alternative}{2012}) and may locate in
-region other than 5'end of miRNAs (Hafner et al.
+also showed that binding location either indicate functionality of
+miRNA:target interaction or affect abundance of targets. It has been
+shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence
+that is perfectly complementary with targets, seed region at the 5' end
+of miRNAs (Bartel \protect\hyperlink{ref-bartel_micrornas:_2009}{2009};
+Grimson et al. \protect\hyperlink{ref-grimson_microrna_2007}{2007}). On
+the other hand, some researchers have reported that seed sequence of
+miRNA can have mismatches or bulged/wobble nucleotides (Chi, Hannon, and
+Darnell \protect\hyperlink{ref-chi2012alternative}{2012}) and may locate
+in region other than 5'end of miRNAs (Hafner et al.
 \protect\hyperlink{ref-hafner_transcriptome-wide_2010}{2010}; Helwak et
@@
 other neighborhoods (\autoref{fig1}d), consequently have potential to
-effect whole network due to ``ripple effect''.
+affect whole network due to ``ripple effect''.
 
@@
 \caption{Schematic presentation of mechanism of network based model. a)
-In steady state, miRNAs (M: gray triangle) repress targets (G: circle
-coconut) according to proportion of their targets' expression. b) Two
-fold increase in transcript level of Gene2 (G2) acts as a trigger (shown
-with red in figure). c) Distribution of miRNA1 (M1) changes. d) The
-change at expression (shown with light red because minor changes are
-occured on their expression values.) of common target effects changes of
-proportional distribution of miRNA2 (M2). Expression values are rounded
-to integers for simplicity}\label{fig1}
+In steady state, miRNAs (M: triangles) repress targets (G: circles)
+according to proportion of target expression level. b) Two fold increase
+in transcript level of Gene2 (G2) acts as a trigger (shown in red). c)
+Distribution of miRNA1 (M1) changes. d) The change at expression of
+common target affects changes of proportional distribution of miRNA2
+(M2). Expression values are rounded to integers for simplicity. Gray
+values on edges indicate initial distribution, red values indicate
+increase and blue values indicate decrease in distribution. Shades of
+circles indicate different levels of increase.}\label{fig1}
 }
@@
 miRNA expression (\(C_m\)) in local neighborhood is distributed among
-targets using individual gene expression levels (\(C_{gi}\)) (\#FIX
-isn't \(C_g\) gene expressions in the group is equal to total of
-\(C_{gi}\) in group? answer: its my false. now its corect.). For the
+targets using individual gene expression levels (\(C_{gi}\)). For the
 genes targeted by multiple miRNAs, cooperative activity of miRNAs on a
@@
 calculating overall miRNA activity. We classified factors into two
-categories. Factors effecting binding determine interaction between
-miRNA and target and they alter amount of miRNA sequestered to target.
-Degradation efficiency factors assign amount of degradated target amount
-in miRNA:target pairs.(\#FIX fix this sentence: answer:). In other
-words, binding factors exert their influence before or during binding,
-efficiency factors exert their influence after binding. In the
-literature, binding free energy (Cao and Chen
-\protect\hyperlink{ref-cao_predicting_2012}{2012} ; Helwak et al.
+categories. Binding factors determine interaction between miRNA and
+target and they alter amount of miRNA sequestered to target. Efficiency
+factors dictate degradation efficieny of miRNA on its target and define
+amount of target to be degraded. In other words, binding factors exert
+their influence before or during binding, efficiency factors exert their
+influence after binding. In the literature, binding free energy (Cao and
+Chen \protect\hyperlink{ref-cao_predicting_2012}{2012} ; Helwak et al.
 \protect\hyperlink{ref-helwak_mapping_2013}{2013}) and seed type (Werfel
@@
 miRNA:target interactions are described as factors effecting binding
-affinity. Efficiency factors determine how many of miRNA:target
-complexes will result in inhibition and binding region on target
-drastically effect miRNA degradation efficiency (Hausser et al.
+affinity. Also, binding region (e.g, 5'UTR, 3'UTR, CDS) was shown to
+drastically affect miRNA degradation efficiency (Hausser et al.
 \protect\hyperlink{ref-hausser_analysis_2013}{2013}; Helwak et al.
-\protect\hyperlink{ref-helwak_mapping_2013}{2013}). Both binding and
-efficiency factors are normalized to their maximum values and scaled to
-{[}0,1{]} interval. The normalized values of factors take into account
-to determine binding activity and miRNA efficiency on targets
-(\autoref{fig2}). Binding affinities (activity, \(Eff\)) of miRNAs on
-each individual gene are calculated as shown in equation \eqref{eq:3};
-\(C_m\), miRNA expression in the group; \(C_{gi}\), individual gene
-expression; gi, individual gene (\autoref{fig2}c). \begin{equation}
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}). Since binding or
+degradation efficiency from literature have various and broad ranges,
+both binding and efficiency factors are normalized to their maximum
+values and scaled to {[}0,1{]} interval. The normalized values of
+factors are used to determine binding activity and miRNA efficiency on
+targets (\autoref{fig2}). Binding affinities (activity, \(Eff\)) of
+miRNAs on each individual gene are calculated as shown in equation
+\eqref{eq:3}; \(C_m\), miRNA expression in the group; \(C_{gi}\),
+individual gene expression; gi, individual gene (\autoref{fig2}c).
+\begin{equation}
 Eff_{gi}= C_m \times E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}/(\sum_{1}^{i} E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}) \tag{3}\label{eq:3}
@@
 region efficiency coefficient; \(E^\prime\), normalized values of energy
-coefficient.}\label{fig2}
+coefficient. Numbering on edges match the pair order in
+\autoref{tab:one}}\label{fig2}
 }
@@
 
-\iffalse
-
-\%\% (\#TODO the equation is not considering binding factors, right?. We
-should explain the calculation made in simulate function here, in simple
-sentences. the function looks like this: \%\%
-input\_graph\%\textgreater{}\% \%\%
-tidygraph::activate(edges)\%\textgreater{}\% \%\%
-group\_by(to)\%\textgreater{}\% \%\% mutate(mirna\_count\_per\_comp =
-mirna\_count\_current\emph{comp\_count\_current}afff\_factor/sum(comp\_count\_current\emph{afff\_factor),
-mirna\_count\_per\_comp = ifelse(is.na(mirna\_count\_per\_comp), 0,
-mirna\_count\_per\_comp))\%\textgreater{}\% \%\%
-ungroup()\%\textgreater{}\% \%\% mutate(effect\_pre = effect\_current,
-effect\_current =
-mirna\_count\_per\_comp}degg\_factor)\%\textgreater{}\% \%\%
-group\_by(from)\%\textgreater{}\% \%\% mutate(comp\_count\_pre =
-ifelse(comp\_count\_current\textless{} 0, 1, comp\_count\_current),
-comp\_count\_current = comp\_count\_pre -
-(sum(effect\_current)-sum(effect\_pre)), comp\_count\_current =
-ifelse(comp\_count\_current\textless{} 0, 1,
-comp\_count\_current))\%\textgreater{}\% \%\%
-ungroup()\%\textgreater{}\% \%\% mutate(comp\_count\_list =
-pmap(list(comp\_count\_list, comp\_count\_current), c), effect\_list =
-pmap(list(effect\_list, effect\_current), c), mirna\_count\_list =
-pmap(list(mirna\_count\_list, mirna\_count\_current), c)) \%\% ) \fi
-
-I shown in figure and explained at following. I didn't fully understand
-whether it is enough or not.
-
-After miRNA binds to its target, but might not repress to bound target.
-The functionality of bound miRNA on target depends on efficiency factors
-like region that is binding sequence of miRNA on its target. Exact
-repression efficiency of miRNA is calculated according to equation
-\eqref{eq:4} (\autoref{fig2}d); \(RE^\prime_{gi}\), normalized values of
-region efficiency coefficient between miRNA and gene. The cooperative
-repression activity of miRNAs to their common targets is figured out as
-shown in \autoref{fig2}e.
+Not all miRNA:target binding events result in degradation of target. The
+degradation of target by bound miRNA depends on efficiency factors such
+as binding region. Exact repression efficiency of miRNA is calculated
+according to equation \eqref{eq:4} (\autoref{fig2}d);
+\(RE^\prime_{gi}\), normalized values of region efficiency coefficient
+between miRNA and gene. The cooperative repression activity of miRNAs to
+their common targets is figured out as shown in \autoref{fig2}e.
 
@@
 miRNA and genes in tumor and normal tissue of single patient are
-retrieved from The Cancer Genome Atlas
-\href{https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga}{TCGA}(\#TODO).
+retrieved from TCGA Research Network: \url{https://www.cancer.gov/tcga}.
 High-throughput experimental datasets which are provided miRNA:gene
@@
 miRNA and gene expression datasets via miRNA:target gene dataset
-retrieved from \ldots{} (\#FIX name of database and its citation)
-answer-\textgreater{}suplementary datasets of (Helwak et al.
+retrieved from published datasets (Helwak et al.
 \protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
@@
 description of network construction and its code is available in
-Supplementary data (\#TODO burada link nasıl olacak, word dosyası veya
-PDF dosyası ismi mi vermemiz mi gerekiyor, yoksa sadece Sup Data
-denilmesi yeterli mi? ilgili dosya: TCGA\_E9-A1N5\_article.Rmd). ABCC1
-gene, over-expressed in tumor tissue, was selected as trigger for
-simulation of integrated dataset. After simulation of network, we have
-compared simulation results and tumor tissue expression levels.
+Supplementary Materials and Methods. We have determined the most
+effective nodes in network and fold changes of genes as compared tumor
+and normal tissue samples of the same patient.As a result of comparison
+of these two datasets, we selected a gene , SERPINE2, as trigger for
+simulation of model (see Supplementary Materials and Methods).
 
@@
 
-\hypertarget{networks-with-single-factor}{%
-\subsubsection{Networks with single
-factor}\label{networks-with-single-factor}}
+\hypertarget{network-with-mirnacerna-expression-level}{%
+\subsubsection{Network with miRNA:ceRNA expression
+level}\label{network-with-mirnacerna-expression-level}}
 
@@
 interact with another miRNA of the trigger gene change in relation to
-their expression. However, expression level change of common target (G4)
-in network has been observed different from G1 that has the same
-expression value with G4 target at initial conditions because more than
-one miRNA repress this target (\autoref{fig1}d) (\#FIX last sentence
-needs some clarification, after ``but cooperative\ldots{}'').
+their expression. However, changes in expression levels of G4 and G1
+differ although they have same initial conditions. This is due to fact
+that G4 is targeted by two miRNAs, M1 and M2 (\autoref{fig1}d).
 
@@
 Sharp \protect\hyperlink{ref-bosson_endogenous_2014}{2014}; Denzler et
-al. \protect\hyperlink{ref-denzler_assessing_2014}{2014}). \href{}{}
+al. \protect\hyperlink{ref-denzler_assessing_2014}{2014}).
 
@@
 
-\hypertarget{competing-endogenous-rnatarget-networks-based-on-interaction-factors}{%
-\subsubsection{Competing Endogenous RNA:Target networks based on
-interaction
-factors}\label{competing-endogenous-rnatarget-networks-based-on-interaction-factors}}
-
-\href{}{\#FIX}: \textless{}\textgreater{} ``title should be more
-informative''-\textgreater{} answer.
+\hypertarget{cernatarget-networks-based-on-interaction-factors}{%
+\subsubsection{ceRNA:Target networks based on interaction
+factors}\label{cernatarget-networks-based-on-interaction-factors}}
 
@@
 binding and repression have changed.
+
+{[}FIXED by selcen{]} When the model run based on expression values of
+miRNAs and ceRNAs (genes), proportional distribution of miRNA on targets
+is determined in accordance with \eqref{eq:1, eq:2}. In that case, the
+system is observed as shown in \autoref{fig1}a at steady-state. However,
+if interaction factors are taken into account in the approach, specified
+interaction factors (i.e, binding factors such as seed type and energy)
+affect distribution of miRNA on targets as calculated in \eqref{eq:3}
+(shown in \autoref{fig2}c). After distributed miRNA efficiency was
+re-calculated with degradation factor(s) (i.e region effect) according
+to \eqref{eq:4}, steady-state condition of network is obtained (shown in
+\autoref{fig2}d and \autoref{fig3}a). For instance, proportional
+distribution of G1:M1 interaction in \autoref{fig1}a is calculated lower
+than the same pair (G1:M1) in the other approach that includes
+interaction factors (\autoref{fig2}c). But, a decrease is observed in
+repression efficiency of M1 miRNA on G1 in interaction factor based
+network approach (\autoref{fig2}d and \autoref{fig3}a).
 
@@
 When model is triggered with two fold increase in expression level of
-common target (G4), the changes of other gene expressions have observed
-more prominent \href{}{}. Furthermore, change in expression level of
-target gene that has strong miRNA repression efficiency resulted in
-evident perturbation in network. On the other hand, it was observed that
-Gene2 was weakly affected from change of Gene4 expression because of its
-weak interaction factors. Since perturbation efficiency of each gene is
-different, we have developed a function which screens each gene in the
-network for their perturbation efficiencies. \textbf{(see details in
-Supplementary File(fig1\_2app.Rmd)}. On the other hand, when we applied
-the method on the minimal dataset, Gene4 has been found to be the most
-efficient element in terms of number of perturbed elements and miRNA2
-(M2) has been found to be causing the highest mean in expression
-changes. {[}{]}{[}\#TODO2{]}
+common target (G4), more prominent changes were observed in gene
+expression levels in network compared to changes caused by G2 increase.
+Furthermore, change in expression level of target gene that has strong
+miRNA repression efficiency resulted in evident perturbation in network.
+On the other hand, it was observed that Gene2 was weakly affected from
+change of Gene4 expression because of its weak interaction factors.
+Since perturbation efficiency of each gene is different, we have
+developed a function which screens each gene in the network for their
+perturbation efficiencies. \textbf{(see details in Supplementary
+Materials and Methods}. On the other hand, when we applied the method on
+the minimal dataset, Gene4 has been found to be the most efficient
+element in terms of number of perturbed elements and miRNA2 (M2) has
+been found to be causing the highest mean in expression changes.
 
-We used our model to simulate a real dataset which contains thousands of
-genes and hundreds of miRNAs. Our model can successfully simulate
-perturbations in such large network despite complex behaviors and
-struggle to reach steady-state \textbf{(see details in Supplementary
-File(TCGA\_E9-A1N5\_article.Rmd)}. Simulations show that change in
-expression level of single gene has potential to effect whole network,
-perturbing almost all nodes. These observations are in accordance with
-competing endogenous RNA hypothesis where genes targeted with many
-common miRNAs subsequently transmit perturbation to neighboring groups.
+\hypertarget{simulation-of-large-and-realistic-network}{%
+\subsubsection{Simulation of large and realistic
+network}\label{simulation-of-large-and-realistic-network}}
+
+Apart from tools predicting miRNA:target pairs, there are databases that
+provide experimentally validated miRNA:target interaction datasets. The
+reliability in these datasets range from weak (e.g, western blot) to
+strong (e.g, CLiP, CLASH). High-throughput methods ensure handling of
+AGO (Argounate Protein) interacted miRNAs (CLiP) and their targets as
+well (CLEAR-CLiP, CLASH). We constructed a large ceRNA network utilizing
+high-throughput datasets containing chimeric reading of miRNA:target
+pairs (Helwak et al. \protect\hyperlink{ref-helwak_mapping_2013}{2013};
+Moore et al. \protect\hyperlink{ref-moore_mirnatarget_2015}{2015}) which
+expose exact complementary seed regions.
+
+Although these experimental sources include exact binding sites, they do
+not imply that each miRNA binding event is functional (Liu and Wang
+\protect\hyperlink{ref-liu2019prediction}{2019}). Seed type, energy,
+binding region affects functionality of miRNA drastically. In order to
+construct realistic network, we incorporated such factors in the
+network. Energy values in miRNA:target pairs represented by
+high-throughput studies (Helwak et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
+\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}), seed type
+(Grimson et al. \protect\hyperlink{ref-grimson_microrna_2007}{2007};
+Bartel \protect\hyperlink{ref-bartel_micrornas:_2009}{2009}; Betel et
+al. \protect\hyperlink{ref-betel2010comprehensive}{2010}) and binding
+region effect (Hausser et al.
+\protect\hyperlink{ref-hausser_analysis_2013}{2013}; Helwak et al.
+\protect\hyperlink{ref-helwak_mapping_2013}{2013}) data were collected
+and integrated into network.
+
+The network was finalized by integrating gene and miRNA expression
+levels from TCGA determined by RNA-Seq. Our model can successfully
+simulate perturbations in such large network despite complex behaviors
+and struggle to reach steady-state \textbf{(see details in Supplementary
+Materials and Methods)}. Simulations show that change in expression
+level of single gene has potential to affect whole network, perturbing
+almost all nodes. These observations are in accordance with competing
+endogenous RNA hypothesis where genes targeted with many common miRNAs
+subsequently transmit perturbation to neighboring groups.
+
+\href{}{\#TODO} We have observed that the results of simulation with
+SERPINE2 gene, a serine protease inhibitor, one of the overexpressed
+genes of breast cancer (Yang et al.
+\protect\hyperlink{ref-yang2018expression}{2018}; Candia et al.
+\protect\hyperlink{ref-candia2006protease}{2006}) is not convenient with
+tumor tissue expression values, when we compared the tumor gene
+expression dataset retrieved from TCGA and simulation results obtained
+via our approach. Simulation results were not expected to be consistent
+with tumor tissue expression values because large number of abnormal
+regulations were observed in these tissues. So, it might not said that a
+single gene is responsible for entire regulations in cancer cells.
+Besides, the all other factors such as up/down regulation of miRNAs or
+other genes were ignored. Therefore, it would be more useful to test our
+approach with an ideal dataset that includes entire informations of gene
+expression values before and after the regulation of a known gene at
+existing tissue, and miRNA expression values of this tissue at initial
+conditions.
+
+{[}fixed by selcen{]}
+
+\hypertarget{discussion}{%
+\subsection{Discussion}\label{discussion}}
 
@@
 that distant ceRNAs can interact with each other via indirect links, and
-the interactions are effected depending on distance between ceRNAs or
+the interactions are affected depending on distance between ceRNAs or
 topological features of network (Nitzan et al.
@@
 miRNA based approach may not be suitable for understanding regulations
-of ceRNA interactions.
-
-{[}\#Add{]}More recently, Silveira et al.~have developed miRmapper
-package for using with R programme (Silveira et al.
-\protect\hyperlink{ref-da2018mirmapper}{2018}). Researchers have
-utilized the adjacency matrix to associate miRNAs using differentially
-expressed genes and found similarities of miRNAs with this approach.
-Authors who utilized the approach in dataset of bladder cancer cell line
-have identified significant miRNAs in network, and miRNAs which work
-synergistically.
-
-{[}\#Add{]} Today, there are many of tools that provide predicted or
-experimentally validated miRNA:target interaction dataset. The datasets
-that contain weak, proteomic analysis like western blot, and strong,
-crosslinking and immunoprecipitation CLiP, experimentally evidence are
-more reliable sources than prediction datasets. While in experimentally
-weak supported datasets it is not known that indirect or direct target
-of miRNAs in cell, high-throughput methods ensure handling of AGO
-(Argounate Protein) interacted miRNAs. For this reason we have preferred
-to utilize high-throughput datasets in our application. Especially, we
-collected datasets that depend on chimeric (Helwak et al.
-\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
-\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}) reading of miRNA:
-target pairs from among these datasets because they contain exact
-complementary seed regions.
-
-{[}\#Add{]} Although these experimental sources include exact binding
-sites, they do not provides which miRNA is functional or not (Liu and
-Wang \protect\hyperlink{ref-liu2019prediction}{2019}). We have used the
-features of bound miRNA:target pairs and scaled these parameters inside
-groups that is consisted according to, ceRNA (Competing Endogenous RNA),
-genes targeted by the same miRNA. In scaling process, we specified the
-interaction factors as numeric values in accordance with previous
-studies. Energy values in miRNA:target pairs are represented by
-high-throughput studies (Helwak et al.
-\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
-\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}) which are utilized
-in this study. On the other hand, we have specified the other
-interaction factors, seed type and location of binding region on the
-target, as numeric values based on the previous studies.(Grimson et al.
-\protect\hyperlink{ref-grimson_microrna_2007}{2007}) have compared the
-seed types' effect on target repression with few miRNA had canonical
-seed pairing in their study. Additionally, (Bartel
-\protect\hyperlink{ref-bartel_micrornas:_2009}{2009}) and (Betel et al.
-\protect\hyperlink{ref-betel2010comprehensive}{2010}) have studied on
-functional and non-functional seed interactions. Based on results of
-these studies we have arranged seed types of miRNA:target interactions
-as numeric values. We also have redefined location of binding region on
-the target as numeric values, based on studies of (Hausser et al.
-\protect\hyperlink{ref-hausser_analysis_2013}{2013}) and (Helwak et al.
-\protect\hyperlink{ref-helwak_mapping_2013}{2013}). With this process,
-we have handled this entegrated dataset in context of competitor
-behaviours and functionality of interactions.
-
-{[}\#Add{]}The main advantage of the method, ceRNAnetsim package, is
-providing various functions which contain optionally arguments. So,
-users can run the functions in accordance with their own data. For
-example, the functions can work without interaction factors or with a
-simple interaction factor that is specified by user. Furthermore, it can
-also be operated according to a different parameter(s) provided by the
-user.In this study, we have preferred to run model, shown in
-\emph{Multifactorial calculations in miRNA:target network} section.
-
-{[}\#TODO answer{]} We have observed that the results of simulation with
-ABCC1 gene, \emph{Multidrug resistance-associated protein 1}, which is
-one of the most significant factor to develop resistance against
-chemotherapotic agents(Atalay, Demirkazik, and Gunduz
-\protect\hyperlink{ref-atalay2008role}{2008}; Lu et al.
-\protect\hyperlink{ref-lu2015microrna}{2015}; Atalay et al.
-\protect\hyperlink{ref-atalay2006multidrug}{2006}) is not convenient
-with tumor tissue expression values, when we compared the tumor gene
-expression dataset retrieved from
-\href{https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga}{TCGA}
-and simulation results obtained via our approach. Simulation results
-were not expected to be consistent with tumor tissue expression values
-because large number of abnormal regulations were observed in these
-tissues. So, it might not said that a single gene is responsible for
-entire regulations in cancer cells. Besides, the all other factors such
-as up/down regulation of miRNAs or other genes were ignored. Therefore,
-it would be more useful to test our approach with an ideal dataset that
-includes entire informations of gene expression values before and after
-the regulation of a known gene at existing tissue, and miRNA expression
-values of this tissue at initial conditions.
+of ceRNA interactions. Additionally, miRmapper package (Silveira et al.
+\protect\hyperlink{ref-da2018mirmapper}{2018}) utilizes an adjacency
+matrix to associate miRNAs using differentially expressed genes and
+identifies significant miRNAs in network.
 
@@
 structure is flexible and can integrate additional regulation elements
-as they become available. In the future, with developments in
-experimental techniques about miRNAs:target interaction dynamica, more
-accurate results can be obtained from our model resulting in better
+as they become available.
+
+\hypertarget{conclusion}{%
+\subsection{Conclusion}\label{conclusion}}
+
+The main advantage of ceRNAnetsim package is providing various functions
+which can be run in accordance with user data. For example, the
+functions can work without interaction factors or with a simple
+interaction factor that is specified by user. Furthermore, it can also
+be operated according to a different parameter(s) provided by the
+user.In this study, we have preferred to run model, shown in
+\emph{Multifactorial calculations in miRNA:target network} section.
+
+In the future, with developments in experimental techniques about
+miRNAs:target interaction dynamics, more accurate results can be
+obtained from our model. Also, our model can integrate emerging miRNA
+effectors such as circRNAs. Consequently, our package may lead to better
 understanding and predictions of abnormal regulations and pathways
@@
 \url{https://doi.org/10.1016/j.ymeth.2015.04.012}.
+
+\leavevmode\hypertarget{ref-candia2006protease}{}%
+Candia, Britny J, William C Hines, Christopher M Heaphy, Jeffrey K
+Griffith, and Robert A Orlando. 2006. ``Protease Nexin-1 Expression Is
+Altered in Human Breast Cancer.'' \emph{Cancer Cell International} 6
+(1): 16.
 
@@
 
+\leavevmode\hypertarget{ref-yang2018expression}{}%
+Yang, Ying, Xiangke Xin, Xing Fu, and Danmei Xu. 2018. ``Expression
+Pattern of Human Serpine2 in a Variety of Human Tumors.'' \emph{Oncology
+Letters} 15 (4): 4523--30.
+
 \end{document}

### 2019-09-17 article.Rmd
revision of SERPINE changes

@@
 Detailed description of network construction and its code is available in Supplementary Materials and Methods.
-We have determined the most effective nodes in network and fold changes of genes as compared tumor and normal tissue samples of the same patient.As a result of comparison of these two datasets, we selected a gene , SERPINE2, as trigger for simulation of model (see Supplementary Materials and Methods).
+SERPINE', one of the most effective nodes, has been used as trigger in ceRNA network simulation in breast cancer patient network.
+Relevant steps and codes are available at  Supplementary Materials and Methods.
 
@@
 [#TODO]: <> "figure1, 2 ve 3 uzerinden farkli dagilim ve degradasyon açıklaması buraya. G2 uzerinden olmasa iyi olur"
-[FIXED by selcen]
+
 When the model run based on expression values of miRNAs and ceRNAs (genes), proportional distribution of miRNA on targets is determined in accordance with \eqref{eq:1, eq:2}.
@@
 
-[#TODO] We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach.
+We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach.
 Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues.
@@
 Therefore, it would be more useful to test our approach with an ideal dataset that includes entire informations of gene expression values before and after the regulation of a known gene at existing tissue, and miRNA expression values of this tissue at initial conditions.
-
-[#TODO]: <> "run perturbation screening on tcga set, find a gene then couple sentences about that gene"
-[fixed by selcen]
 

### 2019-09-17 Supplementary_mat_met.Rmd
sup met edit and spell check

@@
 
-# 1. Defined functions for ceRNA models and workflow of method
+# 1. Functions defined for ceRNA models and workflow of method
 We defined the functions that can be used with R programming.
@@
 
-**Convertion of dataset**: priming_graph function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
+**Conversion of dataset**: `priming_graph()` function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
 If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
@@
 This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
-It is assumed that the initial target amounts in the dataset is observed after the reppressive activity of miRNAs in steady-state.
+It is assumed that the initial target amounts in the dataset is observed after the repressive activity of miRNAs in steady-state.
 
-![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow1.png){#fig1}
+![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow1.png){#figs1}
 
@@
 
-**Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficieny coefficients are calculated according to this acceptance.
-In the network that is found in steady-state conditions, the change is applied to the graph object for disturbtion of steady-state.
-To provide the disturbtion in the network the workflow offer two methods: update_variables and update_how.
+**Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficiency coefficients are calculated according to this acceptance.
+In the network that is found in steady-state conditions, the change is applied to the graph object for distribution of steady-state.
+To provide the distribution in the network the workflow offer two methods: update_variables and update_how.
 The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
@@
 
-**Updating current values of variables**: After variables updating in edge varibles, these are carried to node variables.
+**Updating current values of variables**: After variables updating in edge variables, these are carried to node variables.
 Current and previous values of variables are stored as node variables with update_variables function.
@@
 
-The node elements in the dataset are handled as two type; repressives (miRNAs) and competings (targets).
+The node elements in the dataset are handled as two type; repressive (miRNAs) and competing (targets).
 It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used.
@@
 The functions that are used in the approach are developed with R programming so as can be used with other packages.
-These are can be found in the github repository [ceRNAnetsim github page](https://github.com/selcenari/ceRNAnetsim) and improved with contributions of others.
+These are can be found in the Github repository [ceRNAnetsim Github page](https://github.com/selcenari/ceRNAnetsim) and improved with contributions from others.
 
@@
 
-See Figure S1 in Supplementary Tables file.
+See Table S1 in Supplementary Tables file.
 
@@
 
-- 2. We have obtained graph after change on Gene2 expression as followings:
+- 2. We have obtained graph after change on Gene2 expression as following:
 
@@
 
-- 3. We have determined regulations after Gene2 Upregulation:
+- 3. We have determined regulations after Gene2 upregulation:
 
-```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.cap = "Sequential iteration of minsampdata", fig.subcap= c("First response of system to Gene2 upregulation (2nd iteration)", "Spreading of perturbation on system (3th iteration)")}
+```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of minsamp data.
+a) First response of system to Gene2 upregulation (2nd iteration).
+b) Spreading of perturbation on system (3th iteration)", fig.subcap= NULL}
 
@@
 
-We have made the same analysis in present of interaction factors.
+We have made the same analysis in presence of interaction factors.
 
@@
 
-```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.cap = "Sequential iteration of minsampdata with interaction factors", fig.subcap= c("First response of system to Gene2 upregulation (2nd iteration)", "Spreading of perturbation on system (3th iteration)")}
+```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.cap = "Sequential iteration of minsamp data with interaction factors", fig.subcap= c("First response of system to Gene2 upregulation (2nd iteration)", "Spreading of perturbation on system (3th iteration)")}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
@@
 
-When the graphs which were resulted from analyses were examined, it was observed that behaviours were same.
+When the graphs which were resulted from analyses were examined, it was observed that behaviors were same.
 But, when the results were analysed in terms of expression values, the regulation differences can be observed.
@@
 
-There are hundreds of defined miRNAs for human, so this results in presence of common targets of miRNAs in cells.
-Therefore, we have analysed perturbation efficiency of common target in *minsamp* dataset.
+Genes targeted by multiple miRNAs (referred to as "common target") are of special interest since they are subject to cooperative effect.
+Also, they perturb more than one neighborhood.
+In our small dataset, minsamp, Gene4 is regulated by two miRNAs.
+Let's simulate perturbation effects triggered by Gene4.
 
-```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.cap = "Perturbation of Gene4 on minsampdata with interaction factors", fig.subcap= c("1st iteration after Gene4 upregulation", "2nd iteration after Gene4 perturbation")}
+```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.cap = "Perturbation of Gene4 on minsamp data with interaction factors", fig.subcap= c("1st iteration after Gene4 upregulation", "2nd iteration after Gene4 perturbation")}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
@@
 
-## Determination of perturbation efficiencies efficiencies of elements in system.
+## Determination of perturbation efficiencies of elements in system.
 
@@
 `TCGAbiolinks` package provides to obtain data for whole number of given barcode(s) at once.
-But, we preferred to download them seperately to show datasets.
+But, we preferred to download them separately to show datasets.
 
@@
 + The datasets were combined.
-+ We commited the interaction factors as numeric values according to previous studies. (We added the interaction factors and their numeric values at Supplementary tables S2-3)
++ We committed the interaction factors as numeric values according to previous studies. (We added the interaction factors and their numeric values at Supplementary tables S2-3)
+
+[#TODO çok karmaşık komutlar için ayrı bir klasör hazırlanabilir. burayada link verilebilir: "experimentalmirnagene" tablosu repodaki X.R script ile elde edilmiştir" gibi]
 
@@
 ```
+
+[#TODO SERPINE2]
+
 Secondly, we have determined the most important nodes of network.
@@
 
-The dynamics of the approach are shown in package viggnettes [link](https://selcenari.github.io/ceRNAnetsim/articles/convenient_iteration.html).
+The dynamics of the approach are shown in package  [vignettes](https://selcenari.github.io/ceRNAnetsim/articles/convenient_iteration.html).
 
@@
 Actually, we have developed to provide a new approach mirna mediated regulation networks.
-This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and cohorent miRNA:target regulation approach.
+This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and coherent miRNA:target regulation approach.
 

### 2019-09-17 Supplementary_mat_met.Rmd
eval

--- 2019-09-17 Supplementary_mat_met.Rmd
@@
 
-```{r, fig.cap= "Percentage of affected nodes of each iteration for SERPINE2 Gene"}
+```{r, fig.cap= "Percentage of affected nodes of each iteration for SERPINE2 Gene", eval=FALSE}
 as.data.frame(E9GE_mirnagenenormal)%>%
@@
 
-```{r}
+```{r, eval=FALSE}
 as.data.frame(E9GE_mirnagenenormal)%>%
@@
 
-```{r}
+```{r, eval=FALSE}
 
@@
 ### 2.5.3 Comparison of simulation results and tumor tissue expression values
-```{r}
+```{r, eval=FALSE}
 E9GE_mirnagenetumor%>%

### 2019-09-17 article.Rmd

@@
 Binding factors determine interaction between miRNA and target and they alter amount of miRNA sequestered to target.
-Efficiency factors dictate degradation efficieny of miRNA on its target and define amount of target to be degraded.
+Efficiency factors dictate degradation efficiency of miRNA on its target and define amount of target to be degraded.
 In other words, binding factors exert their influence before or during binding, efficiency factors exert their influence after binding.
@@
 Detailed description of network construction and its code is available in Supplementary Materials and Methods.
-SERPINE', one of the most effective nodes, has been used as trigger in ceRNA network simulation in breast cancer patient network.
+SERPINE2, one of the most effective nodes, has been used as trigger in ceRNA network simulation in breast cancer patient network.
 Relevant steps and codes are available at  Supplementary Materials and Methods.
@@
 However, in presence of interaction factors in miRNA:target interaction network (shown in \autoref{tab:one}), miRNAs are distributed according to target:total target ratio and affinity factors in first like shown in \autoref{fig2}c.
-After affinity mediated propotional distribution of miRNA expression, degradation factor is considered to specify count of repressive miRNA in pairs because entire bound miRNA target pairs might not be resulted with degradation (It is figured out shown in \autoref{fig2}d).
+After affinity mediated proportional distribution of miRNA expression, degradation factor is considered to specify count of repressive miRNA in pairs because entire bound miRNA target pairs might not be resulted with degradation (It is figured out shown in \autoref{fig2}d).
 For this reason, miRNA1 (M1) has more weak repressive effect (shown as edge variables) on target Gene2 \autoref{fig3}a in comparison with miRNA1 (M1):Gene2 (G2) interaction in \autoref{fig1}a.
-On the other hand, it is possible that taking into accont of these interaction factors could cause increasing miRNA repression activity like in miRNA2:Gene4 interaction \autoref{fig3}a.
+On the other hand, it is possible that taking into account of these interaction factors could cause increasing miRNA repression activity like in miRNA2:Gene4 interaction \autoref{fig3}a.
 When expression of Gene2 (G2) increased, expression values of all genes also changed differentially because of contribution of efficiency factors (\autoref{fig3} b,c,d).

### 2019-09-22 Supplementary_additional_file.Rmd
additional files

---
title: "Supplementary Experimental Data File"
author: "Selcen Arı"
date: "22 09 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
bibliography: ref.bib
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}     \renewcommand{\thefigure}{S\arabic{figure}}}
  \usepackage{subfig}
---

```{r, include=FALSE, message=FALSE, warning=FALSE}

library(tidygraph)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(readxl)

```

# Arrangement of CLASH dataset [@helwak_mapping_2013]

CLASH dataset was retrieved from [PubMed](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3650559/)

```{r, eval=FALSE}

clashelwak <- read.table("mmc1.txt", comment.char = "#", header = TRUE, skip = 1)

#hg19
```

## Query of Human Genome 19.

```{r,  eval=FALSE}
#HG19
listEnsemblArchives()
listMarts(host = 'http://grch37.ensembl.org')
ensemblgrch37 = useMart(host='http://grch37.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL',
                    dataset='hsapiens_gene_ensembl')
hg19 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene_id", "strand" ), mart = ensemblgrch37)

```

## Adding miRNA and gene information

```{r,  eval=FALSE}

clashelwak <- clashelwak%>%
 separate(microRNA_name, c("Barcode", "Database", "mirna_name", "type"), sep = "_")%>%
  separate(mRNA_name, c("Ensembl_Gene_Id", "Ensembl_Transcript_Id", "Hugo_Symbol", "mRNA_Type"), sep = "_")

```

MiRNA releases are obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) is utilised.

```{r, eval=FALSE}
read.table("mirbasehg38.txt", comment.char = "#")%>%
  filter(V3 != "miRNA_primary_transcript")%>%
  separate(V9, c("ID", "Alias", "Name", "Precusor"), sep = ";")%>%
  mutate(ID = substr(ID, 4, length(ID)), Alias = substr(Alias, 7, length(Alias)), Name = substr(Name, 6, length(Name)), Precusor = substr(Precusor, 14, length(Precusor)))%>%
  dplyr::select(chr= V1, start = V4, end = V5, strand = V7, ID, Alias, Name, Precusor)->mirbasehg38

```

CLASH dataset is published in miRBase erlease 15 and Human Genome 19 version.

```{r, eval=FALSE}
read_tsv("mirna_mature.txt", col_names = FALSE)%>%
  filter(startsWith(X2, "hsa"))%>%
  dplyr::select(mirna_ID = X2, mirbase_ID = X3)%>%
  inner_join(mirbasehg38%>% dplyr::select(ID, Name), by= c("mirbase_ID"= "ID"))%>%
  dplyr::select(mirbase_ID, Name)%>%
  distinct()%>%
  inner_join(clashelwak, by = c("mirbase_ID"= "Barcode"))%>%
  dplyr::select(Name, miRNA_seq, Ensembl_Gene_Id, Ensembl_Transcript_Id, Hugo_Symbol, mRNA_seq_extended, chimeras_decompressed, seed_type, seed_basepairs,folding_class, seq_ID, folding_energy, X5.UTR, CDS, X3.UTR)%>%
  inner_join(hg19, by = c("Ensembl_Gene_Id"= "ensembl_gene_id", "Ensembl_Transcript_Id"="ensembl_transcript_id", "Hugo_Symbol"="hgnc_symbol"))%>%
  mutate(region1 = ifelse(X5.UTR == "1", "5UTR", " "), region2= ifelse(X3.UTR == "1", "3UTR", " "),region3= ifelse(CDS =="1", "CDS", " "))%>%
  unite(region, c(region1, region2, region3), sep = "||")%>%
  dplyr::select(chromosome_name, start_position, end_position,strand, Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, mRNA_seq_extended, Name, miRNA_seq, seq_ID, seed_type,seed_basepairs, folding_class, folding_energy, region)%>%
  as_tibble() -> clashelwakfinal

```

## Converting CLASH data to human genome 38 build.

```{r,  eval=FALSE}
lift19 <- clashelwakfinal%>%
  dplyr::select(1,2,3)%>%
  unite(start_end, c("start_position", "end_position"), sep = "-")%>%
  mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
  unite(chromosome_name, c("Chromosome","start_end"), sep = ":")

write_tsv(lift19, "lift19.txt")

# Lift over process is made via UCSC liftover tool. (https://genome.ucsc.edu/cgi-bin/hgLiftOver)

lift19_del <- read_tsv("deleted_lift19.txt")
colnames(lift19_del)[1] <- "chromosome_loc"

lift19_del <- lift19_del%>%
  dplyr::filter(startsWith(chromosome_loc, "chr"))%>%
  separate(chromosome_loc, c("Chr", "End"), "-", remove = TRUE)%>%
  separate(Chr, c("Chr", "Start"), ":", remove = TRUE)

lift19_del$Start <- as.numeric(lift19_del$Start)
lift19_del$End <- as.numeric(lift19_del$End)

clashelwakfinal <-clashelwakfinal%>%
  mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
  dplyr::anti_join(lift19_del, by = c("Chromosome"="Chr", "start_position"="Start", "end_position"="End"))

hg38clash <- read.delim("hg38clashcomp.txt", header = FALSE)

clashelwakfinal <- clashelwakfinal%>% 
  bind_cols(hg38clash)

colnames(clashelwakfinal)[18] <- "HG38build_loc"

clashelwakfinal <- clashelwakfinal%>%
  dplyr::mutate(Genom_build= rep("hg19"))
str(clashelwakfinal)

# Arrangement in dataset

clashelwakfinal%>% dplyr::select(cluster=seq_ID, chromosome = Chromosome, start_position, end_position, strand, hgnc_symbol = Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, target_seq = mRNA_seq_extended, miRNA= Name, miR_seq =miRNA_seq, seed_type, seed_type2= seed_basepairs, seed_type3= folding_class, Energy = folding_energy, HG38build_loc, Genom_build, region) -> clashelwakfinal

clashelwakfinal$cluster <- as.character(clashelwakfinal$cluster)
clashelwakfinal$strand <- as.character(clashelwakfinal$strand)
clashelwakfinal$target_seq <- as.character(clashelwakfinal$target_seq)
clashelwakfinal$miR_seq <- as.character(clashelwakfinal$miR_seq)
clashelwakfinal$seed_type <- as.character(clashelwakfinal$seed_type)
clashelwakfinal$HG38build_loc <- as.character(clashelwakfinal$HG38build_loc)
clashelwakfinal$seed_type2 <- as.numeric(clashelwakfinal$seed_type2)
clashelwakfinal$seed_type3 <- as.character(clashelwakfinal$seed_type3)
```

## Interpreting the CLASH seed structures in dataset

```{r,  eval=FALSE}
clashelwakfinal%>%
  mutate(seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="I", paste0(seed_type2, "-mer"), seed_type),
         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="II", paste0(seed_type2, "-mer_noncanonical"), seed_type), 
         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="III", paste0(seed_type2, "-mer_noncanonical"), seed_type),
         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="IV", paste0(seed_type2, "-mer_noncanonical"), seed_type),
         seed_type= ifelse (startsWith(seed_type, "no") ,"none", seed_type))%>%
  dplyr::select(-seed_type2, -seed_type3)->clashelwakfinal

```

## Arrangement of CLEAR-CLiP Dataset [@moore_mirnatarget_2015]

CLASH dataset was retrieved from [Nature web page](nature.com/articles/ncomms9864)

```{r,  eval=FALSE}
clearclip <- read_xlsx("CLEAR-CLIP.xlsx")

#Clearclip hg18
```

## Query of Human Genome 18

```{r,  eval=FALSE}
#HG18
listEnsemblArchives()
listMarts(host = 'may2009.archive.ensembl.org')
ensembl54 = useMart(host='may2009.archive.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL',
                    dataset='hsapiens_gene_ensembl')

hg18 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene", "strand" ), mart = ensembl54)
```

## Adding Genome Information to dataset

```{r,  eval=FALSE}
clearclipfinal <- hg18%>%
  inner_join(clearclip, by= c("entrezgene"= "gene.id", "hgnc_symbol"= "gene.symbol"))%>%
  distinct()
```

## Converting human genome build

```{r,   eval=FALSE}
lift18 <- clearclipfinal%>%
  unite(start_end, c("start_position", "end_position"), sep = "-")%>%
  unite(location, c("chr","start_end"), sep = ":")%>%
  dplyr::select(location)

write_tsv(lift18, "lift18.txt")

deleted_lift18 <- read_tsv("deleted_lift18.txt")

colnames(deleted_lift18)[1] <- "Chromosome_loc"

deleted_lift18 <- deleted_lift18%>%
  dplyr::filter(startsWith(Chromosome_loc, "chr"))%>%
  separate(Chromosome_loc, c("Chr", "End"), "-", remove = TRUE)%>%
  separate(Chr, c("Chr", "Start"), ":", remove = TRUE)

deleted_lift18$Start <- as.numeric(deleted_lift18$Start)
deleted_lift18$End <- as.numeric(deleted_lift18$End)

clearclipfinal <- clearclipfinal%>%
  dplyr::anti_join(deleted_lift18, by = c("chr"="Chr", "start_position"="Start", "end_position"="End"))

hg38clearclip<- read.delim("hg38clearclip.txt", header = FALSE)

clearclipfinal <- clearclipfinal%>% 
  bind_cols(hg38clearclip)

colnames(clearclipfinal)[28] <- "HG38build_loc"

clearclipfinal <- clearclipfinal%>%
  dplyr::mutate(Genom_build= rep("hg18"))

clearclipfinal%>%
  dplyr::select(cluster=cluster.ID, chromosome = chr, start_position, end_position, strand = strand.y, hgnc_symbol, Ensembl_Gene_Id =ensembl_gene_id, Ensembl_Transcript_Id =ensembl_transcript_id, target_seq = target.map, miRNA, miR_seq = miR.map, seed_type = 'seed match', Energy = MFE, HG38build_loc, Genom_build, region) -> clearclipfinal
```

## Seed type manipulation

```{r,  eval=FALSE}
data_frame(seed_type= c("5mer_1","5mer_2", "5mer_3", "6mer", "6mer.indel", "6mer.mm", "6mer_off.mm", "6merA1", "6merA1.indel", "6merA1.mm", "7merA1", "7merA1.indel", "7merA1.mm", "7merm8", "7merm8.indel", "7merm8,mm", "8mer", "8mer.indel", "8mer.mm", "NA" ),
           seed_type_com= c("5-mer", "5-mer_noncanonical", "5-mer_noncanonical", "6-mer", "6-mer_noncanonical", "6-mer_noncanonical", "6-mer_noncanonical", "6-merA1", "6-merA1_noncanonical","6-merA1_noncanonical", "7-merA1", "7-merA1_noncanonical","7-merA1_noncanonical", "7-mer-8m", "7-mer-8m_noncanonical", "7-mer-8m_noncanonical", "8-mer", "8-mer_noncanonical", "8-mer_noncanonical", "none" )) ->clipdata_seed

clearclipfinal%>%
  inner_join(clipdata_seed, by = "seed_type")%>%
  dplyr::select(1:11, seed_type = seed_type_com, Energy, HG38build_loc, Genom_build, region)-> clearclipfinal

clearclipfinal$HG38build_loc <- as.character(clearclipfinal$HG38build_loc)

```

## Integration of two experimental dataset

```{r,  eval=FALSE}

bind_rows(clashelwakfinal, clearclipfinal)%>%distinct() -> experimentalmirnagene

```

## Adding Coefficients of Interaction factors

```{r,  eval=FALSE}
experimentalmirnagene <- experimentalmirnagene%>%
  mutate(region2  = str_replace_all(region, "NA", ""), region3 = str_replace_all(region2, "\\|", ""), region = str_replace_all(region3, c("3'UTR"="3UTR", "5'UTR"="5UTR")))%>%
  dplyr::select(-region2, -region3)%>%
  mutate(region_effect = as.double(ifelse(region %in% c("3UTRCDS", "CDS3UTR", "5UTR3UTR", "CDS5UTR3UTR", "CDS3UTRintron"), "0.93", ifelse(region %in% c("CDS", "CDSintron"), "0.42", ifelse(region %in% c("3UTR","3UTRintron"), "0.84", ifelse(region %in% c("5UTR", "5UTRintron"), "0.01", ifelse(region %in% c("5UTRCDS", "CDS5UTR"), "0.42", ifelse(region %in% c("intron",""), "0.01", ifelse(region %in% c("exon_unclassified", ""), "0.2", NA)))))))))

```

```{r,  eval=FALSE}

seed_type_effect <- data_frame( seed_type = c("5-mer", "5-mer_noncanonical", "6-mer", "6-mer_noncanonical", "6-merA1", "6-merA1_noncanonical", "7-mer", "7-mer_noncanonical", "7-merA1", "7-merA1_noncanonical", "7-mer-8m", "7-mer-8m_noncanonical", "8-mer", "8-mer_noncanonical", "9-mer", "9-mer_noncanonical", "none"),
                                seed_type_effect= c(0.05, 0.04, 0.07, 0.05, 0.07, 0.05, 0.23, 0.19, 0.19, 0.16, 0.25, 0.21, 0.43, 0.35, 0.43,0.35, 0.01)
                                )

experimentalmirnagene%>%
  inner_join(seed_type_effect, by= "seed_type") -> experimentalmirnagene
```

```{r, eval=FALSE}
## Saving dataset
saveRDS(experimentalmirnagene,"data/experimentalmirnagene.RDS")

```

```{r}
readRDS("data/experimentalmirnagene.RDS")->experimentalmirnagene
experimentalmirnagene
```

# REFERENCES

### 2019-09-22 Supplementary_figures.Rmd
additional files

--- 2019-09-17 Supplementary_mat_met.Rmd
@@
 ---
-title: "Supplementary Materials And Methods"
-author: "Selcen Ari"
-date: "7 Nisan 2019"
+title: "Supplementary Figures"
+author: "Selcen Arı"
+date: "17 09 2019"
 output:
@@
 
-# 1. Functions defined for ceRNA models and workflow of method
-We defined the functions that can be used with R programming.
-Briefly, these functions process a given miRNA:gene dataset and convert to graph object.
-All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
-The functions and steps of approach are explained as following (Figure S1) :
-
-**Conversion of dataset**: `priming_graph()` function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
-If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
-The formulations that are given in equations (1-4) are performed in this function.
-This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
-It is assumed that the initial target amounts in the dataset is observed after the repressive activity of miRNAs in steady-state.
-
 ![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow1.png){#figs1}
-
-**Transition of variables in graph**: In the previous step, the calculations are performed in the edge variables of the graph object.
-However, the graph object allows to use node variables, while the node features are handled to the graph.
-In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.
-
-**Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficiency coefficients are calculated according to this acceptance.
-In the network that is found in steady-state conditions, the change is applied to the graph object for distribution of steady-state.
-To provide the distribution in the network the workflow offer two methods: update_variables and update_how.
-The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
-The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
-
-**Updating current values of variables**: After variables updating in edge variables, these are carried to node variables.
-Current and previous values of variables are stored as node variables with update_variables function.
-
-**Simulation of competing behavior of targets**: After the change in the steady-state conditions, the network elements try to gain steady-state again.
-This process progresses as repeating of regulations after the spreading the changes in the network.
-In this step, simulation of regulations according to given cycle count in simulate function is applied.
-After each simulation cycle, the miRNA repression values are re-calculated and the current values of competing elements are found and saved.
-The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node data.
-
-The node elements in the dataset are handled as two type; repressive (miRNAs) and competing (targets).
-It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used.
-If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user.
-On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger.
-The functions that are used in the approach are developed with R programming so as can be used with other packages.
-These are can be found in the Github repository [ceRNAnetsim Github page](https://github.com/selcenari/ceRNAnetsim) and improved with contributions from others.
 
@@
 
-+ load *minsamp* data
-
-```{r}
-data("minsamp")
-
-minsamp
-```
-
-See Table S1 in Supplementary Tables file.
-
-## *minsamp* dataset analysis in lack of interaction factors.
-
-Firstly, we have analysed minimal data without interaction factors between miRNA:target.
-
-- 1. We have evaluated graph in the steady state conditions as followings:
+## *minsamp* dataset analysis at lack of interaction factors.
 
@@
 ```
-
-- 2. We have obtained graph after change on Gene2 expression as following:
 
@@
 
-- 3. We have determined regulations after Gene2 upregulation:
-
 ```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of minsamp data.
@@
   update_how("Gene2", 2)%>%
-  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "Regulations after Gene2 Upregulation", 2)
+  simulate(cycle=1)%>%
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "a)")
+
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
+  update_how("Gene2", 2)%>%
+  simulate(2)%>%
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "b)")
 
@@
 
-Note that the regulations are colored according to expression changes of present and a previous value.
-So, it can be observed that whole gene expressions increase in comparison of initial steady-state.
-The overall regulations of gene expressions are as followings:
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)%>%
-  simulate(2)%>%
-  activate(edges)%>%
-  as_tibble()%>%
-  select(Competing_name,comp_count_list,effect_list)%>%
-  unnest()
-```
-
 ## *minsamp* dataset analysis with interaction factors.
-
-We have made the same analysis in presence of interaction factors.
 
@@
 
-```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.cap = "Sequential iteration of minsamp data with interaction factors", fig.subcap= c("First response of system to Gene2 upregulation (2nd iteration)", "Spreading of perturbation on system (3th iteration)")}
+```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of minsamp data with interaction factors a)First response of system to Gene2 upregulation (2nd iteration).
+b)Spreading of perturbation on system (3th iteration)"}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
   update_how("Gene2", 2)%>%
-  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation with interaction factors",2)
-```
+  simulate(cycle = 1)%>%
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "a)")
 
-When the graphs which were resulted from analyses were examined, it was observed that behaviors were same.
-But, when the results were analysed in terms of expression values, the regulation differences can be observed.
-
-```{r, message=FALSE, warning=FALSE}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
   update_how("Gene2", 2)%>%
-  simulate(3)%>%
-  activate(edges)%>%
-  as_tibble()%>%
-  select(Competing_name,comp_count_list,effect_list)%>%
-  unnest()
+  simulate(cycle = 2)%>%
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "b)")
+
 ```
@@
 
-Genes targeted by multiple miRNAs (referred to as "common target") are of special interest since they are subject to cooperative effect.
-Also, they perturb more than one neighborhood.
-In our small dataset, minsamp, Gene4 is regulated by two miRNAs.
-Let's simulate perturbation effects triggered by Gene4.
+```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Perturbation of Gene4 on minsamp data with interaction factors.
+a) Netwrk at steady-state.
+b) Upregulation of Gene4.
+c) Primary response of network to upregulation of Gene4.
+d) Re-regulation of whole nodes on system (3th iteration)"}
 
-```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.cap = "Perturbation of Gene4 on minsamp data with interaction factors", fig.subcap= c("1st iteration after Gene4 upregulation", "2nd iteration after Gene4 perturbation")}
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "a)")
+
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
   update_how("Gene4", 2)%>%
-  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Common Gene-Gene4 Upregulation without interaction factors",2)
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "b)")
+
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+  update_how("Gene4", 2)%>%
+  simulate(cycle=1)%>%
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "c)")
+
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+  update_how("Gene4", 2)%>%
+  simulate(cycle=2)%>%
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "d)")
 ```
 
-The common target perturbation  (increasing to two fold at Gene4 expression in presence of interaction factors) resulted in more prominent efficiency at the same conditions (shown in following).
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene4", 2)%>%
-  simulate(3)%>%
-  activate(edges)%>%
-  as_tibble()%>%
-  select(Competing_name,comp_count_list,effect_list)%>%
-  unnest()
-```
-
-## Determination of perturbation efficiencies of elements in system.
-
-```{r, message=FALSE, warning=FALSE}
-
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)-> sample_graph
-
-find_node_perturbation(sample_graph, how = 2, cycle = 3, limit = 0.1)
-
-```
-
-# 2. Obtaining breast cancer dataset and integration
-
-This section describes how to apply `ceRNAnetsim` package on a breast cancer patient miRNA:target interaction dataset. Before the approach, we obtained three datasets and combined them.
-
-```{r, echo=FALSE, message= FALSE, warning=FALSE}
-# Install the packages which is necessary to obtain and process datasets.
-#install.packages("BiocManager")
-#BiocManager::install("SummarizedExperiment")
-#BiocManager::install("TCGAbiolinks")
-library(TCGAbiolinks)
-library(SummarizedExperiment)
-```
-
-## 2.1 How to get gene expression counts of TCGA-E9-A1N5 patient.
-
-We have obtained the gene expression values of patient using [TCGAbiolinks](https://www.bioconductor.org/packages/release/bioc/html/TCGAbiolinks.html) package from [Bioconductor](https://www.bioconductor.org/).
-For this process, we have followed the instructions of the package.
-`TCGAbiolinks` package provides to obtain data for whole number of given barcode(s) at once.
-But, we preferred to download them separately to show datasets.
-
-+ Obtain to gene expression counts of tumor tissue.
-
-```{r,message= FALSE, warning=FALSE, eval=FALSE}
-
-BCP_tumor <- GDCquery(project = "TCGA-BRCA",
-                  data.category = "Transcriptome Profiling",
-                  data.type = "Gene Expression Quantification", 
-                  workflow.type = "HTSeq - Counts",
-                  barcode = "TCGA-E9-A1N5-01A-11R-A14D-07")
-
-GDCdownload(BCP_tumor)
-BCPGE_tumor <- GDCprepare(BCP_tumor)
-
-as.data.frame(assay(BCPGE_tumor))%>%
-  mutate(ensembl_gene_id = rownames(.))%>%
-  dplyr::inner_join(as.data.frame(rowData(BCPGE_tumor)), by = "ensembl_gene_id")%>%
-  dplyr::select(ensembl_gene_id, external_gene_name, 1)-> TCGA_E9_A1N5_tumor
-
-colnames(TCGA_E9_A1N5_tumor)[3] <- "GE_tumor"
-
-head(TCGA_E9_A1N5_tumor)
-```
-
-+ Obtain to gene expression counts of normal tissue.
-
-```{r, message= FALSE, warning=FALSE, eval=FALSE}
-BCP_normal <- GDCquery(project = "TCGA-BRCA",
-                  data.category = "Transcriptome Profiling",
-                  data.type = "Gene Expression Quantification", 
-                  workflow.type = "HTSeq - Counts",
-                  barcode = "TCGA-E9-A1N5-11A-41R-A14D-07")
-GDCdownload(BCP_normal)
-
-BCPGE_normal <- GDCprepare(BCP_normal)
-
-as.data.frame(assay(BCPGE_normal))%>%
-  mutate(ensembl_gene_id = rownames(.))%>%
-  dplyr::inner_join(as.data.frame(rowData(BCPGE_normal)), by = "ensembl_gene_id")%>%
-  dplyr::select(ensembl_gene_id, external_gene_name, 1)-> TCGA_E9_A1N5_normal
-
-colnames(TCGA_E9_A1N5_normal)[3] <- "GE_normal"
-
-head(TCGA_E9_A1N5_normal)
-
-```
-
-## 2.2 How to get miRNA expression counts of TCGA-E9-A1N5 patient.
-
-We have used `TCGAbiolinks` package to obtain miRNA expression quantification.
-The query gives read count of miRNA as isoform chromosome coordination.
-The data also contains mature miRNA information.
-So, we processed data to attain -5p -3p isoform information using mirbase release21 dataset.
-
-+ Get the mirbase id of mature miRNA:
-
-We downloaded the mirbase release 21 dataset from [mirbase](http://www.mirbase.org/ftp.shtml) and processed the patient mirna expression datasets as following:
-
-```{r, message= FALSE, warning=FALSE}
-library(readr)
-read_tsv("hsa_mirna.txt", comment = "#", col_names = FALSE)%>%
-  dplyr::select(mirna_type= X3, definition = X9)%>%
-  filter(!endsWith(mirna_type, "primary_transcript"))%>%
-  tidyr::separate(definition, c("ID", "Alias", "Name", "Derivated"), sep = ";")%>%
-  dplyr::select(Alias, Name)%>%
-  tidyr::separate(Alias, c("trash1", "ID"), sep = "=")%>%
-  tidyr::separate(Name, c("trash2", "Name"), sep = "=")%>%
-  dplyr::select(-trash1, -trash2)-> mirbase_id_conv
-
-head(mirbase_id_conv)
-```
-
-+ Obtain the miRNA expression of tumor tissue of patient:
-
-```{r, message= FALSE, warning=FALSE, eval=FALSE}
-BCP_mirnatumor <- GDCquery(project = "TCGA-BRCA",
-                  data.category = "Transcriptome Profiling",
-                  data.type = "Isoform Expression Quantification",
-                  workflow.type = "BCGSC miRNA Profiling",
-                  barcode = "TCGA-E9-A1N5-01A-11R-A14C-13")
-
-GDCdownload(BCP_mirnatumor)
-
-GDCprepare(BCP_mirnatumor)%>%
-  as.data.frame()%>%
-  dplyr::select(miRNA_ID, read_count, reads_per_million_miRNA_mapped, miRNA_region)%>%
-  dplyr::filter(startsWith(miRNA_region, "mature"))%>%
-  dplyr::mutate(mirbase_id =str_remove(miRNA_region, "mature,"))%>%
-  dplyr::select(-miRNA_region)%>%
-  dplyr::inner_join(mirbase_id_conv, by = c("mirbase_id"="ID"))%>%
-  dplyr::select(miRNA_name = Name, read_count, reads_per_million_miRNA_mapped)%>%
-  dplyr::group_by(miRNA_name)%>%
-  mutate(read_count= sum(read_count), reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
-  dplyr::ungroup()%>%
-  distinct() -> BCPME_mirnatumor
-
-head(BCPME_mirnatumor)
-```
-
-+ Obtain the miRNA expression of normal tissue of patient:
-
-```{r, message= FALSE, warning=FALSE, eval=FALSE}
-BCP_mirnanormal <- GDCquery(project = "TCGA-BRCA",
-                  data.category = "Transcriptome Profiling",
-                  data.type = "Isoform Expression Quantification",
-                  workflow.type = "BCGSC miRNA Profiling",
-                  barcode = "TCGA-E9-A1N5-11A-41R-A14C-13")
-
-GDCdownload(BCP_mirnanormal)
-# a616435d-0b69-48ac-813d-5d75ad9b85eb.mirbase21.isoforms.quantification.txt
-
-GDCprepare(BCP_mirnanormal)%>%
-  as.data.frame()%>%
-  dplyr::select(miRNA_ID, read_count, reads_per_million_miRNA_mapped, miRNA_region)%>%
-  dplyr::filter(startsWith(miRNA_region, "mature"))%>%
-  dplyr::mutate(mirbase_id =str_remove(miRNA_region, "mature,"))%>%
-  dplyr::select(-miRNA_region)%>%
-  dplyr::inner_join(mirbase_id_conv, by = c("mirbase_id"="ID"))%>%
-  dplyr::select(miRNA_name = Name, read_count, reads_per_million_miRNA_mapped)%>%
-  dplyr::group_by(miRNA_name)%>%
-  mutate(read_count= sum(read_count), reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
-  dplyr::ungroup()%>%
-  distinct() -> BCPME_mirnanormal
-
-head(BCPME_mirnanormal)
-```
-
-## 2.3 Get the high-throughput experimental miRNA:target dataset.
-
-There are various datasets about miRNA:target pairs such as miRTarBase, DianaTools, miRecords, miRWalk etc.
-Some of these present the experimentally supported miRNA target pairs or only predicted ones.
-The experimentally supported datasets generally provides weak evidence for interactions.
-For these reasons, we obtained the high-throughput experimental miRNA:target dataset from two studies performed by Helwak et al.
-and Moore et al.
-These steps were not handle in this file because they contain many processes.
-
-Briefly these datasets contain various common information about miRNA:target interactions such as the miRNA name, miRNAsequence, target name, target sequence, their chromosomal locations, binding location on the target sequence, binding free energy, seed structure.
-But these datasets provides the informations with different data structures.So we followed the steps:
-
-+ The datasets were directly downloaded from supplementary data files of the studies.
-+ It was provided that the datasets are converted to same human genome build.
-+ The seed type information was organized as the same style.
-+ The datasets were combined.
-+ We committed the interaction factors as numeric values according to previous studies. (We added the interaction factors and their numeric values at Supplementary tables S2-3)
-
-[#TODO çok karmaşık komutlar için ayrı bir klasör hazırlanabilir. burayada link verilebilir: "experimentalmirnagene" tablosu repodaki X.R script ile elde edilmiştir" gibi]
-
-Finally, we have obtained the experimentally supported miRNA:target dataset.
-
-```{r, message= FALSE, warning=FALSE}
-data("experimentalmirnagene")
-
-head(experimentalmirnagene)
-
-```
-
-The methods about miRNA:target interactions are based a basic principle that is reading after isolation of miRNA:target chimerics.
-The datasets contain all the chimeric miRNA:target structures found in the medium during the experiment.
-On the other hand, it could be said that the reading is performed as snapshot.Because of that, the methods can provide different chimeric interactions the same miRNA:target pair.
-We have preferred to select most effective interaction parameters for the same miRNA:target pairs that can exhibit various interactions.
-The step is performed as:
-
-```{r, message= FALSE, warning=FALSE, eval=FALSE}
-experimentalmirnagene%>%
-  dplyr::select(miRNA, Ensembl_Gene_Id, hgnc_symbol, Energy, seed_type_effect, region_effect)%>%
-  distinct()%>%
-  group_by(Ensembl_Gene_Id, miRNA)%>%
-  mutate(seed_type_effect= ifelse(seed_type_effect==max(seed_type_effect), seed_type_effect, max(seed_type_effect)), Energy = ifelse(Energy==min(Energy), Energy, min(Energy)), region_effect= ifelse(region_effect==max(region_effect), region_effect, max(region_effect)))%>%
-  distinct()-> tocombine_mirnagene
-
-head(tocombine_mirnagene)
-```
-
-## 2.4 Combine the dataset
-
-```{r, message= FALSE, warning=FALSE, eval=FALSE}
-BCPME_mirnanormal%>%
-  dplyr::inner_join(tocombine_mirnagene, by = c("miRNA_name"="miRNA"))%>%
-  dplyr::inner_join(TCGA_E9_A1N5_normal, by = c("Ensembl_Gene_Id"="ensembl_gene_id", "hgnc_symbol"="external_gene_name"))%>%
-  distinct()%>%
-  dplyr::select(hgnc_symbol, miRNA_name, mirna_RPM= reads_per_million_miRNA_mapped, GE_normal, Energy, seed_type_effect, region_effect)-> E9GE_mirnagenenormal
-
-```
+## Determination of iteration for Upregulation of SERPINE2 gene from Breast cancer patient dataset
 
 ```{r}
-data("E9GE_mirnagenenormal")
+E9GE_mirnagenenormal<-readRDS("data/E9GE_mirnagenenormal.rda")
 head(E9GE_mirnagenenormal)
 ```
-
-```{r, message= FALSE, warning=FALSE, eval=FALSE}
-BCPME_mirnatumor%>%
-  dplyr::inner_join(tocombine_mirnagene, by = c("miRNA_name"="miRNA"))%>%
-  dplyr::inner_join(TCGA_E9_A1N5_tumor, by = c("Ensembl_Gene_Id"="ensembl_gene_id", "hgnc_symbol"="external_gene_name"))%>%
-  distinct()%>%
-  dplyr::select(hgnc_symbol, miRNA_name, mirna_RPM= reads_per_million_miRNA_mapped, GE_tumor, Energy, seed_type_effect, region_effect)-> E9GE_mirnagenetumor
-
-```
-
-```{r}
-data("E9GE_mirnagenetumor")
-head(E9GE_mirnagenetumor)
-```
-
-## 2.5 Selection of trigger node
-
-We have compared two datasets that are obtained for the tumor and normal tissue samples of same patient. We tried to change expression of a gene in normal tissue as the same level in the tumor tissue.
-
-For this step, we have determined the changes of the gene expression in terms of fold change:
-
-```{r}
-E9GE_mirnagenetumor%>%
-  dplyr::select(hgnc_symbol, GE_tumor)%>%
-  dplyr::inner_join((E9GE_mirnagenenormal%>%dplyr::select(hgnc_symbol, GE_normal)), by ="hgnc_symbol")%>%
-  dplyr::mutate(FC= GE_tumor/GE_normal)%>%
-  distinct()%>%
-  filter(FC>2.5,  FC<3.5)-> three_fold_change
-
-# ABCC1 gene has 5827 read count in tumor tissue although 1420 in normal tissue (FC=4.10)
-```
-
-[#TODO SERPINE2]
-
-Secondly, we have determined the most important nodes of network.
-We applied `find_node_perturbation` function.
-We only defined nodes that affect the other nodes more than 1.05 fold change with 10 iteration when they increase 3 fold.
-
-```{r, include=FALSE, eval=FALSE}
-as.data.frame(E9GE_mirnagenenormal)%>%
-  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
-  find_node_perturbation(how = 3, cycle = 10, fast = 5)%>%
-  filter(!is.na(perturbation_efficiency))->perturbationofnetwork
-
-perturbationofnetwork
-```
-This dataset, perturbationofnetwork, includes 382 genes of totally 423 nodes.
-
-Result of this, we obtained common nodes of these two datasets (i.e perturbationofnetwork and three_fold_change) and selected a gene, SERPINE2.
-
-```{r, eval=FALSE}
-three_fold_change%>%
-  inner_join(perturbationofnetwork, by = c("hgnc_symbol"="name"))
-
-#Selected node is SERPINE2
-```
-## 2.5 Approach of Method into Combined Datasets
-
-We selected SERPINE2 gene for simulation of regulation on network.
-
-### 2.5.1 Find iteration of simulation
 
@@
  simulate(150) %>%
- find_iteration(limit=1, plot= TRUE) #limit=1 describes the change that is not taken into account.
+ find_iteration(limit=0, plot= TRUE) #limit=0 describes the change that is not taken into account.
 ```
-
-The node amount of changed gene on the system in terms of percentage were shown in above.
-As seen, firstly, the changed gene count increase.
-The system which contains the hundreds of miRNAs and thousands of genes can slowly gain the steady-state again.
-At first glance, it can be assumed that when all nodes in the system are reached, stable state will be provided.
-However, although all nodes are reached, the nodes competing with each other cause the edits to continue for a while.
-
-The dynamics of the approach are shown in package  [vignettes](https://selcenari.github.io/ceRNAnetsim/articles/convenient_iteration.html).
-
-So, we offered an approach about to find iteration.
-`find_iteration` function does not give the iteration to gain steady-state, but it gives the iteration which has maximum affected node counts.
-The function is applied as following:
-
-```{r}
-as.data.frame(E9GE_mirnagenenormal)%>%
- priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
- update_how("SERPINE2",2.75) %>%
- simulate(100) %>%
- find_iteration(limit=1, plot= FALSE)
-
-#31
-```
-
-### 2.5.2 Simulation of dataset
-
-We tried to apply two fold of the point that SERPINE2 has maximum affected genes on network.
-
-```{r}
-
-as.data.frame(E9GE_mirnagenenormal)%>%
-  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
-  update_how("SERPINE2",2.75)%>%
-  simulate(62)
-
-```
-
-```{r}
-as.data.frame(E9GE_mirnagenenormal)%>%
-  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
-  update_how("SERPINE2",2.75)%>%
-  simulate(62)%>%
-  as_tibble()%>%
-  select(name, initial_count, count_current)->simulation_results
-```
-
-### 2.5.3 Comparison of simulation results and tumor tissue expression values
-```{r}
-E9GE_mirnagenetumor%>%
-  dplyr::select(hgnc_symbol, GE_tumor)%>%
-  dplyr::inner_join((E9GE_mirnagenenormal%>%dplyr::select(hgnc_symbol, GE_normal)), by ="hgnc_symbol")%>%
-  inner_join(simulation_results, by= c("hgnc_symbol"="name"))%>%
-  distinct()
-```
-
-Actually, we have developed to provide a new approach mirna mediated regulation networks.
-This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and coherent miRNA:target regulation approach.
-
-# 3. REFERENCES

### 2019-09-22 Supplementary_mat_met.Rmd
additional files

--- 2019-09-17 Supplementary_mat_met.Rmd
@@
 All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
-The functions and steps of approach are explained as following (Figure S1) :
+The functions and steps of approach are explained as following (Figure S1 in Supplementary Figures) :
 
@@
 It is assumed that the initial target amounts in the dataset is observed after the repressive activity of miRNAs in steady-state.
-
-![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow1.png){#figs1}
 
@@
 
-- 1. We have evaluated graph in the steady state conditions as followings:
+- 1. We have evaluated graph in the steady state conditions as followings (Figure S2 in Supplementary Figures):
 
-```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset in Steady-state", fig.width=5, fig.height=4}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset in steady-state conditions")
+```{r, message=FALSE, warning=FALSE}
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)
 ```
 
-- 2. We have obtained graph after change on Gene2 expression as following:
+- 2. We have obtained graph after change on Gene2 expression as following (Figure S3 in Supplementary Figures):
 
-```{r, message=FALSE, warning=FALSE, fig.keep='all', fig.cap= "Gene2 Upregulation on Minimal Dataset", fig.width=5, fig.height=4}
+```{r, message=FALSE, warning=FALSE}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors")
+  update_how("Gene2", 2)
 ```
 
-- 3. We have determined regulations after Gene2 upregulation:
+- 3. We have determined regulations after Gene2 upregulation (Figure S4 in Supplementary Figures):
 
-```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of minsamp data.
-a) First response of system to Gene2 upregulation (2nd iteration).
-b) Spreading of perturbation on system (3th iteration)", fig.subcap= NULL}
+```{r, message=FALSE, warning=FALSE}
 
@@
   update_how("Gene2", 2)%>%
-  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "Regulations after Gene2 Upregulation", 2)
+  simulate(cycle= 2)
 
@@
 
-We have made the same analysis in presence of interaction factors.
+We have made the same analysis in presence of interaction factors (Sequentially shown at Figure S5-7 in Supplementary Figures).
 
-```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset with interaction factors in Steady-state", fig.width=5, fig.height=4}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset with interaction factors in steady-state conditions")
+```{r, message=FALSE, warning=FALSE}
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)
 
@@
 
-```{r, message=FALSE, warning=FALSE, fig.cap= " When Gene2 is upregulated on Minimal Dataset with interaction factors", fig.width=5, fig.height=4}
+```{r, message=FALSE, warning=FALSE}
+
+priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+  update_how("Gene2", 2)
+```
+
+```{r, message=FALSE, warning=FALSE}
+
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
   update_how("Gene2", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation with interaction factors")
-```
-
-```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.cap = "Sequential iteration of minsamp data with interaction factors", fig.subcap= c("First response of system to Gene2 upregulation (2nd iteration)", "Spreading of perturbation on system (3th iteration)")}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene2", 2)%>%
-  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation with interaction factors",2)
+  simulate(cycle=2)
 ```
@@
 In our small dataset, minsamp, Gene4 is regulated by two miRNAs.
-Let's simulate perturbation effects triggered by Gene4.
+Let's simulate perturbation effects triggered by Gene4 (Shown at Figure S8 in Supplementary Figures .
 
-```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.cap = "Perturbation of Gene4 on minsamp data with interaction factors", fig.subcap= c("1st iteration after Gene4 upregulation", "2nd iteration after Gene4 perturbation")}
+```{r, message=FALSE, warning=FALSE}
+
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
   update_how("Gene4", 2)%>%
-  simulate_vis(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Common Gene-Gene4 Upregulation without interaction factors",2)
+  simulate(cycle = 2)
 ```
@@
 + The datasets were combined.
-+ We committed the interaction factors as numeric values according to previous studies. (We added the interaction factors and their numeric values at Supplementary tables S2-3)
-
-[#TODO çok karmaşık komutlar için ayrı bir klasör hazırlanabilir. burayada link verilebilir: "experimentalmirnagene" tablosu repodaki X.R script ile elde edilmiştir" gibi]
++ We committed the interaction factors as numeric values according to previous studies.
+(We added the interaction factors and their numeric values at Supplementary files, Process on miRNA:target pairs dataset for process codes and Supplementary Tables for used numeric values)
 
@@
 ```{r, message= FALSE, warning=FALSE}
-data("experimentalmirnagene")
+
+experimentalmirnagene<-readRDS("data/experimentalmirnagene.RDS")
 
@@
 ```{r}
-data("E9GE_mirnagenenormal")
+#saveRDS(E9GE_mirnagenenormal, "data/E9GE_mirnagenenormal.rda")
+
+E9GE_mirnagenenormal<-readRDS("data/E9GE_mirnagenenormal.rda")
 head(E9GE_mirnagenenormal)
@@
 ```{r}
-data("E9GE_mirnagenetumor")
+#saveRDS(E9GE_mirnagenetumor, "data/E9GE_mirnagenetumor.rda")
+
+E9GE_mirnagenetumor<-readRDS("data/E9GE_mirnagenetumor.rda")
 head(E9GE_mirnagenetumor)
@@
 
-We have compared two datasets that are obtained for the tumor and normal tissue samples of same patient. We tried to change expression of a gene in normal tissue as the same level in the tumor tissue.
-
-For this step, we have determined the changes of the gene expression in terms of fold change:
-
-```{r}
-E9GE_mirnagenetumor%>%
-  dplyr::select(hgnc_symbol, GE_tumor)%>%
-  dplyr::inner_join((E9GE_mirnagenenormal%>%dplyr::select(hgnc_symbol, GE_normal)), by ="hgnc_symbol")%>%
-  dplyr::mutate(FC= GE_tumor/GE_normal)%>%
-  distinct()%>%
-  filter(FC>2.5,  FC<3.5)-> three_fold_change
-
-# ABCC1 gene has 5827 read count in tumor tissue although 1420 in normal tissue (FC=4.10)
-```
-
-[#TODO SERPINE2]
-
-Secondly, we have determined the most important nodes of network.
+We have determined the most important nodes of network.
 We applied `find_node_perturbation` function.
@@
 ```{r, include=FALSE, eval=FALSE}
-as.data.frame(E9GE_mirnagenenormal)%>%
+
+as_data_frame(E9GE_mirnagenenormal)%>%
   priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
@@
 
-perturbationofnetwork
+perturbationofnetwork%>%
+  filter(type== "Competing")%>%
+  arrange(desc(perturbation_efficiency))
 ```
-This dataset, perturbationofnetwork, includes 382 genes of totally 423 nodes.
+This dataset, perturbationofnetwork, includes totally 423 effective nodes. We selected SERPINE2 gene to perturb the network because it is the most efficient node gene in network.
 
-Result of this, we obtained common nodes of these two datasets (i.e perturbationofnetwork and three_fold_change) and selected a gene, SERPINE2.
-
-```{r, eval=FALSE}
-three_fold_change%>%
-  inner_join(perturbationofnetwork, by = c("hgnc_symbol"="name"))
-
-#Selected node is SERPINE2
-```
 ## 2.5 Approach of Method into Combined Datasets
@@
 
-```{r, fig.cap= "Percentage of affected nodes of each iteration for SERPINE2 Gene"}
-as.data.frame(E9GE_mirnagenenormal)%>%
- priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
- update_how("SERPINE2",2.75) %>%
- simulate(150) %>%
- find_iteration(limit=1, plot= TRUE) #limit=1 describes the change that is not taken into account.
-```
-
-The node amount of changed gene on the system in terms of percentage were shown in above.
+The node amount of changed gene on the system in terms of percentage were shown at Figure S9 in Supplementary Figures.
 As seen, firstly, the changed gene count increase.
@@
 ```{r}
+library(ceRNAnetsim)
 as.data.frame(E9GE_mirnagenenormal)%>%
  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
- update_how("SERPINE2",2.75) %>%
- simulate(100) %>%
- find_iteration(limit=1, plot= FALSE)
+ update_how("SERPINE2",2.75)%>%
+ simulate(50) %>%
+ find_iteration(limit=0, plot= FALSE)
 
-#31
 ```
@@
 ```{r}
-
 as.data.frame(E9GE_mirnagenenormal)%>%
@@
   update_how("SERPINE2",2.75)%>%
-  simulate(62)
+  simulate(32)
 
@@
   update_how("SERPINE2",2.75)%>%
-  simulate(62)%>%
+  simulate(32)%>%
   as_tibble()%>%
-  select(name, initial_count, count_current)->simulation_results
+  dplyr::select(name, initial_count, count_current)->simulation_results
 ```

### 2019-09-22 Supplementary_tables.Rmd
additional files

@@
 
-|seed_type| Coefficient |
-|-----------|------------|
-| `5-mer`   | `0.05` |
-| `5-mer_noncanonical`   | `0.04` |
-| `6-mer`   | `0.07` |
-| `6-mer_noncanonical`   | `0.05` |
-| `6-merA1`   | `0.07` |
-| `6-merA1_noncanonical`   | `0.05` |
-| `7-mer`   | `0.23` |
-| `7-mer_noncanonical`   | `0.19` |
-| `7-merA1`   | `0.19` |
-| `7-merA1_noncanonical`   | `0.16` |
-| `7-mer-8m`   | `0.25` |
-| `7-mer-8m_nonacnonical`   | `0.21` |
-| `8-mer`   | `0.43` |
-| `8-mer_noncanonical`   | `0.35` |
-| `9-mer`   | `0.43` |
-| `9-mer_noncanonical`   | `0.35` |
-| `none`   | `0.01` |
+\begin{table}[ht]
+\centering
+\caption{Efficiency factors for seed types.}
+\begin{tabular}{rlr}
+  \hline
+ & seed\_type & seed\_type\_effect \\ 
+  \hline
+1 & 6-mer\_noncanonical & 0.05 \\ 
+  2 & 9-mer & 0.43 \\ 
+  3 & 6-mer & 0.07 \\ 
+  4 & 8-mer & 0.43 \\ 
+  5 & 7-mer & 0.23 \\ 
+  6 & none & 0.01 \\ 
+  7 & 5-mer\_noncanonical & 0.04 \\ 
+  8 & 5-mer & 0.05 \\ 
+  9 & 6-merA1\_noncanonical & 0.05 \\ 
+  10 & 7-mer-8m\_noncanonical & 0.21 \\ 
+  11 & 7-mer-8m & 0.25 \\ 
+  12 & 8-mer\_noncanonical & 0.35 \\ 
+  13 & 7-merA1\_noncanonical & 0.16 \\ 
+  14 & 7-merA1 & 0.19 \\ 
+  15 & 6-merA1 & 0.07 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
-Table: Efficiency factors for seed types.
+\begin{table}[ht]
+\centering
+\caption{ Efficiency factors for binding regions on targets}
+\begin{tabular}{rlr}
+  \hline
+ & region & region\_effect \\ 
+  \hline
+1 & 3UTR & 0.84 \\ 
+  2 & CDS & 0.42 \\ 
+  3 & 3UTRCDS & 0.93 \\ 
+  4 & 5UTR & 0.01 \\ 
+  5 & 5UTRCDS & 0.42 \\ 
+  6 &  & 0.01 \\ 
+  7 & intron & 0.01 \\ 
+  8 & CDS3UTR & 0.93 \\ 
+  9 & CDS5UTR & 0.42 \\ 
+  10 & exon\_unclassified & 0.20 \\ 
+  11 & CDS3UTRintron & 0.93 \\ 
+  12 & 3UTRintron & 0.84 \\ 
+  13 & CDSintron & 0.42 \\ 
+  14 & 5UTRintron & 0.01 \\ 
+  15 & 5UTR3UTR & 0.93 \\ 
+  16 & CDS5UTR3UTR & 0.93 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
-|location of binding region on target| Coefficient |
-|-----------|------------|
-| `3'UTR`   | `0.84` |
-| `CDS`   | `0.42` |
-| `5'UTR`   | `0.01` |
-
-Table: Efficiency factors for binding regions on targets
-
-```{r, eval=FALSE, include=FALSE}
-data("E9GE_mirnagenenormal")
-
-xtable(head(E9GE_mirnagenenormal),caption = "Example of E9GE_mirnagenenormal dataset" )
-```
-
-|hgnc_symbol| miRNA_name | mirna_RPM | GE_normal | Energy | seed_type_effect | region_effect |
+|Hugo_Symbol| miRNA_name | mirna_RPM | GE_normal | energy | seed_type_effect | region_effect |
 |-----------|------------|----------|----------------|---------|-----------|----------|
-| CCNG1 | hsa-let-7a-5p | 111204.15 | 5245.00 | -25.10 | 0.05 | 0.84 |
-| DICER1 | hsa-let-7a-5p | 111204.15 | 3285.00 | -24.40 | 0.43 | 0.42 | 
-| SESN1 | hsa-let-7a-5p | 111204.15 | 1179.00 | -22.20 | 0.05 | 0.93 |
-| NIPBL | hsa-let-7a-5p | 111204.15 | 4503.00 | -22.10 | 0.05 | 0.42 | 
-| INTS12 | hsa-let-7a-5p | 111204.15 | 600.00 | -21.90 | 0.05 | 0.42 | 
-| FNIP1 | hsa-let-7a-5p | 111204.15 | 1248.00 | -21.80 | 0.43 | 0.84 |
+| ENAH | hsa-let-7a-5p | 111204.15 | 7540 | -22.70 | 0.01 | 0.00 |
+| GALNT2 | hsa-let-7a-5p | 111204.15 | 2824 | -20.50 | 0.14 | 0.00 | 
+| RLF | hsa-let-7a-5p | 111204.15 | 1144 | -18.00 | 0.03 | 0.20 |
+| MAST2 | hsa-let-7a-5p | 111204.15 | 2640 | -22.60 | 0.24 | 0.00 | 
+| DOCK1 | hsa-let-7a-5p | 111204.15 | 4826 | -20.50 | 0.01 | 0.00 | 
+| ZBTB16 | hsa-let-7a-5p | 111204.15 | 315 | -24.30 | 0.14 | 0.00 |
 
@@
 
+|Variebles| Structure | Means|
+|-----------|------------|------------|
+| **Node Variables**
+| `name`  | character | node name |
+| `type`  | character |Competing or miRNA |
+| `node_id`  | numeric | in on graph object |
+| `initial_count`   | numeric | Initial Expression value of node |
+| `count_pre`   | numeric | Expression value of node at previous regulation |
+| `count_current`   | numeric | Existing expression value of node |
+| `changes_variable`   | character | Regulation of node (Up, down or steady) |
+|-----------|------------|------------|
+| **Edge Variables**
+| `Competing name` | character | name of genes|
+| `miRNA name` | character | name of miRNAs|
+| `Competing expression` | numeric | Expression values of competing elements at steady-state |
+| `miRNA expression` | numeric | Expression values of miRNA elements at steady-state |
+| `energy` | numeric | coefficient of miRNA:target interactions (binding affinity)|
+| `seed type` | numeric | coefficient of miRNA:target interactions (binding affinity)|
+| `region` | numeric | coefficient of miRNA:target interactions (degradation efficiency)|
+| `afff factor` | numeric| coefficient scaled and combined affinity factor |
+| `degg factor` | numeric |coefficient scaled and combined degradation factor| 
+| `comp_count_list` | list | list of competing expression for each iteration | 
+| `comp_count` | numeric | _pre: competing expression at previous iteration; _current: competing expression at present iteration |
+| `mirna_count_list` | list | list of miRNA expression for each iteration | 
+| `mirna_count` | numeric | _pre: miRNA expression at previous iteration; _current: miRNA expression at present iteration |
+| `effect` | numeric | _pre: total miRNA reppressive effect on individual target at previous iteration ; _current: miRNA reppressive effect on individual target at present iteration |
+| `effect_list` | list | list of miRNA reppressive effect on individual target for each iteration |
+
+Table: The context graph object during the process.
+
 ## REFERENCES

### 2019-09-23 Supplementary_additional_file.Rmd
article edit

@@
 title: "Supplementary Experimental Data File"
-author: "Selcen Arı"
+author: 
+- Selcen Ari
+- Alper Yilmaz 
 date: "22 09 2019"
@@
 library(readxl)
+knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
 

### 2019-09-23 Supplementary_figures.Rmd
article edit

@@
 title: "Supplementary Figures"
-author: "Selcen Arı"
+author: 
+- Selcen Ari
+- Alper Yilmaz 
 date: "17 09 2019"
@@
 library(stringr)
+knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
 
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset in steady-state conditions")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4",
+            title = "Minimal dataset in steady-state conditions")
 ```
@@
   update_how("Gene2", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation without interaction factors")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", 
+            title = "Gene2 Upregulation without interaction factors")
 ```
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", title = "Minimal dataset with interaction factors in steady-state conditions")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", 
+            title = "Minimal dataset with interaction factors in steady-state conditions")
 
@@
   update_how("Gene2", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "Gene2 Upregulation with interaction factors")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", 
+            title = "Gene2 Upregulation with interaction factors")
 ```
@@
  simulate(150) %>%
- find_iteration(limit=0, plot= TRUE) #limit=0 describes the change that is not taken into account.
+ find_iteration(limit=1, plot= TRUE) 
+#limit=0 describes the change that is not taken into account.
 ```

### 2019-09-23 Supplementary_mat_met.Rmd
article edit

@@
 title: "Supplementary Materials And Methods"
-author: "Selcen Ari"
+author: 
+- Selcen Ari
+- Alper Yilmaz 
 date: "7 Nisan 2019"
@@
 library(stringr)
+library(readr)
+knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
 
@@
 
-```{r, echo=FALSE, message= FALSE, warning=FALSE}
+```{r, echo=FALSE, message= FALSE, warning=FALSE, eval=FALSE}
 # Install the packages which is necessary to obtain and process datasets.
@@
 ```{r, message= FALSE, warning=FALSE}
-library(readr)
+
 read_tsv("hsa_mirna.txt", comment = "#", col_names = FALSE)%>%
@@
 ```{r}
-library(ceRNAnetsim)
+
 as.data.frame(E9GE_mirnagenenormal)%>%
@@
  simulate(50) %>%
- find_iteration(limit=0, plot= FALSE)
+ find_iteration(limit=1, plot= FALSE)
 
@@
   update_how("SERPINE2",2.75)%>%
-  simulate(32)
+  simulate(62)
 
@@
   update_how("SERPINE2",2.75)%>%
-  simulate(32)%>%
+  simulate(62)%>%
   as_tibble()%>%
@@
 This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and coherent miRNA:target regulation approach.
-
-# 3. REFERENCES

### 2019-09-23 Supplementary_tables.Rmd
article edit

@@
 title: "Supplementary Tables"
-author: "Selcen Arı"
+author: 
+- Selcen Ari
+- Alper Yilmaz 
 date: "16 09 2019"
@@
 header-includes: 
-  \usepackage{float} \floatplacement{figure}{H} 
+  \usepackage{tabularx}
+  \usepackage{float}
+  \usepackage{caption}
   \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
@@
 
+\beginsupplement
+
 ```{r, eval=FALSE, include=FALSE}
-#install.packages("kable")
 library(xtable)
+knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
+library(gt)
+
 ```
-
-\beginsupplement
 
@@
 
+ \begin{table}[ht]
+\centering
+\caption{ miRNA:target pairs supported by High-throughput Experiments}
+\begin{tabular}{lll}
+  \hline
+ Variable & Structure & Means \\
+  \hline
+   cluster    & character & Barcode from experimentally method \\
+   chromosome    & character & Chromosome of Terget gene from raw data \\
+   start\_position    & numeric & Gene start position from raw data \\
+   end\_position     & numeric & Gene end position from raw data \\
+   strand     & chracter & Gene strand \\
+   hgnc\_symbol     & character & Gene name (Symbol) \\
+   Ensembl\_Gene\_Id     & character & Ensembl Gene Id of gene \\
+   Ensembl\_Transcript\_Id   & character & Ensembl transcript id of mRNA of Target gene \\
+   target\_seq   & character & mRNA sequences targeted by miRNA \\
+   miRNA   & character & miRNA id (from miRBase vesion 21 ) \\
+   miR\_seq   & character & miRNA sequence \\
+   seed\_type   & character & seed type of miRNA:target interaction \\
+   Energy   & numeric & Energy of miRNA:target binding \\
+   HG38build\_loc   & character & Recent chromosomal location of Gene \\
+   Genom\_build   & character& Genom build of given chromosome, start and end positions \\
+   region   & character & interaction location on target \\ 
+   region\_effect   & numeric & Coefficient of location on target \\ 
+   seed\_type\_effect   & numeric & Coefficient for seed sequence of miRNA:target interaction \\
+   \hline
+\end{tabular}
+\end{table}
+
 As a result of simulation a dataset, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at followings.
 
-|Variebles| Structure | Means|
-|-----------|------------|------------|
-| **Node Variables**
-| `name`  | character | node name |
-| `type`  | character |Competing or miRNA |
-| `node_id`  | numeric | in on graph object |
-| `initial_count`   | numeric | Initial Expression value of node |
-| `count_pre`   | numeric | Expression value of node at previous regulation |
-| `count_current`   | numeric | Existing expression value of node |
-| `changes_variable`   | character | Regulation of node (Up, down or steady) |
-|-----------|------------|------------|
-| **Edge Variables**
-| `Competing name` | character | name of genes|
-| `miRNA name` | character | name of miRNAs|
-| `Competing expression` | numeric | Expression values of competing elements at steady-state |
-| `miRNA expression` | numeric | Expression values of miRNA elements at steady-state |
-| `energy` | numeric | coefficient of miRNA:target interactions (binding affinity)|
-| `seed type` | numeric | coefficient of miRNA:target interactions (binding affinity)|
-| `region` | numeric | coefficient of miRNA:target interactions (degradation efficiency)|
-| `afff factor` | numeric| coefficient scaled and combined affinity factor |
-| `degg factor` | numeric |coefficient scaled and combined degradation factor| 
-| `comp_count_list` | list | list of competing expression for each iteration | 
-| `comp_count` | numeric | _pre: competing expression at previous iteration; _current: competing expression at present iteration |
-| `mirna_count_list` | list | list of miRNA expression for each iteration | 
-| `mirna_count` | numeric | _pre: miRNA expression at previous iteration; _current: miRNA expression at present iteration |
-| `effect` | numeric | _pre: total miRNA reppressive effect on individual target at previous iteration ; _current: miRNA reppressive effect on individual target at present iteration |
-| `effect_list` | list | list of miRNA reppressive effect on individual target for each iteration |
+ \begin{table}[ht]
+\centering
+\caption{ The context graph object during the process.}
+\begin{tabularx}{\textwidth}{llX}
+  \hline
+ Variables & Structure & Means \\
+  \hline
+   \emph{Node Variables} & & \\
+   name    & character & node name \\
+   type    & character &Competing or miRNA \\
+   node\_id    & numeric & in on graph object \\
+   initial\_count     & numeric & Initial Expression value of node \\
+   count\_pre     & numeric & Expression value of node at previous regulation \\
+   count\_current     & numeric & Existing expression value of node \\
+   changes\_variable     & character & Regulation of node (Up, down or steady) \\
+   \hline
+   \emph{Edge Variables} & & \\
+   Competing name   & character & name of genes\\
+   miRNA name   & character & name of miRNAs \\
+   Competing expression   & numeric & Expression values of competing elements at steady-state \\
+   miRNA expression   & numeric & Expression values of miRNA elements at steady-state \\
+   energy   & numeric & coefficient of miRNA:target interactions (binding affinity) \\
+   seed type   & numeric & coefficient of miRNA:target interactions (binding affinity) \\
+   region   & numeric & coefficient of miRNA:target interactions (degradation efficiency) \\
+   afff factor   & numeric& coefficient scaled and combined affinity factor \\
+   degg factor   & numeric &coefficient scaled and combined degradation factor \\ 
+   comp\_count\_list   & list & list of competing expression for each iteration \\ 
+   comp\_count   & numeric & pre: competing expression at previous iteration; current: competing expression at present iteration \\
+   mirna\_count\_list   & list & list of miRNA expression for each iteration \\
+   mirna\_count   & numeric & pre: miRNA expression at previous iteration; current: miRNA expression at present iteration \\
+   effect   & numeric & pre: total miRNA reppressive effect on individual target at previous iteration ; current: miRNA reppressive effect on individual   target at present iteration \\
+   effect\_list   & list & list of miRNA reppressive effect on individual target for each iteration \\
+  \hline
+\end{tabularx}
+\end{table}
 
-Table: The context graph object during the process.
+\newpage
 

### 2019-09-28 NAR_form_article.Rmd
article NAR form

--- 2019-09-17 article.Rmd
@@
 ---
-title: "Network based multifactorial modelling microRNA-target interaction."
+title: Network based multifactorial modelling microRNA-target interaction.
 author:
-- Selcen Ari^[Department of Bioengineering, Yildiz Technical University, Istanbul,
-  Turkey]
-- Alper Yilmaz^[Department of Bioengineering, Yildiz Technical University, Istanbul,
-  Turkey]
-date: "`r format(Sys.time(), '%d %B %Y')`"
-output:
-  pdf_document:
-    keep_tex: yes
-  word_document: default
-  html_document: default
-link-citations: yes
-cls: NAR.cls
-template: NAR-sample.tex
-bibliography: ref.bib
----
+  - name: Selcen Ari
+    affiliation: "1"
+  - name: Alper Yilmaz
+    affiliation: "1"
+    correspond: "To whom correspondence should be addressed. Phone: +90 212 383 4627 Fax: +90 212 383 4625  Email: alyilmaz@yildiz.edu.tr"
+    separator: "and"
+address:
+  - code: 1
+    address: Department of Bioengineering, Yildiz Technical University, Istanbul, Turkey
 
-## Abstract
-
+abstract: |
   Competing endogenous RNA (ceRNA) regulations and crosstalk between various types of non-coding RNA in human is remarkable in means of miRNA regulation.
@@
 The model is scalable and can be plugged in with emerging miRNA effectors such as circRNAs.
-The model is available as R package [ceRNAnetsim](https://github.com/selcenari/ceRNAnetsim).
+The model is available as R package ceRNAnetsim https://github.com/selcenari/ceRNAnetsim.
 
-## Introduction
+bibliography: NAR-sample.bib
+output: rticles::nar_article
+header-includes: 
+  \usepackage{float} \floatplacement{figure}{H} 
+  \usepackage[colorlinks=true,
+            linkcolor=blue,
+            urlcolor=blue,
+            citecolor=blue]{hyperref}
+  \usepackage[T1]{fontenc}
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE)
+```
+
+# INTRODUCTION
 
@@
 Various computational and experimental studies have tackled the problem of unraveling ceRNA:miRNA interactions.
+
+\enlargethispage{-65.1pt}
+
 It has been observed that when abundance of one of the targets of miR-122 was increased, the other targets' expression also slightly increased as a result of decreasing repressive activity of miR-122 on them [@denzler_assessing_2014].
@@
 
-## Methods
-### Construction of miRNA:target network
+# MATERIALS AND METHODS
+
+## Construction of miRNA:target network
 miRNA and target pairs per line should be provided as edge list to construct the network.
@@
 
-### Triggering perturbation and subsequent calculations
-Initially, the network is assumed in steady-state (\autoref{fig1}a) condition and needs least one trigger for initiating calculations.
-The trigger can be a change in expression level of one or more genes (\autoref{fig1}b).
-After a trigger, the network undergoes iterative cycle of calculations at each of which distribution of miRNA in local neighborhood is recalculated (\autoref{fig1}c).
+## Triggering perturbation and subsequent calculations
+Initially, the network is assumed in steady-state (Figure \ref{fig:fig1}A) condition and needs least one trigger for initiating calculations.
+The trigger can be a change in expression level of one or more genes (Figure \ref{fig:fig1}B).
+After a trigger, the network undergoes iterative cycle of calculations at each of which distribution of miRNA in local neighborhood is recalculated (Figure \ref{fig:fig1}C).
 Based on new miRNA distribution, expression level of each node (i.e.
@@
 In this case, the common target acts as trigger for the other group (the targets interacted with M2 miRNA) in the network.
-Due to common targeted elements, the change in one neighborhood spreads to other neighborhoods (\autoref{fig1}d), consequently have potential to affect whole network due to "ripple effect".
+Due to common targeted elements, the change in one neighborhood spreads to other neighborhoods (Figure \ref{fig:fig1}D), consequently have potential to affect whole network due to "ripple effect".
 
-![Schematic presentation of mechanism of network based model.
-a) In steady state, miRNAs (M: triangles) repress targets (G: circles) according to proportion of target expression level.
-b) Two fold increase in transcript level of Gene2 (G2) acts as a trigger (shown in red).
-c) Distribution of miRNA1 (M1) changes.
-d) The change at expression of common target affects changes of proportional distribution of miRNA2 (M2).
+\begin{figure*}[ht]
+\begin{center}
+\includegraphics[width=15cm,height=10cm]{Fig1.jpg}
+\end{center}
+\caption{Schematic presentation of mechanism of network based model. 
+    \textbf{A)} In steady state, miRNAs (M: triangles) repress targets (G: circles) according to proportion of target expression level. 
+    \textbf{B)} Two fold increase in transcript level of Gene2 (G2) acts as a trigger (shown in red). 
+    \textbf{C)} Distribution of miRNA1 (M1) changes. 
+    \textbf{D)} The change at expression of common target affects changes of proportional distribution of miRNA2 (M2).
 Expression values are rounded to integers for simplicity.
 Gray values on edges indicate initial distribution, red values indicate increase and blue values indicate decrease in distribution.
-Shades of circles indicate different levels of increase.](Fig1.jpg){#fig1}
+Shades of circles indicate different levels of increase.}
+\label{fig:fig1}
+\end{figure*}
 
@@
 The repression efficiency of a miRNA on the individual target ($Eff_{gi}$) is calculated according to equation \eqref{eq:1}; where miRNA expression ($C_m$) in local neighborhood is distributed among targets using individual gene expression levels ($C_{gi}$).
-For the genes targeted by multiple miRNAs, cooperative activity of miRNAs on a target gene, $R$, is calculated by summing repression activity of each miRNA (Equation \eqref{eq:2}).
+For the genes targeted by multiple miRNAs, cooperative activity of miRNAs on a target gene, $R$, is calculated by summing repression activity of each miRNA Equation \eqref{eq:2}.
+\begin{equation*} 
+    Eff_{gi}= C_m \times C_{gi}/\sum_{1}^{i} C_{gi} \tag{1}\label{eq:1}
+\end{equation*}
 
-\begin{equation} 
-    Eff_{gi}= C_m \times C_{gi}/\sum_{1}^{i} C_{gi} \tag{1}\label{eq:1}
-\end{equation}
+\begin{equation*}
+   R_{gi}= Eff_{i1} + Eff_{i2} ... \tag{2}\label{eq:2}
+\end{equation*}
 
-\begin{equation}
-   R_{gi}= Eff_{i1} + Eff_{i2} ... \tag{2}\label{eq:2}
-\end{equation}
-
-### Multifactorial calculations in miRNA:target network
+## Multifactorial calculations in miRNA:target network
 Interactions between miRNAs and their targets can be affected from various factors.
@@
 Since binding or degradation efficiency from literature have various and broad ranges, both binding and efficiency factors are normalized to their maximum values and scaled to [0,1] interval.
-The normalized values of factors are used to determine binding activity and miRNA efficiency on targets (\autoref{fig2}).
-Binding affinities (activity, $Eff$) of miRNAs on each individual gene are calculated as shown in equation \eqref{eq:3}; $C_m$, miRNA expression in the group; $C_{gi}$, individual gene expression; gi, individual gene (\autoref{fig2}c).
+The normalized values of factors are used to determine binding activity and miRNA efficiency on targets (Figure \ref{fig:fig2}).
+Binding affinities (activity, $Eff$) of miRNAs on each individual gene are calculated as shown in equation \eqref{eq:3}; $C_m$, miRNA expression in the group; $C_{gi}$, individual gene expression; gi, individual gene (Figure \ref{fig:fig2}C).
+
 \begin{equation}
@@
 
-![Calculations to determine of miRNA binding and repression efficiency.
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=9cm,height=17cm]{Fig2.jpg}
+\end{center}
+\caption{Calculations to determine of miRNA binding and repression efficiency.
 $G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE^\prime$, normalized values of seed type efficiency coefficient; $RE^\prime$, normalized values of region efficiency coefficient; $E^\prime$, normalized values of energy coefficient.
-Numbering on edges match the pair order in \autoref{tab:one}](fig2.jpg){#fig2}
+Numbering on edges match the pair order in Table \ref{tab:one} }
+\label{fig:fig2}
+\end{figure}
 
@@
 The degradation of target by bound miRNA depends on efficiency factors such as binding region.
-Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:4} (\autoref{fig2}d); $RE^\prime_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
-The cooperative repression activity of miRNAs to their common targets is figured out as shown in \autoref{fig2}e.
+Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:4} (Figure \ref{fig:fig2}D); $RE^\prime_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
+The cooperative repression activity of miRNAs to their common targets is figured out as shown in Figure \ref{fig:fig2}E.
 
-### Breast cancer patient dataset
+## Breast cancer patient dataset
 
 We have applied our model in a real dataset for which experimental measurements of various factors were available.
-Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network: https://www.cancer.gov/tcga.
+Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network https://www.cancer.gov/tcga.
 High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015].
@@
 
-## Results and Conclusions
-### Network with miRNA:ceRNA expression level
+# RESULTS
+
+## Network with miRNA:ceRNA expression level
 We have developed a network-based approach to assess effects of expression level changes in competitive ceRNA regulation.
@@
 Our approach can effortlessly calculate effects of expression changes when abundance levels of miRNAs and targets is only available factor.
-In sample network given in Figure (\autoref{fig1}), after an increase in expression level of a gene (G2), expression values of other genes also changed due to redistribution of miRNA among its targets.
+In sample network given in Figure (\ref{fig:fig1}), after an increase in expression level of a gene (G2), expression values of other genes also changed due to redistribution of miRNA among its targets.
 Previous studies have shown that if a gene abundance increases in ceRNA system, expression levels of genes targeted by shared miRNA are also affected [@lai_understanding_2016; @salmena_cerna_2011; @tay_multilayered_2014].
@@
 However, changes in expression levels of G4 and G1 differ although they have same initial conditions.
-This is due to fact that G4 is targeted by two miRNAs, M1 and M2 (\autoref{fig1}d).
+This is due to fact that G4 is targeted by two miRNAs, M1 and M2 (Figure \ref{fig:fig1}D).
 
@@
 
-![Target regulations with interaction parameters.
-a) In the steady-state the repression activity of miRNAs on the targets after binding and repression efficiency.
-b) The changes the repression activities after increasing of G2 expression.
-c) Perturbation of primary neighborhoods of M1 miRNA (M1 miRNA group).
-d) Regulation of gene expression of other gene group via triggering target (common target between M1 and M2).
-](Fig3.jpg){#fig3}
+\begin{figure*}[ht]
+\begin{center}
+\includegraphics[width=15cm,height=8cm]{Fig3.jpg}
+\end{center}
+\caption{Target regulations with interaction parameters. 
+     \textbf{A)} In the steady-state the repression activity of miRNAs on the targets after binding and repression efficiency. 
+     \textbf{B)} The changes the repression activities after increasing of G2 expression. 
+     \textbf{C)} Perturbation of primary neighborhoods of M1 miRNA (M1 miRNA group). 
+     \textbf{D)} Regulation of gene expression of other gene group via triggering target (common target between M1 and M2)}
+\label{fig:fig3}
+\end{figure*}
 
-### ceRNA:Target networks based on interaction factors
+## ceRNA:Target networks based on interaction factors
 Earlier studies reported that miRNA regulatory interactions are affected by different parameters.
@@
 Therefore, we have integrated aforementioned interaction parameters that could be useful for predicting miRNA repression activity more accurately.
-The sample dataset used in \autoref{fig1} is recalculated with additional factors \autoref{tab:one} in effect with same trigger, two fold increase in Gene2 (\autoref{fig3}).
-When the factors were taken into account in the system, miRNA efficiencies varied as shown in \autoref{fig3}a.
+The sample dataset used in Figure \ref{fig:fig1} is recalculated with additional factors Table \ref{tab:one} in effect with same trigger, two fold increase in Gene2 (Figure \ref{fig:fig3}).
+When the factors were taken into account in the system, miRNA efficiencies varied as shown in Figure \ref{fig:fig3}A.
 Although the miRNA:target expression ratios in steady-state were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed.
@@
 
-When the model run based on expression values of miRNAs and ceRNAs (genes), proportional distribution of miRNA on targets is determined in accordance with \eqref{eq:1, eq:2}.
-In that case, the system is observed as shown in \autoref{fig1}a at steady-state.
-However, if interaction factors are taken into account in the approach, specified interaction factors (i.e, binding factors such as seed type and energy) affect distribution of miRNA on targets as calculated in \eqref{eq:3} (shown in \autoref{fig2}c).
-After distributed miRNA efficiency was re-calculated with degradation factor(s) (i.e region effect) according to \eqref{eq:4}, steady-state condition of network is obtained (shown in \autoref{fig2}d and \autoref{fig3}a).
-For instance, proportional distribution of G1:M1 interaction in \autoref{fig1}a is calculated lower than the same pair (G1:M1) in the other approach that includes interaction factors (\autoref{fig2}c).
-But, a decrease is observed in repression efficiency of M1 miRNA on G1 in interaction factor based network approach (\autoref{fig2}d and \autoref{fig3}a).
+When the model run based on expression values of miRNAs and ceRNAs (genes), proportional distribution of miRNA on targets is determined in accordance with equations \eqref{eq:1} \eqref{eq:2}.
+In that case, the system is observed as shown in Figure \ref{fig:fig1}A at steady-state.
+However, if interaction factors are taken into account in the approach, specified interaction factors (i.e, binding factors such as seed type and energy) affect distribution of miRNA on targets as calculated in equation \eqref{eq:3} (shown in Figure \ref{fig:fig2}C).
+After distributed miRNA efficiency was re-calculated with degradation factor(s) (i.e region effect) according to equation \eqref{eq:4}, steady-state condition of network is obtained (shown in Figure \ref{fig:fig2}D and Figure \ref{fig:fig3}A).
+For instance, proportional distribution of G1:M1 interaction in Figure \ref{fig:fig1}A is calculated lower than the same pair (G1:M1) in the other approach that includes interaction factors (Figure \ref{fig:fig2}C).
+But, a decrease is observed in repression efficiency of M1 miRNA on G1 in interaction factor based network approach (Figure \ref{fig:fig2}D and \ref{fig:fig3}A).
 
-It is considered that entire miRNAs in the system affect targets according to target:total target ratio \autoref{fig1}a, when interaction factors did not take into account.
-However, in presence of interaction factors in miRNA:target interaction network (shown in \autoref{tab:one}), miRNAs are distributed according to target:total target ratio and affinity factors in first like shown in \autoref{fig2}c.
-After affinity mediated proportional distribution of miRNA expression, degradation factor is considered to specify count of repressive miRNA in pairs because entire bound miRNA target pairs might not be resulted with degradation (It is figured out shown in \autoref{fig2}d).
-For this reason, miRNA1 (M1) has more weak repressive effect (shown as edge variables) on target Gene2 \autoref{fig3}a in comparison with miRNA1 (M1):Gene2 (G2) interaction in \autoref{fig1}a.
-On the other hand, it is possible that taking into account of these interaction factors could cause increasing miRNA repression activity like in miRNA2:Gene4 interaction \autoref{fig3}a.
-When expression of Gene2 (G2) increased, expression values of all genes also changed differentially because of contribution of efficiency factors (\autoref{fig3} b,c,d).
+It is considered that entire miRNAs in the system affect targets according to target:total target ratio Figure \ref{fig:fig1}A, when interaction factors did not take into account.
+However, in presence of interaction factors in miRNA:target interaction network (shown in Table \ref{tab:one}), miRNAs are distributed according to target:total target ratio and affinity factors in first like shown in Figure \ref{fig:fig2}C.
+After affinity mediated proportional distribution of miRNA expression, degradation factor is considered to specify count of repressive miRNA in pairs because entire bound miRNA target pairs might not be resulted with degradation (It is figured out shown in Figure \ref{fig:fig2}D).
+For this reason, miRNA1 (M1) has more weak repressive effect (shown as edge variables) on target Gene2 Figure \ref{fig:fig3}A in comparison with miRNA1 (M1):Gene2 (G2) interaction in Figure \ref{fig:fig1}A.
+On the other hand, it is possible that taking into account of these interaction factors could cause increasing miRNA repression activity like in miRNA2:Gene4 interaction Figure \ref{fig:fig3}A.
+When expression of Gene2 (G2) increased, expression values of all genes also changed differentially because of contribution of efficiency factors (Figure \ref{fig:fig3} B,C,D).
 We have considered that in our approach energy and seed type of pairs is significant for binding and targeted region is important for repression.
 
-\begin{table}[ht]
+\begin{table*}[ht]
 \centering
-\caption{Expression values of elements and interaction factors of miRNA:target interactions in \autoref{fig3}} 
+\caption{Expression values of elements and interaction factors of miRNA:target interactions} 
 \begin{tabular}{rllrrrrr}
@@
 \label{tab:one}
-\end{table}
+\end{table*}
 
-In regulation of miRNA target sample system (\autoref{fig3}), miRNA (M1) repressive efficiency on the primary triggering gene (G2) is low in steady-state.
+In regulation of miRNA target sample system (Figure \ref{fig:fig3}), miRNA (M1) repressive efficiency on the primary triggering gene (G2) is low in steady-state.
 So it has been observed that the regulatory activities of miRNA on the targets are weak after the increase of miRNA target G2.
@@
 Since perturbation efficiency of each gene is different, we have developed a function which screens each gene in the network for their perturbation efficiencies.
-**(see details in Supplementary Materials and Methods**.
+*(see details in Supplementary Materials and Methods*.
 On the other hand, when we applied the method on the minimal dataset, Gene4 has been found to be the most efficient element in terms of number of perturbed elements and miRNA2 (M2) has been found to be causing the highest mean in expression changes.
 
-### Simulation of large and realistic network
+## Simulation of large and realistic network
 
@@
 The network was finalized by integrating gene and miRNA expression levels from TCGA determined by RNA-Seq.
-Our  model can successfully simulate perturbations in such large network despite complex behaviors and struggle to reach steady-state **(see details in Supplementary Materials and Methods)**.
+Our  model can successfully simulate perturbations in such large network despite complex behaviors and struggle to reach steady-state *(see details in Supplementary Materials and Methods)*.
 Simulations show that change in expression level of single gene has potential to affect whole network, perturbing almost all nodes.
@@
 
-## Discussion
+# DISCUSSION
 
@@
 
-## Conclusion
+# CONCLUSION
 
@@
 
-## References
+# ACKNOWLEDGEMENTS
+
+Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
+Text. Text. Text. Text.
+
+### Conflict of interest statement.
+
+None declared.
+
+\newpage

### 2019-09-28 NAR_form_article.tex
article NAR form

--- 2019-09-28 NAR_form_article.Rmd
@@
----
-title: Network based multifactorial modelling microRNA-target interaction.
-author:
-  - name: Selcen Ari
-    affiliation: "1"
-  - name: Alper Yilmaz
-    affiliation: "1"
-    correspond: "To whom correspondence should be addressed. Phone: +90 212 383 4627 Fax: +90 212 383 4625  Email: alyilmaz@yildiz.edu.tr"
-    separator: "and"
-address:
-  - code: 1
-    address: Department of Bioengineering, Yildiz Technical University, Istanbul, Turkey
+\documentclass[a4,center,fleqn]{NAR}
 
-abstract: |
-  Competing endogenous RNA (ceRNA) regulations and crosstalk between various types of non-coding RNA in human is remarkable in means of miRNA regulation.
-Many studies have pointed out that an alteration in miRNA:target interaction can result in unexpected changes due to indirect and complex interactions.
-In this paper, we defined a new network-based model that handles miRNA:ceRNA interactions with expression values.
-Our model is able to handle miRNA interaction factors such as seed type, binding energy, if provided.
-Our approach is able to reveal that a perturbation in an element of network affects whole competing elements differently and cooperative efficiencies of miRNAs on common targets could be calculated.
-Our findings emphasized importance of miRNA:target ratios being crucial, as reported by previous studies.
-We have showed that the competing elements which have the same or close expression values may not be affected equally from the perturbation because of repression functionality depended on interaction factors of miRNA target pairs.
-We applied the model to real sample consisting of breast cancer gene and miRNA expression dataset and experimental miRNA:target interaction dataset all generated via high throughput sequencing methods.
-A gene over-expressed in tumor tissue, namely *SERPINE2*, is used as perturbing element.
-We have observed that change in expression level of single gene in miRNA:target network is sufficient to perturb regulations in whole network, due to unforeseen and unpredicted regulation which are only visible when considered in network context.
-Therefore, this model helps unveiling the crosstalk between elements in miRNA:target network where abundance of target and sponge effect are taken into account.
-The model is scalable and can be plugged in with emerging miRNA effectors such as circRNAs.
-The model is available as R package ceRNAnetsim https://github.com/selcenari/ceRNAnetsim.
+% Enter dates of publication
+\copyrightyear{2008}
+\pubdate{31 July 2009}
+\pubyear{2009}
+\jvolume{37}
+\jissue{12}
 
-bibliography: NAR-sample.bib
-output: rticles::nar_article
-header-includes: 
-  \usepackage{float} \floatplacement{figure}{H} 
-  \usepackage[colorlinks=true,
-            linkcolor=blue,
-            urlcolor=blue,
-            citecolor=blue]{hyperref}
-  \usepackage[T1]{fontenc}
----
+%\articlesubtype{This is the article type (optional)}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-```
+\usepackage{float}
 
-# INTRODUCTION
+\floatplacement{figure}{H}
 
-MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key regulator of gene expression through various post-transcriptional mechanisms.
-Although the mechanisms by which miRNA represses are not fully understood, miRNAs predominantly repress their targets.
-Repressive activities of miRNAs vary depending on many factors that are significant to miRNA:target interactions.
-These factors include miRNA:target binding energy, binding location in target sequence, base pairing types between miRNA and target, abundance of miRNAs and targets [@grimson_microrna_2007].
-Binding energies of miRNA:target complexes vary based on nucleotide context and determine folding stability of complex [@cao_predicting_2012].
-It has been demonstrated that the binding energy between miRNA and target indicates stability or affinity of complex [@helwak_mapping_2013] and does not directly determine repressive activity of miRNA [@cao_predicting_2012].
-Early studies have argued that 2-8 nt sequence, seed, located in miRNA 5'end bind to specific sequence located in 3'UTR of its target [@bartel_micrornas_2004; @lewis_conserved_2005].
-In recent studies, it has been shown that miRNAs can interact with targets via sequences located in regions such as 5’UTR or CDS [@hausser_analysis_2013; @helwak_mapping_2013; @moore_mirnatarget_2015].
-These studies also showed that  binding location either indicate functionality of miRNA:target interaction or affect abundance of targets.
-It has been shown that miRNAs exhibit repressive activity via, 6-8 nt long sequence that is perfectly complementary with targets, seed region at the 5’ end of miRNAs [@bartel_micrornas:_2009; @grimson_microrna_2007].
-On the other hand, some researchers have reported that seed sequence of miRNA can have mismatches or bulged/wobble nucleotides [@chi2012alternative] and may locate in region other than 5'end of miRNAs [@hafner_transcriptome-wide_2010; @helwak_mapping_2013].
-On top of all these factors, abundance of miRNAs and targets and miRNA:target ratio in cells predominantly affect efficiency of miRNA:target interaction [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
+\usepackage[colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue]{hyperref} \usepackage[T1]{fontenc}
 
-As it is possible for miRNAs to suppress multiple targets, an individual mRNA molecule can also be targeted by multiple miRNAs.
-In that case, the targeted mRNAs exhibit competitor behavior, that is hypothesized as competing endogenous RNAs (ceRNAs) [@ala_integrated_2013; @cesana_deciphering_2013], against their miRNAs.
-Briefly, Ala et al.
-have explained the ceRNA hypothesis as disturbance of the other target when one of the targets on a steady-state system that included one miRNA and two target was perturbed with expression change [@ala_integrated_2013].
-Regarding interaction between miRNAs and their target in a cell, explaining and predicting results of an individual perturbation is difficult due to complexity of interactions.
-Various computational and experimental studies have tackled the problem of unraveling ceRNA:miRNA interactions.
+\begin{document}
+
+\title{Network based multifactorial modelling microRNA-target interaction.}
+
+\author{
+    Selcen Ari\,$^{1}$
+   and Alper Yilmaz\,$^{1}$ \footnote{To whom correspondence should be addressed. Phone: +90 212 383 4627 Fax:
++90 212 383 4625 Email:
+\href{mailto:alyilmaz@yildiz.edu.tr}{\nolinkurl{alyilmaz@yildiz.edu.tr}}} 
+}
+
+\address{
+   $^{1}$ Department of Bioengineering, Yildiz Technical University, Istanbul,
+Turkey 
+}
+
+\history{%
+Received January 1, 2009;
+Revised February 1, 2009;
+Accepted March 1, 2009}
+
+\maketitle
+
+\begin{abstract}
+Competing endogenous RNA (ceRNA) regulations and crosstalk between
+various types of non-coding RNA in human is remarkable in means of miRNA
+regulation. Many studies have pointed out that an alteration in
+miRNA:target interaction can result in unexpected changes due to
+indirect and complex interactions. In this paper, we defined a new
+network-based model that handles miRNA:ceRNA interactions with
+expression values. Our model is able to handle miRNA interaction factors
+such as seed type, binding energy, if provided. Our approach is able to
+reveal that a perturbation in an element of network affects whole
+competing elements differently and cooperative efficiencies of miRNAs on
+common targets could be calculated. Our findings emphasized importance
+of miRNA:target ratios being crucial, as reported by previous studies.
+We have showed that the competing elements which have the same or close
+expression values may not be affected equally from the perturbation
+because of repression functionality depended on interaction factors of
+miRNA target pairs. We applied the model to real sample consisting of
+breast cancer gene and miRNA expression dataset and experimental
+miRNA:target interaction dataset all generated via high throughput
+sequencing methods. A gene over-expressed in tumor tissue, namely
+\emph{SERPINE2}, is used as perturbing element. We have observed that
+change in expression level of single gene in miRNA:target network is
+sufficient to perturb regulations in whole network, due to unforeseen
+and unpredicted regulation which are only visible when considered in
+network context. Therefore, this model helps unveiling the crosstalk
+between elements in miRNA:target network where abundance of target and
+sponge effect are taken into account. The model is scalable and can be
+plugged in with emerging miRNA effectors such as circRNAs. The model is
+available as R package ceRNAnetsim
+\url{https://github.com/selcenari/ceRNAnetsim}.
+\end{abstract}
+
+\section{INTRODUCTION}
+
+MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key
+regulator of gene expression through various post-transcriptional
+mechanisms. Although the mechanisms by which miRNA represses are not
+fully understood, miRNAs predominantly repress their targets. Repressive
+activities of miRNAs vary depending on many factors that are significant
+to miRNA:target interactions. These factors include miRNA:target binding
+energy, binding location in target sequence, base pairing types between
+miRNA and target, abundance of miRNAs and targets
+\citep{grimson_microrna_2007}. Binding energies of miRNA:target
+complexes vary based on nucleotide context and determine folding
+stability of complex \citep{cao_predicting_2012}. It has been
+demonstrated that the binding energy between miRNA and target indicates
+stability or affinity of complex \citep{helwak_mapping_2013} and does
+not directly determine repressive activity of miRNA
+\citep{cao_predicting_2012}. Early studies have argued that 2-8 nt
+sequence, seed, located in miRNA 5'end bind to specific sequence located
+in 3'UTR of its target
+\citep{bartel_micrornas_2004, lewis_conserved_2005}. In recent studies,
+it has been shown that miRNAs can interact with targets via sequences
+located in regions such as 5'UTR or CDS
+\citep{hausser_analysis_2013, helwak_mapping_2013, moore_mirnatarget_2015}.
+These studies also showed that binding location either indicate
+functionality of miRNA:target interaction or affect abundance of
+targets. It has been shown that miRNAs exhibit repressive activity via,
+6-8 nt long sequence that is perfectly complementary with targets, seed
+region at the 5' end of miRNAs
+\citep{bartel_micrornas:_2009, grimson_microrna_2007}. On the other
+hand, some researchers have reported that seed sequence of miRNA can
+have mismatches or bulged/wobble nucleotides \citep{chi2012alternative}
+and may locate in region other than 5'end of miRNAs
+\citep{hafner_transcriptome-wide_2010, helwak_mapping_2013}. On top of
+all these factors, abundance of miRNAs and targets and miRNA:target
+ratio in cells predominantly affect efficiency of miRNA:target
+interaction
+\citep{arvey_target_2010, bosson_endogenous_2014, denzler_assessing_2014}.
+
+As it is possible for miRNAs to suppress multiple targets, an individual
+mRNA molecule can also be targeted by multiple miRNAs. In that case, the
+targeted mRNAs exhibit competitor behavior, that is hypothesized as
+competing endogenous RNAs (ceRNAs)
+\citep{ala_integrated_2013, cesana_deciphering_2013}, against their
+miRNAs. Briefly, Ala et al.~have explained the ceRNA hypothesis as
+disturbance of the other target when one of the targets on a
+steady-state system that included one miRNA and two target was perturbed
+with expression change \citep{ala_integrated_2013}. Regarding
+interaction between miRNAs and their target in a cell, explaining and
+predicting results of an individual perturbation is difficult due to
+complexity of interactions. Various computational and experimental
+studies have tackled the problem of unraveling ceRNA:miRNA interactions.
 
@@
 
-It has been observed that when abundance of one of the targets of miR-122 was increased, the other targets' expression also slightly increased as a result of decreasing repressive activity of miR-122 on them [@denzler_assessing_2014].
-Bosson et al.
-have developed a mathematical model for changes on total target pool concentration after grouping targets according to affinity and demonstrated that miRNA activity correlated with affinity between miRNA and target [@bosson_endogenous_2014].
-Cooperative efficiency of miRNAs as well as competitor behaviors of targets were also studied and it has been demonstrated to be crucial for regulating available mRNA levels of targets [@denzler_impact_2016].
-MiRNA:target interactions have been modeled as stoichiometric and catalytic mechanisms and Figliuzzi et al.
-have recommended handling models in network context [@figliuzzi_micrornas_2013].
-The model that can explain miRNA target interaction through topological features has been applied at bipartite network by Nitzan et al.
-[@nitzan_interactions_2014].
-Robinson and Henderson applied the model that handles miRNA:target direct and indirect interactions via common targeting miRNA of genes and target of miRNAs, at bipartite network.
-It has been demonstrated that all miRNAs and targets in the network can interact with each other through common miRNAs and genes, without interaction between the same type of nodes [@robinson_modelling_2018].Associated genes that are targets of the same miRNAs have been found with help of correlation of gene expression changes in recent algorithm [@markus_list_sponge_2017].
-List et al.
-have specified that their approach can be useful for ceRNA studies and published their approach as an R package.
+It has been observed that when abundance of one of the targets of
+miR-122 was increased, the other targets' expression also slightly
+increased as a result of decreasing repressive activity of miR-122 on
+them \citep{denzler_assessing_2014}. Bosson et al.~have developed a
+mathematical model for changes on total target pool concentration after
+grouping targets according to affinity and demonstrated that miRNA
+activity correlated with affinity between miRNA and target
+\citep{bosson_endogenous_2014}. Cooperative efficiency of miRNAs as well
+as competitor behaviors of targets were also studied and it has been
+demonstrated to be crucial for regulating available mRNA levels of
+targets \citep{denzler_impact_2016}. MiRNA:target interactions have been
+modeled as stoichiometric and catalytic mechanisms and Figliuzzi et
+al.~have recommended handling models in network context
+\citep{figliuzzi_micrornas_2013}. The model that can explain miRNA
+target interaction through topological features has been applied at
+bipartite network by Nitzan et al.~\citep{nitzan_interactions_2014}.
+Robinson and Henderson applied the model that handles miRNA:target
+direct and indirect interactions via common targeting miRNA of genes and
+target of miRNAs, at bipartite network. It has been demonstrated that
+all miRNAs and targets in the network can interact with each other
+through common miRNAs and genes, without interaction between the same
+type of nodes \citep{robinson_modelling_2018}.Associated genes that are
+targets of the same miRNAs have been found with help of correlation of
+gene expression changes in recent algorithm
+\citep{markus_list_sponge_2017}. List et al.~have specified that their
+approach can be useful for ceRNA studies and published their approach as
+an R package.
 
-# MATERIALS AND METHODS
+\section{MATERIALS AND METHODS}
 
-## Construction of miRNA:target network
-miRNA and target pairs per line should be provided as edge list to construct the network.
-At each line minimum required information is expression levels of miRNA and the target.
-If available, additional data about factors effecting binding or efficiency of miRNA can be provided as separate columns.
-After construction of the network, amount of miRNA per target is calculated and kept as edge data.
-Simply, a target will sequester miRNA proportional to its ratio amount among other targets.
-If additional criteria effecting the binding of miRNA to its target is provided, distribution of miRNA will be calculated accordingly.
-Target can be mRNA or any other ceRNA (circRNA, ncRNA, etc.) thus, throughout the manuscript terms target, gene and ceRNA are used interchangeably.
+\subsection{Construction of miRNA:target network}
 
-## Triggering perturbation and subsequent calculations
-Initially, the network is assumed in steady-state (Figure \ref{fig:fig1}A) condition and needs least one trigger for initiating calculations.
-The trigger can be a change in expression level of one or more genes (Figure \ref{fig:fig1}B).
-After a trigger, the network undergoes iterative cycle of calculations at each of which distribution of miRNA in local neighborhood is recalculated (Figure \ref{fig:fig1}C).
-Based on new miRNA distribution, expression level of each node (i.e.
-ceRNA) is updated.
-This update results in change of expression value of common target (G4) in system.
-In this case, the common target acts as trigger for the other group (the targets interacted with M2 miRNA) in the network.
-Due to common targeted elements, the change in one neighborhood spreads to other neighborhoods (Figure \ref{fig:fig1}D), consequently have potential to affect whole network due to "ripple effect".
+miRNA and target pairs per line should be provided as edge list to
+construct the network. At each line minimum required information is
+expression levels of miRNA and the target. If available, additional data
+about factors effecting binding or efficiency of miRNA can be provided
+as separate columns. After construction of the network, amount of miRNA
+per target is calculated and kept as edge data. Simply, a target will
+sequester miRNA proportional to its ratio amount among other targets. If
+additional criteria effecting the binding of miRNA to its target is
+provided, distribution of miRNA will be calculated accordingly. Target
+can be mRNA or any other ceRNA (circRNA, ncRNA, etc.) thus, throughout
+the manuscript terms target, gene and ceRNA are used interchangeably.
+
+\subsection{Triggering perturbation and subsequent calculations}
+
+Initially, the network is assumed in steady-state (Figure
+\ref{fig:fig1}A) condition and needs least one trigger for initiating
+calculations. The trigger can be a change in expression level of one or
+more genes (Figure \ref{fig:fig1}B). After a trigger, the network
+undergoes iterative cycle of calculations at each of which distribution
+of miRNA in local neighborhood is recalculated (Figure \ref{fig:fig1}C).
+Based on new miRNA distribution, expression level of each node
+(i.e.~ceRNA) is updated. This update results in change of expression
+value of common target (G4) in system. In this case, the common target
+acts as trigger for the other group (the targets interacted with M2
+miRNA) in the network. Due to common targeted elements, the change in
+one neighborhood spreads to other neighborhoods (Figure
+\ref{fig:fig1}D), consequently have potential to affect whole network
+due to ``ripple effect''.
 
@@
 
-During calculations, following assumptions were adopted; 1) Transcription and degradation rates of miRNAs are steady and equal.
-2) All available miRNAs are recycled as in miRNA:ceRNA binding, target is degraded and miRNA is unaffected.
-3) ceRNA targets also have stable transcription and degradation rates and these rates are equal.
+During calculations, following assumptions were adopted; 1)
+Transcription and degradation rates of miRNAs are steady and equal. 2)
+All available miRNAs are recycled as in miRNA:ceRNA binding, target is
+degraded and miRNA is unaffected. 3) ceRNA targets also have stable
+transcription and degradation rates and these rates are equal.
 
-The repression efficiency of a miRNA on the individual target ($Eff_{gi}$) is calculated according to equation \eqref{eq:1}; where miRNA expression ($C_m$) in local neighborhood is distributed among targets using individual gene expression levels ($C_{gi}$).
-For the genes targeted by multiple miRNAs, cooperative activity of miRNAs on a target gene, $R$, is calculated by summing repression activity of each miRNA Equation \eqref{eq:2}.
-\begin{equation*} 
+The repression efficiency of a miRNA on the individual target
+(\(Eff_{gi}\)) is calculated according to equation \eqref{eq:1}; where
+miRNA expression (\(C_m\)) in local neighborhood is distributed among
+targets using individual gene expression levels (\(C_{gi}\)). For the
+genes targeted by multiple miRNAs, cooperative activity of miRNAs on a
+target gene, \(R\), is calculated by summing repression activity of each
+miRNA Equation \eqref{eq:2}. \begin{equation*} 
     Eff_{gi}= C_m \times C_{gi}/\sum_{1}^{i} C_{gi} \tag{1}\label{eq:1}
@@
 
-## Multifactorial calculations in miRNA:target network
-Interactions between miRNAs and their targets can be affected from various factors.
-So, our model integrates multiple factors when calculating overall miRNA activity.
-We classified factors into two categories.
-Binding factors determine interaction between miRNA and target and they alter amount of miRNA sequestered to target.
-Efficiency factors dictate degradation efficiency of miRNA on its target and define amount of target to be degraded.
-In other words, binding factors exert their influence before or during binding, efficiency factors exert their influence after binding.
-In the literature, binding free energy [@cao_predicting_2012 ; @helwak_mapping_2013] and seed type [@werfel_preferential_2017] in miRNA:target interactions are described as factors effecting binding affinity.
-Also, binding region (e.g, 5'UTR, 3'UTR, CDS) was shown to drastically affect miRNA degradation efficiency [@hausser_analysis_2013; @helwak_mapping_2013].
-Since binding or degradation efficiency from literature have various and broad ranges, both binding and efficiency factors are normalized to their maximum values and scaled to [0,1] interval.
-The normalized values of factors are used to determine binding activity and miRNA efficiency on targets (Figure \ref{fig:fig2}).
-Binding affinities (activity, $Eff$) of miRNAs on each individual gene are calculated as shown in equation \eqref{eq:3}; $C_m$, miRNA expression in the group; $C_{gi}$, individual gene expression; gi, individual gene (Figure \ref{fig:fig2}C).
+\subsection{Multifactorial calculations in miRNA:target network}
+
+Interactions between miRNAs and their targets can be affected from
+various factors. So, our model integrates multiple factors when
+calculating overall miRNA activity. We classified factors into two
+categories. Binding factors determine interaction between miRNA and
+target and they alter amount of miRNA sequestered to target. Efficiency
+factors dictate degradation efficiency of miRNA on its target and define
+amount of target to be degraded. In other words, binding factors exert
+their influence before or during binding, efficiency factors exert their
+influence after binding. In the literature, binding free energy
+\citetext{\citealp[
+]{cao_predicting_2012}; \citealp{helwak_mapping_2013}} and seed type
+\citep{werfel_preferential_2017} in miRNA:target interactions are
+described as factors effecting binding affinity. Also, binding region
+(e.g, 5'UTR, 3'UTR, CDS) was shown to drastically affect miRNA
+degradation efficiency
+\citep{hausser_analysis_2013, helwak_mapping_2013}. Since binding or
+degradation efficiency from literature have various and broad ranges,
+both binding and efficiency factors are normalized to their maximum
+values and scaled to {[}0,1{]} interval. The normalized values of
+factors are used to determine binding activity and miRNA efficiency on
+targets (Figure \ref{fig:fig2}). Binding affinities (activity, \(Eff\))
+of miRNAs on each individual gene are calculated as shown in equation
+\eqref{eq:3}; \(C_m\), miRNA expression in the group; \(C_{gi}\),
+individual gene expression; gi, individual gene (Figure
+\ref{fig:fig2}C).
 
@@
 
-Not all miRNA:target binding events result in degradation of target.
-The degradation of target by bound miRNA depends on efficiency factors such as binding region.
-Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:4} (Figure \ref{fig:fig2}D); $RE^\prime_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
-The cooperative repression activity of miRNAs to their common targets is figured out as shown in Figure \ref{fig:fig2}E.
+Not all miRNA:target binding events result in degradation of target. The
+degradation of target by bound miRNA depends on efficiency factors such
+as binding region. Exact repression efficiency of miRNA is calculated
+according to equation \eqref{eq:4} (Figure \ref{fig:fig2}D);
+\(RE^\prime_{gi}\), normalized values of region efficiency coefficient
+between miRNA and gene. The cooperative repression activity of miRNAs to
+their common targets is figured out as shown in Figure \ref{fig:fig2}E.
 
-## Breast cancer patient dataset
+\subsection{Breast cancer patient dataset}
 
-We have applied our model in a real dataset for which experimental measurements of various factors were available.
-Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network https://www.cancer.gov/tcga.
-High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015].
-We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from published datasets [@helwak_mapping_2013; @moore_mirnatarget_2015].
-Detailed description of network construction and its code is available in Supplementary Materials and Methods.
-SERPINE2, one of the most effective nodes, has been used as trigger in ceRNA network simulation in breast cancer patient network.
-Relevant steps and codes are available at  Supplementary Materials and Methods.
+We have applied our model in a real dataset for which experimental
+measurements of various factors were available. Expression levels of
+miRNA and genes in tumor and normal tissue of single patient are
+retrieved from TCGA Research Network \url{https://www.cancer.gov/tcga}.
+High-throughput experimental datasets which are provided miRNA:gene
+target pairs with interaction factors
+\citep{helwak_mapping_2013, moore_mirnatarget_2015}. We have combined
+miRNA and gene expression datasets via miRNA:target gene dataset
+retrieved from published datasets
+\citep{helwak_mapping_2013, moore_mirnatarget_2015}. Detailed
+description of network construction and its code is available in
+Supplementary Materials and Methods. SERPINE2, one of the most effective
+nodes, has been used as trigger in ceRNA network simulation in breast
+cancer patient network. Relevant steps and codes are available at
+Supplementary Materials and Methods.
 
-[#TODO]: <> "about formulas, 2) in formulas eq:1 and eq:2, C_m or Eff_gi is not intuitive, is there consensus terms about effect or nth gene?"
+\section{RESULTS}
 
-# RESULTS
+\subsection{Network with miRNA:ceRNA expression level}
 
-## Network with miRNA:ceRNA expression level
-We have developed a network-based approach to assess effects of expression level changes in competitive ceRNA regulation.
-The basic mode of miRNA repression activity has been based on miRNA and target abundance in various researches [@arvey_target_2010; @denzler_assessing_2014].
-Our approach can effortlessly calculate effects of expression changes when abundance levels of miRNAs and targets is only available factor.
-In sample network given in Figure (\ref{fig:fig1}), after an increase in expression level of a gene (G2), expression values of other genes also changed due to redistribution of miRNA among its targets.
-Previous studies have shown that if a gene abundance increases in ceRNA system, expression levels of genes targeted by shared miRNA are also affected [@lai_understanding_2016; @salmena_cerna_2011; @tay_multilayered_2014].
-It was observed that expression levels of primary neighborhoods which do not interact with another miRNA of the trigger gene change in relation to their expression.
-However, changes in expression levels of G4 and G1 differ although they have same initial conditions.
-This is due to fact that G4 is targeted by two miRNAs, M1 and M2 (Figure \ref{fig:fig1}D).
+We have developed a network-based approach to assess effects of
+expression level changes in competitive ceRNA regulation. The basic mode
+of miRNA repression activity has been based on miRNA and target
+abundance in various researches
+\citep{arvey_target_2010, denzler_assessing_2014}. Our approach can
+effortlessly calculate effects of expression changes when abundance
+levels of miRNAs and targets is only available factor. In sample network
+given in Figure (\ref{fig:fig1}), after an increase in expression level
+of a gene (G2), expression values of other genes also changed due to
+redistribution of miRNA among its targets. Previous studies have shown
+that if a gene abundance increases in ceRNA system, expression levels of
+genes targeted by shared miRNA are also affected
+\citep{lai_understanding_2016, salmena_cerna_2011, tay_multilayered_2014}.
+It was observed that expression levels of primary neighborhoods which do
+not interact with another miRNA of the trigger gene change in relation
+to their expression. However, changes in expression levels of G4 and G1
+differ although they have same initial conditions. This is due to fact
+that G4 is targeted by two miRNAs, M1 and M2 (Figure \ref{fig:fig1}D).
 
-Genes targeted by multiple miRNAs act as a trigger for adjacent local neighborhood of targeting miRNAs, causing changes in expression levels of genes outside the local neighborhood of original trigger gene.
-Therefore, primary expression level change in gene (G2) causes changes in other group of genes (G5 and G6) even though original trigger gene (G2) and genes in other group are not targeted by common miRNA.
-In addition, as shown by ceRNA hypothesis model of Ala et al., after the increase of gene expression level of G2, the miRNA that is found in the same group (M1) tended to be less repressive on its remaining targets (G1, G3 and G4).
-It's important to note that the changes in gene expression levels will have more pronounced effect if miRNA:target ratio is high, i.e., more miRNA available per target, which was reported in previous findings [@arvey_target_2010; @bosson_endogenous_2014; @denzler_assessing_2014].
+Genes targeted by multiple miRNAs act as a trigger for adjacent local
+neighborhood of targeting miRNAs, causing changes in expression levels
+of genes outside the local neighborhood of original trigger gene.
+Therefore, primary expression level change in gene (G2) causes changes
+in other group of genes (G5 and G6) even though original trigger gene
+(G2) and genes in other group are not targeted by common miRNA. In
+addition, as shown by ceRNA hypothesis model of Ala et al., after the
+increase of gene expression level of G2, the miRNA that is found in the
+same group (M1) tended to be less repressive on its remaining targets
+(G1, G3 and G4). It's important to note that the changes in gene
+expression levels will have more pronounced effect if miRNA:target ratio
+is high, i.e., more miRNA available per target, which was reported in
+previous findings
+\citep{arvey_target_2010, bosson_endogenous_2014, denzler_assessing_2014}.
 
@@
 
-## ceRNA:Target networks based on interaction factors
-Earlier studies reported that miRNA regulatory interactions are affected by different parameters.
-For example, Xu et al.
-have investigated the importance of seed pairing type between miRNAs and their targets and target site location by using proteomics dataset [@xu_characterization_2014].
-They have proposed that the features of binding between miRNA and target can be critical for miRNA efficiency.
-In addition, binding energy between miRNA and targets is a significant determinant for miRNA efficiency and it has been reported that strength of miRNA:target interactions is depended on binding energy of complexes [@breda_quantifying_2015].
-Similarly, another study revealed that affinity is correlated with seed pairing of miRNA:target pairs and suggested that affinity is correlated with number of canonical seed base pairing [@bosson_endogenous_2014].
-Therefore, we have integrated aforementioned interaction parameters that could be useful for predicting miRNA repression activity more accurately.
-The sample dataset used in Figure \ref{fig:fig1} is recalculated with additional factors Table \ref{tab:one} in effect with same trigger, two fold increase in Gene2 (Figure \ref{fig:fig3}).
-When the factors were taken into account in the system, miRNA efficiencies varied as shown in Figure \ref{fig:fig3}A.
-Although the miRNA:target expression ratios in steady-state were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed.
+\subsection{ceRNA:Target networks based on interaction factors}
 
-[#TODO]: <> "figure1, 2 ve 3 uzerinden farkli dagilim ve degradasyon açıklaması buraya. G2 uzerinden olmasa iyi olur"
+Earlier studies reported that miRNA regulatory interactions are affected
+by different parameters. For example, Xu et al.~have investigated the
+importance of seed pairing type between miRNAs and their targets and
+target site location by using proteomics dataset
+\citep{xu_characterization_2014}. They have proposed that the features
+of binding between miRNA and target can be critical for miRNA
+efficiency. In addition, binding energy between miRNA and targets is a
+significant determinant for miRNA efficiency and it has been reported
+that strength of miRNA:target interactions is depended on binding energy
+of complexes \citep{breda_quantifying_2015}. Similarly, another study
+revealed that affinity is correlated with seed pairing of miRNA:target
+pairs and suggested that affinity is correlated with number of canonical
+seed base pairing \citep{bosson_endogenous_2014}. Therefore, we have
+integrated aforementioned interaction parameters that could be useful
+for predicting miRNA repression activity more accurately. The sample
+dataset used in Figure \ref{fig:fig1} is recalculated with additional
+factors Table \ref{tab:one} in effect with same trigger, two fold
+increase in Gene2 (Figure \ref{fig:fig3}). When the factors were taken
+into account in the system, miRNA efficiencies varied as shown in Figure
+\ref{fig:fig3}A. Although the miRNA:target expression ratios in
+steady-state were same in comparison with the sample dataset without
+factors, efficiency of binding and repression have changed.
 
-When the model run based on expression values of miRNAs and ceRNAs (genes), proportional distribution of miRNA on targets is determined in accordance with equations \eqref{eq:1} \eqref{eq:2}.
-In that case, the system is observed as shown in Figure \ref{fig:fig1}A at steady-state.
-However, if interaction factors are taken into account in the approach, specified interaction factors (i.e, binding factors such as seed type and energy) affect distribution of miRNA on targets as calculated in equation \eqref{eq:3} (shown in Figure \ref{fig:fig2}C).
-After distributed miRNA efficiency was re-calculated with degradation factor(s) (i.e region effect) according to equation \eqref{eq:4}, steady-state condition of network is obtained (shown in Figure \ref{fig:fig2}D and Figure \ref{fig:fig3}A).
-For instance, proportional distribution of G1:M1 interaction in Figure \ref{fig:fig1}A is calculated lower than the same pair (G1:M1) in the other approach that includes interaction factors (Figure \ref{fig:fig2}C).
-But, a decrease is observed in repression efficiency of M1 miRNA on G1 in interaction factor based network approach (Figure \ref{fig:fig2}D and \ref{fig:fig3}A).
+When the model run based on expression values of miRNAs and ceRNAs
+(genes), proportional distribution of miRNA on targets is determined in
+accordance with equations \eqref{eq:1} \eqref{eq:2}. In that case, the
+system is observed as shown in Figure \ref{fig:fig1}A at steady-state.
+However, if interaction factors are taken into account in the approach,
+specified interaction factors (i.e, binding factors such as seed type
+and energy) affect distribution of miRNA on targets as calculated in
+equation \eqref{eq:3} (shown in Figure \ref{fig:fig2}C). After
+distributed miRNA efficiency was re-calculated with degradation
+factor(s) (i.e region effect) according to equation \eqref{eq:4},
+steady-state condition of network is obtained (shown in Figure
+\ref{fig:fig2}D and Figure \ref{fig:fig3}A). For instance, proportional
+distribution of G1:M1 interaction in Figure \ref{fig:fig1}A is
+calculated lower than the same pair (G1:M1) in the other approach that
+includes interaction factors (Figure \ref{fig:fig2}C). But, a decrease
+is observed in repression efficiency of M1 miRNA on G1 in interaction
+factor based network approach (Figure \ref{fig:fig2}D and
+\ref{fig:fig3}A).
 
-It is considered that entire miRNAs in the system affect targets according to target:total target ratio Figure \ref{fig:fig1}A, when interaction factors did not take into account.
-However, in presence of interaction factors in miRNA:target interaction network (shown in Table \ref{tab:one}), miRNAs are distributed according to target:total target ratio and affinity factors in first like shown in Figure \ref{fig:fig2}C.
-After affinity mediated proportional distribution of miRNA expression, degradation factor is considered to specify count of repressive miRNA in pairs because entire bound miRNA target pairs might not be resulted with degradation (It is figured out shown in Figure \ref{fig:fig2}D).
-For this reason, miRNA1 (M1) has more weak repressive effect (shown as edge variables) on target Gene2 Figure \ref{fig:fig3}A in comparison with miRNA1 (M1):Gene2 (G2) interaction in Figure \ref{fig:fig1}A.
-On the other hand, it is possible that taking into account of these interaction factors could cause increasing miRNA repression activity like in miRNA2:Gene4 interaction Figure \ref{fig:fig3}A.
-When expression of Gene2 (G2) increased, expression values of all genes also changed differentially because of contribution of efficiency factors (Figure \ref{fig:fig3} B,C,D).
-We have considered that in our approach energy and seed type of pairs is significant for binding and targeted region is important for repression.
+It is considered that entire miRNAs in the system affect targets
+according to target:total target ratio Figure \ref{fig:fig1}A, when
+interaction factors did not take into account. However, in presence of
+interaction factors in miRNA:target interaction network (shown in Table
+\ref{tab:one}), miRNAs are distributed according to target:total target
+ratio and affinity factors in first like shown in Figure
+\ref{fig:fig2}C. After affinity mediated proportional distribution of
+miRNA expression, degradation factor is considered to specify count of
+repressive miRNA in pairs because entire bound miRNA target pairs might
+not be resulted with degradation (It is figured out shown in Figure
+\ref{fig:fig2}D). For this reason, miRNA1 (M1) has more weak repressive
+effect (shown as edge variables) on target Gene2 Figure \ref{fig:fig3}A
+in comparison with miRNA1 (M1):Gene2 (G2) interaction in Figure
+\ref{fig:fig1}A. On the other hand, it is possible that taking into
+account of these interaction factors could cause increasing miRNA
+repression activity like in miRNA2:Gene4 interaction Figure
+\ref{fig:fig3}A. When expression of Gene2 (G2) increased, expression
+values of all genes also changed differentially because of contribution
+of efficiency factors (Figure \ref{fig:fig3} B,C,D). We have considered
+that in our approach energy and seed type of pairs is significant for
+binding and targeted region is important for repression.
 
@@
 
-In regulation of miRNA target sample system (Figure \ref{fig:fig3}), miRNA (M1) repressive efficiency on the primary triggering gene (G2) is low in steady-state.
-So it has been observed that the regulatory activities of miRNA on the targets are weak after the increase of miRNA target G2.
-When model is triggered with two fold increase in expression level of common target (G4), more prominent changes were observed in gene expression levels in network compared to changes caused by G2 increase.
-Furthermore, change in expression level of target gene that has strong miRNA repression efficiency resulted in evident perturbation in network.
-On the other hand, it was observed that Gene2 was weakly affected from change of Gene4 expression because of its weak interaction factors.
-Since perturbation efficiency of each gene is different, we have developed a function which screens each gene in the network for their perturbation efficiencies.
-*(see details in Supplementary Materials and Methods*.
-On the other hand, when we applied the method on the minimal dataset, Gene4 has been found to be the most efficient element in terms of number of perturbed elements and miRNA2 (M2) has been found to be causing the highest mean in expression changes.
+In regulation of miRNA target sample system (Figure \ref{fig:fig3}),
+miRNA (M1) repressive efficiency on the primary triggering gene (G2) is
+low in steady-state. So it has been observed that the regulatory
+activities of miRNA on the targets are weak after the increase of miRNA
+target G2. When model is triggered with two fold increase in expression
+level of common target (G4), more prominent changes were observed in
+gene expression levels in network compared to changes caused by G2
+increase. Furthermore, change in expression level of target gene that
+has strong miRNA repression efficiency resulted in evident perturbation
+in network. On the other hand, it was observed that Gene2 was weakly
+affected from change of Gene4 expression because of its weak interaction
+factors. Since perturbation efficiency of each gene is different, we
+have developed a function which screens each gene in the network for
+their perturbation efficiencies. \emph{(see details in Supplementary
+Materials and Methods}. On the other hand, when we applied the method on
+the minimal dataset, Gene4 has been found to be the most efficient
+element in terms of number of perturbed elements and miRNA2 (M2) has
+been found to be causing the highest mean in expression changes.
 
-## Simulation of large and realistic network
+\subsection{Simulation of large and realistic network}
 
-Apart from tools predicting miRNA:target pairs, there are databases that provide  experimentally validated miRNA:target interaction datasets.
-The reliability in these datasets range from weak (e.g, western blot) to strong (e.g, CLiP, CLASH).
-High-throughput methods ensure handling of AGO (Argounate Protein) interacted miRNAs (CLiP) and their targets as well (CLEAR-CLiP, CLASH).
-We constructed a large ceRNA network utilizing high-throughput datasets containing chimeric reading of miRNA:target pairs [@helwak_mapping_2013; @moore_mirnatarget_2015] which expose exact complementary seed regions.
+Apart from tools predicting miRNA:target pairs, there are databases that
+provide experimentally validated miRNA:target interaction datasets. The
+reliability in these datasets range from weak (e.g, western blot) to
+strong (e.g, CLiP, CLASH). High-throughput methods ensure handling of
+AGO (Argounate Protein) interacted miRNAs (CLiP) and their targets as
+well (CLEAR-CLiP, CLASH). We constructed a large ceRNA network utilizing
+high-throughput datasets containing chimeric reading of miRNA:target
+pairs \citep{helwak_mapping_2013, moore_mirnatarget_2015} which expose
+exact complementary seed regions.
 
-Although these experimental sources include exact binding sites, they do not imply that each miRNA binding event is functional [@liu2019prediction].
-Seed type, energy, binding region affects functionality of miRNA drastically.
-In order to construct realistic network, we incorporated such factors in the network.
-Energy  values in miRNA:target pairs represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015], seed type [@grimson_microrna_2007; @bartel_micrornas:_2009; @betel2010comprehensive] and binding region effect [@hausser_analysis_2013; @helwak_mapping_2013] data were collected and integrated  into network.
+Although these experimental sources include exact binding sites, they do
+not imply that each miRNA binding event is functional
+\citep{liu2019prediction}. Seed type, energy, binding region affects
+functionality of miRNA drastically. In order to construct realistic
+network, we incorporated such factors in the network. Energy values in
+miRNA:target pairs represented by high-throughput studies
+\citep{helwak_mapping_2013, moore_mirnatarget_2015}, seed type
+\citep{grimson_microrna_2007, bartel_micrornas:_2009, betel2010comprehensive}
+and binding region effect
+\citep{hausser_analysis_2013, helwak_mapping_2013} data were collected
+and integrated into network.
 
-The network was finalized by integrating gene and miRNA expression levels from TCGA determined by RNA-Seq.
-Our  model can successfully simulate perturbations in such large network despite complex behaviors and struggle to reach steady-state *(see details in Supplementary Materials and Methods)*.
-Simulations show that change in expression level of single gene has potential to affect whole network, perturbing almost all nodes.
-These observations are in accordance with competing endogenous RNA hypothesis where genes targeted with many common miRNAs subsequently transmit perturbation to neighboring groups.
+The network was finalized by integrating gene and miRNA expression
+levels from TCGA determined by RNA-Seq. Our model can successfully
+simulate perturbations in such large network despite complex behaviors
+and struggle to reach steady-state \emph{(see details in Supplementary
+Materials and Methods)}. Simulations show that change in expression
+level of single gene has potential to affect whole network, perturbing
+almost all nodes. These observations are in accordance with competing
+endogenous RNA hypothesis where genes targeted with many common miRNAs
+subsequently transmit perturbation to neighboring groups.
 
-We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach.
-Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues.
-So, it might not said that a single gene is responsible for entire regulations in cancer cells.
-Besides, the all other factors such as up/down regulation of miRNAs or other genes were ignored.
-Therefore, it would be more useful to test our approach with an ideal dataset that includes entire informations of gene expression values before and after the regulation of a known gene at existing tissue, and miRNA expression values of this tissue at initial conditions.
+We have observed that the results of simulation with SERPINE2 gene, a
+serine protease inhibitor, one of the overexpressed genes of breast
+cancer \citep{yang2018expression, candia2006protease} is not convenient
+with tumor tissue expression values, when we compared the tumor gene
+expression dataset retrieved from TCGA and simulation results obtained
+via our approach. Simulation results were not expected to be consistent
+with tumor tissue expression values because large number of abnormal
+regulations were observed in these tissues. So, it might not said that a
+single gene is responsible for entire regulations in cancer cells.
+Besides, the all other factors such as up/down regulation of miRNAs or
+other genes were ignored. Therefore, it would be more useful to test our
+approach with an ideal dataset that includes entire informations of gene
+expression values before and after the regulation of a known gene at
+existing tissue, and miRNA expression values of this tissue at initial
+conditions.
 
-# DISCUSSION
+\section{DISCUSSION}
 
-Network based approaches for analyzing miRNA:target interactions have been developed in earlier studies.
-An initial  attempt demonstrated the ceRNA crosstalk in a network-like minimal interaction structure with concentrations of ceRNA and miRNAs [@figliuzzi_micrornas_2013].
-Next, a network based kinetic model integrating miRNA and target rates of transcription, degradation, binding and unbinding was developed [@nitzan_interactions_2014] using high throughput experimental dataset about miR-92a depletion [@helwak_mapping_2013].
-It was demonstrated that distant ceRNAs can interact with each other via indirect links, and the interactions are affected depending on distance between ceRNAs or topological features of network [@nitzan_interactions_2014].
-More recently, an approach to detect ceRNA interaction by using the miRNA expression, gene expression and common miRNAs between gene targets was developed [@markus_list_sponge_2017] which was effective in analyzing genes through miRNAs.
-Based on the observations that a miRNA can exhibit strong functionality to a target but may not against an other, the authors have concluded that existing miRNA based approach may not be suitable for understanding regulations of ceRNA interactions.
-Additionally, miRmapper package [@da2018mirmapper] utilizes an adjacency matrix to associate miRNAs using differentially expressed genes and identifies significant miRNAs in network.
+Network based approaches for analyzing miRNA:target interactions have
+been developed in earlier studies. An initial attempt demonstrated the
+ceRNA crosstalk in a network-like minimal interaction structure with
+concentrations of ceRNA and miRNAs \citep{figliuzzi_micrornas_2013}.
+Next, a network based kinetic model integrating miRNA and target rates
+of transcription, degradation, binding and unbinding was developed
+\citep{nitzan_interactions_2014} using high throughput experimental
+dataset about miR-92a depletion \citep{helwak_mapping_2013}. It was
+demonstrated that distant ceRNAs can interact with each other via
+indirect links, and the interactions are affected depending on distance
+between ceRNAs or topological features of network
+\citep{nitzan_interactions_2014}. More recently, an approach to detect
+ceRNA interaction by using the miRNA expression, gene expression and
+common miRNAs between gene targets was developed
+\citep{markus_list_sponge_2017} which was effective in analyzing genes
+through miRNAs. Based on the observations that a miRNA can exhibit
+strong functionality to a target but may not against an other, the
+authors have concluded that existing miRNA based approach may not be
+suitable for understanding regulations of ceRNA interactions.
+Additionally, miRmapper package \citep{da2018mirmapper} utilizes an
+adjacency matrix to associate miRNAs using differentially expressed
+genes and identifies significant miRNAs in network.
 
-In our approach, we have not taken into account transcription, degradation or binding rates of elements in network.
-Because, although it is known as the miRNAs are highly stable, the transcription and degradation rates of miRNAs change depending on cellular conditions [@ruegger_microrna_2012].
-However, due to lack of datasets including degradation and transcription rates of miRNAs at specific cellular conditions, such factors are neglected in our model.
-Additionally, other regulation parameters such as gene-gene interactions and transcription factors are ignored but the network structure is flexible and can integrate additional regulation elements as they become available.
+In our approach, we have not taken into account transcription,
+degradation or binding rates of elements in network. Because, although
+it is known as the miRNAs are highly stable, the transcription and
+degradation rates of miRNAs change depending on cellular conditions
+\citep{ruegger_microrna_2012}. However, due to lack of datasets
+including degradation and transcription rates of miRNAs at specific
+cellular conditions, such factors are neglected in our model.
+Additionally, other regulation parameters such as gene-gene interactions
+and transcription factors are ignored but the network structure is
+flexible and can integrate additional regulation elements as they become
+available.
 
-# CONCLUSION
+\section{CONCLUSION}
 
-The main advantage of ceRNAnetsim package is providing various functions which can be run in accordance with user data.
-For example, the functions can work without interaction factors or with a simple interaction factor that is specified by user.
-Furthermore, it can also be operated according to a different parameter(s) provided by the user.In this study, we have preferred to run model, shown in *Multifactorial calculations in miRNA:target network* section.
+The main advantage of ceRNAnetsim package is providing various functions
+which can be run in accordance with user data. For example, the
+functions can work without interaction factors or with a simple
+interaction factor that is specified by user. Furthermore, it can also
+be operated according to a different parameter(s) provided by the
+user.In this study, we have preferred to run model, shown in
+\emph{Multifactorial calculations in miRNA:target network} section.
 
-In the future, with developments in experimental techniques about miRNAs:target interaction dynamics, more accurate results can be obtained from our model.
-Also, our model can integrate emerging miRNA effectors such as circRNAs.
-Consequently, our package may lead to better understanding and predictions of abnormal regulations and pathways underlying diseases or conditions.
+In the future, with developments in experimental techniques about
+miRNAs:target interaction dynamics, more accurate results can be
+obtained from our model. Also, our model can integrate emerging miRNA
+effectors such as circRNAs. Consequently, our package may lead to better
+understanding and predictions of abnormal regulations and pathways
+underlying diseases or conditions.
 
-# ACKNOWLEDGEMENTS
+\section{ACKNOWLEDGEMENTS}
 
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text.
+Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
+Text. Text. Text.
 
-### Conflict of interest statement.
+\subsubsection{Conflict of interest statement.}
 
@@
 \newpage
+
+\bibliographystyle{nar}
+\bibliography{NAR-sample.bib}
+
+\end{document}

### 2019-09-28 Supplementary_mat_met.Rmd
article NAR form

@@
 ```{r}
-data("minsamp")
+minsamp<-readRDS("data/minsamp.RDS")
 
@@
 ```{r}
-#saveRDS(E9GE_mirnagenenormal, "data/E9GE_mirnagenenormal.rda")
-
-E9GE_mirnagenenormal<-readRDS("data/E9GE_mirnagenenormal.rda")
+E9GE_mirnagenenormal<-readRDS("data/E9GE_mirnagenenormal.RDS")
 head(E9GE_mirnagenenormal)
@@
 ```{r}
-#saveRDS(E9GE_mirnagenetumor, "data/E9GE_mirnagenetumor.rda")
+#saveRDS(E9GE_mirnagenetumor, "data/E9GE_mirnagenetumor.RDS")
 
-E9GE_mirnagenetumor<-readRDS("data/E9GE_mirnagenetumor.rda")
+E9GE_mirnagenetumor<-readRDS("data/E9GE_mirnagenetumor.RDS")
 head(E9GE_mirnagenetumor)

### 2019-09-29 NAR_form_article.Rmd
scanning

@@
 At each line minimum required information is expression levels of miRNA and the target.
-If available, additional data about factors effecting binding or efficiency of miRNA can be provided as separate columns.
+If available, additional data about factors effecting binding or efficiency of miRNA can be provided as separate columns (see Figure S1 in Supplementary Tables).
 After construction of the network, amount of miRNA per target is calculated and kept as edge data.
@@
 In this case, the common target acts as trigger for the other group (the targets interacted with M2 miRNA) in the network.
-Due to common targeted elements, the change in one neighborhood spreads to other neighborhoods (Figure \ref{fig:fig1}D), consequently have potential to affect whole network due to "ripple effect".
+Due to common targeted elements, the change in one neighborhood spreads to other neighborhoods (Figure \ref{fig:fig1}D), consequently have potential to affect whole network due to "ripple effect" (see Section 2 in Supplementary Materials and Methods).
 
@@
 Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:4} (Figure \ref{fig:fig2}D); $RE^\prime_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
-The cooperative repression activity of miRNAs to their common targets is figured out as shown in Figure \ref{fig:fig2}E.
+The cooperative repression activity of miRNAs to their common targets is figured out as shown in Figure \ref{fig:fig2}E (see Section 3 in Supplementary Materials and Methods).
 
@@
 Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network https://www.cancer.gov/tcga.
-High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015].
+High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015] (Detailed codes are available at Section 6 in Supplementary Materials and Methods).
 We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from published datasets [@helwak_mapping_2013; @moore_mirnatarget_2015].
@@
 It was observed that expression levels of primary neighborhoods which do not interact with another miRNA of the trigger gene change in relation to their expression.
-However, changes in expression levels of G4 and G1 differ although they have same initial conditions.
-This is due to fact that G4 is targeted by two miRNAs, M1 and M2 (Figure \ref{fig:fig1}D).
 
@@
 Although the miRNA:target expression ratios in steady-state were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed.
+On the other hand, changes in expression levels of G4 and G1 differ although they have same initial conditions (Detailed codes are available at Sections 3-4 in Supplementary Materials and Methods).
+This is due to fact that G4 is targeted by two miRNAs, M1 and M2 (Figures \ref{fig:fig3}D) and changed interaction factors.
 
@@
 
-It is considered that entire miRNAs in the system affect targets according to target:total target ratio Figure \ref{fig:fig1}A, when interaction factors did not take into account.
+It is considered that entire miRNAs in the system affect targets according to target:total target ratio Figure \ref{fig:fig1}A, when interaction factors were not taken into account.
 However, in presence of interaction factors in miRNA:target interaction network (shown in Table \ref{tab:one}), miRNAs are distributed according to target:total target ratio and affinity factors in first like shown in Figure \ref{fig:fig2}C.
@@
 On the other hand, it was observed that Gene2 was weakly affected from change of Gene4 expression because of its weak interaction factors.
-Since perturbation efficiency of each gene is different, we have developed a function which screens each gene in the network for their perturbation efficiencies.
-*(see details in Supplementary Materials and Methods*.
-On the other hand, when we applied the method on the minimal dataset, Gene4 has been found to be the most efficient element in terms of number of perturbed elements and miRNA2 (M2) has been found to be causing the highest mean in expression changes.
+Since perturbation efficiency of each gene is different, we have developed a function which screens each gene in the network for their perturbation efficiencies (Detailed codes are available at Section 4 in Supplementary Materials and Methods).
+On the other hand, when we applied the method on the minimal dataset (i.e.
+Gene4 is trigger element), Gene4 has been found to be the most efficient element in terms of number of perturbed elements and miRNA2 (M2) has been found to be causing the highest mean in expression changes.
 
@@
 The network was finalized by integrating gene and miRNA expression levels from TCGA determined by RNA-Seq.
-Our  model can successfully simulate perturbations in such large network despite complex behaviors and struggle to reach steady-state *(see details in Supplementary Materials and Methods)*.
+Our  model can successfully simulate perturbations in such large network despite complex behaviors and struggle to reach steady-state (see details at Section 6 in Supplementary Materials and Methods).
 Simulations show that change in expression level of single gene has potential to affect whole network, perturbing almost all nodes.
@@
 
-We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach.
+We have analysed the realistic network in context of perturbation efficiencies of nodes.
+This process were run with 3 fold upregulation and 10 iteration in conditions where elements that have less than five percent change are not taken into account (Codes are found at Section 6.5 in Supplementary Materials and methods).
+With this approach, SERPINE is found as one of the most effective gene nodes in network.
+We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach (see Section 6.6 in Supplementary Materials and methods).
 Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues.

### 2019-09-29 NAR_form_article.tex
scanning

@@
 about factors effecting binding or efficiency of miRNA can be provided
-as separate columns. After construction of the network, amount of miRNA
-per target is calculated and kept as edge data. Simply, a target will
-sequester miRNA proportional to its ratio amount among other targets. If
-additional criteria effecting the binding of miRNA to its target is
-provided, distribution of miRNA will be calculated accordingly. Target
-can be mRNA or any other ceRNA (circRNA, ncRNA, etc.) thus, throughout
-the manuscript terms target, gene and ceRNA are used interchangeably.
+as separate columns (see Figure S1 in Supplementary Tables). After
+construction of the network, amount of miRNA per target is calculated
+and kept as edge data. Simply, a target will sequester miRNA
+proportional to its ratio amount among other targets. If additional
+criteria effecting the binding of miRNA to its target is provided,
+distribution of miRNA will be calculated accordingly. Target can be mRNA
+or any other ceRNA (circRNA, ncRNA, etc.) thus, throughout the
+manuscript terms target, gene and ceRNA are used interchangeably.
 
@@
 \ref{fig:fig1}D), consequently have potential to affect whole network
-due to ``ripple effect''.
+due to ``ripple effect'' (see Section 2 in Supplementary Materials and
+Methods).
 
@@
 between miRNA and gene. The cooperative repression activity of miRNAs to
-their common targets is figured out as shown in Figure \ref{fig:fig2}E.
+their common targets is figured out as shown in Figure \ref{fig:fig2}E
+(see Section 3 in Supplementary Materials and Methods).
 
@@
 target pairs with interaction factors
-\citep{helwak_mapping_2013, moore_mirnatarget_2015}. We have combined
-miRNA and gene expression datasets via miRNA:target gene dataset
-retrieved from published datasets
+\citep{helwak_mapping_2013, moore_mirnatarget_2015} (Detailed codes are
+available at Section 6 in Supplementary Materials and Methods). We have
+combined miRNA and gene expression datasets via miRNA:target gene
+dataset retrieved from published datasets
 \citep{helwak_mapping_2013, moore_mirnatarget_2015}. Detailed
@@
 not interact with another miRNA of the trigger gene change in relation
-to their expression. However, changes in expression levels of G4 and G1
-differ although they have same initial conditions. This is due to fact
-that G4 is targeted by two miRNAs, M1 and M2 (Figure \ref{fig:fig1}D).
+to their expression.
 
@@
 steady-state were same in comparison with the sample dataset without
-factors, efficiency of binding and repression have changed.
+factors, efficiency of binding and repression have changed. On the other
+hand, changes in expression levels of G4 and G1 differ although they
+have same initial conditions (Detailed codes are available at Sections
+3-4 in Supplementary Materials and Methods). This is due to fact that G4
+is targeted by two miRNAs, M1 and M2 (Figures \ref{fig:fig3}D) and
+changed interaction factors.
 
@@
 according to target:total target ratio Figure \ref{fig:fig1}A, when
-interaction factors did not take into account. However, in presence of
+interaction factors were not taken into account. However, in presence of
 interaction factors in miRNA:target interaction network (shown in Table
@@
 have developed a function which screens each gene in the network for
-their perturbation efficiencies. \emph{(see details in Supplementary
-Materials and Methods}. On the other hand, when we applied the method on
-the minimal dataset, Gene4 has been found to be the most efficient
-element in terms of number of perturbed elements and miRNA2 (M2) has
-been found to be causing the highest mean in expression changes.
+their perturbation efficiencies (Detailed codes are available at Section
+4 in Supplementary Materials and Methods). On the other hand, when we
+applied the method on the minimal dataset (i.e.~Gene4 is trigger
+element), Gene4 has been found to be the most efficient element in terms
+of number of perturbed elements and miRNA2 (M2) has been found to be
+causing the highest mean in expression changes.
 
@@
 simulate perturbations in such large network despite complex behaviors
-and struggle to reach steady-state \emph{(see details in Supplementary
-Materials and Methods)}. Simulations show that change in expression
-level of single gene has potential to affect whole network, perturbing
-almost all nodes. These observations are in accordance with competing
-endogenous RNA hypothesis where genes targeted with many common miRNAs
-subsequently transmit perturbation to neighboring groups.
+and struggle to reach steady-state (see details at Section 6 in
+Supplementary Materials and Methods). Simulations show that change in
+expression level of single gene has potential to affect whole network,
+perturbing almost all nodes. These observations are in accordance with
+competing endogenous RNA hypothesis where genes targeted with many
+common miRNAs subsequently transmit perturbation to neighboring groups.
 
-We have observed that the results of simulation with SERPINE2 gene, a
-serine protease inhibitor, one of the overexpressed genes of breast
-cancer \citep{yang2018expression, candia2006protease} is not convenient
-with tumor tissue expression values, when we compared the tumor gene
+We have analysed the realistic network in context of perturbation
+efficiencies of nodes. This process were run with 3 fold upregulation
+and 10 iteration in conditions where elements that have less than five
+percent change are not taken into account (Codes are found at Section
+6.5 in Supplementary Materials and methods). With this approach, SERPINE
+is found as one of the most effective gene nodes in network. We have
+observed that the results of simulation with SERPINE2 gene, a serine
+protease inhibitor, one of the overexpressed genes of breast cancer
+\citep{yang2018expression, candia2006protease} is not convenient with
+tumor tissue expression values, when we compared the tumor gene
 expression dataset retrieved from TCGA and simulation results obtained
-via our approach. Simulation results were not expected to be consistent
-with tumor tissue expression values because large number of abnormal
+via our approach (see Section 6.6 in Supplementary Materials and
+methods). Simulation results were not expected to be consistent with
+tumor tissue expression values because large number of abnormal
 regulations were observed in these tissues. So, it might not said that a

### 2019-09-29 Supplementary_mat_met.Rmd
scanning

@@
 
-## *minsamp* dataset analysis in lack of interaction factors.
+# 2. *minsamp* dataset analysis in lack of interaction factors.
 
@@
 
-## *minsamp* dataset analysis with interaction factors.
+# 3. *minsamp* dataset analysis with interaction factors.
 
@@
 
-## Common target perturbation in *minsamp* dataset.
+# 4. Common target perturbation in *minsamp* dataset.
 
@@
 In our small dataset, minsamp, Gene4 is regulated by two miRNAs.
-Let's simulate perturbation effects triggered by Gene4 (Shown at Figure S8 in Supplementary Figures .
+Let's simulate perturbation effects triggered by Gene4 (Shown at Figure S8 in Supplementary Figures) .
 
@@
 
-## Determination of perturbation efficiencies of elements in system.
+# 5. Determination of perturbation efficiencies of elements in system.
 
@@
 
-# 2. Obtaining breast cancer dataset and integration
+# 6. Obtaining breast cancer dataset and integration
 
@@
 
-## 2.1 How to get gene expression counts of TCGA-E9-A1N5 patient.
+## 6.1 How to get gene expression counts of TCGA-E9-A1N5 patient.
 
@@
 
-## 2.2 How to get miRNA expression counts of TCGA-E9-A1N5 patient.
+## 6.2 How to get miRNA expression counts of TCGA-E9-A1N5 patient.
 
@@
 
-## 2.3 Get the high-throughput experimental miRNA:target dataset.
+## 6.3 Get the high-throughput experimental miRNA:target dataset.
 
@@
 + We committed the interaction factors as numeric values according to previous studies.
-(We added the interaction factors and their numeric values at Supplementary files, Process on miRNA:target pairs dataset for process codes and Supplementary Tables for used numeric values)
+(We added the interaction factors and their numeric values.
+The codes about integration of high-throughput experimental studies are available at Supplementary additional file and numeric values of interaction factors can be found at Table S2-3 in Supplementary Tables file.)
 
@@
 
-## 2.4 Combine the dataset
+## 6.4 Combine the dataset
 
@@
 
-## 2.5 Selection of trigger node
+## 6.5 Selection of trigger node
 
@@
 
-## 2.5 Approach of Method into Combined Datasets
+## 6.6 Approach of Method into Combined Datasets
 
@@
 
-### 2.5.1 Find iteration of simulation
+### 6.6.1 Find iteration of simulation
 
@@
 
-### 2.5.2 Simulation of dataset
+### 6.6.2 Simulation of dataset
 
@@
 
-### 2.5.3 Comparison of simulation results and tumor tissue expression values
+### 6.6.3 Comparison of simulation results and tumor tissue expression values
 ```{r}

### 2019-09-30 NAR_form_article.Rmd
Figure defining

@@
 At each line minimum required information is expression levels of miRNA and the target.
-If available, additional data about factors effecting binding or efficiency of miRNA can be provided as separate columns (see Figure S1 in Supplementary Tables).
+If available, additional data about factors effecting binding or efficiency of miRNA can be provided as separate columns (see Table S1 in Supplementary Tables).
 After construction of the network, amount of miRNA per target is calculated and kept as edge data.
 Simply, a target will sequester miRNA proportional to its ratio amount among other targets.
-If additional criteria effecting the binding of miRNA to its target is provided, distribution of miRNA will be calculated accordingly.
+If additional criteria effecting the binding of miRNA to its target is provided, distribution of miRNA will be calculated accordingly (available at Table S6 in Supplementary Tables).
 Target can be mRNA or any other ceRNA (circRNA, ncRNA, etc.) thus, throughout the manuscript terms target, gene and ceRNA are used interchangeably.
@@
 ## Triggering perturbation and subsequent calculations
-Initially, the network is assumed in steady-state (Figure \ref{fig:fig1}A) condition and needs least one trigger for initiating calculations.
-The trigger can be a change in expression level of one or more genes (Figure \ref{fig:fig1}B).
-After a trigger, the network undergoes iterative cycle of calculations at each of which distribution of miRNA in local neighborhood is recalculated (Figure \ref{fig:fig1}C).
+Initially, the network is assumed in steady-state (Figure \ref{fig:fig1}A and Figure S2) condition and needs least one trigger for initiating calculations.
+The trigger can be a change in expression level of one or more genes (Figure \ref{fig:fig1}B and Figure S3).
+After a trigger, the network undergoes iterative cycle of calculations at each of which distribution of miRNA in local neighborhood is recalculated (Figure \ref{fig:fig1}C and Figure S4A).
 Based on new miRNA distribution, expression level of each node (i.e.
@@
 In this case, the common target acts as trigger for the other group (the targets interacted with M2 miRNA) in the network.
-Due to common targeted elements, the change in one neighborhood spreads to other neighborhoods (Figure \ref{fig:fig1}D), consequently have potential to affect whole network due to "ripple effect" (see Section 2 in Supplementary Materials and Methods).
+Due to common targeted elements, the change in one neighborhood spreads to other neighborhoods (Figure \ref{fig:fig1}D and Figure S4B), consequently have potential to affect whole network due to "ripple effect" (see Section 2 in Supplementary Materials and Methods).
 
@@
 \caption{Schematic presentation of mechanism of network based model. 
-    \textbf{A)} In steady state, miRNAs (M: triangles) repress targets (G: circles) according to proportion of target expression level. 
-    \textbf{B)} Two fold increase in transcript level of Gene2 (G2) acts as a trigger (shown in red). 
-    \textbf{C)} Distribution of miRNA1 (M1) changes. 
-    \textbf{D)} The change at expression of common target affects changes of proportional distribution of miRNA2 (M2).
+    (\textbf{A}) In steady state, miRNAs (M: triangles) repress targets (G: circles) according to proportion of target expression level. 
+    (\textbf{B}) Two fold increase in transcript level of Gene2 (G2) acts as a trigger (shown in red). 
+    (\textbf{C}) Distribution of miRNA1 (M1) changes. 
+    (\textbf{D)}) The change at expression of common target affects changes of proportional distribution of miRNA2 (M2).
 Expression values are rounded to integers for simplicity.
@@
 In other words, binding factors exert their influence before or during binding, efficiency factors exert their influence after binding.
-In the literature, binding free energy [@cao_predicting_2012 ; @helwak_mapping_2013] and seed type [@werfel_preferential_2017] in miRNA:target interactions are described as factors effecting binding affinity.
+In the literature, binding free energy [@cao_predicting_2012; @helwak_mapping_2013] and seed type [@werfel_preferential_2017] in miRNA:target interactions are described as factors effecting binding affinity.
 Also, binding region (e.g, 5'UTR, 3'UTR, CDS) was shown to drastically affect miRNA degradation efficiency [@hausser_analysis_2013; @helwak_mapping_2013].
@@
 Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:4} (Figure \ref{fig:fig2}D); $RE^\prime_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
-The cooperative repression activity of miRNAs to their common targets is figured out as shown in Figure \ref{fig:fig2}E (see Section 3 in Supplementary Materials and Methods).
+The cooperative repression activity of miRNAs to their common targets is figured out as shown in Figure \ref{fig:fig2}E.
 
@@
 High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015] (Detailed codes are available at Section 6 in Supplementary Materials and Methods).
-We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from published datasets [@helwak_mapping_2013; @moore_mirnatarget_2015].
-Detailed description of network construction and its code is available in Supplementary Materials and Methods.
+We have combined miRNA and gene expression datasets via miRNA:target gene dataset retrieved from published datasets [@helwak_mapping_2013; @moore_mirnatarget_2015] (Section 6 in Supplementary Materials and Methods and Table S4 in Supplementary Tables).
 SERPINE2, one of the most effective nodes, has been used as trigger in ceRNA network simulation in breast cancer patient network.
@@
 \caption{Target regulations with interaction parameters. 
-     \textbf{A)} In the steady-state the repression activity of miRNAs on the targets after binding and repression efficiency. 
-     \textbf{B)} The changes the repression activities after increasing of G2 expression. 
-     \textbf{C)} Perturbation of primary neighborhoods of M1 miRNA (M1 miRNA group). 
-     \textbf{D)} Regulation of gene expression of other gene group via triggering target (common target between M1 and M2)}
+     (\textbf{A}) In the steady-state the repression activity of miRNAs on the targets after binding and repression efficiency. 
+     (\textbf{B}) The changes the repression activities after increasing of G2 expression. 
+     (\textbf{C}) Perturbation of primary neighborhoods of M1 miRNA (M1 miRNA group). 
+     (\textbf{D}) Regulation of gene expression of other gene group via triggering target (common target between M1 and M2)}
 \label{fig:fig3}
@@
 Therefore, we have integrated aforementioned interaction parameters that could be useful for predicting miRNA repression activity more accurately.
-The sample dataset used in Figure \ref{fig:fig1} is recalculated with additional factors Table \ref{tab:one} in effect with same trigger, two fold increase in Gene2 (Figure \ref{fig:fig3}).
-When the factors were taken into account in the system, miRNA efficiencies varied as shown in Figure \ref{fig:fig3}A.
+The sample dataset used in Figure \ref{fig:fig1} is recalculated with additional factors Table \ref{tab:one} in effect with same trigger (Detailed codes can be found at Section 3 in Supplementary Materials and Methods), two fold increase in Gene2 (Figure \ref{fig:fig3}B, and Figure S6).
+When the factors were taken into account in the system, miRNA efficiencies varied as shown in Figure \ref{fig:fig3}A (Figure S5).
 Although the miRNA:target expression ratios in steady-state were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed.
-On the other hand, changes in expression levels of G4 and G1 differ although they have same initial conditions (Detailed codes are available at Sections 3-4 in Supplementary Materials and Methods).
+On the other hand, changes in expression levels of G4 (Detailed codes are available at Section 4 in Supplementary Materials and Methods; and Figure S8) and G1 differ although they have same initial conditions.
 This is due to fact that G4 is targeted by two miRNAs, M1 and M2 (Figures \ref{fig:fig3}D) and changed interaction factors.
@@
 On the other hand, it is possible that taking into account of these interaction factors could cause increasing miRNA repression activity like in miRNA2:Gene4 interaction Figure \ref{fig:fig3}A.
-When expression of Gene2 (G2) increased, expression values of all genes also changed differentially because of contribution of efficiency factors (Figure \ref{fig:fig3} B,C,D).
+When expression of Gene2 (G2) increased, expression values of all genes also changed differentially because of contribution of efficiency factors (Figure \ref{fig:fig3} B,C,D and Figure S6-7).
 We have considered that in our approach energy and seed type of pairs is significant for binding and targeted region is important for repression.
@@
 
-In regulation of miRNA target sample system (Figure \ref{fig:fig3}), miRNA (M1) repressive efficiency on the primary triggering gene (G2) is low in steady-state.
+In regulation of miRNA target sample system (Figure \ref{fig:fig3} and Figure S5-7), miRNA (M1) repressive efficiency on the primary triggering gene (G2) is low in steady-state.
 So it has been observed that the regulatory activities of miRNA on the targets are weak after the increase of miRNA target G2.
@@
 On the other hand, it was observed that Gene2 was weakly affected from change of Gene4 expression because of its weak interaction factors.
-Since perturbation efficiency of each gene is different, we have developed a function which screens each gene in the network for their perturbation efficiencies (Detailed codes are available at Section 4 in Supplementary Materials and Methods).
-On the other hand, when we applied the method on the minimal dataset (i.e.
-Gene4 is trigger element), Gene4 has been found to be the most efficient element in terms of number of perturbed elements and miRNA2 (M2) has been found to be causing the highest mean in expression changes.
+Since perturbation efficiency of each gene is different, we have developed a function which screens each gene in the network for their perturbation efficiencies.
+When we applied the method on the minimal dataset (i.e.
+Gene4 is trigger element), Gene4 has been found to be the most efficient element in terms of number of perturbed elements and miRNA2 (M2) has been found to be causing the highest mean in expression changes (see Section 5 in Supplementary Materials and Methods).
 
@@
 This process were run with 3 fold upregulation and 10 iteration in conditions where elements that have less than five percent change are not taken into account (Codes are found at Section 6.5 in Supplementary Materials and methods).
-With this approach, SERPINE is found as one of the most effective gene nodes in network.
+With this approach, SERPINE2 has been found as one of the most effective gene nodes in network.
+We have observed that gaining steady-state in realistic network takes a lot of time because of complex interactions, as expected, and SERPINE2 exhibit highest efficiency at iteration 31 with given conditions (see Section 6.6 in Supplementary Materials and Methos and Figure S9 in Supplementary Figures).
+We ran the model with two fold of iteration which is observed as most effective.
 We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach (see Section 6.6 in Supplementary Materials and methods).

### 2019-09-30 NAR_form_article.tex
Figure defining

@@
 about factors effecting binding or efficiency of miRNA can be provided
-as separate columns (see Figure S1 in Supplementary Tables). After
+as separate columns (see Table S1 in Supplementary Tables). After
 construction of the network, amount of miRNA per target is calculated
@@
 criteria effecting the binding of miRNA to its target is provided,
-distribution of miRNA will be calculated accordingly. Target can be mRNA
-or any other ceRNA (circRNA, ncRNA, etc.) thus, throughout the
-manuscript terms target, gene and ceRNA are used interchangeably.
+distribution of miRNA will be calculated accordingly (available at Table
+S6 in Supplementary Tables). Target can be mRNA or any other ceRNA
+(circRNA, ncRNA, etc.) thus, throughout the manuscript terms target,
+gene and ceRNA are used interchangeably.
 
@@
 Initially, the network is assumed in steady-state (Figure
-\ref{fig:fig1}A) condition and needs least one trigger for initiating
-calculations. The trigger can be a change in expression level of one or
-more genes (Figure \ref{fig:fig1}B). After a trigger, the network
-undergoes iterative cycle of calculations at each of which distribution
-of miRNA in local neighborhood is recalculated (Figure \ref{fig:fig1}C).
-Based on new miRNA distribution, expression level of each node
-(i.e.~ceRNA) is updated. This update results in change of expression
-value of common target (G4) in system. In this case, the common target
-acts as trigger for the other group (the targets interacted with M2
-miRNA) in the network. Due to common targeted elements, the change in
-one neighborhood spreads to other neighborhoods (Figure
-\ref{fig:fig1}D), consequently have potential to affect whole network
-due to ``ripple effect'' (see Section 2 in Supplementary Materials and
-Methods).
+\ref{fig:fig1}A and Figure S2) condition and needs least one trigger for
+initiating calculations. The trigger can be a change in expression level
+of one or more genes (Figure \ref{fig:fig1}B and Figure S3). After a
+trigger, the network undergoes iterative cycle of calculations at each
+of which distribution of miRNA in local neighborhood is recalculated
+(Figure \ref{fig:fig1}C and Figure S4A). Based on new miRNA
+distribution, expression level of each node (i.e.~ceRNA) is updated.
+This update results in change of expression value of common target (G4)
+in system. In this case, the common target acts as trigger for the other
+group (the targets interacted with M2 miRNA) in the network. Due to
+common targeted elements, the change in one neighborhood spreads to
+other neighborhoods (Figure \ref{fig:fig1}D and Figure S4B),
+consequently have potential to affect whole network due to ``ripple
+effect'' (see Section 2 in Supplementary Materials and Methods).
 
@@
 influence after binding. In the literature, binding free energy
-\citetext{\citealp[
-]{cao_predicting_2012}; \citealp{helwak_mapping_2013}} and seed type
+\citep{cao_predicting_2012, helwak_mapping_2013} and seed type
 \citep{werfel_preferential_2017} in miRNA:target interactions are
@@
 between miRNA and gene. The cooperative repression activity of miRNAs to
-their common targets is figured out as shown in Figure \ref{fig:fig2}E
-(see Section 3 in Supplementary Materials and Methods).
+their common targets is figured out as shown in Figure \ref{fig:fig2}E.
 
@@
 dataset retrieved from published datasets
-\citep{helwak_mapping_2013, moore_mirnatarget_2015}. Detailed
-description of network construction and its code is available in
-Supplementary Materials and Methods. SERPINE2, one of the most effective
-nodes, has been used as trigger in ceRNA network simulation in breast
-cancer patient network. Relevant steps and codes are available at
-Supplementary Materials and Methods.
+\citep{helwak_mapping_2013, moore_mirnatarget_2015} (Section 6 in
+Supplementary Materials and Methods and Table S4 in Supplementary
+Tables). SERPINE2, one of the most effective nodes, has been used as
+trigger in ceRNA network simulation in breast cancer patient network.
+Relevant steps and codes are available at Supplementary Materials and
+Methods.
 
@@
 dataset used in Figure \ref{fig:fig1} is recalculated with additional
-factors Table \ref{tab:one} in effect with same trigger, two fold
-increase in Gene2 (Figure \ref{fig:fig3}). When the factors were taken
-into account in the system, miRNA efficiencies varied as shown in Figure
-\ref{fig:fig3}A. Although the miRNA:target expression ratios in
-steady-state were same in comparison with the sample dataset without
-factors, efficiency of binding and repression have changed. On the other
-hand, changes in expression levels of G4 and G1 differ although they
-have same initial conditions (Detailed codes are available at Sections
-3-4 in Supplementary Materials and Methods). This is due to fact that G4
-is targeted by two miRNAs, M1 and M2 (Figures \ref{fig:fig3}D) and
-changed interaction factors.
+factors Table \ref{tab:one} in effect with same trigger (Detailed codes
+can be found at Section 3 in Supplementary Materials and Methods), two
+fold increase in Gene2 (Figure \ref{fig:fig3}B, and Figure S6). When the
+factors were taken into account in the system, miRNA efficiencies varied
+as shown in Figure \ref{fig:fig3}A (Figure S5). Although the
+miRNA:target expression ratios in steady-state were same in comparison
+with the sample dataset without factors, efficiency of binding and
+repression have changed. On the other hand, changes in expression levels
+of G4 (Detailed codes are available at Section 4 in Supplementary
+Materials and Methods; and Figure S8) and G1 differ although they have
+same initial conditions. This is due to fact that G4 is targeted by two
+miRNAs, M1 and M2 (Figures \ref{fig:fig3}D) and changed interaction
+factors.
 
@@
 values of all genes also changed differentially because of contribution
-of efficiency factors (Figure \ref{fig:fig3} B,C,D). We have considered
-that in our approach energy and seed type of pairs is significant for
-binding and targeted region is important for repression.
+of efficiency factors (Figure \ref{fig:fig3} B,C,D and Figure S6-7). We
+have considered that in our approach energy and seed type of pairs is
+significant for binding and targeted region is important for repression.
 
@@
 
-In regulation of miRNA target sample system (Figure \ref{fig:fig3}),
-miRNA (M1) repressive efficiency on the primary triggering gene (G2) is
-low in steady-state. So it has been observed that the regulatory
-activities of miRNA on the targets are weak after the increase of miRNA
-target G2. When model is triggered with two fold increase in expression
-level of common target (G4), more prominent changes were observed in
-gene expression levels in network compared to changes caused by G2
-increase. Furthermore, change in expression level of target gene that
-has strong miRNA repression efficiency resulted in evident perturbation
-in network. On the other hand, it was observed that Gene2 was weakly
-affected from change of Gene4 expression because of its weak interaction
-factors. Since perturbation efficiency of each gene is different, we
-have developed a function which screens each gene in the network for
-their perturbation efficiencies (Detailed codes are available at Section
-4 in Supplementary Materials and Methods). On the other hand, when we
-applied the method on the minimal dataset (i.e.~Gene4 is trigger
-element), Gene4 has been found to be the most efficient element in terms
-of number of perturbed elements and miRNA2 (M2) has been found to be
-causing the highest mean in expression changes.
+In regulation of miRNA target sample system (Figure \ref{fig:fig3} and
+Figure S5-7), miRNA (M1) repressive efficiency on the primary triggering
+gene (G2) is low in steady-state. So it has been observed that the
+regulatory activities of miRNA on the targets are weak after the
+increase of miRNA target G2. When model is triggered with two fold
+increase in expression level of common target (G4), more prominent
+changes were observed in gene expression levels in network compared to
+changes caused by G2 increase. Furthermore, change in expression level
+of target gene that has strong miRNA repression efficiency resulted in
+evident perturbation in network. On the other hand, it was observed that
+Gene2 was weakly affected from change of Gene4 expression because of its
+weak interaction factors. Since perturbation efficiency of each gene is
+different, we have developed a function which screens each gene in the
+network for their perturbation efficiencies. When we applied the method
+on the minimal dataset (i.e.~Gene4 is trigger element), Gene4 has been
+found to be the most efficient element in terms of number of perturbed
+elements and miRNA2 (M2) has been found to be causing the highest mean
+in expression changes (see Section 5 in Supplementary Materials and
+Methods).
 
@@
 percent change are not taken into account (Codes are found at Section
-6.5 in Supplementary Materials and methods). With this approach, SERPINE
-is found as one of the most effective gene nodes in network. We have
-observed that the results of simulation with SERPINE2 gene, a serine
-protease inhibitor, one of the overexpressed genes of breast cancer
+6.5 in Supplementary Materials and methods). With this approach,
+SERPINE2 has been found as one of the most effective gene nodes in
+network. We have observed that gaining steady-state in realistic network
+takes a lot of time because of complex interactions, as expected, and
+SERPINE2 exhibit highest efficiency at iteration 31 with given
+conditions (see Section 6.6 in Supplementary Materials and Methos and
+Figure S9 in Supplementary Figures). We ran the model with two fold of
+iteration which is observed as most effective. We have observed that the
+results of simulation with SERPINE2 gene, a serine protease inhibitor,
+one of the overexpressed genes of breast cancer
 \citep{yang2018expression, candia2006protease} is not convenient with

### 2019-09-30 Supplementary_figures.Rmd
article NAR form
scanning
Figure defining

@@
 
+```{r}
+minsamp<-readRDS("data/minsamp.RDS")
+```
+
 ```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset in Steady-state", fig.width=5, fig.height=4}
@@
 ```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of minsamp data.
-a) First response of system to Gene2 upregulation (2nd iteration).
-b) Spreading of perturbation on system (3th iteration)", fig.subcap= NULL}
+A) First response of system to Gene2 upregulation (2nd iteration).
+B) Spreading of perturbation on system (3th iteration)", fig.subcap= NULL}
 
@@
   simulate(cycle=1)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "a)")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "A")
 
@@
   simulate(2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "b)")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "B")
 
@@
 
-```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of minsamp data with interaction factors a)First response of system to Gene2 upregulation (2nd iteration).
-b)Spreading of perturbation on system (3th iteration)"}
+```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of minsamp data with interaction factors A)First response of system to Gene2 upregulation (2nd iteration).
+B)Spreading of perturbation on system (3th iteration)"}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
@@
   simulate(cycle = 1)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "a)")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "A")
 
@@
   simulate(cycle = 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "b)")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "B")
 
@@
 ```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Perturbation of Gene4 on minsamp data with interaction factors.
-a) Netwrk at steady-state.
-b) Upregulation of Gene4.
-c) Primary response of network to upregulation of Gene4.
-d) Re-regulation of whole nodes on system (3th iteration)"}
+A) Network at steady-state.
+B) Upregulation of Gene4.
+C) Primary response of network to upregulation of Gene4.
+D) Re-regulation of whole nodes on system (3th iteration)"}
 
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "a)")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "A")
 
@@
   update_how("Gene4", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "b)")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "B")
 
@@
   simulate(cycle=1)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "c)")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "C")
 
@@
   simulate(cycle=2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "d)")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "D")
 ```
@@
 ```{r}
-E9GE_mirnagenenormal<-readRDS("data/E9GE_mirnagenenormal.rda")
+E9GE_mirnagenenormal<-readRDS("data/E9GE_mirnagenenormal.RDS")
 head(E9GE_mirnagenenormal)

### 2019-09-30 Supplementary_mat_met.Rmd
Figure defining

@@
 # 1. Functions defined for ceRNA models and workflow of method
+
 We defined the functions that can be used with R programming.
@@
 
+ A part of dataset can be found at Table S4 in Supplementary Tables.
+
 ```{r}
@@
 So, we offered an approach about to find iteration.
-`find_iteration` function does not give the iteration to gain steady-state, but it gives the iteration which has maximum affected node counts.
+`find_iteration` function does not give the iteration to gain steady-state, but it gives the iteration which has maximum affected node counts .
 The function is applied as following:

### 2019-10-08 NAR_form_article.Rmd
eps convertion of pics
article check revisions

@@
 \begin{center}
-\includegraphics[width=15cm,height=10cm]{Fig1.jpg}
+\includegraphics[width=16cm,height=9cm]{Fig1.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=9cm,height=17cm]{Fig2.jpg}
+\includegraphics[width=9cm,height=17cm]{Fig2.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=15cm,height=8cm]{Fig3.jpg}
+\includegraphics[width=15cm,height=8cm]{Fig3.eps}
 \end{center}
@@
 Although the miRNA:target expression ratios in steady-state were same in comparison with the sample dataset without factors, efficiency of binding and repression have changed.
-On the other hand, changes in expression levels of G4 (Detailed codes are available at Section 4 in Supplementary Materials and Methods; and Figure S8) and G1 differ although they have same initial conditions.
-This is due to fact that G4 is targeted by two miRNAs, M1 and M2 (Figures \ref{fig:fig3}D) and changed interaction factors.
+On the other hand, changes in expression levels of G4 (Detailed codes are available in Figure S8 and Section 4 in Supplementary Materials and Methods) and G1 differ although they have same initial conditions.
+This is due to fact that G4 is targeted by two miRNAs, M1 and M2 (Figure \ref{fig:fig3}D) and changed interaction factors.
 
@@
 We have analysed the realistic network in context of perturbation efficiencies of nodes.
-This process were run with 3 fold upregulation and 10 iteration in conditions where elements that have less than five percent change are not taken into account (Codes are found at Section 6.5 in Supplementary Materials and methods).
+This process was run with 3 fold upregulation and 10 iteration in conditions where elements that have less than five percent change are not taken into account (Codes are found at Section 6.5 in Supplementary Materials and Methods).
 With this approach, SERPINE2 has been found as one of the most effective gene nodes in network.
@@
 We ran the model with two fold of iteration which is observed as most effective.
-We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach (see Section 6.6 in Supplementary Materials and methods).
+We have observed that the results of simulation with SERPINE2 gene, a serine protease inhibitor, one of the overexpressed genes of breast cancer [@yang2018expression; @candia2006protease] is not convenient with tumor tissue expression values, when we compared the tumor gene expression dataset retrieved from TCGA and simulation results obtained via our approach (see Section 6.6 in Supplementary Materials and Methods).
 Simulation results were not expected to be consistent with tumor tissue expression values because large number of abnormal regulations were observed in these tissues.

### 2019-10-08 NAR_form_article.Rmd
delete todos

@@
 ## Breast cancer patient dataset
-
 We have applied our model in a real dataset for which experimental measurements of various factors were available.
@@
 Relevant steps and codes are available at Supplementary Materials and Methods.
-
-[#TODO]: <> "about formulas, 2) in formulas eq:1 and eq:2, C_m or Eff_gi is not intuitive, is there consensus terms about effect or nth gene?"
 
@@
 This is due to fact that G4 is targeted by two miRNAs, M1 and M2 (Figure \ref{fig:fig3}D) and changed interaction factors.
-
-[#TODO]: <> "figure1, 2 ve 3 uzerinden farkli dagilim ve degradasyon açıklaması buraya. G2 uzerinden olmasa iyi olur"
 
@@
 
-# ACKNOWLEDGEMENTS
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text.
-
 ### Conflict of interest statement.

### 2019-10-08 NAR_form_article.tex
eps convertion of pics
delete todos

@@
 repression have changed. On the other hand, changes in expression levels
-of G4 (Detailed codes are available at Section 4 in Supplementary
-Materials and Methods; and Figure S8) and G1 differ although they have
+of G4 (Detailed codes are available in Figure S8 and Section 4 in
+Supplementary Materials and Methods) and G1 differ although they have
 same initial conditions. This is due to fact that G4 is targeted by two
-miRNAs, M1 and M2 (Figures \ref{fig:fig3}D) and changed interaction
+miRNAs, M1 and M2 (Figure \ref{fig:fig3}D) and changed interaction
 factors.
@@
 We have analysed the realistic network in context of perturbation
-efficiencies of nodes. This process were run with 3 fold upregulation
-and 10 iteration in conditions where elements that have less than five
+efficiencies of nodes. This process was run with 3 fold upregulation and
+10 iteration in conditions where elements that have less than five
 percent change are not taken into account (Codes are found at Section
-6.5 in Supplementary Materials and methods). With this approach,
+6.5 in Supplementary Materials and Methods). With this approach,
 SERPINE2 has been found as one of the most effective gene nodes in
@@
 via our approach (see Section 6.6 in Supplementary Materials and
-methods). Simulation results were not expected to be consistent with
+Methods). Simulation results were not expected to be consistent with
 tumor tissue expression values because large number of abnormal
@@
 
-\section{ACKNOWLEDGEMENTS}
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text.
-
 \subsubsection{Conflict of interest statement.}

### 2019-10-11 NAR_form_article.Rmd
typo check

@@
 Apart from tools predicting miRNA:target pairs, there are databases that provide  experimentally validated miRNA:target interaction datasets.
-The reliability in these datasets range from weak (e.g, western blot) to strong (e.g, CLiP, CLASH).
-High-throughput methods ensure handling of AGO (Argounate Protein) interacted miRNAs (CLiP) and their targets as well (CLEAR-CLiP, CLASH).
+The reliability in these datasets range from weak (e.g, western blot) to strong (.g, CLiP, CLASH).
+High-throughput methods ensure handling of AGO (Argonaute Protein) interacted miRNAs (CLiP) and their targets as well (CLEAR-CLiP, CLASH).
 We constructed a large ceRNA network utilizing high-throughput datasets containing chimeric reading of miRNA:target pairs [@helwak_mapping_2013; @moore_mirnatarget_2015] which expose exact complementary seed regions.
@@
 With this approach, SERPINE2 has been found as one of the most effective gene nodes in network.
-We have observed that gaining steady-state in realistic network takes a lot of time because of complex interactions, as expected, and SERPINE2 exhibit highest efficiency at iteration 31 with given conditions (see Section 6.6 in Supplementary Materials and Methos and Figure S9 in Supplementary Figures).
+We have observed that gaining steady-state in realistic network takes a lot of time because of complex interactions, as expected, and SERPINE2 exhibit highest efficiency at iteration 31 with given conditions (see Section 6.6 in Supplementary Materials and Methods and Figure S9 in Supplementary Figures).
 We ran the model with two fold of iteration which is observed as most effective.

### 2019-10-11 Supplementary_additional_file.Rmd
scanning
Figure defining
typo check

@@
 
-MiRNA releases are obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) is utilised.
+MiRNA releases are obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) was downloaded.
 
@@
 
-CLASH dataset is published in miRBase erlease 15 and Human Genome 19 version.
+CLASH dataset is published in miRBase release 15 and Human Genome 19 version.
 
@@
 
+<!-- #TODO maybe purrr can be used, instead of repeating as.character -->
+
 ## Interpreting the CLASH seed structures in dataset
@@
 
-CLASH dataset was retrieved from [Nature web page](nature.com/articles/ncomms9864)
+CLASH dataset was retrieved from [Nature article](nature.com/articles/ncomms9864)
 
@@
 
+<!-- #TODO the sections above needs some text explaining what is done -->
+
 ## Adding Coefficients of Interaction factors
+
+Numeric values of interaction factors are shown in Table S2-3 in Supplementary Tables file.
 
@@
 
+The context of dataset is shown in Table S5 in Supplementary Tables.
+
 # REFERENCES

### 2019-10-11 Supplementary_figures.Rmd
typo check

@@
 
-## *minsamp* dataset analysis at lack of interaction factors.
+## *minsamp* dataset analysis in absence of interaction factors.
 
@@
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)%>%
-  simulate(cycle=1)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "A")
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression) %>%
+  update_how("Gene2", 2) %>%
+  simulate(cycle=1) %>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            Downregulation = "blue", 
+            title = "A")
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression)%>%
   update_how("Gene2", 2)%>%
   simulate(2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "B")
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            Downregulation = "blue", 
+            title = "B")
 

### 2019-10-11 Supplementary_mat_met.Rmd
typo check

@@
 On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger.
-The functions that are used in the approach are developed with R programming so as can be used with other packages.
-These are can be found in the Github repository [ceRNAnetsim Github page](https://github.com/selcenari/ceRNAnetsim) and improved with contributions from others.
+The functions that are used in the this study are developed with R and their source code is available in [Github repository](https://github.com/selcenari/ceRNAnetsim).
 
@@
 
-When the graphs which were resulted from analyses were examined, it was observed that behaviors were same.
-But, when the results were analysed in terms of expression values, the regulation differences can be observed.
+When the graphs were examined, it was observed that behaviors were same. But, when the results were analysed in terms of expression values, the regulation differences can be observed.
 
@@
 
-We downloaded the mirbase release 21 dataset from [mirbase](http://www.mirbase.org/ftp.shtml) and processed the patient mirna expression datasets as following:
+We downloaded the mirbase release 21 dataset from [mirbase](http://www.mirbase.org/ftp.shtml) and processed the patient miRNA expression datasets as following:
 
@@
 ```
-This dataset, perturbationofnetwork, includes totally 423 effective nodes. We selected SERPINE2 gene to perturb the network because it is the most efficient node gene in network.
+This dataset, `perturbationofnetwork`, includes totally 423 effective nodes. We selected SERPINE2 gene to perturb the network because it is the most efficient node gene in network.
 
@@
 
-Actually, we have developed to provide a new approach mirna mediated regulation networks.
+Actually, we have developed to provide a new approach miRNA mediated regulation networks.
 This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and coherent miRNA:target regulation approach.

### 2019-10-11 Supplementary_tables.Rmd
scanning
typo check

@@
 
-Table: *minsamp* sample dataset that includes interaction factors.
+Table: *minsamp* sample dataset that includes interaction factors. The parameters which affect miRNA:target interactions (i.e. seed type, region, energy) are provided in sample dataset.
 
@@
 We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
-With this process, we have handled this entegrated dataset in context of competitor behaviours and functionality of interactions.
+With this process, we have handled this integrated dataset in context of competitor behaviors and functionality of interactions.
 
@@
    cluster    & character & Barcode from experimentally method \\
-   chromosome    & character & Chromosome of Terget gene from raw data \\
+   chromosome    & character & Chromosome of Target gene from raw data \\
    start\_position    & numeric & Gene start position from raw data \\
    end\_position     & numeric & Gene end position from raw data \\
-   strand     & chracter & Gene strand \\
+   strand     & character & Gene strand \\
    hgnc\_symbol     & character & Gene name (Symbol) \\
@@
    target\_seq   & character & mRNA sequences targeted by miRNA \\
-   miRNA   & character & miRNA id (from miRBase vesion 21 ) \\
+   miRNA   & character & miRNA id (from miRBase version 21 ) \\
    miR\_seq   & character & miRNA sequence \\
@@
    HG38build\_loc   & character & Recent chromosomal location of Gene \\
-   Genom\_build   & character& Genom build of given chromosome, start and end positions \\
+   Genome\_build   & character& Genome build of given chromosome, start and end positions \\
    region   & character & interaction location on target \\ 
@@
 
-As a result of simulation a dataset, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at followings.
+As a result of simulation a dataset, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at following.
 

### 2019-10-14 Supplementary_additional_file.Rmd
fixed todos

@@
 library(readxl)
+library(biomaRt)
 knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
@@
 
-# Arrangement of CLASH dataset [@helwak_mapping_2013]
+# Arrangement of CLASH dataset
 
-CLASH dataset was retrieved from [PubMed](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3650559/)
+CLASH dataset was retrieved from [PubMed](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3650559/) [@helwak_mapping_2013].
 
@@
 
-clashelwak <- read.table("mmc1.txt", comment.char = "#", header = TRUE, skip = 1)
+clashelwak <- read.table("mmc1.txt", comment.char = "#", header = TRUE, skip = 1, stringsAsFactors = FALSE)
 
@@
 ## Query of Human Genome 19.
+
+Human genome 19 information was handled through biomaRt package.
 
@@
 
+There are different liftover methods for conversion among Human Genome builds. We preffered to use [UCSC liftover tool](https://genome.ucsc.edu/cgi-bin/hgLiftOver)
+
 ```{r,  eval=FALSE}
+
+#Obtaining chromosomal locations from miRNA:target interaction dataset.
+
 lift19 <- clashelwakfinal%>%
@@
 
-# Lift over process is made via UCSC liftover tool. (https://genome.ucsc.edu/cgi-bin/hgLiftOver)
+#After we searched this file in UCSC browser, the output loaded (lift19_del deleted regions on the HG38 genome build; hg38clashcomp new locations of the genes on HG38)
 
@@
 
+#removing deleted location from CLASH dataset
+
 clashelwakfinal <-clashelwakfinal%>%
@@
 
-hg38clash <- read.delim("hg38clashcomp.txt", header = FALSE)
+hg38clash <- read.delim("hg38clashcomp.txt", header = FALSE, stringsAsFactors = FALSE)
+
+# adding new location information:
 
@@
   dplyr::mutate(Genom_build= rep("hg19"))
-str(clashelwakfinal)
 
@@
 
-clashelwakfinal$cluster <- as.character(clashelwakfinal$cluster)
 clashelwakfinal$strand <- as.character(clashelwakfinal$strand)
-clashelwakfinal$target_seq <- as.character(clashelwakfinal$target_seq)
-clashelwakfinal$miR_seq <- as.character(clashelwakfinal$miR_seq)
-clashelwakfinal$seed_type <- as.character(clashelwakfinal$seed_type)
-clashelwakfinal$HG38build_loc <- as.character(clashelwakfinal$HG38build_loc)
-clashelwakfinal$seed_type2 <- as.numeric(clashelwakfinal$seed_type2)
-clashelwakfinal$seed_type3 <- as.character(clashelwakfinal$seed_type3)
+
+str(clashelwakfinal)
 ```
-
-<!-- #TODO maybe purrr can be used, instead of repeating as.character -->
 
@@
 ```{r,   eval=FALSE}
+
+#Obtaining chromosomal locations from miRNA:target interaction dataset.
+
 lift18 <- clearclipfinal%>%
@@
 write_tsv(lift18, "lift18.txt")
+
+#After we searched this file in UCSC browser, the output loaded (deleted_lift18 deleted regions on the HG38 genome build; hg38clearclip new locations of the genes on HG38)
 
@@
 
+#removing deleted location from CLEAR-CLiP dataset
+
 clearclipfinal <- clearclipfinal%>%
@@
 
-hg38clearclip<- read.delim("hg38clearclip.txt", header = FALSE)
+hg38clearclip<- read.delim("hg38clearclip.txt", header = FALSE, stringsAsFactors = FALSE)
 
@@
 
+# adding new location information:
+
 clearclipfinal <- clearclipfinal%>%
@@
 
+# Arrangement in dataset
+
 clearclipfinal%>%
   dplyr::select(cluster=cluster.ID, chromosome = chr, start_position, end_position, strand = strand.y, hgnc_symbol, Ensembl_Gene_Id =ensembl_gene_id, Ensembl_Transcript_Id =ensembl_transcript_id, target_seq = target.map, miRNA, miR_seq = miR.map, seed_type = 'seed match', Energy = MFE, HG38build_loc, Genom_build, region) -> clearclipfinal
+
 ```
 
-## Seed type manipulation
+## Seed type manipulation in CLEAR-CLiP dataset
+
+In CLEAR-CLiP dataset, seed types were shown in detail. We adjusted as canonical and non-canonical.
 
@@
   dplyr::select(1:11, seed_type = seed_type_com, Energy, HG38build_loc, Genom_build, region)-> clearclipfinal
-
-clearclipfinal$HG38build_loc <- as.character(clearclipfinal$HG38build_loc)
 
@@
 
-<!-- #TODO the sections above needs some text explaining what is done -->
-
 ## Adding Coefficients of Interaction factors
 
-Numeric values of interaction factors are shown in Table S2-3 in Supplementary Tables file.
+Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study.
+On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study.
+Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions.
+Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values.
+We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
+With this process, we have handled this integrated dataset in context of competitor behaviors and functionality of interactions.
+
+In this step we added numeric intraction values at followings
+
+Fistly, we organized these values due to the fact that the regions were defined differently in two datasets. After that, region effect was added as numeric values (shown in Table S3).
 
@@
 ```
+
+Secondly, we organized seed type interactions in *Seed type manipulation* section for CLEAR-CLiP dataset to show as found in CLASH dataset.
+Same type formatted values added dataset as numeric values (shown in Table S2).
 

### 2019-10-14 Supplementary_tables.Rmd
fixed todos

@@
 
-Note: Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study.
-On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study.
-Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions.
-Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values.
-We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
-With this process, we have handled this integrated dataset in context of competitor behaviors and functionality of interactions.
+Some of information about miRNA:target interactions were exhibited directly by high-throughput studies. On the other haned, we were examined other interacion factors based on different studies.
+
++ [@helwak_mapping_2013; @moore_mirnatarget_2015] reported the energy values in miRNA:target interactions.
++ Comparisons of canonical seed types were evaluated by study of [@grimson_microrna_2007], while functional and non-functional seed interactions were studied by [@bartel_micrornas:_2009] and [@betel2010comprehensive].
++ Numeric definition of target region location effect was performed based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013]
 

### 2019-10-15 NAR_form_article.Rmd
xps to eps conversion

@@
 \begin{center}
-\includegraphics[width=16cm,height=9cm]{Fig1.eps}
+\includegraphics[width=21cm,height=12cm]{fig1_new.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=9cm,height=17cm]{Fig2.eps}
+\includegraphics[width=9cm,height=17cm]{fig2_new.eps}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=15cm,height=8cm]{Fig3.eps}
+\includegraphics[width=15cm,height=8cm]{fig3_new.eps}
 \end{center}

### 2019-10-15 NAR_form_article.tex
xps to eps conversion

@@
 reliability in these datasets range from weak (e.g, western blot) to
-strong (e.g, CLiP, CLASH). High-throughput methods ensure handling of
-AGO (Argounate Protein) interacted miRNAs (CLiP) and their targets as
-well (CLEAR-CLiP, CLASH). We constructed a large ceRNA network utilizing
+strong (.g, CLiP, CLASH). High-throughput methods ensure handling of AGO
+(Argonaute Protein) interacted miRNAs (CLiP) and their targets as well
+(CLEAR-CLiP, CLASH). We constructed a large ceRNA network utilizing
 high-throughput datasets containing chimeric reading of miRNA:target
@@
 SERPINE2 exhibit highest efficiency at iteration 31 with given
-conditions (see Section 6.6 in Supplementary Materials and Methos and
+conditions (see Section 6.6 in Supplementary Materials and Methods and
 Figure S9 in Supplementary Figures). We ran the model with two fold of

### 2019-12-06 NAR_form_article.Rmd
addition
06 dec dit on article

@@
 The model is scalable and can be plugged in with emerging miRNA effectors such as circRNAs.
-The model is available as R package ceRNAnetsim https://github.com/selcenari/ceRNAnetsim.
+The model is available as R package ceRNAnetsim (https://github.com/selcenari/ceRNAnetsim).
 
@@
 
-# INTRODUCTION
+# \textbf{INTRODUCTION}
 
-MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key regulator of gene expression through various post-transcriptional mechanisms.
+MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key regulator of gene expression through various post-transcriptional mechanisms [@brennecke2005principles].
 Although the mechanisms by which miRNA represses are not fully understood, miRNAs predominantly repress their targets.
@@
 
-# MATERIALS AND METHODS
+# \textbf{MATERIALS AND METHODS}
 
@@
 \begin{center}
-\includegraphics[width=21cm,height=12cm]{fig1_new.eps}
+\includegraphics[width=15cm,height=8cm]{fig1_new.eps}
 \end{center}
@@
     (\textbf{C}) Distribution of miRNA1 (M1) changes. 
-    (\textbf{D)}) The change at expression of common target affects changes of proportional distribution of miRNA2 (M2).
+    (\textbf{D}) The change at expression of common target affects changes of proportional distribution of miRNA2 (M2).
 Expression values are rounded to integers for simplicity.
@@
 
+Not all miRNA:target binding events result in degradation of target.
+The degradation of target by bound miRNA depends on efficiency factors such as binding region.
+Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:4} (Figure \ref{fig:fig2}D); $RE^\prime_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
+The cooperative repression activity of miRNAs to their common targets is figured out as shown in Figure \ref{fig:fig2}E.
+
 \begin{figure}[ht]
@@
 
-Not all miRNA:target binding events result in degradation of target.
-The degradation of target by bound miRNA depends on efficiency factors such as binding region.
-Exact repression efficiency of miRNA is calculated according to equation \eqref{eq:4} (Figure \ref{fig:fig2}D); $RE^\prime_{gi}$, normalized values of region efficiency coefficient between miRNA and gene.
-The cooperative repression activity of miRNAs to their common targets is figured out as shown in Figure \ref{fig:fig2}E.
-
 ## Breast cancer patient dataset
 We have applied our model in a real dataset for which experimental measurements of various factors were available.
-Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network https://www.cancer.gov/tcga.
+Expression levels of miRNA and genes in tumor and normal tissue of single patient are retrieved from TCGA Research Network; https://www.cancer.gov/tcga.
 High-throughput experimental datasets which are provided miRNA:gene target pairs with interaction factors [@helwak_mapping_2013; @moore_mirnatarget_2015] (Detailed codes are available at Section 6 in Supplementary Materials and Methods).
@@
 
-# RESULTS
+# \textbf{RESULTS}
 
@@
   \hline
- & competing & miRNA & Competing expression & miRNA expression & seed type & region & energy \\ 
+ & Competing & miRNA & Competing Expression & miRNA Expression & Seed Type & Region & Energy \\ 
   \hline
@@
 
-# DISCUSSION
+# Additions to be fixed
+
+Additionally, miRNA:gene expression dataset which was integrated via miRTarBase [@chou2017mirtarbase] was analysed in mean of perturbation efficiency to determine crucial genes on the network.
+We have observed that two miRNAs and 488 genes are effective in terms of perturbation efficiency, when we ran our model with more than 15% efficiency on target expression.
+miR-10b was investigated via HMDD database [@huang2018hmdd] and mostly found as metastatic biomarker.
+On the other hand, it was observed that miR-148a is critical for promotion of apoptosis and supression of breast cancer cell growth.
+Additionaly, to obtain the number of genes in a narrow range, the model was ran with 17% efficency on target expression and 297 genes and one miRNA were observed as highly perturbation-effective.
+We analysed these genes via eDGAR, a database of Disease-Gene Associations with annotated Relationships among genes [@babbi2017edgar] that supplies different observations about crucial genes such as direct physical interactions, other shared genes with these interactions, protein-protein interactions, GO terms etc.
+With our model, we obtained 3 (TP53, BRCA1, AKT1) of 27 critical breast cancer genes and these genes show direct physical interaction with 21 genes.
+33 shared genes which are observed on these direct interactions were also found in our effective perturbation elements.We have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4, GATA3, CTBP1) transcription factors (TFs) that are common in 44 critical TFs which are provided by eDGAR dataset.
+44 of regulated genes via these TFs were found in our highly effective genes.
+it has been observed that these genes cover 22.5% of 297 perturbation-effective genes.
+
+When we analysed these genes via biomaRt OMIM query, it has been observed that our model provides enriched 1.45 fold results with hypergeometric p-values 3.19e-05.
+
+# \textbf{DISCUSSION}
 
@@
 
-# CONCLUSION
+# \textbf{CONCLUSION}
 
@@
 None declared.
-
-\newpage

### 2019-12-06 NAR_form_article.tex
addition
06 dec dit on article

@@
 
-\usepackage{float}
-
-\floatplacement{figure}{H}
-
-\usepackage[colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue]{hyperref} \usepackage[T1]{fontenc}
+\usepackage{float} \floatplacement{figure}{H} \usepackage[colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue]{hyperref} \usepackage[T1]{fontenc}
 
@@
 available as R package ceRNAnetsim
-\url{https://github.com/selcenari/ceRNAnetsim}.
+(\url{https://github.com/selcenari/ceRNAnetsim}).
 \end{abstract}
 
-\section{INTRODUCTION}
+\section{\texorpdfstring{\textbf{INTRODUCTION}}{}}
 
@@
 regulator of gene expression through various post-transcriptional
-mechanisms. Although the mechanisms by which miRNA represses are not
-fully understood, miRNAs predominantly repress their targets. Repressive
-activities of miRNAs vary depending on many factors that are significant
-to miRNA:target interactions. These factors include miRNA:target binding
-energy, binding location in target sequence, base pairing types between
-miRNA and target, abundance of miRNAs and targets
-\citep{grimson_microrna_2007}. Binding energies of miRNA:target
-complexes vary based on nucleotide context and determine folding
-stability of complex \citep{cao_predicting_2012}. It has been
+mechanisms \citep{brennecke2005principles}. Although the mechanisms by
+which miRNA represses are not fully understood, miRNAs predominantly
+repress their targets. Repressive activities of miRNAs vary depending on
+many factors that are significant to miRNA:target interactions. These
+factors include miRNA:target binding energy, binding location in target
+sequence, base pairing types between miRNA and target, abundance of
+miRNAs and targets \citep{grimson_microrna_2007}. Binding energies of
+miRNA:target complexes vary based on nucleotide context and determine
+folding stability of complex \citep{cao_predicting_2012}. It has been
 demonstrated that the binding energy between miRNA and target indicates
@@
 
-\section{MATERIALS AND METHODS}
+\section{\texorpdfstring{\textbf{MATERIALS AND METHODS}}{}}
 
@@
 
+Not all miRNA:target binding events result in degradation of target. The
+degradation of target by bound miRNA depends on efficiency factors such
+as binding region. Exact repression efficiency of miRNA is calculated
+according to equation \eqref{eq:4} (Figure \ref{fig:fig2}D);
+\(RE^\prime_{gi}\), normalized values of region efficiency coefficient
+between miRNA and gene. The cooperative repression activity of miRNAs to
+their common targets is figured out as shown in Figure \ref{fig:fig2}E.
+
 \begin{figure}[ht]
@@
 
-Not all miRNA:target binding events result in degradation of target. The
-degradation of target by bound miRNA depends on efficiency factors such
-as binding region. Exact repression efficiency of miRNA is calculated
-according to equation \eqref{eq:4} (Figure \ref{fig:fig2}D);
-\(RE^\prime_{gi}\), normalized values of region efficiency coefficient
-between miRNA and gene. The cooperative repression activity of miRNAs to
-their common targets is figured out as shown in Figure \ref{fig:fig2}E.
-
 \subsection{Breast cancer patient dataset}
@@
 miRNA and genes in tumor and normal tissue of single patient are
-retrieved from TCGA Research Network \url{https://www.cancer.gov/tcga}.
+retrieved from TCGA Research Network; \url{https://www.cancer.gov/tcga}.
 High-throughput experimental datasets which are provided miRNA:gene
@@
 
-\section{RESULTS}
+\section{\texorpdfstring{\textbf{RESULTS}}{}}
 
@@
 
-\section{DISCUSSION}
+\section{Additions to be fixed}
+
+Additionally, miRNA:gene expression dataset which was integrated via
+miRTarBase \citep{chou2017mirtarbase} was analysed in mean of
+perturbation efficiency to determine crucial genes on the network. We
+have observed that two miRNAs and 488 genes are effective in terms of
+perturbation efficiency, when we ran our model with more than 15\%
+efficiency on target expression. miR-10b was investigated via HMDD
+database \citep{huang2018hmdd} and mostly found as metastatic biomarker.
+On the other hand, it was observed that miR-148a is critical for
+promotion of apoptosis and supression of breast cancer cell growth.
+Additionaly, to obtain the number of genes in a narrow range, the model
+was ran with 17\% efficency on target expression and 297 genes and one
+miRNA were observed as highly perturbation-effective. We analysed these
+genes via eDGAR, a database of Disease-Gene Associations with annotated
+Relationships among genes \citep{babbi2017edgar} that supplies different
+observations about crucial genes such as direct physical interactions,
+other shared genes with these interactions, protein-protein
+interactions, GO terms etc. With our model, we obtained 3 (TP53, BRCA1,
+AKT1) of 27 critical breast cancer genes and these genes show direct
+physical interaction with 21 genes. 33 shared genes which are observed
+on these direct interactions were also found in our effective
+perturbation elements.We have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4,
+GATA3, CTBP1) transcription factors (TFs) that are common in 44 critical
+TFs which are provided by eDGAR dataset. 44 of regulated genes via these
+TFs were found in our highly effective genes. it has been observed that
+these genes cover 22.5\% of 297 perturbation-effective genes.
+
+When we analysed these genes via biomaRt OMIM query, it has been
+observed that our model provides enriched 1.45 fold results with
+hypergeometric p-values 3.19e-05.
+
+\section{\texorpdfstring{\textbf{DISCUSSION}}{}}
 
@@
 
-\section{CONCLUSION}
+\section{\texorpdfstring{\textbf{CONCLUSION}}{}}
 
@@
 
-\newpage
-
 \bibliographystyle{nar}

### 2019-12-06 Supplementary_mat_met.Rmd
06 dec dit on article

@@
 library(ceRNAnetsim)
+library(stringr)
 ```
@@
 ```
-This dataset, `perturbationofnetwork`, includes totally 423 effective nodes. We selected SERPINE2 gene to perturb the network because it is the most efficient node gene in network.
+This dataset, `perturbationofnetwork`, includes totally 420 effective nodes. We selected SERPINE2 gene to perturb the network because it is the most efficient node gene in network.
 

### 2019-12-09 NAR_form_article.Rmd
revise last paragraph

@@
 
-## Simulation of large and realistic network
+## Simulation of a large and realistic network
 
@@
 
-# Additions to be fixed
+# Genes causing widespread perturbation
 
 Additionally, miRNA:gene expression dataset which was integrated via miRTarBase [@chou2017mirtarbase] was analysed in mean of perturbation efficiency to determine crucial genes on the network.
-We have observed that two miRNAs and 488 genes are effective in terms of perturbation efficiency, when we ran our model with more than 15% efficiency on target expression.
-miR-10b was investigated via HMDD database [@huang2018hmdd] and mostly found as metastatic biomarker.
-On the other hand, it was observed that miR-148a is critical for promotion of apoptosis and supression of breast cancer cell growth.
-Additionaly, to obtain the number of genes in a narrow range, the model was ran with 17% efficency on target expression and 297 genes and one miRNA were observed as highly perturbation-effective.
-We analysed these genes via eDGAR, a database of Disease-Gene Associations with annotated Relationships among genes [@babbi2017edgar] that supplies different observations about crucial genes such as direct physical interactions, other shared genes with these interactions, protein-protein interactions, GO terms etc.
+We have observed that one miRNA and 297 genes have high perturbation efficiency, when we ran our model with more than 17% efficiency on target expression.
+The miRNA, miR-10b was investigated via HMDD database [@huang2018hmdd] and was found as metastatic biomarker.
+The genes with high perturbation efficiency were analysed via eDGAR, a database of Disease-Gene Associations with annotated relationships among genes [@babbi2017edgar] that supplies different observations about crucial genes such as direct physical interactions, other shared genes with these interactions, protein-protein interactions, GO terms etc.
 With our model, we obtained 3 (TP53, BRCA1, AKT1) of 27 critical breast cancer genes and these genes show direct physical interaction with 21 genes.
-33 shared genes which are observed on these direct interactions were also found in our effective perturbation elements.We have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4, GATA3, CTBP1) transcription factors (TFs) that are common in 44 critical TFs which are provided by eDGAR dataset.
+33 shared genes which are observed on these direct interactions were also found in our effective perturbation elements.
+We have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4, GATA3, CTBP1) transcription factors (TFs) that are common in 44 critical TFs which are provided by eDGAR dataset.
 44 of regulated genes via these TFs were found in our highly effective genes.

### 2019-12-20 NAR_form_article.Rmd
Adding discussion final part and selcen's name

@@
 author:
-  - name: Selcen Ari
+  - name: Selcen Ari Yuka
     affiliation: "1"
@@
 
+In breast cancer sample, all the genes with high perturbation efficiency were found to be all connected to miR10b.
+This can be explained by very high expression level of that miRNA.
+In a large network where miRNAs have balanced expression, perturbation analysis have potential to reveal genes with high perturbation efficiency more widsespread and dispersed in the network.
+
 # \textbf{CONCLUSION}

### 2019-12-20 Supplementary_figures.Rmd
Add highly perturbing elements to supp figure

@@
 ```
+
+## Highly effective perturbing elements in network
+
+![Interactions of genes that have high perturbation efficiency.
+Critical gene, represents crucial genes for breast cancer,provided by eDGAR; TFs, transcription factors;  Regulated by TFs, regulated genes with these transcription factors; Physical interaction with critical genes; shared genes in physical interactions between critical genes ](highly_effective_genes.png){#figs10}

### 2019-12-20 NAR_form_article.Rmd
add refer to Figure S10

@@
 The genes with high perturbation efficiency were analysed via eDGAR, a database of Disease-Gene Associations with annotated relationships among genes [@babbi2017edgar] that supplies different observations about crucial genes such as direct physical interactions, other shared genes with these interactions, protein-protein interactions, GO terms etc.
-With our model, we obtained 3 (TP53, BRCA1, AKT1) of 27 critical breast cancer genes and these genes show direct physical interaction with 21 genes.
+With our model, we obtained 3 (TP53, BRCA1, AKT1) of 27 critical breast cancer genes and these genes show direct physical interaction with 21 genes (Figure S10).
 33 shared genes which are observed on these direct interactions were also found in our effective perturbation elements.
-We have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4, GATA3, CTBP1) transcription factors (TFs) that are common in 44 critical TFs which are provided by eDGAR dataset.
+We have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4, GATA3, CTBP1) transcription factors (TFs) that are common in 44 critical TFs which are provided by eDGAR dataset (Figure S10).
 44 of regulated genes via these TFs were found in our highly effective genes.

### 2019-12-27 NAR_form_article.Rmd
NAR_submission files 27 Dec 2019

@@
 author:
-  - name: Selcen Ari Yuka
+  - name: Selcen Ari
     affiliation: "1"
@@
 abstract: |
-  Competing endogenous RNA (ceRNA) regulations and crosstalk between various types of non-coding RNA in human is remarkable in means of miRNA regulation.
-Many studies have pointed out that an alteration in miRNA:target interaction can result in unexpected changes due to indirect and complex interactions.
-In this paper, we defined a new network-based model that handles miRNA:ceRNA interactions with expression values.
-Our model is able to handle miRNA interaction factors such as seed type, binding energy, if provided.
-Our approach is able to reveal that a perturbation in an element of network affects whole competing elements differently and cooperative efficiencies of miRNAs on common targets could be calculated.
-Our findings emphasized importance of miRNA:target ratios being crucial, as reported by previous studies.
-We have showed that the competing elements which have the same or close expression values may not be affected equally from the perturbation because of repression functionality depended on interaction factors of miRNA target pairs.
-We applied the model to real sample consisting of breast cancer gene and miRNA expression dataset and experimental miRNA:target interaction dataset all generated via high throughput sequencing methods.
-A gene over-expressed in tumor tissue, namely *SERPINE2*, is used as perturbing element.
-We have observed that change in expression level of single gene in miRNA:target network is sufficient to perturb regulations in whole network, due to unforeseen and unpredicted regulation which are only visible when considered in network context.
-Therefore, this model helps unveiling the crosstalk between elements in miRNA:target network where abundance of target and sponge effect are taken into account.
-The model is scalable and can be plugged in with emerging miRNA effectors such as circRNAs.
-The model is available as R package ceRNAnetsim (https://github.com/selcenari/ceRNAnetsim).
+  Competing endogenous RNA (ceRNA) regulations and crosstalk between various types of non-coding RNA in human is an important and under-explored subject.
+Several studies have pointed out that an alteration in miRNA:target interaction can result in unexpected changes due to indirect and complex interactions.
+In this paper, we defined a new network-based model that incorporates miRNA:ceRNA interactions with expression values.
+Our approach is able to calculate network-wide effects after a perturbation in expression level of a single element while utilizing miRNA interaction factors such as seed type, binding energy.
+We applied the model to existing breast cancer mRNA, miRNA expression and high throughput experimental miRNA:target interaction dataset.
+A gene over-expressed in tumor tissue, namely *SERPINE2*, was identified as a perturbing element which affect regulation of many nodes in whole network when its expression level changes.
+To conclude, this model helps unveiling the crosstalk between elements in miRNA:target network where abundance of target and sponge effect are taken into account.
+The model has potential to reveal unforeseen and unpredicted regulation which are only visible when considered in network context.
+Our tool is scalable and can be plugged in with emerging miRNA effectors such as circRNAs and available as R package ceRNAnetsim (https://github.com/selcenari/ceRNAnetsim)
 
-bibliography: NAR-sample.bib
+bibliography: NAR-sample
 output: rticles::nar_article
@@
 knitr::opts_chunk$set(echo = FALSE)
+options(tinytex.verbose = TRUE)
 ```
@@
 Briefly, Ala et al.
-have explained the ceRNA hypothesis as disturbance of the other target when one of the targets on a steady-state system that included one miRNA and two target was perturbed with expression change [@ala_integrated_2013].
+have explained the ceRNA hypothesis as disturbance of the other target when one of the targets on a steady-state system that included one miRNA
+
+\enlargethispage{-65.1pt}
+
+and two target was perturbed with expression change [@ala_integrated_2013].
 Regarding interaction between miRNAs and their target in a cell, explaining and predicting results of an individual perturbation is difficult due to complexity of interactions.
 Various computational and experimental studies have tackled the problem of unraveling ceRNA:miRNA interactions.
-
-\enlargethispage{-65.1pt}
 
@@
 
-# Genes causing widespread perturbation
+## Genes causing widespread perturbation
 
@@
 44 of regulated genes via these TFs were found in our highly effective genes.
-it has been observed that these genes cover 22.5% of 297 perturbation-effective genes.
-
-When we analysed these genes via biomaRt OMIM query, it has been observed that our model provides enriched 1.45 fold results with hypergeometric p-values 3.19e-05.
+It has been observed that these genes cover 22.5% of 297 perturbation-effective genes.
+When we analysed these genes via biomaRt OMIM query, they show 1.45 fold enrichment of OMIM genes (hypergeometric p-value 3.19e-05).
 

### 2019-12-27 NAR_form_article.tex
NAR_submission files 27 Dec 2019

@@
 % Enter dates of publication
-\copyrightyear{2008}
-\pubdate{31 July 2009}
-\pubyear{2009}
-\jvolume{37}
-\jissue{12}
+\copyrightyear{X}
+\pubdate{X}
+\pubyear{X}
+\jvolume{X}
+\jissue{X}
 
@@
 \usepackage{float} \floatplacement{figure}{H} \usepackage[colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue]{hyperref} \usepackage[T1]{fontenc}
+
+%\usepackage[scaled]{helvet}
+%\renewcommand\familydefault{\sfdefault}
 
@@
 \author{
-    Selcen Ari\,$^{1}$
-   and Alper Yilmaz\,$^{1}$ \footnote{To whom correspondence should be addressed. Phone: +90 212 383 4627 Fax:
-+90 212 383 4625 Email:
-\href{mailto:alyilmaz@yildiz.edu.tr}{\nolinkurl{alyilmaz@yildiz.edu.tr}}} 
+    Selcen Ari Yuka\,$^{1}$ \footnote{To whom correspondence should be addressed. Email: selcenay@yildiz.edu.tr}
+   and Alper Yilmaz\,$^{1}$ \footnote{Correspondence may also be addressed to Wieteke Alper Yilmaz. Phone: +90 212 383 4627 Fax: +90 212 383 4625 Email: alyilmaz@yildiz.edu.tr}
 }
@@
 \history{%
-Received January 1, 2009;
-Revised February 1, 2009;
-Accepted March 1, 2009}
+X;
+X;
+X}
 
@@
 Competing endogenous RNA (ceRNA) regulations and crosstalk between
-various types of non-coding RNA in human is remarkable in means of miRNA
-regulation. Many studies have pointed out that an alteration in
-miRNA:target interaction can result in unexpected changes due to
-indirect and complex interactions. In this paper, we defined a new
-network-based model that handles miRNA:ceRNA interactions with
-expression values. Our model is able to handle miRNA interaction factors
-such as seed type, binding energy, if provided. Our approach is able to
-reveal that a perturbation in an element of network affects whole
-competing elements differently and cooperative efficiencies of miRNAs on
-common targets could be calculated. Our findings emphasized importance
-of miRNA:target ratios being crucial, as reported by previous studies.
-We have showed that the competing elements which have the same or close
-expression values may not be affected equally from the perturbation
-because of repression functionality depended on interaction factors of
-miRNA target pairs. We applied the model to real sample consisting of
-breast cancer gene and miRNA expression dataset and experimental
-miRNA:target interaction dataset all generated via high throughput
-sequencing methods. A gene over-expressed in tumor tissue, namely
-\emph{SERPINE2}, is used as perturbing element. We have observed that
-change in expression level of single gene in miRNA:target network is
-sufficient to perturb regulations in whole network, due to unforeseen
-and unpredicted regulation which are only visible when considered in
-network context. Therefore, this model helps unveiling the crosstalk
-between elements in miRNA:target network where abundance of target and
-sponge effect are taken into account. The model is scalable and can be
-plugged in with emerging miRNA effectors such as circRNAs. The model is
-available as R package ceRNAnetsim
-(\url{https://github.com/selcenari/ceRNAnetsim}).
+various types of non-coding RNA in human is an important and
+under-explored subject. Several studies have pointed out that an
+alteration in miRNA:target interaction can result in unexpected changes
+due to indirect and complex interactions. In this paper, we defined a
+new network-based model that incorporates miRNA:ceRNA interactions with
+expression values. Our approach is able to calculate network-wide
+effects after a perturbation in expression level of a single element
+while utilizing miRNA interaction factors such as seed type, binding
+energy. We applied the model to existing breast cancer mRNA, miRNA
+expression and high throughput experimental miRNA:target interaction
+dataset. A gene over-expressed in tumor tissue, namely \emph{SERPINE2},
+was identified as a perturbing element which affect regulation of many
+nodes in whole network when its expression level changes. To conclude,
+this model helps unveiling the crosstalk between elements in
+miRNA:target network where abundance of target and sponge effect are
+taken into account. The model has potential to reveal unforeseen and
+unpredicted regulation which are only visible when considered in network
+context. Our tool is scalable and can be plugged in with emerging miRNA
+effectors such as circRNAs and available as R package ceRNAnetsim
+(\url{https://github.com/selcenari/ceRNAnetsim})
 \end{abstract}
@@
 disturbance of the other target when one of the targets on a
-steady-state system that included one miRNA and two target was perturbed
-with expression change \citep{ala_integrated_2013}. Regarding
-interaction between miRNAs and their target in a cell, explaining and
-predicting results of an individual perturbation is difficult due to
-complexity of interactions. Various computational and experimental
-studies have tackled the problem of unraveling ceRNA:miRNA interactions.
+steady-state system that included one miRNA
 
 \enlargethispage{-65.1pt}
+
+and two target was perturbed with expression change
+\citep{ala_integrated_2013}. Regarding interaction between miRNAs and
+their target in a cell, explaining and predicting results of an
+individual perturbation is difficult due to complexity of interactions.
+Various computational and experimental studies have tackled the problem
+of unraveling ceRNA:miRNA interactions.
 
@@
 
-\subsection{Simulation of large and realistic network}
+\subsection{Simulation of a large and realistic network}
 
@@
 
-\section{Additions to be fixed}
+\subsection{Genes causing widespread perturbation}
 
@@
 perturbation efficiency to determine crucial genes on the network. We
-have observed that two miRNAs and 488 genes are effective in terms of
-perturbation efficiency, when we ran our model with more than 15\%
-efficiency on target expression. miR-10b was investigated via HMDD
-database \citep{huang2018hmdd} and mostly found as metastatic biomarker.
-On the other hand, it was observed that miR-148a is critical for
-promotion of apoptosis and supression of breast cancer cell growth.
-Additionaly, to obtain the number of genes in a narrow range, the model
-was ran with 17\% efficency on target expression and 297 genes and one
-miRNA were observed as highly perturbation-effective. We analysed these
-genes via eDGAR, a database of Disease-Gene Associations with annotated
-Relationships among genes \citep{babbi2017edgar} that supplies different
-observations about crucial genes such as direct physical interactions,
-other shared genes with these interactions, protein-protein
-interactions, GO terms etc. With our model, we obtained 3 (TP53, BRCA1,
-AKT1) of 27 critical breast cancer genes and these genes show direct
-physical interaction with 21 genes. 33 shared genes which are observed
-on these direct interactions were also found in our effective
-perturbation elements.We have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4,
-GATA3, CTBP1) transcription factors (TFs) that are common in 44 critical
-TFs which are provided by eDGAR dataset. 44 of regulated genes via these
-TFs were found in our highly effective genes. it has been observed that
-these genes cover 22.5\% of 297 perturbation-effective genes.
-
-When we analysed these genes via biomaRt OMIM query, it has been
-observed that our model provides enriched 1.45 fold results with
-hypergeometric p-values 3.19e-05.
+have observed that one miRNA and 297 genes have high perturbation
+efficiency, when we ran our model with more than 17\% efficiency on
+target expression. The miRNA, miR-10b was investigated via HMDD database
+\citep{huang2018hmdd} and was found as metastatic biomarker. The genes
+with high perturbation efficiency were analysed via eDGAR, a database of
+Disease-Gene Associations with annotated relationships among genes
+\citep{babbi2017edgar} that supplies different observations about
+crucial genes such as direct physical interactions, other shared genes
+with these interactions, protein-protein interactions, GO terms etc.
+With our model, we obtained 3 (TP53, BRCA1, AKT1) of 27 critical breast
+cancer genes and these genes show direct physical interaction with 21
+genes (Figure S10). 33 shared genes which are observed on these direct
+interactions were also found in our effective perturbation elements. We
+have obtained 7 (KLF4, TP53, BRCA1, CREB, ID4, GATA3, CTBP1)
+transcription factors (TFs) that are common in 44 critical TFs which are
+provided by eDGAR dataset (Figure S10). 44 of regulated genes via these
+TFs were found in our highly effective genes. It has been observed that
+these genes cover 22.5\% of 297 perturbation-effective genes.When we 
+analysed these genes via biomaRt OMIM query, they show 1.45 fold 
+enrichment of OMIM genes (hypergeometric p-value 3.19e-05).
 
@@
 
+In breast cancer sample, all the genes with high perturbation efficiency
+were found to be all connected to miR10b. This can be explained by very
+high expression level of that miRNA. In a large network where miRNAs
+have balanced expression, perturbation analysis have potential to reveal
+genes with high perturbation efficiency more widsespread and dispersed
+in the network.
+
 \section{\texorpdfstring{\textbf{CONCLUSION}}{}}
@@
 
-\bibliographystyle{nar}
-\bibliography{NAR-sample.bib}
+\bibliographystyle{NAR-natbib}
+\bibliography{NAR-sample}
 

### 2019-12-27 Supplementary_additional_file.Rmd
NAR_submission files 27 Dec 2019

@@
 author: 
-- Selcen Ari
+- Selcen Ari Yuka
 - Alper Yilmaz 
-date: "22 09 2019"
+date: "`r Sys.Date()`"
 output:
@@
   html_document: default
-  word_document: default
-bibliography: ref.bib
+  word_document: default#
+bibliography: NAR-sample.bib
 header-includes: 
@@
 library(biomaRt)
-knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
 
@@
 
-clashelwak <- read.table("mmc1.txt", comment.char = "#", header = TRUE, skip = 1, stringsAsFactors = FALSE)
+clashelwak <- read.table("mmc1.txt", comment.char = "#", header = TRUE, 
+                         skip = 1, stringsAsFactors = FALSE)
 
@@
                     dataset='hsapiens_gene_ensembl')
-hg19 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene_id", "strand" ), mart = ensemblgrch37)
+hg19 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id",
+                             "chromosome_name", "start_position","end_position", 
+                             "hgnc_symbol", "entrezgene_id", "strand" ), mart = ensemblgrch37)
 
@@
 clashelwak <- clashelwak%>%
- separate(microRNA_name, c("Barcode", "Database", "mirna_name", "type"), sep = "_")%>%
-  separate(mRNA_name, c("Ensembl_Gene_Id", "Ensembl_Transcript_Id", "Hugo_Symbol", "mRNA_Type"), sep = "_")
+ separate(microRNA_name, c("Barcode", "Database", 
+                           "mirna_name", "type"), sep = "_")%>%
+  separate(mRNA_name, c("Ensembl_Gene_Id", "Ensembl_Transcript_Id", 
+                        "Hugo_Symbol", "mRNA_Type"), sep = "_")
 
@@
   separate(V9, c("ID", "Alias", "Name", "Precusor"), sep = ";")%>%
-  mutate(ID = substr(ID, 4, length(ID)), Alias = substr(Alias, 7, length(Alias)), Name = substr(Name, 6, length(Name)), Precusor = substr(Precusor, 14, length(Precusor)))%>%
-  dplyr::select(chr= V1, start = V4, end = V5, strand = V7, ID, Alias, Name, Precusor)->mirbasehg38
+  mutate(ID = substr(ID, 4, length(ID)), 
+         Alias = substr(Alias, 7, length(Alias)), 
+         Name = substr(Name, 6, length(Name)), 
+         Precusor = substr(Precusor, 14, length(Precusor)))%>%
+  dplyr::select(chr= V1, start = V4, end = V5, 
+                strand = V7, ID, Alias, Name, Precusor)->mirbasehg38
 
@@
   dplyr::select(mirna_ID = X2, mirbase_ID = X3)%>%
-  inner_join(mirbasehg38%>% dplyr::select(ID, Name), by= c("mirbase_ID"= "ID"))%>%
+  inner_join(mirbasehg38%>% dplyr::select(ID, Name), 
+             by= c("mirbase_ID"= "ID"))%>%
   dplyr::select(mirbase_ID, Name)%>%
@@
   inner_join(clashelwak, by = c("mirbase_ID"= "Barcode"))%>%
-  dplyr::select(Name, miRNA_seq, Ensembl_Gene_Id, Ensembl_Transcript_Id, Hugo_Symbol, mRNA_seq_extended, chimeras_decompressed, seed_type, seed_basepairs,folding_class, seq_ID, folding_energy, X5.UTR, CDS, X3.UTR)%>%
-  inner_join(hg19, by = c("Ensembl_Gene_Id"= "ensembl_gene_id", "Ensembl_Transcript_Id"="ensembl_transcript_id", "Hugo_Symbol"="hgnc_symbol"))%>%
-  mutate(region1 = ifelse(X5.UTR == "1", "5UTR", " "), region2= ifelse(X3.UTR == "1", "3UTR", " "),region3= ifelse(CDS =="1", "CDS", " "))%>%
+  dplyr::select(Name, miRNA_seq, Ensembl_Gene_Id, 
+                Ensembl_Transcript_Id, Hugo_Symbol, 
+                mRNA_seq_extended, chimeras_decompressed, 
+                seed_type, seed_basepairs,folding_class, 
+                seq_ID, folding_energy, 
+                X5.UTR, CDS, X3.UTR)%>%
+  inner_join(hg19, by = c("Ensembl_Gene_Id"= "ensembl_gene_id", 
+                          "Ensembl_Transcript_Id"="ensembl_transcript_id", 
+                          "Hugo_Symbol"="hgnc_symbol"))%>%
+  mutate(region1 = ifelse(X5.UTR == "1", "5UTR", " "), 
+         region2= ifelse(X3.UTR == "1", "3UTR", " "),
+         region3= ifelse(CDS =="1", "CDS", " "))%>%
   unite(region, c(region1, region2, region3), sep = "||")%>%
-  dplyr::select(chromosome_name, start_position, end_position,strand, Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, mRNA_seq_extended, Name, miRNA_seq, seq_ID, seed_type,seed_basepairs, folding_class, folding_energy, region)%>%
+  dplyr::select(chromosome_name, start_position, 
+                end_position, strand, Hugo_Symbol, 
+                Ensembl_Gene_Id, Ensembl_Transcript_Id, 
+                mRNA_seq_extended, Name, miRNA_seq, 
+                seq_ID, seed_type,seed_basepairs, 
+                folding_class, folding_energy, region)%>%
   as_tibble() -> clashelwakfinal
@@
   mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
-  dplyr::anti_join(lift19_del, by = c("Chromosome"="Chr", "start_position"="Start", "end_position"="End"))
+  dplyr::anti_join(lift19_del, by = c("Chromosome"="Chr", 
+                                      "start_position"="Start", 
+                                      "end_position"="End"))
 
@@
 
-clashelwakfinal%>% dplyr::select(cluster=seq_ID, chromosome = Chromosome, start_position, end_position, strand, hgnc_symbol = Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, target_seq = mRNA_seq_extended, miRNA= Name, miR_seq =miRNA_seq, seed_type, seed_type2= seed_basepairs, seed_type3= folding_class, Energy = folding_energy, HG38build_loc, Genom_build, region) -> clashelwakfinal
+clashelwakfinal%>% dplyr::select(cluster=seq_ID, chromosome = Chromosome, 
+                                 start_position, end_position, 
+                                 strand, hgnc_symbol = Hugo_Symbol, 
+                                 Ensembl_Gene_Id, Ensembl_Transcript_Id, 
+                                 target_seq = mRNA_seq_extended, miRNA= Name, 
+                                 miR_seq =miRNA_seq, seed_type, 
+                                 seed_type2= seed_basepairs, 
+                                 seed_type3= folding_class, 
+                                 Energy = folding_energy, 
+                                 HG38build_loc, 
+                                 Genom_build, region) -> clashelwakfinal
 
@@
 clashelwakfinal%>%
-  mutate(seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="I", paste0(seed_type2, "-mer"), seed_type),
-         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="II", paste0(seed_type2, "-mer_noncanonical"), seed_type), 
-         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="III", paste0(seed_type2, "-mer_noncanonical"), seed_type),
-         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="IV", paste0(seed_type2, "-mer_noncanonical"), seed_type),
+  mutate(seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="I", 
+                            paste0(seed_type2, "-mer"), seed_type),
+         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="II", 
+                            paste0(seed_type2, "-mer_noncanonical"), seed_type), 
+         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="III", 
+                            paste0(seed_type2, "-mer_noncanonical"), seed_type),
+         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="IV", 
+                            paste0(seed_type2, "-mer_noncanonical"), seed_type),
          seed_type= ifelse (startsWith(seed_type, "no") ,"none", seed_type))%>%
@@
 
-hg18 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene", "strand" ), mart = ensembl54)
+hg18 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id",
+                             "chromosome_name", "start_position","end_position", 
+                             "hgnc_symbol", "entrezgene", "strand" ), mart = ensembl54)
 ```
@@
 clearclipfinal <- hg18%>%
-  inner_join(clearclip, by= c("entrezgene"= "gene.id", "hgnc_symbol"= "gene.symbol"))%>%
+  inner_join(clearclip, by= c("entrezgene"= "gene.id", 
+                              "hgnc_symbol"= "gene.symbol"))%>%
   distinct()
@@
 clearclipfinal <- clearclipfinal%>%
-  dplyr::anti_join(deleted_lift18, by = c("chr"="Chr", "start_position"="Start", "end_position"="End"))
+  dplyr::anti_join(deleted_lift18, by = c("chr"="Chr", 
+                                          "start_position"="Start", 
+                                          "end_position"="End"))
 
-hg38clearclip<- read.delim("hg38clearclip.txt", header = FALSE, stringsAsFactors = FALSE)
+hg38clearclip<- read.delim("hg38clearclip.txt", header = FALSE, 
+                           stringsAsFactors = FALSE)
 
@@
 clearclipfinal%>%
-  dplyr::select(cluster=cluster.ID, chromosome = chr, start_position, end_position, strand = strand.y, hgnc_symbol, Ensembl_Gene_Id =ensembl_gene_id, Ensembl_Transcript_Id =ensembl_transcript_id, target_seq = target.map, miRNA, miR_seq = miR.map, seed_type = 'seed match', Energy = MFE, HG38build_loc, Genom_build, region) -> clearclipfinal
+  dplyr::select(cluster=cluster.ID, chromosome = chr, start_position, 
+                end_position, strand = strand.y, hgnc_symbol, Ensembl_Gene_Id =ensembl_gene_id, 
+                Ensembl_Transcript_Id =ensembl_transcript_id, target_seq = target.map, 
+                miRNA, miR_seq = miR.map, seed_type = 'seed match', Energy = MFE, 
+                HG38build_loc, Genom_build, region) -> clearclipfinal
 
@@
 ```{r,  eval=FALSE}
-data_frame(seed_type= c("5mer_1","5mer_2", "5mer_3", "6mer", "6mer.indel", "6mer.mm", "6mer_off.mm", "6merA1", "6merA1.indel", "6merA1.mm", "7merA1", "7merA1.indel", "7merA1.mm", "7merm8", "7merm8.indel", "7merm8,mm", "8mer", "8mer.indel", "8mer.mm", "NA" ),
-           seed_type_com= c("5-mer", "5-mer_noncanonical", "5-mer_noncanonical", "6-mer", "6-mer_noncanonical", "6-mer_noncanonical", "6-mer_noncanonical", "6-merA1", "6-merA1_noncanonical","6-merA1_noncanonical", "7-merA1", "7-merA1_noncanonical","7-merA1_noncanonical", "7-mer-8m", "7-mer-8m_noncanonical", "7-mer-8m_noncanonical", "8-mer", "8-mer_noncanonical", "8-mer_noncanonical", "none" )) ->clipdata_seed
+data_frame(seed_type= c("5mer_1","5mer_2", "5mer_3", "6mer", "6mer.indel", "6mer.mm", 
+                        "6mer_off.mm", "6merA1", "6merA1.indel", "6merA1.mm", "7merA1", 
+                        "7merA1.indel", "7merA1.mm", "7merm8", "7merm8.indel", "7merm8,mm", 
+                        "8mer", "8mer.indel", "8mer.mm", "NA" ),
+           seed_type_com= c("5-mer", "5-mer_noncanonical", "5-mer_noncanonical", "6-mer", 
+                            "6-mer_noncanonical", "6-mer_noncanonical", "6-mer_noncanonical", 
+                            "6-merA1", "6-merA1_noncanonical","6-merA1_noncanonical", "7-merA1",
+                            "7-merA1_noncanonical","7-merA1_noncanonical", "7-mer-8m", 
+                            "7-mer-8m_noncanonical", "7-mer-8m_noncanonical", "8-mer", 
+                            "8-mer_noncanonical", "8-mer_noncanonical", "none" )) ->clipdata_seed
 
@@
   inner_join(clipdata_seed, by = "seed_type")%>%
-  dplyr::select(1:11, seed_type = seed_type_com, Energy, HG38build_loc, Genom_build, region)-> clearclipfinal
+  dplyr::select(1:11, seed_type = seed_type_com, Energy, 
+                HG38build_loc, Genom_build, region)-> clearclipfinal
 
@@
 Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study.
-On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study.
-Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions.
+On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.Grimson et al.
+[-@grimson_microrna_2007] have compared the effect of seed types on target repression with few miRNA had canonical seed pairing in their study.
+Additionally, Bartel et al.
+[-@bartel_micrornas:_2009] and Betel et al.
+[-@betel2010comprehensive] have studied on functional and non-functional seed interactions.
 Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values.
-We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
+We also have redefined location of binding region on the target as numeric values, based on studies of Hausser et al.
+[-@hausser_analysis_2013] and Helwak et al.
+[-@helwak_mapping_2013].
 With this process, we have handled this integrated dataset in context of competitor behaviors and functionality of interactions.
@@
 experimentalmirnagene <- experimentalmirnagene%>%
-  mutate(region2  = str_replace_all(region, "NA", ""), region3 = str_replace_all(region2, "\\|", ""), region = str_replace_all(region3, c("3'UTR"="3UTR", "5'UTR"="5UTR")))%>%
+  mutate(region2  = str_replace_all(region, "NA", ""), 
+         region3 = str_replace_all(region2, "\\|", ""), 
+         region = str_replace_all(region3, c("3'UTR"="3UTR", "5'UTR"="5UTR")))%>%
   dplyr::select(-region2, -region3)%>%
-  mutate(region_effect = as.double(ifelse(region %in% c("3UTRCDS", "CDS3UTR", "5UTR3UTR", "CDS5UTR3UTR", "CDS3UTRintron"), "0.93", ifelse(region %in% c("CDS", "CDSintron"), "0.42", ifelse(region %in% c("3UTR","3UTRintron"), "0.84", ifelse(region %in% c("5UTR", "5UTRintron"), "0.01", ifelse(region %in% c("5UTRCDS", "CDS5UTR"), "0.42", ifelse(region %in% c("intron",""), "0.01", ifelse(region %in% c("exon_unclassified", ""), "0.2", NA)))))))))
+  mutate(region_effect = as.double(ifelse(region %in% c("3UTRCDS", "CDS3UTR", "5UTR3UTR", "CDS5UTR3UTR", "CDS3UTRintron"), 
+                                  "0.93", ifelse(region %in% c("CDS", "CDSintron"), 
+                                  "0.42", ifelse(region %in% c("3UTR","3UTRintron"), 
+                                  "0.84", ifelse(region %in% c("5UTR", "5UTRintron"), 
+                                  "0.01", ifelse(region %in% c("5UTRCDS", "CDS5UTR"),
+                                  "0.42", ifelse(region %in% c("intron",""), 
+                                  "0.01", ifelse(region %in% c("exon_unclassified", ""), "0.2", NA)))))))))
 
@@
 
-seed_type_effect <- data_frame( seed_type = c("5-mer", "5-mer_noncanonical", "6-mer", "6-mer_noncanonical", "6-merA1", "6-merA1_noncanonical", "7-mer", "7-mer_noncanonical", "7-merA1", "7-merA1_noncanonical", "7-mer-8m", "7-mer-8m_noncanonical", "8-mer", "8-mer_noncanonical", "9-mer", "9-mer_noncanonical", "none"),
-                                seed_type_effect= c(0.05, 0.04, 0.07, 0.05, 0.07, 0.05, 0.23, 0.19, 0.19, 0.16, 0.25, 0.21, 0.43, 0.35, 0.43,0.35, 0.01)
+seed_type_effect <- data_frame( seed_type = c("5-mer", "5-mer_noncanonical", 
+                                              "6-mer", "6-mer_noncanonical", 
+                                              "6-merA1", "6-merA1_noncanonical", 
+                                              "7-mer", "7-mer_noncanonical", "7-merA1", 
+                                              "7-merA1_noncanonical", "7-mer-8m", 
+                                              "7-mer-8m_noncanonical", "8-mer", 
+                                              "8-mer_noncanonical", "9-mer", 
+                                              "9-mer_noncanonical", "none"),
+                                seed_type_effect= c(0.05, 0.04, 0.07, 0.05, 
+                                                    0.07, 0.05, 0.23, 0.19, 
+                                                    0.19, 0.16, 0.25, 0.21, 
+                                                    0.43, 0.35, 0.43,0.35, 0.01)
                                 )

### 2019-12-27 Supplementary_figures.Rmd
NAR_submission files 27 Dec 2019

@@
 author: 
-- Selcen Ari
+- Selcen Ari Yuka
 - Alper Yilmaz 
-date: "17 09 2019"
+date: "`r Sys.Date()`"
 output:
@@
   word_document: default
-bibliography: ref.bib
 header-includes: 
@@
 library(stringr)
-knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
 
@@
 ```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset in Steady-state", fig.width=5, fig.height=4}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4",
-            title = "Minimal dataset in steady-state conditions")
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression)%>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4",
+            title = "Minimal dataset in\nsteady-state conditions")
 ```
@@
 ```{r, message=FALSE, warning=FALSE, fig.keep='all', fig.cap= "Gene2 Upregulation on Minimal Dataset", fig.width=5, fig.height=4}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression)%>%
   update_how("Gene2", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", 
-            title = "Gene2 Upregulation without interaction factors")
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            title = "Gene2 Upregulation\nwithout interaction factors")
 ```
@@
 ```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset with interaction factors in Steady-state", fig.width=5, fig.height=4}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", 
-            title = "Minimal dataset with interaction factors in steady-state conditions")
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            title = "Minimal dataset with interaction\nfactors in steady-state conditions")
 
@@
 ```{r, message=FALSE, warning=FALSE, fig.cap= " When Gene2 is upregulated on Minimal Dataset with interaction factors", fig.width=5, fig.height=4}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
   update_how("Gene2", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", 
-            title = "Gene2 Upregulation with interaction factors")
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            title = "Gene2 Upregulation\nwith interaction factors")
 ```
@@
 B)Spreading of perturbation on system (3th iteration)"}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
   update_how("Gene2", 2)%>%
   simulate(cycle = 1)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "A")
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            Downregulation = "blue", 
+            title = "A")
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
   update_how("Gene2", 2)%>%
   simulate(cycle = 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "B")
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            Downregulation = "blue", 
+            title = "B")
 
@@
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "A")
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            title = "A")
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
   update_how("Gene4", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "B")
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            title = "B")
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
   update_how("Gene4", 2)%>%
   simulate(cycle=1)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "C")
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            title = "C")
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
   update_how("Gene4", 2)%>%
   simulate(cycle=2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "D")
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            title = "D")
 ```
 
-## Determination of iteration for Upregulation of SERPINE2 gene from Breast cancer patient dataset
+## Determination of optimum iteration for Upregulation of SERPINE2 gene from Breast cancer patient dataset
 
@@
 as.data.frame(E9GE_mirnagenenormal)%>%
- priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
+ priming_graph(competing_count = GE_normal, 
+               miRNA_count = mirna_RPM, 
+               aff_factor = c(Energy, seed_type_effect), 
+               deg_factor = region_effect)%>%
  update_how("SERPINE2",2.75) %>%

### 2019-12-27 Supplementary_mat_met.Rmd
NAR_submission files 27 Dec 2019

@@
 author: 
-- Selcen Ari
+- Selcen Ari Yuka
 - Alper Yilmaz 
-date: "7 Nisan 2019"
+date: "`r Sys.Date()`"
 output:
@@
   word_document: default
-bibliography: ref.bib
+bibliography: NAR-sample.bib
 header-includes: 
@@
 library(readr)
-knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
+#knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
 
@@
 ```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression)
 ```
@@
 ```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression)%>%
   update_how("Gene2", 2)
@@
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression)%>%
   update_how("Gene2", 2)%>%
@@
 ```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression)%>%
   update_how("Gene2", 2)%>%
@@
 ```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)
 
@@
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
   update_how("Gene2", 2)
@@
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
   update_how("Gene2", 2)%>%
@@
 ```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
   update_how("Gene2", 2)%>%
@@
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
   update_how("Gene4", 2)%>%
@@
 ```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)%>%
   update_how("Gene4", 2)%>%
@@
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)-> sample_graph
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)-> sample_graph
 
@@
   mutate(ensembl_gene_id = rownames(.))%>%
-  dplyr::inner_join(as.data.frame(rowData(BCPGE_tumor)), by = "ensembl_gene_id")%>%
+  dplyr::inner_join(as.data.frame(rowData(BCPGE_tumor)), 
+                    by = "ensembl_gene_id")%>%
   dplyr::select(ensembl_gene_id, external_gene_name, 1)-> TCGA_E9_A1N5_tumor
@@
   mutate(ensembl_gene_id = rownames(.))%>%
-  dplyr::inner_join(as.data.frame(rowData(BCPGE_normal)), by = "ensembl_gene_id")%>%
+  dplyr::inner_join(as.data.frame(rowData(BCPGE_normal)), 
+                    by = "ensembl_gene_id")%>%
   dplyr::select(ensembl_gene_id, external_gene_name, 1)-> TCGA_E9_A1N5_normal
@@
   dplyr::select(-miRNA_region)%>%
-  dplyr::inner_join(mirbase_id_conv, by = c("mirbase_id"="ID"))%>%
-  dplyr::select(miRNA_name = Name, read_count, reads_per_million_miRNA_mapped)%>%
+  dplyr::inner_join(mirbase_id_conv, 
+                    by = c("mirbase_id"="ID"))%>%
+  dplyr::select(miRNA_name = Name, 
+                read_count, 
+                reads_per_million_miRNA_mapped)%>%
   dplyr::group_by(miRNA_name)%>%
-  mutate(read_count= sum(read_count), reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
+  mutate(read_count= sum(read_count), 
+         reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
   dplyr::ungroup()%>%
@@
   as.data.frame()%>%
-  dplyr::select(miRNA_ID, read_count, reads_per_million_miRNA_mapped, miRNA_region)%>%
+  dplyr::select(miRNA_ID, 
+                read_count, 
+                reads_per_million_miRNA_mapped, 
+                miRNA_region)%>%
   dplyr::filter(startsWith(miRNA_region, "mature"))%>%
@@
 For these reasons, we obtained the high-throughput experimental miRNA:target dataset from two studies performed by Helwak et al.
-and Moore et al.
-These steps were not handle in this file because they contain many processes.
+[-@helwak_mapping_2013] and Moore et al.
+[-@moore_mirnatarget_2015] These steps were not handle in this file because they contain many processes.
 
@@
   group_by(Ensembl_Gene_Id, miRNA)%>%
-  mutate(seed_type_effect= ifelse(seed_type_effect==max(seed_type_effect), seed_type_effect, max(seed_type_effect)), Energy = ifelse(Energy==min(Energy), Energy, min(Energy)), region_effect= ifelse(region_effect==max(region_effect), region_effect, max(region_effect)))%>%
+  mutate(seed_type_effect= ifelse(seed_type_effect==max(seed_type_effect), 
+                                  seed_type_effect, max(seed_type_effect)), 
+         Energy = ifelse(Energy==min(Energy), Energy, min(Energy)), 
+         region_effect= ifelse(region_effect==max(region_effect), 
+                               region_effect, max(region_effect)))%>%
   distinct()-> tocombine_mirnagene
@@
 BCPME_mirnanormal%>%
-  dplyr::inner_join(tocombine_mirnagene, by = c("miRNA_name"="miRNA"))%>%
-  dplyr::inner_join(TCGA_E9_A1N5_normal, by = c("Ensembl_Gene_Id"="ensembl_gene_id", "hgnc_symbol"="external_gene_name"))%>%
+  dplyr::inner_join(tocombine_mirnagene, 
+                    by = c("miRNA_name"="miRNA"))%>%
+  dplyr::inner_join(TCGA_E9_A1N5_normal, 
+                    by = c("Ensembl_Gene_Id"="ensembl_gene_id", 
+                           "hgnc_symbol"="external_gene_name"))%>%
   distinct()%>%
-  dplyr::select(hgnc_symbol, miRNA_name, mirna_RPM= reads_per_million_miRNA_mapped, GE_normal, Energy, seed_type_effect, region_effect)-> E9GE_mirnagenenormal
+  dplyr::select(hgnc_symbol, 
+                miRNA_name, 
+                mirna_RPM= reads_per_million_miRNA_mapped, 
+                GE_normal, 
+                Energy, 
+                seed_type_effect, 
+                region_effect)-> E9GE_mirnagenenormal
 
@@
   dplyr::inner_join(tocombine_mirnagene, by = c("miRNA_name"="miRNA"))%>%
-  dplyr::inner_join(TCGA_E9_A1N5_tumor, by = c("Ensembl_Gene_Id"="ensembl_gene_id", "hgnc_symbol"="external_gene_name"))%>%
+  dplyr::inner_join(TCGA_E9_A1N5_tumor, 
+                    by = c("Ensembl_Gene_Id"="ensembl_gene_id", 
+                           "hgnc_symbol"="external_gene_name"))%>%
   distinct()%>%
-  dplyr::select(hgnc_symbol, miRNA_name, mirna_RPM= reads_per_million_miRNA_mapped, GE_tumor, Energy, seed_type_effect, region_effect)-> E9GE_mirnagenetumor
+  dplyr::select(hgnc_symbol, 
+                miRNA_name, 
+                mirna_RPM= reads_per_million_miRNA_mapped, 
+                GE_tumor, 
+                Energy, 
+                seed_type_effect, 
+                region_effect)-> E9GE_mirnagenetumor
 
@@
 as_data_frame(E9GE_mirnagenenormal)%>%
-  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
+  priming_graph(competing_count = GE_normal, 
+                miRNA_count = mirna_RPM, 
+                aff_factor = c(Energy, seed_type_effect), 
+                deg_factor = region_effect)%>%
   find_node_perturbation(how = 3, cycle = 10, fast = 5)%>%
@@
 as.data.frame(E9GE_mirnagenenormal)%>%
- priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
+ priming_graph(competing_count = GE_normal, 
+               miRNA_count = mirna_RPM, 
+               aff_factor = c(Energy, seed_type_effect), 
+               deg_factor = region_effect)%>%
  update_how("SERPINE2",2.75)%>%
@@
 as.data.frame(E9GE_mirnagenenormal)%>%
-  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
+  priming_graph(competing_count = GE_normal, 
+                miRNA_count = mirna_RPM, 
+                aff_factor = c(Energy, seed_type_effect), 
+                deg_factor = region_effect)%>%
   update_how("SERPINE2",2.75)%>%
@@
   dplyr::select(hgnc_symbol, GE_tumor)%>%
-  dplyr::inner_join((E9GE_mirnagenenormal%>%dplyr::select(hgnc_symbol, GE_normal)), by ="hgnc_symbol")%>%
-  inner_join(simulation_results, by= c("hgnc_symbol"="name"))%>%
+  dplyr::inner_join((E9GE_mirnagenenormal%>%dplyr::select(hgnc_symbol, GE_normal)), 
+                    by ="hgnc_symbol")%>%
+  inner_join(simulation_results, 
+             by= c("hgnc_symbol"="name"))%>%
   distinct()
@@
 This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and coherent miRNA:target regulation approach.
+
+# REFERENCES

### 2019-12-27 Supplementary_tables.Rmd
NAR_submission files 27 Dec 2019

@@
 author: 
-- Selcen Ari
+- Selcen Ari Yuka
 - Alper Yilmaz 
-date: "16 09 2019"
-output:
-  pdf_document: default
-  html_document: default
-  word_document: default
-bibliography: ref.bib
+date: "`r Sys.Date()`"
+output: pdf_document 
+bibliography: NAR-sample.bib
 header-includes: 
@@
 library(xtable)
-knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
+knitr::opts_chunk$set(tidy.opts=list(width.cutoff=30),tidy=TRUE)
 library(gt)
@@
 Table: *minsamp* sample dataset that includes interaction factors. The parameters which affect miRNA:target interactions (i.e. seed type, region, energy) are provided in sample dataset.
-
-Some of information about miRNA:target interactions were exhibited directly by high-throughput studies. On the other haned, we were examined other interacion factors based on different studies.
-
-+ [@helwak_mapping_2013; @moore_mirnatarget_2015] reported the energy values in miRNA:target interactions.
-+ Comparisons of canonical seed types were evaluated by study of [@grimson_microrna_2007], while functional and non-functional seed interactions were studied by [@bartel_micrornas:_2009] and [@betel2010comprehensive].
-+ Numeric definition of target region location effect was performed based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013]
 
@@
   5 & 5UTRCDS & 0.42 \\ 
-  6 &  & 0.01 \\ 
+  6 & none & 0.01 \\ 
   7 & intron & 0.01 \\ 
@@
 
-|Hugo_Symbol| miRNA_name | mirna_RPM | GE_normal | energy | seed_type_effect | region_effect |
+Some of information about miRNA:target interactions were exhibited directly by high-throughput studies. On the other haned, we were examined other interacion factors based on different studies.
+
++ Helwak et al. [-@helwak_mapping_2013] and Moore et al.[- @moore_mirnatarget_2015] reported the energy values in miRNA:target interactions.
++ Comparisons of canonical seed types were evaluated by study of Grimson et al.
+[-@grimson_microrna_2007], while functional and non-functional seed interactions were studied by Bartel et al.
+[-@bartel_micrornas:_2009] and Betel et al.
+[-@betel2010comprehensive] (Table S2).
++ Numeric definition of target region location effect was performed based on studies of Hausser et al. [-@hausser_analysis_2013] and Helwak et al. [-@helwak_mapping_2013] (Table S3).
+
+|Hugo Symbol| miRNA name | mirna RPM | GE normal | energy | seed type effect | region effect |
 |-----------|------------|----------|----------------|---------|-----------|----------|
@@
 Table: Example of E9GE_mirnagenenormal dataset.
+Hugo Symbol, gene symbol; miRNA name, name of miRNA that interacted with specified gene; mirna RPM; mirna count in read per million; seed type and region effect, numeric seed type and region coefficients for specified miRNA:gene interactions based on Table S2 and S3.
 
+\newpage
  \begin{table}[ht]
 \centering
-\caption{ miRNA:target pairs supported by High-throughput Experiments}
+\caption{ Context of miRNA:target pairs dataset supported by High-throughput Experiments}
 \begin{tabular}{lll}
@@
 \centering
-\caption{ The context graph object during the process.}
+\caption{ The context of graph object during the process.}
 \begin{tabularx}{\textwidth}{llX}
@@
 
-## REFERENCES
+# REFERENCES

### 2019-12-27 NAR_form_article.Rmd

@@
 ## Genes causing widespread perturbation 
-
 Additionally, miRNA:gene expression dataset which was integrated via miRTarBase [@chou2017mirtarbase] was analysed in mean of perturbation efficiency to determine crucial genes on the network.

### 2020-08-14 Supplementary_fig_tab.Rmd
article current files

---
title: "Supplementary Figures and Tables"
author: 
- Selcen Ari
- Alper Yilmaz 
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
bibliography: NAR-sample.bib
header-includes: 
  \usepackage{tabularx}
  \usepackage{float}
  \floatplacement{figure}{H}
  \usepackage{caption}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---
\tableofcontents

\beginsupplement

# 1. Supplementary Figures

![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow1.png){#figs1}

```{r, include=FALSE, message=FALSE, warning=FALSE}

library(tidygraph)
library(dplyr)
library(tidyr)
library(stringr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
#knitr::opts_chunk$set(fig.pos = 'H')
knitr::opts_chunk$set(cache=FALSE)

```

```{r, message=FALSE, warning=FALSE}
library(ceRNAnetsim)
```

## 1.1 *minsamp* dataset analysis in absence of interaction factors.

```{r}
data("minsamp")
```

```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset in Steady-state", fig.width=5, fig.height=4}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4",
            title = " ")
```

```{r, message=FALSE, warning=FALSE, fig.keep='all', fig.cap= "Gene2 Upregulation on Minimal Dataset", fig.width=5, fig.height=4}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
  update_how("Gene2", 2)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", 
            title = " ")
```

```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of minsamp data. A) First response of system to Gene2 upregulation (2nd iteration). B) Spreading of perturbation on system (3th iteration)", fig.subcap= NULL}

priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression) %>%
  update_how("Gene2", 2) %>%
  simulate(cycle=1) %>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            Downregulation = "blue", 
            title = "A")

priming_graph(minsamp, 
              competing_count = Competing_expression, 
              miRNA_count = miRNA_expression)%>%
  update_how("Gene2", 2)%>%
  simulate(2)%>%
  vis_graph(Competing_color = "navajowhite3", 
            mirna_color = "ivory4", 
            Upregulation = "red", 
            Downregulation = "blue", 
            title = "B")

```

## 1.2 *minsamp* dataset analysis with interaction factors.

```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset with interaction factors in Steady-state", fig.width=5, fig.height=4}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", 
            title = " ")

```

```{r, message=FALSE, warning=FALSE, fig.cap= " When Gene2 is upregulated on Minimal Dataset with interaction factors", fig.width=5, fig.height=4}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_how("Gene2", 2)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", 
            title = " ")
```

```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of minsamp data with interaction factors A)First response of system to Gene2 upregulation (2nd iteration). B)Spreading of perturbation on system (3th iteration)"}
priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_how("Gene2", 2)%>%
  simulate(cycle = 1)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "A")

priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_how("Gene2", 2)%>%
  simulate(cycle = 2)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "B")

```

## 1.3 Common target perturbation in *minsamp* dataset.

```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Perturbation of Gene4 on minsamp data with interaction factors. A) Network at steady-state. B) Upregulation of Gene4. C) Primary response of network to upregulation of Gene4. D) Re-regulation of whole nodes on system (3th iteration)"}

priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "A")

priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_how("Gene4", 2)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "B")

priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_how("Gene4", 2)%>%
  simulate(cycle=1)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "C")

priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
  update_how("Gene4", 2)%>%
  simulate(cycle=2)%>%
  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "D")
```

![Gene analysis of transcription factor association and KEGG pathway, obtained from experimental miRNA:target dataset ](experimental_kegg.png){#figs9}

# 2 Supplementary Tables

```{r, eval=FALSE, include=FALSE}
library(xtable)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
library(gt)
knitr::opts_chunk$set(cache=FALSE)

```

## 2.1 *minsamp* dataset

|competing| miRNA | Competing_expression | miRNA_expression | seed_type | region | energy
|-----------|------------|----------|----------------|---------|-----------|----------|
| `Gene1`   | `Mir1` | 10000 | 1000 | 0.43 | 0.30 | -20 |
| `Gene2`   | `Mir1` | 10000 | 1000 | 0.43 | 0.01 | -15 |
| `Gene3`   | `Mir1` | 5000  | 1000 | 0.32 | 0.40 | -14 |
| `Gene4`   | `Mir1` | 10000 | 1000 | 0.23 | 0.50 | -10 |
| `Gene4`   | `Mir2` | 10000 | 2000 | 0.35 | 0.90 | -12 |
| `Gene5`   | `Mir2` | 5000  | 2000 | 0.05 | 0.40 | -11 |
| `Gene6`   | `Mir2` | 10000 | 2000 | 0.01 | 0.80 | -25 |

Table: *minsamp* sample dataset that includes interaction factors. The parameters which affect miRNA:target interactions (i.e. seed type, region, energy) are provided in sample dataset.

## 2.2 Significant factors in miRNA:target interactions

Some of information about miRNA:target interactions were exhibited directly by high-throughput studies. On the other hand, we were examined other interacion factors based on different studies.

+ [@helwak_mapping_2013; @moore_mirnatarget_2015] reported the energy values in miRNA:target interactions.
+ Comparisons of canonical seed types were evaluated by study of [@grimson_microrna_2007], while functional and non-functional seed interactions were studied by [@bartel_micrornas:_2009] and [@betel2010comprehensive]. 
+ Numeric definition of target region location effect was performed based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013]

|seed type | seed type effect 
|----------|---------------|
| 6-mer_noncanonical | 0.05 | 
| 9-mer | 0.43 | 
| 6-mer | 0.07 | 
| 8-mer | 0.43 | 
| 7-mer | 0.23 | 
| none | 0.01 | 
| 5-mer_noncanonical | 0.04 | 
| 5-mer | 0.05 | 
| 6-merA1_noncanonical | 0.05 | 
| 7-mer-8m_noncanonical | 0.21 | 
| 7-mer-8m | 0.25 | 
| 8-mer_noncanonical | 0.35 | 
| 7-merA1_noncanonical | 0.16 | 
| 7-merA1 | 0.19 | 
| 6-merA1 | 0.07 |

Table: Efficiency factors for seed types.

| region | region effect
|---------|-------------|
| 3UTR | 0.84 |
| CDS | 0.42 | 
| 3UTRCDS | 0.93 | 
| 5UTR | 0.01 |
| 5UTRCDS | 0.42 | 
| none  | 0.01 | 
| intron | 0.01 | 
| CDS3UTR | 0.93 | 
| CDS5UTR | 0.42 | 
| exon_unclassified | 0.20 | 
| CDS3UTRintron | 0.93 | 
| 3UTRintron | 0.84 | 
| CDSintron | 0.42 | 
| 5UTRintron | 0.01 | 
| 5UTR3UTR | 0.93 | 
| CDS5UTR3UTR | 0.93 |

Table: Efficiency factors for binding regions on targets

## 2.3 Content of High-throughput experimental studies

|  Variable | Structure | Means 
|-----------|-----------|-------|
|  cluster    | character | Barcode from experimentally method |
|  chromosome    | character | Chromosome of Target gene from raw data |
|   start_position    | numeric | Gene start position from raw data |
|   end_position     | numeric | Gene end position from raw data |
|   strand     | character | Gene strand |
|   hgnc_symbol     | character | Gene name (Symbol) |
|   Ensembl_Gene_Id     | character | Ensembl Gene Id of gene |
|   Ensembl_Transcript_Id   | character | Ensembl transcript id of mRNA of Target gene |
|   target_seq   | character | mRNA sequences targeted by miRNA |
|   miRNA   | character | miRNA id (from miRBase version 21 ) |
|   miR_seq   | character | miRNA sequence |
|   seed_type   | character | seed type of miRNA:target interaction |
|   Energy   | numeric | Energy of miRNA:target binding |
|   HG38build_loc   | character | Recent chromosomal location of Gene |
|   Genome_build   | character| Genome build of given chromosome, start and end positions |
|   region   | character | interaction location on target | 
|   region_effect   | numeric | Coefficient of location on target | 
|   seed_type_effect   | numeric | Coefficient for seed sequence of miRNA:target interaction |

Table: miRNA:target pairs supported by High-throughput Experiments

## 2.4  Variables of network object during simulation

As a result of simulation a dataset, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at following.

 \begin{table}[ht]
\centering
\caption{ The context graph object during the process.}
\begin{tabularx}{\textwidth}{llX}
  \hline
 Variables & Structure & Means \\
  \hline
   \emph{Node Variables} & & \\
   name    & character & node name \\
   type    & character &Competing or miRNA \\
   node\_id    & numeric & in on graph object \\
   initial\_count     & numeric & Initial Expression value of node \\
   count\_pre     & numeric & Expression value of node at previous regulation \\
   count\_current     & numeric & Existing expression value of node \\
   changes\_variable     & character & Regulation of node (Up, down or steady) \\
   \hline
   \emph{Edge Variables} & & \\
   Competing name   & character & name of genes\\
   miRNA name   & character & name of miRNAs \\
   Competing expression   & numeric & Expression values of competing elements at steady-state \\
   miRNA expression   & numeric & Expression values of miRNA elements at steady-state \\
   energy   & numeric & coefficient of miRNA:target interactions (binding affinity) \\
   seed type   & numeric & coefficient of miRNA:target interactions (binding affinity) \\
   region   & numeric & coefficient of miRNA:target interactions (degradation efficiency) \\
   afff factor   & numeric& coefficient scaled and combined affinity factor \\
   degg factor   & numeric &coefficient scaled and combined degradation factor \\ 
   comp\_count\_list   & list & list of competing expression for each iteration \\ 
   comp\_count   & numeric & pre: competing expression at previous iteration; current: competing expression at present iteration \\
   mirna\_count\_list   & list & list of miRNA expression for each iteration \\
   mirna\_count   & numeric & pre: miRNA expression at previous iteration; current: miRNA expression at present iteration \\
   effect   & numeric & pre: total miRNA reppressive effect on individual target at previous iteration ; current: miRNA reppressive effect on individual   target at present iteration \\
   effect\_list   & list & list of miRNA reppressive effect on individual target for each iteration \\
  \hline
\end{tabularx}
\end{table}

\newpage

# REFERENCES

### 2020-08-14 recovering1.Rmd
article current files

---
title: "First part"
output: html_notebook
---
```{r}
#install.packages("dplyr")
#install.packages("tidyverse")
#install.packages("tidygraph")
#install.packages("janitor")
library(janitor)
library(tidyverse)
library(purrr)
library(readr)

```

```{r, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("ceRNAnetsim")
```

```{r}
#BiocManager::install("TCGAbiolinks")
#BiocManager::install("biomaRt")
library(biomaRt)
library(TCGAbiolinks)
library(ceRNAnetsim)
library(dplyr)
```

#Firstly, experimental miRNA:gene pairs analysis

```{r}
list.files("important_ceRNA_per_patient/") # Bu bizim ilk truba analizimiz.
```

```{r}

list.files("experimental_lim_1") # tüm hastalar için experimental lim=1 olarak çalıştırdık.

list.files("experimental_lim_2_50/") # Bu ise seksen yedi hastanın 2-50 aralığındaki sonuçlarını içeren data

```

```{r}
readr::read_tsv("experimental_lim_1/TCGA-3C-AAAU-01A_perturbation.tsv")%>%
  #dplyr::filter(!is.na(perturbation_efficiency))%>%
  #filter(startsWith(name, "ENSG"))->test
  as_tibble()%>%
  dplyr::select(name)->nodes_experimental

#örnek olarak 366 genden bahsediyoruz.
nodes_experimental%>%
  filter(startsWith(name, "hsa"))
```

```{r}
list.files("experimental_lim_1")%>%
  as_tibble()%>%
  mutate(patient_ids= substr(value, 0, nchar(value)-17))%>%
  dplyr::select(patient_ids)%>%
  filter(str_detect(patient_ids, "-11A"))-> normal_tissues

substr(normal_tissues$patient_ids,0, nchar(normal_tissues$patient_ids)-4)

list.files("experimental_lim_1")%>%
  as_tibble()%>%
  mutate(patient_ids= substr(value, 0, nchar(value)-17))%>%
  dplyr::select(patient_ids)%>%
  filter(str_detect(patient_ids, "-01"))%>% #tumor
  mutate(patient_barcode = substr(patient_ids,0, nchar(patient_ids)-4))%>%
  filter(patient_barcode %in% substr(normal_tissues$patient_ids,0, nchar(normal_tissues$patient_ids)-4))%>%
  filter(endsWith(patient_ids, "01A"))%>%
  dplyr::select(patient_ids)->tumor_tissue #üc tane direkt normal doku var onları elemek için aşağıdakikomutları yazdım

normal_tissues%>%
  filter(substr(patient_ids,0, nchar(patient_ids)-4) %in% substr(tumor_tissue$patient_ids,0, nchar(tumor_tissue$patient_ids)-4))%>%
  bind_rows(tumor_tissue)-> matched_patients

write.csv(matched_patients, "matched_patients.csv")

```

```{r}
for ( i in list.files("experimental_lim_1")){

  file <- as_tibble(readr::read_tsv(paste0("experimental_lim_1/", i)))

  colnames(file)[2] <- paste0(substr(i, 0, nchar(i)-16), "PE")
  colnames(file)[3] <- paste0(substr(i, 0, nchar(i)-16), "PC")

  nodes_experimental <- nodes_experimental%>%
    left_join(file, by = "name")

}

nodes_experimental
```

#Tüm hasta sonuçları

```{r}
nodes_experimental%>%
  pivot_longer(cols = contains("TCGA"), names_to = "situation", values_to = "values")%>%
  separate(situation, into = c("patient_id", "analysis_situ"), sep = "_")%>%
  mutate(analysis_situ = ifelse(analysis_situ == "PE", "perturbation_efficiency", "perturbed_node_count"))-> experimental_res_lim1
```

#87 hastanın sonuçları

```{r}
experimental_res_lim1%>%
  filter(patient_id %in% matched_patients$patient_ids)-> experimental_res_lim1
```

```{r}
experimental_res_lim1%>%
  mutate(tissue_type = ifelse(endsWith(patient_id, "11A"), "Normal", "Tumor"))%>%
  filter(analysis_situ == "perturbation_efficiency")%>%
  group_by(name, tissue_type)%>%
  summarise(effective_sample_count=sum(!is.na(values)))%>%
  pivot_wider(id_cols = name, names_from = tissue_type, values_from = effective_sample_count)%>%
  mutate_each(funs(./87))

#durumu kıyaslayabilmek için üstteki komutu yazdım.

#Alttaki de spesifikliğini belli edebilmek için

experimental_res_lim1%>%
  mutate(tissue_type = ifelse(endsWith(patient_id, "11A"), "normal", "tumor"))%>%
  filter(analysis_situ == "perturbation_efficiency")%>%
  group_by(name, tissue_type)%>%
  summarise(effective_sample_count=sum(!is.na(values)))%>%
  mutate(situ = ifelse(effective_sample_count>50, paste0(tissue_type, "_effective"), "non"))%>%
  mutate(situ = ifelse(situ== "tumor_effective" & situ== "normal_effective", "effective_inboth", situ))

```

```{r}
experimental_res_lim1%>%
  filter(analysis_situ== "perturbed_node_count")%>%
  mutate(tissue_type = ifelse(endsWith(patient_id, "11A"), "Normal", "Tumor"))%>%
  filter(tissue_type == "Tumor")->tumor_perturbed_count

experimental_res_lim1%>%
  filter(analysis_situ== "perturbed_node_count")%>%
  mutate(tissue_type = ifelse(endsWith(patient_id, "11A"), "Normal", "Tumor"))%>%
  filter(tissue_type == "Normal")->normal_perturbed_count
```

#Venn diagram

```{r}
tumor_perturbed_count%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  select(name)%>%
  filter(!startsWith(name, "ENS"))%>%
  pull()->lim1_tumor_mirna
#308 gen var 54 mirna

normal_perturbed_count%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  select(name)%>%
  filter(!startsWith(name, "ENS"))%>%
  pull()-> lim1_normal_mirna
```

```{r}
#install.packages("VennDiagram")
#install.packages("RColorBrewer")
library(VennDiagram)
library(RColorBrewer)

myCol <- brewer.pal(4, "Pastel2")
m_plot1 <- venn.diagram(
  x = list(lim1_tumor_gene,lim1_normal_gene, lim1_tumor_mirna, lim1_normal_mirna),
  category.names = c("Genes in tumor" , "Genes in normal " , "miRNAs in tumor", "miRNAs in normal"),
  filename = NULL,

  output=TRUE,

  lwd = 2,
  lty = 'blank',
  fill = myCol,

  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.pos = c(-57, -27, 25, 57),
  cat.dist = c(0.055, 0.055, 0.085, 0.085),
  #cat.col = myCol,
  cat.fontfamily = "Helvetica",
  main.fontfamily ="Helvetica",
  fontfamily = "Helvetica"
)

ggsave(my_plot1, filename = 'lim1_perturbation_efficiency.png', width = 7, height = 6, device=cairo_ps)
```

```{r}
tumor_perturbed_count%>%
  mutate(patient_id = substr(patient_id, 0, nchar(patient_id)-4))%>%
  left_join(mutate(normal_perturbed_count, patient_id = substr(patient_id, 0, nchar(patient_id)-4)), by = c("name", "patient_id", "analysis_situ"), suffix = c("_tumor", "_normal"))%>%
  dplyr::select(-contains("tissue_type"))%>%
  mutate(values_tumor = ifelse(is.na(values_tumor), 0, values_tumor), values_normal = ifelse(is.na(values_normal), 0, values_normal))%>%
  mutate(FC= log2((values_tumor+1)/(values_normal+1)))
```

# lim 5 ile de yapalım: Heatmap'e gitmek için. OLMADIIIII

```{r}
readr::read_tsv("experimental_lim_2_50/TCGA-3C-AAAU-01A_5_perturbation.tsv")%>%
  #dplyr::filter(!is.na(perturbation_efficiency))%>%
  #filter(startsWith(name, "ENSG"))->test
  as_tibble()%>%
  dplyr::select(name)->nodes_experimental

#örnek olarak 366 genden bahsediyoruz.
nodes_experimental%>%
  filter(startsWith(name, "hsa"))

list.files("experimental_lim_2_50")%>%
  as_tibble()%>%
  filter(str_detect(value, "_2_"))
```

```{r}
for ( i in list.files("experimental_lim_2_50")){

  file <- as_tibble(readr::read_tsv(paste0("experimental_lim_2_50/", i)))

  colnames(file)[2] <- paste0(substr(i, 0, nchar(i)-16), "PE")
  colnames(file)[3] <- paste0(substr(i, 0, nchar(i)-16), "PC")

    nodes_experimental <- nodes_experimental%>%
     left_join(file, by = "name")
}

nodes_experimental
```

#heat_map yapamadım. Neden dersek biz burada bir elli tümör dokusu için analiz yapmışız???? Neden????

Bu nedenle aslında ben FC kıyaslaması yapamıyorum. Yani bu olmuyor.

```{r}
nodes_experimental%>%
  dplyr::select(name, contains("_10_"))%>%
  tidyr::pivot_longer(cols = contains("TCGA"), names_to = "situation", values_to = "values")%>%
  tidyr::separate(situation, into = c("patient_id", "analysis_situ"), sep = "_")%>%
  mutate(analysis_situ = ifelse(analysis_situ == "PE", "perturbation_efficiency", "perturbed_node_count"))%>%
  dplyr::select(patient_id)%>%
  distinct()
```

```{r}
BiocManager::install("RDAVIDWebService")
library(RDAVIDWebService)
```

```{r}

bind_rows(tumor_perturbed_count%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  filter(startsWith(name, "ENS"))%>%
  mutate(rank = n/87, tissue= "tumor")%>%
    ungroup(),
  normal_perturbed_count%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  filter(startsWith(name, "ENS"))%>%
  mutate(rank = n/87, tissue= "normal")%>%
    ungroup())%>%
  dplyr::select(-n)%>%
  group_by(name)%>%
  count(name)->speciality

bind_rows(tumor_perturbed_count%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  filter(startsWith(name, "ENS"))%>%
  mutate(rank = n/87, tissue= "tumor")%>%
    ungroup(),
  normal_perturbed_count%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  filter(startsWith(name, "ENS"))%>%
  mutate(rank = n/87, tissue= "normal")%>%
    ungroup())%>%
  left_join(speciality, by ="name")%>%
  unite(type, n.y, tissue)%>%
  mutate(type = ifelse(startsWith(type, "2"), "common_both", type), 
         type = ifelse(startsWith(type, "1"), paste(substr(type, 3, nchar(type)), "_specific"), type))%>%
  dplyr::rename("sample_count"= n.x)%>%
  mutate(name = str_trim(name, side="both"))->genes_lim1                                     #-> special_nodes_experimental

genes_lim1%>%
  select(name)%>%
  distinct()%>%
  left_join(hg19, by =c("name"="ensembl_gene_id"))->genes_lim1

```

```{r}
bind_rows(tumor_perturbed_count%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  filter(startsWith(name, "hsa"))%>%
  mutate(rank = n/87, tissue= "tumor")%>%
    ungroup(),
  normal_perturbed_count%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  filter(startsWith(name, "hsa"))%>%
  mutate(rank = n/87, tissue= "normal")%>%
    ungroup())%>%
  dplyr::select(-n)%>%
  group_by(name)%>%
  count(name)->speciality

bind_rows(tumor_perturbed_count%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  filter(startsWith(name, "hsa"))%>%
  mutate(rank = n/87, tissue= "tumor")%>%
    ungroup(),
  normal_perturbed_count%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  filter(startsWith(name, "hsa"))%>%
  mutate(rank = n/87, tissue= "normal")%>%
    ungroup())%>%
  left_join(speciality, by ="name")%>%
  unite(type, n.y, tissue)%>%
  mutate(type = ifelse(startsWith(type, "2"), "common_both", type), 
         type = ifelse(startsWith(type, "1"), paste(substr(type, 3, nchar(type)), "_specific"), type))%>%
  dplyr::rename("sample_count"= n.x)%>%
  mutate(name = str_trim(name, side="both"))-> mirnas_lim1

```

# Experimental results network: ÇOK BÜYÜK.

```{r}
readRDS("important_ceRNA_per_patient/tocombinepairs.RDS")->tocombinepairs
readRDS("important_ceRNA_per_patient/gene_exp_data.RDS")-> gene_exp_data
readRDS("important_ceRNA_per_patient/mirna_exp_data.RDS")-> mirna_exp_data
readRDS("important_ceRNA_per_patient/general_BRCA_ceRNAs.RDS")-> general_BRCA_ceRNAs

colnames(gene_exp_data)[2] # bir hasta seçtim.

tocombinepairs%>%
  right_join(select(gene_exp_data, Ensembl_Gene_Id, 'TCGA-3C-AAAU-01A'), by="Ensembl_Gene_Id")%>%
  right_join(select(mirna_exp_data, miRNA, 'TCGA-3C-AAAU-01A'), by ="miRNA", suffix= c("Gene_expression", "miRNA_expression"))%>%
  select(Ensembl_Gene_Id, miRNA, Gene_expression= paste0('TCGA-3C-AAAU-01A', "Gene_expression"), miRNA_expression = paste0('TCGA-3C-AAAU-01A', "miRNA_expression"), Energy, seed_type_effect, region_effect)%>%
  filter(Ensembl_Gene_Id %in% general_BRCA_ceRNAs$geneA, Ensembl_Gene_Id %in% special_nodes_experimental$name)%>%
  priming_graph(competing_count = Gene_expression, 
                miRNA_count = miRNA_expression, 
                aff_factor = c(Energy, seed_type_effect),
                deg_factor = region_effect)%>%
  vis_graph()

```

#David api sorunu çıkıyor. Ben de manuel yaptım. dosyalar: KEGG_experimental_lim1.txt ve Biogrid_int_experimental_lim1.txt

```{r}
genes_lim1%>%
  dplyr::select(name)%>%
  distinct()%>%
  write_tsv("david_query.tsv")
```

```{r}
read_tsv("Biogrid_int_experimental_lim1.txt")%>%
  filter(PValue < 0.05)%>%
  dplyr::select(Term, Genes)%>%
  mutate(aaa=str_split(Genes, pattern = ","))%>%
  unnest()%>%
  mutate(gene = str_trim(aaa, side="both"))%>%
  select(gene)%>%
  distinct()%>%
  left_join(hg19, by =c("gene"="ensembl_gene_id"))%>%
  mutate(interaction="biogrid protein interaction")%>%
  select(name=gene, 2,3)-> biogrid_int

#238

238/362

# Yine ribozomal protein ağırlıklı olarak gözlenmiş.

read_tsv("Biogrid_int_experimental_lim1.txt")%>%
  filter(PValue < 0.05)%>%
  filter(str_detect(Term, "ribosomal"))%>%
   dplyr::select(Term, Genes)%>%
  mutate(aaa=str_split(Genes, pattern = ","))%>%
  unnest()%>%
  dplyr::select(aaa)%>%
  distinct()

54/238

```

```{r}
read_tsv("KEGG_experimental_lim1.txt")%>%
  filter(PValue < 0.05)%>%
  dplyr::select(Term, Genes)%>%
  mutate(aaa=str_split(Genes, pattern = ","))%>%
  unnest()%>%
  dplyr::select(Term, gene=aaa)%>%
  mutate(gene = str_trim(gene, side="both"))%>%
  distinct()%>%
  separate(Term, into= c("code", "pathway_name"), sep = ":")%>%
  left_join(hg19, by =c("gene"="ensembl_gene_id"))%>%
  select(name = gene, hgnc_symbol, interaction=pathway_name)->kegg_analysis

69/362

```

# Edgar analizi

Conversion için

```{r, eval=FALSE}
#BiocManager::install("biomaRt")
library(biomaRt)
listEnsemblArchives()
listMarts(host = "http://apr2020.archive.ensembl.org")
ensembl_con = useMart(host = "http://apr2020.archive.ensembl.org", biomart = "ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl")
hg19 <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"), mart = ensembl_con)

hg19
```

```{r}

read_tsv("edgar_critical.txt")
read_delim("edgar_critical.txt", delim = "\t", col_names = "text", skip = 1)%>%
  separate(text, into = c("critical_gene", "OMIM", "Band", "assoc_dis_num", "disease_code"), sep = "\t")%>%
  mutate(critical_gene = str_trim(critical_gene, side="both"))-> edgar_critcals

```

```{r}
read_delim("string_TFs.txt", delim = "\t", col_names = "text", skip = 1)%>%
  separate(text, into = c("critical_gene", "Gene2", "Direct", "Interaction", "shared_gene_num", "shared_genes"), sep = "\t")%>%
  mutate(shared_gene = str_split(shared_genes, pattern = ","))%>%
  unnest()%>%
  select(critical_gene, Gene2, shared_gene)%>%
  distinct()%>%
  mutate(shared_gene=str_trim(shared_gene, side = "both"))->TFs

```

```{r}
genes_lim1%>% #kritik gen yok
  mutate(interaction = ifelse(hgnc_symbol %in% TFs$shared_gene, "TF associated", NA),
         interaction = ifelse(hgnc_symbol %in% TFs$Gene2, "TFs", interaction),
         interaction = ifelse(hgnc_symbol %in% edgar_critcals$critical_gene, "critical", interaction))%>%
  bind_rows(kegg_analysis)%>%
  filter(!is.na(interaction))->experimental_res

```

sankey diagram

```{r}
#install.packages("networkD3")
#install.packages("svglite")
#install.packages("htmlwidgets")
library(svglite)
library(networkD3)
library(htmlwidgets)

experimental_res%>%
  select(source = hgnc_symbol, target = interaction)%>%
  mutate(value = 1)->links

nodes <- data.frame(
  name=c(as.character(links$source), as.character(links$target)) %>% 
    unique()
)

links$IDsource <- match(links$source, nodes$name)-1 
links$IDtarget <- match(links$target, nodes$name)-1

p<- sankeyNetwork(Links = links, Nodes = nodes, Source = "IDsource", Target = "IDtarget", 
              Value = "value", NodeID = "name", fontSize = 75, nodeWidth = 50,
              fontFamily = "sans-serif", iterations = 0,  width = 5000, height = 7000, margin = list("right" = 1000))

#p <- htmlwidgets::prependContent(p, htmltools::tags$h1("Highly perturbing genes constructed from experimental miRNA:gene pairs"))

p<- onRender(p,
  '
  function(el,x) {
    d3.select(el)
      .selectAll(".node text")
      .attr("x", x.options.nodeWidth - 50)
      .attr("text-anchor", "end")
  }
  '
)

saveWidget(p, file="sankeyColor1.html")
```

#disgenet analizi

```{r}
read_tsv("disgenet_breast_car.tsv")%>%
  filter(!str_detect(Gene, "MIR"), !str_detect(Gene, "LNC"))%>%                        #6387 dan 672'sinin score 0,5
  filter(Gene %in% genes_lim1$hgnc_symbol)%>%
  filter(Score_gda>0.1)                   #25 tane 362'de 178 tanesi burada var.
mirnas_lim1
```

# HMDD mirna analizi
```{r}
mirnas_lim1%>%
  select(name)%>%
  distinct()%>%
 mutate(name= str_replace_all(name, "miR", "mir"))%>%
  mutate(name= ifelse(str_detect(name,"miR"), str_replace(name, "miR", "mir"), name))%>%
  separate(name, into =c("human", "value1", "value2", "isoform"), sep = "-")%>%
  mutate(name= str_c(human, value1, value2, sep = "-"), name = paste0(name, ";"))%>%
  select(name)%>%
  distinct()      #62'den 56'e düştü.
 # write_tsv("HMDD_query.tsv")

#47'si meme kanseri ile ilişkilendirilmiş.

readxl::read_xls("HMDD_res.xls", col_names = FALSE)%>%
  bind_rows(readxl::read_xls("HMDD_res1.xls", col_names = FALSE))%>%
  bind_rows(readxl::read_xls("HMDD_res2.xls", col_names = FALSE))%>%
  clean_names()%>%
  filter(str_detect(x3, "Breast"))->HMDD_res

# meme kanseri ilişkili olanlar:

mirnas_lim1%>%
 mutate(name= str_replace_all(name, "miR", "mir"))%>%
  mutate(name= ifelse(str_detect(name,"miR"), str_replace(name, "miR", "mir"), name))%>%
  separate(name, into =c("human", "value1", "value2", "isoform"), sep = "-")%>%
  mutate(name= str_c(human, value1, value2, sep = "-"))%>%
  select(name, ceRNAetsim_rank = rank, type)%>%
  left_join(HMDD_res, by = c("name"="x1"))%>%
  #filter(is.na(x2))%>%
  select(-x4)%>%
  distinct()
47/56

# bulunamayanlar

mirnas_lim1%>%
 mutate(name= str_replace_all(name, "miR", "mir"))%>%
  mutate(name= ifelse(str_detect(name,"miR"), str_replace(name, "miR", "mir"), name))%>%
  separate(name, into =c("human", "value1", "value2", "isoform"), sep = "-")%>%
  mutate(name= str_c(human, value1, value2, sep = "-"))%>%
  select(name, ceRNAetsim_rank = rank, type)%>%
  anti_join(HMDD_res, by = c("name"="x1"))%>%
  select(name, type)%>%
  distinct()

```

### 2020-08-14 recovery2_mirtarbase.Rmd
article current files

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(janitor)
library(tidyverse)
library(purrr)
library(readr)
library(biomaRt)
library(TCGAbiolinks)
library(ceRNAnetsim)
library(dplyr)
library(xlsx)
#install.packages("installr")
#library(installr)
#install.MikTeX(page_with_download_url = "https://miktex.org/download")
library(latexpdf)
library(tinytex)
```

```{r}
list.files("mirtarbase_lim_1/results")

read_tsv("mirtarbase_lim_1/results/TCGA-A7-A0CE-01A_1_perturbation_test.tsv")%>%
 select(name)->node_mirtarbase

#3265 gen, 581 miRNA

mirtarbasegene_mat

```

#Parsing

```{r}
for ( i in list.files("mirtarbase_lim_1/results")){

  file <- as_tibble(readr::read_tsv(paste0("mirtarbase_lim_1/results/", i)))

  colnames(file)[2] <- paste0(substr(i, 0, nchar(i)-23), "PE")
  colnames(file)[3] <- paste0(substr(i, 0, nchar(i)-23), "PC")

  node_mirtarbase <- node_mirtarbase%>%
    left_join(file, by = "name")

}

node_mirtarbase
```
#87 hasta sonuçları

```{r}
node_mirtarbase%>%
  pivot_longer(cols = contains("TCGA"), names_to = "situation", values_to = "values")%>%
  separate(situation, into = c("patient_id", "analysis_situ"), sep = "_")%>%
  mutate(analysis_situ = ifelse(analysis_situ == "PE", "perturbation_efficiency", "perturbed_node_count"))%>%
  filter(patient_id %in% matched_patients$patient_ids)-> mirtarbase_lim1_res

mirtarbase_lim1_res
```

```{r}
mirtarbase_lim1_res%>%
  pivot_wider(id_cols = c("name", "patient_id"), names_from = "analysis_situ", values_from= "values")

```

#Venn diagram

```{r}
mirtarbase_lim1_res%>%
  filter(analysis_situ== "perturbed_node_count")%>%
  mutate(tissue_type = ifelse(endsWith(patient_id, "11A"), "Normal", "Tumor"))%>%
  filter(tissue_type == "Tumor")->tumor_perturbed_count_lim1

mirtarbase_lim1_res%>%
  filter(analysis_situ== "perturbed_node_count")%>%
  mutate(tissue_type = ifelse(endsWith(patient_id, "11A"), "Normal", "Tumor"))%>%
  filter(tissue_type == "Normal")->normal_perturbed_count_lim1
```

```{r}
tumor_perturbed_count_lim1%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  select(name)%>%
  filter(!startsWith(name, "ENS"))%>%
  pull()->lim1_tumor_mirna

normal_perturbed_count_lim1%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  select(name)%>%
  filter(!startsWith(name, "ENS"))%>%
  pull()-> lim1_normal_mirna
```

```{r}
#install.packages("VennDiagram")
#install.packages("RColorBrewer")
library(VennDiagram)
library(RColorBrewer)

myCol <- brewer.pal(4, "Pastel2")
my_plot2 <- venn.diagram(
  x = list(lim1_tumor_gene,lim1_normal_gene, lim1_tumor_mirna, lim1_normal_mirna),
  category.names = c("Genes in tumor" , "Genes in normal " , "miRNAs in tumor", "miRNAs in normal"),
  filename = NULL,

  output=TRUE,

  lwd = 2,
  lty = 'blank',
  fill = myCol,

  cat.fontface = "bold",
  cat.default.pos = "outer",
  cat.pos = c(-57, -27, 25, 57),
  cat.dist = c(0.055, 0.055, 0.085, 0.085),
  #cat.col = myCol,
  cat.fontfamily = "Helvetica",
  main.fontfamily ="Helvetica",
  fontfamily = "Helvetica"
)

ggsave(my_plot2, filename = 'lim1_perturbation_efficiency_mirtarbase.eps', width = 8, height = 6, device=cairo_ps)
```

# Specifity analysis

experimental olanda mirna ve gen ayrı yapmıştım. Bunda birleştirip yaptım.  BU arkadaşların orijinalleri recovering1 dosyasında var gerekirse bakabilirsin.

```{r}
bind_rows(tumor_perturbed_count_lim1%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  #filter(startsWith(name, "ENS"))%>%
  mutate(rank = n/87, tissue= "tumor")%>%
    ungroup(),
  normal_perturbed_count_lim1%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  #filter(startsWith(name, "ENS"))%>%
  mutate(rank = n/87, tissue= "normal")%>%
    ungroup())%>%
  dplyr::select(-n)%>%
  group_by(name)%>%
  count(name)->speciality

bind_rows(tumor_perturbed_count_lim1%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  #filter(startsWith(name, "ENS"))%>%
  mutate(rank = n/87, tissue= "tumor")%>%
    ungroup(),
  normal_perturbed_count_lim1%>%
  filter(!is.na(values))%>%
  group_by(name)%>%
  count(name)%>%
  filter(n> 50)%>%
  #filter(startsWith(name, "ENS"))%>%
  mutate(rank = n/87, tissue= "normal")%>%
    ungroup())%>%
  left_join(speciality, by ="name")%>%
  unite(type, n.y, tissue)%>%
  mutate(speciality = ifelse(startsWith(type, "2"), "common_both", type), 
         speciality = ifelse(startsWith(speciality, "1"), paste(substr(type, 3, nchar(speciality)), "_specific"), speciality),
         count_intissue = paste0(substr(type, 3, nchar(type)), ": ", n.x))%>%
  select(1, count_intissue, ceRNAnetsim_rank =rank, speciality)%>%
  arrange(name)%>%
  mutate(name = str_trim(name, side="both"))-> special_nodes_mirtar_lim1

```

#hasta sayılarına göre 50 'den fazla hastada aktif olan nodelar

```{r}
mirtarbase_lim1_res%>%
  mutate(tissue_type = ifelse(endsWith(patient_id, "11A"), "Normal", "Tumor"))%>%
  filter(analysis_situ == "perturbation_efficiency")%>%
  group_by(name, tissue_type)%>%
  summarise(effective_sample_count=sum(!is.na(values)))%>%
  pivot_wider(id_cols = name, names_from = tissue_type, values_from = effective_sample_count)%>%
  #mutate_each(funs(./87))%>%
  ungroup()%>%
  filter(Normal> 50 | Tumor >50)-> high_efficient
```

# Tümör ve normal doku ayrımı

```{r}
mirtarbase_lim1_res%>%
  mutate(tissue_type = ifelse(endsWith(patient_id, "11A"), "Normal", "Tumor"))%>%
  filter(tissue_type =="Normal")%>%
  mutate(patient_id = substr(patient_id, 0, nchar(patient_id)-4))%>%
  pivot_wider(id_cols =c(name, patient_id), names_from = analysis_situ, values_from = values)->normal_tissue_lim1

mirtarbase_lim1_res%>%
  mutate(tissue_type = ifelse(endsWith(patient_id, "11A"), "Normal", "Tumor"))%>%
  filter(tissue_type =="Tumor")%>%
  mutate(patient_id = substr(patient_id, 0, nchar(patient_id)-4))%>%
  pivot_wider(id_cols =c(name, patient_id), names_from = analysis_situ, values_from = values)->tumor_tissue_lim1

```

# log2FC> 1 ve log2FC < -1

```{r}
tumor_tissue_lim1%>%
    left_join(normal_tissue_lim1, by = c("name", "patient_id"), suffix = c("_tumor", "_normal"))%>%
  mutate_at(vars(c("perturbation_efficiency_tumor", "perturbed_node_count_tumor", "perturbation_efficiency_normal", "perturbed_node_count_normal")),
             funs(as.numeric(str_replace_na(., "0"))))%>%
  filter( perturbed_node_count_tumor != 0 | perturbed_node_count_normal != 0 )%>%
    mutate(perturbed_FC = log2((perturbed_node_count_tumor+1)/(perturbed_node_count_normal+1)))%>%
  filter(name %in% high_efficient$name, startsWith(name, "hsa"))%>%
  filter(perturbed_FC> 1 | perturbed_FC < -1 )

  #-> lim1_mirna

  ggplot(aes(x = patient_id, y = name, fill = perturbed_FC))+
  geom_tile(aes(colour = perturbed_FC, fill = perturbed_FC))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0))+
  scale_colour_gradientn(colours = c("blue", "white", "red"), aesthetics = c("colour", "fill"))+
  ylab("Highly pertubing miRNA nodes")+
  xlab("Patient ids")+
  ggsave("test.svg", width = 12, height = 5)
```

#DAVID QUERY: KEGG and Protein interaction analysis.

Çok fazla gen olduğu için sadece tümör veya normal spesifik olanları analiz edeceğim.

Bu da çok saçma oluyor. Ben en iyisi FC analizine göre olanları bir analiz edeyim.

```{r}
special_nodes_mirtar_lim1%>%
  select(name,type)%>%
  filter(startsWith(name, "ENS"), str_detect(type, "specific"))%>%
  distinct()%>%
  left_join(hg19, by =c("name"="ensembl_gene_id"))->genes_lim1_toquery

genes_lim1_toquery%>%
  select(name)%>%
  write_tsv("mirtarbase_david_query.tsv")

special_nodes_mirtar_lim1%>%
  filter(startsWith(name, "ENS"), str_detect(type, "specific"))%>%
  select(name)%>%
  mutate(name= str_trim(name, side = "both"))%>%
  distinct()%>%
  write_tsv("mirtarbase_david_query.tsv")

tumor_tissue_lim1%>%
    left_join(normal_tissue_lim1, by = c("name", "patient_id"), suffix = c("_tumor", "_normal"))%>%
  mutate_at(vars(c("perturbation_efficiency_tumor", "perturbed_node_count_tumor", "perturbation_efficiency_normal", "perturbed_node_count_normal")),
             funs(as.numeric(str_replace_na(., "0"))))%>%
  filter( perturbed_node_count_tumor != 0 | perturbed_node_count_normal != 0 )%>%
    mutate(perturbed_FC = log2((perturbed_node_count_tumor+1)/(perturbed_node_count_normal+1)))%>%
  filter(name %in% high_efficient$name, !startsWith(name, "hsa"))%>%
  filter(perturbed_FC> 1 | perturbed_FC < -1 )%>%
  select(name)%>%
  filter(startsWith(name, "ENSG"))%>%
  mutate(name =str_trim(name, side ="both"))%>%
  distinct()%>%                                        # pull()-> FC_genes
  write_tsv("mirtarbase_david_query.tsv")

#Buradan bir FC_genes elde ettim. Anlamlı olsun diye. Bunların kegg ve biogrid analizini yaptım.

```

Yukarıda üç tip analiz yaptım davidden. tümör ve normal specific, FC sınırı uygulanmış tüm genler ve tüm genler.

Tümör ve normal spesifikler çok anlamsız olacaktı. FC sınırı uygulanmış ve uygulanmamış arasında çok anlamlı bir fark yoktu. ikisini de kullanabilirim. Bunlardan ben bir supplementary_data yaptım.

*** Bu nedenle log2(FC) < -1 ve log2(FC) > 1 şartıyla olanlarla analize devam ettim. 2 katı ve yarısı anlamında

#David Query at non-common nodes:

```{r}
special_nodes_mirtar_lim1%>%
  filter(str_detect(speciality, "spec"), startsWith(name, "ENS"))%>%
  select(name)%>%
  write_tsv("mirtarbase_tissue_spe_david.tsv")

# p<0.05 olan üç tane pathway buldu ????? O yüzden bunu yapmam dünya saçması.

```

# SUPPLEMENTARY DATA:

Burada annotation summary oluşturdum. 1091 gen için bunu yaptım. log2(FC) < -1 ve log2(FC) > 1 : 1091 gen kaldı.

+ Excell'in ilk sayfasında hangi node'lar tümör spesifik,  normal spesifik veya iki dokuda ortak onu verdim.
+ İkinci sayfasında ise p_value < 0.05 olan kegg datası.
+ Üçüncü sayfasında p_value < 0.05 biogrid interaction datası.

```{r}

read_tsv("kegg_mirtarbase_lim1.txt")%>%
  filter(PValue < 0.05)%>%
  clean_names()%>%
  select(term, Gene_Count = count, p_value, genes, fold_enrichment)%>%
  mutate(genes2 = str_split(genes, ", "))%>%
  unnest()%>%
  mutate(genes2 = str_trim(genes2, side = "both"))%>%
  left_join(hg19, by = c("genes2"="ensembl_gene_id"))%>%
  select(genes2)%>%
  distinct()%>%
  pull()->kegg_vec

read_tsv("biogrid_mirtarbase_lim1.txt")%>%
  filter(PValue < 0.05)%>%
  clean_names()%>%
  select(term, Gene_Count = count, p_value, genes, fold_enrichment)%>%
  mutate(genes2 = str_split(genes, ", "))%>%
  unnest()%>%
  mutate(genes2 = str_trim(genes2, side = "both"))%>%
  left_join(hg19, by = c("genes2"="ensembl_gene_id"))%>%
  select(genes2)%>%
  distinct()%>%
  pull()->biogrid_vec

special_nodes_mirtar_lim1%>%
  filter(startsWith(name, "ENSG"), name %in% FC_genes)%>%
  left_join(hg19, by =c("name"="ensembl_gene_id"))%>%
  mutate(name = str_trim(name, side = "both"))%>%
  mutate(annotation = ifelse(name %in% kegg_vec, "KEGG_PATHWAY", ""),
         annotation2 = ifelse(name %in% biogrid_vec, "Biogrid_Interaction", ""))%>%
  mutate(annotation = str_c(annotation, annotation2, sep = ","),
         annotation = ifelse(startsWith(annotation, ",") | endsWith(annotation, ",") , str_replace(annotation, ",", ""), annotation),
         annotation = str_replace(annotation, ",", ", "))%>%
  select(name, hgnc_symbol, count_intissue, ceRNAnetsim_rank, speciality, annotation)%>%
  write.xlsx("Annotation_summary_supplement.xlsx", sheetName = "Effective_Genes")
```

```{r}

read_tsv("kegg_mirtarbase_lim1.txt")%>%
  filter(PValue < 0.05)%>%
  clean_names()%>%
  select(term, Gene_Count = count, p_value, genes, fold_enrichment)%>%
  mutate(genes2 = str_split(genes, ", "))%>%
  unnest()%>%
  mutate(genes2 = str_trim(genes2, side = "both"))%>%
  left_join(hg19, by = c("genes2"="ensembl_gene_id"))%>%
  group_by(term, Gene_Count, p_value, fold_enrichment)%>%
  summarize(gene_ensembl_id = paste(genes2,collapse=", "),  hgnc_symbol=paste(hgnc_symbol,collapse=", "))%>%
  ungroup()%>%
  arrange(desc(-p_value))%>%
  write.xlsx("Annotation_summary_supplement.xlsx", sheetName = "KEGG_pathway_analysis", append = TRUE)
```

```{r}
read_tsv("biogrid_mirtarbase_lim1.txt")%>%
  filter(PValue < 0.05)%>%
  clean_names()%>%
  select(term, Gene_Count = count, p_value, genes, fold_enrichment)%>%
  mutate(genes2 = str_split(genes, ", "))%>%
  unnest()%>%
  mutate(genes2 = str_trim(genes2, side = "both"))%>%
  left_join(hg19, by = c("genes2"="ensembl_gene_id"))%>%
  group_by(term, Gene_Count, p_value, fold_enrichment)%>%
  summarize(gene_ensembl_id = paste(genes2,collapse=", "),  hgnc_symbol=paste(hgnc_symbol,collapse=", "))%>%
  ungroup()%>%
  arrange(desc(-p_value))%>%
  write.xlsx("Annotation_summary_supplement.xlsx", sheetName = "Biogrid_protein_interaction_analysis", append = TRUE)
```

# Disgenet analysis

```{r}

lim1_gene%>%
  select(name)%>%
  distinct()%>%
  left_join(hg19, by =c("name"="ensembl_gene_id"))->to_disgenet

read_tsv("disgenet_breast_car.tsv")%>%
  filter(!str_detect(Gene, "MIR"), !str_detect(Gene, "LNC"))%>%                        #6387 dan 672'sinin score 0,5
  filter(Gene %in% to_disgenet$hgnc_symbol)#%>%
  filter(Score_gda>0.1)    #score 0.1den büyük 105 adet ve 582 burada

```

#HMDD analysis

```{r}
lim1_mirna%>%
  select(name)%>%
  distinct()%>%
 mutate(name= str_replace_all(name, "miR", "mir"))%>%
  mutate(name= ifelse(str_detect(name,"miR"), str_replace(name, "miR", "mir"), name))%>%
  separate(name, into =c("human", "value1", "value2", "isoform"), sep = "-")%>%
  mutate(name= str_c(human, value1, value2, sep = "-"), name = paste0(name, ";"))%>%
  select(name)%>%
  distinct()%>%      #26'dan 24'e düştü.
  write_tsv("HMDD_query_mirtarbase.tsv")
```

```{r}
readxl::read_xls("HMDD_mirtarbase1.xls", col_names = FALSE)%>%
  bind_rows(readxl::read_xls("HMDD_mirtarbase2.xls", col_names = FALSE))%>%
  clean_names()%>%
  filter(str_detect(x3, "Breast"))->HMDD_mirtarbase

HMDD_mirtarbase%>%
  #filter(x6 == "YES")%>%
  select(x1)%>%
  distinct()
```

```{r}
22/24
```

--------------------------------------------------

#Lim =5 analizi
Önce node'ları bulalım:

```{r}
read_tsv("mirtarbase_lim_1/results/TCGA-A7-A0CE-01A_1_perturbation_test.tsv")%>%
 select(name)->node_mirtarbase
```

```{r}

list.files("mirtarbase_lim_5_20/results")%>%
  as_tibble()%>%
  filter(str_detect(value, "_5_"))%>%
  pull()->lim5_patient

for ( i in lim5_patient){

  file <- as_tibble(readr::read_tsv(paste0("mirtarbase_lim_5_20/results/", i)))

  colnames(file)[2] <- paste0(substr(i, 0, nchar(i)-23), "PE")
  colnames(file)[3] <- paste0(substr(i, 0, nchar(i)-23), "PC")

  node_mirtarbase <- node_mirtarbase%>%
    left_join(file, by = "name")

}

node_mirtarbase

```

```{r}

node_mirtarbase%>%
  pivot_longer(cols = contains("TCGA"), names_to = "situation", values_to = "values")%>%
  separate(situation, into = c("patient_id", "analysis_situ"), sep = "_")%>%
  mutate(analysis_situ = ifelse(analysis_situ == "PE", "perturbation_efficiency", "perturbed_node_count"))%>%
  filter(patient_id %in% matched_patients$patient_ids)-> mirtarbase_lim5_res

```

hasta sayılarına göre 50 'den fazla hastada aktif olan arkadaşlar

```{r}
mirtarbase_lim5_res%>%
  mutate(tissue_type = ifelse(endsWith(patient_id, "11A"), "Normal", "Tumor"))%>%
  filter(analysis_situ == "perturbation_efficiency")%>%
  group_by(name, tissue_type)%>%
  summarise(effective_sample_count=sum(!is.na(values)))%>%
  pivot_wider(id_cols = name, names_from = tissue_type, values_from = effective_sample_count)%>%
  #mutate_each(funs(./87))%>%
  ungroup()%>%
  filter(Normal> 50 | Tumor >50)-> high_efficient_lim5
```

# Lim=5 için mirna analizi

```{r}
mirtarbase_lim5_res%>%
  mutate(tissue_type = ifelse(endsWith(patient_id, "11A"), "Normal", "Tumor"))%>%
  filter(tissue_type =="Normal")%>%
  mutate(patient_id = substr(patient_id, 0, nchar(patient_id)-4))%>%
  pivot_wider(id_cols =c(name, patient_id), names_from = analysis_situ, values_from = values)->normal_tissue_lim5

mirtarbase_lim5_res%>%
  mutate(tissue_type = ifelse(endsWith(patient_id, "11A"), "Normal", "Tumor"))%>%
  filter(tissue_type =="Tumor")%>%
  mutate(patient_id = substr(patient_id, 0, nchar(patient_id)-4))%>%
  pivot_wider(id_cols =c(name, patient_id), names_from = analysis_situ, values_from = values)->tumor_tissue_lim5

```

```{r}
tumor_tissue_lim5%>%
    left_join(normal_tissue_lim5, by = c("name", "patient_id"), suffix = c("_tumor", "_normal"))%>%
  mutate_at(vars(c("perturbation_efficiency_tumor", "perturbed_node_count_tumor", "perturbation_efficiency_normal", "perturbed_node_count_normal")),
             funs(as.numeric(str_replace_na(., "0"))))%>%
  filter( perturbed_node_count_tumor != 0 | perturbed_node_count_normal != 0 )%>%
    mutate(perturbed_FC = log2((perturbed_node_count_tumor+1)/(perturbed_node_count_normal+1)))%>%
  filter(name %in% high_efficient$name, startsWith(name, "hsa"))%>%
  filter(perturbed_FC> 1 | perturbed_FC < -1 )%>%

  #->lim5_mirna

  ggplot(aes(x = patient_id, y = name, fill = perturbed_FC))+
  geom_tile(aes(colour = perturbed_FC, fill = perturbed_FC))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0))+
  scale_colour_gradientn(colours = c("blue", "white", "red"), aesthetics = c("colour", "fill"))+
  ylab("Highly pertubing miRNA nodes")+
  xlab("Patient ids")#+ggsave("test.svg", width = 12, height = 5)
```

# Lim=1 ve Lim=5 için mirna FC analizi

```{r}

lim1_mirna%>%
  full_join(lim5_mirna, by =c("name", "patient_id"), suffix= c("_lim1", "_lim5"))%>%
  select(1,2, perturbed_FC_lim1, perturbed_FC_lim5)%>%
  pivot_longer(cols = c("perturbed_FC_lim1", "perturbed_FC_lim5"), names_to = "situ", values_to = "FC_insitu")%>%
  ggplot(aes(x = patient_id, y = name, fill = FC_insitu))+
  geom_tile(aes(colour = FC_insitu, fill = FC_insitu))+
  theme_test()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0),
        plot.title = element_text(hjust = 0.5, size =20), axis.title = element_text(size =20))+
  scale_colour_gradientn(colours = c("blue", "white", "red"), aesthetics = c("colour", "fill"), na.value = "white")+
  ylab("Highly pertubing miRNA nodes")+
  xlab("Patient ids")+
  ggtitle( "miRNA perturbation efficiency comparison in the cases of lim = 1 and lim = 5")+
  facet_grid(rows = "situ")+
  ggsave("mirtarbase_mirna_lim1_lim5_comp.eps", width = 15, height = 8)

```

# Lim =1 ve Lim =5 için special gene analizi:

Special yapmasam çok korkunç oluyor. Yapsam çok saçma

```{r}
tumor_tissue_lim5%>%
    left_join(normal_tissue_lim5, by = c("name", "patient_id"), suffix = c("_tumor", "_normal"))%>%
  mutate_at(vars(c("perturbation_efficiency_tumor", "perturbed_node_count_tumor", "perturbation_efficiency_normal", "perturbed_node_count_normal")),
             funs(as.numeric(str_replace_na(., "0"))))%>%
  filter( perturbed_node_count_tumor != 0 | perturbed_node_count_normal != 0 )%>%
    mutate(perturbed_FC = log2((perturbed_node_count_tumor+1)/(perturbed_node_count_normal+1)))%>%
  filter(name %in% high_efficient$name, !startsWith(name, "hsa"))%>%
  filter(perturbed_FC> 1 | perturbed_FC < -1 )-> lim5_gene

#Aynısını lim1 için de yaptık.
#elde ettiğimiz genleri ise special olan genler olarak filtreledik.

#special node'lar lim değerine göre değişmedi.
special_nodes_mirtar_lim1%>%
  filter(str_detect(speciality, "spec"))%>%
  select(name)%>%
  distinct()%>%pull()-> gene_spec

lim1_gene%>%
  full_join(lim5_gene, by =c("name", "patient_id"), suffix= c("_lim1", "_lim5"))%>%
  select(1,2, perturbed_FC_lim1, perturbed_FC_lim5)%>%
  pivot_longer(cols = c("perturbed_FC_lim1", "perturbed_FC_lim5"), names_to = "situ", values_to = "FC_insitu")%>%
  #filter(name %in% gene_spec)%>%
  ggplot(aes(x = patient_id, y = name, fill = FC_insitu))+
  geom_tile(aes(colour = FC_insitu, fill = FC_insitu))+
  theme_test()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust=0),
        axis.text.y = element_text(angle = 0, vjust = 0, hjust=0, size = 5),
        plot.title = element_text(hjust = 0.5, size =20), axis.title = element_text(size =20))+
  scale_colour_gradientn(colours = c("blue", "white", "red"), aesthetics = c("colour", "fill"), na.value = "white")+
  ylab("Highly pertubing miRNA nodes")+
  xlab("Patient ids")+
  ggtitle( "Gene perturbation efficiency comparison in the cases of lim = 1 and lim = 5")+
  facet_grid()+
  ggsave("mirtarbase_gene_lim1_lim5_ comp.svg", width = 12, height = 40)

#güzel bir grafik olmadı görüğümüz gibi. Bunu supplementary datada vereceğim o yüzden.

```

# how 0.33 analizi

### 2020-08-14 Supplementary_mat_met.Rmd
article current files

--- 2019-10-11 Supplementary_mat_met.Rmd
@@
 - Alper Yilmaz 
-date: "7 Nisan 2019"
+date: "`r Sys.Date()`"
 output:
@@
 ---
+\tableofcontents
 
@@
 **Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficiency coefficients are calculated according to this acceptance.
-In the network that is found in steady-state conditions, the change is applied to the graph object for distribution of steady-state.
-To provide the distribution in the network the workflow offer two methods: update_variables and update_how.
+In the network that is found in steady-state conditions, the change is applied to the graph object for perturbation of steady-state.
+To provide the perturbation in the network the work-flow offer two methods: *update_variables* and *update_how*.
 The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
 The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
-
-**Updating current values of variables**: After variables updating in edge variables, these are carried to node variables.
-Current and previous values of variables are stored as node variables with update_variables function.
 
@@
 On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger.
-The functions that are used in the this study are developed with R and their source code is available in [Github repository](https://github.com/selcenari/ceRNAnetsim).
+The functions that are used in the this study are developed with R and are available in [Bioconductor](https://www.bioconductor.org/packages/release/bioc/html/ceRNAnetsim.html).
 
@@
 knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
+knitr::opts_chunk$set(warning = FALSE)
 
@@
 ```{r, message=FALSE, warning=FALSE}
-#install.packages("devtools")
-#devtools::install_github("selcenari/ceRNAnetsim")
 library(ceRNAnetsim)
@@
 ```{r}
-minsamp<-readRDS("data/minsamp.RDS")
-
+data("minsamp")
 minsamp
@@
 
-See Table S1 in Supplementary Tables file.
+See Table S1 in Supplementary Figures and Tables file.
 
@@
 
-When the graphs were examined, it was observed that behaviors were same. But, when the results were analysed in terms of expression values, the regulation differences can be observed.
+When the graphs were examined, it was observed that behaviors (i.e.
+fluctuating trend) were same.
+But, when the results were analysed in terms of expression values, the regulation differences can be observed.
 
@@
 
-The common target perturbation  (increasing to two fold at Gene4 expression in presence of interaction factors) resulted in more prominent efficiency at the same conditions (shown in following).
+The common target perturbation  (increasing to two fold at Gene4 expression in presence of interaction factors) resulted in more prominent efficiency at the same conditions (shown at following).
 
@@
 
-find_node_perturbation(sample_graph, how = 2, cycle = 3, limit = 0.1)
+find_node_perturbation(sample_graph, how = 2, cycle = 3, limit = 0.1)%>%
+  as_tibble()%>%
+  select(name, perturbation_efficiency, perturbed_count)
 
@@
 
-# 6. Obtaining breast cancer dataset and integration
+# 6. Additional data manipulation steps
 
-This section describes how to apply `ceRNAnetsim` package on a breast cancer patient miRNA:target interaction dataset. Before the approach, we obtained three datasets and combined them.
+## 6.1 Arrangement of CLASH dataset
 
-```{r, echo=FALSE, message= FALSE, warning=FALSE, eval=FALSE}
-# Install the packages which is necessary to obtain and process datasets.
-#install.packages("BiocManager")
-#BiocManager::install("SummarizedExperiment")
-#BiocManager::install("TCGAbiolinks")
-library(TCGAbiolinks)
-library(SummarizedExperiment)
+CLASH dataset was retrieved from [PubMed](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3650559/) [@helwak_mapping_2013].
+
+```{r, eval=FALSE}
+
+clashelwak <- read.table("mmc1.txt", comment.char = "#", header = TRUE, skip = 1, stringsAsFactors = FALSE)
+
+#hg19
 ```
 
-## 6.1 How to get gene expression counts of TCGA-E9-A1N5 patient.
+### Query of Human Genome 19.
 
-We have obtained the gene expression values of patient using [TCGAbiolinks](https://www.bioconductor.org/packages/release/bioc/html/TCGAbiolinks.html) package from [Bioconductor](https://www.bioconductor.org/).
-For this process, we have followed the instructions of the package.
-`TCGAbiolinks` package provides to obtain data for whole number of given barcode(s) at once.
-But, we preferred to download them separately to show datasets.
+Human genome 19 information was handled through biomaRt package.
 
-+ Obtain to gene expression counts of tumor tissue.
-
-```{r,message= FALSE, warning=FALSE, eval=FALSE}
-
-BCP_tumor <- GDCquery(project = "TCGA-BRCA",
-                  data.category = "Transcriptome Profiling",
-                  data.type = "Gene Expression Quantification", 
-                  workflow.type = "HTSeq - Counts",
-                  barcode = "TCGA-E9-A1N5-01A-11R-A14D-07")
-
-GDCdownload(BCP_tumor)
-BCPGE_tumor <- GDCprepare(BCP_tumor)
-
-as.data.frame(assay(BCPGE_tumor))%>%
-  mutate(ensembl_gene_id = rownames(.))%>%
-  dplyr::inner_join(as.data.frame(rowData(BCPGE_tumor)), by = "ensembl_gene_id")%>%
-  dplyr::select(ensembl_gene_id, external_gene_name, 1)-> TCGA_E9_A1N5_tumor
-
-colnames(TCGA_E9_A1N5_tumor)[3] <- "GE_tumor"
-
-head(TCGA_E9_A1N5_tumor)
-```
-
-+ Obtain to gene expression counts of normal tissue.
-
-```{r, message= FALSE, warning=FALSE, eval=FALSE}
-BCP_normal <- GDCquery(project = "TCGA-BRCA",
-                  data.category = "Transcriptome Profiling",
-                  data.type = "Gene Expression Quantification", 
-                  workflow.type = "HTSeq - Counts",
-                  barcode = "TCGA-E9-A1N5-11A-41R-A14D-07")
-GDCdownload(BCP_normal)
-
-BCPGE_normal <- GDCprepare(BCP_normal)
-
-as.data.frame(assay(BCPGE_normal))%>%
-  mutate(ensembl_gene_id = rownames(.))%>%
-  dplyr::inner_join(as.data.frame(rowData(BCPGE_normal)), by = "ensembl_gene_id")%>%
-  dplyr::select(ensembl_gene_id, external_gene_name, 1)-> TCGA_E9_A1N5_normal
-
-colnames(TCGA_E9_A1N5_normal)[3] <- "GE_normal"
-
-head(TCGA_E9_A1N5_normal)
+```{r,  eval=FALSE}
+#HG19
+listEnsemblArchives()
+listMarts(host = 'http://grch37.ensembl.org')
+ensemblgrch37 = useMart(host='http://grch37.ensembl.org', 
+                     biomart='ENSEMBL_MART_ENSEMBL',
+                    dataset='hsapiens_gene_ensembl')
+hg19 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene_id", "strand" ), mart = ensemblgrch37)
 
@@
 
-## 6.2 How to get miRNA expression counts of TCGA-E9-A1N5 patient.
+### Adding miRNA and gene information
 
-We have used `TCGAbiolinks` package to obtain miRNA expression quantification.
-The query gives read count of miRNA as isoform chromosome coordination.
-The data also contains mature miRNA information.
-So, we processed data to attain -5p -3p isoform information using mirbase release21 dataset.
+```{r,  eval=FALSE}
 
-+ Get the mirbase id of mature miRNA:
-
-We downloaded the mirbase release 21 dataset from [mirbase](http://www.mirbase.org/ftp.shtml) and processed the patient miRNA expression datasets as following:
-
-```{r, message= FALSE, warning=FALSE}
-
-read_tsv("hsa_mirna.txt", comment = "#", col_names = FALSE)%>%
-  dplyr::select(mirna_type= X3, definition = X9)%>%
-  filter(!endsWith(mirna_type, "primary_transcript"))%>%
-  tidyr::separate(definition, c("ID", "Alias", "Name", "Derivated"), sep = ";")%>%
-  dplyr::select(Alias, Name)%>%
-  tidyr::separate(Alias, c("trash1", "ID"), sep = "=")%>%
-  tidyr::separate(Name, c("trash2", "Name"), sep = "=")%>%
-  dplyr::select(-trash1, -trash2)-> mirbase_id_conv
-
-head(mirbase_id_conv)
-```
-
-+ Obtain the miRNA expression of tumor tissue of patient:
-
-```{r, message= FALSE, warning=FALSE, eval=FALSE}
-BCP_mirnatumor <- GDCquery(project = "TCGA-BRCA",
-                  data.category = "Transcriptome Profiling",
-                  data.type = "Isoform Expression Quantification",
-                  workflow.type = "BCGSC miRNA Profiling",
-                  barcode = "TCGA-E9-A1N5-01A-11R-A14C-13")
-
-GDCdownload(BCP_mirnatumor)
-
-GDCprepare(BCP_mirnatumor)%>%
-  as.data.frame()%>%
-  dplyr::select(miRNA_ID, read_count, reads_per_million_miRNA_mapped, miRNA_region)%>%
-  dplyr::filter(startsWith(miRNA_region, "mature"))%>%
-  dplyr::mutate(mirbase_id =str_remove(miRNA_region, "mature,"))%>%
-  dplyr::select(-miRNA_region)%>%
-  dplyr::inner_join(mirbase_id_conv, by = c("mirbase_id"="ID"))%>%
-  dplyr::select(miRNA_name = Name, read_count, reads_per_million_miRNA_mapped)%>%
-  dplyr::group_by(miRNA_name)%>%
-  mutate(read_count= sum(read_count), reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
-  dplyr::ungroup()%>%
-  distinct() -> BCPME_mirnatumor
-
-head(BCPME_mirnatumor)
-```
-
-+ Obtain the miRNA expression of normal tissue of patient:
-
-```{r, message= FALSE, warning=FALSE, eval=FALSE}
-BCP_mirnanormal <- GDCquery(project = "TCGA-BRCA",
-                  data.category = "Transcriptome Profiling",
-                  data.type = "Isoform Expression Quantification",
-                  workflow.type = "BCGSC miRNA Profiling",
-                  barcode = "TCGA-E9-A1N5-11A-41R-A14C-13")
-
-GDCdownload(BCP_mirnanormal)
-# a616435d-0b69-48ac-813d-5d75ad9b85eb.mirbase21.isoforms.quantification.txt
-
-GDCprepare(BCP_mirnanormal)%>%
-  as.data.frame()%>%
-  dplyr::select(miRNA_ID, read_count, reads_per_million_miRNA_mapped, miRNA_region)%>%
-  dplyr::filter(startsWith(miRNA_region, "mature"))%>%
-  dplyr::mutate(mirbase_id =str_remove(miRNA_region, "mature,"))%>%
-  dplyr::select(-miRNA_region)%>%
-  dplyr::inner_join(mirbase_id_conv, by = c("mirbase_id"="ID"))%>%
-  dplyr::select(miRNA_name = Name, read_count, reads_per_million_miRNA_mapped)%>%
-  dplyr::group_by(miRNA_name)%>%
-  mutate(read_count= sum(read_count), reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
-  dplyr::ungroup()%>%
-  distinct() -> BCPME_mirnanormal
-
-head(BCPME_mirnanormal)
-```
-
-## 6.3 Get the high-throughput experimental miRNA:target dataset.
-
-There are various datasets about miRNA:target pairs such as miRTarBase, DianaTools, miRecords, miRWalk etc.
-Some of these present the experimentally supported miRNA target pairs or only predicted ones.
-The experimentally supported datasets generally provides weak evidence for interactions.
-For these reasons, we obtained the high-throughput experimental miRNA:target dataset from two studies performed by Helwak et al.
-and Moore et al.
-These steps were not handle in this file because they contain many processes.
-
-Briefly these datasets contain various common information about miRNA:target interactions such as the miRNA name, miRNAsequence, target name, target sequence, their chromosomal locations, binding location on the target sequence, binding free energy, seed structure.
-But these datasets provides the informations with different data structures.So we followed the steps:
-
-+ The datasets were directly downloaded from supplementary data files of the studies.
-+ It was provided that the datasets are converted to same human genome build.
-+ The seed type information was organized as the same style.
-+ The datasets were combined.
-+ We committed the interaction factors as numeric values according to previous studies.
-(We added the interaction factors and their numeric values.
-The codes about integration of high-throughput experimental studies are available at Supplementary additional file and numeric values of interaction factors can be found at Table S2-3 in Supplementary Tables file.)
-
-Finally, we have obtained the experimentally supported miRNA:target dataset.
-
-```{r, message= FALSE, warning=FALSE}
-
-experimentalmirnagene<-readRDS("data/experimentalmirnagene.RDS")
-
-head(experimentalmirnagene)
+clashelwak <- clashelwak%>%
+ separate(microRNA_name, c("Barcode", "Database", "mirna_name", "type"), sep = "_")%>%
+  separate(mRNA_name, c("Ensembl_Gene_Id", "Ensembl_Transcript_Id", "Hugo_Symbol", "mRNA_Type"), sep = "_")
 
@@
 
-The methods about miRNA:target interactions are based a basic principle that is reading after isolation of miRNA:target chimerics.
-The datasets contain all the chimeric miRNA:target structures found in the medium during the experiment.
-On the other hand, it could be said that the reading is performed as snapshot.Because of that, the methods can provide different chimeric interactions the same miRNA:target pair.
-We have preferred to select most effective interaction parameters for the same miRNA:target pairs that can exhibit various interactions.
-The step is performed as:
+MiRNA releases are obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) was downloaded.
 
-```{r, message= FALSE, warning=FALSE, eval=FALSE}
-experimentalmirnagene%>%
-  dplyr::select(miRNA, Ensembl_Gene_Id, hgnc_symbol, Energy, seed_type_effect, region_effect)%>%
-  distinct()%>%
-  group_by(Ensembl_Gene_Id, miRNA)%>%
-  mutate(seed_type_effect= ifelse(seed_type_effect==max(seed_type_effect), seed_type_effect, max(seed_type_effect)), Energy = ifelse(Energy==min(Energy), Energy, min(Energy)), region_effect= ifelse(region_effect==max(region_effect), region_effect, max(region_effect)))%>%
-  distinct()-> tocombine_mirnagene
-
-head(tocombine_mirnagene)
-```
-
-## 6.4 Combine the dataset
-
-```{r, message= FALSE, warning=FALSE, eval=FALSE}
-BCPME_mirnanormal%>%
-  dplyr::inner_join(tocombine_mirnagene, by = c("miRNA_name"="miRNA"))%>%
-  dplyr::inner_join(TCGA_E9_A1N5_normal, by = c("Ensembl_Gene_Id"="ensembl_gene_id", "hgnc_symbol"="external_gene_name"))%>%
-  distinct()%>%
-  dplyr::select(hgnc_symbol, miRNA_name, mirna_RPM= reads_per_million_miRNA_mapped, GE_normal, Energy, seed_type_effect, region_effect)-> E9GE_mirnagenenormal
+```{r, eval=FALSE}
+read.table("mirbasehg38.txt", comment.char = "#")%>%
+  filter(V3 != "miRNA_primary_transcript")%>%
+  separate(V9, c("ID", "Alias", "Name", "Precusor"), sep = ";")%>%
+  mutate(ID = substr(ID, 4, length(ID)), Alias = substr(Alias, 7, length(Alias)), Name = substr(Name, 6, length(Name)), Precusor = substr(Precusor, 14, length(Precusor)))%>%
+  dplyr::select(chr= V1, start = V4, end = V5, strand = V7, ID, Alias, Name, Precusor)->mirbasehg38
 
@@
 
-A part of dataset can be found at Table S4 in Supplementary Tables.
+CLASH dataset is published in miRBase release 15 and Human Genome 19 version.
 
-```{r}
-E9GE_mirnagenenormal<-readRDS("data/E9GE_mirnagenenormal.RDS")
-head(E9GE_mirnagenenormal)
-```
-
-```{r, message= FALSE, warning=FALSE, eval=FALSE}
-BCPME_mirnatumor%>%
-  dplyr::inner_join(tocombine_mirnagene, by = c("miRNA_name"="miRNA"))%>%
-  dplyr::inner_join(TCGA_E9_A1N5_tumor, by = c("Ensembl_Gene_Id"="ensembl_gene_id", "hgnc_symbol"="external_gene_name"))%>%
+```{r, eval=FALSE}
+read_tsv("mirna_mature.txt", col_names = FALSE)%>%
+  filter(startsWith(X2, "hsa"))%>%
+  dplyr::select(mirna_ID = X2, mirbase_ID = X3)%>%
+  inner_join(mirbasehg38%>% dplyr::select(ID, Name), by= c("mirbase_ID"= "ID"))%>%
+  dplyr::select(mirbase_ID, Name)%>%
   distinct()%>%
-  dplyr::select(hgnc_symbol, miRNA_name, mirna_RPM= reads_per_million_miRNA_mapped, GE_tumor, Energy, seed_type_effect, region_effect)-> E9GE_mirnagenetumor
+  inner_join(clashelwak, by = c("mirbase_ID"= "Barcode"))%>%
+  dplyr::select(Name, miRNA_seq, Ensembl_Gene_Id, Ensembl_Transcript_Id, Hugo_Symbol, mRNA_seq_extended, chimeras_decompressed, seed_type, seed_basepairs,folding_class, seq_ID, folding_energy, X5.UTR, CDS, X3.UTR)%>%
+  inner_join(hg19, by = c("Ensembl_Gene_Id"= "ensembl_gene_id", "Ensembl_Transcript_Id"="ensembl_transcript_id", "Hugo_Symbol"="hgnc_symbol"))%>%
+  mutate(region1 = ifelse(X5.UTR == "1", "5UTR", " "), region2= ifelse(X3.UTR == "1", "3UTR", " "),region3= ifelse(CDS =="1", "CDS", " "))%>%
+  unite(region, c(region1, region2, region3), sep = "||")%>%
+  dplyr::select(chromosome_name, start_position, end_position,strand, Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, mRNA_seq_extended, Name, miRNA_seq, seq_ID, seed_type,seed_basepairs, folding_class, folding_energy, region)%>%
+  as_tibble() -> clashelwakfinal
 
@@
 
-```{r}
-#saveRDS(E9GE_mirnagenetumor, "data/E9GE_mirnagenetumor.RDS")
+### Converting CLASH data to human genome 38 build.
 
-E9GE_mirnagenetumor<-readRDS("data/E9GE_mirnagenetumor.RDS")
-head(E9GE_mirnagenetumor)
+There are different liftover methods for conversion among Human Genome builds. We preffered to use [UCSC liftover tool](https://genome.ucsc.edu/cgi-bin/hgLiftOver)
+
+```{r,  eval=FALSE}
+
+#Obtaining chromosomal locations from miRNA:target interaction dataset.
+
+lift19 <- clashelwakfinal%>%
+  dplyr::select(1,2,3)%>%
+  unite(start_end, c("start_position", "end_position"), sep = "-")%>%
+  mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
+  unite(chromosome_name, c("Chromosome","start_end"), sep = ":")
+
+write_tsv(lift19, "lift19.txt")
+
+#After we searched this file in UCSC browser, the output loaded (lift19_del deleted regions on the HG38 genome build; hg38clashcomp new locations of the genes on HG38)
+
+lift19_del <- read_tsv("deleted_lift19.txt")
+colnames(lift19_del)[1] <- "chromosome_loc"
+
+lift19_del <- lift19_del%>%
+  dplyr::filter(startsWith(chromosome_loc, "chr"))%>%
+  separate(chromosome_loc, c("Chr", "End"), "-", remove = TRUE)%>%
+  separate(Chr, c("Chr", "Start"), ":", remove = TRUE)
+
+lift19_del$Start <- as.numeric(lift19_del$Start)
+lift19_del$End <- as.numeric(lift19_del$End)
+
+#removing deleted location from CLASH dataset
+
+clashelwakfinal <-clashelwakfinal%>%
+  mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
+  dplyr::anti_join(lift19_del, by = c("Chromosome"="Chr", "start_position"="Start", "end_position"="End"))
+
+hg38clash <- read.delim("hg38clashcomp.txt", header = FALSE, stringsAsFactors = FALSE)
+
+# adding new location information:
+
+clashelwakfinal <- clashelwakfinal%>% 
+  bind_cols(hg38clash)
+
+colnames(clashelwakfinal)[18] <- "HG38build_loc"
+
+clashelwakfinal <- clashelwakfinal%>%
+  dplyr::mutate(Genom_build= rep("hg19"))
+
+# Arrangement in dataset
+
+clashelwakfinal%>% dplyr::select(cluster=seq_ID, chromosome = Chromosome, start_position, end_position, strand, hgnc_symbol = Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, target_seq = mRNA_seq_extended, miRNA= Name, miR_seq =miRNA_seq, seed_type, seed_type2= seed_basepairs, seed_type3= folding_class, Energy = folding_energy, HG38build_loc, Genom_build, region) -> clashelwakfinal
+
+clashelwakfinal$strand <- as.character(clashelwakfinal$strand)
+
+str(clashelwakfinal)
 ```
 
-## 6.5 Selection of trigger node
+### Interpreting the CLASH seed structures in dataset
 
-We have determined the most important nodes of network.
-We applied `find_node_perturbation` function.
-We only defined nodes that affect the other nodes more than 1.05 fold change with 10 iteration when they increase 3 fold.
-
-```{r, include=FALSE, eval=FALSE}
-
-as_data_frame(E9GE_mirnagenenormal)%>%
-  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
-  find_node_perturbation(how = 3, cycle = 10, fast = 5)%>%
-  filter(!is.na(perturbation_efficiency))->perturbationofnetwork
-
-perturbationofnetwork%>%
-  filter(type== "Competing")%>%
-  arrange(desc(perturbation_efficiency))
-```
-This dataset, `perturbationofnetwork`, includes totally 423 effective nodes. We selected SERPINE2 gene to perturb the network because it is the most efficient node gene in network.
-
-## 6.6 Approach of Method into Combined Datasets
-
-We selected SERPINE2 gene for simulation of regulation on network.
-
-### 6.6.1 Find iteration of simulation
-
-The node amount of changed gene on the system in terms of percentage were shown at Figure S9 in Supplementary Figures.
-As seen, firstly, the changed gene count increase.
-The system which contains the hundreds of miRNAs and thousands of genes can slowly gain the steady-state again.
-At first glance, it can be assumed that when all nodes in the system are reached, stable state will be provided.
-However, although all nodes are reached, the nodes competing with each other cause the edits to continue for a while.
-
-The dynamics of the approach are shown in package  [vignettes](https://selcenari.github.io/ceRNAnetsim/articles/convenient_iteration.html).
-
-So, we offered an approach about to find iteration.
-`find_iteration` function does not give the iteration to gain steady-state, but it gives the iteration which has maximum affected node counts .
-The function is applied as following:
-
-```{r}
-
-as.data.frame(E9GE_mirnagenenormal)%>%
- priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
- update_how("SERPINE2",2.75)%>%
- simulate(50) %>%
- find_iteration(limit=1, plot= FALSE)
+```{r,  eval=FALSE}
+clashelwakfinal%>%
+  mutate(seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="I", paste0(seed_type2, "-mer"), seed_type),
+         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="II", paste0(seed_type2, "-mer_noncanonical"), seed_type), 
+         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="III", paste0(seed_type2, "-mer_noncanonical"), seed_type),
+         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="IV", paste0(seed_type2, "-mer_noncanonical"), seed_type),
+         seed_type= ifelse (startsWith(seed_type, "no") ,"none", seed_type))%>%
+  dplyr::select(-seed_type2, -seed_type3)->clashelwakfinal
 
@@
 
-### 6.6.2 Simulation of dataset
+## 6.2 Arrangement of CLEAR-CLiP Dataset [@moore_mirnatarget_2015]
 
-We tried to apply two fold of the point that SERPINE2 has maximum affected genes on network.
+CLASH dataset was retrieved from [Nature article](nature.com/articles/ncomms9864)
 
-```{r}
-as.data.frame(E9GE_mirnagenenormal)%>%
-  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
-  update_how("SERPINE2",2.75)%>%
-  simulate(62)
+```{r,  eval=FALSE}
+clearclip <- read_xlsx("CLEAR-CLIP.xlsx")
+
+#Clearclip hg18
+```
+
+### Query of Human Genome 18
+
+```{r,  eval=FALSE}
+#HG18
+listEnsemblArchives()
+listMarts(host = 'may2009.archive.ensembl.org')
+ensembl54 = useMart(host='may2009.archive.ensembl.org', 
+                     biomart='ENSEMBL_MART_ENSEMBL',
+                    dataset='hsapiens_gene_ensembl')
+
+hg18 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene", "strand" ), mart = ensembl54)
+```
+
+### Adding Genome Information to dataset
+
+```{r,  eval=FALSE}
+clearclipfinal <- hg18%>%
+  inner_join(clearclip, by= c("entrezgene"= "gene.id", "hgnc_symbol"= "gene.symbol"))%>%
+  distinct()
+```
+
+### Converting human genome build
+
+```{r,   eval=FALSE}
+
+#Obtaining chromosomal locations from miRNA:target interaction dataset.
+
+lift18 <- clearclipfinal%>%
+  unite(start_end, c("start_position", "end_position"), sep = "-")%>%
+  unite(location, c("chr","start_end"), sep = ":")%>%
+  dplyr::select(location)
+
+write_tsv(lift18, "lift18.txt")
+
+#After we searched this file in UCSC browser, the output loaded (deleted_lift18 deleted regions on the HG38 genome build; hg38clearclip new locations of the genes on HG38)
+
+deleted_lift18 <- read_tsv("deleted_lift18.txt")
+
+colnames(deleted_lift18)[1] <- "Chromosome_loc"
+
+deleted_lift18 <- deleted_lift18%>%
+  dplyr::filter(startsWith(Chromosome_loc, "chr"))%>%
+  separate(Chromosome_loc, c("Chr", "End"), "-", remove = TRUE)%>%
+  separate(Chr, c("Chr", "Start"), ":", remove = TRUE)
+
+deleted_lift18$Start <- as.numeric(deleted_lift18$Start)
+deleted_lift18$End <- as.numeric(deleted_lift18$End)
+
+#removing deleted location from CLEAR-CLiP dataset
+
+clearclipfinal <- clearclipfinal%>%
+  dplyr::anti_join(deleted_lift18, by = c("chr"="Chr", "start_position"="Start", "end_position"="End"))
+
+hg38clearclip<- read.delim("hg38clearclip.txt", header = FALSE, stringsAsFactors = FALSE)
+
+clearclipfinal <- clearclipfinal%>% 
+  bind_cols(hg38clearclip)
+
+colnames(clearclipfinal)[28] <- "HG38build_loc"
+
+# adding new location information:
+
+clearclipfinal <- clearclipfinal%>%
+  dplyr::mutate(Genom_build= rep("hg18"))
+
+# Arrangement in dataset
+
+clearclipfinal%>%
+  dplyr::select(cluster=cluster.ID, chromosome = chr, start_position, end_position, strand = strand.y, hgnc_symbol, Ensembl_Gene_Id =ensembl_gene_id, Ensembl_Transcript_Id =ensembl_transcript_id, target_seq = target.map, miRNA, miR_seq = miR.map, seed_type = 'seed match', Energy = MFE, HG38build_loc, Genom_build, region) -> clearclipfinal
 
@@
 
-```{r}
-as.data.frame(E9GE_mirnagenenormal)%>%
-  priming_graph(competing_count = GE_normal, miRNA_count = mirna_RPM, aff_factor = c(Energy, seed_type_effect), deg_factor = region_effect)%>%
-  update_how("SERPINE2",2.75)%>%
-  simulate(62)%>%
-  as_tibble()%>%
-  dplyr::select(name, initial_count, count_current)->simulation_results
+### Seed type manipulation in CLEAR-CLiP dataset
+
+In CLEAR-CLiP dataset, seed types were shown in detail. We adjusted as canonical and non-canonical.
+
+```{r,  eval=FALSE}
+data_frame(seed_type= c("5mer_1","5mer_2", "5mer_3", "6mer", "6mer.indel", "6mer.mm", "6mer_off.mm", "6merA1", "6merA1.indel", "6merA1.mm", "7merA1", "7merA1.indel", "7merA1.mm", "7merm8", "7merm8.indel", "7merm8,mm", "8mer", "8mer.indel", "8mer.mm", "NA" ),
+           seed_type_com= c("5-mer", "5-mer_noncanonical", "5-mer_noncanonical", "6-mer", "6-mer_noncanonical", "6-mer_noncanonical", "6-mer_noncanonical", "6-merA1", "6-merA1_noncanonical","6-merA1_noncanonical", "7-merA1", "7-merA1_noncanonical","7-merA1_noncanonical", "7-mer-8m", "7-mer-8m_noncanonical", "7-mer-8m_noncanonical", "8-mer", "8-mer_noncanonical", "8-mer_noncanonical", "none" )) ->clipdata_seed
+
+clearclipfinal%>%
+  inner_join(clipdata_seed, by = "seed_type")%>%
+  dplyr::select(1:11, seed_type = seed_type_com, Energy, HG38build_loc, Genom_build, region)-> clearclipfinal
+
 ```
 
-### 6.6.3 Comparison of simulation results and tumor tissue expression values
-```{r}
-E9GE_mirnagenetumor%>%
-  dplyr::select(hgnc_symbol, GE_tumor)%>%
-  dplyr::inner_join((E9GE_mirnagenenormal%>%dplyr::select(hgnc_symbol, GE_normal)), by ="hgnc_symbol")%>%
-  inner_join(simulation_results, by= c("hgnc_symbol"="name"))%>%
-  distinct()
+## 6.3 Integration of two experimental dataset
+
+```{r,  eval=FALSE}
+
+bind_rows(clashelwakfinal, clearclipfinal)%>%distinct() -> experimentalmirnagene
+
 ```
 
-Actually, we have developed to provide a new approach miRNA mediated regulation networks.
-This approach may not explain the whole regulation behaviors between miRNAs and targets but can be first step to more detailed and coherent miRNA:target regulation approach.
+### Adding Coefficients of Interaction factors
+
+Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study.
+On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study.
+Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions.
+Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values.
+We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
+With this process, we have handled this integrated dataset in context of competitor behaviors and functionality of interactions.
+
+In this step we added numeric intraction values at followings
+
+Fistly, we organized these values due to the fact that the regions were defined differently in two datasets. After that, region effect was added as numeric values (shown in Table S3).
+
+```{r,  eval=FALSE}
+experimentalmirnagene <- experimentalmirnagene%>%
+  mutate(region2  = str_replace_all(region, "NA", ""), region3 = str_replace_all(region2, "\\|", ""), region = str_replace_all(region3, c("3'UTR"="3UTR", "5'UTR"="5UTR")))%>%
+  dplyr::select(-region2, -region3)%>%
+  mutate(region_effect = as.double(ifelse(region %in% c("3UTRCDS", "CDS3UTR", "5UTR3UTR", "CDS5UTR3UTR", "CDS3UTRintron"), "0.93", ifelse(region %in% c("CDS", "CDSintron"), "0.42", ifelse(region %in% c("3UTR","3UTRintron"), "0.84", ifelse(region %in% c("5UTR", "5UTRintron"), "0.01", ifelse(region %in% c("5UTRCDS", "CDS5UTR"), "0.42", ifelse(region %in% c("intron",""), "0.01", ifelse(region %in% c("exon_unclassified", ""), "0.2", NA)))))))))
+
+```
+
+Secondly, we organized seed type interactions in *Seed type manipulation* section for CLEAR-CLiP dataset to show as found in CLASH dataset.
+Same type formatted values added dataset as numeric values (shown in Table S2).
+
+```{r,  eval=FALSE}
+
+seed_type_effect <- data_frame( seed_type = c("5-mer", "5-mer_noncanonical", "6-mer", "6-mer_noncanonical", "6-merA1", "6-merA1_noncanonical", "7-mer", "7-mer_noncanonical", "7-merA1", "7-merA1_noncanonical", "7-mer-8m", "7-mer-8m_noncanonical", "8-mer", "8-mer_noncanonical", "9-mer", "9-mer_noncanonical", "none"),
+                                seed_type_effect= c(0.05, 0.04, 0.07, 0.05, 0.07, 0.05, 0.23, 0.19, 0.19, 0.16, 0.25, 0.21, 0.43, 0.35, 0.43,0.35, 0.01)
+                                )
+
+experimentalmirnagene%>%
+  inner_join(seed_type_effect, by= "seed_type") -> experimentalmirnagene
+```
+
+```{r}
+readRDS("data/experimentalmirnagene.RDS")->experimentalmirnagene
+experimentalmirnagene
+```
+
+The context of dataset is shown in Table S5 in Supplementary Tables.
+
+# REFERENCES

### 2020-08-19 recovering1.Rmd
adding recent analysis

@@
 
-```{r}
-BiocManager::install("RDAVIDWebService")
-library(RDAVIDWebService)
-```
+#Special Genes
 
@@
 ```
+
+#Special mirnas
 
@@
 read_delim("string_TFs.txt", delim = "\t", col_names = "text", skip = 1)%>%
-  separate(text, into = c("critical_gene", "Gene2", "Direct", "Interaction", "shared_gene_num", "shared_genes"), sep = "\t")%>%
-  mutate(shared_gene = str_split(shared_genes, pattern = ","))%>%
+  separate(text, into = c("TFs", "disease_assoc", "regulated_genes", "regulated_gene_assoc", "regulated_associated_genes"), sep = "\t")%>%
+  mutate(shared_gene = str_split(regulated_gene_assoc, pattern = ","))%>%
   unnest()%>%
-  select(critical_gene, Gene2, shared_gene)%>%
+  select(TFs, disease_assoc, shared_gene)%>%
   distinct()%>%
   mutate(shared_gene=str_trim(shared_gene, side = "both"))->TFs
+
+TFs%>%
+  left_join(hg19, by = c("shared_gene"="hgnc_symbol"))%>%
+  distinct()%>%
+  filter(!is.na(ensembl_gene_id), ensembl_gene_id %in% highest_efficient_nodes$name)%>%
+  mutate(interaction = "TF associated", term = paste0("TF_", TFs))%>%
+  select(term, genes2=ensembl_gene_id, hgnc_symbol = shared_gene, interaction)%>%
+  distinct()
 
@@
   mutate(interaction = ifelse(hgnc_symbol %in% TFs$shared_gene, "TF associated", NA),
-         interaction = ifelse(hgnc_symbol %in% TFs$Gene2, "TFs", interaction),
+         interaction = ifelse(hgnc_symbol %in% TFs$TFs, "TFs", interaction),
          interaction = ifelse(hgnc_symbol %in% edgar_critcals$critical_gene, "critical", interaction))%>%
@@
   filter(!is.na(interaction))->experimental_res
+
+experimental_res%>%
+  filter(interaction != "TFs", interaction != "TF associated")%>%
+  select(name)%>%
+  distinct()
+
+experimental_res%>%
+  filter(interaction == "TFs"| interaction == "TF associated")%>%
+  left_join(TFs, by = c("hgnc_symbol"="shared_gene"))%>%
+  filter(disease_assoc == "YES")%>%
+  distinct(hgnc_symbol)
+
+#80'i edgar TF ve pathwaylerde bulunmuş, 41 TF ilşkili 4 ü zaten transkripsiyon faktörü, 47'si ise pathwaylerde olmak üzere.
 

### 2020-08-19 recovery2_mirtarbase.Rmd
adding recent analysis

@@
 library(tinytex)
+library(ggraph)
+library(ggplot2)
+library(igraph)
+library(tidygraph)
 ```
@@
 #3265 gen, 581 miRNA
-
-mirtarbasegene_mat
 
@@
   select(name)%>%
-  filter(!startsWith(name, "ENS"))%>%
-  pull()->lim1_tumor_mirna
+  filter(startsWith(name, "ENS"))%>%
+  pull()->lim1_tumor_gene
 
@@
   select(name)%>%
-  filter(!startsWith(name, "ENS"))%>%
-  pull()-> lim1_normal_mirna
+  filter(startsWith(name, "ENS"))%>%
+  pull()-> lim1_normal_gene
 ```
@@
   filter(perturbed_FC> 1 | perturbed_FC < -1 )
+    #-> lim1_mirna
 
-  #-> lim1_mirna
+  #filter(name %in% high_efficient$name)%>%
+  #filter(perturbed_FC> 1 | perturbed_FC < -1 )%>%
+  #select(name)%>%
+  #distinct() ->highest_efficient_nodes
 
@@
 
-```{r}
-22/24
-```
-
 --------------------------------------------------
@@
 
-# how 0.33 analizi
+# sample network
+
+bunun için gen ifadesi vs zaten gerekmiyor combined pairs mirtarbase_lim1 klasöründe node'lardan filtre yapıp edge seçmiş olacağım.
+
+```{r}
+
+node_mirtarbase
+
+readRDS("mirtarbase_lim_1/tocombinepairs_mirtarbase.RDS")%>%
+  filter(miRNA %in% node_mirtarbase$name, Ensembl_Gene_Id %in% node_mirtarbase$name)%>%
+  mutate(miRNA_name = miRNA, competing_name = Ensembl_Gene_Id)%>%
+  as_tbl_graph()
+
+# Tabi ki elimde yine dev gibi bir network var. Şimdi de effective olarak bulduklarımızı filtreleyelim:
+
+# 228. satırdaki chunkta highest olanları buldum.
+
+highest_efficient_nodes
+
+readRDS("mirtarbase_lim_1/tocombinepairs_mirtarbase.RDS")%>%
+  filter(miRNA %in% highest_efficient_nodes$name, Ensembl_Gene_Id %in% highest_efficient_nodes$name)%>%
+  mutate(miRNA_name = miRNA, competing_name = Ensembl_Gene_Id)%>%
+  as_tbl_graph()%>%
+  mutate(node_type = ifelse(startsWith(name, "hsa"), "miRNA", "ceRNA"))%>%
+  ggraph(layout = "fr")+ #fr, graphopt iyi olabilir.
+  geom_edge_link(colour = "seashell2")+
+  geom_node_point(aes(filter= node_type == "ceRNA"), color = "sandybrown", shape = "circle", size= 1.5)+
+  geom_node_point(aes(filter= node_type == "miRNA"), color = "slateblue4", shape = "triangle", size= 3)+
+  theme_graph(base_family = "sans")+
+  geom_node_text(aes(filter= node_type == "miRNA", label = name), size =3, repel = TRUE)
+
+#son olarak kaydedilecek
+ggsave(filename = 'highest_efficient_nodes_mirtarbase.eps', device=cairo_ps, width = 12, height = 7 )
+
+```
+# Hive R plot
+
+transkripsiyon faktörleri
+
+```{r}
+
+readRDS("mirtarbase_lim_1/tocombinepairs_mirtarbase.RDS")%>%
+  filter(miRNA %in% highest_efficient_nodes$name, Ensembl_Gene_Id %in% highest_efficient_nodes$name)%>%
+  select(2,1)->combined_pairs
+
+read_delim("string_TFs.txt", delim = "\t", col_names = "text", skip = 1)%>%
+  separate(text, into = c("TFs", "disease_assoc", "regulated_genes", "regulated_gene_assoc", "regulated_associated_genes"), sep = "\t")%>%
+  mutate(shared_gene = str_split(regulated_gene_assoc, pattern = ","))%>%
+  unnest()%>%
+  select(TFs, disease_assoc, shared_gene)%>%
+  distinct()%>%
+  mutate(shared_gene=str_trim(shared_gene, side = "both"))->TFs
+
+TFs%>%
+  left_join(hg19, by = c("shared_gene"="hgnc_symbol"))%>%
+  distinct()%>%
+  filter(!is.na(ensembl_gene_id), ensembl_gene_id %in% highest_efficient_nodes$name)%>%
+  mutate(interaction = "TF associated", term = paste0("TF_", TFs))%>%
+  select(term, genes2=ensembl_gene_id, hgnc_symbol = shared_gene, interaction)%>%
+  distinct()->mirtarbase_TF_analysis
+
+# 83 gen meme kanseri ilişkili olarak rapor edilmiş. 227 tanesi ise zaten bulunmuş.
+
+```
+
+```{r}
+
+bind_rows((
+read_tsv("kegg_mirtarbase_lim1.txt")%>%
+  filter(PValue < 0.05)%>%
+  clean_names()%>%
+  select(term, Gene_Count = count, p_value, genes)%>%
+  mutate(genes2 = str_split(genes, ", "))%>%
+  unnest()%>%
+  mutate(genes2 = str_trim(genes2, side = "both"), genes2 = ifelse(endsWith(genes2, ","), substr(genes2, 0, (nchar(genes2)-1)), genes2))%>%
+  left_join(hg19, by = c("genes2"="ensembl_gene_id"))%>%
+  select(term, genes2, hgnc_symbol)%>%
+  mutate(interaction = "KEGG_pathways")), (read_tsv("biogrid_mirtarbase_lim1.txt")%>%
+  filter(PValue < 0.05)%>%
+  clean_names()%>%
+  select(term, Gene_Count = count, p_value, genes)%>%
+  mutate(genes2 = str_split(genes, ", "))%>%
+  unnest()%>%
+  mutate(genes2 = str_trim(genes2, side = "both"), genes2 = ifelse(endsWith(genes2, ","), substr(genes2, 0, (nchar(genes2)-1)), genes2))%>%
+  left_join(hg19, by = c("genes2"="ensembl_gene_id"))%>%
+  select(term, genes2, hgnc_symbol)%>%
+  mutate(interaction = "PPI")))-> hive_plot
+
+hive_plot%>%
+  bind_rows(mirtarbase_TF_analysis)->hive_plot
+
+hive_plot%>%
+  filter(startsWith(term, "TF_"))
+
+hive_plot%>%
+  select(interaction)%>%
+  distinct()
+
+hive_plot%>%
+  filter(interaction != "PPI", interaction != "KEGG_pathways")%>%
+  distinct(hgnc_symbol)
+
+# 318'i kegg patwayde
+# 471'i biogrid interaction'larda
+# 227'si ise TF associated Bunun 83'ü meme kanseri ilişkili olarak rapor edimiş.
+
+hive_plot%>%
+  select(hgnc_symbol)%>%
+  distinct()
+
+#607 tanesi
+
+hive_plot%>%
+  arrange(hgnc_symbol)%>%
+  mutate(situ= term)%>%
+  as_tbl_graph()%>%
+  mutate(interaction = "Biogrid protein interaction",
+         interaction = ifelse(startsWith(name, "hsa"), "KEGG Pathway", interaction),
+         interaction = ifelse(startsWith(name, "ENS"), "Gene", interaction),
+         interaction = ifelse(startsWith(name, "TF_"), "TF associated", interaction)
+         )%>%
+  #filter(interaction != "Biogrid protein interaction")%>%
+  ggraph("hive", axis = interaction)+
+  geom_edge_hive(aes(color = factor(situ)), lineend = "butt",
+  linejoin = "round",
+  linemitre = 2,
+  label_alpha = 1,
+  label_parse = FALSE,
+  check_overlap = TRUE) + 
+  geom_axis_hive(color="gray43", size = 2) +
+  coord_fixed()+
+  theme_graph()+
+  theme(legend.position="none")
+
+ggsave(filename = 'highest_efficient_nodes_enrichment_mirtarbase.eps', device=cairo_ps)
+
+```

### 2020-10-20 Reproducible_codes.Rmd
adding revised files

--- 2020-08-14 Supplementary_mat_met.Rmd
@@
 ---
-title: "Supplementary Materials And Methods"
+title: "Reproducible Codes"
 author: 
@@
 
-# 1. Functions defined for ceRNA models and workflow of method
-
-We defined the functions that can be used with R programming.
-Briefly, these functions process a given miRNA:gene dataset and convert to graph object.
-All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
-The functions and steps of approach are explained as following (Figure S1 in Supplementary Figures) :
-
-**Conversion of dataset**: `priming_graph()` function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
-If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
-The formulations that are given in equations (1-4) are performed in this function.
-This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
-It is assumed that the initial target amounts in the dataset is observed after the repressive activity of miRNAs in steady-state.
-
-**Transition of variables in graph**: In the previous step, the calculations are performed in the edge variables of the graph object.
-However, the graph object allows to use node variables, while the node features are handled to the graph.
-In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.
-
-**Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficiency coefficients are calculated according to this acceptance.
-In the network that is found in steady-state conditions, the change is applied to the graph object for perturbation of steady-state.
-To provide the perturbation in the network the work-flow offer two methods: *update_variables* and *update_how*.
-The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
-The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
-
-**Simulation of competing behavior of targets**: After the change in the steady-state conditions, the network elements try to gain steady-state again.
-This process progresses as repeating of regulations after the spreading the changes in the network.
-In this step, simulation of regulations according to given cycle count in simulate function is applied.
-After each simulation cycle, the miRNA repression values are re-calculated and the current values of competing elements are found and saved.
-The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node data.
-
-The node elements in the dataset are handled as two type; repressive (miRNAs) and competing (targets).
-It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used.
-If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user.
-On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger.
-The functions that are used in the this study are developed with R and are available in [Bioconductor](https://www.bioconductor.org/packages/release/bioc/html/ceRNAnetsim.html).
-
 ```{r, include=FALSE, message=FALSE, warning=FALSE}
@@
 
-+ load *minsamp* data
-
-```{r}
-data("minsamp")
-minsamp
+```{r, include=FALSE}
+data("minsamp") # refers Sample and Sample+ data
 ```
@@
 
-# 2. *minsamp* dataset analysis in lack of interaction factors.
+# 2. Sample dataset analysis in lack of interaction factors.
 
@@
 
-# 3. *minsamp* dataset analysis with interaction factors.
+# 3. Sample+ dataset analysis with interaction factors.
 
-We have made the same analysis in presence of interaction factors (Sequentially shown at Figure S5-7 in Supplementary Figures).
+We have made the same analysis in presence of interaction factors (Sequentially shown at Figure S6-8 in Supplementary Figures).
 
@@
 
-# 4. Common target perturbation in *minsamp* dataset.
+# 4. Common target perturbation in Sample+ dataset.
 
@@
 In our small dataset, minsamp, Gene4 is regulated by two miRNAs.
-Let's simulate perturbation effects triggered by Gene4 (Shown at Figure S8 in Supplementary Figures) .
+Let's simulate perturbation effects triggered by Gene4 (Shown at Figure S9 in Supplementary Figures) .
 
@@
 ```
-
-# 5. Determination of perturbation efficiencies of elements in system.
-
-```{r, message=FALSE, warning=FALSE}
-
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)-> sample_graph
-
-find_node_perturbation(sample_graph, how = 2, cycle = 3, limit = 0.1)%>%
-  as_tibble()%>%
-  select(name, perturbation_efficiency, perturbed_count)
-
-```
-
-# 6. Additional data manipulation steps
-
-## 6.1 Arrangement of CLASH dataset
-
-CLASH dataset was retrieved from [PubMed](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3650559/) [@helwak_mapping_2013].
-
-```{r, eval=FALSE}
-
-clashelwak <- read.table("mmc1.txt", comment.char = "#", header = TRUE, skip = 1, stringsAsFactors = FALSE)
-
-#hg19
-```
-
-### Query of Human Genome 19.
-
-Human genome 19 information was handled through biomaRt package.
-
-```{r,  eval=FALSE}
-#HG19
-listEnsemblArchives()
-listMarts(host = 'http://grch37.ensembl.org')
-ensemblgrch37 = useMart(host='http://grch37.ensembl.org', 
-                     biomart='ENSEMBL_MART_ENSEMBL',
-                    dataset='hsapiens_gene_ensembl')
-hg19 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene_id", "strand" ), mart = ensemblgrch37)
-
-```
-
-### Adding miRNA and gene information
-
-```{r,  eval=FALSE}
-
-clashelwak <- clashelwak%>%
- separate(microRNA_name, c("Barcode", "Database", "mirna_name", "type"), sep = "_")%>%
-  separate(mRNA_name, c("Ensembl_Gene_Id", "Ensembl_Transcript_Id", "Hugo_Symbol", "mRNA_Type"), sep = "_")
-
-```
-
-MiRNA releases are obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) was downloaded.
-
-```{r, eval=FALSE}
-read.table("mirbasehg38.txt", comment.char = "#")%>%
-  filter(V3 != "miRNA_primary_transcript")%>%
-  separate(V9, c("ID", "Alias", "Name", "Precusor"), sep = ";")%>%
-  mutate(ID = substr(ID, 4, length(ID)), Alias = substr(Alias, 7, length(Alias)), Name = substr(Name, 6, length(Name)), Precusor = substr(Precusor, 14, length(Precusor)))%>%
-  dplyr::select(chr= V1, start = V4, end = V5, strand = V7, ID, Alias, Name, Precusor)->mirbasehg38
-
-```
-
-CLASH dataset is published in miRBase release 15 and Human Genome 19 version.
-
-```{r, eval=FALSE}
-read_tsv("mirna_mature.txt", col_names = FALSE)%>%
-  filter(startsWith(X2, "hsa"))%>%
-  dplyr::select(mirna_ID = X2, mirbase_ID = X3)%>%
-  inner_join(mirbasehg38%>% dplyr::select(ID, Name), by= c("mirbase_ID"= "ID"))%>%
-  dplyr::select(mirbase_ID, Name)%>%
-  distinct()%>%
-  inner_join(clashelwak, by = c("mirbase_ID"= "Barcode"))%>%
-  dplyr::select(Name, miRNA_seq, Ensembl_Gene_Id, Ensembl_Transcript_Id, Hugo_Symbol, mRNA_seq_extended, chimeras_decompressed, seed_type, seed_basepairs,folding_class, seq_ID, folding_energy, X5.UTR, CDS, X3.UTR)%>%
-  inner_join(hg19, by = c("Ensembl_Gene_Id"= "ensembl_gene_id", "Ensembl_Transcript_Id"="ensembl_transcript_id", "Hugo_Symbol"="hgnc_symbol"))%>%
-  mutate(region1 = ifelse(X5.UTR == "1", "5UTR", " "), region2= ifelse(X3.UTR == "1", "3UTR", " "),region3= ifelse(CDS =="1", "CDS", " "))%>%
-  unite(region, c(region1, region2, region3), sep = "||")%>%
-  dplyr::select(chromosome_name, start_position, end_position,strand, Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, mRNA_seq_extended, Name, miRNA_seq, seq_ID, seed_type,seed_basepairs, folding_class, folding_energy, region)%>%
-  as_tibble() -> clashelwakfinal
-
-```
-
-### Converting CLASH data to human genome 38 build.
-
-There are different liftover methods for conversion among Human Genome builds. We preffered to use [UCSC liftover tool](https://genome.ucsc.edu/cgi-bin/hgLiftOver)
-
-```{r,  eval=FALSE}
-
-#Obtaining chromosomal locations from miRNA:target interaction dataset.
-
-lift19 <- clashelwakfinal%>%
-  dplyr::select(1,2,3)%>%
-  unite(start_end, c("start_position", "end_position"), sep = "-")%>%
-  mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
-  unite(chromosome_name, c("Chromosome","start_end"), sep = ":")
-
-write_tsv(lift19, "lift19.txt")
-
-#After we searched this file in UCSC browser, the output loaded (lift19_del deleted regions on the HG38 genome build; hg38clashcomp new locations of the genes on HG38)
-
-lift19_del <- read_tsv("deleted_lift19.txt")
-colnames(lift19_del)[1] <- "chromosome_loc"
-
-lift19_del <- lift19_del%>%
-  dplyr::filter(startsWith(chromosome_loc, "chr"))%>%
-  separate(chromosome_loc, c("Chr", "End"), "-", remove = TRUE)%>%
-  separate(Chr, c("Chr", "Start"), ":", remove = TRUE)
-
-lift19_del$Start <- as.numeric(lift19_del$Start)
-lift19_del$End <- as.numeric(lift19_del$End)
-
-#removing deleted location from CLASH dataset
-
-clashelwakfinal <-clashelwakfinal%>%
-  mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
-  dplyr::anti_join(lift19_del, by = c("Chromosome"="Chr", "start_position"="Start", "end_position"="End"))
-
-hg38clash <- read.delim("hg38clashcomp.txt", header = FALSE, stringsAsFactors = FALSE)
-
-# adding new location information:
-
-clashelwakfinal <- clashelwakfinal%>% 
-  bind_cols(hg38clash)
-
-colnames(clashelwakfinal)[18] <- "HG38build_loc"
-
-clashelwakfinal <- clashelwakfinal%>%
-  dplyr::mutate(Genom_build= rep("hg19"))
-
-# Arrangement in dataset
-
-clashelwakfinal%>% dplyr::select(cluster=seq_ID, chromosome = Chromosome, start_position, end_position, strand, hgnc_symbol = Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, target_seq = mRNA_seq_extended, miRNA= Name, miR_seq =miRNA_seq, seed_type, seed_type2= seed_basepairs, seed_type3= folding_class, Energy = folding_energy, HG38build_loc, Genom_build, region) -> clashelwakfinal
-
-clashelwakfinal$strand <- as.character(clashelwakfinal$strand)
-
-str(clashelwakfinal)
-```
-
-### Interpreting the CLASH seed structures in dataset
-
-```{r,  eval=FALSE}
-clashelwakfinal%>%
-  mutate(seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="I", paste0(seed_type2, "-mer"), seed_type),
-         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="II", paste0(seed_type2, "-mer_noncanonical"), seed_type), 
-         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="III", paste0(seed_type2, "-mer_noncanonical"), seed_type),
-         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="IV", paste0(seed_type2, "-mer_noncanonical"), seed_type),
-         seed_type= ifelse (startsWith(seed_type, "no") ,"none", seed_type))%>%
-  dplyr::select(-seed_type2, -seed_type3)->clashelwakfinal
-
-```
-
-## 6.2 Arrangement of CLEAR-CLiP Dataset [@moore_mirnatarget_2015]
-
-CLASH dataset was retrieved from [Nature article](nature.com/articles/ncomms9864)
-
-```{r,  eval=FALSE}
-clearclip <- read_xlsx("CLEAR-CLIP.xlsx")
-
-#Clearclip hg18
-```
-
-### Query of Human Genome 18
-
-```{r,  eval=FALSE}
-#HG18
-listEnsemblArchives()
-listMarts(host = 'may2009.archive.ensembl.org')
-ensembl54 = useMart(host='may2009.archive.ensembl.org', 
-                     biomart='ENSEMBL_MART_ENSEMBL',
-                    dataset='hsapiens_gene_ensembl')
-
-hg18 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene", "strand" ), mart = ensembl54)
-```
-
-### Adding Genome Information to dataset
-
-```{r,  eval=FALSE}
-clearclipfinal <- hg18%>%
-  inner_join(clearclip, by= c("entrezgene"= "gene.id", "hgnc_symbol"= "gene.symbol"))%>%
-  distinct()
-```
-
-### Converting human genome build
-
-```{r,   eval=FALSE}
-
-#Obtaining chromosomal locations from miRNA:target interaction dataset.
-
-lift18 <- clearclipfinal%>%
-  unite(start_end, c("start_position", "end_position"), sep = "-")%>%
-  unite(location, c("chr","start_end"), sep = ":")%>%
-  dplyr::select(location)
-
-write_tsv(lift18, "lift18.txt")
-
-#After we searched this file in UCSC browser, the output loaded (deleted_lift18 deleted regions on the HG38 genome build; hg38clearclip new locations of the genes on HG38)
-
-deleted_lift18 <- read_tsv("deleted_lift18.txt")
-
-colnames(deleted_lift18)[1] <- "Chromosome_loc"
-
-deleted_lift18 <- deleted_lift18%>%
-  dplyr::filter(startsWith(Chromosome_loc, "chr"))%>%
-  separate(Chromosome_loc, c("Chr", "End"), "-", remove = TRUE)%>%
-  separate(Chr, c("Chr", "Start"), ":", remove = TRUE)
-
-deleted_lift18$Start <- as.numeric(deleted_lift18$Start)
-deleted_lift18$End <- as.numeric(deleted_lift18$End)
-
-#removing deleted location from CLEAR-CLiP dataset
-
-clearclipfinal <- clearclipfinal%>%
-  dplyr::anti_join(deleted_lift18, by = c("chr"="Chr", "start_position"="Start", "end_position"="End"))
-
-hg38clearclip<- read.delim("hg38clearclip.txt", header = FALSE, stringsAsFactors = FALSE)
-
-clearclipfinal <- clearclipfinal%>% 
-  bind_cols(hg38clearclip)
-
-colnames(clearclipfinal)[28] <- "HG38build_loc"
-
-# adding new location information:
-
-clearclipfinal <- clearclipfinal%>%
-  dplyr::mutate(Genom_build= rep("hg18"))
-
-# Arrangement in dataset
-
-clearclipfinal%>%
-  dplyr::select(cluster=cluster.ID, chromosome = chr, start_position, end_position, strand = strand.y, hgnc_symbol, Ensembl_Gene_Id =ensembl_gene_id, Ensembl_Transcript_Id =ensembl_transcript_id, target_seq = target.map, miRNA, miR_seq = miR.map, seed_type = 'seed match', Energy = MFE, HG38build_loc, Genom_build, region) -> clearclipfinal
-
-```
-
-### Seed type manipulation in CLEAR-CLiP dataset
-
-In CLEAR-CLiP dataset, seed types were shown in detail. We adjusted as canonical and non-canonical.
-
-```{r,  eval=FALSE}
-data_frame(seed_type= c("5mer_1","5mer_2", "5mer_3", "6mer", "6mer.indel", "6mer.mm", "6mer_off.mm", "6merA1", "6merA1.indel", "6merA1.mm", "7merA1", "7merA1.indel", "7merA1.mm", "7merm8", "7merm8.indel", "7merm8,mm", "8mer", "8mer.indel", "8mer.mm", "NA" ),
-           seed_type_com= c("5-mer", "5-mer_noncanonical", "5-mer_noncanonical", "6-mer", "6-mer_noncanonical", "6-mer_noncanonical", "6-mer_noncanonical", "6-merA1", "6-merA1_noncanonical","6-merA1_noncanonical", "7-merA1", "7-merA1_noncanonical","7-merA1_noncanonical", "7-mer-8m", "7-mer-8m_noncanonical", "7-mer-8m_noncanonical", "8-mer", "8-mer_noncanonical", "8-mer_noncanonical", "none" )) ->clipdata_seed
-
-clearclipfinal%>%
-  inner_join(clipdata_seed, by = "seed_type")%>%
-  dplyr::select(1:11, seed_type = seed_type_com, Energy, HG38build_loc, Genom_build, region)-> clearclipfinal
-
-```
-
-## 6.3 Integration of two experimental dataset
-
-```{r,  eval=FALSE}
-
-bind_rows(clashelwakfinal, clearclipfinal)%>%distinct() -> experimentalmirnagene
-
-```
-
-### Adding Coefficients of Interaction factors
-
-Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study.
-On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study.
-Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions.
-Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values.
-We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
-With this process, we have handled this integrated dataset in context of competitor behaviors and functionality of interactions.
-
-In this step we added numeric intraction values at followings
-
-Fistly, we organized these values due to the fact that the regions were defined differently in two datasets. After that, region effect was added as numeric values (shown in Table S3).
-
-```{r,  eval=FALSE}
-experimentalmirnagene <- experimentalmirnagene%>%
-  mutate(region2  = str_replace_all(region, "NA", ""), region3 = str_replace_all(region2, "\\|", ""), region = str_replace_all(region3, c("3'UTR"="3UTR", "5'UTR"="5UTR")))%>%
-  dplyr::select(-region2, -region3)%>%
-  mutate(region_effect = as.double(ifelse(region %in% c("3UTRCDS", "CDS3UTR", "5UTR3UTR", "CDS5UTR3UTR", "CDS3UTRintron"), "0.93", ifelse(region %in% c("CDS", "CDSintron"), "0.42", ifelse(region %in% c("3UTR","3UTRintron"), "0.84", ifelse(region %in% c("5UTR", "5UTRintron"), "0.01", ifelse(region %in% c("5UTRCDS", "CDS5UTR"), "0.42", ifelse(region %in% c("intron",""), "0.01", ifelse(region %in% c("exon_unclassified", ""), "0.2", NA)))))))))
-
-```
-
-Secondly, we organized seed type interactions in *Seed type manipulation* section for CLEAR-CLiP dataset to show as found in CLASH dataset.
-Same type formatted values added dataset as numeric values (shown in Table S2).
-
-```{r,  eval=FALSE}
-
-seed_type_effect <- data_frame( seed_type = c("5-mer", "5-mer_noncanonical", "6-mer", "6-mer_noncanonical", "6-merA1", "6-merA1_noncanonical", "7-mer", "7-mer_noncanonical", "7-merA1", "7-merA1_noncanonical", "7-mer-8m", "7-mer-8m_noncanonical", "8-mer", "8-mer_noncanonical", "9-mer", "9-mer_noncanonical", "none"),
-                                seed_type_effect= c(0.05, 0.04, 0.07, 0.05, 0.07, 0.05, 0.23, 0.19, 0.19, 0.16, 0.25, 0.21, 0.43, 0.35, 0.43,0.35, 0.01)
-                                )
-
-experimentalmirnagene%>%
-  inner_join(seed_type_effect, by= "seed_type") -> experimentalmirnagene
-```
-
-```{r}
-readRDS("data/experimentalmirnagene.RDS")->experimentalmirnagene
-experimentalmirnagene
-```
-
-The context of dataset is shown in Table S5 in Supplementary Tables.
-
-# REFERENCES

### 2020-10-20 Supplementary_fig_tab.Rmd
adding revised files

@@
 header-includes: 
+  \usepackage{float} \floatplacement{figure}{H} 
+  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0}     \renewcommand{\thefigure}{S\arabic{figure}}}
+  \usepackage{subfig}
   \usepackage{tabularx}
-  \usepackage{float}
-  \floatplacement{figure}{H}
   \usepackage{caption}
-  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
 ---
@@
 
-# 1. Supplementary Figures
+# 1. Functions defined for ceRNA models and workflow of method
+
+We defined the functions that can be used with R programming.
+Briefly, these functions process a given miRNA:gene dataset and convert to graph object.
+All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
+The functions and steps of approach are explained as following (Figure S1) :
+
+**Conversion of dataset**: `priming_graph()` function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
+If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
+The formulations that are given in equations (1-4) are performed in this function.
+This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
+It is assumed that the initial target amounts in the dataset is observed after the repressive activity of miRNAs in steady-state.
+
+**Transition of variables in graph**: In the previous step, the calculations are performed in the edge variables of the graph object.
+However, the graph object allows to use node variables, while the node features are handled to the graph.
+In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.
+
+**Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficiency coefficients are calculated according to this acceptance.
+In the network that is found in steady-state conditions, the change is applied to the graph object for perturbation of steady-state.
+To provide the perturbation in the network the work-flow offer two methods: *update_variables* and *update_how*.
+The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
+The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
+
+**Simulation of competing behavior of targets**: After the change in the steady-state conditions, the network elements try to gain steady-state again.
+This process progresses as repeating of regulations after the spreading the changes in the network.
+In this step, simulation of regulations according to given cycle count in simulate function is applied.
+After each simulation cycle, the miRNA repression values are re-calculated and the current values of competing elements are found and saved.
+The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node data.
+
+The node elements in the dataset are handled as two type; repressive (miRNAs) and competing (targets).
+It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used.
+If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user.
+On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger.
+The functions that are used in the this study are developed with R and are available in [Bioconductor](https://www.bioconductor.org/packages/release/bioc/html/ceRNAnetsim.html).
+
+# 2. Supplementary Figures
+
+The codes used to obtain networks in this file can be found on the github page. Images that produced have been referred with codes.
 
@@
 
-## 1.1 *minsamp* dataset analysis in absence of interaction factors.
+## 2.1 Sample dataset analysis in absence of interaction factors.
 
 ```{r}
+# Sample dataset refers to minsamp in ceRNAnetsim package.
 data("minsamp")
@@
 
-```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset in Steady-state", fig.width=5, fig.height=4}
+```{r, message=FALSE, warning=FALSE,echo=FALSE, fig.cap= "Sample Dataset in Steady-state", fig.width=5, fig.height=4}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
@@
 
-```{r, message=FALSE, warning=FALSE, fig.keep='all', fig.cap= "Gene2 Upregulation on Minimal Dataset", fig.width=5, fig.height=4}
+```{r, message=FALSE, warning=FALSE,echo=FALSE,  fig.keep='all', fig.cap= "Gene2 Upregulation on Sample Dataset", fig.width=5, fig.height=4}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
@@
 
-```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of minsamp data.
+```{r, message=FALSE, warning=FALSE, echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of Sample data.
 A) First response of system to Gene2 upregulation (2nd iteration).
@@
 
-## 1.2 *minsamp* dataset analysis with interaction factors.
+## 2.2 Calculations with interaction factors
 
-```{r, message=FALSE, warning=FALSE, fig.cap= "Minimal Dataset with interaction factors in Steady-state", fig.width=5, fig.height=4}
+![Calculations to determine of miRNA binding and repression efficiency.](fig_s5.png){#figs5}
+
+$G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE^\prime$, normalized values of seed type efficiency coefficient; $RE^\prime$, normalized values of region efficiency coefficient; $E^\prime$, normalized values of energy coefficient.
+Numbering on edges match the pair order in Table S1.
+
+## 2.3 Sample+ dataset analysis with interaction factors.
+
+```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.cap= "Sample+ in Steady-state", fig.width=5, fig.height=4}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
@@
 
-```{r, message=FALSE, warning=FALSE, fig.cap= " When Gene2 is upregulated on Minimal Dataset with interaction factors", fig.width=5, fig.height=4}
+```{r, message=FALSE, warning=FALSE, echo=FALSE,  fig.cap= " When Gene2 is upregulated on Sample+", fig.width=5, fig.height=4}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
@@
 
-```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of minsamp data with interaction factors A)First response of system to Gene2 upregulation (2nd iteration).
+```{r, message=FALSE, warning=FALSE,echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of Sample+ A)First response of system to Gene2 upregulation (2nd iteration).
 B)Spreading of perturbation on system (3th iteration)"}
@@
 
-## 1.3 Common target perturbation in *minsamp* dataset.
+## 2.4 Common target perturbation in Sample+ dataset.
 
-```{r, message=FALSE, warning=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Perturbation of Gene4 on minsamp data with interaction factors.
+```{r, message=FALSE, warning=FALSE, echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Perturbation of Gene4 on Sample+.
 A) Network at steady-state.
@@
 
-![Gene analysis of transcription factor association and KEGG pathway, obtained from experimental miRNA:target dataset ](experimental_kegg.png){#figs9}
+![Sankey diagram represents top five KEGG and GO (molecular function and biological process) terms and genes enriched on these terms. Genes with single edge are not shown.](sankey_v2.png){#figs9}
 
-# 2 Supplementary Tables
+# 3 Supplementary Tables
 
@@
 
-## 2.1 *minsamp* dataset
+## 3.1 *Sample+* dataset
 
-|competing| miRNA | Competing_expression | miRNA_expression | seed_type | region | energy
+|Competing| miRNA | Competing Expression | miRNA Expression | Seed Type Effect | Region Effect | Energy
 |-----------|------------|----------|----------------|---------|-----------|----------|
@@
 
-Table: *minsamp* sample dataset that includes interaction factors. The parameters which affect miRNA:target interactions (i.e. seed type, region, energy) are provided in sample dataset.
+Table: The parameters which affect miRNA:target interactions (i.e. seed type, region, energy) are provided in Sample+ dataset, while these factors are not utilized in simulation of Sample dataset.
 
-## 2.2 Significant factors in miRNA:target interactions
+## 3.2 Significant factors in miRNA:target interactions
 
@@
 
-## 2.3 Content of High-throughput experimental studies
+## 3.3 Content of High-throughput experimental studies
 
@@
 
-## 2.4  Variables of network object during simulation
+## 3.4  Variables of network object during simulation
 
@@
   \hline
- Variables & Structure & Means \\
+ Variables & Structure & Description \\
   \hline
@@
 
-\newpage
+# 4. Notes
+
+To evaluate significant nodes through parallel processing in breast cancer patient network, perturbation on all nodes were triggered by 3-fold up-regulation with 10 iteration (cycle) due to small diameter of Real and Real+ networks.
+We analyzed perturbation efficiencies of all nodes in situation that accepts nodes with more than one percent change in expression as re-regulated (i.e, limit=1 argument in *find_node_perturbation()* function).
+
+All codes are available at [github repo](https://github.com/selcenari/ceRNA_multifactorial_model).
 

### 2020-10-20 Supplementary_mat_met.Rmd
adding revised files

@@
 
-# 1. Functions defined for ceRNA models and workflow of method
-
-We defined the functions that can be used with R programming.
-Briefly, these functions process a given miRNA:gene dataset and convert to graph object.
-All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
-The functions and steps of approach are explained as following (Figure S1 in Supplementary Figures) :
-
-**Conversion of dataset**: `priming_graph()` function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
-If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
-The formulations that are given in equations (1-4) are performed in this function.
-This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
-It is assumed that the initial target amounts in the dataset is observed after the repressive activity of miRNAs in steady-state.
-
-**Transition of variables in graph**: In the previous step, the calculations are performed in the edge variables of the graph object.
-However, the graph object allows to use node variables, while the node features are handled to the graph.
-In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.
-
-**Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficiency coefficients are calculated according to this acceptance.
-In the network that is found in steady-state conditions, the change is applied to the graph object for perturbation of steady-state.
-To provide the perturbation in the network the work-flow offer two methods: *update_variables* and *update_how*.
-The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
-The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
-
-**Simulation of competing behavior of targets**: After the change in the steady-state conditions, the network elements try to gain steady-state again.
-This process progresses as repeating of regulations after the spreading the changes in the network.
-In this step, simulation of regulations according to given cycle count in simulate function is applied.
-After each simulation cycle, the miRNA repression values are re-calculated and the current values of competing elements are found and saved.
-The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node data.
-
-The node elements in the dataset are handled as two type; repressive (miRNAs) and competing (targets).
-It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used.
-If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user.
-On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger.
 The functions that are used in the this study are developed with R and are available in [Bioconductor](https://www.bioconductor.org/packages/release/bioc/html/ceRNAnetsim.html).
 
+# 1. Determination of perturbation efficiencies of nodes in network.
 ```{r, include=FALSE, message=FALSE, warning=FALSE}
@@
 
-+ load *minsamp* data
-
-```{r}
-data("minsamp")
-minsamp
+```{r, include=FALSE}
+data("minsamp") # refers Sample and Sample+ data
 ```
 
-See Table S1 in Supplementary Figures and Tables file.
+## Node efficiencies on Sample+ network
 
-# 2. *minsamp* dataset analysis in lack of interaction factors.
-
-Firstly, we have analysed minimal data without interaction factors between miRNA:target.
-
-- 1. We have evaluated graph in the steady state conditions as followings (Figure S2 in Supplementary Figures):
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)
-```
-
-- 2. We have obtained graph after change on Gene2 expression as following (Figure S3 in Supplementary Figures):
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)
-```
-
-- 3. We have determined regulations after Gene2 upregulation (Figure S4 in Supplementary Figures):
-
-```{r, message=FALSE, warning=FALSE}
-
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)%>%
-  simulate(cycle= 2)
-
-```
-
-Note that the regulations are colored according to expression changes of present and a previous value.
-So, it can be observed that whole gene expressions increase in comparison of initial steady-state.
-The overall regulations of gene expressions are as followings:
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)%>%
-  simulate(2)%>%
-  activate(edges)%>%
-  as_tibble()%>%
-  select(Competing_name,comp_count_list,effect_list)%>%
-  unnest()
-```
-
-# 3. *minsamp* dataset analysis with interaction factors.
-
-We have made the same analysis in presence of interaction factors (Sequentially shown at Figure S5-7 in Supplementary Figures).
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)
-
-```
-
-```{r, message=FALSE, warning=FALSE}
-
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene2", 2)
-```
-
-```{r, message=FALSE, warning=FALSE}
-
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene2", 2)%>%
-  simulate(cycle=2)
-```
-
-When the graphs were examined, it was observed that behaviors (i.e.
-fluctuating trend) were same.
-But, when the results were analysed in terms of expression values, the regulation differences can be observed.
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene2", 2)%>%
-  simulate(3)%>%
-  activate(edges)%>%
-  as_tibble()%>%
-  select(Competing_name,comp_count_list,effect_list)%>%
-  unnest()
-```
-
-# 4. Common target perturbation in *minsamp* dataset.
-
-Genes targeted by multiple miRNAs (referred to as "common target") are of special interest since they are subject to cooperative effect.
-Also, they perturb more than one neighborhood.
-In our small dataset, minsamp, Gene4 is regulated by two miRNAs.
-Let's simulate perturbation effects triggered by Gene4 (Shown at Figure S8 in Supplementary Figures) .
-
-```{r, message=FALSE, warning=FALSE}
-
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene4", 2)%>%
-  simulate(cycle = 2)
-```
-
-The common target perturbation  (increasing to two fold at Gene4 expression in presence of interaction factors) resulted in more prominent efficiency at the same conditions (shown at following).
-
-```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene4", 2)%>%
-  simulate(3)%>%
-  activate(edges)%>%
-  as_tibble()%>%
-  select(Competing_name,comp_count_list,effect_list)%>%
-  unnest()
-```
-
-# 5. Determination of perturbation efficiencies of elements in system.
-
-```{r, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
 
@@
 
-# 6. Additional data manipulation steps
+# 2. Additional data manipulation steps
 
-## 6.1 Arrangement of CLASH dataset
+## 2.1 Arrangement of CLASH dataset
 
@@
 
-## 6.2 Arrangement of CLEAR-CLiP Dataset [@moore_mirnatarget_2015]
+## 2.2 Arrangement of CLEAR-CLiP Dataset [@moore_mirnatarget_2015]
 
@@
 
-## 6.3 Integration of two experimental dataset
+## 2.3 Integration of two experimental dataset
 
@@
                                 )
-
 experimentalmirnagene%>%

### 2020-10-26 figure_manuscript/codes_of_figures.Rmd
Reproducible rmd files and datasets for figs

---
title: "Main Figures"
author:
- Selcen Ari
- Alper Yilmaz
date: "`r Sys.Date()`"
output: pdf_document
---

# Figures

This file consists analysis steps and codes of figures in manuscript.

```{r, include=FALSE, message=FALSE, warning=FALSE}
library(janitor)
library(tidyverse)
library(purrr)
library(readr)
library(biomaRt)
library(TCGAbiolinks)
library(ceRNAnetsim)
library(dplyr)
library(ggVennDiagram)
library(ggraph)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
#knitr::opts_chunk$set(fig.pos = 'H')
knitr::opts_chunk$set(cache=FALSE)
```

```{r}

mirtarbase_node_perturbations <- readRDS("mirtarbase_node_perturbations.RDS") # data includes results of find_node_perturbation() function for tumor and normal samples of each patient.

```

## Venn diagrams **Figure 4A**

```{r}
mirtarbase_node_perturbations%>%
  group_by(name)%>%
  summarise(perturbed = sum(perturbed_count >0, na.rm = TRUE))%>%
  ungroup()->gene_dist

mirtarbase_node_perturbations%>%
  left_join(gene_dist, by ="name")%>%
  filter(perturbed> 10, perturbed_count > 78)%>%
  distinct(name)->effective_nodes
```

+ Detecting highly perturbing gene nodes in tumor tissues

```{r}
mirtarbase_node_perturbations%>%
  left_join(gene_dist, by = "name")%>%
  mutate(tissue_type = ifelse(endsWith(file_name, "01A"), "Tumor", "Normal"),
         node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>% 
  filter(tissue_type != "Normal", startsWith(name, "ENS"))%>%
  distinct(name)%>%
  pull()->lim_1_gene_tumor
```
+ Detecting highly perturbing gene nodes in normal tissues

```{r}
mirtarbase_node_perturbations%>%
  left_join(gene_dist, by = "name")%>%
  mutate(tissue_type = ifelse(endsWith(file_name, "01A"), "Tumor", "Normal"),
         node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>% 
  filter(tissue_type == "Normal", startsWith(name, "ENS"))%>%
  distinct(name)%>%
  pull()->lim_1_gene_normal
```

### Venn diagram of genes

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=8, fig.height=8}
x_gene = list(" "= lim_1_gene_tumor, " " = lim_1_gene_normal)

pa1 <- ggVennDiagram(x_gene, label = "count", category.names = c("Tumor","Normal"))+
  scale_fill_gradient(low='bisque1',high = 'lightsalmon')+
  theme(legend.position="none", text = element_text(""), plot.margin = margin(0, 0, 0, 0, "mm"))

pa1

```

+ Detecting highly perturbing mirna nodes in tumor tissues

```{r}

mirtarbase_node_perturbations%>%
  left_join(gene_dist, by = "name")%>%
  mutate(tissue_type = ifelse(endsWith(file_name, "01A"), "Tumor", "Normal"),
         node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>% 
  filter(tissue_type != "Normal", startsWith(name, "hsa"))%>%
  distinct(name)%>%
  pull()->lim_1_mirna_tumor

```

+ Detecting highly perturbing mirna nodes in normal tissues

```{r}

mirtarbase_node_perturbations%>%
  left_join(gene_dist, by = "name")%>%
  mutate(tissue_type = ifelse(endsWith(file_name, "01A"), "Tumor", "Normal"),
         node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>% 
  filter(tissue_type == "Normal", startsWith(name, "hsa"))%>%
  distinct(name)%>%
  pull()->lim_1_mirna_normal

```

```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=8, fig.height=8}
x_mirna = list(" "= lim_1_mirna_tumor, " " = lim_1_mirna_normal)

pa2 <- ggVennDiagram(x_mirna, label = "count", category.names = c("Tumor","Normal"))+
  scale_fill_gradient(low='lightskyblue1',high = 'lightskyblue3')+
  theme(legend.position="none", text = element_text(""), plot.margin = margin(0, 0, 0, 0, "cm"))

pa2
```
## Functional Annotation Analysis

```{r}
Anno_tumor_specific <- readRDS("Anno_tumor_specific.RDS") # includes functional annotation of 29 genes found as tumor-specific.
```

#Fuctional annotation of Tumor specific 29 genes. **Figure 4B**

## Tumor only:

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=10,fig.show='hold'}
Anno_tumor_specific%>%
  filter(startsWith(Category, c("KEGG", "GO")), !str_detect(Category, "CC_DIRECT"))%>%
  filter(!str_detect(Category, "MF_DIRECT"))%>%
  clean_names()%>%
  group_by(category)%>%
  filter(row_number()<6)->tumor_only_annotation

position <- rev(c("hsa05219:Bladder cancer", "hsa05212:Pancreatic cancer", "hsa04510:Focal adhesion", "hsa05206:MicroRNAs in cancer", "hsa05200:Pathways in cancer", "GO:0071230~cellular response to amino acid stimulus", 	"GO:0043154~negative regulation of cysteine-type endopeptidase activity involved in apoptotic process", "GO:0001701~in utero embryonic development", 	"GO:0043066~negative regulation of apoptotic process", "GO:0000122~negative regulation of transcription from RNA polymerase II promoter"))

tumor_only_annotation%>%
  group_by(category)%>%
  arrange(-desc(fold_enrichment))

tumor_only_annotation%>%
  ggplot(aes(x = fold_enrichment, y = term))+
  geom_col(aes(color = category, fill= category))+
  geom_text(aes(label=round(p_value, 5)), hjust = 0.9, color ="darkslategrey", size =3)+
  xlab("Fold Enrichment")+
  ylab("Annotation Term")+
  theme(panel.grid = element_blank(), axis.title = element_text(size =20), axis.text = element_text(size = 12),legend.position = "none")+
  scale_y_discrete(labels=scales::wrap_format(40),position = "left", limits = position)

```

## Network construction with overall functional annotation: **Figure 3**

```{r}

significant_node_graph <- readRDS("significant_node_graph.RDS") #consists network of highly perturbing nodes.
hg19 <- readRDS("hg19.RDS") #For id matching (obtained via bioMaRt package)

```

## Annotation network:

```{r}

Annotation_overall <- readRDS("Annotation_overall.RDS") # consists functional annotation of 145 high perturbing nodes.

Annotation_overall%>%
  filter(startsWith(Category, c("KEGG", "GO")), !str_detect(Category, "CC_DIRECT"))%>%
  filter(!str_detect(Category, "MF_DIRECT"))%>%
  clean_names()%>%
  group_by(category)%>%
  filter(row_number()<6)%>%
  dplyr::select(term, Gene_Count = count, p_value, genes)%>%
  mutate(genes2 = str_split(genes, ", "))%>%
  unnest()%>%
  ungroup%>%
  mutate(genes2 = str_trim(genes2, side = "both"), 
         genes2 = ifelse(endsWith(genes2, ","), substr(genes2, 0, (nchar(genes2)-1)), genes2),
         interaction_type = term, 
         Ensembl_Gene_Id = genes2, 
         competing_name = genes2)%>%
  dplyr::select(interaction= term, Ensembl_Gene_Id, interaction_type, competing_name)->gene_top_annotation_graph

```

```{r, warning=FALSE}

significant_node_graph%>%
  bind_rows(gene_top_annotation_graph)%>%
  as_tbl_graph()%>%
  mutate(type = ifelse(startsWith(name, "hsa"), "KEGG", "GO"),
         type = ifelse(startsWith(name, "hsa-"), "miRNA", type),
         type = ifelse(startsWith(name, "ENSG"), "Gene", type))%>%
  mutate(centrality= centrality_degree(mode = "all"),
         annotation = ifelse(name %in% gene_top_annotation_graph$Ensembl_Gene_Id, "top_annotation", "other"))%>%
  as_tibble()%>%
  filter(annotation == "top_annotation", centrality >10)%>% 
  dplyr::select(name)%>%
  pull()->top_important

significant_node_graph%>%
  bind_rows(gene_top_annotation_graph)%>%
  as_tbl_graph()%>%
  mutate(type = ifelse(startsWith(name, "hsa"), "KEGG", "GO"),
         type = ifelse(startsWith(name, "hsa-"), "miRNA", type),
         type = ifelse(startsWith(name, "ENSG"), "Gene", type))%>%
  mutate(centrality= centrality_degree(mode = "all"),
         annotation = ifelse(name %in% gene_top_annotation_graph$Ensembl_Gene_Id, "Enriched Genes", "other"))%>%
  left_join(hg19, by =c("name"="ensembl_gene_id"))%>%
  filter( type %in% c("miRNA", "Gene"))%>%
  ggraph(layout = "kk")+ 
  geom_edge_link(colour = "seashell2")+
  geom_node_point(color = "seashell2")+
  geom_node_point(aes(filter= annotation == "Enriched Genes", size = centrality, color = "Enriched Genes"), shape = 16)+
  geom_node_point(aes(filter= type == "miRNA", color = "miRNA"), shape = 17, size = 2)+
  geom_node_point(aes(filter= type == "GO", size = centrality, color = "GO"), shape = 18)+
  geom_node_point(aes(filter= type == "KEGG", size = centrality, color = "KEGG"), shape = 15)+
  theme_graph(base_family = "sans")+
  geom_node_text(aes(filter= name %in% top_important, label = hgnc_symbol), size =3.5, repel = TRUE)+
  theme(plot.margin = margin(0, 0, 0, 0, "cm"), legend.position = "none")+
  guides(size=FALSE, shape = FALSE)+
  scale_color_manual(name = "Node types", values = c("Enriched Genes" = "sandybrown", "miRNA"= "dodgerblue2", "GO"= "green", "KEGG"= "red"))

```

# Heat-maps for miRNAs

+Detecting perturbing miRNAs

```{r, warning=FALSE}
mirtarbase_node_perturbations%>%
  left_join(gene_dist, by = "name")%>%
  mutate(tissue_type = ifelse(endsWith(file_name, "01A"), "Tumor", "Normal"),
         node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>%
  filter(tissue_type != "Normal", startsWith(name, "hsa"))%>%
  dplyr::select(file_name, name,tissue_type, perturbed_count)->mirna_tumor

mirtarbase_node_perturbations%>%
  left_join(gene_dist, by = "name")%>%
  mutate(tissue_type = ifelse(endsWith(file_name, "01A"), "Tumor", "Normal"),
         node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>%
  filter(tissue_type == "Normal", startsWith(name, "hsa"))%>%
  dplyr::select(file_name, name,tissue_type, perturbed_count)->mirna_normal

```

## Heat map Figure 4C

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=9,fig.show='hold'}
mirna_tumor%>%
  full_join(mirna_normal, by =c("name", "file_name"), suffix= c("_tumor", "_normal"), fill.na = 0)%>%
  dplyr::select(-3, -5)%>%
  mutate(perturbed_count_tumor = ifelse(is.na(perturbed_count_tumor), 1, perturbed_count_tumor),
         perturbed_count_normal = ifelse(is.na(perturbed_count_normal), 1, perturbed_count_normal),
         log_FC= log2(perturbed_count_tumor/perturbed_count_normal))%>%
  ggplot(aes(x = file_name, y = name, fill = log_FC))+
  geom_tile(aes(colour = log_FC, fill = log_FC))+coord_fixed(ratio = 1)+
  theme_test()+
  theme(axis.text.x = element_blank(), #element_text(angle = 90, vjust = 0, hjust=0)
        plot.title = element_text(hjust = 0.5, size =20), 
        axis.title = element_text(size =20))+
  scale_colour_gradientn(colours = c("green","white", "red"), aesthetics = c("colour", "fill"), na.value = "grey")+
  ylab("Highly pertubing miRNA nodes")+
  xlab("Patients")+
  theme( legend.title = element_blank(), panel.background = element_rect(fill = "white"), plot.margin = margin(0, 0, 0, 0, "cm"))

```

## Heat map Figure 4D

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=10, fig.show='hold'}

mirna_tumor%>%
  full_join(mirna_normal, by =c("name", "file_name"), suffix= c("_tumor", "_normal"), fill.na = 0)%>%
  dplyr::select(-3, -5)%>%
  pivot_longer(cols = c("perturbed_count_tumor", "perturbed_count_normal"), names_to = "situ", values_to = "perturbed_count")%>%
  mutate(perturbed_count_log = log2(perturbed_count),
         situ = str_replace(situ, "perturbed_count_normal", 'Perturbed count in normal'),
         situ = str_replace(situ, "perturbed_count_tumor", 'Perturbed count in tumor'),
         perturbed_count = as.numeric(str_replace_na(perturbed_count, 0)),
         perturbed_count_log = as.numeric(str_replace_na(perturbed_count_log, 0)))%>%
  ggplot(aes(x = file_name, y = name, fill = perturbed_count_log))+
  geom_tile(aes(colour = perturbed_count_log, fill = perturbed_count_log))+coord_fixed(ratio = 1)+
  theme_test()+
  theme(axis.text.x = element_blank(), #element_text(angle = 90, vjust = 0, hjust=0)
        plot.title = element_text(hjust = 0.5, size =20), 
        axis.title = element_text(size =20))+
  scale_colour_gradientn(colours = c("green","red", "black"), aesthetics = c("colour", "fill"), na.value = "grey")+
  ylab("Highly pertubing miRNA nodes")+
  xlab("Patients")+
  facet_grid(rows = "situ")+
  theme( legend.title = element_blank(), panel.background = element_rect(fill = "gray93"), plot.margin = margin(0, 0, 0, 0, "cm"))

```

### 2020-10-30 Supplementary_fig_tab.Rmd
update supplements

@@
 bibliography: NAR-sample.bib
+urlcolor: blue
 header-includes: 
@@
 library(stringr)
+library(ggraph)
+library(ggplot2)
 knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE, echo=FALSE}
 library(ceRNAnetsim)
@@
 
-```{r}
+```{r, echo=FALSE}
 # Sample dataset refers to minsamp in ceRNAnetsim package.
@@
 ## 2.3 Sample+ dataset analysis with interaction factors.
+
+Interaction factors of Sample+ network are available in Table S1.
 
@@
 ![Sankey diagram represents top five KEGG and GO (molecular function and biological process) terms and genes enriched on these terms. Genes with single edge are not shown.](sankey_v2.png){#figs9}
+
+```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=12, fig.height=5,fig.show='hold', fig.cap = "Tumor-specific perturbing nodes log2 transformed expression levels in tumor and normal tissue samples of 87 patients."}
+
+matched_patient_exp_dataset <- readRDS("figures_manuscript/matched_patient_exp_dataset.RDS")
+mirtarbase_gene_res <- readRDS("figures_manuscript/mirtarbase_gene_res.RDS")
+
+matched_patient_exp_dataset%>%
+  filter(Ensembl_Gene_Id %in% pull(filter(mirtarbase_gene_res, situ == "tumor_specific"), Ensembl_Gene_Id))%>%
+  pivot_longer(cols = contains("TCGA"), names_to="patient_id", values_to = "expressions")%>%
+  mutate(patient = ifelse(endsWith(patient_id, "01A"), "tumor", "normal"))%>%
+  left_join(mirtarbase_gene_res, by ="Ensembl_Gene_Id")%>%
+  group_by(patient, Ensembl_Gene_Id)%>%
+  ggplot(aes(x= hgnc_symbol, y=log2(expressions), fill = patient))+
+  #scale_y_log10()+
+  geom_boxplot(position = "dodge")+
+  theme_bw()+
+  theme(axis.text.x = element_text(angle = 90), panel.background = element_rect())+
+  xlab("")
+```
 
@@
 
-|  Variable | Structure | Means 
+|  Variable | Structure | Definition 
 |-----------|-----------|-------|
@@
 
-Table: miRNA:target pairs supported by High-throughput Experiments
+Table: Context of miRNA:target pairs supported by High-throughput Experiments. CLEAR-CLiP and CLASH datasets were integrated as described in Section 2 of Supplementary Material and Method.
 

### 2020-10-30 figure_manuscript/codes_of_figures.Rmd
figure code update

@@
 
-pa1 <- ggVennDiagram(x_gene, label = "count", category.names = c("Tumor","Normal"))+
+pa1 <- ggVennDiagram(x_gene, label = "both", category.names = c("Tumor","Normal"))+
   scale_fill_gradient(low='bisque1',high = 'lightsalmon')+
@@
 pa1
+
+ggsave(filename = 'lim1_PE_gene_mirtarbase.eps', device=cairo_ps, width = 6, height = 3)
 
@@
 ```
+
+```{r, eval=FALSE}
+library(ggpubr)
+
+#pA <- grid.arrange(pa1, pa2)
+
+pA <- ggpubr::ggarrange(pa1, pa2, nrow = 2)
+pA
+#ggsave(filename = 'lim1_PE_gene_mirtarbase.svg', width = 6, height = 6)
+```
+
 ## Functional Annotation Analysis

### 2020-11-02 Supplementary_fig_tab.Rmd
updated supplements

@@
 
-# 1. Functions defined for ceRNA models and workflow of method
-
-We defined the functions that can be used with R programming.
-Briefly, these functions process a given miRNA:gene dataset and convert to graph object.
-All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
-The functions and steps of approach are explained as following (Figure S1) :
-
-**Conversion of dataset**: `priming_graph()` function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
-If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
-The formulations that are given in equations (1-4) are performed in this function.
-This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
-It is assumed that the initial target amounts in the dataset is observed after the repressive activity of miRNAs in steady-state.
-
-**Transition of variables in graph**: In the previous step, the calculations are performed in the edge variables of the graph object.
-However, the graph object allows to use node variables, while the node features are handled to the graph.
-In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.
-
-**Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficiency coefficients are calculated according to this acceptance.
-In the network that is found in steady-state conditions, the change is applied to the graph object for perturbation of steady-state.
-To provide the perturbation in the network the work-flow offer two methods: *update_variables* and *update_how*.
-The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
-The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
-
-**Simulation of competing behavior of targets**: After the change in the steady-state conditions, the network elements try to gain steady-state again.
-This process progresses as repeating of regulations after the spreading the changes in the network.
-In this step, simulation of regulations according to given cycle count in simulate function is applied.
-After each simulation cycle, the miRNA repression values are re-calculated and the current values of competing elements are found and saved.
-The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node data.
-
-The node elements in the dataset are handled as two type; repressive (miRNAs) and competing (targets).
-It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used.
-If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user.
-On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger.
-The functions that are used in the this study are developed with R and are available in [Bioconductor](https://www.bioconductor.org/packages/release/bioc/html/ceRNAnetsim.html).
-
 # 2. Supplementary Figures
-
-The codes used to obtain networks in this file can be found on the github page. Images that produced have been referred with codes.
 
@@
 
-```{r, message=FALSE, warning=FALSE,echo=FALSE, fig.cap= "Sample Dataset in Steady-state", fig.width=5, fig.height=4}
+```{r, message=FALSE, warning=FALSE,echo=FALSE, fig.cap= "Sample Dataset in Steady-state.
+Inıtial expression levels in minsamp network (*Sample* network in manuscript).
+The network contains two miRNAs and 6 genes with arbitrarily chosen expression values.
+Refer to Table S1 for exact expression values", fig.width=5, fig.height=4}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
   vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4",
-            title = " ")
+            title = " ")+
+  geom_node_text(aes(label = name), repel = TRUE, size =3)
 ```
 
-```{r, message=FALSE, warning=FALSE,echo=FALSE,  fig.keep='all', fig.cap= "Gene2 Upregulation on Sample Dataset", fig.width=5, fig.height=4}
+```{r, message=FALSE, warning=FALSE,echo=FALSE,  fig.keep='all', fig.cap= "Gene2 Upregulation on Sample Dataset.
+Expression level of Gene2 is increased from 10,000 to 20,000 in order to demonstrate the calculation steps.", fig.width=5, fig.height=4}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
@@
   vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", 
-            title = " ")
+            title = " ")+
+  geom_node_text(aes(label = name), repel = TRUE, size =3)
 ```
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of Sample data.
+```{r, message=FALSE, warning=FALSE, echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of Sample data after up-regulation of Gene2.
 A) First response of system to Gene2 upregulation (2nd iteration).
@@
             Downregulation = "blue", 
-            title = "A")
+            title = "A")+
+  geom_node_text(aes(label = name), repel = TRUE, size =3)
 
@@
             Downregulation = "blue", 
-            title = "B")
+            title = "B")+
+  geom_node_text(aes(label = name), repel = TRUE, size =3)
 
@@
 
-![Calculations to determine of miRNA binding and repression efficiency.](fig_s5.png){#figs5}
+```{r, echo=FALSE, fig.cap="Calculation of initial miRNA repression activity using interaction parameters in Sample+ network.
+A) Interaction paramaters between genes and miRNAs in Sample+ are shown on network while expression levels can be found in Table S1.
+B) Interaction paramaters were updated after normalisation C) Amount of miRNA distributed to each mRNA according to mRNA levels and affinity parameters are shown on edges.
+D) Values on edges indicate repression activity.
+Red values indicate repression activity affected by region effect (RE) parameter.
+E) Total repression on G4 from two miRNAs is calculated by adding repression values originating from both miRNAs.", fig.show='hold', out.height= "70%", out.width="70%", fig.align = "center"}
+knitr::include_graphics("fig_s5.png")
+
+```
 
@@
 
-Interaction factors of Sample+ network are available in Table S1.
-
-```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.cap= "Sample+ in Steady-state", fig.width=5, fig.height=4}
+```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.cap= "Sample+ in Steady-state. Interaction factors of Sample+ network are available in Table S1.", fig.width=5, fig.height=4}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
   vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", 
-            title = " ")
+            title = " ")+
+  geom_node_text(aes(label = name), repel = TRUE, size =3)
 
@@
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE,  fig.cap= " When Gene2 is upregulated on Sample+", fig.width=5, fig.height=4}
+```{r, message=FALSE, warning=FALSE, echo=FALSE,  fig.cap= "Perturbation in Sample+ network by two-fold increase in Gene2 expression level.", fig.width=5, fig.height=4}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
@@
   vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", 
-            title = " ")
+            title = " ")+
+  geom_node_text(aes(label = name), repel = TRUE, size =3)
 ```
 
-```{r, message=FALSE, warning=FALSE,echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of Sample+ A)First response of system to Gene2 upregulation (2nd iteration).
-B)Spreading of perturbation on system (3th iteration)"}
+```{r, message=FALSE, warning=FALSE,echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of Sample+ A) First response of system to Gene2 upregulation (2nd iteration).
+B) Spreading of perturbation on system (3rd iteration).
+Although visualisation looks similar to Figure S4B, current counts of genes are drastically different."}
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
@@
   simulate(cycle = 1)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "A")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "A")+
+  geom_node_text(aes(label = name), repel = TRUE, size =3)
 
@@
   simulate(cycle = 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "B")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "B")+
+  geom_node_text(aes(label = name), repel = TRUE, size =3)
 
@@
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Perturbation of Gene4 on Sample+.
+```{r, message=FALSE, warning=FALSE, echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Perturbation of Gene4 and its effects on Sample+.
 A) Network at steady-state.
@@
 priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "A")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "A")+
+  geom_node_text(aes(label = name), repel = TRUE, size =3)
 
@@
   update_how("Gene4", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "B")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "B")+
+  geom_node_text(aes(label = name), repel = TRUE, size =3)
 
@@
   simulate(cycle=1)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "C")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "C")+
+  geom_node_text(aes(label = name), repel = TRUE, size =3)
 
@@
   simulate(cycle=2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "D")
+  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "D")+
+  geom_node_text(aes(label = name), repel = TRUE, size =3)
 ```
@@
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=12, fig.height=5,fig.show='hold', fig.cap = "Tumor-specific perturbing nodes log2 transformed expression levels in tumor and normal tissue samples of 87 patients."}
+```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=12, fig.height=5,fig.show='hold', fig.cap = "Log2 transformed expression levels of tumor-specific perturbing nodes in tumor and normal tissue samples of 87 patients."}
 
@@
 
-|  Variable | Structure | Definition 
-|-----------|-----------|-------|
-|  cluster    | character | Barcode from experimentally method |
-|  chromosome    | character | Chromosome of Target gene from raw data |
-|   start_position    | numeric | Gene start position from raw data |
-|   end_position     | numeric | Gene end position from raw data |
-|   strand     | character | Gene strand |
-|   hgnc_symbol     | character | Gene name (Symbol) |
-|   Ensembl_Gene_Id     | character | Ensembl Gene Id of gene |
-|   Ensembl_Transcript_Id   | character | Ensembl transcript id of mRNA of Target gene |
-|   target_seq   | character | mRNA sequences targeted by miRNA |
-|   miRNA   | character | miRNA id (from miRBase version 21 ) |
-|   miR_seq   | character | miRNA sequence |
-|   seed_type   | character | seed type of miRNA:target interaction |
-|   Energy   | numeric | Energy of miRNA:target binding |
-|   HG38build_loc   | character | Recent chromosomal location of Gene |
-|   Genome_build   | character| Genome build of given chromosome, start and end positions |
-|   region   | character | interaction location on target | 
-|   region_effect   | numeric | Coefficient of location on target | 
-|   seed_type_effect   | numeric | Coefficient for seed sequence of miRNA:target interaction |
+|  Variable | Definition 
+|-----------|-------|
+|  cluster    | Barcode from experimental method |
+|  chromosome    | Chromosome of Target gene from raw data |
+|   start_position   | Gene start position from raw data |
+|   end_position    | Gene end position from raw data |
+|   strand    | Gene strand |
+|   hgnc_symbol     | Gene name (Symbol) |
+|   Ensembl_Gene_Id     | Ensembl Gene Id of gene |
+|   Ensembl_Transcript_Id   | Ensembl transcript id of mRNA of Target gene |
+|   target_seq   | mRNA sequences targeted by miRNA |
+|   miRNA   | miRNA id (from miRBase version 21 ) |
+|   miR_seq   | miRNA sequence |
+|   seed_type  | seed type of miRNA:target interaction |
+|   Energy   | Energy of miRNA:target binding |
+|   HG38build_loc    | Recent chromosomal location of Gene |
+|   Genome_build   | Genome build of given chromosome, start and end positions |
+|   region   | interaction location on target | 
+|   region_effect   | Coefficient of location on target | 
+|   seed_type_effect | Coefficient for seed sequence of miRNA:target interaction |
 
@@
 
-As a result of simulation a dataset, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at following.
+As a result of simulation a dataset, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at Table S5.
 
@@
 \caption{ The context graph object during the process.}
-\begin{tabularx}{\textwidth}{llX}
+\begin{tabularx}{\textwidth}{lX}
   \hline
- Variables & Structure & Description \\
+ Variables & Description \\
   \hline
-   \emph{Node Variables} & & \\
-   name    & character & node name \\
-   type    & character &Competing or miRNA \\
-   node\_id    & numeric & in on graph object \\
-   initial\_count     & numeric & Initial Expression value of node \\
-   count\_pre     & numeric & Expression value of node at previous regulation \\
-   count\_current     & numeric & Existing expression value of node \\
-   changes\_variable     & character & Regulation of node (Up, down or steady) \\
+   \emph{Node Variables} &  \\
+   name    &  node name \\
+   type    & Competing or miRNA \\
+   node\_id    &  in on graph object \\
+   initial\_count     & Initial Expression value of node \\
+   count\_pre     &  Expression value of node at previous regulation \\
+   count\_current     &  Existing expression value of node \\
+   changes\_variable     &  Regulation of node (Up, down or steady) \\
    \hline
-   \emph{Edge Variables} & & \\
-   Competing name   & character & name of genes\\
-   miRNA name   & character & name of miRNAs \\
-   Competing expression   & numeric & Expression values of competing elements at steady-state \\
-   miRNA expression   & numeric & Expression values of miRNA elements at steady-state \\
-   energy   & numeric & coefficient of miRNA:target interactions (binding affinity) \\
-   seed type   & numeric & coefficient of miRNA:target interactions (binding affinity) \\
-   region   & numeric & coefficient of miRNA:target interactions (degradation efficiency) \\
-   afff factor   & numeric& coefficient scaled and combined affinity factor \\
-   degg factor   & numeric &coefficient scaled and combined degradation factor \\ 
-   comp\_count\_list   & list & list of competing expression for each iteration \\ 
-   comp\_count   & numeric & pre: competing expression at previous iteration; current: competing expression at present iteration \\
-   mirna\_count\_list   & list & list of miRNA expression for each iteration \\
-   mirna\_count   & numeric & pre: miRNA expression at previous iteration; current: miRNA expression at present iteration \\
-   effect   & numeric & pre: total miRNA reppressive effect on individual target at previous iteration ; current: miRNA reppressive effect on individual   target at present iteration \\
-   effect\_list   & list & list of miRNA reppressive effect on individual target for each iteration \\
+   \emph{Edge Variables} & \\
+   Competing name   &  name of genes\\
+   miRNA name   &  name of miRNAs \\
+   Competing expression   &  Expression values of competing elements at steady-state \\
+   miRNA expression   &  Expression values of miRNA elements at steady-state \\
+   energy   & coefficient of miRNA:target interactions (binding affinity) \\
+   seed type   &  coefficient of miRNA:target interactions (binding affinity) \\
+   region   &  coefficient of miRNA:target interactions (degradation efficiency) \\
+   afff factor   &  coefficient scaled and combined affinity factor \\
+   degg factor   & coefficient scaled and combined degradation factor \\ 
+   comp\_count\_list   & list of competing expression for each iteration \\ 
+   comp\_count   & pre: competing expression at previous iteration; current: competing expression at present iteration \\
+   mirna\_count\_list   & list of miRNA expression for each iteration \\
+   mirna\_count   &  pre: miRNA expression at previous iteration; current: miRNA expression at present iteration \\
+   effect   & pre: total miRNA reppressive effect on individual target at previous iteration ; current: miRNA reppressive effect on individual   target at present iteration \\
+   effect\_list   & list of miRNA reppressive effect on individual target for each iteration \\
   \hline
@@
 
-# 4. Notes
-
-To evaluate significant nodes through parallel processing in breast cancer patient network, perturbation on all nodes were triggered by 3-fold up-regulation with 10 iteration (cycle) due to small diameter of Real and Real+ networks.
-We analyzed perturbation efficiencies of all nodes in situation that accepts nodes with more than one percent change in expression as re-regulated (i.e, limit=1 argument in *find_node_perturbation()* function).
+# 4. Access to code
 
@@
 
+\newpage
+
 # REFERENCES

### 2020-11-02 Supplementary_mat_met.Rmd
updated supplements

@@
 
-# 1. Determination of perturbation efficiencies of nodes in network.
 ```{r, include=FALSE, message=FALSE, warning=FALSE}
@@
 
+# 1. Determination of perturbation efficiencies of nodes in network.
+
+`find_node_perturbation` function runs `calc_perturbation` function for each node with given parameters.
+In this particular example upregulation parameter is two, iteration parameter is three.
+By using formulas described in manuscript perturbation efficiency and number of affected nodes (perturbed count) are calculated.
+Results for Sample+ network are shown below.
+
 ```{r, message=FALSE, warning=FALSE}
 library(ceRNAnetsim)
-```
-
-```{r, include=FALSE}
 data("minsamp") # refers Sample and Sample+ data
-```
-
-## Node efficiencies on Sample+ network
-
-```{r, message=FALSE, warning=FALSE, echo=FALSE}
 
@@
 
-CLASH dataset was retrieved from [PubMed](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3650559/) [@helwak_mapping_2013].
+CLASH dataset was retrieved from [Helwak et. al](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3650559/) [@helwak_mapping_2013].
 
@@
 
-clashelwak <- read.table("mmc1.txt", comment.char = "#", header = TRUE, skip = 1, stringsAsFactors = FALSE)
+clash_url <- "https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3650559/bin/mmc1.txt"
+
+clashelwak <- read.table(clash_url, comment.char = "#", header = TRUE, skip = 1, stringsAsFactors = FALSE)
 
@@
 ```{r, eval=FALSE}
-read.table("mirbasehg38.txt", comment.char = "#")%>%
+mirbase_url <- "ftp://mirbase.org/pub/mirbase/21/genomes/hsa.gff3"
+
+read.table(mirbase_url, comment.char = "#")%>%
   filter(V3 != "miRNA_primary_transcript")%>%
@@
 
-CLASH dataset is published in miRBase release 15 and Human Genome 19 version.
+Since CLASH dataset is published according to Human Genome 19 version, we used miRBase release 15.
 
 ```{r, eval=FALSE}
-read_tsv("mirna_mature.txt", col_names = FALSE)%>%
+mirbase15_url <- "ftp://mirbase.org/pub/mirbase/15/database_files/mirna_mature.txt.gz"
+
+read_tsv(mirbase15_url, col_names = FALSE)%>%
   filter(startsWith(X2, "hsa"))%>%
@@
 
-CLASH dataset was retrieved from [Nature article](nature.com/articles/ncomms9864)
+CLEAR-CLiP dataset was retrieved from [Moore et.al](https://www.nature.com/articles/ncomms9864).
+Please note that we need Supplementary Data 4, however there is error in website, you need to click on Supplementary Data 3 to download the data.
 
 ```{r,  eval=FALSE}
-clearclip <- read_xlsx("CLEAR-CLIP.xlsx")
+clear_clip_url <- "https://static-content.springer.com/esm/art%3A10.1038%2Fncomms9864/MediaObjects/41467_2015_BFncomms9864_MOESM1022_ESM.xlsx"
+
+download.file(url=clear_clip_url, destfile="CLEAR-CLIP.xlsx", mode = "wb")
+
+clearclip <- readxl::read_excel("CLEAR-CLIP.xlsx",1, skip=23)
 
@@
 
+CLASH and CLEAR-CLiP data were processed heavily and resulted in similar format after which they were concatenated.
+
 ```{r,  eval=FALSE}
@@
 
-### Adding Coefficients of Interaction factors
+## 2.4 Adding Coefficients of Interaction factors
 
@@
 
-In this step we added numeric intraction values at followings
+In this step we added numeric interaction values as following
 
-Fistly, we organized these values due to the fact that the regions were defined differently in two datasets. After that, region effect was added as numeric values (shown in Table S3).
+Firstly, we organized these values due to the fact that the regions were defined differently in two dataset. After that, region effect was added as numeric values (shown in Table S3).
 
@@
 
+# 3. Functions defined for ceRNA models and workflow of method
+
+We defined the functions that can be used with R programming.
+Briefly, these functions process a given miRNA:gene dataset and convert to graph object.
+All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
+The functions and steps of approach are explained as following (Figure S1) :
+
+**Conversion of dataset**: `priming_graph()` function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
+If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
+The formulations that are given in equations (1-4) are performed in this function.
+This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
+It is assumed that the initial target amounts in the dataset is observed after the repressive activity of miRNAs in steady-state.
+
+**Transition of variables in graph**: In the previous step, the calculations are performed in the edge variables of the graph object.
+However, the graph object allows to use node variables, while the node features are handled to the graph.
+In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.
+
+**Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficiency coefficients are calculated according to this acceptance.
+In the network that is found in steady-state conditions, the change is applied to the graph object for perturbation of steady-state.
+To provide the perturbation in the network the work-flow offer two methods: *update_variables* and *update_how*.
+The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
+The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
+
+**Simulation of competing behavior of targets**: After the change in the steady-state conditions, the network elements try to gain steady-state again.
+This process progresses as repeating of regulations after the spreading the changes in the network.
+In this step, simulation of regulations according to given cycle count in simulate function is applied.
+After each simulation cycle, the miRNA repression values are re-calculated and the current values of competing elements are found and saved.
+The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node data.
+
+The node elements in the dataset are handled as two type; repressive (miRNAs) and competing (targets).
+It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used.
+If the trigger of the network is a miRNA, it maintains the current value of amount that provides by user.
+On the contrary, it tries to help this process to provide steady-state through the regulations on its amount, if a competing element is used as a trigger.
+The functions that are used in the this study are developed with R and are available in [Bioconductor](https://www.bioconductor.org/packages/release/bioc/html/ceRNAnetsim.html).
+
+# 4. Perturbation analysis of Real/Real+ networks
+
+To evaluate significant nodes through parallel processing in breast cancer patient network, perturbation on all nodes were triggered by 3-fold up-regulation with 10 iteration (cycle) due to small diameter of Real and Real+ networks.
+We analyzed perturbation efficiencies of all nodes in situation that accepts nodes with more than one percent change in expression as re-regulated (i.e, limit=1 argument in *find_node_perturbation()* function).
+
 # REFERENCES

### 2020-11-05 Reproducible_codes_Supp_figures.Rmd
supplement_fig_codes
tidyverse styling and validating code

--- 2020-10-20 Reproducible_codes.Rmd
@@
 ```{r, include=FALSE}
-data("minsamp") # refers Sample and Sample+ data
+data("minsamp") # refers to Sample and Sample+ data
 ```
 
-See Table S1 in Supplementary Figures and Tables file.
+See Table S1 in Supplementary Figure and Table file.
 
@@
 ```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression) %>%
   update_how("Gene2", 2)
@@
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)%>%
-  simulate(cycle= 2)
-
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression) %>%
+  update_how("Gene2", 2) %>%
+  simulate(cycle = 2)
 ```
@@
 ```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)%>%
-  simulate(2)%>%
-  activate(edges)%>%
-  as_tibble()%>%
-  select(Competing_name,comp_count_list,effect_list)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression) %>%
+  update_how("Gene2", 2) %>%
+  simulate(2) %>%
+  activate(edges) %>%
+  as_tibble() %>%
+  select(Competing_name, comp_count_list, effect_list) %>%
   unnest()
@@
 ```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)
-
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region)
 ```
@@
 ```{r, message=FALSE, warning=FALSE}
-
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
   update_how("Gene2", 2)
@@
 ```{r, message=FALSE, warning=FALSE}
-
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene2", 2)%>%
-  simulate(cycle=2)
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
+  update_how("Gene2", 2) %>%
+  simulate(cycle = 2)
 ```
@@
 ```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene2", 2)%>%
-  simulate(3)%>%
-  activate(edges)%>%
-  as_tibble()%>%
-  select(Competing_name,comp_count_list,effect_list)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
+  update_how("Gene2", 2) %>%
+  simulate(3) %>%
+  activate(edges) %>%
+  as_tibble() %>%
+  select(Competing_name, comp_count_list, effect_list) %>%
   unnest()
@@
 ```{r, message=FALSE, warning=FALSE}
-
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene4", 2)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
+  update_how("Gene4", 2) %>%
   simulate(cycle = 2)
@@
 ```{r, message=FALSE, warning=FALSE}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene4", 2)%>%
-  simulate(3)%>%
-  activate(edges)%>%
-  as_tibble()%>%
-  select(Competing_name,comp_count_list,effect_list)%>%
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
+  update_how("Gene4", 2) %>%
+  simulate(3) %>%
+  activate(edges) %>%
+  as_tibble() %>%
+  select(Competing_name, comp_count_list, effect_list) %>%
   unnest()

### 2020-11-05 Supplementary_fig_tab.Rmd
tidyverse styling and validating code

@@
 Refer to Table S1 for exact expression values", fig.width=5, fig.height=4}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4",
-            title = " ")+
-  geom_node_text(aes(label = name), repel = TRUE, size =3)
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression) %>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4",
+            title = " ") +
+  geom_node_text(aes(label = name), repel = TRUE, size = 3)
 ```
@@
 Expression level of Gene2 is increased from 10,000 to 20,000 in order to demonstrate the calculation steps.", fig.width=5, fig.height=4}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", 
-            title = " ")+
-  geom_node_text(aes(label = name), repel = TRUE, size =3)
+priming_graph(minsamp,
+              competing_count = Competing_expression,
+              miRNA_count = miRNA_expression) %>%
+  update_how("Gene2", 2) %>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red",
+            title = " ") +
+  geom_node_text(aes(label = name), repel = TRUE, size = 3)
 ```
@@
               competing_count = Competing_expression, 
-              miRNA_count = miRNA_expression)%>%
-  update_how("Gene2", 2)%>%
-  simulate(2)%>%
+              miRNA_count = miRNA_expression) %>%
+  update_how("Gene2", 2) %>%
+  simulate(2) %>%
   vis_graph(Competing_color = "navajowhite3", 
@@
 ```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.cap= "Sample+ in Steady-state. Interaction factors of Sample+ network are available in Table S1.", fig.width=5, fig.height=4}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", 
-            title = " ")+
-  geom_node_text(aes(label = name), repel = TRUE, size =3)
-
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4",
+            title = " ") +
+  geom_node_text(aes(label = name), repel = TRUE, size = 3)
 ```
@@
 ```{r, message=FALSE, warning=FALSE, echo=FALSE,  fig.cap= "Perturbation in Sample+ network by two-fold increase in Gene2 expression level.", fig.width=5, fig.height=4}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene2", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", 
-            title = " ")+
-  geom_node_text(aes(label = name), repel = TRUE, size =3)
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
+  update_how("Gene2", 2) %>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red",
+            title = " ") +
+  geom_node_text(aes(label = name), repel = TRUE, size = 3)
 ```
@@
 Although visualisation looks similar to Figure S4B, current counts of genes are drastically different."}
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene2", 2)%>%
-  simulate(cycle = 1)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "A")+
-  geom_node_text(aes(label = name), repel = TRUE, size =3)
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
+  update_how("Gene2", 2) %>%
+  simulate(cycle = 1) %>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            Downregulation = "blue", 
+            title = "A") +
+  geom_node_text(aes(label = name), repel = TRUE, size = 3)
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene2", 2)%>%
-  simulate(cycle = 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", Downregulation = "blue", title = "B")+
-  geom_node_text(aes(label = name), repel = TRUE, size =3)
-
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
+  update_how("Gene2", 2) %>%
+  simulate(cycle = 2) %>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            Downregulation = "blue", 
+            title = "B") +
+  geom_node_text(aes(label = name), repel = TRUE, size = 3)
 ```
@@
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "A")+
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            title = "A") +
   geom_node_text(aes(label = name), repel = TRUE, size =3)
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene4", 2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "B")+
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
+  update_how("Gene4", 2) %>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            title = "B") +
   geom_node_text(aes(label = name), repel = TRUE, size =3)
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene4", 2)%>%
-  simulate(cycle=1)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "C")+
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
+  update_how("Gene4", 2) %>%
+  simulate(cycle=1) %>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            title = "C") +
   geom_node_text(aes(label = name), repel = TRUE, size =3)
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)%>%
-  update_how("Gene4", 2)%>%
-  simulate(cycle=2)%>%
-  vis_graph(Competing_color = "navajowhite3", mirna_color = "ivory4", Upregulation = "red", title = "D")+
+priming_graph(minsamp, 
+              competing_count = Competing_expression, 
+              miRNA_count = miRNA_expression, 
+              aff_factor = c(energy, seed_type), 
+              deg_factor = region) %>%
+  update_how("Gene4", 2) %>%
+  simulate(cycle=2) %>%
+  vis_graph(Competing_color = "navajowhite3", 
+            mirna_color = "ivory4", 
+            Upregulation = "red", 
+            title = "D") +
   geom_node_text(aes(label = name), repel = TRUE, size =3)
@@
 
-matched_patient_exp_dataset <- readRDS("figures_manuscript/matched_patient_exp_dataset.RDS")
-mirtarbase_gene_res <- readRDS("figures_manuscript/mirtarbase_gene_res.RDS")
+matched_patient_exp_dataset <- readRDS("figure_manuscript/matched_patient_exp_dataset.RDS")
+mirtarbase_gene_res <- readRDS("figure_manuscript/mirtarbase_gene_res.RDS")
 
-matched_patient_exp_dataset%>%
-  filter(Ensembl_Gene_Id %in% pull(filter(mirtarbase_gene_res, situ == "tumor_specific"), Ensembl_Gene_Id))%>%
-  pivot_longer(cols = contains("TCGA"), names_to="patient_id", values_to = "expressions")%>%
-  mutate(patient = ifelse(endsWith(patient_id, "01A"), "tumor", "normal"))%>%
-  left_join(mirtarbase_gene_res, by ="Ensembl_Gene_Id")%>%
-  group_by(patient, Ensembl_Gene_Id)%>%
-  ggplot(aes(x= hgnc_symbol, y=log2(expressions), fill = patient))+
-  #scale_y_log10()+
-  geom_boxplot(position = "dodge")+
-  theme_bw()+
-  theme(axis.text.x = element_text(angle = 90), panel.background = element_rect())+
+matched_patient_exp_dataset %>%
+  filter(Ensembl_Gene_Id %in% pull(filter(mirtarbase_gene_res, situ == "tumor_specific"), Ensembl_Gene_Id)) %>%
+  pivot_longer(cols = contains("TCGA"), 
+               names_to = "patient_id", 
+               values_to = "expressions") %>%
+  mutate(patient = ifelse(endsWith(patient_id, "01A"), 
+                          "tumor", "normal")) %>%
+  left_join(mirtarbase_gene_res, by = "Ensembl_Gene_Id") %>%
+  group_by(patient, Ensembl_Gene_Id) %>%
+  ggplot(aes(x = hgnc_symbol, y = log2(expressions), fill = patient)) +
+  # scale_y_log10()+
+  geom_boxplot(position = "dodge") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90), 
+        panel.background = element_rect()) +
   xlab("")
@@
 # 3 Supplementary Tables
-
-```{r, eval=FALSE, include=FALSE}
-library(xtable)
-knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
-library(gt)
-knitr::opts_chunk$set(cache=FALSE)
-
-```
 

### 2020-11-05 Supplementary_mat_met.Rmd
tidyverse styling and validating code

@@
 
-priming_graph(minsamp, competing_count = Competing_expression, miRNA_count = miRNA_expression, aff_factor = c(energy, seed_type), deg_factor = region)-> sample_graph
+sample_graph <- priming_graph(minsamp, 
+                              competing_count = Competing_expression, 
+                              miRNA_count = miRNA_expression, 
+                              aff_factor = c(energy, seed_type), 
+                              deg_factor = region)
 
-find_node_perturbation(sample_graph, how = 2, cycle = 3, limit = 0.1)%>%
-  as_tibble()%>%
+find_node_perturbation(sample_graph, 
+                       how = 2, 
+                       cycle = 3, 
+                       limit = 0.1) %>%
+  as_tibble() %>%
   select(name, perturbation_efficiency, perturbed_count)
@@
 
-```{r, eval=FALSE}
+```{r, eval=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-clashelwak <- read.table(clash_url, comment.char = "#", header = TRUE, skip = 1, stringsAsFactors = FALSE)
+clashelwak <- read_tsv(clash_url, comment = "#", skip = 1)
 
@@
 ```{r, eval=FALSE}
-#HG19
-listEnsemblArchives()
-listMarts(host = 'http://grch37.ensembl.org')
-ensemblgrch37 = useMart(host='http://grch37.ensembl.org', 
+library(biomaRt)  #can be installed via BiocManager
+
+#listMarts(host = 'http://grch37.ensembl.org') #HG19
+ensemblgrch37 <- useMart(host='http://grch37.ensembl.org',
                          biomart='ENSEMBL_MART_ENSEMBL',
                          dataset='hsapiens_gene_ensembl')
-hg19 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene_id", "strand" ), mart = ensemblgrch37)
 
+hg19 <- getBM(attributes = c("ensembl_transcript_id", 
+                             "ensembl_gene_id",
+                             "chromosome_name", 
+                             "start_position",
+                             "end_position", 
+                             "hgnc_symbol", 
+                             "entrezgene_id", 
+                             "strand" ), 
+              mart = ensemblgrch37)
 ```
@@
 ```{r,  eval=FALSE}
-
-clashelwak <- clashelwak%>%
- separate(microRNA_name, c("Barcode", "Database", "mirna_name", "type"), sep = "_")%>%
-  separate(mRNA_name, c("Ensembl_Gene_Id", "Ensembl_Transcript_Id", "Hugo_Symbol", "mRNA_Type"), sep = "_")
-
+clashelwak <- clashelwak %>%
+  separate(microRNA_name, 
+           c("Barcode", "Database", "mirna_name", "type"), 
+           sep = "_") %>%
+  separate(mRNA_name, 
+           c("Ensembl_Gene_Id", "Ensembl_Transcript_Id", 
+             "Hugo_Symbol", "mRNA_Type"), 
+           sep = "_") %>%
+  rename_with(~ str_replace(.,"\'","_"), contains("UTR"))
 ```
 
-MiRNA releases are obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) was downloaded.
+miRNA releases are obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) was downloaded.
 
-```{r, eval=FALSE}
+```{r, eval=FALSE, warning=FALSE, message=FALSE}
 mirbase_url <- "ftp://mirbase.org/pub/mirbase/21/genomes/hsa.gff3"
 
-read.table(mirbase_url, comment.char = "#")%>%
-  filter(V3 != "miRNA_primary_transcript")%>%
-  separate(V9, c("ID", "Alias", "Name", "Precusor"), sep = ";")%>%
-  mutate(ID = substr(ID, 4, length(ID)), Alias = substr(Alias, 7, length(Alias)), Name = substr(Name, 6, length(Name)), Precusor = substr(Precusor, 14, length(Precusor)))%>%
-  dplyr::select(chr= V1, start = V4, end = V5, strand = V7, ID, Alias, Name, Precusor)->mirbasehg38
-
+mirbasehg38 <- read_tsv(mirbase_url, comment = "#", col_names = FALSE) %>%
+  filter(X3 != "miRNA_primary_transcript") %>%
+  separate(X9, c("ID", "Alias", "Name", "Precusor"), sep = ";") %>%
+  mutate(ID = substr(ID, 4, length(ID)), 
+         Alias = substr(Alias, 7, length(Alias)), 
+         Name = substr(Name, 6, length(Name)), 
+         Precusor = substr(Precusor, 14, length(Precusor))) %>%
+  dplyr::select(chr = X1, start = X4, end = X5, strand = X7, ID, Alias, Name, Precusor)
 ```
@@
 
-read_tsv(mirbase15_url, col_names = FALSE)%>%
-  filter(startsWith(X2, "hsa"))%>%
-  dplyr::select(mirna_ID = X2, mirbase_ID = X3)%>%
-  inner_join(mirbasehg38%>% dplyr::select(ID, Name), by= c("mirbase_ID"= "ID"))%>%
-  dplyr::select(mirbase_ID, Name)%>%
-  distinct()%>%
-  inner_join(clashelwak, by = c("mirbase_ID"= "Barcode"))%>%
-  dplyr::select(Name, miRNA_seq, Ensembl_Gene_Id, Ensembl_Transcript_Id, Hugo_Symbol, mRNA_seq_extended, chimeras_decompressed, seed_type, seed_basepairs,folding_class, seq_ID, folding_energy, X5.UTR, CDS, X3.UTR)%>%
-  inner_join(hg19, by = c("Ensembl_Gene_Id"= "ensembl_gene_id", "Ensembl_Transcript_Id"="ensembl_transcript_id", "Hugo_Symbol"="hgnc_symbol"))%>%
-  mutate(region1 = ifelse(X5.UTR == "1", "5UTR", " "), region2= ifelse(X3.UTR == "1", "3UTR", " "),region3= ifelse(CDS =="1", "CDS", " "))%>%
-  unite(region, c(region1, region2, region3), sep = "||")%>%
-  dplyr::select(chromosome_name, start_position, end_position,strand, Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, mRNA_seq_extended, Name, miRNA_seq, seq_ID, seed_type,seed_basepairs, folding_class, folding_energy, region)%>%
-  as_tibble() -> clashelwakfinal
-
+clashelwakfinal <- read_tsv(mirbase15_url, col_names = FALSE) %>%
+  filter(str_starts(X2, "hsa")) %>%
+  dplyr::select(mirna_ID = X2, mirbase_ID = X3) %>%
+  inner_join(mirbasehg38 %>% dplyr::select(ID, Name), 
+             by = c("mirbase_ID" = "ID")) %>%
+  dplyr::select(mirbase_ID, Name) %>%
+  distinct() %>%
+  inner_join(clashelwak, by = c("mirbase_ID" = "Barcode")) %>%
+  dplyr::select(Name, miRNA_seq, Ensembl_Gene_Id, Ensembl_Transcript_Id, 
+                Hugo_Symbol, mRNA_seq_extended, chimeras_decompressed, 
+                seed_type, seed_basepairs, folding_class, seq_ID, 
+                folding_energy, `5_UTR`, CDS, `3_UTR`) %>%
+  inner_join(hg19, by = c("Ensembl_Gene_Id" = "ensembl_gene_id", 
+                          "Ensembl_Transcript_Id" = "ensembl_transcript_id", 
+                          "Hugo_Symbol" = "hgnc_symbol")) %>% 
+  mutate(region1 = ifelse(`5_UTR` == "1", "5UTR", " "), 
+         region2 = ifelse(`3_UTR` == "1", "3UTR", " "), 
+         region3 = ifelse(CDS == "1", "CDS", " ")) %>%
+  unite(region, c(region1, region2, region3), sep = "||") %>% 
+  dplyr::select(chromosome_name, start_position, end_position, strand, 
+                Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, 
+                mRNA_seq_extended, Name, miRNA_seq, seq_ID, seed_type, 
+                seed_basepairs, folding_class, folding_energy, region) 
 ```
@@
 
-lift19 <- clashelwakfinal%>%
-  dplyr::select(1,2,3)%>%
-  unite(start_end, c("start_position", "end_position"), sep = "-")%>%
-  mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
-  unite(chromosome_name, c("Chromosome","start_end"), sep = ":")
+clashelwakfinal %>%
+  dplyr::select(1,2,3) %>%
+  unite(start_end, c("start_position", "end_position"), sep = "-") %>%
+  mutate(Chromosome = paste0("chr",chromosome_name, "")) %>%
+  unite(chromosome_name, c("Chromosome","start_end"), sep = ":") %>% 
+  write_tsv("lift19.txt",col_names = FALSE)
 
-write_tsv(lift19, "lift19.txt")
-
-#After we searched this file in UCSC browser, the output loaded (lift19_del deleted regions on the HG38 genome build; hg38clashcomp new locations of the genes on HG38)
+#After we uploaded this file to UCSC Liftover site, the output files  lift19_del (deleted regions on the HG38 genome build) and hg38clashcomp (new locations of the genes on HG38) were downloaded
 
@@
 
-lift19_del <- lift19_del%>%
-  dplyr::filter(startsWith(chromosome_loc, "chr"))%>%
-  separate(chromosome_loc, c("Chr", "End"), "-", remove = TRUE)%>%
+lift19_del <- lift19_del %>%
+  dplyr::filter(startsWith(chromosome_loc, "chr")) %>%
+  separate(chromosome_loc, c("Chr", "End"), "-", remove = TRUE) %>%
   separate(Chr, c("Chr", "Start"), ":", remove = TRUE)
@@
 clashelwakfinal <-clashelwakfinal%>%
-  mutate(Chromosome = paste0("chr",chromosome_name, ""))%>%
-  dplyr::anti_join(lift19_del, by = c("Chromosome"="Chr", "start_position"="Start", "end_position"="End"))
+  mutate(Chromosome = paste0("chr",chromosome_name, "")) %>%
+  dplyr::anti_join(lift19_del, by = c("Chromosome"="Chr",
+                                      "start_position"="Start",
+                                      "end_position"="End"))
 
@@
 
-clashelwakfinal <- clashelwakfinal%>% 
+clashelwakfinal <- clashelwakfinal %>% 
   bind_cols(hg38clash)
@@
 
-clashelwakfinal <- clashelwakfinal%>%
+clashelwakfinal <- clashelwakfinal %>%
   dplyr::mutate(Genom_build= rep("hg19"))
@@
 
-clashelwakfinal%>% dplyr::select(cluster=seq_ID, chromosome = Chromosome, start_position, end_position, strand, hgnc_symbol = Hugo_Symbol, Ensembl_Gene_Id, Ensembl_Transcript_Id, target_seq = mRNA_seq_extended, miRNA= Name, miR_seq =miRNA_seq, seed_type, seed_type2= seed_basepairs, seed_type3= folding_class, Energy = folding_energy, HG38build_loc, Genom_build, region) -> clashelwakfinal
+clashelwakfinal <- clashelwakfinal %>% 
+  dplyr::select(cluster=seq_ID, 
+                chromosome = Chromosome,
+                start_position, end_position, strand,
+                hgnc_symbol = Hugo_Symbol,
+                Ensembl_Gene_Id, Ensembl_Transcript_Id,
+                target_seq = mRNA_seq_extended,
+                miRNA = Name,
+                miR_seq = miRNA_seq,
+                seed_type,
+                seed_type2 = seed_basepairs,
+                seed_type3 = folding_class,
+                Energy = folding_energy,
+                HG38build_loc, Genom_build, region)
 
@@
 ```{r,  eval=FALSE}
-clashelwakfinal%>%
-  mutate(seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="I", paste0(seed_type2, "-mer"), seed_type),
-         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="II", paste0(seed_type2, "-mer_noncanonical"), seed_type), 
-         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="III", paste0(seed_type2, "-mer_noncanonical"), seed_type),
-         seed_type= ifelse (seed_type == "noncanonical_seed"& seed_type2 >4 &seed_type3=="IV", paste0(seed_type2, "-mer_noncanonical"), seed_type),
-         seed_type= ifelse (startsWith(seed_type, "no") ,"none", seed_type))%>%
-  dplyr::select(-seed_type2, -seed_type3)->clashelwakfinal
+clashelwakfinal <- clashelwakfinal %>%
+  mutate(seed_type= ifelse (seed_type == "noncanonical_seed" & seed_type2 >4 & seed_type3=="I", 
+                            paste0(seed_type2, "-mer"), 
+                            seed_type),
+         seed_type= ifelse (seed_type == "noncanonical_seed" & seed_type2 >4 & seed_type3=="II", 
+                            paste0(seed_type2, "-mer_noncanonical"), 
+                            seed_type), 
+         seed_type= ifelse (seed_type == "noncanonical_seed" & seed_type2 >4 & seed_type3=="III", 
+                            paste0(seed_type2, "-mer_noncanonical"), 
+                            seed_type),
+         seed_type= ifelse (seed_type == "noncanonical_seed" & seed_type2 >4 & seed_type3=="IV", 
+                            paste0(seed_type2, "-mer_noncanonical"), 
+                            seed_type),
+         seed_type= ifelse (startsWith(seed_type, "no") ,
+                            "none", 
+                            seed_type))%>%
+  dplyr::select(-seed_type2, -seed_type3)
 
@@
 #HG18
-listEnsemblArchives()
-listMarts(host = 'may2009.archive.ensembl.org')
-ensembl54 = useMart(host='may2009.archive.ensembl.org', 
-                     biomart='ENSEMBL_MART_ENSEMBL',
-                    dataset='hsapiens_gene_ensembl')
 
-hg18 <- getBM(attributes = c("ensembl_transcript_id","ensembl_gene_id","chromosome_name", "start_position","end_position", "hgnc_symbol", "entrezgene", "strand" ), mart = ensembl54)
+#listMarts(host = 'http://may2009.archive.ensembl.org')
+
+ensembl54 <- useMart(biomart='ENSEMBL_MART_ENSEMBL',
+                    dataset='hsapiens_gene_ensembl',
+                    host='may2009.archive.ensembl.org')
+ensembl54 <- useEnsembl(biomart = 'genes', 
+                       dataset = 'hsapiens_gene_ensembl',
+                       version = 54)
+listMarts(host="may2009.archive.ensembl.org")
+useMart("ensembl",host="may2009.archive.ensembl.org", dataset = "hsapiens_gene_ensembl")
+httr_config <- httr::config(ssl_cipher_list = "DEFAULT@SECLEVEL=1")
+
+mart <- useEnsembl(biomart = "ensembl", 
+                   dataset = "hsapiens_gene_ensembl", 
+                   version = "54")
+
+hg18 <- getBM(attributes = c("ensembl_transcript_id",
+                             "ensembl_gene_id",
+                             "chromosome_name", 
+                             "start_position",
+                             "end_position", 
+                             "hgnc_symbol", 
+                             "entrezgene", 
+                             "strand" ), 
+              mart = ensembl54)
 ```
@@
 ```{r,  eval=FALSE}
-clearclipfinal <- hg18%>%
-  inner_join(clearclip, by= c("entrezgene"= "gene.id", "hgnc_symbol"= "gene.symbol"))%>%
+clearclipfinal <- hg18 %>%
+  inner_join(clearclip, by= c("entrezgene"= "gene.id", 
+                              "hgnc_symbol"= "gene.symbol")) %>%
   distinct()
@@
 
-lift18 <- clearclipfinal%>%
-  unite(start_end, c("start_position", "end_position"), sep = "-")%>%
-  unite(location, c("chr","start_end"), sep = ":")%>%
+lift18 <- clearclipfinal %>%
+  unite(start_end, c("start_position", "end_position"), sep = "-") %>%
+  unite(location, c("chr","start_end"), sep = ":") %>%
   dplyr::select(location)
@@
 
-deleted_lift18 <- deleted_lift18%>%
-  dplyr::filter(startsWith(Chromosome_loc, "chr"))%>%
-  separate(Chromosome_loc, c("Chr", "End"), "-", remove = TRUE)%>%
+deleted_lift18 <- deleted_lift18 %>%
+  dplyr::filter(startsWith(Chromosome_loc, "chr")) %>%
+  separate(Chromosome_loc, c("Chr", "End"), "-", remove = TRUE) %>%
   separate(Chr, c("Chr", "Start"), ":", remove = TRUE)
@@
 
-clearclipfinal <- clearclipfinal%>%
-  dplyr::anti_join(deleted_lift18, by = c("chr"="Chr", "start_position"="Start", "end_position"="End"))
+clearclipfinal <- clearclipfinal %>%
+  dplyr::anti_join(deleted_lift18, by = c("chr"="Chr", 
+                                          "start_position"="Start", 
+                                          "end_position"="End"))
 
@@
 
-clearclipfinal <- clearclipfinal%>% 
+clearclipfinal <- clearclipfinal %>% 
   bind_cols(hg38clearclip)
@@
 
-clearclipfinal <- clearclipfinal%>%
+clearclipfinal <- clearclipfinal %>%
   dplyr::mutate(Genom_build= rep("hg18"))
@@
 
-clearclipfinal%>%
-  dplyr::select(cluster=cluster.ID, chromosome = chr, start_position, end_position, strand = strand.y, hgnc_symbol, Ensembl_Gene_Id =ensembl_gene_id, Ensembl_Transcript_Id =ensembl_transcript_id, target_seq = target.map, miRNA, miR_seq = miR.map, seed_type = 'seed match', Energy = MFE, HG38build_loc, Genom_build, region) -> clearclipfinal
+clearclipfinal %>%
+  dplyr::select(cluster=cluster.ID, chromosome = chr, start_position, 
+                end_position, strand = strand.y, hgnc_symbol, 
+                Ensembl_Gene_Id =ensembl_gene_id, 
+                Ensembl_Transcript_Id =ensembl_transcript_id, 
+                target_seq = target.map, miRNA, 
+                miR_seq = miR.map, 
+                seed_type = 'seed match', Energy = MFE, 
+                HG38build_loc, Genom_build, region) -> clearclipfinal
 
@@
 ```{r,  eval=FALSE}
-data_frame(seed_type= c("5mer_1","5mer_2", "5mer_3", "6mer", "6mer.indel", "6mer.mm", "6mer_off.mm", "6merA1", "6merA1.indel", "6merA1.mm", "7merA1", "7merA1.indel", "7merA1.mm", "7merm8", "7merm8.indel", "7merm8,mm", "8mer", "8mer.indel", "8mer.mm", "NA" ),
-           seed_type_com= c("5-mer", "5-mer_noncanonical", "5-mer_noncanonical", "6-mer", "6-mer_noncanonical", "6-mer_noncanonical", "6-mer_noncanonical", "6-merA1", "6-merA1_noncanonical","6-merA1_noncanonical", "7-merA1", "7-merA1_noncanonical","7-merA1_noncanonical", "7-mer-8m", "7-mer-8m_noncanonical", "7-mer-8m_noncanonical", "8-mer", "8-mer_noncanonical", "8-mer_noncanonical", "none" )) ->clipdata_seed
+clipdata_seed <- data_frame(seed_type= c("5mer_1","5mer_2", "5mer_3", "6mer", "6mer.indel", "6mer.mm", "6mer_off.mm", "6merA1", "6merA1.indel", "6merA1.mm", "7merA1", "7merA1.indel", "7merA1.mm", "7merm8", "7merm8.indel", "7merm8,mm", "8mer", "8mer.indel", "8mer.mm", "NA" ),
+           seed_type_com= c("5-mer", "5-mer_noncanonical", "5-mer_noncanonical", "6-mer", "6-mer_noncanonical", "6-mer_noncanonical", "6-mer_noncanonical", "6-merA1", "6-merA1_noncanonical","6-merA1_noncanonical", "7-merA1", "7-merA1_noncanonical","7-merA1_noncanonical", "7-mer-8m", "7-mer-8m_noncanonical", "7-mer-8m_noncanonical", "8-mer", "8-mer_noncanonical", "8-mer_noncanonical", "none" ))
 
-clearclipfinal%>%
-  inner_join(clipdata_seed, by = "seed_type")%>%
-  dplyr::select(1:11, seed_type = seed_type_com, Energy, HG38build_loc, Genom_build, region)-> clearclipfinal
+clearclipfinal <- clearclipfinal %>%
+  inner_join(clipdata_seed, by = "seed_type") %>%
+  dplyr::select(1:11, seed_type = seed_type_com, Energy, HG38build_loc, Genom_build, region)
 
@@
 
-bind_rows(clashelwakfinal, clearclipfinal)%>%distinct() -> experimentalmirnagene
-
+experimentalmirnagene <- bind_rows(clashelwakfinal, clearclipfinal) %>%
+  distinct()  
 ```
@@
 ```{r,  eval=FALSE}
-experimentalmirnagene <- experimentalmirnagene%>%
-  mutate(region2  = str_replace_all(region, "NA", ""), region3 = str_replace_all(region2, "\\|", ""), region = str_replace_all(region3, c("3'UTR"="3UTR", "5'UTR"="5UTR")))%>%
-  dplyr::select(-region2, -region3)%>%
-  mutate(region_effect = as.double(ifelse(region %in% c("3UTRCDS", "CDS3UTR", "5UTR3UTR", "CDS5UTR3UTR", "CDS3UTRintron"), "0.93", ifelse(region %in% c("CDS", "CDSintron"), "0.42", ifelse(region %in% c("3UTR","3UTRintron"), "0.84", ifelse(region %in% c("5UTR", "5UTRintron"), "0.01", ifelse(region %in% c("5UTRCDS", "CDS5UTR"), "0.42", ifelse(region %in% c("intron",""), "0.01", ifelse(region %in% c("exon_unclassified", ""), "0.2", NA)))))))))
-
+experimentalmirnagene <- experimentalmirnagene %>%
+  mutate(region2 = str_replace_all(region, "NA", ""), 
+         region3 = str_replace_all(region2, "\\|", ""), 
+         region = str_replace_all(region3, c("3'UTR" = "3UTR", "5'UTR" = "5UTR"))) %>%
+  dplyr::select(-region2, -region3) %>%
+  mutate(region_effect = case_when(
+    region %in% c("3UTRCDS", "CDS3UTR", "5UTR3UTR", "CDS5UTR3UTR", "CDS3UTRintron") ~ "0.93",
+    region %in% c("CDS", "CDSintron") ~ "0.42",
+    region %in% c("3UTR", "3UTRintron") ~ "0.84",
+    region %in% c("5UTR", "5UTRintron") ~ "0.01",
+    region %in% c("intron", "") ~ "0.01",
+    region %in% c("exon_unclassified", "") ~ "0.2",
+    TRUE ~ NA
+  ) %>% 
+    mutate(region_effect = as.double(region_effect))
 ```
@@
                                 )
-experimentalmirnagene%>%
-  inner_join(seed_type_effect, by= "seed_type") -> experimentalmirnagene
+experimentalmirnagene <- experimentalmirnagene %>%
+  inner_join(seed_type_effect, by= "seed_type")
 ```
@@
 ```{r}
-readRDS("data/experimentalmirnagene.RDS")->experimentalmirnagene
+experimentalmirnagene <- readRDS("data/experimentalmirnagene.RDS")
 experimentalmirnagene

### 2020-11-05 Supplementary_mat_met.Rmd
liftover data added, cosmetic changes

@@
 
-#After we uploaded this file to UCSC Liftover site, the output files  lift19_del (deleted regions on the HG38 genome build) and hg38clashcomp (new locations of the genes on HG38) were downloaded
+#In liftover site, Original Assembly was chosen as hg19 and New Assebly was chosen as hg38.
+After we uploaded "lift19.txt" file to UCSC Liftover site, the output "View Conversions" was downloaded as hg38clashcomp (new locations of the genes on HG38), "Display failure file" was download as "lift19_del.txt" (deleted regions on the HG38 genome build)
 
-lift19_del <- read_tsv("deleted_lift19.txt")
-colnames(lift19_del)[1] <- "chromosome_loc"
-
-lift19_del <- lift19_del %>%
-  dplyr::filter(startsWith(chromosome_loc, "chr")) %>%
+lift19_del <- read_tsv("figure_manuscript/lift19_del.txt", 
+                       col_names = FALSE, 
+                       comment = "#") %>% 
+  rename(chromosome_loc = X1) %>% 
+  dplyr::filter(str_starts(chromosome_loc, "chr")) %>% 
   separate(chromosome_loc, c("Chr", "End"), "-", remove = TRUE) %>%
-  separate(Chr, c("Chr", "Start"), ":", remove = TRUE)
-
-lift19_del$Start <- as.numeric(lift19_del$Start)
-lift19_del$End <- as.numeric(lift19_del$End)
+  separate(Chr, c("Chr", "Start"), ":", remove = TRUE) %>% 
+  mutate(Start=as.numeric(Start),
+         End=as.numeric(End))
 
@@
 
-clashelwakfinal <-clashelwakfinal%>%
+clashelwakfinal <- clashelwakfinal %>%
   mutate(Chromosome = paste0("chr",chromosome_name, "")) %>%
@@
                                       "end_position"="End"))
-
-hg38clash <- read.delim("hg38clashcomp.txt", header = FALSE, stringsAsFactors = FALSE)
-
 # adding new location information:
@@
 clashelwakfinal <- clashelwakfinal %>% 
-  bind_cols(hg38clash)
-
-colnames(clashelwakfinal)[18] <- "HG38build_loc"
-
-clashelwakfinal <- clashelwakfinal %>%
-  dplyr::mutate(Genom_build= rep("hg19"))
+  bind_cols(read_tsv("figure_manuscript/hg38clashcomp.txt", 
+                      col_names = FALSE))
 
@@
 clashelwakfinal <- clashelwakfinal %>% 
+  rename(HG38build_loc = X1) %>% 
+  mutate(Genom_build="hg19") %>% 
   dplyr::select(cluster=seq_ID, 
@@
                 Energy = folding_energy,
-                HG38build_loc, Genom_build, region)
-
-clashelwakfinal$strand <- as.character(clashelwakfinal$strand)
-
-str(clashelwakfinal)
+                HG38build_loc, Genom_build, region) %>% 
+  mutate(strand = as.character(strand))
 ```
@@
 
-deleted_lift18 <- read_tsv("deleted_lift18.txt")
-
-colnames(deleted_lift18)[1] <- "Chromosome_loc"
-
-deleted_lift18 <- deleted_lift18 %>%
-  dplyr::filter(startsWith(Chromosome_loc, "chr")) %>%
+deleted_lift18 <- read_tsv("figure_manuscript/lift18_del.txt",
+                           comment = "#",
+                           col_names = FALSE) %>% 
+  rename(Chromosome_loc = X1) %>% 
+  dplyr::filter(str_starts(Chromosome_loc, "chr")) %>%
   separate(Chromosome_loc, c("Chr", "End"), "-", remove = TRUE) %>%
-  separate(Chr, c("Chr", "Start"), ":", remove = TRUE)
-
-deleted_lift18$Start <- as.numeric(deleted_lift18$Start)
-deleted_lift18$End <- as.numeric(deleted_lift18$End)
+  separate(Chr, c("Chr", "Start"), ":", remove = TRUE) %>% 
+  mutate(Start = as.numeric(Start),
+         End = as.numeric(End))
 
@@
 
-hg38clearclip<- read.delim("hg38clearclip.txt", header = FALSE, stringsAsFactors = FALSE)
-
-clearclipfinal <- clearclipfinal %>% 
-  bind_cols(hg38clearclip)
-
-colnames(clearclipfinal)[28] <- "HG38build_loc"
+hg38clearclip <- read_tsv("figure_manuscript/hg38clearclip.txt",
+                          col_names = FALSE)
 
 # adding new location information:
-
 clearclipfinal <- clearclipfinal %>% 
-  dplyr::mutate(Genom_build= rep("hg18"))
-
-# Arrangement in dataset
-
-clearclipfinal %>%
+  bind_cols(hg38clearclip) %>% 
+  rename(HG38build_loc = X1) %>% 
+  mutate(Genom_build="hg18") %>%
   dplyr::select(cluster=cluster.ID, chromosome = chr, start_position, 
@@
                 seed_type = 'seed match', Energy = MFE, 
-                HG38build_loc, Genom_build, region) -> clearclipfinal
+                HG38build_loc, Genom_build, region)
 

### 2020-11-09 Supplementary_mat_met.Rmd
HG18 solved

@@
 #HG18
+library(biomaRt)
 
-#listMarts(host = 'http://may2009.archive.ensembl.org')
+listEnsemblArchives()
 
-ensembl54 <- useMart(biomart='ENSEMBL_MART_ENSEMBL',
-                    dataset='hsapiens_gene_ensembl',
-                    host='may2009.archive.ensembl.org')
-ensembl54 <- useEnsembl(biomart = 'genes', 
-                       dataset = 'hsapiens_gene_ensembl',
-                       version = 54)
-listMarts(host="may2009.archive.ensembl.org")
-useMart("ensembl",host="may2009.archive.ensembl.org", dataset = "hsapiens_gene_ensembl")
-httr_config <- httr::config(ssl_cipher_list = "DEFAULT@SECLEVEL=1")
+#listMarts(host = 'may2009.archive.ensembl.org')
 
-mart <- useEnsembl(biomart = "ensembl", 
+ensembl54 = useEnsembl(host='may2009.archive.ensembl.org', 
+                        biomart = "ensembl", 
                          dataset = "hsapiens_gene_ensembl", 
-                   version = "54")
+                         mirror = "useast")
 
@@
                               "hgnc_symbol", 
-                             "entrezgene", 
+                              "entrezgene_id", 
                               "strand" ), 
                               mart = ensembl54)
+
 ```

### 2020-11-10 Supplementary_mat_met.Rmd
entrez_id check
after adding liftover data, update code

@@
 clearclipfinal <- hg18 %>%
-  inner_join(clearclip, by= c("entrezgene"= "gene.id", 
+  inner_join(clearclip, by= c("entrezgene_id"= "gene.id", 
                               "hgnc_symbol"= "gene.symbol")) %>%
@@
 write_tsv(lift18, "lift18.txt")
+
+```
+Please note that we have provided the hg18 liftover dataset under the data folder (lift18), since server error may occur.
+
+```{r,  eval=FALSE}
+
+lift18 <- read_tsv("data/lift18.txt")
 

### 2020-11-12 Supplementary_mat_met.Rmd
supp mat and methods are reproducible now.. yay!

@@
 
-#In liftover site, Original Assembly was chosen as hg19 and New Assebly was chosen as hg38.
+#In liftover site, Original Assembly was chosen as hg19 and New Assembly was chosen as hg38.
 After we uploaded "lift19.txt" file to UCSC Liftover site, the output "View Conversions" was downloaded as hg38clashcomp (new locations of the genes on HG38), "Display failure file" was download as "lift19_del.txt" (deleted regions on the HG38 genome build)
 
-lift19_del <- read_tsv("figure_manuscript/lift19_del.txt", 
+lift19_del <- read_tsv("data/lift19_del.txt", 
                        col_names = FALSE, 
@@
 clashelwakfinal <- clashelwakfinal %>% 
-  bind_cols(read_tsv("figure_manuscript/hg38clashcomp.txt", 
+  bind_cols(read_tsv("data/hg38clashcomp.txt", 
                       col_names = FALSE))
@@
 
-listEnsemblArchives()
-
 #listMarts(host = 'may2009.archive.ensembl.org')
-
 ensembl54 = useEnsembl(host='may2009.archive.ensembl.org', 
@@
                          dataset = "hsapiens_gene_ensembl", 
-                         mirror = "useast")
+                         mirror = "asia")
 
@@
                               mart = ensembl54)
+
+# should there be any connection problems, read the pre-saved version of hg18 gene information by commenting out the following line
+
+# hg18 <- read_tsv("data/hg18.txt") %>%
+#   rename_with(~ tolower(.x) %>%
+#                 str_replace_all(" ","_") %>%
+#                 str_remove("_\\(bp\\)")) %>%
+#   rename(start_position=gene_start,
+#          end_position=gene_end) %>% 
+#   filter(!is.na(entrezgene_id)) %>% 
+#   filter(!is.na(hgnc_symbol))
 
@@
 
-write_tsv(lift18, "lift18.txt")
+# you can save current state of lift18 data frame
+# write_tsv(lift18, "data/lift18.txt")
 
 ```
-Please note that we have provided the hg18 liftover dataset under the data folder (lift18), since server error may occur.
+
+Please note that we have provided the hg18 liftover dataset under the data folder (lift18), in case of server errors/changes may occur.
+If you had problems generating lift18 data frame, please run the following code.
+
+```{r, eval=FALSE}
+lift18 <- read_tsv("data/lift18.txt")
+```
 
@@
 
-lift18 <- read_tsv("data/lift18.txt")
+# In liftover site, Original Assembly was chosen as hg18 and New Assembly was chosen as hg38.
+After we uploaded "lift18.txt" file to UCSC Liftover site, the output "View Conversions" was downloaded as "hg38clearclip.txt" (new locations of the genes on HG38), "Display failure file" was download as "lift18_del.txt" (deleted regions on the HG38 genome build)
 
-#After we searched this file in UCSC browser, the output loaded (deleted_lift18 deleted regions on the HG38 genome build; hg38clearclip new locations of the genes on HG38)
-
-deleted_lift18 <- read_tsv("figure_manuscript/lift18_del.txt",
+deleted_lift18 <- read_tsv("data/lift18_del.txt",
                            comment = "#",
@@
 
-hg38clearclip <- read_tsv("figure_manuscript/hg38clearclip.txt",
+hg38clearclip <- read_tsv("data/hg38clearclip.txt",
                           col_names = FALSE)
@@
     region %in% c("3UTRCDS", "CDS3UTR", "5UTR3UTR", "CDS5UTR3UTR", "CDS3UTRintron") ~ "0.93",
+    region %in% c("5UTRCDS", "CDS5UTR") ~ "0.42",
     region %in% c("CDS", "CDSintron") ~ "0.42",
@@
     region %in% c("exon_unclassified", "") ~ "0.2",
-    TRUE ~ NA
-  ) %>% 
+    TRUE ~ NA_character_
+  )) %>% 
     mutate(region_effect = as.double(region_effect))
@@
 ```{r}
-experimentalmirnagene <- readRDS("data/experimentalmirnagene.RDS")
-experimentalmirnagene
+# the data frame generated up to now should match the data frame used in manuscript so the output should be "zero" different rows.
+setdiff(experimentalmirnagene,readRDS("data/experimentalmirnagene.RDS"))
 ```

### 2020-11-12 Supplementary_mat_met.Rmd
remove clash and clear-clip discrepancies

@@
 
-```{r}
+The methods about miRNA:target interactions are based a basic principle that is reading after isolation of miRNA:target chimerics.
+The datasets contain all the chimeric miRNA:target structures found in the medium during the experiment.
+On the other hand, it could be said that the reading is performed as snapshot.Because of that, the methods can provide different chimeric interactions the same miRNA:target pair.
+We have preferred to select most effective interaction parameters for the same miRNA:target pairs that can exhibit various interactions.
+The step is performed as:
+
+```{r eval=FALSE}
+experimentalmirnagene <- experimentalmirnagene %>%
+  dplyr::select(miRNA, Ensembl_Gene_Id, hgnc_symbol, Energy, seed_type_effect, region_effect) %>%
+  distinct() %>%
+  group_by(Ensembl_Gene_Id, miRNA) %>%
+  mutate(
+    seed_type_effect = ifelse(seed_type_effect == max(seed_type_effect),
+                              seed_type_effect, 
+                              max(seed_type_effect) ),
+    Energy = ifelse(Energy == min(Energy), 
+                    Energy, 
+                    min(Energy)),
+    region_effect = ifelse(region_effect == max(region_effect),
+                           region_effect, 
+                           max(region_effect) ) ) %>%
+  distinct()
+```
+
+```{r, eval=FALSE}
 # the data frame generated up to now should match the data frame used in manuscript so the output should be "zero" different rows.

### 2020-11-13 Supplementary_fig_tab.Rmd
mir30a figure

@@
 
+```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=5, fig.height=5, fig.show='hold', fig.cap = "Log2 transformed expression level of miR-30a-5p in tumor and normal tissue samples of 87 patients."}
+
+matched_patient_miRNA_exp_dataset <- readRDS("figure_manuscript/matched_patient_miRNA_exp_dataset.RDS")
+
+significant_mirna_nodes <- readRDS("figure_manuscript/significant_mirna_nodes.RDS")
+
+matched_patient_miRNA_exp_dataset%>%
+  filter(miRNA %in% significant_mirna_nodes)%>%
+  filter(miRNA == "hsa-miR-30a-5p")%>%
+  pivot_longer(cols = contains("TCGA"), 
+               names_to = "patient_id", 
+               values_to = "expressions") %>%
+  mutate(patient = ifelse(endsWith(patient_id, "01A"), 
+                          "tumor", "normal")) %>%
+  group_by(patient, miRNA) %>%
+  ggplot(aes(x = miRNA, y = log2(expressions), fill = patient)) +
+  # scale_y_log10()+
+  geom_boxplot(position = "dodge") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90), 
+        panel.background = element_rect(),
+        legend.title = element_blank()) +
+  xlab("")
+
+```
+
 # 3 Supplementary Tables

### 2020-11-13 Supplementary_fig_tab.Rmd
resize fig S12

@@
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=5, fig.height=5, fig.show='hold', fig.cap = "Log2 transformed expression level of miR-30a-5p in tumor and normal tissue samples of 87 patients."}
+```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=4, fig.height=4, fig.show='hold', fig.cap = "Log2 transformed expression level of miR-30a-5p in tumor and normal tissue samples of 87 patients."}
 

### 2021-01-21 Supplementary_fig_tab.Rmd
updated supplementary files

@@
 
-![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps 2-6 is saved and updated continuously.](workflow1.png){#figs1}
+![Workflow for simulation of competing endogenous RNA regulations. Graph object in steps is saved and updated continuously in simulation.](workflow2.jpg){#figs1}
 
@@
 library(ggplot2)
+library(purrr)
+library(ggsignif)
 knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
@@
 knitr::opts_chunk$set(cache=FALSE)
+source("C:/Users/SELCEN/Desktop/article/R/vis_graph.R")
+source("C:/Users/SELCEN/Desktop/article/R/simulate_vis.R")
+source("C:/Users/SELCEN/Desktop/article/R/simulate.R")
 
@@
 
-```{r, echo=FALSE, fig.cap="Calculation of initial miRNA repression activity using interaction parameters in Sample+ network.
+```{r, echo=FALSE, fig.cap="Calculation of initial miRNA repression level using interaction parameters in Sample+ network.
 A) Interaction paramaters between genes and miRNAs in Sample+ are shown on network while expression levels can be found in Table S1.
 B) Interaction paramaters were updated after normalisation C) Amount of miRNA distributed to each mRNA according to mRNA levels and affinity parameters are shown on edges.
-D) Values on edges indicate repression activity.
-Red values indicate repression activity affected by region effect (RE) parameter.
+D) Values on edges indicate repression level.
+Red values indicate repression level affected by region effect (RE) parameter.
 E) Total repression on G4 from two miRNAs is calculated by adding repression values originating from both miRNAs.", fig.show='hold', out.height= "70%", out.width="70%", fig.align = "center"}
@@
 
+comparisons <-  list(c("GNAS_normal", "GNAS_tumor"), c("MYH9_normal", "MYH9_tumor"), c("SLC7A5_normal", "SLC7A5_tumor"), c("NDRG1_normal", "NDRG1_tumor"), c("TRPS1_normal", "TRPS1_tumor"), c("CDK6_normal", "CDK6_tumor"), c("SYNGR2_normal", "SYNGR2_tumor"), c("CCND1_normal", "CCND1_tumor"), c("SOD2_normal", "SOD2_tumor"), c("VEGFA_normal", "VEGFA_tumor"), c("RGS2_normal", "RGS2_tumor"), c("DNMT1_normal", "DNMT1_tumor"), c("NOTCH2_normal", "NOTCH2_tumor"), c("THBS1_normal", "THBS1_tumor"), c("IGF1R_normal", "IGF1R_tumor"), c("RHOB_normal", "RHOB_tumor"), c("EIF4EBP2_normal", "EIF4EBP2_tumor"), c("SERPINH1_normal", "SERPINH1_tumor"), c("HNRNPK_normal", "HNRNPK_tumor"), c("ARL6IP1_normal", "ARL6IP1_tumor"), c("ZNF217_normal", "ZNF217_tumor"), c("SMAD2_normal", "SMAD2_tumor"), c("RCC2_normal", "RCC2_tumor"), c("NPM1_normal", "NPM1_tumor"), c("RPS23_normal", "RPS23_tumor"), c("COL4A1_normal", "COL4A1_tumor"), c("COL5A2_normal",   "COL5A2_tumor"))
+
 matched_patient_exp_dataset %>%
@@
   group_by(patient, Ensembl_Gene_Id) %>% 
-  ggplot(aes(x = hgnc_symbol, y = log2(expressions), fill = patient)) +
+  mutate(genes= str_c(hgnc_symbol, "_", patient))%>%# distinct(hgnc_symbol)%>% ungroup()%>% select(hgnc_symbol) %>%pull()->gene_list
+  ggplot(aes(x = genes, y = log2(expressions), fill = patient)) +
   # scale_y_log10()+
   geom_boxplot(position = "dodge") +
+  ggsignif::geom_signif(comparisons = comparisons, map_signif_level=TRUE, tip_length = 0.02, test = "wilcox.test", y_position = 19.5)+
   theme_bw() +
@@
   xlab("")
+
 ```
@@
 
+```
+
+## 2.5 Runtime Analysis
+
+```{r, echo=FALSE, fig.cap="Simulation runtime comparison of sampled networks with size 500, 1000, 5000 and 10000", fig.show='hold', out.height= "50%", out.width="50%", fig.align = "center"}
+knitr::include_graphics("runtime_analysis/simulation_runtime_analysis.png")
+```
+
+```{r, echo=FALSE, fig.cap="Perturbation efficiency evaluation function runtime comparison of sampled networks with size 500, 1000, 5000 and 10000", fig.show='hold', out.height= "50%", out.width="50%", fig.align = "center"}
+knitr::include_graphics("runtime_analysis/perturbation_runtime_analysis.png")
 ```
@@
 
+## 3.5 Comparison of perturbation efficiency between Sample and Sample+ dataset
+
+| Data set || Sample+ || Sample 
+|-----------|-------|---------|------------| ------------|
+| **Name** |  **PE**  | **PC** | **PE** | **PC** 
+| Gene1 | 0.132 | 2 | 0.268 | 3 |
+| Gene2 | 0.198 | 3 | 0.268 | 3 |
+| Gene3 | 0.0555 | 2 | 0.150 | 3 |
+| Gene4 | 0.197 | 4 | 0.870 | 5 |
+| Gene5 | 0.143 | 1 | 0.358 | 2 |
+| Gene6 | 0.131 | 1 | 0.619 | 2 |
+| Mir1 | 0.806 | 3 | 1.638 | 4 |
+| Mir2 | 2.80 | 3 | 3.432 | 3 |
+
+Table: Perturbation efficiencies of nodes in Sample+ and Sample data set. PE, perturbation efficiency; PC, perturbed node count.
+
 # 4. Access to code

### 2021-01-21 Supplementary_mat_met.Rmd
updated supplementary files

@@
 `find_node_perturbation` function runs `calc_perturbation` function for each node with given parameters.
-In this particular example upregulation parameter is two, iteration parameter is three.
-By using formulas described in manuscript perturbation efficiency and number of affected nodes (perturbed count) are calculated.
-Results for Sample+ network are shown below.
+In this particular example up-regulation parameter was two, iteration parameter was three.
+By using formulas described in manuscript perturbation efficiency and number of affected nodes (perturbed count) were calculated.
+Results for *Sample* network are shown below.
+
+## 1.1  Sample Data set:
 
@@
 
+sample_graph <- priming_graph(minsamp, 
+                              competing_count = Competing_expression, 
+                              miRNA_count = miRNA_expression)
+
+find_node_perturbation(sample_graph, 
+                       how = 2, 
+                       cycle = 3, 
+                       limit = 0.1) %>%
+  as_tibble() %>%
+  select(name, perturbation_efficiency, perturbed_count)
+
+```
+
+## 1.2 Sample+ Data set:
+
+Results for *Sample+* network are shown below.
+
+```{r, message=FALSE, warning=FALSE}
 sample_graph <- priming_graph(minsamp, 
@@
 
-## 2.1 Arrangement of CLASH dataset
+## 2.1 Arrangement of CLASH data set
 
-CLASH dataset was retrieved from [Helwak et. al](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3650559/) [@helwak_mapping_2013].
+CLASH data set was retrieved from [Helwak et. al](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3650559/) [@helwak_mapping_2013].
 
@@
 
-miRNA releases are obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) was downloaded.
+miRNA releases were obtained from [miRBase](http://www.mirbase.org/ftp.shtml). In this step, release 21 (in Human genome 38) was downloaded.
 
@@
 
-Since CLASH dataset is published according to Human Genome 19 version, we used miRBase release 15.
+Since CLASH data set was published according to Human Genome 19 version, we used miRBase release 15.
 
@@
 
-There are different liftover methods for conversion among Human Genome builds. We preffered to use [UCSC liftover tool](https://genome.ucsc.edu/cgi-bin/hgLiftOver)
+There are different liftover methods for conversion among Human Genome builds. We preferred to use [UCSC liftover tool](https://genome.ucsc.edu/cgi-bin/hgLiftOver)
 
@@
 
-#Obtaining chromosomal locations from miRNA:target interaction dataset.
+#Obtaining chromosomal locations from miRNA:target interaction data set.
 
@@
 
-#removing deleted location from CLASH dataset
+#removing deleted location from CLASH data set
 
@@
 
-# Arrangement in dataset
+# Arrangement in data set
 
@@
 
-### Interpreting the CLASH seed structures in dataset
+### Interpreting the CLASH seed structures in data set
 
@@
 
-## 2.2 Arrangement of CLEAR-CLiP Dataset [@moore_mirnatarget_2015]
+## 2.2 Arrangement of CLEAR-CLiP data set [@moore_mirnatarget_2015]
 
-CLEAR-CLiP dataset was retrieved from [Moore et.al](https://www.nature.com/articles/ncomms9864).
-Please note that we need Supplementary Data 4, however there is error in website, you need to click on Supplementary Data 3 to download the data.
+CLEAR-CLiP data set was retrieved from [Moore et.al](https://www.nature.com/articles/ncomms9864).
+Please note that we need Supplementary Data 4, however there was error in website, you need to click on Supplementary Data 3 to download the data.
 
@@
 
-### Adding Genome Information to dataset
+### Adding Genome Information to data set
 
@@
 
-Please note that we have provided the hg18 liftover dataset under the data folder (lift18), in case of server errors/changes may occur.
+Please note that we have provided the hg18 liftover data set under the data folder (lift18), in case of server errors/changes may occur.
 If you had problems generating lift18 data frame, please run the following code.
@@
 
-#removing deleted location from CLEAR-CLiP dataset
+#removing deleted location from CLEAR-CLiP data set
 
@@
 
-### Seed type manipulation in CLEAR-CLiP dataset
+### Seed type manipulation in CLEAR-CLiP data set
 
-In CLEAR-CLiP dataset, seed types were shown in detail. We adjusted as canonical and non-canonical.
+In CLEAR-CLiP data set, seed types were shown in detail. We adjusted as canonical and non-canonical.
 
@@
 
-## 2.3 Integration of two experimental dataset
+## 2.3 Integration of two experimental data set
 
@@
 
-Energy values in miRNA:target pairs are represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which are utilized in this study.
+Energy values in miRNA:target pairs were represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which were utilized in this study.
 On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study.
@@
 We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
-With this process, we have handled this integrated dataset in context of competitor behaviors and functionality of interactions.
+With this process, we have handled this integrated data set in context of competitor behaviors and functionality of interactions.
 
@@
 
-Firstly, we organized these values due to the fact that the regions were defined differently in two dataset. After that, region effect was added as numeric values (shown in Table S3).
+Firstly, we organized these values due to the fact that the regions were defined differently in two data set. After that, region effect was added as numeric values (shown in Table S3).
 
@@
 
-Secondly, we organized seed type interactions in *Seed type manipulation* section for CLEAR-CLiP dataset to show as found in CLASH dataset.
-Same type formatted values added dataset as numeric values (shown in Table S2).
+Secondly, we organized seed type interactions in *Seed type manipulation* section for CLEAR-CLiP data set to show as found in CLASH data set.
+Same type formatted values added data set as numeric values (shown in Table S2).
 
@@
 The methods about miRNA:target interactions are based a basic principle that is reading after isolation of miRNA:target chimerics.
-The datasets contain all the chimeric miRNA:target structures found in the medium during the experiment.
+The data sets contain all the chimeric miRNA:target structures found in the medium during the experiment.
 On the other hand, it could be said that the reading is performed as snapshot.Because of that, the methods can provide different chimeric interactions the same miRNA:target pair.
 We have preferred to select most effective interaction parameters for the same miRNA:target pairs that can exhibit various interactions.
-The step is performed as:
+The step was performed as:
 
@@
 
-The context of dataset is shown in Table S5 in Supplementary Tables.
+The context of data set was shown in Table S5 in Supplementary Tables.
 
@@
 We defined the functions that can be used with R programming.
-Briefly, these functions process a given miRNA:gene dataset and convert to graph object.
-All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that are given in previous section.
+Briefly, these functions process a given miRNA:gene data set and convert to graph object.
+All values that are significant in miRNA:target interactions are stored in edge variables and processed with formulations that were suggested in manuscript.
 The functions and steps of approach are explained as following (Figure S1) :
 
-**Conversion of dataset**: `priming_graph()` function processes the given dataset that includes competing elements in first variable and repressive element in second variable.
+**Conversion of data set**: `priming_graph()` function processes the given data set that includes competing nodes in first variable and repressive miRNA nodes in second variable.
 If the affinity and/or degradation factors are specified in the function, factors are taken into account, are processed with defaults in vice versa.
 The formulations that are given in equations (1-4) are performed in this function.
-This step gives the graph object which contains efficiency values of miRNA:competing target pairs in steady-state in terms of amount.
-It is assumed that the initial target amounts in the dataset is observed after the repressive activity of miRNAs in steady-state.
+This step gives the graph object which contains *repression activity* values of miRNAs on paired target in steady-state.
 
-**Transition of variables in graph**: In the previous step, the calculations are performed in the edge variables of the graph object.
-However, the graph object allows to use node variables, while the node features are handled to the graph.
-In this direction, update_nodes function carries the amount values to node variables This step must be applied with “once” option because it is primary process.
+**Selecting trigger in graph**: The data set is assumed as steady-state in previous step and the interaction parameters are calculated.
+In the network at steady-state conditions, the change is applied to the graph object for perturbation of steady-state.
+To provide the perturbation in the network the work-flow offer two methods: *update_variables* and *update_how*.
+The first, a new data set that is contained competing and miRNA names and current values of these are processed.
+The second option, the amount of the given node name in update_how function, is changed according to “how” argument.
 
-**Trigger change in graph**: The dataset are assumed as steady-state in previous step and the efficiency coefficients are calculated according to this acceptance.
-In the network that is found in steady-state conditions, the change is applied to the graph object for perturbation of steady-state.
-To provide the perturbation in the network the work-flow offer two methods: *update_variables* and *update_how*.
-The first, a new dataset that is contained competing and repressive element names and current values of these can be processed with update_variables.
-The second option, the amount of the given node name in update_how function can be changed according to “how” argument.
+**Simulation of competing behavior of targets**: After the change in the steady-state conditions, the network object tries to reach steady-state again.
+This process progresses as repeating of regulations after the spreading the changes in the network.
+In this step, simulation of regulations according to given cycle count in *simulate* function is applied.
+After each simulation cycle, the miRNA repression activities are re-calculated and the current values of competing elements are found and saved.
+The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node variables.
 
-**Simulation of competing behavior of targets**: After the change in the steady-state conditions, the network elements try to gain steady-state again.
-This process progresses as repeating of regulations after the spreading the changes in the network.
-In this step, simulation of regulations according to given cycle count in simulate function is applied.
-After each simulation cycle, the miRNA repression values are re-calculated and the current values of competing elements are found and saved.
-The process is performed in the edge data and at the same time outputs of the calculations are carried from edge to node data.
-
-The node elements in the dataset are handled as two type; repressive (miRNAs) and competing (targets).
+The nodes in the data set are handled as two type; miRNAs and competing (i.e., genes or targets).
 It is assumed in approach that while targets are degrading or inhibiting by miRNAs continuously, miRNAs reversibly used.

### 2021-01-24 Supplementary_fig_tab.Rmd
updated supplementary figure and table

@@
 
-# 4. Access to code
+# 4. Notes and Access to code
 
-All codes are available at [github repo](https://github.com/selcenari/ceRNA_multifactorial_model).
++ *Table S7*: represents functional annotations of all highly perturbing genes from simulations of network retrieved from miRTarBase.
+
++ *Table S8*: represents functional annotations of tumor specific highly perturbing genes achieved from simulations of network retrieved from miRTarBase.
+
++ *Table S9*: represents functional annotations of tumor specific highly perturbing genes achieved from simulations of network retrieved from miRNA:target pairs that are validated by high-throughput experimental studies.
+
++ All codes are available at [github repo](https://github.com/selcenari/ceRNA_multifactorial_model).
 

### 2021-01-27 Supplementary_fig_tab.Rmd
supplementary files again

--- 2021-01-21 Supplementary_fig_tab.Rmd
@@
 
-## 2.1 Sample dataset analysis in absence of interaction factors.
+## 2.1 Sample data set analysis in absence of interaction factors.
 
 ```{r, echo=FALSE}
-# Sample dataset refers to minsamp in ceRNAnetsim package.
+# Sample data set refers to minsamp in ceRNAnetsim package.
 data("minsamp")
@@
 
-```{r, message=FALSE, warning=FALSE,echo=FALSE, fig.cap= "Sample Dataset in Steady-state.
+```{r, message=FALSE, warning=FALSE,echo=FALSE, fig.cap= "Sample Data set in Steady-state.
 Inıtial expression levels in minsamp network (*Sample* network in manuscript).
@@
 
-```{r, message=FALSE, warning=FALSE,echo=FALSE,  fig.keep='all', fig.cap= "Gene2 Upregulation on Sample Dataset.
+```{r, message=FALSE, warning=FALSE,echo=FALSE,  fig.keep='all', fig.cap= "Gene2 Upregulation on Sample Data set.
 Expression level of Gene2 is increased from 10,000 to 20,000 in order to demonstrate the calculation steps.", fig.width=5, fig.height=4}
@@
 ```{r, message=FALSE, warning=FALSE, echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2,  fig.show='hold', fig.cap = "Sequential iteration of Sample data after up-regulation of Gene2.
-A) First response of system to Gene2 upregulation (2nd iteration).
+A) First response of network to Gene2 upregulation (2nd iteration).
 B) Spreading of perturbation on system (3th iteration)", fig.subcap= NULL}
@@
 
-```{r, echo=FALSE, fig.cap="Calculation of initial miRNA repression level using interaction parameters in Sample+ network.
+```{r, echo=FALSE, fig.cap="Calculation of initial miRNA repression level (counts) using interaction parameters in Sample+ network.
 A) Interaction paramaters between genes and miRNAs in Sample+ are shown on network while expression levels can be found in Table S1.
-B) Interaction paramaters were updated after normalisation C) Amount of miRNA distributed to each mRNA according to mRNA levels and affinity parameters are shown on edges.
-D) Values on edges indicate repression level.
-Red values indicate repression level affected by region effect (RE) parameter.
-E) Total repression on G4 from two miRNAs is calculated by adding repression values originating from both miRNAs.", fig.show='hold', out.height= "70%", out.width="70%", fig.align = "center"}
+B) Interaction paramaters were updated after normalisation C) Amount of miRNA distributed to each mRNA according to mRNA levels and affinity parameters (Energy and Seed Type Effect) are shown on edges.
+D) Values on edges indicate degredation level (couns).
+Red values indicate degredation level affected by region effect (RE) parameter.
+E) Total repression on G4 from two miRNAs is calculated by summing repression values originating from both miRNAs.", fig.show='hold', out.height= "70%", out.width="70%", fig.align = "center"}
 knitr::include_graphics("fig_s5.png")
@@
 
-$G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE^\prime$, normalized values of seed type efficiency coefficient; $RE^\prime$, normalized values of region efficiency coefficient; $E^\prime$, normalized values of energy coefficient.
+$G$, Gene; $M$, miRNA; $STE$, seed type effect; $RE$, Region Effect; $E$, Energy; $STE^\prime$, normalized values of seed type effect; $RE^\prime$, normalized values of region effect parameter; $E^\prime$, normalized values of Energy parameter.
 Numbering on edges match the pair order in Table S1.
 
-## 2.3 Sample+ dataset analysis with interaction factors.
+## 2.3 Sample+ data set analysis with interaction factors.
 
@@
 
-```{r, message=FALSE, warning=FALSE,echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of Sample+ A) First response of system to Gene2 upregulation (2nd iteration).
+```{r, message=FALSE, warning=FALSE,echo=FALSE, out.width='50%', fig.asp=1, fig.ncol = 2, fig.show='hold', fig.cap = "Sequential iteration of Sample+ A) First response of network to Gene2 upregulation (2nd iteration).
 B) Spreading of perturbation on system (3rd iteration).
@@
 
-## 2.4 Common target perturbation in Sample+ dataset.
+## 2.4 Common target perturbation in Sample+ data set.
 
@@
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=12, fig.height=5,fig.show='hold', fig.cap = "Log2 transformed expression levels of tumor-specific perturbing nodes in tumor and normal tissue samples of 87 patients."}
+```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=12, fig.height=5,fig.show='hold', fig.cap = "Log2 transformed expression levels of tumor-specific perturbing nodes in tumor and normal tissue samples of 87 patients for Real network."}
 
@@
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=4, fig.height=4, fig.show='hold', fig.cap = "Log2 transformed expression level of miR-30a-5p in tumor and normal tissue samples of 87 patients."}
+```{r, message=FALSE, warning=FALSE, echo=FALSE, fig.width=4, fig.height=4, fig.show='hold', fig.cap = "Log2 transformed expression level of miR-30a-5p in tumor and normal tissue samples of 87 patients for Real data set."}
 
@@
 
-## 3.1 *Sample+* dataset
+## 3.1 *Sample+* data set
 
@@
 
-Table: The parameters which affect miRNA:target interactions (i.e. seed type, region, energy) are provided in Sample+ dataset, while these factors are not utilized in simulation of Sample dataset.
+Table: The parameters which affect miRNA:target interactions (i.e. seed type, region, energy) are provided in Sample+ data set, while these factors are not utilized in simulation of Sample data set.
 
@@
 
-Some of information about miRNA:target interactions were exhibited directly by high-throughput studies. On the other hand, we were examined other interacion factors based on different studies.
+Some of information about miRNA:target interactions were exhibited directly by high-throughput studies. On the other hand, we were examined other interaction parameters based on different studies.
 
@@
 + Comparisons of canonical seed types were evaluated by study of [@grimson_microrna_2007], while functional and non-functional seed interactions were studied by [@bartel_micrornas:_2009] and [@betel2010comprehensive].
-+ Numeric definition of target region location effect was performed based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013]
++ Numeric definition of target region location effect was performed based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
 
@@
 
-Table: Context of miRNA:target pairs supported by High-throughput Experiments. CLEAR-CLiP and CLASH datasets were integrated as described in Section 2 of Supplementary Material and Method.
+Table: Context of miRNA:target pairs supported by High-throughput Experiments. CLEAR-CLiP and CLASH data sets were integrated as described in Section 2 of Supplementary Material and Method.
 
@@
 
-As a result of simulation a dataset, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at Table S5.
+As a result of simulation a data set, a graph object is obtained that includes various variables in edge and node data. A graph object includes variables at Table S5.
 
@@
 
-## 3.5 Comparison of perturbation efficiency between Sample and Sample+ dataset
+## 3.5 Comparison of perturbation efficiency between Sample and Sample+ data set
 
@@
 
-# 4. Access to code
+# 4. Notes and Access to code
 
-All codes are available at [github repo](https://github.com/selcenari/ceRNA_multifactorial_model).
++ *Table S7*: refers functional annotations of all highly perturbing genes from simulations of network retrieved from miRTarBase.
+
++ *Table S8*: refers functional annotations of tumor specific highly perturbing genes achieved from simulations of network retrieved from miRTarBase.
+
++ *Table S9*: refers functional annotations of tumor specific highly perturbing genes achieved from simulations of network retrieved from miRNA:target pairs that are validated by high-throughput experimental studies.
+
++ All codes are available at [github repo](https://github.com/selcenari/ceRNA_multifactorial_model).
 

### 2021-01-27 Supplementary_mat_met.Rmd
supplementary files again

@@
 Energy values in miRNA:target pairs were represented by high-throughput studies [@helwak_mapping_2013; @moore_mirnatarget_2015] which were utilized in this study.
-On the other hand, we have specified the other interaction factors, seed type and location of binding region on the target, as numeric values based on the previous studies.[@grimson_microrna_2007] have compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study.
+On the other hand, we specified the other interaction factors (seed type and location of binding region on the target) as numeric values based on the previous studies.[@grimson_microrna_2007] had compared the seed types' effect on target repression with few miRNA had canonical seed pairing in their study.
 Additionally, [@bartel_micrornas:_2009] and [@betel2010comprehensive] have studied on functional and non-functional seed interactions.
-Based on results of these studies we have arranged seed types of miRNA:target interactions as numeric values.
-We also have redefined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
-With this process, we have handled this integrated data set in context of competitor behaviors and functionality of interactions.
+Based on results of these studies we compiled seed types of miRNA:target interactions as numeric values.
+We also have re-defined location of binding region on the target as numeric values, based on studies of [@hausser_analysis_2013] and [@helwak_mapping_2013].
+With this process, we handled this integrated data set in context of competitor behaviors and functionality of interactions.
 
@@
 
-Secondly, we organized seed type interactions in *Seed type manipulation* section for CLEAR-CLiP data set to show as found in CLASH data set.
-Same type formatted values added data set as numeric values (shown in Table S2).
+Secondly, we arranged seed type parameters, as in *Seed type manipulation* section, of CLEAR-CLiP data set to show as found in CLASH data set.
+The values added data set as numeric values (shown in Table S2).
 
@@
 
-The methods about miRNA:target interactions are based a basic principle that is reading after isolation of miRNA:target chimerics.
-The data sets contain all the chimeric miRNA:target structures found in the medium during the experiment.
-On the other hand, it could be said that the reading is performed as snapshot.Because of that, the methods can provide different chimeric interactions the same miRNA:target pair.
+The approaches to detect miRNA:target interactions are based a basic principle that is reading after isolation of miRNA:target chimerics.
+The data sets contain all readings from the chimeric miRNA:target pairs, isolated at a particular time.
+On the other hand, it could be said that the reading is performed in a snapshot fashion.
+Because of that, the methods can provide different chimeric interactions the same miRNA:target pair.
 We have preferred to select most effective interaction parameters for the same miRNA:target pairs that can exhibit various interactions.

### 2021-01-31 figure_manuscript/codes_of_figures.Rmd
update codes of figures

@@
          node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
-  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>% 
+  filter(node_type == "effective", perturbed_count > 78, perturbed > 10)%>% 
   filter(tissue_type != "Normal", startsWith(name, "ENS"))%>%
@@
          node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
-  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>% 
+  filter(node_type == "effective", perturbed_count > 78, perturbed > 10)%>% 
   filter(tissue_type == "Normal", startsWith(name, "ENS"))%>%
@@
          node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
-  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>% 
+  filter(node_type == "effective", perturbed_count > 78, perturbed > 10)%>% 
   filter(tissue_type != "Normal", startsWith(name, "hsa"))%>%
@@
          node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
-  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>% 
+  filter(node_type == "effective", perturbed_count > 78, perturbed > 10)%>% 
   filter(tissue_type == "Normal", startsWith(name, "hsa"))%>%
@@
 
+ggsave(filename = 'mirtarbase_annotation.svg', width = 8, height = 6)
+
 ```
 
-## Network construction with overall functional annotation: **Figure 3**
+-----------------
+## Network construction with overall functional annotation: Additional figure important genes from functional Annotation
 
@@
 
+-----------------
+
 # Heat-maps for miRNAs
@@
          node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
-  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>%
+  filter(node_type == "effective", perturbed_count > 78, perturbed > 10)%>%
   filter(tissue_type != "Normal", startsWith(name, "hsa"))%>%
@@
          node_type = ifelse(name %in% effective_nodes$name, "effective", "non_effective"))%>%
-  filter(node_type == "effective", perturbed_count > 78, perturbed > 1)%>%
+  filter(node_type == "effective", perturbed_count > 78, perturbed > 10)%>%
   filter(tissue_type == "Normal", startsWith(name, "hsa"))%>%
@@
   ggplot(aes(x = file_name, y = name, fill = log_FC))+
-  geom_tile(aes(colour = log_FC, fill = log_FC))+coord_fixed(ratio = 1)+
+  geom_tile(aes(colour = log_FC, fill = log_FC))+
   theme_test()+
   theme(axis.text.x = element_blank(), #element_text(angle = 90, vjust = 0, hjust=0)
-        plot.title = element_text(hjust = 0.5, size =20), 
-        axis.title = element_text(size =20))+
+        plot.title = element_text(hjust = 0.5))+
   scale_colour_gradientn(colours = c("green","white", "red"), aesthetics = c("colour", "fill"), na.value = "grey")+
@@
   xlab("Patients")+
-  theme( legend.title = element_blank(), panel.background = element_rect(fill = "white"), plot.margin = margin(0, 0, 0, 0, "cm"))
+  theme( legend.title = element_blank(), panel.background = element_rect(fill = "white"))
 
@@
   ggplot(aes(x = file_name, y = name, fill = perturbed_count_log))+
-  geom_tile(aes(colour = perturbed_count_log, fill = perturbed_count_log))+coord_fixed(ratio = 1)+
+  geom_tile(aes(colour = perturbed_count_log, fill = perturbed_count_log))+
   theme_test()+
   theme(axis.text.x = element_blank(), #element_text(angle = 90, vjust = 0, hjust=0)
-        plot.title = element_text(hjust = 0.5, size =20), 
-        axis.title = element_text(size =20))+
+        plot.title = element_text(hjust = 0.5))+
   scale_colour_gradientn(colours = c("green","red", "black"), aesthetics = c("colour", "fill"), na.value = "grey")+
@@
   facet_grid(rows = "situ")+
-  theme( legend.title = element_blank(), panel.background = element_rect(fill = "gray93"), plot.margin = margin(0, 0, 0, 0, "cm"))
+  theme( legend.title = element_blank(), panel.background = element_rect(fill = "gray93"))
 

### 2021-02-01 figure_manuscript/codes_of_figures.Rmd
reproducible codes of figures

@@
 
------------------
+----------------- This figure was removed from manuscript.
 ## Network construction with overall functional annotation: Additional figure important genes from functional Annotation
@@
   filter(tissue_type != "Normal", startsWith(name, "hsa"))%>%
+  mutate(file_name= str_remove(file_name, "-01A"))%>%
   dplyr::select(file_name, name,tissue_type, perturbed_count)->mirna_tumor
@@
   filter(tissue_type == "Normal", startsWith(name, "hsa"))%>%
+  mutate(file_name= str_remove(file_name, "-11A"))%>%
   dplyr::select(file_name, name,tissue_type, perturbed_count)->mirna_normal
@@
   xlab("Patients")+
-  theme( legend.title = element_blank(), panel.background = element_rect(fill = "white"))
+  theme( legend.title = element_blank(), panel.background = element_rect(fill = "white"),  plot.margin = margin(0, 0, 0, 0, "cm"))
+
+ggsave(filename = 'mirna_comparison.svg', width = 8, height = 4)
 
@@
   facet_grid(rows = "situ")+
-  theme( legend.title = element_blank(), panel.background = element_rect(fill = "gray93"))
+  theme( legend.title = element_blank(), panel.background = element_rect(fill = "gray93"),  plot.margin = margin(0, 0, 0, 0, "cm"))
 
+ggsave(filename = 'mirna_comparison_normal_tumor.svg', width = 8, height = 7)
 ```

### 2021-02-19 figure_manuscript/codes_of_figures.Rmd
Updating manuscript figure codes

@@
 ```{r, include=FALSE, message=FALSE, warning=FALSE}
-library(janitor)
 library(tidyverse)
@@
 library(ceRNAnetsim)
-library(dplyr)
 library(ggVennDiagram)
 library(ggraph)
+library(janitor)
 knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
@@
 
-mirtarbase_node_perturbations <- readRDS("mirtarbase_node_perturbations.RDS") # data includes results of find_node_perturbation() function for tumor and normal samples of each patient.
+mirtarbase_node_perturbations <- readRDS("mirtarbase_node_perturbations.RDS") 
+# data includes results of find_node_perturbation() 
+#function for tumor and normal samples of each patient.
 
@@
 
-significant_node_graph <- readRDS("significant_node_graph.RDS") #consists network of highly perturbing nodes.
+significant_node_graph <- readRDS("significant_node_graph.RDS") 
+#consists network of highly perturbing nodes.
 hg19 <- readRDS("hg19.RDS") #For id matching (obtained via bioMaRt package)
@@
 
-Annotation_overall <- readRDS("Annotation_overall.RDS") # consists functional annotation of 145 high perturbing nodes.
+Annotation_overall <- readRDS("Annotation_overall.RDS") 
+# consists functional annotation of 145 high perturbing nodes.
 
@@
   left_join(hg19, by =c("name"="ensembl_gene_id"))%>%
-  filter( type %in% c("miRNA", "Gene"))%>%
+  filter( type %in% c("miRNA", "Gene"))-> annotation_graph
+
+annotation_graph%>%
   ggraph(layout = "kk")+ 
@@
   geom_node_point(color = "seashell2")+
-  geom_node_point(aes(filter= annotation == "Enriched Genes", size = centrality, color = "Enriched Genes"), shape = 16)+
-  geom_node_point(aes(filter= type == "miRNA", color = "miRNA"), shape = 17, size = 2)+
-  geom_node_point(aes(filter= type == "GO", size = centrality, color = "GO"), shape = 18)+
-  geom_node_point(aes(filter= type == "KEGG", size = centrality, color = "KEGG"), shape = 15)+
+  geom_node_point(aes(filter= annotation == "Enriched Genes", 
+                      size = centrality, color = "Enriched Genes"), 
+                  shape = 16)+
+  geom_node_point(aes(filter= type == "miRNA", color = "miRNA"), 
+                  shape = 17, 
+                  size = 2)+
+  geom_node_point(aes(filter= type == "GO", size = centrality, color = "GO"), 
+                  shape = 18)+
+  geom_node_point(aes(filter= type == "KEGG", size = centrality, color = "KEGG"), 
+                  shape = 15)+
   theme_graph(base_family = "sans")+
-  geom_node_text(aes(filter= name %in% top_important, label = hgnc_symbol), size =3.5, repel = TRUE)+
+  geom_node_text(aes(filter= name %in% top_important, label = hgnc_symbol), 
+                 size =3.5, 
+                 repel = TRUE)+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"), legend.position = "none")+
   guides(size=FALSE, shape = FALSE)+
-  scale_color_manual(name = "Node types", values = c("Enriched Genes" = "sandybrown", "miRNA"= "dodgerblue2", "GO"= "green", "KEGG"= "red"))
+  scale_color_manual(name = "Node types", 
+                     values = c("Enriched Genes" = "sandybrown",
+                                "miRNA"= "dodgerblue2",
+                                "GO"= "green", "KEGG"= "red"))
+
+```
+
+Note: Alternatively, network can also be visualized by using Cytoscape (v 3.8.2) and Rcy3 package from Bioconductor.
+
+```{r, eval=FALSE}
+annotation_graph
+#BiocManager::install("RCy3")
+
+library(RCy3)
+
+# After starting cytoscape at desktop, following commands are run:
+
+RCy3::cytoscapePing()
+createNetworkFromIgraph(annotation_graph)
 
@@
         plot.title = element_text(hjust = 0.5))+
-  scale_colour_gradientn(colours = c("green","white", "red"), aesthetics = c("colour", "fill"), na.value = "grey")+
+  scale_colour_gradientn(colours = c("green","white", "red"), 
+                         aesthetics = c("colour", "fill"), na.value = "grey")+
   ylab("Highly pertubing miRNA nodes")+
   xlab("Patients")+
-  theme( legend.title = element_blank(), panel.background = element_rect(fill = "white"),  plot.margin = margin(0, 0, 0, 0, "cm"))
+  theme( legend.title = element_blank(), panel.background = element_rect(fill = "white"), 
+         plot.margin = margin(0, 0, 0, 0, "cm"))
 
@@
   dplyr::select(-3, -5)%>%
-  pivot_longer(cols = c("perturbed_count_tumor", "perturbed_count_normal"), names_to = "situ", values_to = "perturbed_count")%>%
+  pivot_longer(cols = c("perturbed_count_tumor", "perturbed_count_normal"), 
+               names_to = "situ", values_to = "perturbed_count")%>%
   mutate(perturbed_count_log = log2(perturbed_count),
@@
   facet_grid(rows = "situ")+
-  theme( legend.title = element_blank(), panel.background = element_rect(fill = "gray93"),  plot.margin = margin(0, 0, 0, 0, "cm"))
+  theme( legend.title = element_blank(), panel.background = element_rect(fill = "gray93"),
+         plot.margin = margin(0, 0, 0, 0, "cm"))
 

### 2021-02-19 Supplementary_mat_met.Rmd
final versions of supplementary files
updating supplementay_mat_met files

@@
 
-```{r, include=FALSE, message=FALSE, warning=FALSE}
+```{r message=FALSE, warning=FALSE, include=FALSE}
 
@@
 ```
-
 ## 1.2 Sample+ Data set:
@@
 clearclipfinal <- clearclipfinal %>% 
-  bind_cols(hg38clearclip) %>% 
+  dplyr::bind_cols(hg38clearclip) %>% 
   rename(HG38build_loc = X1) %>% 
-  mutate(Genom_build="hg18") %>%
+  dplyr::mutate(Genom_build="hg18") %>%
   dplyr::select(cluster=cluster.ID, chromosome = chr, start_position, 
@@
                            max(region_effect) ) ) %>%
+  ungroup()%>%
   distinct()
+
+saveRDS(experimentalmirnagene, file = "data/experimentalmirnagene.RDS")
+
 ```
@@
 # the data frame generated up to now should match the data frame used in manuscript so the output should be "zero" different rows.
-setdiff(experimentalmirnagene,readRDS("data/experimentalmirnagene.RDS"))
+dplyr::setdiff(experimentalmirnagene,readRDS("data/experimentalmirnagene.RDS"))
 ```

### 2021-02-19 multiple_samples_reproducible_codes.Rmd
example reprodcible codes for multiple samples

---
title: "Reproducible multiple sample script"
author:
- Selcen Ari
- Alper Yilmaz
date: "`r Sys.Date()`"
output: pdf_document
---

# Generating networks for multiple sample

This file contains an example of competing endogenous RNA analysis with ceRNAnetsim package.

```{r, include=FALSE, message=FALSE, warning=FALSE}
#required packages
library(tidyverse)
library(tidygraph)
library(readr)
library(biomaRt)
library(TCGAbiolinks)
library(ceRNAnetsim)
library(ggraph)
library(janitor)
library(SummarizedExperiment)
library(ceRNAnetsim)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# 1. Downloading data for analysis

```{r}

#query for Ovarian and Breast Carcinoma in TCGA. In this example TCGAbiolinks package was used.

query_gene_exp <- GDCquery(project = c("TCGA-BRCA", "TCGA-OV"),
                  data.category = "Transcriptome Profiling",
                  data.type = "Gene Expression Quantification", 
                  workflow.type = "HTSeq - Counts")

query_mirna_exp <- GDCquery(project = c("TCGA-BRCA", "TCGA-OV"),
                  data.category = "Transcriptome Profiling",
                  data.type = "Isoform Expression Quantification",
                  workflow.type = "BCGSC miRNA Profiling")

#finding samples which have both miRNA and gene expression dataset

getResults(query_gene_exp, cols = c("project","cases"))%>%
  mutate(barcode = substr(cases, 1,12))-> gene_exp_cases

getResults(query_mirna_exp, cols = c("project","cases"))%>%
  mutate(barcode = substr(cases, 1,12))-> mirna_exp_cases

# For demostration purposes we select 10 random patients per cancer project.
set.seed(1234)  
gene_exp_cases%>%
  inner_join(mirna_exp_cases, by = c("project", "barcode"))%>%
  dplyr::select(project, barcode)%>%
  distinct()%>%
  group_by(project)%>%
  sample_n(10)%>%
  pull(barcode)->selected_barcodes
```

## query for selected patient barcodes

```{r}
query_gene_exp_selected <- GDCquery(project = c("TCGA-BRCA", "TCGA-OV"),
                  data.category = "Transcriptome Profiling",
                  data.type = "Gene Expression Quantification", 
                  workflow.type = "HTSeq - Counts",
                  barcode = selected_barcodes)

query_mirna_exp_selected <- GDCquery(project = c("TCGA-BRCA", "TCGA-OV"),
                  data.category = "Transcriptome Profiling",
                  data.type = "Isoform Expression Quantification",
                  workflow.type = "BCGSC miRNA Profiling",
                  barcode = selected_barcodes)
```

## Downloading and preparing data

```{r}
GDCdownload(query_gene_exp_selected)
GDCdownload(query_mirna_exp_selected)

gene_exps <- GDCprepare(query_gene_exp_selected)
mirna_exps <- GDCprepare(query_mirna_exp_selected)

```

## Preparing gene expression dataset

```{r}
#preparation of gene expression dataset

as.data.frame(assay(gene_exps))%>%
  mutate(ensembl_gene_id = rownames(.))%>%
  dplyr::left_join(as.data.frame(rowData(gene_exps)), 
                    by = "ensembl_gene_id")%>%
  dplyr::select(Ensembl_Gene_Id = ensembl_gene_id, external_gene_name, 1:21)->gene_exp_to_be_analyzed

#fixing sample ids
cases <- which(substring(names(gene_exp_to_be_analyzed),1,4) %in% "TCGA")
names(gene_exp_to_be_analyzed)[cases] <- substr(names(gene_exp_to_be_analyzed)[cases], 1,12)

head(gene_exp_to_be_analyzed)
```

## Preparing miRNA expression dataset

```{r}

# miRNA expression dataset contains miRNAs with mirbase ids. So, firstly, mirbase_id_coversion dataset is downloaded and processed.

mirbase_url <- "ftp://mirbase.org/pub/mirbase/21/genomes/hsa.gff3"

read_tsv(mirbase_url, comment = "#", col_names = FALSE) %>%
  dplyr::select(mirna_type= X3, definition = X9)%>%
  filter(!endsWith(mirna_type, "primary_transcript"))%>%
  tidyr::separate(definition, c("ID", "Alias", "Name", "Derivated"), sep = ";")%>%
  dplyr::select(Alias, Name)%>%
  tidyr::separate(Alias, c("trash1", "ID"), sep = "=")%>%
  tidyr::separate(Name, c("trash2", "Name"), sep = "=")%>%
  dplyr::select(-trash1, -trash2)-> mirbase_id_conv

# preparation of miRNA expression dataset. 
# We used miRBase (Version 21) to obtain miRBase id (like MIMAT0000) for each mature isoform and 
# aggregated readr per million for each isoform.

mirna_exps%>%
  as.data.frame()%>%
  dplyr::select(miRNA_ID, 
                read_count, 
                reads_per_million_miRNA_mapped, 
                miRNA_region, 
                barcode)%>%
  dplyr::filter(startsWith(miRNA_region, "mature"))%>%
  dplyr::mutate(mirbase_id =str_remove(miRNA_region, "mature,"))%>%
  dplyr::select(-miRNA_region)%>%
  dplyr::inner_join(mirbase_id_conv, 
                    by = c("mirbase_id"="ID"))%>%
  dplyr::group_by(Name, barcode)%>%
  mutate(read_count= sum(read_count), 
         reads_per_million_miRNA_mapped = sum(reads_per_million_miRNA_mapped))%>%
  dplyr::ungroup()%>%
  dplyr::select(miRNA = Name, reads_per_million_miRNA_mapped, barcode)%>% # reads_per_million_miRNA_mapped used for ceRNA analysis (optionally)
  distinct()%>%
  tidyr::pivot_wider(names_from = "barcode", values_from = "reads_per_million_miRNA_mapped")-> mirna_exp_to_be_analyzed

#fixing sample ids
cases <- which(substring(names(mirna_exp_to_be_analyzed),1,4) %in% "TCGA")
names(mirna_exp_to_be_analyzed)[cases] <- substr(names(mirna_exp_to_be_analyzed)[cases], 1,12)

head(mirna_exp_to_be_analyzed)

```

# Performing competing endogenous RNA (ceRNAnetsim) analysis

Firstly, miRNA:gene pair dataset must be defined. Here, dataset of miRNA:gene pairs which were obtained from high-throughput experimental studies is used as an example. Note that in manuscript SPONGE analysis was preformed to refine miRNA:gene pairs. In this demonstration we omitted SPONGE analysis and use bare

```{r}

experimentalmirnagene <- readRDS("data/experimentalmirnagene.RDS")

graph_list = list()

for(i in selected_barcodes){

current_network <- experimentalmirnagene%>%
  right_join(dplyr::select(gene_exp_to_be_analyzed, Ensembl_Gene_Id, i), by="Ensembl_Gene_Id")%>%
  right_join(dplyr::select(mirna_exp_to_be_analyzed, miRNA, i), by ="miRNA", suffix= c("Gene_expression", "miRNA_expression"))%>%
  dplyr::select(Ensembl_Gene_Id, miRNA, Gene_expression= paste0(i, "Gene_expression"), miRNA_expression = paste0(i, "miRNA_expression"))%>%
  filter(!is.na(Gene_expression), !is.na(miRNA_expression))%>% 
  filter(Gene_expression != 0, miRNA_expression != 0)%>%
  priming_graph(competing_count = Gene_expression, 
                miRNA_count = miRNA_expression)

graph_list[[i]] <- current_network

}
```

Now we have miRNA:gene networks for 20 patients in a list. So any of them can be used with various functions provided by ceRNAnetsim package. Below we just printing out network of a single patient.

```{r}
graph_list$`TCGA-S3-AA17`
```
